### 2010-05-04 doc/paper.tex
Big update to use underlying mongodb; added mapping db; modified all services; simplified results generation (will be done in mongodb).
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@392 4525493e-7705-40b1-a816-d608a930855b

\documentclass[a4paper]{jpconf}
\usepackage{graphicx}
\begin{document}
\title{The CMS Data Aggregation System}

\author{Valentin Kuznetsov}
\address{Cornell University, Ithaca, New York, USA}
\ead{vkuznet@gmail.com}

\author{Dave Evans}
\address{Fermilab, Batavia, Illinois, USA}
\ead{evansde@fnal.gov}

\author{Simon Matson}
\address{Bristol University, Bristol, UK}
\ead{s.metson@bristol.ac.uk}

%In a large modern enterprise, information is almost inevitably distributed among several database management systems. Despite considerable attention from the research community, relatively few commercial systems have attempted to address this issue. This article describes the technology that enables clients of IBM's federated database engine to access and integrate the data and specialized computational capabilities of a wide range of relational and non­relational data sources.

\begin{abstract}
A metadata plays significant role in a large modern enterprises, research experiments,
digital libraries where it comes from different sources and distributed in a 
variety of forms and digital formats. It is organized and managed by constantly
evolving software using both relational and nonrelational data sources. There is
a big demand to access information from multiple sources.
Here we discuss a new data aggregation system which deliver and cache information 
from different relational and nonrelational data sources on a concrete example 
of large scale High-Energy physics experiment.
\end{abstract}

\newpage

\section{Introduction}
The CERN, the European Organization for Nuclear Research, plays a leading
role in fundamental studies of physics. But apart from scientific community 
it is known as a place where World Wide Web was born. At that time the 
information look-up among hyperlinks represents a certain challenge for scientists.
Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
Physics, promising to deliver a few PB of data each year. Scientists are
facing another set of problems from distributed computing to multi-dimentional
view on information retrieval. One of the aspect of such research is an efficient
and at the same time precise look-up of produced meta-data, which comes in variety 
of forms and digital formats. As was pointed out in \cite{Amr} a mixed content and 
mixed metadata and metadata consistency should be considered as a whole in design 
of the system to successful information discovery.

The CMS, the Compact Muon Solenoid experiment, operated at LHC,
represents an heterogeneous environment of distributing computing, relational and
nonrelational data sources where we face this problem. The computing resources
are distributed among almost 40 countries, 183 institutions and more then 3000 physicists.
Broad variety of RDMS systems at different data centers collect a various
metadata information, including data location and transfer, detector conditions,
calibrations constants, simulation information, etc. Moreover the development
of those components were done in parallel by different group and technology
tools. Therefore a wide spectrum of data services and
data formats represents a problem of information discovery.

\section{Related Work}
Even though the idea of querying relation databases via keyword based search
algorithm is not knew it is still under significant activity in computer
sceince domain. A few
alternative solutions has been proposed to address this issue. On one
side, the federated DB \cite{FedDB} unify data coming from different
RDMS into another DB where SQL queries can be placed to search desired
data. While on another approaches of querying relational
DB using keyword search algorithms has been proposed \cite{DBXplorer, QueryAnswer}.
In former case, you still face with understanding the underlying schema and
imposing relation condition in your query, while in another an exact match
of provided keywords is expected. Even though in last case you can 
intuitively explore information presented in databases, in some cases
it is not enough. For example, how to deal with numbers? What number is represent?
The row ID, the value of the field?

Our users can be classified as proficient
users who understand semantics of the problem and want precise answers
for their queries, at the same time do not need or should not be worried
about underlying schema of underlying databases. Moreover, some information
can be represented in multiple DBs and our users want to place 
queries across those DBs.

In \cite{DBS-QL} a simple, intuitive and flexible query language was introduced 
on top of the CMS data-bookkeeping system. It represented a power of SQL while
hiding underlying relational schema from the end users. As a results
a human questions were intuitively mapped into simple queries. For example,
a question
{\it I'm looking for files who contain data taken on certain date and located at
particular site} was represented as simple as
\begin{verbatim}
find file where date 2009-01-02 11:59 CET and site = T2
\end{verbatim}
We wanted to expand this approach and apply such queries across multiple
data services. Here we discuss a new system, the Data Aggregation System (DAS),
developed in CMS collaboration to address this issue. We start with
discussion of CMS data model and data-services. In section \ref{DAS}
we provides a detailed description of underlying design and components
used in this system. And represents results in section \ref{Results}.
\section{CMS data model\label{DataModel}}
\section{Data Aggregation System\label{DAS}}
\section{Results\label{Results}}

\section{Summary}

\section{Acknowledgements}

This work was supported by the National Science Foundation and Department of Energy of the United States of America. Fermilab is operated by Fermi Research Alliance, LLC under Contract
No. DE-AC02-07CH11359 with the United States Department of Energy.

\section*{References}
\begin{thebibliography}{9}
\bibitem{DBS} A. Afaq, et. al. ``The CMS Dataset Bookkeeping Service'', CHEP 2007 
\bibitem{DBS07} A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
``A multi-dimensional view on information retrieval of CMS data'', CHEP 2007
\bibitem{DD} https://cmsweb.cern.ch/dbs\_discovery

\bibitem{Arms}
C. R. Arms, W. Y. Arms, ``Mixed Content and Mixed Metadata 
Information Discovery in a Messy World'',
chapter from ``Metadata in Practice'', ALA Editions, 2004
\bibitem{DBXplorer}
Sanjay Agrawal, Surajit Chaudhuri, Gautam Das: DBXplorer: A System for
Keyword-Based Search over Relational Databases. ICDE 2002: 5-16

\bibitem{QueryAnswer}
Georgia Koutrika, Alkis Simitsis, Yannis E. Ioannidis: Pr\'{e}cis: The Essence of
a Query Answer. ICDE 2006: 69-78

\bibitem{FedDB}
L. Haas, E. Lin,
``IBM Federated Database Technology'', \\
http://www.ibm.com/developerworks/data/library/techarticle/0203haas/0203haas.html

\bibitem{CouchDB}
http://couchdb.apache.org/

\bibitem{MongoDB}
http://www.mongodb.org/

\end{thebibliography}

\end{document}

### 2010-05-04 doc/workstatement.tex
Big update to use underlying mongodb; added mapping db; modified all services; simplified results generation (will be done in mongodb).
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@392 4525493e-7705-40b1-a816-d608a930855b

\documentclass[a4paper]{jpconf}
\usepackage{graphicx}
\begin{document}
\title{The CMS Data Aggregation System}

\author{Valentin Kuznetsov}
\address{Cornell University, Ithaca, New York, USA}
\ead{vkuznet@gmail.com}

\begin{abstract}
A metadata plays significant role in a large modern enterprises, research experiments,
digital libraries where it comes from different sources and distributed in a 
variety of forms and digital formats. It is organized and managed by constantly
evolving software using both relational and nonrelational data sources. There is
a big demand almost in any industry to access information from multiple 
heterogeneous sources.

The CMS experiment at CERN, represents a particular example where data discovery
plays significant role in daily acitivities of physycists. 
It is expected that starting 2009 it will collect around a few PB of data every 
year of its operations. This reflects into
several hundren of TB of metadata produced by experiment. 
We propose to design and implement a new data aggregation system to search for
data from different relational and non-relational data sources within CMS
experiment environment.
\end{abstract}

\newpage

\section{Introduction}
The CMS experiment is one of the major experiments at CERN. It is expected
that it will start collecting its data in late 2009. Apart from data management
issue the data discovery will play significant role in any analysis and
production workflow tools. In CMS many databases, both
proprietary (ORACLE) and open-source (MySQL, Postgres, SQLite), are used
to store different sort of meta-data. This includes data bookkeeping information,
data location information, run summary information, data quality information,
detector conditions, etc. The information is accessed in a form of
various data-services and presented to the users in variety of data-formats.
Such heterogeneous and distributed environment represents not only manageable
issue, but also inefficient and cumbersome data look-up capabilities. It is desired
to look-up information from more then one data-service and have uniform data
representation across multiple data-services. As a goal we are targeting
towards the Query Language which can be used to place queries across
one or more data-services. Its benefits can be adopted to data-discovery and
production workflow tools.
%An efficient search over one or
%multiple databases without explosure of underlying schema is a key 
%ingredient of any activity in CMS, from production workflow tools to data analysis.

Traditionally RDMS provides a powerful Structured Query Language tool to
query underlying DB. But it requires a full knowledge of DB schema as well
as explicitly specifying relation information in a query. Such approach is
used for creation of specific data-services or APIs which users can use,
but doesn't provide a way for end-users to place a query against
underlying DB. Recently we've been developed DBS-QL language \cite{DBSQL}
which demonstrates that queries can be placed against single DB in
flexible and intuitive way by end-users without any knowledge about
underlying schema. So it becomes desired to extend
this capability over other existing CMS DBs and have ability to
place queries across multiple data-services.

\section{Previous activity}
Cornell group has been played a significant role in design and implementation of
central CMS data bookkeeping system \cite{DBS}. 
%After original workshop
%hosted at Cornell in 2006 \cite{CornellWorkshop} we provided a roadmap 
%document \cite{roadmap} with guidelines to CMS DBS system. 
We also took a leading role in design and 
development of DBS query language \cite{DBSQL, DBS07} and CMS data discovery
service \cite{DD}.

\section{Data Aggregation Service}
Upon success of Data Discovery service \cite{DD} it was proposed
to extend this activity to other data services. A new project called
Data Aggregation Service was born, see \cite{DAS}. Under its unbrella
all meta-data can be queried using DAS Query Language similar to
DBS one \cite{DBSQL}. While such behavior is desired from  a user point of
view, its implementation represents a ``hot'' research topic in
Computer Science community, see \cite{Arms, DBXplorer, QueryAnswer, FedDB}.
Even though success of Information Retrieval (IR) and RDMS systems
are obvious, there is no a common way to utilize both sources of
information via simple, intuitive keyword search. Such demand has
been presented in various enterprise industries,
including health care, medical, financial domains.
While we cannot fully adopt keyword search (due to specific
domain features, such as find data who has bfield 0.4 and runs with
a specific data quality information), the adoption of DBS-QL
seems very reasonable choice.

%A keyword search over distributed relational and non-relational 
%data sources represents a demand in various enterprise industries, 
%including health care, medical, financial domains.

Upon preliminary studies, see \cite{DAS}, we explored a
system based on combination of RDMS and object-oriented databases
which will be capabable to provide desired requirements. Moreover,
it can serve as a cache buffer for variety of CMS data-services
and used by both production workflow tools and end-users.

We propose to design and develop DAS system based on existing
technologies, such as RDMS, IBM federated DB \cite{FedDB} or 
document-oriented databases \cite{CouchDB, MongoDB}, within
domain of CMS experiment and existing workflow APIs. Initially
the following data-services will be cosidered to be participants in
DAS system: data bookkeeping system, data location system, run summary, 
dashboard, data quality and luminosity DBs.
Each data-service will be gradualy added to DAS extending DAS
QL and amount of data processed and cached by DAS. The system will
undergoes to the analysis and testing, via DAS analytics, to address pre-fetching
strategies and scalability issues.

\section{Estimates}
A design and guidelines of DAS system has been summarized on
\cite{DAS} twiki page. Here we represent a main activities within
DAS project.

\begin{itemize}
\item
{\bf DAS core}: 1 FTE\\
Design and implementation of DAS core system. It comprises of the following
components:
\begin{itemize}
\item implementation of DAS request server
\item implementation of DAS Mapping DB
\item implementation of cache plugins, including hot and raw-caches, access
to underlying caching layer, etc.
\item implementation of data-service plugins, e.g. DBS plugin, phedex plugin, etc.
\item implementation of data populator, pre-fetch strategies, etc.
\item administrative toolkit
\item stress and scalability testing
\item unit tests for DAS components
\end{itemize}

\item
{\bf Data-service components}: 0.5 FTE\\
Work on plugable modules to access data-services,
including DBS, Phedex, LumiDB, RunSummary, DQ services.
This work mostly requires coordination between CMS project
leaders, ensuring production of DAS compliant services and data formats or
developing appropriate wrappers to existing legacy data-services. 
It can be paralelized among data-service developers.

\item
{\bf DAS query language}: 0.2 FTE \\
Design and implementation of DAS QL (Query Language);
work on QL syntax and semantics, lexer, parser.
We need to evaluate existing technologies, such as ANTRL or
other lexer/parser tools. Design and implement DAS QL syntax and
semantics; provide mapping between existing data-service APIs and
DAS QL.

\item
{\bf Web interface}: 0.2-0.5 FTE \\
Design and develop web interface for DAS discovery service.
We need to follow CMS SLA guidelines \cite{SLA} on providing secure
and integrated web service within existing CMS infrastructure
on cmsweb site. We must use existing CMS webtools framework
and infrastructure.

\item
{\bf DAS analytics}: 0.2-0.5 FTE \\
Design and develop DAS analytics DB, based on analysis of
common CMS queries to different data-services.
This work includes studies of common CMS queries to
various data-services, DBS, Phedex, RunSummary, etc. As a result
a pre-fetch strategies to DAS should be provided.
\end{itemize}

\section{Hardware requirements}
Initial prototype has been proven to utilize a multi-core technologies
allowing to take a full power of existing hardware. In stress tests
the 8-core machine, 2.33 GHz/core and 16GB of RAM has been
used. We plan to use it as a DAS request/processing server. In addition
we need to allocate significant disk space for caching purposes.
For example, existing size of DBS system is 27 GB (tables) and 37 GB (index).
So we should rely on a few TB of cache for DAS, possible extandable over
time and load on a system.

\section*{References}
\begin{thebibliography}{9}
\bibitem{DBSQL}  A. Afaq, V. Kuznetsov, L. Lueking, D. Riley, V. Sekhri,
''The CMS Dataset Bookkeeping Service Query Language (DBSql)''
To be published in CHEP 2009 procedings.
%\bibitem{CornellWorkshop}
%https://wiki.lepp.cornell.edu/hep/bin/view/DBSWorkshop/

\bibitem{DBS} A. Afaq, et. al. ``The CMS Dataset Bookkeeping Service'', CHEP 2007 
%\bibitem{roadmap} https://twiki.cern.ch/twiki/pub/CMS/DBS-TDR/roadmap.pdf
\bibitem{DBS07} A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
``A multi-dimensional view on information retrieval of CMS data'', CHEP 2007
\bibitem{DD} https://cmsweb.cern.ch/dbs\_discovery
\bibitem{DAS} https://twiki.cern.ch/twiki/bin/view/CMS/DMWMDataAggregationService
\bibitem{Arms}
C. R. Arms, W. Y. Arms, ``Mixed Content and Mixed Metadata 
Information Discovery in a Messy World'',
chapter from ``Metadata in Practice'', ALA Editions, 2004

\bibitem{DBXplorer}
Sanjay Agrawal, Surajit Chaudhuri, Gautam Das: DBXplorer: A System for
Keyword-Based Search over Relational Databases. ICDE 2002: 5-16

\bibitem{QueryAnswer}
Georgia Koutrika, Alkis Simitsis, Yannis E. Ioannidis: Pr\'{e}cis: The Essence of
a Query Answer. ICDE 2006: 69-78

\bibitem{FedDB}
L. Haas, E. Lin,
``IBM Federated Database Technology'', \\
http://www.ibm.com/developerworks/data/library/techarticle/0203haas/0203haas.html

\bibitem{CouchDB}
http://couchdb.apache.org/

\bibitem{MongoDB}
http://www.mongodb.org/

\bibitem{SLA} 
https://twiki.cern.ch/twiki/bin/view/CMS/DMWTServiceLevelAgreement

\end{thebibliography}

\end{document}

### 2010-05-04 src/python/DAS/core/das_viewmanager.py

@@ -5,8 +5,8 @@
 View manager class.
 """
 
-__revision__ = "$Id: das_viewmanager.py,v 1.6 2009/07/15 15:58:27 valya Exp $"
-__version__ = "$Revision: 1.6 $"
+__revision__ = "$Id: das_viewmanager.py,v 1.7 2009/09/01 01:42:44 valya Exp $"
+__version__ = "$Revision: 1.7 $"
 __author__ = "Valentin Kuznetsov"
 
 # system modules
@@ -48,7 +48,7 @@ class User(BASE):
     id          = Column(Integer, primary_key=True)
     login       = Column(Text, nullable=False, unique=True)
     fullname    = Column(Text, nullable=False, unique=True)
-    contact     = Column(Text, nullable=True, unique=True)
+    contact     = Column(Text, nullable=True)
     group_id    = Column(Integer, ForeignKey('groups.id'))
     group       = relation(Group, backref=backref('groups', order_by=id))
 
@@ -111,10 +111,10 @@ class DASViewManager(object):
         self.engine  = create_engine(dbengine, echo=False)
         self.session = sessionmaker(bind=self.engine)
         if  not dbfile:
-            self.create_table()
+            self.create_db()
         else: # sqlite case
             if  not os.path.isfile(dbfile):
-                self.create_table()
+                self.create_db()
         if  config.has_key('sum_views'):
             login = 'dascore'
             fullname = 'DAS'
@@ -124,10 +124,9 @@ class DASViewManager(object):
                 if  not existing_views.has_key(viewname):
                     self.create(viewname, viewdef, login, fullname, group)
 
-    def create_table(self):
+    def create_db(self):
         """Create DB tables based on ORM objects"""
         metadata = MetaData()
-#        table    = View.__table__
         BASE.metadata.create_all(self.engine)
 
     def create(self, viewname, query, 

### 2010-05-04 src/python/DAS/services/dashboard/dashboard_service.py

@@ -11,14 +11,18 @@ __author__ = "Valentin Kuznetsov"
 import time
 import types
 import urllib
+try:
+    # Python 2.5
+    import xml.etree.ElementTree as ET
+except:
+    # prior requires elementtree
+    import elementtree.ElementTree as ET
 from DAS.services.abstract_service import DASAbstractService
-from DAS.utils.utils import map_validator
-from DAS.services.dashboard.dashboard_parser import parser
-from DAS.core.das_mapping import jsonparser4key
+from DAS.utils.utils import map_validator, dasheader
 
 def convert_datetime(sec):
     """Convert seconds since epoch to date format used in dashboard"""
-    return time.strftime("%Y-%m-%d %H:%M:%S",time.gmtime(sec))
+    return time.strftime("%Y-%m-%d %H:%M:%S", time.gmtime(sec))
 
 class DashboardService(DASAbstractService):
     """
@@ -27,30 +31,34 @@ class DashboardService(DASAbstractService):
     def __init__(self, config):
         DASAbstractService.__init__(self, 'dashboard', config)
         self.headers = {'Accept': 'text/xml'}
-        self.map = {
-            'jobsummary-plot-or-table' : {
-                'keys' : ['jobsummary'],
-                'params' : {
-                    'user': '',
-                    'site': '',
-                    'ce': '',
-                    'submissiontool': '',
-                    'dataset': '',
-                    'application': '',
-                    'rb': '',
-                    'activity': '',
-                    'grid': '',
-                    'date1': '',
-                    'date2': '',
-                    'jobtype': '',
-                    'tier': '',
-                    'check': 'submitted',
-                }
-            }
-        }
+        self.map = self.dasmapping.servicemap(self.name)
         map_validator(self.map)
 
-    def api(self, query, cond_dict=None):
+    def parser(self, api, data, params=None):
+        """
+        Dashboard XML parser, it returns a list of dict rows, e.g.
+        [{'file':value, 'run':value}, ...]
+        """
+        try:
+            elem  = ET.fromstring(data)
+        except:
+            print "data='%s'" % data
+            raise Exception('Unable to parse dashboard output')
+        for i in elem:
+            if  i.tag == 'summaries':
+                for j in i:
+                    row = {}
+                    for k in j.getchildren():
+                        name = k.tag
+                        row[name] = k.text
+                    if  params:
+                        for key, val in params.items():
+                            if  not row.has_key(key):
[…]

### 2010-05-04 src/python/DAS/services/dbs/dbs_service.py

@@ -4,14 +4,18 @@
 """
 DBS service
 """
-__revision__ = "$Id: dbs_service.py,v 1.7 2009/07/10 19:26:11 valya Exp $"
-__version__ = "$Revision: 1.7 $"
+__revision__ = "$Id: dbs_service.py,v 1.8 2009/09/01 01:42:45 valya Exp $"
+__version__ = "$Revision: 1.8 $"
 __author__ = "Valentin Kuznetsov"
 
-import types
 from DAS.services.abstract_service import DASAbstractService
-from DAS.services.dbs.dbs_parser import parser, parser_dbshelp
-from DAS.utils.utils import genkey, map_validator
+from DAS.utils.utils import map_validator
+try:
+    # Python 2.5
+    import xml.etree.ElementTree as ET
+except:
+    # prior requires elementtree
+    import elementtree.ElementTree as ET
 
 class DBSService(DASAbstractService):
     """
@@ -19,53 +23,64 @@ class DBSService(DASAbstractService):
     """
     def __init__(self, config):
         DASAbstractService.__init__(self, 'dbs', config)
-        self.results = []
-        self.params  = {'apiversion': 'DBS_2_0_6', 'api': 'executeQuery'}
-        self._keys   = None
-        # DBS uses DBSServlet and API passed as parameter, so we don't
-        # define api in map and rather use url w/ DBSServlet
-        self.map     = {
-            '' : {
-                'keys' : self.dbs_keys(),
-                'params' : self.params
-            }
-        }
+        self.reserved = ['api', 'apiversion']
+        self.map = self.dasmapping.servicemap(self.name, 'javaservlet')
         map_validator(self.map)
 
-    def parameters(self):
-        """
-        Return mapped service parameters, we overwrite this function since 
-        in DBS case input parameters for query are equal to DBS keys
-        """
-        return self.dbs_keys()
+    def transform_tag(self, system, tag, cache):
+        """Transform given tag name into DAS notation"""
+        if  cache.has_key(tag):
+            newkey = cache[tag]
+        else:
+            newkey = self.dasmapping.notation2das(system, tag)
+            cache[tag] = newkey
+        return newkey
 
-    def dbs_keys(self):
-        """
-        Fetch DBS-QL keys and store them in memcache.
-        """
-        if  self._keys:
-            return self._keys
-        url    = self.url
-        params = dict(self.params)
-        params['api'] = 'getHelp'
-        res = self.getdata(url, params)
[…]

### 2010-05-04 src/python/DAS/services/lumidb/lumidb_service.py

@@ -2,8 +2,8 @@
 """
 LumiDB service
 """
-__revision__ = "$Id: lumidb_service.py,v 1.7 2009/06/05 14:12:37 valya Exp $"
-__version__ = "$Revision: 1.7 $"
+__revision__ = "$Id: lumidb_service.py,v 1.8 2009/09/01 01:42:46 valya Exp $"
+__version__ = "$Revision: 1.8 $"
 __author__ = "Valentin Kuznetsov"
 
 from DAS.services.abstract_service import DASAbstractService
@@ -12,63 +12,5 @@ from DAS.utils.utils import map_validator
 class LumiDBService(DASAbstractService):
     def __init__(self, config):
         DASAbstractService.__init__(self, 'lumidb', config)
-        self.map = {
-            'findTrgPathByRun' : {
-                'api' : {'api':'findTrgPathByRun'},
-                'keys': ['trigpath'],
-                'params' : {'run_number':''}
-            },
-            'findIntegratedLuminosity' : {
-                'api' : {'api':'findIntegratedLuminosity'},
-                'keys': ['intlumi'],
-                'params' : {'run_number':'', 'tag':'', 'hlt_path':''}
-            },
-            'findAvgIntegratedLuminosity' : {
-                'api' : {'api':'findAvgIntegratedLuminosity'},
-                'keys': ['avglumi'],
-                'params' : {'run_number':'', 'tag':'', 'hlt_path':''}
-            },
-            'findIntRawLumi' : {
-                'api' : {'api':'findIntRawLumi'},
-                'keys': ['intrawlumi'],
-                'params' : {'run_number':''}
-            },
-            'findL1Counts' : {
-                'api' : {'api':'findL1Counts'},
-                'keys': ['L1counts'],
-                'params' : {'run_number':'', 'cond_name':''}
-            },
-            'findHLTCounts' : {
-                'api' : {'api':'findHLTCounts'},
-                'keys': ['HLTcounts'],
-                'params' : {'run_number':'', 'path_name':'', 'count_type':''}
-            },
-            'findRawLumi' : {
-                'api' : {'api':'findRawLumi'},
-                'keys': ['rawlumi'],
-                'params' : {'run_number':'', 'tag':''}
-            },
-            'listLumiByBunch' : {
-                'api' : {'api':'listLumiByBunch'},
-                'keys': ['lumibybunch'],
-                'params' : {'run_number':'', 'lumi_section_number':'', 
-                            'option':''}
-            },
-            'listLumiSummary' : {
[…]

### 2010-05-04 src/python/DAS/services/phedex/phedex_service.py

@@ -4,12 +4,19 @@
 """
 Phedex service
 """
-__revision__ = "$Id: phedex_service.py,v 1.8 2009/07/22 20:40:11 valya Exp $"
-__version__ = "$Revision: 1.8 $"
+__revision__ = "$Id: phedex_service.py,v 1.9 2009/09/01 01:42:46 valya Exp $"
+__version__ = "$Revision: 1.9 $"
 __author__ = "Valentin Kuznetsov"
 
 from DAS.services.abstract_service import DASAbstractService
-from DAS.utils.utils import map_validator, splitlist
+from DAS.utils.utils import map_validator
+import types
+try:
+    # Python 2.6
+    import json
+except:
+    # Prior to 2.6 requires simplejson
+    import simplejson as json
 
 class PhedexService(DASAbstractService):
     """
@@ -17,86 +24,64 @@ class PhedexService(DASAbstractService):
     """
     def __init__(self, config):
         DASAbstractService.__init__(self, 'phedex', config)
-        self.map = {
-            'blockReplicas' : {
-                'keys': ['block', 'node', 'site', 
-                         'block.size', 'block.numfiles', 'block.replica',
-                         'block.custodial', 'block.is_open',
-                         'block.complete', 'block.createdate', 'block.moddate'],
-                'params' : {'se':'', 'block':'', 'node':''}
-            },
-#            'fileReplicas' : {
-#                'keys': ['block', 'node', 'site',
-#                         'block.size', 'block.numfiles',
-#                         'file', 'file.checksum',
-#                         'file.node', 'file.origin_node'],
-#                'params' : {'se':'', 'block':'required', 'node':''}
-#            },
-            'fileReplicas' : {
-                'keys': ['file', 'file.checksum',
-                         'file.node', 'file.origin_node'],
-                'params' : {'se':'', 'block':'required', 'node':''}
-            },
-            'nodes' : {
-                'keys': ['site', 'node', 
-                         'node.storage', 'node.kind'],
-                'params' : {'node':'', 'noempty':''}
-            },
-            'lfn2pfn' : {
-                'keys': ['file.pfn', 'file', 'node', 'file.protocol',
-                         'file.custodial'],
-                'params' : {'node':'required', 'lfn':'required', 
-                            'destination':'', 'protocol':'srmv2'}
-            },
-            'tfc' : {
-                'keys': ['tfc', 'tfc.protocol', 'tfc.element_name'],
-                'params' : {'node':'required'}
[…]

### 2010-05-04 src/python/DAS/services/runsum/runsum_parser.py

@@ -4,8 +4,8 @@
 """
 RunSummary XML parser
 """
-__revision__ = "$Id: runsum_parser.py,v 1.2 2009/06/04 14:10:18 valya Exp $"
-__version__ = "$Revision: 1.2 $"
+__revision__ = "$Id: runsum_parser.py,v 1.3 2009/09/01 01:42:47 valya Exp $"
+__version__ = "$Revision: 1.3 $"
 __author__ = "Valentin Kuznetsov"
 
 try:
@@ -26,13 +26,18 @@ def parser(data):
             row = {}
             for j in i:
                 if  j.tag:
-                    row[j.tag] = j.text
+                    newkey = self.dasmapping.notation2das(system, j.tag)
+                    row[newkey] = j.text
+#                    row[j.tag] = j.text
                 nrow = {}
                 for k in j.getchildren():
                     if  k.tag:
-                        nrow[k.tag] = k.text
+                        nkey = self.dasmapping.notation2das(system, k.tag)
+                        nrow[nkey] = k.text
+#                        nrow[k.tag] = k.text
                 if  nrow:
-                    row[j.tag] = nrow
+                    row[newkey] = nrow
+#                    row[j.tag] = nrow
             yield row
 
 #

### 2010-05-04 src/python/DAS/utils/das_config.py

@@ -5,8 +5,8 @@
 Config utilities
 """
 
-__revision__ = "$Id: das_config.py,v 1.21 2009/07/24 01:08:14 valya Exp $"
-__version__ = "$Revision: 1.21 $"
+__revision__ = "$Id: das_config.py,v 1.22 2009/09/01 01:42:47 valya Exp $"
+__version__ = "$Revision: 1.22 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -45,7 +45,6 @@ def das_readconfig(dasconfig=None):
     configdict['couch_lifetime'] = config.getint('couch', 'lifetime')
     configdict['couch_cleantime'] = config.getint('couch', 'cleantime')
 
-    configdict['mongocache_dir'] = config.get('mongocache', 'dir', '')
     configdict['mongocache_dbhost'] = config.get('mongocache', 'dbhost', 'localhost')
     configdict['mongocache_dbport'] = int(config.get('mongocache', 'dbport', '27017'))
     configdict['mongocache_dbname'] = config.get('mongocache', 'dbname', 'das')
@@ -62,6 +61,9 @@ def das_readconfig(dasconfig=None):
 
     configdict['views_engine'] = config.get('views', 'db_engine', None)
     configdict['views_dir'] = config.get('views', 'dir', '')
+
+    configdict['mapping_db_engine'] = config.get('mapping_db', 'db_engine', None)
+    configdict['mapping_db_dir'] = config.get('mapping_db', 'dir', '')
 
     configdict['rawcache'] = config.get('das', 'rawcache', None)
     configdict['hotcache'] = config.get('das', 'hotcache', None)
@@ -106,7 +108,7 @@ def das_writeconfig():
     config.add_section('das')
     config.set('das', 'systems', '%s' % systems)
     config.set('das', 'verbose', 0)
-    config.set('das', 'rawcache', 'DASFilecache')
+    config.set('das', 'rawcache', 'DASMongocache')
 #    config.set('das', 'hotcache', 'DASMemcache')
     config.set('das', 'hotcache', 'DASFilecache')
     config.set('das', 'logdir', '/tmp')
@@ -122,7 +124,6 @@ def das_writeconfig():
     config.set('couch', 'cleantime', 2*60*60) # in seconds
 
     config.add_section('mongocache')
-    config.set('mongocache', 'dir', '/tmp/db')
[…]

### 2010-05-04 src/python/DAS/utils/utils.py

@@ -5,8 +5,8 @@
 General set of useful utilities used by DAS
 """
 
-__revision__ = "$Id: utils.py,v 1.23 2009/07/22 20:40:11 valya Exp $"
-__version__ = "$Revision: 1.23 $"
+__revision__ = "$Id: utils.py,v 1.24 2009/09/01 01:42:47 valya Exp $"
+__version__ = "$Revision: 1.24 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -28,6 +28,33 @@ class dict_of_none (dict):
         """Assign missing key to None"""
         return None
 
+def dict_value(dict, prim_key):
+    """
+    Find value in given dictionary for given primary key. The primary
+    key can be composed one, e.g. a.b.
+    """
+    try:
+        if  prim_key.find('.') != -1:
+            value = dict
+            for key in prim_key.split('.'):
+                value = value[key]
+        else:
+            value = dict[prim_key]
+        return value
+    except:
+        msg  = 'Unable to look-up key=%s\n' % prim_key
+        msg += 'dict=%s' % str(dict)
+        raise Exception(msg)
+def merge_dict(dict1, dict2):
+    """
+    Merge content of two dictionaries w/ default list value for keys.
+    """
+    merged_dict = {}
+    for dictionary in [dict1, dict2]:
+        for key, value in dictionary.items():
+            merged_dict.setdefault(key,[]).append(value) 
+    return merged_dict
+
 def splitlist(ilist, nentries):
     """
     Split input list into a list of lists with nentries
@@ -38,6 +65,40 @@ def splitlist(ilist, nentries):
         if  jdx > len(ilist):
             jdx = len(ilist)
         yield ilist[idx:jdx]
+
+def dasheader(system, query, api, url, args, ctime, expire, ver):
+    """
+    Return DAS header (dict) wrt DAS specifications, see
+    https://twiki.cern.ch/twiki/bin/view/CMS/DMWMDataAggregationService
+    #DAS_data_service_compliance
+    """
+    timestamp = time.time()
+    dasdict = dict(system=[system], timestamp=timestamp,
+                url=url, ctime=ctime, query=[query],
+                params={api:args}, version=ver,
+                expire=timestamp+expire, api=[api])
+    return dict(das=dasdict)
+
+def update_dasheader(header, system, query, api, url, args, ctime, expire, ver):
+    """
+    Update existing dasheader in MognoDB
+    """
+    systems = header['das']['system']
[…]

### 2010-05-04 test/das_mapping_t.py

@@ -1,104 +1,135 @@
 #!/usr/bin/env python
-#pylint: disable-msg=c0301,c0103
+#pylint: disable-msg=C0301,C0103
 
 """
-unit test for DAS mapping set of tools
+Unit test for DAS mappingdb class
 """
 
+import os
+import time
 import unittest
-from DAS.core.das_mapping import translate, jsonparser4key, json2das
-from DAS.core.das_mapping import das2api, result2das
+from DAS.utils.das_config import das_readconfig
+from DAS.utils.logger import DASLogger
+from DAS.core.das_mapping import DASMapping
 
-class testDASMapping(unittest.TestCase):
+class testDASMappingMgr(unittest.TestCase):
     """
-    A test class for the DAS mapping
+    A test class for the DAS mappingdb class
     """
-    def test_mapconfig(self): 
-        """test DAS translate routine"""
-        result = das2api('phedex', 'site')
-        expect = ['se']
-        self.assertEqual(expect, result)
+    def setUp(self):
+        """
+        set up DAS core module
+        """
+        debug    = 0
+        self.db  = 'test_mapping.db'
+        config   = das_readconfig()
+        logger   = DASLogger(verbose=debug, stdout=debug)
+        config['logger']  = logger
+        config['verbose'] = debug
+        config['mapping_db_engine'] = 'sqlite:///%s' % self.db
+        self.mgr = DASMapping(config)
 
-    def test_jsonparser4key(self): 
-        """test DAS jsonparser routine"""
-        rep1 = {'bytes':1, 'files':1, 'se':'T2'}
-        rep2 = {'bytes':1, 'files':1, 'se':'T3'}
-        rec1 = {'bytes':1, 'files':1, 'name':'ABC', 'replica':rep1}
-        rec2 = {'bytes':2, 'files':2, 'name':'CDE', 'replica':[rep1, rep2]}
-        jsondict = {'service': {'block': [rec1, rec2]},'timestamp':123}
-        result = [i for i in jsonparser4key(jsondict, 'block.name')]
-        expect = ['ABC', 'CDE']
-        self.assertEqual(expect, result)
+    def tearDown(self):
+        """Invoke after each test"""
+        try:
+            os.remove(self.db)
+        except:
+            pass
 
-        result = [i for i in jsonparser4key(jsondict, 'block.bytes')]
-        expect = [1, 2]
-        self.assertEqual(expect, result)
[…]

### 2010-05-04 test/qlparser_t.py

@@ -90,12 +90,19 @@ class testQLParser(unittest.TestCase):
 
         q = "find runs where run>1 and run <= 200 "
         result = mongo_exp(ql.conditions(q))
-        expect = [{'run': {'$gt' : '1', '$lte' : '200'}}]
-        print "\nMOGNO", q, result, expect
+        expect = {'run': {'$gt' : '1', '$lte' : '200'}}
         self.assertEqual(result, expect)
 
-        q = "find runs where DQFlagList=Tracker_Global=GOOD&Tracker_Local1=1"
-        result = ql.params(q)
+        q = "find runs where dataset=/a/b/c and DQFlagList=Tracker_Global=GOOD&Tracker_Local1=1"
+        result = ql.params(q)['conditions']
+        expect = [{'value': '/a/b/c', 'key': 'dataset', 'op': '='}, 'and', 
+        {'value': 'Tracker_Global=GOOD&Tracker_Local1=1', 'key': 'DQFlagList', 'op': '='}]
+        self.assertEqual(result, expect)
+
+        q = "find runs where DQFlagList=Tracker_Global=GOOD&Tracker_Local1<=1"
+        result = ql.params(q)['conditions']
+        expect = [{'value': 'Tracker_Global=GOOD&Tracker_Local1<=1', 'key': 'DQFlagList', 'op': '='}]
+        self.assertEqual(result, expect)
 
         q = "find dataset where ((run=1 or run=2) or dataset=bla) and site=cern order by block"
         result = ql.selkeys(q)
@@ -127,21 +134,25 @@ class testQLParser(unittest.TestCase):
         q = "find phedex:block where block=bla"
         result = ql.params(q)
         expect = {'functions': {}, 'selkeys': ['block'], 
-                'unique_services': ['phedex'], 'order_by_list': [], 
-                'order_by_order': None, 'services': {'phedex': ['block']}, 
-                'daslist': [{'phedex': 'find block where block=bla'}], 
+                'unique_services': ['dbs', 'phedex'], 'order_by_list': [], 
+                'order_by_order': None, 
+                'services': {'phedex': ['block'], 'dbs': ['block']}, 
+                'dasqueries': {'phedex': ['find block where block = bla'],
+                               'dbs': ['find block where block = bla']}, 
                 'conditions': [{'value': 'bla', 'key': 'block', 'op': '='}], 
-                'allkeys': ['block'], 'unique_keys': ['block']}
[…]

### 2010-05-04 test/utils_t.py

@@ -6,16 +6,51 @@ Unit test for DAS QL parser
 """
 
 import unittest
-from DAS.utils.utils import cartesian_product
+from DAS.utils.utils import cartesian_product, dasheader, update_dasheader
 from DAS.utils.utils import genresults, transform_dict2list
 from DAS.utils.utils import sitename, add2dict, map_validator
 from DAS.utils.utils import splitlist, gen_key_tuples, sort_data
+from DAS.utils.utils import dict_value, merge_dict
 
 class testUtils(unittest.TestCase):
     """
     A test class for the DAS utils module
     """
+    def test_merge_dict(self):
+        """Test merge_dict"""
+        dict1 = {'block':{'name':'AAA', 'b':{'c':1}, 'size':2}, 'das':{'system':'dbs'}}
+        dict2 = {'block':{'name':'AAA', 'x':{'y':1}, 'z':1, 'size':2}, 'das':{'system':'phedex'}}
+        result = merge_dict(dict1, dict2)
+        expect = {'block': [{'b': {'c': 1}, 'name': 'AAA', 'size': 2}, 
+        {'x': {'y': 1}, 'z': 1, 'name': 'AAA', 'size': 2}], 
+        'das': [{'system': 'dbs'}, {'system': 'phedex'}]}
+        self.assertEqual(expect, result)
+
+    def test_dict_value(self):
+        """Test dict_value"""
+        dict = {'a':{'b':{'c':1}}, 'd':2}
+        result = dict_value(dict, 'a.b.c')
+        expect = 1
+        self.assertEqual(expect, result)
+
+        result = dict_value(dict, 'd')
+        expect = 2
+        self.assertEqual(expect, result)
+        
+    def test_dasheader(self):
+        """Test DAS header"""
+        expect = ['dbs']
+        header = dasheader('dbs', 'q1', 'api1', 'url1', 'args1', 'ct1', 10, 1)
+        self.assertEqual(expect, header['das']['system'])
+
+        expect = ['dbs', 'phedex']
+        update_dasheader(header, 'phedex', 'q2', 'api2', 'url2', 'args2', 'ct2', 10, 1)
+        self.assertEqual(expect, header['das']['system'])
+        expect = {'api1':'args1', 'api2':'args2'}
+        self.assertEqual(expect, header['das']['params'])
+
     def test_cartesian_product(self):
+        """Test cartesian product function"""
[…]

### 2010-05-04 src/python/DAS/services/sitedb/sitedb_service.py
Revisit look-up of primary key and extend API to accept api name and input value. Based on those I can able to look-up appropriate primary key (if multiple has been assigned to data-service API). For instance I should be able to distinguish this case: site=T1_UK means site.name versus site=srm.cern.ch means site.se
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@393 4525493e-7705-40b1-a816-d608a930855b

@@ -4,8 +4,8 @@
 """
 SiteDB service
 """
-__revision__ = "$Id: sitedb_service.py,v 1.11 2009/09/01 01:42:47 valya Exp $"
-__version__ = "$Revision: 1.11 $"
+__revision__ = "$Id: sitedb_service.py,v 1.12 2009/09/01 17:06:16 valya Exp $"
+__version__ = "$Revision: 1.12 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -22,10 +22,24 @@ class SiteDBService(DASAbstractService):
         self.map = self.dasmapping.servicemap(self.name)
         map_validator(self.map)
 
+#    def clean_params(self, api, params):
+#        """
+#        Clean all parameters to get as much as possible information
+#        from SiteDB. Skip only those which marked as required.
+#        """
+#        args = {}
+#        for key, val in self.map[api]['params'].items():
+#            if  val != 'required':
+#                args[key] = val
+#            else:
+#                args[key] = params[key]
+#        return args
+
     def parser(self, api, data, params=None):
         """
         Parser for SiteDB JSON data-services
         """
+        cache = {}
         jsondict = eval(data)
         pat = re.compile('T[0-9]_')
         for key, val in jsondict.items():
@@ -58,84 +72,5 @@ class SiteDBService(DASAbstractService):
                         if  val.find('.') != -1: # SE or CE
                             if  not row.has_key('se'):
                                 row['se'] = val
+            self.row2das(self.name, row)
             yield {'site': row}
-
-    def se2cms(self, site_se):
-        """
-        Convert SE to CMS names.
-        """
-        args = {'name': site_se}
-        jsondict = self.call_service_api('SEtoCMSName', args)
-        for value in jsondict.values():
-            return value['name']
-
-    def cms2se(self, cmsname):
-        """
-        Convert CMS name to SE.
-        """
-        args = {'name':cmsname}
-        jsondict = self.call_service_api('CMSNametoSE', args)
-        selist = []
-        for value in jsondict.values():
-            se = value['name']
-            if  not selist.count(se):
-                selist.append(se)
-        return selist
-
-    def findapi(self, key, site):
-        """
-        Find SiteDB api corresponding to given key
-        """
-        api  = ""
-        args = {}
-        for api, val in self.map.items():
-            keys = val['keys']
-            args = val['params']
-            if  keys.count(key):
-                for k in args.keys():
-                    if  api == 'SEtoCMSName':
-                        args['name'] = self.site2se(site)
[…]

### 2010-05-04 src/python/DAS/utils/utils.py
Adjust dict merging. Now values are added as single items in a list.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@394 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 General set of useful utilities used by DAS
 """
 
-__revision__ = "$Id: utils.py,v 1.24 2009/09/01 01:42:47 valya Exp $"
-__version__ = "$Revision: 1.24 $"
+__revision__ = "$Id: utils.py,v 1.25 2009/09/01 18:21:31 valya Exp $"
+__version__ = "$Revision: 1.25 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -31,13 +31,18 @@ class dict_of_none (dict):
 def dict_value(dict, prim_key):
     """
     Find value in given dictionary for given primary key. The primary
-    key can be composed one, e.g. a.b.
+    key can be composed one, e.g. a.b. If found value is a list type
+    take first element of the list and look-up over there the primary key.
     """
     try:
         if  prim_key.find('.') != -1:
             value = dict
             for key in prim_key.split('.'):
-                value = value[key]
+                if  type(value) is types.ListType:
+                    value = value[0][key]
+                    break
+                else:
+                    value = value[key]
         else:
             value = dict[prim_key]
         return value
@@ -45,14 +50,28 @@ def dict_value(dict, prim_key):
         msg  = 'Unable to look-up key=%s\n' % prim_key
         msg += 'dict=%s' % str(dict)
         raise Exception(msg)
+
 def merge_dict(dict1, dict2):
     """
     Merge content of two dictionaries w/ default list value for keys.
-    """
+    Original code was:
     merged_dict = {}
     for dictionary in [dict1, dict2]:
         for key, value in dictionary.items():
             merged_dict.setdefault(key,[]).append(value) 
+    return merged_dict
+    where I changed append(value) on merging lists or adding new
+    value into the list based on value type.
+    """
+    merged_dict = {}
+    for dictionary in [dict1, dict2]:
+        for key, value in dictionary.items():
+            dict_value = merged_dict.setdefault(key, [])
+            if  type(value) is types.ListType:
+                dict_value += value
+            else:
+                dict_value += [value]
+            merged_dict[key] = dict_value
     return merged_dict
 
 def splitlist(ilist, nentries):

### 2010-05-04 test/utils_t.py

@@ -36,6 +36,17 @@ class testUtils(unittest.TestCase):
         result = dict_value(dict, 'd')
         expect = 2
         self.assertEqual(expect, result)
+
+        dict = {'a' : [{'b':1}, {'b':2}]}
+        result = dict_value(dict, 'a.b')
+        expect = 1
+        self.assertEqual(expect, result)
+        
+        dict1 = {'a' : [{'b':1, 'c':1}, {'b':2, 'c':2}]}
+        dict2 = {'a' : {'b':1, 'e':1}}
+        result = merge_dict(dict1, dict2)
+        expect = {'a': [{'c': 1, 'b': 1}, {'c': 2, 'b': 2}, {'b': 1, 'e': 1}]}
+        self.assertEqual(expect, result)
         
     def test_dasheader(self):
         """Test DAS header"""

### 2010-05-04 src/python/DAS/tools/das_mapping_db.py
A few adjustments to DAS mapping populator
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@395 4525493e-7705-40b1-a816-d608a930855b

@@ -4,10 +4,11 @@
 """
 DAS command line interface
 """
-__revision__ = "$Id: das_mapping_db.py,v 1.1 2009/09/01 01:42:47 valya Exp $"
-__version__ = "$Revision: 1.1 $"
+__revision__ = "$Id: das_mapping_db.py,v 1.2 2009/09/01 18:23:09 valya Exp $"
+__version__ = "$Revision: 1.2 $"
 __author__ = "Valentin Kuznetsov"
 
+import os
 import sys
 from optparse import OptionParser
 from DAS.core.das_core import DASCore
@@ -58,6 +59,11 @@ if __name__ == '__main__':
     logger = DASLogger(verbose=opts.debug, stdout=opts.debug)
     config = dict(logger=logger, verbose=opts.debug,
         mapping_db_engine= 'sqlite:///%s' % opts.dbfile)
+
+    if  os.path.isfile(opts.dbfile) and not opts.listapis and not opts.listkeys:
+        print "remove", opts.dbfile
+        os.remove(opts.dbfile)
+
     mgr = DASMapping(config)
 
     if  opts.listapis:
@@ -217,7 +223,7 @@ if __name__ == '__main__':
     # nodes API
     api = 'nodes'
     params = {'node':'*', 'noempty':''}
-    daskeys = dict(site='site.name', node='site.name')
+    daskeys = dict(site='site.name')
     api2das = [('node', 'site', "re.compile('([a-zA-Z0-9]+\.){2}')")]
     mgr.add_api(system, api, params, daskeys, api2das)
     # lfn2pfn API
@@ -235,6 +241,7 @@ if __name__ == '__main__':
     mgr.add_notation(system, 'files', 'nfiles')
     mgr.add_notation(system, 'events', 'nevents')
     mgr.add_notation(system, 'lfn', 'name')
+    mgr.add_notation(system, 'node', 'site')
     ##### END OF PHEDEX
 
     # SiteDB
@@ -271,6 +278,7 @@ if __name__ == '__main__':
 
     api = 'CMSNametoSE'
     params = {'name':''}
+#    daskeys = dict(site='site.se')
     daskeys = dict(site='site.name')
     api2das = [('name', 'site', "re.compile('^T[0-3]_')")]
     mgr.add_api(system, api, params, daskeys, api2das)
@@ -281,6 +289,7 @@ if __name__ == '__main__':
[…]

### 2010-05-04 src/python/DAS/web/DASSearch.py
Re-factor web methods
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@445 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 DAS web interface, based on WMCore/WebTools
 """
 
-__revision__ = "$Id: DASSearch.py,v 1.18 2009/09/09 18:48:41 valya Exp $"
-__version__ = "$Revision: 1.18 $"
+__revision__ = "$Id: DASSearch.py,v 1.19 2009/09/16 20:37:18 valya Exp $"
+__version__ = "$Revision: 1.19 $"
 __author__ = "Valentin Kuznetsov"
 
 # system modules
@@ -35,7 +35,7 @@ from WMCore.WebTools.Page import exposejson, exposexml, exposedasplist
 # DAS modules
 from DAS.core.das_core import DASCore
 from DAS.utils.utils import getarg
-from DAS.web.utils import urllib2_request
+from DAS.web.utils import urllib2_request, json2html
 
 import sys
 if sys.version_info < (2, 5):
@@ -54,7 +54,6 @@ def ajax_response(msg):
     page  = """<ajax-response>"""
     page += "<div>" + msg + "</div>"
     page += "</ajax-response>"
-    print page
     return page
 
 class DASSearch(TemplatedPage):
@@ -190,11 +189,29 @@ class DASSearch(TemplatedPage):
         page = self.templatepage('das_searchform', input=uinput, msg=msg)
         return page
 
-    def result(self, kwargs):
+    def nresults(self, kwargs):
         """
         invoke DAS search call, parse results and return them to
         web methods
         """
+        url     = self.cachesrv
+        uinput  = getarg(kwargs, 'input', '')
+        params  = {'query':uinput}
+        path    = '/rest/nresults'
+        headers = {"Accept": "application/json"}
+        result  = self.decoder.decode(
+        urllib2_request('GET', url+path, params, headers=headers))
+        if  type(result) is types.StringType:
+            data = json.loads(result)
+        else:
+            data = result
+        if  data['results']['status'] == 'success':
+            return data['results']['nresults']
+        self.send_request('POST', kwargs)
+        return 0
+
+    def send_request(self, method, kwargs):
+        "Send POST request to server with provided parameters"
         url     = self.cachesrv
         uinput  = getarg(kwargs, 'input', '')
         format  = getarg(kwargs, 'format', '')
@@ -204,70 +221,48 @@ class DASSearch(TemplatedPage):
         sdir    = getarg(kwargs, 'dir', 'asc')
         params  = {'query':uinput, 'idx':idx, 'limit':limit, 
[…]

### 2010-05-04 src/python/DAS/web/DASSearch.py
Added not_ready template
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@446 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 DAS web interface, based on WMCore/WebTools
 """
 
-__revision__ = "$Id: DASSearch.py,v 1.19 2009/09/16 20:37:18 valya Exp $"
-__version__ = "$Revision: 1.19 $"
+__revision__ = "$Id: DASSearch.py,v 1.20 2009/09/17 18:51:39 valya Exp $"
+__version__ = "$Revision: 1.20 $"
 __author__ = "Valentin Kuznetsov"
 
 # system modules
@@ -278,7 +278,7 @@ class DASSearch(TemplatedPage):
         total   = self.nresults(kwargs)
         if  not total:
             ctime   = (time.time()-time0)
-            page    = 'DAS processing your request, please try in a few moments ...'
+            page    = self.templatepage('not_ready')
             page    = self.page(form + page, ctime=ctime)
             return page
 
@@ -303,7 +303,9 @@ class DASSearch(TemplatedPage):
                     api    = item['api'][idx]
                     system = item['system'][idx]
                     key    = item['selection_keys'][idx]
-                    data   = row[key][jdx]
+                    data   = row[key]
+                    if  type(data) is types.ListType:
+                        data = data[jdx]
                     if  type(data) is types.ListType:
                         data = data[idx]
                     pad = ""
@@ -347,7 +349,9 @@ class DASSearch(TemplatedPage):
                     api    = item['api'][idx]
                     system = item['system'][idx]
                     key    = item['selection_keys'][idx]
-                    data   = row[key][jdx]
+                    data   = row[key]
+                    if  type(data) is types.ListType:
+                        data = data[jdx]
                     if  type(data) is types.ListType:
                         data = data[idx]
                     pad = ""
@@ -381,7 +385,7 @@ class DASSearch(TemplatedPage):
         if  not total:
             ctime   = (time.time()-time0)
             form    = self.form(uinput)
-            page    = 'DAS processing your request, please try in a few moments ...'
+            page    = self.templatepage('not_ready')
             page    = self.page(form + page, ctime=ctime)
             return page
 

### 2010-05-04 doc/paper.tex
Modify DAS pictures and other tweaks
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@447 4525493e-7705-40b1-a816-d608a930855b

@@
 Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
-Physics, promising to deliver a few PB of data each year. Scientists are
+Physics (HEP), promising to deliver a few PB of data each year. Scientists are
 facing another set of problems from distributed computing to multi-dimentional
@@
 RDMS into another DB where SQL queries can be placed to search desired
-data. While on another approaches of querying relational
-DB using keyword search algorithms has been proposed \cite{DBXplorer, QueryAnswer}.
-In former case, you still face with understanding the underlying schema and
-imposing relation condition in your query, while in another an exact match
-of provided keywords is expected. Even though in last case you can 
-intuitively explore information presented in databases, in some cases
-it is not enough. For example, how to deal with numbers? What number is represent?
-The row ID, the value of the field?
-
-Our users can be classified as proficient
+data. While another approaches of querying relational
+DB(s) using keyword search algorithms has been proposed recently, see
+\cite{DBXplorer, QueryAnswer}.
+In former case, you still face with understanding of the underlying schema and
+imposing relational conditions in your query, while queries can be
+expressed via simple keyword based search, even though exact match
+of provided keywords is expected. Although keyword based search is
+intuitive and easy to use, with respect to HEP domain it is not sufficient. 
+For example, how to deal with numbers? What a single number means in a user input?
+The row ID, the value of the field, the sum of files or their size? In other
+words a further step is required to understand up-coming request without
+sacrifies the output results match. Our users can be classified as proficient
 users who understand semantics of the problem and want precise answers
-for their queries, at the same time do not need or should not be worried
-about underlying schema of underlying databases. Moreover, some information
-can be represented in multiple DBs and our users want to place 
-queries across those DBs.
+for their queries. They are more interesting in precise answers to their
+questions rather ranking output.
 
@@
 We wanted to expand this approach and apply such queries across multiple
-data services. Here we discuss a new system, the Data Aggregation System (DAS),
+data services and different DB back-ends.
+
+Here we discuss a new system, the Data Aggregation System (DAS),
 developed in CMS collaboration to address this issue. We start with
-discussion of CMS data model and data-services. In section \ref{DAS}
+discussion of CMS data model and existing data-services, see \ref{DataModel}. 
+In section \ref{DAS}
 we provides a detailed description of underlying design and components
-used in this system. And represents results in section \ref{Results}.
+used in DAS system. Our results are summarized in section \ref{Results}.
+
 \section{CMS data model\label{DataModel}}
+
 \section{Data Aggregation System\label{DAS}}
+Our goal was to create scalable and pluggable system for existing data-services
+within CMS computing infrastructure. We knew that all our data are delivered 
+via data-services who take care of their availability. Therefore, we treat 
+DAS as aggregation and cache service for existing data-services who should
+take care of the following:
+\begin{itemize}
+\item an individual data-service can have an own set of security policies;
+\item a dynamic nature of data-service APIs, due to continiously changing
+detector and software environments;
+\item a broad variety of data-formats;
+\item distributed nature of data-services;
+\item heterogeneous software environment.
+\item support flexible and intuitive Query Language (QL) to easy look-up our
+data;
+\item pluggable interface to existing and up-coming data-service APIs;
+\end{itemize}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=150mm]{DAS_architecture.pdf}
+\caption{
+DAS architecture.
+}
+\label{DAS_arch}
+\end{figure}
+
+Based on this set of constrains and requirements we come up with the following
+DAS components, see Figure \ref{DAS_arch}:
+\begin{itemize}
+\item DAS core system which is support pluggable modules to retrieve data from
+CMS data-services, e.g. Data Bookkeeping System (DBS), data location service (Phedex),
+etc.;
+\item DAS caching system which has dual purpose: to store ``raw'' results coming
+out from data-services, a raw-cache, and aggregated results produced by DAS core into
+hot-cache;
+\item DAS request service to handle user requests
+\item DAS mapping DB to store mapping between data-services and DAS notations and
+data-representation;
+\item DAS analytics DB to collect query statistics and establish pre-fetching 
+strategies.
+\end{itemize}
+We will discuss individual components in more details in up-coming sub-sections.
+
+\subsection{Dealing with data-services}
+The DAS supports a pluggable approach to add a new CMS sub-system into a game by
+identifying appropriate APIs, input and output parameters and keys, data-service
+format and mapping between data-service, DAS and other sub-system. For instance,
+the Data Bookkeeping System (DBS) and data location system (Phedex) store information about
+a file. In former case, a file name was stored as well as its relations to other physics 
+objects, such as run and luminosity, while in later case the Phedex store information 
+about file location. When user specify a file as their selection key, based on
+information store in DAS mapping DB we were able to identify appropriate set of
+data-service APIs, look-up data, re-format their output to DAS notations and
+store it into DAS raw-cache. Even though we support a data retrieval on demand
+basis we also store user queries into DAS analytics DB to identify
+a common patterns. Based on this information we organized a pre-fetching
+populators. Once data were coming to raw-cache an additional step of
+aggregation was applied. If other data with identical keys were presented in
+a raw-cache we merge it them together. For instance upon the following query
+$find file where ... $ a file information from DBS and Phedex systems were
+retrieved. Since both data-services output contained file name as a common
+key, their output where merged and final result was stored into raw-cache.
+This merged (aggregated) results were shipped back to client and stored into
+hot-cache (memcached) to speed up a DAS web client application. Each DAS data
+object contains a standard DAS header, which identify look-up time,
+request url, API, parameters and method as well as response version, expiration
+time and checksum of the output.
+
+We designed DAS interface to be fully REST complaint \cite{REST}. It is not
+only simplify development of DAS clients (CLI, web interface, populator, etc.)
+but also consider DAS itself as own sub-component, i.e. place queries against
+DAS content, rather then data-service one.
+
+\subsection{Query Language\label{DAS-QL}}
+Upon success of DBS-QL \cite{DBS-QL} we adopted its syntax to DAS with a few minor
+changes. Here we briefly recall its syntax:
+\begin{equation}
+action\,\,\,
+key_1,\, key_2,\, ...\,\,\, where\,\,\,
+\langle key\rangle\, 
+\langle op\rangle\, 
+\langle value\rangle \,\,\, and|or ...
+\label{QL_syntax}
+\end{equation}
+Each DAS QL expression starts with special keyword indicating an action such as 
+$find,\, plot,\, summary$, followed by a set of selection keys. The action identify
+the behavior of the query, find to look-up data, plot to plot found data, summary to
+apply a summary view over provided selection keys. An optional conditions can be 
+specified via $where$ keyword following by condition expression.
+Each condition expression represented via $key\,\,\, operator\,\,\, value$ triplet, where
+$key$ is DAS selection key, $operator$ is conditional operator, 
+$=,\, >=,\, <,\, in,\, between$ and $value$ is a value applied to selection key.
+We also support brackets as well as $and$ and $or$ logical operators.
+The DAS selection and condition keys were identified based on analysis 
+of terminology used by our users. For instance, $file$ was used to either
+select file name or used in condition to identify file name provided by user.
+Here is a simple example of DAS QL expression:
+$$
+find\,\,\, file,\, run\,\,\, where\,\,\, site=T1\_CH\_CERN
+$$
+which represents a natural language question: 
+{\it find files and runs located at T1\_CH\_CERN site}.
+
+\subsection{Mapping}
+As been discussed above we used a DAS mapping DB to map data-service API into
+DAS as well as map provide a relations between different data-services.
+Upon identification of APIs which participate in DAS activity 
+we store a mapping between data-service notations
+into DAS mapping DB. This allows to map user input query parameters into
+data-service API metrics as well as map back data-service output into DAS names.
+For instance, the DBS system used $logical\_file\_name$ notation, while in Phedex
+the same entity was named as $lfn$. Since both of them were referred to
+logical file name, we map both notations into $file$ notation in DAS.\footnote{We
+also deal with physical file names, but still used $file$ to refer to them in
+DAS notation, since simple regular expressions were used to identify $file$
+meaning from the provided value.}
+For each service we provide a mapping from its own data-format (JSON, XML, etc.) into
+common data-format used in DAS.\footnote{We used JSON as a DAS data format}
+We also store a mapping between DAS keywords and data-service API names to
+establish appropriate calls to data-service upon provided DAS query. For example,
+when user specifies $file$ in their DAS query, we invoke DBS listFiles and Phedex
+fileReplicas APIs. Since all mapping kept in RDMS we were able to abstract the 
+DAS core library calls from hardcoded data-service API calls and make it
+transparent to any participating data-services. Here is a logic of DAS core
+library:
+\begin{itemize}
+\item parse input DAS query
+\item map provided selection and condition keys into data-service APIs and
+input parameters
+\item invoke data-service APIs
+\item parse output results by mapping output notation into DAS keys
+\item store and aggregate them into DAS raw-cache
+\end{itemize}
+
+\subsection{Caching system and Analytics}
+The DAS caching system was divided into two different caches: raw-cache and hot-cache.
+The purpose of raw-cache was to store results coming from data services for
+as long as specified in data-service response header. While the purpose of hot-cache
+was to keep final DAS results corresponding to provided user query. We use a populator
+services (jobs) to feed raw-cache with highly requested queries. Here we will only
+discuss the raw-cache organization. 
+\begin{figure}[htb]
+\centering
+\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
+\caption{
+DAS cache system.
+}
+\label{DAS_cache}
+\end{figure}
+As it shown in figure \ref{DAS_cache} the DAS cache server has dual purpose.
+On one side it talks to DAS Analytics DB to trigger jobs for worker robot
+who update cache with most popular queries, on another it talks to DAS
+raw-cache back-end to either stored mapped results or merge them with 
+existing DAS objects in a cache. We should note that we use on-demand
+approach to answer user queries, but at the same time with a help of
+DAS Analytics DB we force worker robots (populators) to pre-fetch
+the most popular queries up-front.
+
+We evaluated several technologies
+as a choice for DAS cache back-end, including relational MySQL DB \cite{MySQL} 
+and object oriented CouchDB \cite{CouchDB} and MongoDB \cite{MongoDB}.
+Each of them had its own pros and cons. For instance usage
+of MySQL provided a SQL queries (as in federated DBs, see \cite{FedDB}), but
+required auto-schema generation on a fly.\footnote{We did not apply any
+restriction to DAS QL and did not know in advance what our users will be
+querying.} While, using CouchDB and MongoDB, we were able to store
+data-service output naturally as documents objects into DB. Finally we decided
+to use MongoDB due to its flexible and
+powerful querying syntax, which were easy to map into DAS QL expressions.
+
 \section{Results\label{Results}}
+DAS was deployed with the following CMS data-services:
+\begin{itemize}
+\item DBS \cite{DBS}, Data Bookkeeping System which collect information
+about CMS meta-data;
+\item PhEDEx \cite{Phedex}, Physics Experiment Data Export project which
+provides the data placement and the file transfer system for the CMS experiment;
+\item RunSummary DB \cite{RunSummary} collects information about run and triggers
+conditions during data-taking;
+\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
+\item DQ \cite{DQ}, Data Quality data-service which collects infromation
+about detector conditions during data taking;
+\item Overview \cite{Overview}, collects information about CMS
+transfer rate, etc.;
+\item Dashboard \cite{Dashboard} project for LHC experiments aims to 
+provide a single entry point to the monitoring data collected from the 
+distributed computing systems of the LHC virtual organization.
+\end{itemize}
+Each data-service has its own scope, size and lifetime. For instance the data
+in Phedex are transient due to contant migration of CMS data. The DBS system
+were divided into dozen of individual DBs, whose size varied at the level of
+few tens of GBs.
 
@@
 ``A multi-dimensional view on information retrieval of CMS data'', CHEP 2007
+\bibitem{DBS-QL} V. Kuznetsov, D. Riley, A. Afaq, V. Sekhri, Y. Guo, L. Lueking,
+``The CMS DBS Query Language'', CHEP 2009
 \bibitem{DD} https://cmsweb.cern.ch/dbs\_discovery
@@
 
+\bibitem{MySQL}
+http://www.mysql.com/
+
 \bibitem{CouchDB}
@@
 
+\bibitem{REST}
+Fielding, Roy Thomas ``Architectural Styles and the Design of 
+Network-based Software Architectures'', Doctoral dissertation, 2000,
+University of California, Irvine
+
+\bibitem{Phedex}
+Rehn et. al.,
+``PhEDEx high-throughput data transfer management system'', CHEP06, Mumbai, India.
+
+\bibitem{RunSummary}
+
+\bibitem{SiteDB}
+
+\bibitem{LumiDB}
+https://twiki.cern.ch/twiki/bin/view/CMS/CMS-DMWM-DBS-Luminosity
+
+\bibitem{DQ}
+
+\bibitem{Overview}
+https://cmsweb.cern.ch/overview/
+
+\bibitem{Dashboard}
+J. Andreeva, et. al,
+``Experiment Dashboard – The Monitoring System for the LHC Experiments'',
+GMW’07, June 25, 2007, Monterey, California, USA.
+
 \end{thebibliography}

### 2010-05-04 src/python/DAS/core/das_analytics_db.py
Mongo has a bug in update/inc, so it can't increment counter for all records at once. Bug suppose to be fixed in next release. Meanwhile I get all records and update one by one.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@448 4525493e-7705-40b1-a816-d608a930855b

@@ -7,8 +7,8 @@ DAS analytics DB
 
 from __future__ import with_statement
 
-__revision__ = "$Id: das_analytics_db.py,v 1.7 2009/09/14 20:34:32 valya Exp $"
-__version__ = "$Revision: 1.7 $"
+__revision__ = "$Id: das_analytics_db.py,v 1.8 2009/09/18 00:10:07 valya Exp $"
+__version__ = "$Revision: 1.8 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -78,7 +78,15 @@ class DASAnalytics(object):
         msg = 'DASAnalytics::update(%s, %s)' % (system, query)
         self.logger.info(msg)
         cond = {'query':query, 'system':system}
-        self.col.update(cond, {'$inc' : {'counter':1}})
+        # TODO:
+        # MongoDB has a bug, http://jira.mongodb.org/browse/SERVER-268 
+        # which prevent from bulk update of all records, uncomment the
+        # line below once it's fixed
+#        self.col.update(cond, {'$inc' : {'counter':1}})
+        # meanwhile we'll retrieve all records and update them individually
+        for record in self.col.find(cond):
+            self.col.update({'_id':record['_id']}, {'$inc' : {'counter':1}})
+        
 
     def list_systems(self):
         """

### 2010-05-04 src/python/DAS/core/das_core.py
Can be initialized with external config file, rather reading default one. Can be useful in populators
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@449 4525493e-7705-40b1-a816-d608a930855b

@@ -12,8 +12,8 @@ combine them together for presentation layer (CLI or WEB).
 
 from __future__ import with_statement
 
-__revision__ = "$Id: das_core.py,v 1.33 2009/09/16 20:34:28 valya Exp $"
-__version__ = "$Revision: 1.33 $"
+__revision__ = "$Id: das_core.py,v 1.34 2009/09/18 14:06:29 valya Exp $"
+__version__ = "$Revision: 1.34 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -51,10 +51,13 @@ class DASCore(object):
     service_keys = {'service':{list of keys]}
     service_maps = {('service1', 'service2'):'key'}
     """
-    def __init__(self, debug=None):
-        dasconfig    = das_readconfig()
-        verbose      = dasconfig['verbose']
-        self.stdout  = debug
+    def __init__(self, config=None, debug=None):
+        if  config:
+            dasconfig = config
+        else:
+            dasconfig = das_readconfig()
+        verbose       = dasconfig['verbose']
+        self.stdout   = debug
         if  type(debug) is types.IntType:
             self.verbose = debug
             dasconfig['verbose'] = debug
@@ -170,6 +173,7 @@ class DASCore(object):
         self.das_aggregation = {} # determine at run-time
         if  self.verbose:
             self.timer.record('DASCore.__init__')
+        self.dasconfig = dasconfig
 
     def keys(self):
         """

### 2010-05-04 src/python/DAS/core/das_robot.py
First version of DAS populator class and its main counterpart
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@450 4525493e-7705-40b1-a816-d608a930855b

@@ -1,0 +1,192 @@
+#!/usr/bin/env python
+#-*- coding: ISO-8859-1 -*-
+
+"""
+DAS robot (populator) class. Code based on
+http://www.jejik.com/articles/2007/02/a_simple_unix_linux_daemon_in_python/
+"""
+
+__revision__ = "$Id: das_robot.py,v 1.1 2009/09/18 14:08:11 valya Exp $"
+__version__ = "$Revision: 1.1 $"
+__author__ = "Valentin Kuznetsov"
+
+import os
+import sys
+import time
+import atexit
+from signal import SIGTERM 
+
+# DAS modules
+from DAS.utils.utils import genkey, getarg
+from DAS.core.das_core import DASCore
+
+class Robot(object):
+    """
+    DAS Robot (daemon) class to fetch data from provided URL/API
+    and store them into DAS cache.
+    """
+    def __init__(self, query=None, sleep=600):
+        self.dascore     = DASCore()
+        config           = self.dascore.dasconfig
+        config['stdout'] = '/tmp/das-robot.txt'
+        config['stderr'] = '/tmp/das-robot.txt'
+        logdir       = getarg(config, 'logdir', '/tmp')
+        self.pidfile = os.path.join(logdir, 'robot-%s.pid' % genkey(query))
+
+        if (hasattr(os, "devnull")):
+            devnull = os.devnull
+        else:
+            devnull = "/dev/null"
+
+        self.stdin   = devnull
+        self.stdout  = getarg(config, 'stdout', devnull)
+        self.stderr  = getarg(config, 'stderr', devnull)
+        self.query   = query
+        self.sleep   = sleep
+
+    def daemonize(self):
+        """
+        do the UNIX double-fork magic, see Stevens' "Advanced 
+        Programming in the UNIX Environment" for details (ISBN 0201563177)
+        http://www.erlenstar.demon.co.uk/unix/faq_2.html#SEC16
+        """
+        try: 
+            pid = os.fork() 
+            if  pid > 0:
+                # exit first parent
+                sys.exit(0) 
+        except OSError, e: 
+            sys.stderr.write("fork #1 failed: %d (%s)\n" % (e.errno, e.strerror))
+            sys.exit(1)
+
+        # decouple from parent environment
+        os.chdir("/") 
+        os.umask(0) 
+        os.setsid() 
+
+        # do second fork
+        try: 
+            pid = os.fork() 
+            if  pid > 0:
+                # exit from second parent
[…]

### 2010-05-04 src/python/DAS/core/basemanager.py
Remove obsolete code
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@453 4525493e-7705-40b1-a816-d608a930855b

@@ -1,39 +1,0 @@
-#!/usr/bin/env python
-#-*- coding: ISO-8859-1 -*-
-
-"""
-Abstract manager class.
-"""
-
-__revision__ = "$Id: basemanager.py,v 1.3 2009/05/27 20:28:03 valya Exp $"
-__version__ = "$Revision: 1.3 $"
-__author__ = "Valentin Kuznetsov"
-
-class BaseManager(object):
-    """
-    Base manager class used in DAS. It defines result, keys
-    method used by DAS core.
-    """
-    def __init__(self, dasconfig):
-        self.verbose        = dasconfig['verbose']
-        self.logger         = dasconfig['logger']
-        self.cache_servers  = dasconfig['cache_servers']
-        self.cache_lifetime = dasconfig['cache_lifetime']
-        self.couch_servers  = dasconfig['couch_servers']
-        self.couch_lifetime = dasconfig['couch_lifetime']
-        self.filecache_dir  = dasconfig['filecache_dir']
-        self.filecache_lifetime = dasconfig['filecache_lifetime']
-        self.service_keys   = {}
-
-    def keys(self):
-        """
-        Base method to return service keys
-        """
-        return self.service_keys
-
-    def result(self, query=None):
-        """
-        Base method to return results of the query
-        """
-        print query
-        return

### 2010-05-04 test/basemanager_t.py

@@ -1,37 +1,0 @@
-#!/usr/bin/env python
-#pylint: disable-msg=c0301,c0103
-
-"""
-unit test for base manager module
-"""
-
-import unittest
-from DAS.utils.das_config import das_readconfig
-from DAS.core.basemanager import BaseManager
-from DAS.utils.logger import DASLogger
-
-class testBaseManager(unittest.TestCase):
-    """
-    a test class for the das base manager module
-    """
-    def setUp(self):
-        """
-        set up das core module
-        """
-        dasconfig = das_readconfig()
-        logger = DASLogger()
-        dasconfig['logger'] = logger
-        self.mgr = BaseManager(dasconfig)
-
-    def test_result(self):                          
-        """test base manager result method"""
-        query  = "find site where site=T2_UK"
-        result = self.mgr.result(query)
-        expect = None
-        self.assertEqual(expect, result)
-#
-# main
-#
-if __name__ == '__main__':
-    unittest.main()
-

### 2010-05-04 test/qlparser_t.py
Remove obsolete test
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@454 4525493e-7705-40b1-a816-d608a930855b

@@ -75,13 +75,13 @@ class testQLParser(unittest.TestCase):
 
     def test_qlparser(self):
         """test QLParser class"""
-        imap = {'dbs':['dataset', 'run', 'site', 'block'],
-                'phedex':['block', 'replica', 'site'], 
+        imap = {'dbs':['dataset', 'run', 'file', 'block'],
+                'phedex':['block', 'file', 'site'], 
                 'sitedb': ['admin', 'site'],
                 'dq': ['runs', 'DQFlagList'],
                 'dashboard': ['jobsummary']}
-        params = {'dbs':['dataset', 'run', 'site', 'block'], 
-                'phedex':['block', 'replica', 'site'], 
+        params = {'dbs':['dataset', 'run', 'file', 'block'], 
+                'phedex':['block', 'file', 'site'], 
                 'sitedb':['admin', 'site'], 
                 'dq':['run','dataset'],
                 'dashboard':['site']}
@@ -137,10 +137,11 @@ class testQLParser(unittest.TestCase):
                 'unique_services': ['dbs', 'phedex'], 'order_by_list': [], 
                 'order_by_order': None, 
                 'services': {'phedex': ['block'], 'dbs': ['block']}, 
+                'query' : 'find block where block=bla',
                 'dasqueries': {'phedex': ['find block where block = bla'],
                                'dbs': ['find block where block = bla']}, 
                 'conditions': [{'value': 'bla', 'key': 'block', 'op': '='}], 
-                'allkeys': ['block'], 'unique_keys': ['block', 'site']}
+                'allkeys': ['block'], 'unique_keys': ['block', 'file']}
         self.assertEqual(result, expect)
 
         q = "find dbs:block where block=bla"
@@ -150,9 +151,10 @@ class testQLParser(unittest.TestCase):
                 'dasqueries': {'dbs': ['find block where block = bla'], 
                                'phedex': ['find block where block = bla']}, 
                 'order_by_order': None, 
+                'query': 'find block where block=bla',
                 'services': {'dbs': ['block'], 'phedex': ['block']}, 
                 'conditions': [{'value': 'bla', 'key': 'block', 'op': '='}], 
-                'allkeys': ['block'], 'unique_keys': ['block', 'site']}
+                'allkeys': ['block'], 'unique_keys': ['block', 'file']}
         self.assertEqual(result, expect)
 
         q = "find jobsummary where date last 25h"
@@ -163,30 +165,6 @@ class testQLParser(unittest.TestCase):
 
[…]

### 2010-05-04 doc/paper.tex
Work on paper
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@456 4525493e-7705-40b1-a816-d608a930855b

@@
 
-\author{Simon Matson}
+\author{Simon Metson}
 \address{Bristol University, Bristol, UK}
@@
 \begin{abstract}
-A metadata plays significant role in a large modern enterprises, research experiments,
+A meta-data plays significant role in a large modern enterprises, research experiments,
 digital libraries where it comes from different sources and distributed in a 
 variety of forms and digital formats. It is organized and managed by constantly
-evolving software using both relational and nonrelational data sources. There is
-a big demand to access information from multiple sources.
-Here we discuss a new data aggregation system which deliver and cache information 
-from different relational and nonrelational data sources on a concrete example 
-of large scale High-Energy physics experiment.
+evolving software using both relational and non-relational data sources. There is
+a big demand to enable information discovery from multiple sources.
+Here we discuss a new data aggregation system which consume and deliver information 
+from different relational and non-relational data sources on a concrete example 
+of large scale, distributed system of CMS physics experiment at LHC.
 \end{abstract}
@@
 The CERN, the European Organization for Nuclear Research, plays a leading
-role in fundamental studies of physics. But apart from scientific community 
+role in fundamental studies of physics. Apart from scientific community 
 it is known as a place where World Wide Web was born. At that time the 
@@
 Physics (HEP), promising to deliver a few PB of data each year. Scientists are
-facing another set of problems from distributed computing to multi-dimentional
+facing another set of problems from distributed computing to multi-dimensional
 view on information retrieval. One of the aspect of such research is an efficient
 and at the same time precise look-up of produced meta-data, which comes in variety 
-of forms and digital formats. As was pointed out in \cite{Amr} a mixed content and 
-mixed metadata and metadata consistency should be considered as a whole in design 
+of forms and digital formats. As was pointed out in \cite{Arms} a mixed content and 
+mixed meta-data and meta-data consistency should be considered as a whole in design 
 of the system to successful information discovery.
@@
 represents an heterogeneous environment of distributing computing, relational and
-nonrelational data sources where we face this problem. The computing resources
-are distributed among almost 40 countries, 183 institutions and more then 3000 physicists.
+non-relational data sources where we are facing this problem. The computing resources
+are distributed among almost 40 countries, 183 institutions and more than 3000 physicists.
 Broad variety of RDMS systems at different data centers collect a various
-metadata information, including data location and transfer, detector conditions,
+meta-data information, including data location and transfer, detector conditions,
 calibrations constants, simulation information, etc. Moreover the development
@@
 tools. Therefore a wide spectrum of data services and
-data formats represents a problem of information discovery.
+data formats represents a challenge of efficient information discovery.
 
-\section{Related Work}
+Here we present a work on Data Aggregation System (DAS), the system designed
+to provide ability to query, search and aggregate information from different 
+data-services as well as serves as a caching layer for them, 
+preserving their security policies. We organize
+our discussion as following. In section \ref{RelatedWork} we discuss
+related work in domain of keyword search over relation data sources.
+The section \ref{DataModel} provides review of CMS data model. In section
+\ref{DAS} we outline architecture of the DAS system, including discussion of its
+various components. Finally our results are summarized in section \ref{Results}.
+
+\section{Related Work\label{RelatedWork}}
 Even though the idea of querying relation databases via keyword based search
-algorithm is not knew it is still under significant activity in computer
-sceince domain. A few
-alternative solutions has been proposed to address this issue. On one
-side, the federated DB \cite{FedDB} unify data coming from different
-RDMS into another DB where SQL queries can be placed to search desired
-data. While another approaches of querying relational
-DB(s) using keyword search algorithms has been proposed recently, see
-\cite{DBXplorer, QueryAnswer}.
-In former case, you still face with understanding of the underlying schema and
-imposing relational conditions in your query, while queries can be
-expressed via simple keyword based search, even though exact match
+algorithms is not knew it is still under significant activity in computer
+science domain. A few alternative solutions has been proposed to address this issue
+in last several years. The federated DB \cite{FedDB} by IBM unifies data coming 
+from different RDMS into another DB where SQL queries can be placed to search desired
+data. The other approaches \cite{DBXplorer, QueryAnswer}, influenced by great success
+of search engines, explored ability to use keyword search algorithms over relataional RDMS
+by indexing DB content. In former case, you still face a problem with understanding 
+of the underlying schema and imposing relational conditions in your query, while 
+in later queries can be expressed via simple keyword based search, even though exact match
 of provided keywords is expected. Although keyword based search is
-intuitive and easy to use, with respect to HEP domain it is not sufficient. 
-For example, how to deal with numbers? What a single number means in a user input?
-The row ID, the value of the field, the sum of files or their size? In other
-words a further step is required to understand up-coming request without
-sacrifies the output results match. Our users can be classified as proficient
-users who understand semantics of the problem and want precise answers
-for their queries. They are more interesting in precise answers to their
-questions rather ranking output.
-
-In \cite{DBS-QL} a simple, intuitive and flexible query language was introduced 
-on top of the CMS data-bookkeeping system. It represented a power of SQL while
+intuitive and easy to use, with respect to HEP and other domains it is not sufficient. 
+For example, it does provide provide ability to place conditions expressions between
+keys (DB entities/attributes) and their values. The proximity of results becomes very poor
+or impossible in case of numbers, aggregate functions, etc. To address this issues
+an attempt to build a simple, intuitive and flexible query language was introduced
+in \cite{DBS-QL}. It represented a power of SQL while
 hiding underlying relational schema from the end users. As a results
 a human questions were intuitively mapped into simple queries. For example,
-a question
+the question
 {\it I'm looking for files who contain data taken on certain date and located at
@@
 \end{verbatim}
-We wanted to expand this approach and apply such queries across multiple
-data services and different DB back-ends.
+We wanted to expand this approach to handle broad variety of relational and
+non-relational data-source in CMS within our distributed non-homogeneous environment.
 
-Here we discuss a new system, the Data Aggregation System (DAS),
-developed in CMS collaboration to address this issue. We start with
-discussion of CMS data model and existing data-services, see \ref{DataModel}. 
-In section \ref{DAS}
-we provides a detailed description of underlying design and components
-used in DAS system. Our results are summarized in section \ref{Results}.
+%Here we discuss a new system, the Data Aggregation System (DAS),
+%developed in CMS collaboration to address this issue. We start with
+%discussion of CMS data model and existing data-services, see \ref{DataModel}. 
+%In section \ref{DAS}
+%we provides a detailed description of underlying design and components
+%used in DAS system. Our results are summarized in section \ref{Results}.
 
 \section{CMS data model\label{DataModel}}
+The CMS distributed computing and data model \cite{CMSDataModel} 
+is designed to process
+and efficiently manage a PBs of data expected to be produced each year
+at LHC. The computing resources provided by members of CMS
+collaboration are geographically distributed, 
+interconnected via high throughput networks and operated by means 
+of Grid software. The model allows to cover broad variety of
+hardware, mass and storage elements and configuration of
+clusters. To accomodate the efficient data processing we use 
+a multi-Tier distributed model, where specific tasks of data taking,
+processing, archival and distribution assigned to each tier based
+on CMS data model. For example, the Tier-0 center at CERN responsible
+for data archives coming out of detector, prompt first pass reconstruction
+and data distribution to Tier-1 centers. The 7 Tier-1 centers
+located in France, Germany, Ialy, Spain, Taiwan, United Kingdom and United States
+keep a portion (copy) of data delivered by Tier-0 for further processing.
+They provide storage and high priority analysis jobs at their facilities.
+The Tier-2 centers, located at more then 50 sites around the world,
+are dedicated for user analysis tasks and production of simulated data.
+
+A broad variety of data-services being designed and developed to
+maintain detector and production operations, including detector
+conditions databases, data location and bookkeeping services,
+data transfer and job monitoring tasks. Even though majority of them
+are located at CERN, it was never been a requirement in CMS computing
+and data model. For instance, the production teams operated at Tier-1,2
+centers set up local services for data bookkeeping and operational
+tasks. Based on their budget resources the choice of back-end DB were
+up to site managers. Therefore we designed our software to support different
+DB back-ends and provide tools for migrating data across them.
+
+Once such conglomerate of data-services starts operating an obvious
+question arise: how to find out desired information across multiple data-services
+in our distributed environment? Even though individual data-services were designed
+to answer specific question about data they serve the ability to relate and search
+information among them was human task. The growing amount of information
+and desire to make cross-service queries force us to design and develop a new
+type of system, the Data Aggregation System.
 
 \section{Data Aggregation System\label{DAS}}
-Our goal was to create scalable and pluggable system for existing data-services
-within CMS computing infrastructure. We knew that all our data are delivered 
-via data-services who take care of their availability. Therefore, we treat 
-DAS as aggregation and cache service for existing data-services who should
-take care of the following:
+The design of DAS system was based on previous studies of CMS 
+data bookkeeping system (DBS) usage patterns \cite{DBS07}. We carefully analyzed user
+queries placed to DBS, their patterns, frequencies and latency. As a result of this
+we created a data discovery service \cite{DD} which came through several iteration
+of UI. At the end the end-user interface was based on DBS Query Language \cite{DBS-QL}
+with broad variety of template for presentation layer. A quick adoption
+and wide spread of DBS QL give us confidence of choosen approach and QL syntax.
+
+Starting from those ground accomplishments we created DAS system as
+scalable and pluggable layer sitting on top of existing data-services
+within CMS computing infrastructure. Data availability were deligated to individual
+data-services. Therefore, we treat DAS as aggregation and cache service 
+with the following characteristics:
 \begin{itemize}
-\item an individual data-service can have an own set of security policies;
-\item a dynamic nature of data-service APIs, due to continiously changing
-detector and software environments;
-\item a broad variety of data-formats;
-\item distributed nature of data-services;
-\item heterogeneous software environment.
 \item support flexible and intuitive Query Language (QL) to easy look-up our
@@
 \item pluggable interface to existing and up-coming data-service APIs;
+\item support heterogeneous software environment and distributed nature of data-services;
+\item preserve security policies of individual data-services while provide unique
+authentication schema for end-users;
+\item be flexible to adjust to dynamic nature of data-service APIs, 
+due to continuously changing detector and software environments;
+\item be transparent to data-service data formats.
 \end{itemize}
 
+%\begin{figure}[htb]
+%\centering
+%\includegraphics[width=150mm]{DAS_architecture.pdf}
+%\caption{
+%DAS architecture.
+%}
+%\label{DAS_arch}
+%\end{figure}
 \begin{figure}[htb]
 \centering
-\includegraphics[width=150mm]{DAS_architecture.pdf}
+\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
 \caption{
-DAS architecture.
+DAS architecture diagram. It consists of DAS cache server, mapping and analytics DBs.
+Once query was placed to DAS an associated
+data-service API was invoked, triggering in DAS cache population. Once data become
+available in DAS cache user were able to see the results. This so-called on demand
+query retrieval system, where data were retrieved and store into cache from
+back-end data service. We use DAS analytics DB to establish cache population
+for most common queries by using DAS robots.
 }
-\label{DAS_arch}
+\label{DAS_cache}
 \end{figure}
 
-Based on this set of constrains and requirements we come up with the following
-DAS components, see Figure \ref{DAS_arch}:
+\noindent
+Figure \ref{DAS_cache} summarize a overview of DAS architecture who consist of the
+following components:
 \begin{itemize}
-\item DAS core system which is support pluggable modules to retrieve data from
-CMS data-services, e.g. Data Bookkeeping System (DBS), data location service (Phedex),
-etc.;
-\item DAS caching system which has dual purpose: to store ``raw'' results coming
-out from data-services, a raw-cache, and aggregated results produced by DAS core into
-hot-cache;
-\item DAS request service to handle user requests
-\item DAS mapping DB to store mapping between data-services and DAS notations and
-data-representation;
-\item DAS analytics DB to collect query statistics and establish pre-fetching 
+\item DAS core library with support of pluggable modules for data retrieval;
+\item DAS caching layer with long-term cache for aggregated results and short-term
+cache for presentation layer;
+\item DAS request service to handle user requests;
+\item DAS mapping DB with DAS/data-services mapping, notations mapping and
+mapping of DAS objects into data-presentation layer;
+\item DAS analytics DB with query statistics to enable pre-fetching 
 strategies.
 \end{itemize}
-We will discuss individual components in more details in up-coming sub-sections.
+Further we proceed with a discussion of individual DAS components.
 
 \subsection{Dealing with data-services}
-The DAS supports a pluggable approach to add a new CMS sub-system into a game by
-identifying appropriate APIs, input and output parameters and keys, data-service
-format and mapping between data-service, DAS and other sub-system. For instance,
-the Data Bookkeeping System (DBS) and data location system (Phedex) store information about
-a file. In former case, a file name was stored as well as its relations to other physics 
-objects, such as run and luminosity, while in later case the Phedex store information 
-about file location. When user specify a file as their selection key, based on
-information store in DAS mapping DB we were able to identify appropriate set of
+DAS provided pluggable system to add a new CMS data-service. To do that we
+identify appropriate APIs, input and output parameters and keys, data-service
+format and mapping between data-service, DAS and other sub-systems. 
+This information were stored into DAS mapping DB. For instance,
+the Data Bookkeeping System (DBS) and data location system (Phedex) both
+provide information about CMS data files. In former case, DBS keeps information between
+files and other physics  objects, such as run and luminosity, while Phedex provides 
+information about file location, such as site and details about storage elements.
+Upon user request to find information about a file, based on forementioned
+information we were able to identify appropriate set of
 data-service APIs, look-up data, re-format their output to DAS notations and
-store it into DAS raw-cache. Even though we support a data retrieval on demand
+store results into DAS cache.\footebnote{Even though we supported a data retrieval on demand
 basis we also store user queries into DAS analytics DB to identify
 a common patterns. Based on this information we organized a pre-fetching
-populators. Once data were coming to raw-cache an additional step of
-aggregation was applied. If other data with identical keys were presented in
-a raw-cache we merge it them together. For instance upon the following query
-$find file where ... $ a file information from DBS and Phedex systems were
+strategies.}
+Once data were retreived from data-services an additional step to aggregate
+matched objects in DAS cache were applied. For instance in above examples
+file information from DBS and Phedex systems were
 retrieved. Since both data-services output contained file name as a common
@@
 We designed DAS interface to be fully REST complaint \cite{REST}. It is not
-only simplify development of DAS clients (CLI, web interface, populator, etc.)
+only simplify development of DAS clients (CLI, web interface, robot, etc.)
 but also consider DAS itself as own sub-component, i.e. place queries against
@@
 fileReplicas APIs. Since all mapping kept in RDMS we were able to abstract the 
-DAS core library calls from hardcoded data-service API calls and make it
+DAS core library calls from hard coded data-service API calls and make it
 transparent to any participating data-services. Here is a logic of DAS core
@@
 as long as specified in data-service response header. While the purpose of hot-cache
-was to keep final DAS results corresponding to provided user query. We use a populator
+was to keep final DAS results corresponding to provided user query. We use a robot
 services (jobs) to feed raw-cache with highly requested queries. Here we will only
 discuss the raw-cache organization. 
-\begin{figure}[htb]
-\centering
-\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
-\caption{
-DAS cache system.
-}
-\label{DAS_cache}
-\end{figure}
+%\begin{figure}[htb]
+%\centering
+%\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
+%\caption{
+%DAS cache system.
+%}
+%\label{DAS_cache}
+%\end{figure}
 As it shown in figure \ref{DAS_cache} the DAS cache server has dual purpose.
@@
 approach to answer user queries, but at the same time with a help of
-DAS Analytics DB we force worker robots (populators) to pre-fetch
+DAS Analytics DB we force worker robots (robots) to pre-fetch
 the most popular queries up-front.
@@
 \item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
-\item DQ \cite{DQ}, Data Quality data-service which collects infromation
+\item DQ \cite{DQ}, Data Quality data-service which collects information
 about detector conditions during data taking;
@@
 Each data-service has its own scope, size and lifetime. For instance the data
-in Phedex are transient due to contant migration of CMS data. The DBS system
+in Phedex are transient due to constant migration of CMS data. The DBS system
 were divided into dozen of individual DBs, whose size varied at the level of
@@
 
-\section{Acknowledgements}
+\section{Acknowledgments}
 
@@
 \begin{thebibliography}{9}
-\bibitem{DBS} A. Afaq, et. al. ``The CMS Dataset Bookkeeping Service'', CHEP 2007 
-\bibitem{DBS07} A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
-``A multi-dimensional view on information retrieval of CMS data'', CHEP 2007
-\bibitem{DBS-QL} V. Kuznetsov, D. Riley, A. Afaq, V. Sekhri, Y. Guo, L. Lueking,
-``The CMS DBS Query Language'', CHEP 2009
-\bibitem{DD} https://cmsweb.cern.ch/dbs\_discovery
-
 \bibitem{Arms}
@@
 chapter from ``Metadata in Practice'', ALA Editions, 2004
+
+\bibitem{FedDB}
+L. Haas, E. Lin,
+``IBM Federated Database Technology'', \\
+http://www.ibm.com/developerworks/data/library/techarticle/0203haas/0203haas.html
+
 \bibitem{DBXplorer}
@@
 
-\bibitem{FedDB}
-L. Haas, E. Lin,
-``IBM Federated Database Technology'', \\
-http://www.ibm.com/developerworks/data/library/techarticle/0203haas/0203haas.html
+\bibitem{CMSDataModel} A. Fanfani et. al.,
+``Distributed Analysis in CMS'', to be published in Journal of Grid Computing.
+
+\bibitem{DBS-QL} V. Kuznetsov, D. Riley, A. Afaq, V. Sekhri, Y. Guo, L. Lueking,
+``The CMS DBS Query Language'', CHEP 2009
+
+\bibitem{DBS} A. Afaq, et. al. ``The CMS Dataset Bookkeeping Service'', CHEP 2007
+
+\bibitem{DBS07} A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
+``A multi-dimensional view on information retrieval of CMS data'', CHEP 2007
+
+\bibitem{DD} https://cmsweb.cern.ch/dbs\_discovery
 
@@
 
+\bibitem{DBSearch}
+D. Konopnicki, O. Shmueli,
+``Database-Inspired Search'', 
+Proc. of the 31st VLDB Conference, 2005.
 \end{thebibliography}

### 2010-05-04 doc/paper.tex
Finish the sentence before previous commit
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@457 4525493e-7705-40b1-a816-d608a930855b

@@
 matched objects in DAS cache were applied. For instance in above examples
-file information from DBS and Phedex systems were
-retrieved. Since both data-services output contained file name as a common
-key, their output where merged and final result was stored into raw-cache.
-This merged (aggregated) results were shipped back to client and stored into
-hot-cache (memcached) to speed up a DAS web client application. Each DAS data
+file information came from DBS and Phedex systems. 
+Since both data-services output contained file name as a common
+key, their output where merged and final result was stored into DAS cache.
+This merged (aggregated) results represented a DAS data object. Each DAS data
 object contains a standard DAS header, which identify look-up time,

### 2010-05-04 src/python/DAS/tools/das_mapping_db.py
Correct patterns for DAS keywords, added presentation record to be used for web UI
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@460 4525493e-7705-40b1-a816-d608a930855b

@@ -4,8 +4,8 @@
 """
 DAS command line interface
 """
-__revision__ = "$Id: das_mapping_db.py,v 1.5 2009/09/14 20:39:08 valya Exp $"
-__version__ = "$Revision: 1.5 $"
+__revision__ = "$Id: das_mapping_db.py,v 1.6 2009/09/29 20:45:29 valya Exp $"
+__version__ = "$Revision: 1.6 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -77,6 +77,14 @@ if __name__ == '__main__':
         keys = mgr.list_daskeys(opts.system)
         print keys
         sys.exit(0)
+    # daskeys defines a mapping between daskey used in DAS-QL and DAS record key
+    # e.g. block means block.name, the pattern defines regular expression to which
+    # daskey should satisfy
+    #
+    # api2das defines mapping between API input parameter and DAS-QL key,
+    # dict(api_param='storage_element_name', das_key='site', 
+    #      pattern="re.compile('([a-zA-Z0-9]+\.){2}')"),
+    # here pattern is applied to passed api_param
 
     ##### DBS
     apiversion = 'DBS_2_0_8'
@@ -244,8 +252,10 @@ if __name__ == '__main__':
         'api' : dict(name=api, params=params),
         'daskeys' : [dict(key='block', map='block.name', pattern='')],
         'api2das' : [
-                dict(api_param='se', das_key='site', pattern="re.compile('^T[0-3]_')"),
-                dict(api_param='node', das_key='site', pattern="re.compile('([a-zA-Z0-9]+\.){2}')"),
+                dict(api_param='node', das_key='site', pattern="re.compile('^T[0-3]_')"),
+                dict(api_param='node', das_key='site.name', pattern="re.compile('^T[0-3]_')"),
+                dict(api_param='se', das_key='site', pattern="re.compile('([a-zA-Z0-9]+\.){2}')"),
+                dict(api_param='se', das_key='site.se', pattern="re.compile('([a-zA-Z0-9]+\.){2}')"),
                 dict(api_param='block', das_key='block', pattern=""),
         ]
     }
@@ -258,8 +268,10 @@ if __name__ == '__main__':
         'api' : dict(name=api, params=params),
         'daskeys' : [dict(key='file', map='file.name', pattern='')],
         'api2das' : [
-                dict(api_param='se', das_key='site', pattern="re.compile('^T[0-3]_')"),
[…]

### 2010-05-04 src/python/DAS/web/DASCacheModel.py
Work on presentation layer of how to represent DAS objects on a page
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@476 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 DAS cache RESTfull model, based on WMCore/WebTools
 """
 
-__revision__ = "$Id: DASCacheModel.py,v 1.14 2009/09/16 20:36:56 valya Exp $"
-__version__ = "$Revision: 1.14 $"
+__revision__ = "$Id: DASCacheModel.py,v 1.15 2009/10/02 15:31:29 valya Exp $"
+__version__ = "$Revision: 1.15 $"
 __author__ = "Valentin Kuznetsov"
 
 # system modules
@@ -103,12 +103,11 @@ def checkargs(func):
     wrapper.exposed = True
     return wrapper
 
-def worker(item):
+def worker(query, expire):
     """
     Worker function which invoke DAS core to update cache for input query
     """
     dascore = DASCore()
-    query, expire = item
     status  = dascore.update_cache(query, expire)
     return status
 
@@ -163,16 +162,19 @@ class DASCacheModel(RESTModel):
         data = {'server_method':'nresults'}
         if  kwargs.has_key('query'):
             query = kwargs['query']
+            query = self.dascore.mongoparser.dasql2mongo(query)
             data.update({'status':'success'})
             if  hasattr(self.dascore, 'cache'):
-                if  self.dascore.cache.incache(query):
-                    data['nresults'] = self.dascore.cache.nresults(query)
+                res = self.dascore.cache.nresults(query)
+                if  res:
+                    data['nresults'] = res
                 else:
                     data['status'] = 'not found'
             else:
-                if  self.dascore.in_raw_cache(query):
+                res = self.dascore.in_raw_cache_nresults(query)
+                if  res:
                     data['status'] = 'success'
-                    data['nresults'] = self.dascore.in_raw_cache_nresults(query)
+                    data['nresults'] = res
                 else:
                     data['status'] = 'not found'
         else:
@@ -190,6 +192,7 @@ class DASCacheModel(RESTModel):
         data = {'server_method':'request'}
         if  kwargs.has_key('query'):
             query = kwargs['query']
+            query = self.dascore.mongoparser.dasql2mongo(query)
             idx   = getarg(kwargs, 'idx', 0)
             limit = getarg(kwargs, 'limit', 0)
             skey  = getarg(kwargs, 'skey', '')
@@ -245,6 +248,7 @@ class DASCacheModel(RESTModel):
         data = {'server_method':'create'}
         if  kwargs.has_key('query'):
             query  = kwargs['query']
[…]

### 2010-05-04 doc/paper.tex
Updated abstract based on Simon suggestion
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@478 4525493e-7705-40b1-a816-d608a930855b

@@
 \begin{abstract}
-A meta-data plays significant role in a large modern enterprises, research experiments,
-digital libraries where it comes from different sources and distributed in a 
-variety of forms and digital formats. It is organized and managed by constantly
-evolving software using both relational and non-relational data sources. There is
-a big demand to enable information discovery from multiple sources.
-Here we discuss a new data aggregation system which consume and deliver information 
-from different relational and non-relational data sources on a concrete example 
-of large scale, distributed system of CMS physics experiment at LHC.
+%A meta-data plays significant role in a large modern enterprises, research experiments,
+%digital libraries where it comes from different sources and distributed in a 
+%variety of forms and digital formats. It is organized and managed by constantly
+%evolving software using both relational and non-relational data sources. There is
+%a big demand to enable information discovery from multiple sources.
+%Here we discuss a new data aggregation system which consume and deliver information 
+%from different relational and non-relational data sources on a concrete example 
+%of large scale, distributed system of CMS physics experiment at LHC.
+
+Meta-data plays signiﬁcant role in a large modern enterprises, 
+research experiments, digital libraries where it comes from different 
+sources and distributed in a variety of forms and digital formats. 
+It is organized and managed by constantly evolving software using 
+both relational and non-relational data sources. There is a big 
+demand to enable information discovery from multiple sources.
+
+Here we discuss a new data aggregation system which consumes, 
+indexes and delivers information from different relational and 
+non-relational data sources to answer cross data source queries 
+examining metadata associated to petabytes of 
+experimental data. We examine simplicity of keyword-based system 
+with precise answers of SQL queries under new system where 
+aggregated information being collected from various sources 
+allowing end-users to place dynamic queries and trigger information 
+retrieval on demand. Such close to real-time system based on studies 
+of use cases of the CMS particle physics experiment at the LHC which 
+uses a large scale, distributed computing system to motivate the work.
+
+%Here we discuss a new data aggregation system which consumes, 
+%indexes and delivers information from different relational and 
+%non-relational data sources to answer cross data source queries 
+%examining metadata associated to petabytes of 
+%experimental data. We follow the use cases of the CMS particle 
+%physics experiment at the LHC which uses a large scale, distributed 
+%computing system system to motivate the work.
+
 \end{abstract}
@@
 in our distributed environment? Even though individual data-services were designed
-to answer specific question about data they serve the ability to relate and search
+to answer specific questions about data they serve, the ability to relate and search
 information among them was human task. The growing amount of information
@@
 The design of DAS system was based on previous studies of CMS 
-data bookkeeping system (DBS) usage patterns \cite{DBS07}. We carefully analyzed user
+data bookkeeping system (DBS) usage patterns \cite{DBS, DBS07}. We carefully analyzed user
 queries placed to DBS, their patterns, frequencies and latency. As a result of this
 we created a data discovery service \cite{DD} which came through several iteration
-of UI. At the end the end-user interface was based on DBS Query Language \cite{DBS-QL}
+of UI. At the end the end-user interface was based on DBS Query Language (DBS-QL)\cite{DBS-QL}
 with broad variety of template for presentation layer. A quick adoption
-and wide spread of DBS QL give us confidence of choosen approach and QL syntax.
+and wide spread of DBS-QL give us confidence of choosen approach and QL syntax.
 
 Starting from those ground accomplishments we created DAS system as
-scalable and pluggable layer sitting on top of existing data-services
+scalable and pluggable layer sitting on top of the existing data-services
 within CMS computing infrastructure. Data availability were deligated to individual
@@
 Once query was placed to DAS an associated
-data-service API was invoked, triggering in DAS cache population. Once data become
-available in DAS cache user were able to see the results. This so-called on demand
+data-service API was invoked, triggering DAS cache population. Once data become
+available in DAS cache user were able to see the results. This on demand
 query retrieval system, where data were retrieved and store into cache from
@@
 data-service APIs, look-up data, re-format their output to DAS notations and
-store results into DAS cache.\footebnote{Even though we supported a data retrieval on demand
+store results into DAS cache.\footnote{Even though we supported a data retrieval on demand
 basis we also store user queries into DAS analytics DB to identify
@@
 
+\bibitem{DBS-QL} V. Kuznetsov, D. Riley, A. Afaq, V. Sekhri, Y. Guo, L. Lueking,
+``The CMS DBS Query Language'', CHEP 2009
+
 \bibitem{CMSDataModel} A. Fanfani et. al.,
 ``Distributed Analysis in CMS'', to be published in Journal of Grid Computing.
-
-\bibitem{DBS-QL} V. Kuznetsov, D. Riley, A. Afaq, V. Sekhri, Y. Guo, L. Lueking,
-``The CMS DBS Query Language'', CHEP 2009
 

### 2010-05-04 src/python/DAS/core/das_mapping_db.py
Adjust init message
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@479 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 DAS mapping DB
 """
 
-__revision__ = "$Id: das_mapping_db.py,v 1.5 2009/09/29 20:48:11 valya Exp $"
-__version__ = "$Revision: 1.5 $"
+__revision__ = "$Id: das_mapping_db.py,v 1.6 2009/10/02 18:58:41 valya Exp $"
+__version__ = "$Revision: 1.6 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -32,7 +32,7 @@ class DASMapping(object):
         self.dbname  = getarg(config, 'mapping_dbname', 'mapping')
         self.colname = 'db'
 
-        msg = "Init DAS mapping %s:%s@%s" \
+        msg = "DASMapping::__init__ %s:%s@%s" \
         % (self.dbhost, self.dbport, self.dbname)
         self.logger.info(msg)
         

### 2010-05-04 src/python/DAS/core/das_mongocache.py

@@ -5,8 +5,8 @@
 DAS mongocache wrapper.
 """
 
-__revision__ = "$Id: das_mongocache.py,v 1.12 2009/10/02 15:04:05 valya Exp $"
-__version__ = "$Revision: 1.12 $"
+__revision__ = "$Id: das_mongocache.py,v 1.13 2009/10/02 18:59:10 valya Exp $"
+__version__ = "$Revision: 1.13 $"
 __author__ = "Valentin Kuznetsov"
 
 import time
@@ -74,7 +74,7 @@ class DASMongocache(Cache):
         self.logger  = config['logger']
         self.verbose = config['verbose']
 
-        msg = "Init mongocache %s:%s@%s" \
+        msg = "DASMongocache::__init__ %s:%s@%s" \
         % (self.dbhost, self.dbport, self.dbname)
         self.logger.info(msg)
 

### 2010-05-04 src/python/DAS/core/das_core.py
Rearrage code initialization, remove obsolete comments
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@480 4525493e-7705-40b1-a816-d608a930855b

@@ -12,8 +12,8 @@ combine them together for presentation layer (CLI or WEB).
 
 from __future__ import with_statement
 
-__revision__ = "$Id: das_core.py,v 1.36 2009/10/02 15:07:22 valya Exp $"
-__version__ = "$Revision: 1.36 $"
+__revision__ = "$Id: das_core.py,v 1.37 2009/10/02 18:59:51 valya Exp $"
+__version__ = "$Revision: 1.37 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -52,9 +52,7 @@ class DASTimer(object):
 
 class DASCore(object):
     """
-    DAS core class:
-    service_keys = {'service':{list of keys]}
-    service_maps = {('service1', 'service2'):'key'}
+    DAS core class.
     """
     def __init__(self, config=None, debug=None):
         if  config:
@@ -78,12 +76,17 @@ class DASCore(object):
         self.logger = DASLogger(idir=logdir, verbose=self.verbose, stdout=debug)
         dasconfig['logger'] = self.logger
 
+        # define Mapping/Analytics/Parser in this order since Parser depends
+        # on first two
         dasmapping = DASMapping(dasconfig)
         dasconfig['dasmapping'] = dasmapping
         self.mapping = dasmapping
 
         self.analytics = DASAnalytics(dasconfig)
         dasconfig['dasanalytics'] = self.analytics
+
+        self.mongoparser = MongoParser(dasconfig)
+        dasconfig['mongoparser'] = self.mongoparser
 
         self.viewmgr = DASViewManager(dasconfig)
 
@@ -99,7 +102,7 @@ class DASCore(object):
             klassobj = '%s(dasconfig)' % klass
             setattr(self, 'rawcache', eval(klassobj))
             dasconfig['rawcache'] = self.rawcache
-            self.logger.info('DASCore::__init__, rawcache=%s' % klass)
+            self.logger.info('DASCore::__init__ rawcache=%s' % klass)
         else:
             msg = 'DAS configuration file does not provide rawcache'
             raise Exception(msg)
@@ -114,7 +117,7 @@ class DASCore(object):
             klassobj = '%s(dasconfig)' % klass
             setattr(self, 'hotcache', eval(klassobj))
             self.cache   = self.hotcache
-            self.logger.info('DASCore::__init__, hotcache=%s' % klass)
+            self.logger.info('DASCore::__init__ hotcache=%s' % klass)
 
         # plug-in architecture: loop over registered data-services in
[…]

### 2010-05-04 src/python/DAS/web/DASSearch.py
Added convert2ui method which will be used to make DAS record to UI conversion based on maps stored in Mapping DB for UI presentation. Tested with simple case of site query. Now record is shown as CMSName: T1_CH_CERN, Site Admin Email: Victor.Zhiltsov@cern.ch, Site Admin Role: Data Manager, DAS record: show, hide which provides invormation about site DAS record and access to full DAS JSON record. Looks like exactly what we need to represent DAS records on a web.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@555 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 DAS web interface, based on WMCore/WebTools
 """
 
-__revision__ = "$Id: DASSearch.py,v 1.24 2009/10/14 15:19:05 valya Exp $"
-__version__ = "$Revision: 1.24 $"
+__revision__ = "$Id: DASSearch.py,v 1.25 2009/10/23 19:43:03 valya Exp $"
+__version__ = "$Revision: 1.25 $"
 __author__ = "Valentin Kuznetsov"
 
 # system modules
@@ -14,6 +14,7 @@ import os
 import time
 import types
 import urllib
+from itertools import groupby
 from cherrypy import expose
 from cherrypy.lib.static import serve_file
 
@@ -34,7 +35,7 @@ from WMCore.WebTools.Page import exposejson, exposexml, exposedasplist
 
 # DAS modules
 from DAS.core.das_core import DASCore
-from DAS.utils.utils import getarg
+from DAS.utils.utils import getarg, access
 from DAS.web.utils import urllib2_request, json2html
 
 import sys
@@ -289,6 +290,19 @@ class DASSearch(TemplatedPage):
         rows = self.result(kwargs)
         return rows
 
+    def convert2ui(self, idict):
+        """
+        Convert input row (dict) into UI presentation
+        """
+        for key in idict.keys():
+            if  key == 'das' or key == '_id':
+                continue
+            for item in self.dasmapping.presentation(key):
+                daskey = item['das']
+                uikey  = item['ui']
+                for value in access(idict, daskey):
+                    yield uikey, value
+
     @expose
     def listview(self, kwargs):
         """
@@ -313,10 +327,11 @@ class DASSearch(TemplatedPage):
         page    = self.templatepage('das_nrecords', **ndict)
         for nrecord in range(0, len(rows)):
             row = rows[nrecord]
-            if  nrecord % 2:
-                style = "white"
-            else:
-                style = "blue" 
+            style = "white"
+#            if  nrecord % 2:
+#                style = "white"
+#            else:
+#                style = "blue" 
             page += '<div class="%s"><hr class="line" /><b>Record #%s</b><br />' % (style, nrecord)
             das = row['das']
             if  type(das) is types.DictType:
@@ -336,10 +351,15 @@ class DASSearch(TemplatedPage):
                     except:
                         data = row
                         key  = 'N/A'
+                    gen = self.convert2ui(data)
+                    for uikey, value in [k for k, g in groupby(gen)]:
[…]

### 2010-05-04 doc/paper.tex
New round of changes for paper
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@556 4525493e-7705-40b1-a816-d608a930855b

@@
 Meta-data plays signiﬁcant role in a large modern enterprises, 
-research experiments, digital libraries where it comes from different 
+research experiments and digital libraries where it comes from different 
 sources and distributed in a variety of forms and digital formats. 
 It is organized and managed by constantly evolving software using 
-both relational and non-relational data sources. There is a big 
-demand to enable information discovery from multiple sources.
+both relational and non-relational data sources. 
+Even though we can access it via viriety of data-services, twikies, blogs, etc.,
+there is no coherent way for information discovery from different sources
+in heterogeneous, distributed environment with different security policies enforced.
 
@@
 indexes and delivers information from different relational and 
-non-relational data sources to answer cross data source queries 
-examining metadata associated to petabytes of 
-experimental data. We examine simplicity of keyword-based system 
-with precise answers of SQL queries under new system where 
-aggregated information being collected from various sources 
-allowing end-users to place dynamic queries and trigger information 
-retrieval on demand. Such close to real-time system based on studies 
-of use cases of the CMS particle physics experiment at the LHC which 
-uses a large scale, distributed computing system to motivate the work.
-
-%Here we discuss a new data aggregation system which consumes, 
-%indexes and delivers information from different relational and 
-%non-relational data sources to answer cross data source queries 
-%examining metadata associated to petabytes of 
-%experimental data. We follow the use cases of the CMS particle 
-%physics experiment at the LHC which uses a large scale, distributed 
-%computing system system to motivate the work.
+non-relational data sources to answer cross data-source queries 
+to explore metadata associated to petabytes of experimental data. 
+We combine simplpicity of keyword-based search with precision of RDMS
+under a new system where aggregated information being collected from various sources 
+allowing end-users to place dynamic queries, get precise answers and 
+trigger information retrieval on demand. Such close to real-time system 
+based on studies  of use cases of the CMS particle physics experiment at 
+the LHC which uses a large scale, distributed computing system to motivate the work.
 
@@
 The CERN, the European Organization for Nuclear Research, plays a leading
-role in fundamental studies of physics. Apart from scientific community 
-it is known as a place where World Wide Web was born. At that time the 
-information look-up among hyperlinks represents a certain challenge for scientists.
+role in fundamental studies of physics. It is also known as a place where
+World Wide Web was born to help researches share information among each other
+via concept of hypertext.
 Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
-Physics (HEP), promising to deliver a few PB of data each year. Scientists are
-facing another set of problems from distributed computing to multi-dimensional
-view on information retrieval. One of the aspect of such research is an efficient
-and at the same time precise look-up of produced meta-data, which comes in variety 
-of forms and digital formats. As was pointed out in \cite{Arms} a mixed content and 
-mixed meta-data and meta-data consistency should be considered as a whole in design 
-of the system to successful information discovery.
+Physics (HEP), promising to deliver a few PB of data each year. 
+At this scale scientists are facing a new set of problems in an area of
+``information discovery'' within heterogeneous, distributed environment.
+The data and associated meta-data are produced in variety of forms and digital formats.
+They are stored into relational and non-releational data-sources, such as 
+RDMS systems, document oriented databases, blogs, twikies, customized applications, etc. 
+and its access is restricted by broad variety of security policies. 
+Working in such environment is not easy and require a lot of knowledge, but users
+(in this case physicists) are always looking for simple, intuitive and flexible
+tool to look-up their desired data. A well-known solutions, such as web-services,
+blogs, twikies and search engines targeted and tighted to specific data
+sources and end-users are left with a way to bookkeep and relatate information
+from them.
 
-The CMS, the Compact Muon Solenoid experiment, operated at LHC,
-represents an heterogeneous environment of distributing computing, relational and
-non-relational data sources where we are facing this problem. The computing resources
-are distributed among almost 40 countries, 183 institutions and more than 3000 physicists.
-Broad variety of RDMS systems at different data centers collect a various
-meta-data information, including data location and transfer, detector conditions,
-calibrations constants, simulation information, etc. Moreover the development
-of those components were done in parallel by different group and technology
-tools. Therefore a wide spectrum of data services and
-data formats represents a challenge of efficient information discovery.
+Here we present a work on Data Aggregation System (DAS) designed for
+CMS High-Energy Experiment operated at LHC which provide
+ability to query, search and aggregate information from different 
+data-services as well as serves as a caching layer for them, 
+preserving their security policies.
 
-Here we present a work on Data Aggregation System (DAS), the system designed
-to provide ability to query, search and aggregate information from different 
-data-services as well as serves as a caching layer for them, 
-preserving their security policies. We organize
+We organize
 our discussion as following. In section \ref{RelatedWork} we discuss
-related work in domain of keyword search over relation data sources.
-The section \ref{DataModel} provides review of CMS data model. In section
+related work in a domain of keyword search over relation data sources.
+The section \ref{DataModel} describe CMS experiment and its data model. In section
 \ref{DAS} we outline architecture of the DAS system, including discussion of its
 various components. Finally our results are summarized in section \ref{Results}.
+
+%As was pointed out in \cite{Arms} a mixed content and 
+%mixed meta-data and meta-data consistency should be considered as a whole in design 
+%of the system to successful information discovery.
 
@@
 by indexing DB content. In former case, you still face a problem with understanding 
-of the underlying schema and imposing relational conditions in your query, while 
+of underlying schema and imposing relational conditions in your query, while 
 in later queries can be expressed via simple keyword based search, even though exact match
@@
 We wanted to expand this approach to handle broad variety of relational and
-non-relational data-source in CMS within our distributed non-homogeneous environment.
-
-%Here we discuss a new system, the Data Aggregation System (DAS),
-%developed in CMS collaboration to address this issue. We start with
-%discussion of CMS data model and existing data-services, see \ref{DataModel}. 
-%In section \ref{DAS}
-%we provides a detailed description of underlying design and components
-%used in DAS system. Our results are summarized in section \ref{Results}.
+non-relational data-sources in CMS within our distributed heterogeneous environment.
+At the same time we wanted to simplify it up to the level that users will be
+able to place free text-based queries (keyword based search) to look-up their desired data.
 
 \section{CMS data model\label{DataModel}}
+The CMS, the Compact Muon Solenoid experiment \cite{CMS} 
+is one of the two large general-purpose particle physics detectors built on 
+the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
+It is designed to explore the frontier of High Energy Physics and provide physicsists
+ability to look at conditions which were present at early stage of our Universe.
+More then 3000 physicists from 183 institutions prepresenting almost 
+38 countries are involved in design, construction and maintenance of the experiment.
+
 The CMS distributed computing and data model \cite{CMSDataModel} 
@@
 tasks. Based on their budget resources the choice of back-end DB were
-up to site managers. Therefore we designed our software to support different
+up to site managers. Therefore the CMS software were designed to support different
 DB back-ends and provide tools for migrating data across them.
@@
 to answer specific questions about data they serve, the ability to relate and search
-information among them was human task. The growing amount of information
+information among them was a human task. The growing amount of information
 and desire to make cross-service queries force us to design and develop a new
@@
 The design of DAS system was based on previous studies of CMS 
-data bookkeeping system (DBS) usage patterns \cite{DBS, DBS07}. We carefully analyzed user
-queries placed to DBS, their patterns, frequencies and latency. As a result of this
-we created a data discovery service \cite{DD} which came through several iteration
-of UI. At the end the end-user interface was based on DBS Query Language (DBS-QL)\cite{DBS-QL}
-with broad variety of template for presentation layer. A quick adoption
-and wide spread of DBS-QL give us confidence of choosen approach and QL syntax.
+data bookkeeping system (DBS) use cases \cite{DBS, DBS07}. We carefully analyzed user
+queries placed to DBS, their patterns, frequencies and latency. As a result
+we created a data discovery service \cite{DD} which came through several iterations
+of design choices. At the end the end-user interface was based on DBS 
+Query Language (DBS-QL)\cite{DBS-QL} which support dynamic queries against
+DBS back-end and had syntax very closed to free-text based search queries. 
+The DBS-QL \cite{DBS-QL} syntax was based on SQL query language with one important
+difference that we took away from end-users any knowledge about relations of underlying schema.
+It was achieved by usage of short-path algorithm to establish necessary
+table joins based on provided information in user queries and auto-loaded
+schema from DB back-end and knowledge of foreign key relationships.
+A quick adoption and wide spread of DBS-QL give us confidence of choosen approach 
+and provide a basis for building DAS system.
 
-Starting from those ground accomplishments we created DAS system as
-scalable and pluggable layer sitting on top of the existing data-services
-within CMS computing infrastructure. Data availability were deligated to individual
-data-services. Therefore, we treat DAS as aggregation and cache service 
-with the following characteristics:
+As was pointed out in \cite{Arms} a mixed content and 
+mixed meta-data and meta-data consistency should be considered as a whole in design 
+of the system to successful information discovery. 
+Starting from this ground we designed DAS system as
+additional layer on top of the existing data-services
+within CMS computing infrastructure by requiring the following:
 \begin{itemize}
@@
 due to continuously changing detector and software environments;
-\item be transparent to data-service data formats.
+\item retrieve data on demand and aggregate them if necessary across
+multiple data-services;
+\item be transparent to data-service data formats;
+\item support legacy application, data-services and APIs.
 \end{itemize}
@@
 \noindent
-Figure \ref{DAS_cache} summarize a overview of DAS architecture who consist of the
+The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of the
 following components:
 \begin{itemize}
-\item DAS core library with support of pluggable modules for data retrieval;
-\item DAS caching layer with long-term cache for aggregated results and short-term
-cache for presentation layer;
-\item DAS request service to handle user requests;
-\item DAS mapping DB with DAS/data-services mapping, notations mapping and
-mapping of DAS objects into data-presentation layer;
-\item DAS analytics DB with query statistics to enable pre-fetching 
+\item core library with support of pluggable modules for data retrieval;
+\item caching layer to store data-service output and aggregated results;
+\item request service to handle user requests;
+\item mapping DB to keep information about data-service APIs, their
+notations and mapping to DAS;
+\item analytics DB with query statistics to enable pre-fetching 
 strategies.
 \end{itemize}
-Further we proceed with a discussion of individual DAS components.
 
-\subsection{Dealing with data-services}
-DAS provided pluggable system to add a new CMS data-service. To do that we
-identify appropriate APIs, input and output parameters and keys, data-service
-format and mapping between data-service, DAS and other sub-systems. 
-This information were stored into DAS mapping DB. For instance,
-the Data Bookkeeping System (DBS) and data location system (Phedex) both
-provide information about CMS data files. In former case, DBS keeps information between
-files and other physics  objects, such as run and luminosity, while Phedex provides 
-information about file location, such as site and details about storage elements.
-Upon user request to find information about a file, based on forementioned
-information we were able to identify appropriate set of
-data-service APIs, look-up data, re-format their output to DAS notations and
-store results into DAS cache.\footnote{Even though we supported a data retrieval on demand
-basis we also store user queries into DAS analytics DB to identify
-a common patterns. Based on this information we organized a pre-fetching
-strategies.}
-Once data were retreived from data-services an additional step to aggregate
-matched objects in DAS cache were applied. For instance in above examples
-file information came from DBS and Phedex systems. 
-Since both data-services output contained file name as a common
-key, their output where merged and final result was stored into DAS cache.
-This merged (aggregated) results represented a DAS data object. Each DAS data
-object contains a standard DAS header, which identify look-up time,
-request url, API, parameters and method as well as response version, expiration
-time and checksum of the output.
+The DAS core communicates with different data-services and retrieve
+data on demand. This was done by invoking appropriate data-service APIs
+upon user requests. All user queries were written to Analytics DB
+for further analysis. The data-service output where re-mapped into
+DAS notations and if necessary aggregated inside of DAS cache.
+Since we didn't know a-priory what kind of data users will ask, 
+what kind of queries they will place against the system and what
+level of aggregation will be applied the choice
+of RDMS as DAS back-end was problematic\footnote{In addition DAS 
+didn't impose any requirements for transactions and persistency of the data.} 
+and we look around for other options. Through the analysis of available IT solutions, 
+such as file-based and memory caches, 
+key-value databases, documented-oriented databases we made our choice in favor 
+of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
+MongoDB \cite{MongoDB}. Both of them provide schema free document
+storage, replication and fail-over features, but our choice in favor of 
+MongoDB was obvious because its support of dynamic queries, 
+full indexies, including on inner objects and embedded arrays,
+and auto-sharding. Our prelimiary benchmarks shown that it can sustain
+the desired load and size for our meta-data information. We use MongoDB for
+the three main components: DAS cache, Mapping and Analytics databases. All of them
+will be discussed in forthcoming section.
 
-We designed DAS interface to be fully REST complaint \cite{REST}. It is not
-only simplify development of DAS clients (CLI, web interface, robot, etc.)
-but also consider DAS itself as own sub-component, i.e. place queries against
-DAS content, rather then data-service one.
-
-\subsection{Query Language\label{DAS-QL}}
-Upon success of DBS-QL \cite{DBS-QL} we adopted its syntax to DAS with a few minor
-changes. Here we briefly recall its syntax:
-\begin{equation}
-action\,\,\,
-key_1,\, key_2,\, ...\,\,\, where\,\,\,
-\langle key\rangle\, 
-\langle op\rangle\, 
-\langle value\rangle \,\,\, and|or ...
-\label{QL_syntax}
-\end{equation}
-Each DAS QL expression starts with special keyword indicating an action such as 
-$find,\, plot,\, summary$, followed by a set of selection keys. The action identify
-the behavior of the query, find to look-up data, plot to plot found data, summary to
-apply a summary view over provided selection keys. An optional conditions can be 
-specified via $where$ keyword following by condition expression.
-Each condition expression represented via $key\,\,\, operator\,\,\, value$ triplet, where
-$key$ is DAS selection key, $operator$ is conditional operator, 
-$=,\, >=,\, <,\, in,\, between$ and $value$ is a value applied to selection key.
-We also support brackets as well as $and$ and $or$ logical operators.
-The DAS selection and condition keys were identified based on analysis 
-of terminology used by our users. For instance, $file$ was used to either
-select file name or used in condition to identify file name provided by user.
-Here is a simple example of DAS QL expression:
-$$
-find\,\,\, file,\, run\,\,\, where\,\,\, site=T1\_CH\_CERN
-$$
-which represents a natural language question: 
-{\it find files and runs located at T1\_CH\_CERN site}.
-
-\subsection{Mapping}
-As been discussed above we used a DAS mapping DB to map data-service API into
-DAS as well as map provide a relations between different data-services.
-Upon identification of APIs which participate in DAS activity 
-we store a mapping between data-service notations
-into DAS mapping DB. This allows to map user input query parameters into
+DAS Mapping DB holds mapping between data-service APIs and DAS keys. 
+Upon identification of APIs which will participate in DAS activity 
+we store to Mapping DB API name, input parameters, their type and accept 
+values as regular expression patterns as well as mapping between data-service 
+notations into DAS keys. This allows to map user input query parameters into
 data-service API metrics as well as map back data-service output into DAS names.
-For instance, the DBS system used $logical\_file\_name$ notation, while in Phedex
+For instance, the DBS system used $logical\_file\_name$ notation, while in PhEDEx
+\cite{PhEDEx}
 the same entity was named as $lfn$. Since both of them were referred to
-logical file name, we map both notations into $file$ notation in DAS.\footnote{We
+logical file name, we map both notations into DAS key $file$.\footnote{We
 also deal with physical file names, but still used $file$ to refer to them in
@@
 For each service we provide a mapping from its own data-format (JSON, XML, etc.) into
-common data-format used in DAS.\footnote{We used JSON as a DAS data format}
-We also store a mapping between DAS keywords and data-service API names to
+common data-format used in DAS.\footnote{We used JSON as a DAS data format.}
+We also store a mapping between DAS keys and data-service API names to
 establish appropriate calls to data-service upon provided DAS query. For example,
-when user specifies $file$ in their DAS query, we invoke DBS listFiles and Phedex
-fileReplicas APIs. Since all mapping kept in RDMS we were able to abstract the 
-DAS core library calls from hard coded data-service API calls and make it
-transparent to any participating data-services. Here is a logic of DAS core
-library:
-\begin{itemize}
-\item parse input DAS query
-\item map provided selection and condition keys into data-service APIs and
-input parameters
-\item invoke data-service APIs
-\item parse output results by mapping output notation into DAS keys
-\item store and aggregate them into DAS raw-cache
-\end{itemize}
+when user specifies $file$ in their DAS query, we invoke DBS listFiles and PhEDEx
+fileReplicas APIs.
 
-\subsection{Caching system and Analytics}
-The DAS caching system was divided into two different caches: raw-cache and hot-cache.
-The purpose of raw-cache was to store results coming from data services for
-as long as specified in data-service response header. While the purpose of hot-cache
-was to keep final DAS results corresponding to provided user query. We use a robot
-services (jobs) to feed raw-cache with highly requested queries. Here we will only
-discuss the raw-cache organization. 
-%\begin{figure}[htb]
-%\centering
-%\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
-%\caption{
-%DAS cache system.
-%}
-%\label{DAS_cache}
-%\end{figure}
-As it shown in figure \ref{DAS_cache} the DAS cache server has dual purpose.
-On one side it talks to DAS Analytics DB to trigger jobs for worker robot
-who update cache with most popular queries, on another it talks to DAS
-raw-cache back-end to either stored mapped results or merge them with 
-existing DAS objects in a cache. We should note that we use on-demand
-approach to answer user queries, but at the same time with a help of
-DAS Analytics DB we force worker robots (robots) to pre-fetch
-the most popular queries up-front.
+The DAS Analytics DB plays special role for DAS. It collects information
+about user requests placed to the system. Each request was recorded. Upon its
+decomposition into set of selection keys and conditions we also record
+API name and API input parameters and keep updating their counter for repeated
+calls. This allows to keep track of frequency of API calls and establish pre-fetch
+strategy for most requested APIs.
 
-We evaluated several technologies
-as a choice for DAS cache back-end, including relational MySQL DB \cite{MySQL} 
-and object oriented CouchDB \cite{CouchDB} and MongoDB \cite{MongoDB}.
-Each of them had its own pros and cons. For instance usage
-of MySQL provided a SQL queries (as in federated DBs, see \cite{FedDB}), but
-required auto-schema generation on a fly.\footnote{We did not apply any
-restriction to DAS QL and did not know in advance what our users will be
-querying.} While, using CouchDB and MongoDB, we were able to store
-data-service output naturally as documents objects into DB. Finally we decided
-to use MongoDB due to its flexible and
-powerful querying syntax, which were easy to map into DAS QL expressions.
+The DAS caching system was used to store data-service results and aggregated
+information in form of DAS records. As it shown in figure \ref{DAS_cache} 
+the DAS cache server has dual purpose.
+On one side it talks to DAS Analytics DB to trigger jobs for worker robots
+who update cache with most popular queries, on another side it talks to DAS
+cache back-end to either stored mapped results or merge them with 
+existing DAS objects in a cache.
+
+Each user request placed to the system was parsed and set of selection
+and condition keys were identified. They were mapped into set of
+data service APIs and their input parameters. DAS
+cache was checked for a similar set of selection and condition
+keys which can be placed earlier.\footnote{Since DAS doesn't hold
+any data by default we cannot rely on data being present in a cache.
+Moreover if we look-up for data first, we cannot tell if it was complete
+set since earlier queries may retrieve only portion of the requested data.}
+If the superset was found a data were
+retrieved from the cache, otherwise a data-service APIs were
+invoked and the results were collected and aggregated in DAS cache.
+
+%Now we present overall logic of DAS workflow and details of individual components.
+%Upon request to DAS the following steps were executed: 
+%\begin{itemize}
+%\item parse input DAS query
+%\item map provided selection and condition keys into data-service APIs and
+%input parameters
+%\item look-up similar queries in DAS cache 
+%\item invoke data-service APIs
+%\item parse output results by mapping output notation into DAS keys
+%\item store and aggregate them into DAS raw-cache
+%\end{itemize}
+
+Apart from previous DBS-QL syntax we decided to fully support free text-based queries for end-users.
+All of them were transformed into MongoDB query syntax, who was used internally in
+DAS core layer. It represents a set of selection keys and conditions as a JSON document.
+For example, a query
+$dataset\,\,site=T1\_CH\_CERN$
+where transformed into 
+$\{"fields" : ["dataset"], "spec" : \{ "site" :"T1\_CH\_CERN"\}\}$.
+Such syntax was perfect fit for us, since it allows us to store all queries
+as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
+MongoDB query syntax is very reach. It supports advanced queries with 
+boolean expressions, conditional operators, regular expression, etc. providing
+a real power for end-user queries.
+
+\subsection{Data-services and aggregation}
+DAS provided a pluggable system to add new CMS data-services. To do that we
+identify appropriate APIs, input and output parameters, data-service
+format and their mapping to DAS keys as well as relation to other sub-systems. 
+This information was stored into DAS Mapping DB. A data-service plugin consists
+of its DAS map and individual data parser. The DAS Mapping DB becomes an authorative
+source of information about all data-services participating in DAS and their
+relationships. Based on this information we were able to retrieve appropriate data
+from different data-services, re-map them into DAS notations and
+aggregate them on demand.
+For instance, the Data Bookkeeping System (DBS) and data location system (PhEDEx)
+provided information about CMS data files. In former case, DBS holds information about
+files and their relationship to other physics objects, such as run and luminosity, 
+while PhEDEx stores information about the file location, such as sites which
+holds copy of the files and details about storage elements.
+Upon a user request to find a file, we were able to identify appropriate set of
+DBS and PhEDEx APIs, look-up data, re-format their output to DAS notations and
+store results into DAS cache.
+Once data were retreived from data-services an additional step to aggregate
+matched objects in DAS cache were applied. For instance, both data-service outputs 
+contained a file name in them. We use it as a common key to merge records in DAS cache.
+This merged (aggregated) results represented a DAS data objects. Each DAS data
+object contains a standard DAS header, which identify look-up time,
+request url, input parameters and API method as well as response version, expiration
+time and checksum of the output.
+
+We used REST model \cite{REST} for all data requests to DAS.
+It allows us to share core library and other components during development of 
+various DAS clients, e.g. CLI, web interface, robots, etc. and hides
+complexity of access to individual data-services and security policies.
+For instance, to access on-line Run Summary information we were forced to use
+Grid certificates. Instead of propagating user certificate to data-service
+the DAS works as a proxy and use host certficiate to serve such requests.
 
 \section{Results\label{Results}}
-DAS was deployed with the following CMS data-services:
+DAS was deployed plugins to work with the following CMS data-services:
 \begin{itemize}
@@
 about CMS meta-data;
-\item PhEDEx \cite{Phedex}, Physics Experiment Data Export project which
+\item PhEDEx \cite{PhEDEx}, Physics Experiment Data Export project which
 provides the data placement and the file transfer system for the CMS experiment;
+\item SiteDB \cite{SiteDB} is a catalogue of all CMS computing Tiers. 
+It records the CMS resources at the site, the resources pledged for the 
+future, and it keeps track of CMS personnel at each site, including the 
+roles and duties they fulfill in the collaboration;
 \item RunSummary DB \cite{RunSummary} collects information about run and triggers
@@
 Each data-service has its own scope, size and lifetime. For instance the data
-in Phedex are transient due to constant migration of CMS data. The DBS system
-were divided into dozen of individual DBs, whose size varied at the level of
-few tens of GBs.
+in PhEDEx are transient due to constant migration of CMS data with daily DB size of
+around few tens of GBs. The DBS system were divided into dozen of individual DBs, 
+who has been used by different production teams for various bookkeeping tasks. Typical
+size of individual DBS instance was around of fwe GBs, with overall DBS size of
+around of 50 GBs.\footnote{This include table and index sizes.}
+
+At the moment we don't have exact numbers for amount of data we will store to DAS,
+but based on test performed during CRAFT09 exercise we produced 0.3PB of data which
+translate into $\sim50$ GB of meta-data stored into DBS system. Based on this
+we can anticipate to collect $\sim500$GB of meta-data per year during data-taking period
+at LHC.
+%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
+%billion events, 330TB in 32 datasets:
+%AlCa/Calibration RAW: 926 million events, 29TB
+%Physics RAW: 1.24 billion events, 147TB
+%PromptReco: 1.24 billion events, 111TB
+%Bulk ALCARECO: 2.34 billion events, 11TB
+%Express FEVT: 105 million events, 24TB
+%Express ALCARECO: 153 million events, 0.99TB
+%HLTMON FEVTHLTALL: 31 million events, 7.4TB
+%Above event numbers have signicant overlap by design, 
+%2.2 billion unique events, including 524 million cosmic muon
+%triggers
+% 0.3PB - 50GB meta-data in DBS
+% 3PB/y - x GB
+
+We did preliminary studies of DAS performance based on information stored
+in 2 systems, DBS and PhEDEx. We benchmark time required to collect and 
+aggregate block objects information from both system, see Table \ref{DAS_benchmark}.
+
+\begin{table*}[hbt]
+\centering
+\begin{tabular}{llllll}\hline
+\hline
+
+System & Format & Size & Records & Elapsed time & Elapsed time \\
+& & & & no cached data & w/ cached data \\
+\hline
+DAS (DBS) & XML & 187MB & 346820 & 165 sec & 2.53 sec \\
+DAS (PhEDEx) & JSON & 164MB & 230035 & 177 sec & 2.42 sec \\
+DAS (aggregate) & JSON & 357MB & 348866 & 386 sec & 41.7 sec \\
+\hline
+\hline
+\end{tabular}
+\caption{DAS benchmark results. We performed a stress test with 2 CMS system, which
+both deliver block object information. The total number of DAS records is calculated
+as number of aggregated records and left over from both systems.}
+\label{DAS_benchmark}
+\end{table*}
+
+We measured elapsed time as the following:
+\begin{itemize}
+\item[]
+Elapsed time = retrieval time + parsing time + remapping time 
+        + cache insertion/indexing time 
+        + (aggregation time) + (output creation time)
+\end{itemize}
+Here the retrieval time is a time required to access data from remote data-services,
+parsing time is a time required to read and parse received data, remapping time
+is a time to convert notations used by data-service to DAS ones for every object
+we parse, cache insertion and indexing time represents time spend to add objects into
+the DAS cache, aggregation time is required to merge objects into DAS records based
+on their common key (block name in this case) and output creation time 
+is a time required to write DAS records to disk. Please note that last two
+components, aggregation and output creation time, were only applied to
+final DAS step and not to individual components.
+
+The individual DAS components (DBS and PhEDEx) shown roughly the same performance.
+For example, in case of PhEDEx the total time spent in DAS was 177 seconds, where
+majority of the time was spent on retrieval, 83 seconds, and parsing, 29 seconds.
+The remaining 65 seconds was spent on remapping, cache insertion, indexing. The
+numbers for DBS system were a like. As you can see from Table \ref{DAS_benchmark},
+the time spent to read DAS records (last column) from cache was quite
+reasonable, roughly 2.5 seconds from both systems. While final time to
+get DAS records on disk was below 42 seconds. This was mostly due to I/O operations.
+
+Based on this results we measured that we can achieve 900 docs/sec for creation of DAS
+in a cache and 8500 docs/sec for reading DAS records back, which is suitable for our
+needs.\footnote{We've been satisfied with such performance,
+even though our measurements shown that it can still be speed up by factor of
+10 if we will use C++ DB driver.}
 
 \section{Summary}
+We presented new data aggregation service (DAS) developed for High-Energy experiment
+at LHC, CERN, Geneva, Switzerland. It was designed to providing caching and
+aggregation layer on top of existing relational and non-relation data-service
+in mostly real time fashion. All the data were retrieved on demand basis,
+while a pre-fetch of most common queries is in our plans. The DAS system reach
+first is under commissioning phase right now and prelimiary performance studies
+shown it can accommodate needs of CMS experiment. The CMS experiment started to
+collect data from November 2009 and we expect to write about PB of data each
+year to tape. In addition the Monte-Carlo samples will be produced at this scale.
+Based on test studies performed in CMS, it can translate to the order of
+$\sim500$GB of meta-data each year. The DAS should sustain such load and provide
+interface as data discovery service in CMS.
 
@@
 \begin{thebibliography}{9}
-\bibitem{Arms}
-C. R. Arms, W. Y. Arms, ``Mixed Content and Mixed Metadata 
-Information Discovery in a Messy World'',
-chapter from ``Metadata in Practice'', ALA Editions, 2004
-
 \bibitem{FedDB}
@@
 
-\bibitem{CMSDataModel} A. Fanfani et. al.,
+\bibitem{CMS} CMS Collaboration R. Adolphi et al., The CMS experiment at the CERN LHC, JINST, 0803, S08004 (2008)
+
+\bibitem{CMSDataModel} 
+.Grandi, D.Stickland, L.Taylor et al., The CMS Computing Model, CERN-LHCC-2004- 035/G-083 (2004);
+A. Fanfani et. al.,
 ``Distributed Analysis in CMS'', to be published in Journal of Grid Computing.
@@
 \bibitem{DBS} A. Afaq, et. al. ``The CMS Dataset Bookkeeping Service'', CHEP 2007 
+J.Phys.Conf.Ser, 119, 072001 (2008)
 
@@
 \bibitem{DD} https://cmsweb.cern.ch/dbs\_discovery
+
+\bibitem{Arms}
+C. R. Arms, W. Y. Arms, ``Mixed Content and Mixed Metadata 
+Information Discovery in a Messy World'',
+chapter from ``Metadata in Practice'', ALA Editions, 2004
 
@@
 
-\bibitem{Phedex}
+\bibitem{PhEDEx}
 Rehn et. al.,
 ``PhEDEx high-throughput data transfer management system'', CHEP06, Mumbai, India.
+R. Egeland et al., Data transfer infrastructure for CMS data taking, Proceedings of Science,
+PoS(ACAT08)033 (2008) ;
+L. Tuura et al., Scaling CMS data transfer system for LHC start-up, J.Phys.Conf.Ser, 119, 072030 (2008)
 
 \bibitem{RunSummary}
+https://cmswbm.web.cern.ch/cmswbm/cmsdb/servlet/RunSummary
 
 \bibitem{SiteDB}
+https://cmsweb.cern.ch/sitedb/
 
@@
 \bibitem{DQ}
+Data-Quality DB reference
 

### 2010-05-04 doc/paper.tex
Added ATLAS AMI web portal reference for completeness
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@557 4525493e-7705-40b1-a816-d608a930855b

@@
 an attempt to build a simple, intuitive and flexible query language was introduced
-in \cite{DBS-QL}. It represented a power of SQL while
+in \cite{DBS-QL}.\footnote{The ATLAS experiment at CERN, has been developed
+AMI web portal \cite{AMI} with similar functionality.}
+It represented a power of SQL while
 hiding underlying relational schema from the end users. As a results
@@
 
+\bibitem{AMI}
+Altas AMI web portal:
+http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html
+
 \bibitem{CMS} CMS Collaboration R. Adolphi et al., The CMS experiment at the CERN LHC, JINST, 0803, S08004 (2008)

### 2010-05-04 doc/paper.tex
More updates
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@558 4525493e-7705-40b1-a816-d608a930855b

@@
 and auto-sharding. Our prelimiary benchmarks shown that it can sustain
-the desired load and size for our meta-data information. We use MongoDB for
-the three main components: DAS cache, Mapping and Analytics databases. All of them
-will be discussed in forthcoming section.
+the desired load and size for our meta-data information. We use MongoDB as a back-end 
+for the three main components: Mapping, Analytics and DAS cache databases.
 
@@
 Upon identification of APIs which will participate in DAS activity 
-we store to Mapping DB API name, input parameters, their type and accept 
+we store to Mapping DB the API name, input parameters, their type and accept 
 values as regular expression patterns as well as mapping between data-service 
@@
 data-service API metrics as well as map back data-service output into DAS names.
-For instance, the DBS system used $logical\_file\_name$ notation, while in PhEDEx
+For instance, the DBS system \cite{DBS}
+used {\it logical\_file\_name} notation, while in PhEDEx
 \cite{PhEDEx}
-the same entity was named as $lfn$. Since both of them were referred to
-logical file name, we map both notations into DAS key $file$.\footnote{We
-also deal with physical file names, but still used $file$ to refer to them in
-DAS notation, since simple regular expressions were used to identify $file$
+the same entity was named as {\it lfn}. Since both of them were referred to
+logical file name, we map both notations into DAS key {\it file}.\footnote{We
+also deal with physical file names, but still used {\it file} to refer to them in
+DAS notation, since simple regular expressions were used to identify {\it file}
 meaning from the provided value.}
@@
 establish appropriate calls to data-service upon provided DAS query. For example,
-when user specifies $file$ in their DAS query, we invoke DBS listFiles and PhEDEx
+when user specifies {\it file} in their DAS query, we invoke DBS listFiles and PhEDEx
 fileReplicas APIs.
 
-The DAS Analytics DB plays special role for DAS. It collects information
+The DAS Analytics DB plays special role. It collects information
 about user requests placed to the system. Each request was recorded. Upon its
@@
 information in form of DAS records. As it shown in figure \ref{DAS_cache} 
-the DAS cache server has dual purpose.
-On one side it talks to DAS Analytics DB to trigger jobs for worker robots
-who update cache with most popular queries, on another side it talks to DAS
-cache back-end to either stored mapped results or merge them with 
-existing DAS objects in a cache.
+the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
+and aggregate data over there. The independent set of daemons, DAS robots, are
+constantly monitor Analytics DB for most popular requests and pre-fetch most
+common data in DAS cache.
 
@@
 For example, a query
-$dataset\,\,site=T1\_CH\_CERN$
+{\it dataset site=T1\_CH\_CERN}
 where transformed into 
-$\{"fields" : ["dataset"], "spec" : \{ "site" :"T1\_CH\_CERN"\}\}$.
+{\it \{"fields" : ["dataset"], "spec" : \{"site" : "T1\_CH\_CERN"\}\}}.
 Such syntax was perfect fit for us, since it allows us to store all queries
@@
 boolean expressions, conditional operators, regular expression, etc. providing
-a real power for end-user queries.
+a real power for our end-users.
 
 \subsection{Data-services and aggregation}
-DAS provided a pluggable system to add new CMS data-services. To do that we
-identify appropriate APIs, input and output parameters, data-service
+DAS provided a pluggable interface to add new CMS data-services. 
+Each data-service plugin consists of its DAS map and individual data parser. 
+The data-service DAS map consists of information about data-service APIs,
+input and output parameters, data-service
 format and their mapping to DAS keys as well as relation to other sub-systems. 
-This information was stored into DAS Mapping DB. A data-service plugin consists
-of its DAS map and individual data parser. The DAS Mapping DB becomes an authorative
+The information was stored into DAS Mapping DB. Below you can see a typical
+record of SiteDB \cite{SiteDB} data-service:
+\begin{verbatim}
+{"api": {"params": {"name": ""}, "name": "CMSNametoSE"}, 
+ "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
+              "api_param": "name", 
+              "das_key": "site"}], 
+ "_id": "4ae1f8dfe2194e4581000014", 
+ "system": "sitedb", 
+ "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
+              "key": "site", 
+              "map": "site.name"}]}
+\end{verbatim}
+It contains information about API name and parameters, mapping between API input
+parameter and DAS key, CMS system name and set of DAS keys this API is covering.
+We also used regular expression patterns, evaluated by python during run time
+execution, to verify provided set of input paremters and DAS keys.
+
+The DAS Mapping DB becomes an authorative
 source of information about all data-services participating in DAS and their
@@
 contained a file name in them. We use it as a common key to merge records in DAS cache.
-This merged (aggregated) results represented a DAS data objects. Each DAS data
+This merged (aggregated) results represented DAS data objects. Each DAS data
 object contains a standard DAS header, which identify look-up time,
 request url, input parameters and API method as well as response version, expiration
-time and checksum of the output.
+time and checksum of the output. Here is an example of individual DAS record:
+\begin{verbatim}
+{"_id": "4ae1e7ffe2194e4243000003", 
+ "site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
+         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
+ "das": [{"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
+          "ctime": 0.95439004898071289, "expire": 1256362175.7692671, 
+          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSAMName", 
+          "timestamp": 1256318975.7692671, "system": ["sitedb"], 
+          "api": ["CMStoSAMName"], "version": "", "selection_keys": []}, 
+         {"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
+          "ctime": 1.2158100605010986, 
+          "api": ["CMStoSiteName"], 
+          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSiteName", 
+          "timestamp": 1256318974.7774379, "system": ["sitedb"], 
+          "expire": 1256362174.7774379, "version": "", "selection_keys": []}
+        ]
+}
+\end{verbatim}
+It represents aggregated DAS record between two SiteDB APIs, CMStoSAMName and
+CMStoSiteName, where API results are merged under {\it site} key. The information
+under {\it das} key shows DAS header, which contains sufficient amount of
+information about API name, location, execution time and lifetime of the data.
+Using this information was valuable contribution in debugging process of
+data-services themselves, e.g. identification of not-matched records, 
+latency studies, etc.
 
@@
 Grid certificates. Instead of propagating user certificate to data-service
-the DAS works as a proxy and use host certficiate to serve such requests.
+the DAS works as a proxy and used host certficiate to serve such requests.
 
 \section{Results\label{Results}}
-DAS was deployed plugins to work with the following CMS data-services:
+DAS was deployed to work with the following CMS data-services:
 \begin{itemize}
@@
 conditions during data-taking;
-\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
-\item DQ \cite{DQ}, Data Quality data-service which collects information
-about detector conditions during data taking;
-\item Overview \cite{Overview}, collects information about CMS
-transfer rate, etc.;
+%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
+%\item DQ \cite{DQ}, Data Quality data-service which collects information
+%about detector conditions during data taking;
+%\item Overview \cite{Overview}, collects information about CMS
+%transfer rate, etc.;
 \item Dashboard \cite{Dashboard} project for LHC experiments aims to 
@@
 in PhEDEx are transient due to constant migration of CMS data with daily DB size of
-around few tens of GBs. The DBS system were divided into dozen of individual DBs, 
-who has been used by different production teams for various bookkeeping tasks. Typical
-size of individual DBS instance was around of fwe GBs, with overall DBS size of
-around of 50 GBs.\footnote{This include table and index sizes.}
+around few tens of GBs. The DBS system were divided into dozen of individual instances, 
+who has been used by different production teams for various bookkeeping tasks.
+Figure \ref{db_size} shows current size of meta-data from aforementioned data-service.
+\begin{figure}[htb]
+\centering
+\includegraphics[width=100mm]{db_size.pdf}
+\caption{
+Database size of CMS data-services participating in DAS in November 2009, before
+data taking.
+}
+\label{db_size}
+\end{figure}
+%Typical
+%size of individual DBS instance was around of fwe GBs, with overall DBS size of
+%around of 50 GBs.\footnote{This include table and index sizes.}
 
-At the moment we don't have exact numbers for amount of data we will store to DAS,
-but based on test performed during CRAFT09 exercise we produced 0.3PB of data which
-translate into $\sim50$ GB of meta-data stored into DBS system. Based on this
+At the moment we don't have exact numbers for amount of data which will stored to DAS,
+but based on test performed during CRAFT09 \cite{CRAFT09}
+exercise we produced 0.3PB of data which
+translate into $\sim50$ GB of meta-data stored into DBS system. Therefore
 we can anticipate to collect $\sim500$GB of meta-data per year during data-taking period
@@
 \item[]
+{\it
 Elapsed time = retrieval time + parsing time + remapping time 
@@
         + (aggregation time) + (output creation time)
+}
 \end{itemize}
-Here the retrieval time is a time required to access data from remote data-services,
-parsing time is a time required to read and parse received data, remapping time
+Here the {\it retrieval time} is a time required to access data from remote data-services,
+{\it parsing time} is a time required to read and parse received data, {\it remapping time}
 is a time to convert notations used by data-service to DAS ones for every object
-we parse, cache insertion and indexing time represents time spend to add objects into
-the DAS cache, aggregation time is required to merge objects into DAS records based
-on their common key (block name in this case) and output creation time 
+we parse, {\it cache insertion} and {\it indexing time} represents time spend to add objects into
+the DAS cache, {\it aggregation time} is required to merge objects into DAS records based
+on their common key (block name in this case) and {\it output creation time}
 is a time required to write DAS records to disk. Please note that last two
-components, aggregation and output creation time, were only applied to
-final DAS step and not to individual components.
+components, {\it aggregation} and {\it output creation time}, were only applied to
+final DAS step and not to individual DAS components.
 
@@
 majority of the time was spent on retrieval, 83 seconds, and parsing, 29 seconds.
-The remaining 65 seconds was spent on remapping, cache insertion, indexing. The
-numbers for DBS system were a like. As you can see from Table \ref{DAS_benchmark},
-the time spent to read DAS records (last column) from cache was quite
+The remaining 65 seconds was distributed among remapping, cache insertion, indexing. The
+numbers for DBS system were alike. As you can see from Table \ref{DAS_benchmark},
+the time spent to read DAS records (last column) from the cache was quite
 reasonable, roughly 2.5 seconds from both systems. While final time to
@@
 \section{Summary}
-We presented new data aggregation service (DAS) developed for High-Energy experiment
-at LHC, CERN, Geneva, Switzerland. It was designed to providing caching and
+We presented new data aggregation service (DAS) developed for CMS High-Energy experiment
+at LHC, CERN, Geneva, Switzerland. It was designed to provide caching and
 aggregation layer on top of existing relational and non-relation data-service
-in mostly real time fashion. All the data were retrieved on demand basis,
-while a pre-fetch of most common queries is in our plans. The DAS system reach
-first is under commissioning phase right now and prelimiary performance studies
-shown it can accommodate needs of CMS experiment. The CMS experiment started to
+mostly in real time fashion. All the data were retrieved on demand basis,
+while data pre-fetching of most common queries is in our plans. The DAS system 
+is under commissioning phase right now and prelimiary performance studies
+show it can accommodate needs of CMS experiment as general purpose meta-data
+service. The CMS experiment is started to
 collect data from November 2009 and we expect to write about PB of data each
-year to tape. In addition the Monte-Carlo samples will be produced at this scale.
+year to tape. In addition, the Monte-Carlo samples will be produced at this scale.
 Based on test studies performed in CMS, it can translate to the order of
@@
 
-\bibitem{LumiDB}
-https://twiki.cern.ch/twiki/bin/view/CMS/CMS-DMWM-DBS-Luminosity
+%\bibitem{LumiDB}
+%https://twiki.cern.ch/twiki/bin/view/CMS/CMS-DMWM-DBS-Luminosity
 
-\bibitem{DQ}
-Data-Quality DB reference
+%\bibitem{DQ}
+%Data-Quality DB reference
 
-\bibitem{Overview}
-https://cmsweb.cern.ch/overview/
+%\bibitem{Overview}
+%https://cmsweb.cern.ch/overview/
 
@@
 GMW’07, June 25, 2007, Monterey, California, USA.
+
+\bibitem{CRAFT09}
+https://twiki.cern.ch/twiki/bin/view/CMS/CRAFT09AnalysisInfo
 

### 2010-05-04 src/python/DAS/services/dbs/dbs_service.py
One more round w/ notations optimization.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@630 4525493e-7705-40b1-a816-d608a930855b

@@ -4,8 +4,8 @@
 """
 DBS service
 """
-__revision__ = "$Id: dbs_service.py,v 1.17 2009/11/25 18:22:37 valya Exp $"
-__version__ = "$Revision: 1.17 $"
+__revision__ = "$Id: dbs_service.py,v 1.18 2009/11/27 19:17:28 valya Exp $"
+__version__ = "$Revision: 1.18 $"
 __author__ = "Valentin Kuznetsov"
 
 import types
@@ -22,15 +22,20 @@ class DBSService(DASAbstractService):
         self.reserved = ['api', 'apiversion']
         self.map = self.dasmapping.servicemap(self.name, 'javaservlet')
         map_validator(self.map)
+        self.notationmap = self.notations()
+
+    def get_notations(self, api):
+        """Return notations used for given API"""
+        notations = dict(self.notationmap[''])
+        if  self.notationmap.has_key(api):
+            notations.update(self.notationmap[api])
+        return notations
 
     def parser(self, source, api, args=None):
         """
         DBS data-service parser.
         """
-        notationmap = self.notations()
-        notations = notationmap[''] # use api='', i.e. notations valid for all APIs
-        if  notationmap.has_key(api):
-            notations.update(notationmap[api])
+        notations = self.get_notations(api)
         add = None
         if  api == 'listBlocks':
             tag = 'block'
@@ -41,5 +46,6 @@ class DBSService(DASAbstractService):
                 % (self.name, api)
             raise Exception(msg)
         gen = xml_parser(notations, source, tag, add)
+#        gen = xml_parser(source, tag, add)
         for row in gen:
             yield row

### 2010-05-04 doc/paper.tex
Added new numbers
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@631 4525493e-7705-40b1-a816-d608a930855b

@@
 both relational and non-relational data sources. 
-Even though we can access it via viriety of data-services, twikies, blogs, etc.,
+Even though we can access it via set of data-services, twikies, blogs, etc.,
 there is no coherent way for information discovery from different sources
@@
 indexes and delivers information from different relational and 
-non-relational data sources to answer cross data-source queries 
-to explore metadata associated to petabytes of experimental data. 
+non-relational data sources to answer cross data-service queries 
+and explore meta-data associated with petabytes of experimental data. 
 We combine simplpicity of keyword-based search with precision of RDMS
@@
 Physics (HEP), promising to deliver a few PB of data each year. 
-At this scale scientists are facing a new set of problems in an area of
+At this scale scientists are facing with a new set of problems in an area of
 ``information discovery'' within heterogeneous, distributed environment.
 The data and associated meta-data are produced in variety of forms and digital formats.
-They are stored into relational and non-releational data-sources, such as 
-RDMS systems, document oriented databases, blogs, twikies, customized applications, etc. 
-and its access is restricted by broad variety of security policies. 
+They are stored and retrieved from relational and non-releational data-sources, such as 
+RDMS systems, document oriented databases, blogs, twikies, file systems and
+customized applications, etc. Its access is restricted by broad variety of security 
+policies enforced by individual sites and community in general.
 Working in such environment is not easy and require a lot of knowledge, but users
@@
 tool to look-up their desired data. A well-known solutions, such as web-services,
-blogs, twikies and search engines targeted and tighted to specific data
-sources and end-users are left with a way to bookkeep and relatate information
+blogs, twikies and search engines targeted and tighted to a specific data
+sources and end-users are left with a manual task to bookkeep and relatate information
 from them.
@@
 ability to query, search and aggregate information from different 
-data-services as well as serves as a caching layer for them, 
+data-services as well as serve them as a caching layer, 
 preserving their security policies.
@@
 our discussion as following. In section \ref{RelatedWork} we discuss
-related work in a domain of keyword search over relation data sources.
-The section \ref{DataModel} describe CMS experiment and its data model. In section
+related work in a domain of keyword search over relational data sources.
+The section \ref{DataModel} describes CMS experiment and its data model. In section
 \ref{DAS} we outline architecture of the DAS system, including discussion of its
@@
 in last several years. The federated DB \cite{FedDB} by IBM unifies data coming 
-from different RDMS into another DB where SQL queries can be placed to search desired
-data. The other approaches \cite{DBXplorer, QueryAnswer}, influenced by great success
-of search engines, explored ability to use keyword search algorithms over relataional RDMS
-by indexing DB content. In former case, you still face a problem with understanding 
-of underlying schema and imposing relational conditions in your query, while 
-in later queries can be expressed via simple keyword based search, even though exact match
-of provided keywords is expected. Although keyword based search is
-intuitive and easy to use, with respect to HEP and other domains it is not sufficient. 
-For example, it does provide provide ability to place conditions expressions between
-keys (DB entities/attributes) and their values. The proximity of results becomes very poor
-or impossible in case of numbers, aggregate functions, etc. To address this issues
+from different RDMS systems into federated DB where SQL queries can be placed to 
+search desired data. The other approaches \cite{DBXplorer, QueryAnswer}, influenced by 
+great success of search engines, explored ability to use keyword search algorithms 
+over relataional RDMS by indexing DB content. In former case, a problem with understanding 
+of underlying schema and imposing relational conditions in a query still exists, while 
+in later queries can be expressed via simple keyword based search, even though the exact match
+of provided keywords is expected. Although those solutions exist,
+with respect to HEP and other domains they are not sufficient.
+In a case of federeated database exact knowledge of schema is required. This means
+that end-users should be aware of database internal sctructure in order to use SQL.
+While the proposed keyword search based systems
+work in confined environment and do not provide ability to use condition expressions
+over relational data. For instance, as quoted in \cite{QueryAnswer}, if someone asks about 
+{\it Woody Allen} a set of database records can be found to satisfy this request.
+Such approach cannot be used to answer the logical questions, such as 
+{\it find me total number of files whose size more then 20 but less then 100 taken from january
+to february of 2009} which is easy to address using SQL syntax.
+
+To address those issues
 an attempt to build a simple, intuitive and flexible query language was introduced
-in \cite{DBS-QL}.\footnote{The ATLAS experiment at CERN, has been developed
-AMI web portal \cite{AMI} with similar functionality.}
+in \cite{DBS-QL, AMI}.
+%\footnote{The ATLAS experiment at CERN, has been developed
+%AMI web portal \cite{AMI} with similar functionality.}
 It represented a power of SQL while
-hiding underlying relational schema from the end users. As a results
+hiding underlying relational schema from the end-users. As a results
 a human questions were intuitively mapped into simple queries. For example,
@@
 {\it I'm looking for files who contain data taken on certain date and located at
-particular site} was represented as simple as
+particular site} was represented as simple as \cite{DBS-QL}
 \begin{verbatim}
@@
 \end{verbatim}
-We wanted to expand this approach to handle broad variety of relational and
-non-relational data-sources in CMS within our distributed heterogeneous environment.
-At the same time we wanted to simplify it up to the level that users will be
-able to place free text-based queries (keyword based search) to look-up their desired data.
+Such approach not only address common use cases within HEP community, but also
+represented an intuitive mapping between mental model of our users and 
+underlying database back-end.
+
+We wanted to expand this approach further and apply it 
+to handle cross data-service queries from broad variety 
+of relational and non-relational data-sources within 
+our distributed heterogeneous environment.
 
 \section{CMS data model\label{DataModel}}
-The CMS, the Compact Muon Solenoid experiment \cite{CMS} 
+The CMS, Compact Muon Solenoid, experiment \cite{CMS} 
 is one of the two large general-purpose particle physics detectors built on 
@@
 ability to look at conditions which were present at early stage of our Universe.
-More then 3000 physicists from 183 institutions prepresenting almost 
+More then 3000 physicists from 183 institutions reprepresenting almost 
 38 countries are involved in design, construction and maintenance of the experiment.
@@
 is designed to process
-and efficiently manage a PBs of data expected to be produced each year
+and efficiently manage a few PBs of data expected to be produced each year
 at LHC. The computing resources provided by members of CMS
@@
 of Grid software. The model allows to cover broad variety of
-hardware, mass and storage elements and configuration of
-clusters. To accomodate the efficient data processing we use 
+hardware, mass and storage elements and configuration of the
+clusters. To accomodate efficient data processing CMS uses
 a multi-Tier distributed model, where specific tasks of data taking,
-processing, archival and distribution assigned to each tier based
-on CMS data model. For example, the Tier-0 center at CERN responsible
-for data archives coming out of detector, prompt first pass reconstruction
+processing, archival and distribution are assigned to each tier based
+on CMS data model. For example, the Tier-0 center at CERN is responsible
+for archiving the data coming out from the detector, prompt first pass reconstruction
 and data distribution to Tier-1 centers. The 7 Tier-1 centers
 located in France, Germany, Ialy, Spain, Taiwan, United Kingdom and United States
-keep a portion (copy) of data delivered by Tier-0 for further processing.
-They provide storage and high priority analysis jobs at their facilities.
+keep a portion (copy) of the data delivered by Tier-0 for further processing.
+They provide storage and CPU power for high priority analysis jobs.
 The Tier-2 centers, located at more then 50 sites around the world,
@@
 centers set up local services for data bookkeeping and operational
-tasks. Based on their budget resources the choice of back-end DB were
-up to site managers. Therefore the CMS software were designed to support different
+tasks. Based on their resources the choice of back-end DB was
+up to the site managers. Therefore the CMS software were designed to support different
 DB back-ends and provide tools for migrating data across them.
@@
 in our distributed environment? Even though individual data-services were designed
-to answer specific questions about data they serve, the ability to relate and search
-information among them was a human task. The growing amount of information
+to answer specific questions about data they serve, the ability to search and relate
+information among them was tedious human task. The growing amount of information
 and desire to make cross-service queries force us to design and develop a new
@@
 \section{Data Aggregation System\label{DAS}}
-The design of DAS system was based on previous studies of CMS 
-data bookkeeping system (DBS) use cases \cite{DBS, DBS07}. We carefully analyzed user
-queries placed to DBS, their patterns, frequencies and latency. As a result
+The design of DAS system was based on previous studies of CMS Data 
+Bookkeeping System (DBS) use cases \cite{DBS, DBS07}. We carefully analyzed user
+requests and queries placed to DBS, their patterns, frequencies and latency. As a result
 we created a data discovery service \cite{DD} which came through several iterations
 of design choices. At the end the end-user interface was based on DBS 
-Query Language (DBS-QL)\cite{DBS-QL} which support dynamic queries against
-DBS back-end and had syntax very closed to free-text based search queries. 
-The DBS-QL \cite{DBS-QL} syntax was based on SQL query language with one important
-difference that we took away from end-users any knowledge about relations of underlying schema.
-It was achieved by usage of short-path algorithm to establish necessary
-table joins based on provided information in user queries and auto-loaded
-schema from DB back-end and knowledge of foreign key relationships.
-A quick adoption and wide spread of DBS-QL give us confidence of choosen approach 
+Query Language (DBS-QL)\cite{DBS-QL} which supports dynamic queries against
+DBS back-end. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language with one 
+important difference, we took away any knowledge about relations of underlying schema
+from end-users. It was achieved by usage of Dijkstras shortest path algorithm 
+to establish necessary table joins based on provided information in user queries and
+knowledge of foreign key relationships from auto-loaded database schema.
+The DBS-QL keys exposed to end-users were choosen based on jargon used by
+physicists. They're internally mapped into database entitites and attributes, 
+e.g. {\it run} DBS-QL key refers to run number attribute of run summary table.
+A quick adoption and wide usage of DBS-QL gave us confidence of choosen approach 
 and provide a basis for building DAS system.
@@
 of the system to successful information discovery. 
-Starting from this ground we designed DAS system as
+Starting from this ground we designed DAS system as an
 additional layer on top of the existing data-services
-within CMS computing infrastructure by requiring the following:
+within CMS computing infrastructure by imposing the following set of requirements:
 \begin{itemize}
-\item support flexible and intuitive Query Language (QL) to easy look-up our
-data;
+\item support flexible and intuitive Query Language (QL) to support
+both keyword and condition queries;
 \item pluggable interface to existing and up-coming data-service APIs;
 \item support heterogeneous software environment and distributed nature of data-services;
-\item preserve security policies of individual data-services while provide unique
-authentication schema for end-users;
+\item preserve security policies of individual data-services;
 \item be flexible to adjust to dynamic nature of data-service APIs, 
@@
 multiple data-services;
-\item be transparent to data-service data formats;
-\item support legacy application, data-services and APIs.
+\item be transparent to data-service's implementations and data formats;
+\item support legacy data-services, applications and APIs.
 \end{itemize}
@@
 DAS architecture diagram. It consists of DAS cache server, mapping and analytics DBs.
-Once query was placed to DAS an associated
-data-service API was invoked, triggering DAS cache population. Once data become
-available in DAS cache user were able to see the results. This on demand
-query retrieval system, where data were retrieved and store into cache from
-back-end data service. We use DAS analytics DB to establish cache population
-for most common queries by using DAS robots.
+Once query was placed to DAS the associated
+data-service APIs were invoked, triggering DAS cache population. Once data become
+available and aggregated in DAS cache users were able to see the results. This on demand
+system, where data were retrieved and store into cache from
+back-end data services, was supplemented by DAS robots (unix daemons) who
+were consulting DAS analytics database to enable data population into DAS
+cache for most common queries.
 }
@@
 
-The DAS core communicates with different data-services and retrieve
+The DAS core communicates with different data-services and retrieves
 data on demand. This was done by invoking appropriate data-service APIs
 upon user requests. All user queries were written to Analytics DB
-for further analysis. The data-service output where re-mapped into
-DAS notations and if necessary aggregated inside of DAS cache.
+for further analysis. The data-service output were re-mapped into
+DAS notations and, if necessary, aggregated inside of DAS cache.
 Since we didn't know a-priory what kind of data users will ask, 
@@
 of RDMS as DAS back-end was problematic\footnote{In addition DAS 
-didn't impose any requirements for transactions and persistency of the data.} 
+didn't impose any requirements on transactions and persistency of the data.} 
 and we look around for other options. Through the analysis of available IT solutions, 
@@
 key-value databases, documented-oriented databases we made our choice in favor 
-of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
-MongoDB \cite{MongoDB}. Both of them provide schema free document
+of the last technology. Among of them we evaluated CouchDB \cite{CouchDB} and 
+MongoDB \cite{MongoDB}. Both systems provide schema free document
 storage, replication and fail-over features, but our choice in favor of 
-MongoDB was obvious because its support of dynamic queries, 
+MongoDB was obvious because of its support of dynamic queries, 
 full indexies, including on inner objects and embedded arrays,
@@
 values as regular expression patterns as well as mapping between data-service 
-notations into DAS keys. This allows to map user input query parameters into
-data-service API metrics as well as map back data-service output into DAS names.
+notations and DAS keys. It allows us to map user input query parameters into
+data-service API metrics as well as map back data-service output into DAS notations.
 For instance, the DBS system \cite{DBS}
@@
 the same entity was named as {\it lfn}. Since both of them were referred to
-logical file name, we map both notations into DAS key {\it file}.\footnote{We
+the logical file name, we map both notations into DAS key {\it file}.\footnote{We
 also deal with physical file names, but still used {\it file} to refer to them in
-DAS notation, since simple regular expressions were used to identify {\it file}
-meaning from the provided value.}
+DAS notationis, since simple regular expressions were used to identify the
+meaning of the {\it file} keyword from the provided value.}
 For each service we provide a mapping from its own data-format (JSON, XML, etc.) into
@@
 We also store a mapping between DAS keys and data-service API names to
-establish appropriate calls to data-service upon provided DAS query. For example,
-when user specifies {\it file} in their DAS query, we invoke DBS listFiles and PhEDEx
-fileReplicas APIs.
+establish appropriate calls to data-services upon provided DAS query. For example,
+when user specifies {\it file} in their DAS query, we invoke DBS 
+{\it listFiles} and PhEDEx {\it fileReplicas} APIs, respectively.
 
-The DAS Analytics DB plays special role. It collects information
+The DAS Analytics DB plays a special role. It collects information
 about user requests placed to the system. Each request was recorded. Upon its
 decomposition into set of selection keys and conditions we also record
-API name and API input parameters and keep updating their counter for repeated
-calls. This allows to keep track of frequency of API calls and establish pre-fetch
-strategy for most requested APIs.
+API name and API input parameters into Analytics database and 
+keep updating their counter information for repeated calls. 
+This allows to keep track of frequency of API calls and establish 
+pre-fetch strategies for most common queries.
 
@@
 constantly monitor Analytics DB for most popular requests and pre-fetch most
-common data in DAS cache.
+common data into DAS cache.
 
@@
 cache was checked for a similar set of selection and condition
-keys which can be placed earlier.\footnote{Since DAS doesn't hold
+keys which can be used earlier.\footnote{Since DAS doesn't hold
 any data by default we cannot rely on data being present in a cache.
-Moreover if we look-up for data first, we cannot tell if it was complete
-set since earlier queries may retrieve only portion of the requested data.}
+Moreover if we look-up data first, we cannot tell if it was complete
+set or not, since earlier queries may retrieve only portion of the requested data.}
 If the superset was found a data were
@@
 {\it \{"fields" : ["dataset"], "spec" : \{"site" : "T1\_CH\_CERN"\}\}}.
-Such syntax was perfect fit for us, since it allows us to store all queries
+Such syntax was perfect fit for us, since it allows to store all queries
 as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
 MongoDB query syntax is very reach. It supports advanced queries with 
-boolean expressions, conditional operators, regular expression, etc. providing
+boolean expressions, conditional operators, regular expressions, etc. providing
 a real power for our end-users.
@@
 \subsection{Data-services and aggregation}
-DAS provided a pluggable interface to add new CMS data-services. 
-Each data-service plugin consists of its DAS map and individual data parser. 
-The data-service DAS map consists of information about data-service APIs,
+DAS provided an interface to add new CMS data-services via data-service plugins.
+Each plugin consists of its DAS map and data parser suitable for data-service output
+format(s).
+The data-service DAS map contains information about data-service APIs,
 input and output parameters, data-service
 format and their mapping to DAS keys as well as relation to other sub-systems. 
-The information was stored into DAS Mapping DB. Below you can see a typical
-record of SiteDB \cite{SiteDB} data-service:
+Below you can see a typical DAS Mapping DB record for CMS
+SiteDB data-service \cite{SiteDB}\footnote{We used regular expression 
+patterns, evaluated by python during run time
+execution, to validate provided set of input paremters and DAS keys.}:
 \begin{verbatim}
@@
 \end{verbatim}
-It contains information about API name and parameters, mapping between API input
-parameter and DAS key, CMS system name and set of DAS keys this API is covering.
-We also used regular expression patterns, evaluated by python during run time
-execution, to verify provided set of input paremters and DAS keys.
+This information was stored into DAS Mapping DB.
 
@@
 source of information about all data-services participating in DAS and their
-relationships. Based on this information we were able to retrieve appropriate data
+relationships. Based information stored there we were able to retrieve appropriate data
 from different data-services, re-map them into DAS notations and
@@
 while PhEDEx stores information about the file location, such as sites which
-holds copy of the files and details about storage elements.
+holds copy of the files and details about site storage elements.
 Upon a user request to find a file, we were able to identify appropriate set of
@@
 contained a file name in them. We use it as a common key to merge records in DAS cache.
-This merged (aggregated) results represented DAS data objects. Each DAS data
-object contains a standard DAS header, which identify look-up time,
+This merged (aggregated) results represent DAS records. Each DAS
+record contains a standard DAS header, which identify look-up time,
 request url, input parameters and API method as well as response version, expiration
@@
 \end{verbatim}
-It represents aggregated DAS record between two SiteDB APIs, CMStoSAMName and
-CMStoSiteName, where API results are merged under {\it site} key. The information
-under {\it das} key shows DAS header, which contains sufficient amount of
-information about API name, location, execution time and lifetime of the data.
-Using this information was valuable contribution in debugging process of
+This DAS record represents aggregated information between two SiteDB APIs, 
+{\it CMStoSAMName} and {\it CMStoSiteName}, where API results were merged 
+under the {\it site} key. The information provided by {\it das} key shows 
+DAS header, which contains information about API name, location, execution 
+time and lifetime of the data.
+Usage of this information was valuable contribution in debugging process of
 data-services themselves, e.g. identification of not-matched records, 
@@
 
-We used REST model \cite{REST} for all data requests to DAS.
+We used RESTful model \cite{REST} for all DAS cache server APIs.
 It allows us to share core library and other components during development of 
 various DAS clients, e.g. CLI, web interface, robots, etc. and hides
-complexity of access to individual data-services and security policies.
+complexity of access to individual data-services and their security policies
+as well as proper organize caching solution.
 For instance, to access on-line Run Summary information we were forced to use
 Grid certificates. Instead of propagating user certificate to data-service
-the DAS works as a proxy and used host certficiate to serve such requests.
+the DAS cache server works as a proxy to this server and used host certficiate 
+to serve such requests on behalf of the users.
 
@@
 \begin{itemize}
-\item DBS \cite{DBS}, Data Bookkeeping System which collect information
+\item DBS \cite{DBS}, Data Bookkeeping System, which collect information
 about CMS meta-data;
-\item PhEDEx \cite{PhEDEx}, Physics Experiment Data Export project which
+\item PhEDEx \cite{PhEDEx}, Physics Experiment Data Export project, which
 provides the data placement and the file transfer system for the CMS experiment;
-\item SiteDB \cite{SiteDB} is a catalogue of all CMS computing Tiers. 
-It records the CMS resources at the site, the resources pledged for the 
+\item SiteDB \cite{SiteDB} which 
+records the CMS resources at the site, the resources pledged for the 
 future, and it keeps track of CMS personnel at each site, including the 
 roles and duties they fulfill in the collaboration;
-\item RunSummary DB \cite{RunSummary} collects information about run and triggers
+\item RunSummary DB \cite{RunSummary} which collects information about run and triggers
 conditions during data-taking;
@@
 %transfer rate, etc.;
-\item Dashboard \cite{Dashboard} project for LHC experiments aims to 
-provide a single entry point to the monitoring data collected from the 
-distributed computing systems of the LHC virtual organization.
+\item Dashboard \cite{Dashboard} which monitors data collected from the 
+distributed computing systems of the CMS experiment and LHC virtual organization
+in general.
 \end{itemize}
 Each data-service has its own scope, size and lifetime. For instance the data
-in PhEDEx are transient due to constant migration of CMS data with daily DB size of
-around few tens of GBs. The DBS system were divided into dozen of individual instances, 
+in PhEDEx are transient due to constant migration of CMS data, 
+the DBS system were divided into dozen of individual instances, 
 who has been used by different production teams for various bookkeeping tasks.
-Figure \ref{db_size} shows current size of meta-data from aforementioned data-service.
+Figure \ref{db_size} shows current size of meta-data from aforementioned data-services.
 \begin{figure}[htb]
@@
 \caption{
-Database size of CMS data-services participating in DAS in November 2009, before
-data taking.
+Current size of meta-data for CMS data-services participated in DAS, before
+official data-taking. We anticipate to collect $\sim$500GB of meta-data
+per year once LHC starts operating.
 }
@@
 At the moment we don't have exact numbers for amount of data which will stored to DAS,
-but based on test performed during CRAFT09 \cite{CRAFT09}
+but based on test performed during 
+Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}
 exercise we produced 0.3PB of data which
-translate into $\sim50$ GB of meta-data stored into DBS system. Therefore
+translate into $\sim50$ GB of meta-data stored in DBS system. Therefore
 we can anticipate to collect $\sim500$GB of meta-data per year during data-taking period
@@
 
-We did preliminary studies of DAS performance based on information stored
-in 2 systems, DBS and PhEDEx. We benchmark time required to collect and 
-aggregate block objects information from both system, see Table \ref{DAS_benchmark}.
+Preliminary results of DAS performance are shown in Table \ref{DAS_benchmark}.
+We benchmark time required to collect and aggregate block object information
+from DBS and PhEDEx CMS data-services.
 
@@
 \hline
-DAS (DBS) & XML & 187MB & 346820 & 165 sec & 2.53 sec \\
-DAS (PhEDEx) & JSON & 164MB & 230035 & 177 sec & 2.42 sec \\
-DAS (aggregate) & JSON & 357MB & 348866 & 386 sec & 41.7 sec \\
+DAS (DBS) & XML & 187MB & 368716 & 53 sec & 0.92 sec \\
+DAS (PhEDEx) & XML & 164MB & 175884/170131 & 160 sec & 0.94 sec \\
+DAS (aggregate) & JSON & 357MB & 374468 & 256 sec & 44.8 sec \\
 \hline
@@
 \end{tabular}
-\caption{DAS benchmark results. We performed a stress test with 2 CMS system, which
-both deliver block object information. The total number of DAS records is calculated
-as number of aggregated records and left over from both systems.}
+\caption{Time required to fetch, parse and aggregate block information
+from DBS and PhEDEx systems via their APIs. In the case of PhEDEx
+system we show total number of fetched records 175884 together with
+number of records merged in DAS 170131. The total number of DAS records 
+were calculated as number of aggregated records plus left over, non-matched records,
+from both systems.}
 \label{DAS_benchmark}
@@
 
-We measured elapsed time as the following:
+Elapsed time measured during this test consist of the following componenets:
 \begin{itemize}
@@
 {\it
-Elapsed time = retrieval time + parsing time + remapping time 
+Elapsed time = retrieval time + parsing time + re-mapping time 
         + cache insertion/indexing time 
@@
 \end{itemize}
-Here the {\it retrieval time} is a time required to access data from remote data-services,
-{\it parsing time} is a time required to read and parse received data, {\it remapping time}
+Here the {\it retrieval time} is a time required to access data from remote data-service,
+{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
 is a time to convert notations used by data-service to DAS ones for every object
@@
 components, {\it aggregation} and {\it output creation time}, were only applied to
-final DAS step and not to individual DAS components.
+final DAS step and not to time spent in individual data-services.
+
+%DAS execution time (phedex) 160.606897116 sec
+%DAS execution time (dbs) 52.5349619389 sec
+%DAS execution time 256.0670228 sec, Sat, 28 Nov 2009 15:46:39 GMT
+%
+%DAS execution time (phedex) 0.952513933182 sec
+%DAS execution time (dbs) 0.951184034348 sec
+%DAS execution time 44.7982931137 sec, Sat, 28 Nov 2009 15:52:11 GMT
 
 The individual DAS components (DBS and PhEDEx) shown roughly the same performance.
-For example, in case of PhEDEx the total time spent in DAS was 177 seconds, where
-majority of the time was spent on retrieval, 83 seconds, and parsing, 29 seconds.
-The remaining 65 seconds was distributed among remapping, cache insertion, indexing. The
-numbers for DBS system were alike. As you can see from Table \ref{DAS_benchmark},
+Although the time of PhEDEx component was divided among: retrieval + parsing 
+time $\sim$110 sec and merging step $\sim$50 sec.
+
+As you can see from Table \ref{DAS_benchmark},
 the time spent to read DAS records (last column) from the cache was quite
-reasonable, roughly 2.5 seconds from both systems. While final time to
-get DAS records on disk was below 42 seconds. This was mostly due to I/O operations.
+reasonable, roughly 1 second for both systems. While final time to
+get DAS records on disk was about 45 seconds. This was mostly due to I/O and
+conversion between binary data format used in database (BSON) to JSON.
+All operations were performed on a single CPU core and obviously can be done
+in parallel.
 
-Based on this results we measured that we can achieve 900 docs/sec for creation of DAS
-in a cache and 8500 docs/sec for reading DAS records back, which is suitable for our
+Based on this results we measured that we can achieve 7000 docs/sec rate
+for insertion, 3400 docs/sec for aggregation and 8500 docs/sec for reading operation
+in DAS, which is suitable for our 
 needs.\footnote{We've been satisfied with such performance,
-even though our measurements shown that it can still be speed up by factor of
-10 if we will use C++ DB driver.}
+even though our measurements shown that it still can be speed up by few times
+using C++ DB driver instead of python one used in this prototype system.}
 
@@
 at LHC, CERN, Geneva, Switzerland. It was designed to provide caching and
-aggregation layer on top of existing relational and non-relation data-service
+aggregation layer on top of the existing relational and non-relation data-services
 mostly in real time fashion. All the data were retrieved on demand basis,
-while data pre-fetching of most common queries is in our plans. The DAS system 
-is under commissioning phase right now and prelimiary performance studies
-show it can accommodate needs of CMS experiment as general purpose meta-data
-service. The CMS experiment is started to
+while data pre-fetching of most common queries is in our plans. We developed
+prototype in python for DAS system which is under commissioning phase right now 
+and performed performance studies shown in this paper. 
+The CMS experiment is started to
 collect data from November 2009 and we expect to write about PB of data each
 year to tape. In addition, the Monte-Carlo samples will be produced at this scale.
-Based on test studies performed in CMS, it can translate to the order of
-$\sim500$GB of meta-data each year. The DAS should sustain such load and provide
-interface as data discovery service in CMS.
+Based on test studies performed in CMS, we expect that total size of
+meta-data produced by experiment each year will be of the order of
+$\sim500$GB. The DAS should sustain such load and will provide generic 
+``data-discovery'' service for CMS experiment in years to come.
 

### 2010-05-04 doc/paper.tex
Made 1/2 pass through the paper, up to 4.1 chapter. A lot of changes. Simon please have a look.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@632 4525493e-7705-40b1-a816-d608a930855b

@@
 It is organized and managed by constantly evolving software using 
-both relational and non-relational data sources. 
-Even though we can access it via set of data-services, twikies, blogs, etc.,
-there is no coherent way for information discovery from different sources
-in heterogeneous, distributed environment with different security policies enforced.
+both relational and non-relational data sources. Even though we can apply
+information retrieval approach to non-relation data sources,
+we can't do so for relational ones, where information can be accessed via
+pre-established set of data-services.
+%Even though we can access it via set of data-services, twikies, blogs, etc.,
+%the information discovery across them is still a manual task.
+% which prevent
+%information discovery
+%there is no coherent way for information discovery from different sources
+%in heterogeneous, distributed environment with different security policies enforced.
 
@@
 We combine simplpicity of keyword-based search with precision of RDMS
-under a new system where aggregated information being collected from various sources 
+under the new system. The aggregated information being collected from various sources,
 allowing end-users to place dynamic queries, get precise answers and 
@@
 based on studies  of use cases of the CMS particle physics experiment at 
-the LHC which uses a large scale, distributed computing system to motivate the work.
+the LHC where we used a large scale, distributed computing system 
+to motivate the work.
 
@@
 \section{Introduction}
-The CERN, the European Organization for Nuclear Research, plays a leading
-role in fundamental studies of physics. It is also known as a place where
-World Wide Web was born to help researches share information among each other
-via concept of hypertext.
+The European Organization for Nuclear Research, known as CERN, plays a leading
+role in fundamental studies of physics. But it is also known as a place where
+many innovations in the area of computer science were developed, e.g. World Wide Web.
 Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
 Physics (HEP), promising to deliver a few PB of data each year. 
-At this scale scientists are facing with a new set of problems in an area of
-``information discovery'' within heterogeneous, distributed environment.
+At this scale the ``information discovery'' within heterogeneous, distributed 
+environment become a key ingredient of successful data analysis in HEP.
+%scientists are facing with a new set of problems in an area of
+%``information discovery'' within heterogeneous, distributed environment.
 The data and associated meta-data are produced in variety of forms and digital formats.
@@
 RDMS systems, document oriented databases, blogs, twikies, file systems and
-customized applications, etc. Its access is restricted by broad variety of security 
-policies enforced by individual sites and community in general.
-Working in such environment is not easy and require a lot of knowledge, but users
+customized applications, etc. 
+%Its access is restricted by broad variety of security 
+%policies enforced by individual sites and community in general.
+Working in such environment requires a lot of expertise and our users
 (in this case physicists) are always looking for simple, intuitive and flexible
-tool to look-up their desired data. A well-known solutions, such as web-services,
-blogs, twikies and search engines targeted and tighted to a specific data
-sources and end-users are left with a manual task to bookkeep and relatate information
-from them.
+tool(s) to look-up their desired data. A well-known solutions, such as data-services
+and search engines tighted to a specific data sources and end-users are left 
+with a manual task to bookkeep and relatate information from them.
 
 Here we present a work on Data Aggregation System (DAS) designed for
-CMS High-Energy Experiment operated at LHC which provide
+CMS High-Energy Experiment at LHC which provides
 ability to query, search and aggregate information from different 
-data-services as well as serve them as a caching layer, 
-preserving their security policies.
+data-services and trigger information retrieval on demand.
 
-We organize
-our discussion as following. In section \ref{RelatedWork} we discuss
-related work in a domain of keyword search over relational data sources.
-The section \ref{DataModel} describes CMS experiment and its data model. In section
-\ref{DAS} we outline architecture of the DAS system, including discussion of its
-various components. Finally our results are summarized in section \ref{Results}.
+The rest of this paper is organized as following. 
+In section \ref{RelatedWork} we discuss related work in a domain of 
+keyword search over relational data sources.
+Section \ref{DataModel} describes CMS experiment and its data model. In section
+\ref{DAS} we present architecture of the DAS system, including discussion of its
+various components. Finally, our results are summarized in section \ref{Results}.
 
@@
 \section{Related Work\label{RelatedWork}}
-Even though the idea of querying relation databases via keyword based search
-algorithms is not knew it is still under significant activity in computer
-science domain. A few alternative solutions has been proposed to address this issue
-in last several years. The federated DB \cite{FedDB} by IBM unifies data coming 
-from different RDMS systems into federated DB where SQL queries can be placed to 
-search desired data. The other approaches \cite{DBXplorer, QueryAnswer}, influenced by 
-great success of search engines, explored ability to use keyword search algorithms 
-over relataional RDMS by indexing DB content. In former case, a problem with understanding 
-of underlying schema and imposing relational conditions in a query still exists, while 
-in later queries can be expressed via simple keyword based search, even though the exact match
-of provided keywords is expected. Although those solutions exist,
-with respect to HEP and other domains they are not sufficient.
-In a case of federeated database exact knowledge of schema is required. This means
-that end-users should be aware of database internal sctructure in order to use SQL.
-While the proposed keyword search based systems
-work in confined environment and do not provide ability to use condition expressions
-over relational data. For instance, as quoted in \cite{QueryAnswer}, if someone asks about 
-{\it Woody Allen} a set of database records can be found to satisfy this request.
-Such approach cannot be used to answer the logical questions, such as 
-{\it find me total number of files whose size more then 20 but less then 100 taken from january
-to february of 2009} which is easy to address using SQL syntax.
+Even though the idea of keyword queries over the relational database(s)
+%querying relation databases via keyword based search
+%algorithms 
+is not knew it is still under discussion in computer science domain.
+A few alternative solutions has been proposed to address this issue
+in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
+i.e. retrieval of only documents that contain all query keywords, has
+been discussed. It requires that all keywords are matched in a row from
+single table or joined tables. While \cite{QueryAnswer} make a step forward
+by considering information around matched values and presents customized answers to
+end-users over the generated new database. Those and other works\footnote{
+For more information about
+keyword based search over relational databases we redirect users to reference in
+\cite{DBXplorer, QueryAnswer}.} 
+are based on some kind of the graph over the underlying database and
+moreover use a generated database out of the pristine one.
+The proximity of the results in those techniques 
+are mostly confined to the text-based values stored in a database, since the
+keyword based search with numerical values 
+need the precise knowledge of the related entity in a schema. For example, typing
+100 can lead to inappropriate matches, such as row ids, and in order to be
+answered correctly it requires additional keyword, clarifying its context, 
+which should be matched with particular
+entity (table.column) in the underlying database.
+Therefore such approaches cannot be used to answer the logical questions, such as  
+{\it find me total number of files whose size more then 20 but less then 
+100 taken from january to february of 2009} which can be easily accomplished
+using SQL syntax.
 
-To address those issues
-an attempt to build a simple, intuitive and flexible query language was introduced
-in \cite{DBS-QL, AMI}.
-%\footnote{The ATLAS experiment at CERN, has been developed
-%AMI web portal \cite{AMI} with similar functionality.}
-It represented a power of SQL while
-hiding underlying relational schema from the end-users. As a results
-a human questions were intuitively mapped into simple queries. For example,
-the question
+To address those limitations an attempt to build a simple, intuitive and 
+flexible query language was introduced in \cite{DBS-QL}\footnote{A similar but less
+sophisticated approach was done in \cite{AMI}}.
+It represents a power of SQL while hiding underlying relational schema from 
+the end-users. As a results a human questions were intuitively mapped into 
+simple queries. For example, the question
 {\it I'm looking for files who contain data taken on certain date and located at
@@
 \end{verbatim}
-Such approach not only address common use cases within HEP community, but also
-represented an intuitive mapping between mental model of our users and 
-underlying database back-end.
+This solution represents a common use cases within HEP community and provide
+intuitive mapping between mental model of end-users and underlying database back-end.
 
-We wanted to expand this approach further and apply it 
-to handle cross data-service queries from broad variety 
-of relational and non-relational data-sources within 
-our distributed heterogeneous environment.
+At the same time the data aggregation is commonly addressed in the context of
+federated databases, see for example \cite{FedDB}. It unifies data coming 
+from different RDMS systems into federated DB where SQL queries can be placed to 
+search desired data. Even though it would be interesting to apply technique 
+discussed in \cite{DBS-QL} to a federated database it still requires 
+an access to the underlying database schema. This obviously creates a problem
+working in distributed heterogeneous environment where different security policies
+are in place.
+
+To avoid those limitations we decided to build a keyword-search based system around
+existing data-services whose nature and access policies are defined by developers.
+Keeping in mind that meta-data can come in different forms and data formats
+we used a document-oriented database back-end with full support of query language.
+The key ingredients were the setup of the mapping and analytics services which provide
+necessary glue across existing API/notations and user queries. This approach 
+allows us to perform the data aggregation across various data-services, achieve
+precise answers using keyword-based queries
+and organize information retrieval on demand.
+%to store meta-data information 
+%We wanted to expand this approach further and apply it 
+%to handle cross data-service queries from broad variety 
+%of relational and non-relational data-sources within 
+%our distributed heterogeneous environment.
 
 \section{CMS data model\label{DataModel}}
-The CMS, Compact Muon Solenoid, experiment \cite{CMS} 
+The CMS, Compact Muon Solenoid, \cite{CMS} 
 is one of the two large general-purpose particle physics detectors built on 
 the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
-It is designed to explore the frontier of High Energy Physics and provide physicsists
-ability to look at conditions which were present at early stage of our Universe.
+It is designed to explore the frontier of High Energy Physics and provide physicists
+ability to look at conditions presented at early stage of our Universe.
 More then 3000 physicists from 183 institutions reprepresenting almost 
-38 countries are involved in design, construction and maintenance of the experiment.
+38 countries are involved in the design, construction and maintenance of the experiment.
 
@@
 is designed to process
-and efficiently manage a few PBs of data expected to be produced each year
-at LHC. The computing resources provided by members of CMS
+and efficiently manage a few PBs of data expected each year
+in CMS operation at LHC. The computing resources provided by members of CMS
 collaboration are geographically distributed, 
@@
 centers set up local services for data bookkeeping and operational
-tasks. Based on their resources the choice of back-end DB was
+tasks. Based on their resources the choice of database back-end was
 up to the site managers. Therefore the CMS software were designed to support different
@@
 in our distributed environment? Even though individual data-services were designed
-to answer specific questions about data they serve, the ability to search and relate
-information among them was tedious human task. The growing amount of information
-and desire to make cross-service queries force us to design and develop a new
-type of system, the Data Aggregation System.
+to answer specific questions about the data they serve, the ability to search and relate
+information among them was always a tedious human task. The growing amount of information
+and desire to make cross-service queries force us to design and develop new
+Data Aggregation System.
 
 \section{Data Aggregation System\label{DAS}}
-The design of DAS system was based on previous studies of CMS Data 
-Bookkeeping System (DBS) use cases \cite{DBS, DBS07}. We carefully analyzed user
-requests and queries placed to DBS, their patterns, frequencies and latency. As a result
-we created a data discovery service \cite{DD} which came through several iterations
-of design choices. At the end the end-user interface was based on DBS 
-Query Language (DBS-QL)\cite{DBS-QL} which supports dynamic queries against
-DBS back-end. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language with one 
-important difference, we took away any knowledge about relations of underlying schema
-from end-users. It was achieved by usage of Dijkstras shortest path algorithm 
+The design of the DAS system was based on previous studies of CMS Data 
+Bookkeeping System (DBS) \cite{DBS, DBS07}. We carefully analyzed user
+requests and queries placed into the DBS, their patterns, frequencies and latency. 
+As a result we created a data discovery service \cite{DD}. Originally it used MVC
+architecture and later, being re-factored as pure presentation layer on top of 
+the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
+DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
+we took away relations and details on underlying schema from end-users. 
+This was achieved by usaging Dijkstras shortest path algorithm 
 to establish necessary table joins based on provided information in user queries and
 knowledge of foreign key relationships from auto-loaded database schema.
-The DBS-QL keys exposed to end-users were choosen based on jargon used by
+The DBS-QL keys exposed to end-users were choosen based on a jargon used by
 physicists. They're internally mapped into database entitites and attributes, 
-e.g. {\it run} DBS-QL key refers to run number attribute of run summary table.
-A quick adoption and wide usage of DBS-QL gave us confidence of choosen approach 
+e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
+A quick adoption and wide usage of DBS-QL gave us confidence in choosen approach 
 and provide a basis for building DAS system.
@@
 \begin{itemize}
-\item support flexible and intuitive Query Language (QL) to support
-both keyword and condition queries;
+\item support of keyword based search queries with ability to use condition operators;
 \item pluggable interface to existing and up-coming data-service APIs;
-\item support heterogeneous software environment and distributed nature of data-services;
+\item support heterogeneous software environment and distributed nature of the data-services;
 \item preserve security policies of individual data-services;
-\item be flexible to adjust to dynamic nature of data-service APIs, 
-due to continuously changing detector and software environments;
+%\item be flexible to adjust to dynamic nature of data-service APIs, 
+%due to continuously changing detector and software environments;
 \item retrieve data on demand and aggregate them if necessary across
 multiple data-services;
-\item be transparent to data-service's implementations and data formats;
-\item support legacy data-services, applications and APIs.
+\item be transparent to data-services implementations and their data formats;
+%\item support legacy data-services, applications and APIs.
 \end{itemize}
-
-%\begin{figure}[htb]
-%\centering
-%\includegraphics[width=150mm]{DAS_architecture.pdf}
-%\caption{
-%DAS architecture.
-%}
-%\label{DAS_arch}
-%\end{figure}
 \begin{figure}[htb]
@@
 \caption{
-DAS architecture diagram. It consists of DAS cache server, mapping and analytics DBs.
-Once query was placed to DAS the associated
-data-service APIs were invoked, triggering DAS cache population. Once data become
-available and aggregated in DAS cache users were able to see the results. This on demand
-system, where data were retrieved and store into cache from
-back-end data services, was supplemented by DAS robots (unix daemons) who
-were consulting DAS analytics database to enable data population into DAS
-cache for most common queries.
+The DAS architecture diagram. 
+Input query where placed into analytics DB and mapped into set of
+data-service APIs and their parameters. If superset of the query has been
+found in DAS cache, the appropriate results were delivered back to users.
+Otherwise a data-service APIs were invoked, triggering DAS cache population. 
+Once data become available and aggregated in DAS cache users were able 
+to see the results. This on demand system was supplemented by DAS robots 
+(unix daemons) who were consulting DAS analytics database to enable data 
+population into DAS cache for most common query requests.
 }
@@
 \noindent
+Due to this set of requirements, we rule out the choice of RDMS as DAS back-ends
+for several reasons. We agreed to support dynamic queries, data
+aggregation on demand and pluggable data-services. Therefore a choice 
+of concrete schema for data storage was problematic. Moreover we didn't impose
+any transactions and data persistency in DAS. This allows us to look 
+for alternative IT solutions.
+Through the analysis of available options, such as a file-based and memory caches, 
+key-value databases, documented-oriented databases we made our choice in favor 
+of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
+MongoDB \cite{MongoDB}. Both systems provide schema free document
+storage, replication and fail-over features, but our choice in favor of 
+MongoDB was obvious, due to its support of dynamic queries, 
+full indexies, including inner objects and embedded arrays,
+and auto-sharding. Our prelimiary benchmarks shown that it can sustain
+the desired load and size for our meta-data information. We used 
+document-oriented database, MongoDB, as a back-end for the three 
+main DAS components: Mapping, Analytics and DAS cache databases, 
+which will be discussed next.
+
 The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of the
 following components:
-\begin{itemize}
-\item core library with support of pluggable modules for data retrieval;
-\item caching layer to store data-service output and aggregated results;
-\item request service to handle user requests;
-\item mapping DB to keep information about data-service APIs, their
+core library with support of pluggable modules for data retrieval;
+caching layer to store data-service output and aggregated results;
+request service to handle user requests;
+mapping DB to keep information about data-service APIs, their
 notations and mapping to DAS;
-\item analytics DB with query statistics to enable pre-fetching 
+analytics DB for query analysis and identification of pre-fetch 
 strategies.
-\end{itemize}
 
 The DAS core communicates with different data-services and retrieves
-data on demand. This was done by invoking appropriate data-service APIs
-upon user requests. All user queries were written to Analytics DB
-for further analysis. The data-service output were re-mapped into
-DAS notations and, if necessary, aggregated inside of DAS cache.
-Since we didn't know a-priory what kind of data users will ask, 
-what kind of queries they will place against the system and what
-level of aggregation will be applied the choice
-of RDMS as DAS back-end was problematic\footnote{In addition DAS 
-didn't impose any requirements on transactions and persistency of the data.} 
-and we look around for other options. Through the analysis of available IT solutions, 
-such as file-based and memory caches, 
-key-value databases, documented-oriented databases we made our choice in favor 
-of the last technology. Among of them we evaluated CouchDB \cite{CouchDB} and 
-MongoDB \cite{MongoDB}. Both systems provide schema free document
-storage, replication and fail-over features, but our choice in favor of 
-MongoDB was obvious because of its support of dynamic queries, 
-full indexies, including on inner objects and embedded arrays,
-and auto-sharding. Our prelimiary benchmarks shown that it can sustain
-the desired load and size for our meta-data information. We use MongoDB as a back-end 
-for the three main components: Mapping, Analytics and DAS cache databases.
+data on demand. This was done by invoking appropriate data-service plugin
+upon the user request. Each plugin was responsible for invoking
+data-service APIs and parsing retrieved data into the DAS data format
+suitable for storage, JSON. Although the majority of data-services looks alike
+and their logic was abstracted, including parsers, there were a few
+cases where special care was required. For instance,
+we dealt with data-services which were only accessible via private network.
+Therefore, a way to propagate user credentials was required, leading
+to specific plugin implementation.
 
-DAS Mapping DB holds mapping between data-service APIs and DAS keys. 
+%All user queries were written to Analytics DB for further analysis, while
+%mapping between data services were kept in Mapping DB. The data-service 
+%output were re-mapped into DAS notations and, if necessary, aggregated 
+%inside of the DAS cache.
+
+DAS Mapping DB holded the mapping between data-service APIs and DAS keys exposed to
+the users. It created a necessary glue between data-services and identifies
+how their data can be related to each other. Basically Mapping DB was positioned 
+as a {\it registration} service with ability to translate notations among data-services.
 Upon identification of APIs which will participate in DAS activity 
-we store to Mapping DB the API name, input parameters, their type and accept 
-values as regular expression patterns as well as mapping between data-service 
-notations and DAS keys. It allows us to map user input query parameters into
-data-service API metrics as well as map back data-service output into DAS notations.
-For instance, the DBS system \cite{DBS}
-used {\it logical\_file\_name} notation, while in PhEDEx
-\cite{PhEDEx}
-the same entity was named as {\it lfn}. Since both of them were referred to
-the logical file name, we map both notations into DAS key {\it file}.\footnote{We
+we stored their names, input parameters, together with their type and accept 
+values as regular expression patterns into the Mapping DB. 
+Also we stored the mapping between data-service notations and DAS keys. 
+For instance, the CMS DBS system \cite{DBS}
+used {\it logical\_file\_name} notation to identify the name of the
+logical file, while the same entity was called {\it lfn} in CMS PhEDEx system \cite{PhEDEx}.
+Since both of them were referred to the logical file name, 
+we map both notations into DAS key {\it file}.\footnote{We
 also deal with physical file names, but still used {\it file} to refer to them in
-DAS notationis, since simple regular expressions were used to identify the
+DAS notations, since simple regular expressions were used to identify the
 meaning of the {\it file} keyword from the provided value.}
-For each service we provide a mapping from its own data-format (JSON, XML, etc.) into
-common data-format used in DAS.\footnote{We used JSON as a DAS data format.}
-We also store a mapping between DAS keys and data-service API names to
-establish appropriate calls to data-services upon provided DAS query. For example,
-when user specifies {\it file} in their DAS query, we invoke DBS 
+Such keys, e.g. {\it file}, were exposed to the end-users as pre-defined keywords to
+be used in search queries. For instance, end-users were able to type {\it file} in
+order to get all file names known to DAS via data-services as well as
+provide a concrete patterns associated with a file, e.g. {\it file=/abc*}.
+Under the hood, the appropriate data-service APIs were looked up along with appropriate
+parameters to answer those questions. For example,
+in aforementioned example, using {\it file} in DAS query cause invokation of DBS 
 {\it listFiles} and PhEDEx {\it fileReplicas} APIs, respectively.
+This approach gave power to end-users to express the meaning of their queries, especially
+when they deal with numbers and achieve accuracy of results from DAS. 
+For instance, users were able to type {\it run > 10 and run < 100}
+to specify a range of run numbers to be looked up in DAS cache. At the same time
+a full free text based keyword search is under development within MongoDB.
 
-The DAS Analytics DB plays a special role. It collects information
+The DAS Analytics DB played a special role. It collected information
 about user requests placed to the system. Each request was recorded. Upon its
-decomposition into set of selection keys and conditions we also record
-API name and API input parameters into Analytics database and 
+decomposition into set of selection keys and conditions we also recorded
+API name and input parameters into the Analytics database and 
 keep updating their counter information for repeated calls. 
@@
 the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
-and aggregate data over there. The independent set of daemons, DAS robots, are
+and aggregate data over there. The independent set of daemons, DAS robots, were
 constantly monitor Analytics DB for most popular requests and pre-fetch most
@@
 data service APIs and their input parameters. DAS
-cache was checked for a similar set of selection and condition
-keys which can be used earlier.\footnote{Since DAS doesn't hold
+cache was checked for a queries with a similar set of selection and condition
+keys used earlier.\footnote{Since DAS didn't hold
 any data by default we cannot rely on data being present in a cache.
@@
 retrieved from the cache, otherwise a data-service APIs were
-invoked and the results were collected and aggregated in DAS cache.
+invoked and the results were collected and aggregated inside of DAS cache.
 
-%Now we present overall logic of DAS workflow and details of individual components.
-%Upon request to DAS the following steps were executed: 
-%\begin{itemize}
-%\item parse input DAS query
-%\item map provided selection and condition keys into data-service APIs and
-%input parameters
-%\item look-up similar queries in DAS cache 
-%\item invoke data-service APIs
-%\item parse output results by mapping output notation into DAS keys
-%\item store and aggregate them into DAS raw-cache
-%\end{itemize}
-
-Apart from previous DBS-QL syntax we decided to fully support free text-based queries for end-users.
-All of them were transformed into MongoDB query syntax, who was used internally in
+Apart from previous DBS-QL syntax we decided to drop its syntax and allow users
+express their queries in a free text-based form. Such queries were transformed 
+into MongoDB query syntax, who was used internally in
 DAS core layer. It represents a set of selection keys and conditions as a JSON document.
@@
 in DAS, which is suitable for our 
-needs.\footnote{We've been satisfied with such performance,
-even though our measurements shown that it still can be speed up by few times
-using C++ DB driver instead of python one used in this prototype system.}
+needs.\footnote{We were satisfied with performance of the DAS prototype,
+even though it throughput can be increased by few times
+using C++ DB driver instead of python one used in this system.}
 
@@
 \begin{thebibliography}{9}
-\bibitem{FedDB}
-L. Haas, E. Lin,
-``IBM Federated Database Technology'', \\
-http://www.ibm.com/developerworks/data/library/techarticle/0203haas/0203haas.html
-
 \bibitem{DBXplorer}
@@
 http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html
+
+\bibitem{FedDB}
+L. Haas, E. Lin,
+``IBM Federated Database Technology'', \\
+http://www.ibm.com/developerworks/data/library/techarticle/0203haas/0203haas.html
 

### 2010-05-04 doc/paper.tex
Finish chapter 4.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@633 4525493e-7705-40b1-a816-d608a930855b

@@
 
+\subsection{DAS core}
 The DAS core communicates with different data-services and retrieves
 data on demand. This was done by invoking appropriate data-service plugin
-upon the user request. Each plugin was responsible for invoking
-data-service APIs and parsing retrieved data into the DAS data format
-suitable for storage, JSON. Although the majority of data-services looks alike
+upon the user request. Each plugin consists of DAS map,
+data parser(s) suitable for data-service output format(s) into DAS common
+data format (JSON) and was responsible for data-service API calls.
+The data-service DAS map contains information about data-service APIs,
+input and output parameters and their mapping to DAS keys. 
+This information was stored into DAS Mapping DB which we discuss next.
+Below you can see a typical DAS Mapping DB record for CMS
+SiteDB data-service \cite{SiteDB}\footnote{We used regular expression 
+patterns, evaluated during run time, to validate provided set of input 
+paremters and DAS keys.}:
+\begin{verbatim}
+{"api": {"params": {"name": ""}, "name": "CMSNametoSE"}, 
+ "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
+              "api_param": "name", 
+              "das_key": "site"}], 
+ "_id": "4ae1f8dfe2194e4581000014", 
+ "system": "sitedb", 
+ "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
+              "key": "site", 
+              "map": "site.name"}]}
+\end{verbatim}
+Although the majority of data-services looks alike
 and their logic was abstracted, including parsers, there were a few
@@
 
+\subsection{DAS Mapping DB}
 DAS Mapping DB holded the mapping between data-service APIs and DAS keys exposed to
@@
 order to get all file names known to DAS via data-services as well as
-provide a concrete patterns associated with a file, e.g. {\it file=/abc*}.
+provide a concrete patterns associated with a file, e.g. {\it file=abc*}.
 Under the hood, the appropriate data-service APIs were looked up along with appropriate
 parameters to answer those questions. For example,
-in aforementioned example, using {\it file} in DAS query cause invokation of DBS 
+in aforementioned example, using {\it file} in DAS query cause call of DBS 
 {\it listFiles} and PhEDEx {\it fileReplicas} APIs, respectively.
@@
 when they deal with numbers and achieve accuracy of results from DAS. 
-For instance, users were able to type {\it run > 10 and run < 100}
+For instance, users were able to type $run>10\,\,and\,\, run < 100$
 to specify a range of run numbers to be looked up in DAS cache. At the same time
@@
 
+\subsection{DAS Analytics DB}
 The DAS Analytics DB played a special role. It collected information
@@
 
+\subsection{DAS caching system}
 The DAS caching system was used to store data-service results and aggregated
@@
 \subsection{Data-services and aggregation}
-DAS provided an interface to add new CMS data-services via data-service plugins.
-Each plugin consists of its DAS map and data parser suitable for data-service output
-format(s).
-The data-service DAS map contains information about data-service APIs,
-input and output parameters, data-service
-format and their mapping to DAS keys as well as relation to other sub-systems. 
-Below you can see a typical DAS Mapping DB record for CMS
-SiteDB data-service \cite{SiteDB}\footnote{We used regular expression 
-patterns, evaluated by python during run time
-execution, to validate provided set of input paremters and DAS keys.}:
-\begin{verbatim}
-{"api": {"params": {"name": ""}, "name": "CMSNametoSE"}, 
- "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
-              "api_param": "name", 
-              "das_key": "site"}], 
- "_id": "4ae1f8dfe2194e4581000014", 
- "system": "sitedb", 
- "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
-              "key": "site", 
-              "map": "site.name"}]}
-\end{verbatim}
-This information was stored into DAS Mapping DB.
-
-The DAS Mapping DB becomes an authorative
-source of information about all data-services participating in DAS and their
-relationships. Based information stored there we were able to retrieve appropriate data
-from different data-services, re-map them into DAS notations and
-aggregate them on demand.
-For instance, the Data Bookkeeping System (DBS) and data location system (PhEDEx)
-provided information about CMS data files. In former case, DBS holds information about
-files and their relationship to other physics objects, such as run and luminosity, 
-while PhEDEx stores information about the file location, such as sites which
-holds copy of the files and details about site storage elements.
-Upon a user request to find a file, we were able to identify appropriate set of
-DBS and PhEDEx APIs, look-up data, re-format their output to DAS notations and
-store results into DAS cache.
-Once data were retreived from data-services an additional step to aggregate
-matched objects in DAS cache were applied. For instance, both data-service outputs 
-contained a file name in them. We use it as a common key to merge records in DAS cache.
-This merged (aggregated) results represent DAS records. Each DAS
+As we outlined above the DAS Mapping DB becomes an authorative
+source of information about all data-services participating in DAS.
+Based on information stored in Mapping DB we were able to 
+retrieve appropriate data from different data-services, re-map them into 
+DAS notations and aggregate them on demand. For instance, the Data 
+Bookkeeping System (DBS) and data location system (PhEDEx)
+provided information about CMS data files. In former case, DBS 
+stored information about files and their relationship to other 
+physics objects, such as run and luminosity, while PhEDEx kept
+information about the file location, e.g. details about sites 
+holding copy of the files. Upon a user request to find a file, 
+we were able to identify appropriate set of DBS and PhEDEx APIs, 
+retrieve data from them, re-format their output to DAS notations 
+and store results into DAS cache performing aggregation of matched 
+objects. For instance, since both data-services returns in their
+output a file name we used as a common key during aggregation.
+This merged results represented a set of DAS records. Each DAS
 record contains a standard DAS header, which identify look-up time,
-request url, input parameters and API method as well as response version, expiration
-time and checksum of the output. Here is an example of individual DAS record:
+request url, input parameters and API method as well as response 
+version, expiration time and checksum of the output. Usage of this 
+information was valuable contribution in debugging process of
+data-services themselves, e.g. identification of not-matched records, 
+latency studies, etc. Here is an example of DAS record representing
+aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
+and {\it CMStoSiteName}, with API results merged under the {\it site} key.
 \begin{verbatim}
@@
 \end{verbatim}
-This DAS record represents aggregated information between two SiteDB APIs, 
-{\it CMStoSAMName} and {\it CMStoSiteName}, where API results were merged 
-under the {\it site} key. The information provided by {\it das} key shows 
-DAS header, which contains information about API name, location, execution 
-time and lifetime of the data.
-Usage of this information was valuable contribution in debugging process of
-data-services themselves, e.g. identification of not-matched records, 
-latency studies, etc.
 
-We used RESTful model \cite{REST} for all DAS cache server APIs.
-It allows us to share core library and other components during development of 
-various DAS clients, e.g. CLI, web interface, robots, etc. and hides
-complexity of access to individual data-services and their security policies
-as well as proper organize caching solution.
-For instance, to access on-line Run Summary information we were forced to use
-Grid certificates. Instead of propagating user certificate to data-service
-the DAS cache server works as a proxy to this server and used host certficiate 
-to serve such requests on behalf of the users.
+%We used RESTful model \cite{REST} for all DAS cache server APIs.
+%It allows us to share core library and other components during development of 
+%various DAS clients, e.g. CLI, web interface, robots, etc. and hides
+%complexity of access to individual data-services and their security policies
+%as well as proper organize caching solution.
+%For instance, to access on-line Run Summary information we were forced to use
+%Grid certificates. Instead of propagating user certificate to data-service
+%the DAS cache server works as a proxy to this server and used host certficiate 
+%to serve such requests on behalf of the users.
 

### 2010-05-04 doc/paper.tex
Finished results chapter and summary.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@634 4525493e-7705-40b1-a816-d608a930855b

@@
 \section{Results\label{Results}}
-DAS was deployed to work with the following CMS data-services:
+In this section we discuss preliminary restults achieved with DAS system
+using the following CMS data-services:
 \begin{itemize}
-\item DBS \cite{DBS}, Data Bookkeeping System, which collect information
-about CMS meta-data;
-\item PhEDEx \cite{PhEDEx}, Physics Experiment Data Export project, which
-provides the data placement and the file transfer system for the CMS experiment;
-\item SiteDB \cite{SiteDB} which 
-records the CMS resources at the site, the resources pledged for the 
-future, and it keeps track of CMS personnel at each site, including the 
-roles and duties they fulfill in the collaboration;
-\item RunSummary DB \cite{RunSummary} which collects information about run and triggers
-conditions during data-taking;
+\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
+\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
+CMS file transfer system and data-service;
+\item SiteDB \cite{SiteDB}, the CMS site resources data-service;
+\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
 %\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
@@
 %transfer rate, etc.;
-\item Dashboard \cite{Dashboard} which monitors data collected from the 
-distributed computing systems of the CMS experiment and LHC virtual organization
-in general.
+\item Dashboard \cite{Dashboard}, the LHC virtual organization
+monitoring data-service portal;
 \end{itemize}
-Each data-service has its own scope, size and lifetime. For instance the data
-in PhEDEx are transient due to constant migration of CMS data, 
-the DBS system were divided into dozen of individual instances, 
-who has been used by different production teams for various bookkeeping tasks.
-Figure \ref{db_size} shows current size of meta-data from aforementioned data-services.
+Each data-service has its own scope, size and peculiar properties. 
+For instance, the data in PhEDEx are transient due to constant 
+migration of CMS data; the DBS system were divided into dozen 
+of individual instances with own size and lifetime, who has been used by different production 
+teams for various bookkeeping tasks; the RunSummary data-service
+operates on private network accessible at CMS online operation center.
+Figure \ref{db_size} shows the current size of meta-data from aforementioned 
+data-services.
 \begin{figure}[htb]
@@
 Current size of meta-data for CMS data-services participated in DAS, before
-official data-taking. We anticipate to collect $\sim$500GB of meta-data
-per year once LHC starts operating.
+official data-taking. 
+%We anticipate to collect $\sim$500GB of meta-data
+%per year during normal LHC operations.
 }
@@
 \end{figure}
-%Typical
-%size of individual DBS instance was around of fwe GBs, with overall DBS size of
-%around of 50 GBs.\footnote{This include table and index sizes.}
 
-At the moment we don't have exact numbers for amount of data which will stored to DAS,
-but based on test performed during 
-Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}
-exercise we produced 0.3PB of data which
-translate into $\sim50$ GB of meta-data stored in DBS system. Therefore
-we can anticipate to collect $\sim500$GB of meta-data per year during data-taking period
-at LHC.
+At the moment we don't have exact numbers for amount of data which will stored 
+in those data-service and therefor accessible in DAS. Although we project to
+accumulate of the order of 500 GB of meta-data per year during normal LHC and CMS
+operations.\footnote{This number is based on tests 
+performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
 %In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
@@
 
-Preliminary results of DAS performance are shown in Table \ref{DAS_benchmark}.
-We benchmark time required to collect and aggregate block object information
-from DBS and PhEDEx CMS data-services.
+To test DAS performance we measured elapsed time and CPU/RAM utilization
+required to collect and aggregate block object information from DBS and 
+PhEDEx CMS data-services. Preliminary results of those measurements can
+be found in Table \ref{DAS_benchmark}.
 
@@
 
-Elapsed time measured during this test consist of the following componenets:
+Please note that initial DAS implementation was done in Python \cite{Python}
+and all tests were performed on x64 Linux platform with 16GB of RAM where
+1 CPU was used by DAS server and 1 CPU by MongoDB database server. The CPU utilization
+of DAS server was 67\% in average and RAM usage was below 35MB for
+all tests. All tests were performed multiple times to avoid fluctuations. 
+The elapsed time measured during those tests consists of the 
+following componenets:
 \begin{itemize}
@@
 {\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
-is a time to convert notations used by data-service to DAS ones for every object
+is a time required to convert notations used by data-service to DAS ones for every object
 we parse, {\it cache insertion} and {\it indexing time} represents time spend to add objects into
@@
 components, {\it aggregation} and {\it output creation time}, were only applied to
-final DAS step and not to time spent in individual data-services.
+the one of the sub-system (either DBS or PhEDEx) since the first one was injected into
+fresh database, while the second required the merging step.
 
@@
 
-The individual DAS components (DBS and PhEDEx) shown roughly the same performance.
-Although the time of PhEDEx component was divided among: retrieval + parsing 
-time $\sim$110 sec and merging step $\sim$50 sec.
+The individual tests of DAS components (DBS and PhEDEx) shown roughly 
+the same performance. Although the time of PhEDEx component was bigger 
+due to longer retrieval and parsing time.
+The time of PhEDEx component shown in Table \ref{DAS_benchmark} divided 
+among: retrieval + parsing time $\sim$110 sec and merging step $\sim$50 sec.
 
 As you can see from Table \ref{DAS_benchmark},
-the time spent to read DAS records (last column) from the cache was quite
+the time spent to access individual components (last column) was quite
 reasonable, roughly 1 second for both systems. While final time to
 get DAS records on disk was about 45 seconds. This was mostly due to I/O and
-conversion between binary data format used in database (BSON) to JSON.
-All operations were performed on a single CPU core and obviously can be done
-in parallel.
+conversion between binary data format used by MongoDB, BSON, to DAS
+data format, JSON.
 
@@
 for insertion, 3400 docs/sec for aggregation and 8500 docs/sec for reading operation
-in DAS, which is suitable for our 
-needs.\footnote{We were satisfied with performance of the DAS prototype,
-even though it throughput can be increased by few times
-using C++ DB driver instead of python one used in this system.}
+in DAS, which is suitable for our needs. Since all operations were performed 
+on a single CPU core a further improvements are expected. Also, a necesarry
+tuning can be done by using native C++ MongoDB driver instead of python one.
 
 \section{Summary}
-We presented new data aggregation service (DAS) developed for CMS High-Energy experiment
+We presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
 at LHC, CERN, Geneva, Switzerland. It was designed to provide caching and
 aggregation layer on top of the existing relational and non-relation data-services
-mostly in real time fashion. All the data were retrieved on demand basis,
+mostly in real time fashion. All of the data were retrieved on demand basis,
 while data pre-fetching of most common queries is in our plans. We developed
@@
 and performed performance studies shown in this paper. 
-The CMS experiment is started to
-collect data from November 2009 and we expect to write about PB of data each
-year to tape. In addition, the Monte-Carlo samples will be produced at this scale.
+The CMS experiment is started to collect its data in December of 2009 and 
+we expect to accumulate a few PB of real data each year to tape. 
+In addition, the Monte-Carlo samples will be produced at this scale.
 Based on test studies performed in CMS, we expect that total size of
@@
 
-This work was supported by the National Science Foundation and Department of Energy of the United States of America. Fermilab is operated by Fermi Research Alliance, LLC under Contract
+This work was supported by the National Science Foundation and 
+Department of Energy of the United States of America. 
+Fermilab is operated by Fermi Research Alliance, LLC under Contract
 No. DE-AC02-07CH11359 with the United States Department of Energy.
@@
 
+\bibitem{Python} Python programming language, http://www.python.org
+
 \bibitem{DBSearch}

### 2010-05-04 src/python/DAS/tools/das_cli.py
Added printout of loose query
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@635 4525493e-7705-40b1-a816-d608a930855b

@@ -4,13 +4,14 @@
 """
 DAS command line interface
 """
-__revision__ = "$Id: das_cli.py,v 1.21 2009/11/24 16:01:10 valya Exp $"
-__version__ = "$Revision: 1.21 $"
+__revision__ = "$Id: das_cli.py,v 1.22 2009/12/01 15:12:50 valya Exp $"
+__version__ = "$Revision: 1.22 $"
 __author__ = "Valentin Kuznetsov"
 
 import time
 from optparse import OptionParser
 from DAS.core.das_core import DASCore
+from DAS.core.das_mongocache import convert2pattern
 from DAS.utils.utils import dump, genkey
 
 import sys
@@ -109,8 +110,10 @@ if __name__ == '__main__':
     DAS = DASCore(debug=debug)
     if  opts.hash:
         mongo_query = DAS.mongoparser.requestquery(query)
+        loose_query_pat, loose_query = convert2pattern(mongo_query)
         print "DAS-QL query:", query
         print "Mongo query :", mongo_query
+        print "Loose query :", loose_query
         print "query hash  :", genkey(str(mongo_query))
         sys.exit(0)
     sdict = DAS.keys()

### 2010-05-04 src/python/DAS/tools/das_mapping_db.py
Changed patterns for site db input parameters, T[0-3]_ is too wide, it should be T[0-3] since people can provide T1* patterns
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@636 4525493e-7705-40b1-a816-d608a930855b

@@ -4,8 +4,8 @@
 """
 DAS command line interface
 """
-__revision__ = "$Id: das_mapping_db.py,v 1.15 2009/11/19 20:15:35 valya Exp $"
-__version__ = "$Revision: 1.15 $"
+__revision__ = "$Id: das_mapping_db.py,v 1.16 2009/12/01 15:13:50 valya Exp $"
+__version__ = "$Revision: 1.16 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -365,9 +365,9 @@ if __name__ == '__main__':
     params = {'name':''}
     rec = {'system' : system, 
         'api' : dict(name=api, params=params),
-        'daskeys' : [dict(key='site', map='site.name', pattern="re.compile('^T[0-3]_')")],
+        'daskeys' : [dict(key='site', map='site.name', pattern="re.compile('^T[0-3]')")],
         'api2das' : [
-                dict(api_param='name', das_key='site', pattern="re.compile('^T[0-3]_')"),
+                dict(api_param='name', das_key='site', pattern="re.compile('^T[0-3]')"),
         ]
     }
     mgr.add(rec)
@@ -376,9 +376,9 @@ if __name__ == '__main__':
     params = {'name':''}
     rec = {'system' : system, 
         'api' : dict(name=api, params=params),
-        'daskeys' : [dict(key='site', map='site.name', pattern="re.compile('^T[0-3]_')")],
+        'daskeys' : [dict(key='site', map='site.name', pattern="re.compile('^T[0-3]')")],
         'api2das' : [
-                dict(api_param='name', das_key='site', pattern="re.compile('^T[0-3]_')"),
+                dict(api_param='name', das_key='site', pattern="re.compile('^T[0-3]')"),
         ]
     }
     mgr.add(rec)
@@ -387,9 +387,9 @@ if __name__ == '__main__':
     params = {'name':'required'}
     rec = {'system' : system, 
         'api' : dict(name=api, params=params),
-        'daskeys' : [dict(key='site', map='site.name', pattern="re.compile('^T[0-3]_')")],
+        'daskeys' : [dict(key='site', map='site.name', pattern="re.compile('^T[0-3]')")],
         'api2das' : [
-                dict(api_param='name', das_key='site', pattern="re.compile('^T[0-3]_')"),
+                dict(api_param='name', das_key='site', pattern="re.compile('^T[0-3]')"),
         ]
     }
     mgr.add(rec)
[…]

### 2010-05-04 src/python/DAS/services/abstract_service.py
use newline for better visibility of output on stdout
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@642 4525493e-7705-40b1-a816-d608a930855b

@@ -4,8 +4,8 @@
 """
 Abstract interface for DAS service
 """
-__revision__ = "$Id: abstract_service.py,v 1.53 2009/12/02 14:40:15 valya Exp $"
-__version__ = "$Revision: 1.53 $"
+__revision__ = "$Id: abstract_service.py,v 1.54 2009/12/02 15:44:56 valya Exp $"
+__version__ = "$Revision: 1.54 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -203,11 +203,11 @@ class DASAbstractService(object):
         Filter provided apicall wrt existing apicall records in Analytics DB.
         """
         spec = {'apicall.url':url, 'apicall.api':api}
-        msg  = 'DBSAbstractService::pass_apicall, %s, API=%s, args=%s,'\
+        msg  = 'DBSAbstractService::pass_apicall, %s, API=%s, args=%s'\
         % (self.name, api, api_params)
         for row in self.analytics.col.find(spec):
             if  compare_specs(api_params, row['apicall']['api_params']):
-                msg += ' will re-use existing api call with args=%s'\
+                msg += '\nwill re-use existing api call with args=%s'\
                 % row['apicall']['api_params']
                 self.logger.info(msg)
                 return True

### 2010-05-04 src/python/DAS/core/das_mapping_db.py
Added init_notationcache method to allow dynamically load notations if analytics db is empty (as in unit tests).
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@643 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 DAS mapping DB
 """
 
-__revision__ = "$Id: das_mapping_db.py,v 1.18 2009/11/25 18:19:17 valya Exp $"
-__version__ = "$Revision: 1.18 $"
+__revision__ = "$Id: das_mapping_db.py,v 1.19 2009/12/02 21:01:08 valya Exp $"
+__version__ = "$Revision: 1.19 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -40,6 +40,12 @@ class DASMapping(object):
         self.create_db()
 
         self.notationcache = {}
+        self.init_notationcache()
+
+    def init_notationcache(self):
+        """
+        Initialize notation cache by reading notations.
+        """
         for system, notations in self.notations().items():
             for row in notations:
                 key = system, row['api_param']
@@ -251,7 +257,8 @@ class DASMapping(object):
         Return DAS notation map.
         """
         notationmap = {}
-        spec = {'notations':{'$ne':None}}
+#        spec = {'notations':{'$ne':None}}
+        spec = {'notations':{'$exists':True}}
         if  system:
             spec['system'] = system
         for item in self.col.find(spec):
@@ -264,6 +271,8 @@ class DASMapping(object):
         run_number=run. In case when api_param is not presented in DB
         just return it back.
         """
+        if  not self.notationcache:
+            self.init_notationcache()
         if  self.notationcache.has_key((system, api_param)):
             _api, das_name = self.notationcache[(system, api_param)]
             if  _api:

### 2010-05-04 test/das_mapping_t.py
Adjust all unit tests to pass
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@644 4525493e-7705-40b1-a816-d608a930855b

@@ -114,8 +114,8 @@ class testDASMapping(unittest.TestCase):
         # adding notations
         notations = {'system':system, 
             'notations':[
-                    {'api_param':'storage_element_name', 'das_name':'se'},
-                    {'api_param':'number_of_events', 'das_name':'nevents'},
+                    {'api_param':'storage_element_name', 'das_name':'se', 'api':''},
+                    {'api_param':'number_of_events', 'das_name':'nevents', 'api':''},
                         ]
         }
         self.mgr.add(notations)

### 2010-05-04 test/das_viewmanager_t.py

@@ -26,6 +26,7 @@ class testDAS(unittest.TestCase):
         logger   = DASLogger(verbose=debug, stdout=debug)
         config['logger']  = logger
         config['verbose'] = debug
+        config['views_dir'] = os.getcwd()
         config['views_engine'] = 'sqlite:///%s' % self.db
         try: 
             del config['sum_views']

### 2010-05-04 test/qlparser_t.py

@@ -67,19 +67,20 @@ class testQLParser(unittest.TestCase):
         Test requestquery function.
         """
         query = 'block site=T1_CH_CERN'
-        expect = {'fields': ['block'], 
-                  'spec': {'site.name': 'T1_CH_CERN'}}
+        expect = {'fields': None, 
+                  'spec': {'site.name': 'T1_CH_CERN', 'block.name':'*'}}
         result = self.parser.requestquery(query)
         self.assertEqual(expect, result)
 
         query = 'block site=a.b.c'
-        expect = {'fields': ['block'], 
-                  'spec': {'site.se': 'a.b.c'}}
+        expect = {'fields': None, 
+                  'spec': {'site.se': 'a.b.c', 'block.name':'*'}}
         result = self.parser.requestquery(query)
         self.assertEqual(expect, result)
 
         query = 'file,site block=bla'
-        expect = {'fields': ['file', 'site'], 'spec': {'block.name': 'bla'}}
+        expect = {'fields': None, 'spec': {'block.name': 'bla', 
+                        'file.name':'*', 'site.name':'*', 'site.se':'*'}}
         result = self.parser.requestquery(query)
         self.assertEqual(expect, result)
 

### 2010-05-04 test/utils_t.py

@@ -21,29 +21,29 @@ class testUtils(unittest.TestCase):
         """Test merge_dict"""
         dict1  = {'block':{'name':'AAA', 'b':{'c':1}, 'size':2}, 'das':{'system':'dbs'}}
         dict2  = {'block':{'name':'AAA', 'x':{'y':1}, 'z':1, 'size':2}, 'das':{'system':'phedex'}}
-        result = merge_dict(dict1, dict2)
+        merge_dict(dict1, dict2)
         expect = {'block': [{'b': {'c': 1}, 'name': 'AAA', 'size': 2}, 
         {'x': {'y': 1}, 'z': 1, 'name': 'AAA', 'size': 2}], 
         'das': [{'system': 'dbs'}, {'system': 'phedex'}]}
-        self.assertEqual(expect, result)
+        self.assertEqual(expect, dict1)
 
         dict1  = {'test':[1,2]}
         dict2  = {'test':3}
         expect = {'test':[1,2,3]}
-        result = merge_dict(dict1, dict2)
-        self.assertEqual(expect, result)
+        merge_dict(dict1, dict2)
+        self.assertEqual(expect, dict1)
 
         dict1  = {'test':[1,2]}
         dict2  = {'test':[3,4]}
         expect = {'test':[1,2,3,4]}
-        result = merge_dict(dict1, dict2)
-        self.assertEqual(expect, result)
+        merge_dict(dict1, dict2)
+        self.assertEqual(expect, dict1)
 
         dict1  = {'test':1}
         dict2  = {'test':[2,3]}
         expect = {'test':[1,2,3]}
-        result = merge_dict(dict1, dict2)
-        self.assertEqual(expect, result)
+        merge_dict(dict1, dict2)
+        self.assertEqual(expect, dict1)
 
     def test_dict_value(self):
         """Test dict_value"""
@@ -63,9 +63,9 @@ class testUtils(unittest.TestCase):
         
         dict1 = {'a' : [{'b':1, 'c':1}, {'b':2, 'c':2}]}
         dict2 = {'a' : {'b':1, 'e':1}}
-        result = merge_dict(dict1, dict2)
+        merge_dict(dict1, dict2)
         expect = {'a': [{'c': 1, 'b': 1}, {'c': 2, 'b': 2}, {'b': 1, 'e': 1}]}
-        self.assertEqual(expect, result)
+        self.assertEqual(expect, dict1)
         
     def test_adjust_value(self):
         """Test adjust_value"""

### 2010-05-04 doc/paper.tex
Next update: added intro to DAS section, new refs, etc.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@645 4525493e-7705-40b1-a816-d608a930855b

@@
 
-%In a large modern enterprise, information is almost inevitably distributed among several database management systems.
-Despite considerable attention from the research community, relatively few commercial systems have attempted to address this issue.
-This article describes the technology that enables clients of IBM's federated database engine to access and integrate the data and specialized computational capabilities of a wide range of relational and non­relational data sources.
-
 \begin{abstract}
-%A meta-data plays significant role in a large modern enterprises, research experiments,
-%digital libraries where it comes from different sources and distributed in a 
-%variety of forms and digital formats. It is organized and managed by constantly
-%evolving software using both relational and non-relational data sources. There is
-%a big demand to enable information discovery from multiple sources.
-%Here we discuss a new data aggregation system which consume and deliver information 
-%from different relational and non-relational data sources on a concrete example 
-%of large scale, distributed system of CMS physics experiment at LHC.
-
-Meta-data plays signiﬁcant role in a large modern enterprises, 
-research experiments and digital libraries where it comes from different 
+Meta-data plays signiﬁcant role in large modern enterprises, 
+research experiments and digital libraries where it comes from a different 
 sources and distributed in a variety of forms and digital formats. 
@@
 pre-established set of data-services.
-%Even though we can access it via set of data-services, twikies, blogs, etc.,
-%the information discovery across them is still a manual task.
-% which prevent
-%information discovery
-%there is no coherent way for information discovery from different sources
-%in heterogeneous, distributed environment with different security policies enforced.
 
@@
 the LHC where we used a large scale, distributed computing system 
-to motivate the work.
+to motivate this work.
 
@@
 At this scale the ``information discovery'' within heterogeneous, distributed 
-environment become a key ingredient of successful data analysis in HEP.
-%scientists are facing with a new set of problems in an area of
-%``information discovery'' within heterogeneous, distributed environment.
+environment becomes a key ingredient of successful data analysis in HEP.
 The data and associated meta-data are produced in variety of forms and digital formats.
 They are stored and retrieved from relational and non-releational data-sources, such as 
-RDMS systems, document oriented databases, blogs, twikies, file systems and
+RDMS systems, document oriented databases, blogs, twikies, file systems,
 customized applications, etc. 
-%Its access is restricted by broad variety of security 
-%policies enforced by individual sites and community in general.
-Working in such environment requires a lot of expertise and our users
-(in this case physicists) are always looking for simple, intuitive and flexible
+Working in such environment requires a lot of expertise and our users, physicists, 
+are always looking for simple, intuitive and flexible
 tool(s) to look-up their desired data. A well-known solutions, such as data-services
-and search engines tighted to a specific data sources and end-users are left 
+and search engines are tighted to a specific data sources and end-users are left 
 with a manual task to bookkeep and relatate information from them.
@@
 i.e. retrieval of only documents that contain all query keywords, has
-been discussed. It requires that all keywords are matched in a row from
-single table or joined tables. While \cite{QueryAnswer} make a step forward
+been discussed. It requires that all keywords are matched in a row from a
+single table or joined tables. Authors of \cite{QueryAnswer} made a step forward
 by considering information around matched values and presents customized answers to
-end-users over the generated new database. Those and other works\footnote{
-For more information about
-keyword based search over relational databases we redirect users to reference in
-\cite{DBXplorer, QueryAnswer}.} 
+end-users over the generated new database. These and other related 
+works\footnote{For more information about keyword based search over relational 
+databases we redirect users to references in \cite{DBXplorer, QueryAnswer}.} 
 are based on some kind of the graph over the underlying database and
@@
 The proximity of the results in those techniques 
-are mostly confined to the text-based values stored in a database, since the
-keyword based search with numerical values 
-need the precise knowledge of the related entity in a schema. For example, typing
-100 can lead to inappropriate matches, such as row ids, and in order to be
-answered correctly it requires additional keyword, clarifying its context, 
+are mostly confined to the text-based values stored in a database, since 
+usage of numerical values in the input search field is mostly useless without 
+knowledge of the context of the value. For example, typing a value 100 
+in the search input field leads to plenty of inappropriate matches, 
+such as row ids, and in order to be
+answered correctly requires additional keyword, clarifying its meaning, 
 which should be matched with particular
 entity (table.column) in the underlying database.
-Therefore such approaches cannot be used to answer the logical questions, such as  
+Therefore to answer presicely the logical questions, such as  
 {\it find me total number of files whose size more then 20 but less then 
-100 taken from january to february of 2009} which can be easily accomplished
+100} would be hard or almost impossible using keyword search techniques discussed 
+in \cite{DBXplorer, QueryAnswer}, while it can be easily accomplished
 using SQL syntax.
@@
 To address those limitations an attempt to build a simple, intuitive and 
-flexible query language was introduced in \cite{DBS-QL}\footnote{A similar but less
-sophisticated approach was done in \cite{AMI}}.
+flexible query language was introduced in \cite{DBS-QL}.
 It represents a power of SQL while hiding underlying relational schema from 
@@
 This solution represents a common use cases within HEP community and provide
-intuitive mapping between mental model of end-users and underlying database back-end.
+intuitive mapping between mental model of end-users and underlying database schema.
 
@@
 federated databases, see for example \cite{FedDB}. It unifies data coming 
-from different RDMS systems into federated DB where SQL queries can be placed to 
-search desired data. Even though it would be interesting to apply technique 
+from different RDMS systems into the federated DB where SQL queries can be placed to 
+search for desired data. Even though it would be interesting to apply technique 
 discussed in \cite{DBS-QL} to a federated database it still requires 
 an access to the underlying database schema. This obviously creates a problem
-working in distributed heterogeneous environment where different security policies
+in distributed heterogeneous environment where different security policies
 are in place.
@@
 Keeping in mind that meta-data can come in different forms and data formats
-we used a document-oriented database back-end with full support of query language.
+we used a document-oriented database back-end with full support of the query language.
 The key ingredients were the setup of the mapping and analytics services which provide
 necessary glue across existing API/notations and user queries. This approach 
-allows us to perform the data aggregation across various data-services, achieve
+allows us to perform the dynamic data aggregation across various data-services, achieve
 precise answers using keyword-based queries
@@
 the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
-It is designed to explore the frontier of High Energy Physics and provide physicists
-ability to look at conditions presented at early stage of our Universe.
+It is designed to explore the frontier of the High Energy Physics and provide physicists
+ability to look at conditions presented at the early stage of our Universe.
 More then 3000 physicists from 183 institutions reprepresenting almost 
@@
 \section{Data Aggregation System\label{DAS}}
+The goal of the Data Aggregation System was to combine existing data-services
+under one umrella, utilizing their access rules, APIs, security policies, etc., and
+allowing users to perform queries against them. In contrast, approaches
+discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
+enforced a publication step of data into a dedicated database 
+or central repositories \cite{iRODS}. Quite often it is required
+that data should be converted into uniform data formats in order enable
+aggregation, see \cite{OpenArchive}. In DAS architecture the data are provided
+by existing data-services and infrastructure, whose design, implementation and
+security policies are left up to their developers. Therefore DAS does not requires
+a data preservation and transaction capabilities. Instead DAS can be treated
+as proxy or caching layer where information retrieval (data) are triggered dynamically
+by user queries. Such infrastructure allows data sharing across different queries and
+data aggregation using appropriate mapping. Below we discuss details of
+DAS architecture.
+
+\subsection{DAS architecture}
 The design of the DAS system was based on previous studies of CMS Data 
@@
 we took away relations and details on underlying schema from end-users. 
-This was achieved by usaging Dijkstras shortest path algorithm 
+This was achieved by using Dijkstras shortest path algorithm 
 to establish necessary table joins based on provided information in user queries and
@@
 \noindent
-Due to this set of requirements, we rule out the choice of RDMS as DAS back-ends
+Due to this set of requirements, we rule out the choice of RDMS as DAS back-end(s)
 for several reasons. We agreed to support dynamic queries, data
@@
 
+\subsection{DAS queries}
 Apart from previous DBS-QL syntax we decided to drop its syntax and allow users
-express their queries in a free text-based form. Such queries were transformed 
-into MongoDB query syntax, who was used internally in
-DAS core layer. It represents a set of selection keys and conditions as a JSON document.
-For example, a query
+express their queries in a free text-based form. Although we must clarify its
+meaning. We used a pre-defined set of keys known to DAS to describe common
+entities of our data. For example, {\it dataset, file, block, run, event}. Those
+names were used by our users in their daily work operations. So, by typing
+{\it file} we didn't perform a string matching to ``file'' string, but rather
+look-up all file objects in a DAS. Moreover each DAS key has a default attribute.
+For intance, for {\it dataset, block, file} the default attribute 
+was a {\it name}, while for {\it run} and {\it event} the default attribute was 
+a {\it number}. This allows to specify key-value pair as simple as possible, e.g.
+{\it run=123456}, which means run number 123456. At the same time a queyr 
+key, e.g. {\it site} meant to look-up all site objects. Such ability
+to use key-value pairs resolve ambiguity in object values. For instance,
+specifying a run number as {\it run=123456} instruct system to resolve
+data object with run equal to 123456 rather then other objects which may have
+123456 values, e.g. event number or file size. At the same time, to complement
+the stregth of the DAS we do plan to index data objects and add a keyword-based 
+search later.
+
+The free text-based queries were transformed into MongoDB query syntax, 
+who was used internally by DAS. It represents a set of selection 
+keys and conditions as a JSON document. For example, a query
 {\it dataset site=T1\_CH\_CERN}
@@
 \bibitem{DBXplorer}
-Sanjay Agrawal, Surajit Chaudhuri, Gautam Das: DBXplorer: A System for
-Keyword-Based Search over Relational Databases. ICDE 2002: 5-16
+S. Agrawal, S. Chaudhuri, G. Das,
+``DBXplorer: A System for Keyword-Based Search over Relational Databases'',
+ICDE 2002, pp. 5-16.
 
 \bibitem{QueryAnswer}
-Georgia Koutrika, Alkis Simitsis, Yannis E. Ioannidis: Pr\'{e}cis: The Essence of
-a Query Answer. ICDE 2006: 69-78
+G. Koutrika, A. Simitsis, Y. E. Ioannidis,
+``Pr\'{e}cis: The Essence of a Query Answer'',
+ICDE 2006, pp. 69-78.
 
-\bibitem{DBS-QL} V. Kuznetsov, D. Riley, A. Afaq, V. Sekhri, Y. Guo, L. Lueking,
-``The CMS DBS Query Language'', CHEP 2009
+\bibitem{DBS-QL} 
+V. Kuznetsov, D. Riley, A. Afaq, V. Sekhri, Y. Guo, L. Lueking,
+``The CMS DBS Query Language'', CHEP 2009.
 
-\bibitem{AMI}
-Altas AMI web portal:
-http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html
+%\bibitem{AMI}
+%Altas AMI web portal,
+%http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html
 
@@
 
-\bibitem{CMS} CMS Collaboration R. Adolphi et al., The CMS experiment at the CERN LHC, JINST, 0803, S08004 (2008)
+\bibitem{CMS} 
+R. Adolphi et al., 
+``The CMS experiment at the CERN LHC'',
+JINST 0803, S08004 (2008).
 
 \bibitem{CMSDataModel} 
-.Grandi, D.Stickland, L.Taylor et al., The CMS Computing Model, CERN-LHCC-2004- 035/G-083 (2004);
+C. Grandi, D.Stickland, L.Taylor et al.,
+``The CMS Computing Model'',
+CERN-LHCC-2004-035/G-083 (2004);
 A. Fanfani et. al., 
-``Distributed Analysis in CMS'', to be published in Journal of Grid Computing.
+``Distributed Analysis in CMS'', 
+to be published in Journal of Grid Computing.
 
-\bibitem{DBS} A. Afaq, et. al. ``The CMS Dataset Bookkeeping Service'', CHEP 2007 
-J.Phys.Conf.Ser, 119, 072001 (2008)
+\bibitem{iRODS}
+The Integrated Rule-Oriented Data System,
+http://www.irods.org/
 
-\bibitem{DBS07} A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
-``A multi-dimensional view on information retrieval of CMS data'', CHEP 2007
+\bibitem{OpenArchive}
+Open Archives Initiative,
+http://www.openarchives.org/
+
+\bibitem{DBS} 
+A. Afaq, et. al.,
+``The CMS Dataset Bookkeeping Service'', 
+J. Phys. Conf. Ser, 119, 072001 (2008).
+
+\bibitem{DBS07} 
+A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
+``A multi-dimensional view on information retrieval of CMS data'',
+Journal of Physics: Conference Series 119, 072013, (2008).
 
@@
 \bibitem{Arms}
-C. R. Arms, W. Y. Arms, ``Mixed Content and Mixed Metadata 
+C. R. Arms, W. Y. Arms,
+ ``Mixed Content and Mixed Metadata 
 Information Discovery in a Messy World'',
-chapter from ``Metadata in Practice'', ALA Editions, 2004
+chapter from ``Metadata in Practice'', ALA Editions, 2004.
 
@@
 
-\bibitem{REST}
-Fielding, Roy Thomas ``Architectural Styles and the Design of 
-Network-based Software Architectures'', Doctoral dissertation, 2000,
-University of California, Irvine
+%\bibitem{REST}
+%Fielding, Roy Thomas ``Architectural Styles and the Design of 
+%Network-based Software Architectures'', Doctoral dissertation, 2000,
+%University of California, Irvine
 
@@
 Rehn et. al.,
-``PhEDEx high-throughput data transfer management system'', CHEP06, Mumbai, India.
-R. Egeland et al., Data transfer infrastructure for CMS data taking, Proceedings of Science,
-PoS(ACAT08)033 (2008) ;
-L. Tuura et al., Scaling CMS data transfer system for LHC start-up, J.Phys.Conf.Ser, 119, 072030 (2008)
+``PhEDEx high-throughput data transfer management system'', CHEP 2006;
+R. Egeland et al., 
+``Data transfer infrastructure for CMS data taking'', 
+Proceedings of Science, PoS (ACAT08) 033 (2008);
+L. Tuura et al., 
+``Scaling CMS data transfer system for LHC start-up'', 
+J. Phys. Conf. Ser, 119, 072030 (2008).
 
@@
 
-\bibitem{DBSearch}
-D. Konopnicki, O. Shmueli,
-``Database-Inspired Search'', 
-Proc. of the 31st VLDB Conference, 2005.
+%\bibitem{DBSearch}
+%D. Konopnicki, O. Shmueli,
+%``Database-Inspired Search'', 
+%Proc. of the 31st VLDB Conference, 2005.
 \end{thebibliography}

### 2010-05-04 doc/paper.tex
Revisit DAS caching system subsection. Added description of the algorithm
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@646 4525493e-7705-40b1-a816-d608a930855b

@@
 \centering
-\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
+\includegraphics[width=120mm]{DAS_Cache_and_Analytics.pdf}
 \caption{
@@
  "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
-              "api_param": "name", 
-              "das_key": "site"}], 
+              "api_param": "name", "das_key": "site"}], 
  "_id": "4ae1f8dfe2194e4581000014", 
@@
  "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
-              "key": "site", 
-              "map": "site.name"}]}
+              "key": "site", "map": "site.name"}]}
 \end{verbatim}
@@
 \subsection{DAS caching system}
-The DAS caching system was used to store data-service results and aggregated
-information in form of DAS records. As it shown in figure \ref{DAS_cache} 
-the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
-and aggregate data over there. The independent set of daemons, DAS robots, were
-constantly monitor Analytics DB for most popular requests and pre-fetch most
+The DAS caching system was used to dynamically fetch
+and aggregate data upon user requests. We used the following 
+algorithm:
+\begin{description}
+\item[  1.] Decompose query into set of conditions ($conds$) 
+and selection keys ($skeys$)
+\item[  2.] Check cache for query hash, if found: look-up results from cache
+\item[  3.] Identify list of APIs which can serve requested data
+\begin{verbatim}
+apilist = [r(url, api, args) for r in mapping_db(skeys, conds)]
+\end{verbatim}
+\item[  4.] Iterate over APIs
+\begin{verbatim}
+for url, api, args in apilist:
+    if  superset_in_analytics_found(api, args):
+        continue
+    for row in parser(get_data(url, api, args)):
+        yield row
+\end{verbatim}
+\item[  5.] Perform aggrgation by iterating over the previos step
+\begin{verbatim}
+if cache_has(row.match_key):
+   existing_record = get_from_cache(row.match_key)
+   merge(row, existing_record)
+   insert_into_cache(row)
+   delete_from_cache(existing_record)
+else:
+   insert_into_cache(row)
+\end{verbatim}
+\item[  6.] Get results from the cache
+\end{description}
+The data was stored in a form of DAS records into the cache, while
+queries were recorded into Analytics DB for further analysis. We
+also had an option to run independent set of daemons, DAS robots, who
+monitor Analytics DB for most popular requests and pre-fetch the most
 common data into DAS cache. 
+%The interesting part was being able
+%to identify the superset of api/args out of found one in an Analytics DB.
 
-Each user request placed to the system was parsed and set of selection
-and condition keys were identified. They were mapped into set of
-data service APIs and their input parameters. DAS
-cache was checked for a queries with a similar set of selection and condition
-keys used earlier.\footnote{Since DAS didn't hold
-any data by default we cannot rely on data being present in a cache.
-Moreover if we look-up data first, we cannot tell if it was complete
-set or not, since earlier queries may retrieve only portion of the requested data.}
-If the superset was found a data were
-retrieved from the cache, otherwise a data-service APIs were
-invoked and the results were collected and aggregated inside of DAS cache.
+%The DAS caching system was used to store data-service results and aggregated
+%information in form of DAS records. As it shown in figure \ref{DAS_cache} 
+%the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
+%and aggregate data over there. The independent set of daemons, DAS robots, were
+%constantly monitor Analytics DB for most popular requests and pre-fetch most
+%common data into DAS cache.
+
+%Each user request placed to the system was parsed and set of selection
+%and condition keys were identified. They were mapped into set of
+%data service APIs and their input parameters. DAS
+%cache was checked for a queries with a similar set of selection and condition
+%keys used earlier.\footnote{Since DAS didn't hold
+%any data by default we cannot rely on data being present in a cache.
+%Moreover if we look-up data first, we cannot tell if it was complete
+%set or not, since earlier queries may retrieve only portion of the requested data.}
+%If the superset was found a data were
+%retrieved from the cache, otherwise a data-service APIs were
+%invoked and the results were collected and aggregated inside of DAS cache.
 
@@
 a {\it number}. This allows to specify key-value pair as simple as possible, e.g.
-{\it run=123456}, which means run number 123456. At the same time a queyr 
+{\it run=123456}, which means run number 123456. At the same time a query 
 key, e.g. {\it site} meant to look-up all site objects. Such ability
@@
 operates on private network accessible at CMS online operation center.
-Figure \ref{db_size} shows the current size of meta-data from aforementioned 
+Table \ref{db_size} shows the current size of meta-data from aforementioned 
 data-services.
-\begin{figure}[htb]
+%\begin{figure}[htb]
+%\centering
+%\includegraphics[width=100mm]{db_size.pdf}
+%\caption{
+%Current size of meta-data for CMS data-services participated in DAS, before
+%official data-taking. 
+%}
+%\label{db_size}
+%\end{figure}
+
+\begin{table*}[hbt]
 \centering
-\includegraphics[width=100mm]{db_size.pdf}
+\begin{tabular}{llllll}\hline
+\hline
+
+Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
+\hline
+tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
+indexies (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
+\hline
+\hline
+\end{tabular}
 \caption{
@@
 official data-taking. 
-%We anticipate to collect $\sim$500GB of meta-data
-%per year during normal LHC operations.
 }
 \label{db_size}
-\end{figure}
+\end{table*}
 

### 2010-05-04 doc/paper.tex
Further tune-ups
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@648 4525493e-7705-40b1-a816-d608a930855b

@@
-\documentclass[a4paper]{jpconf}
-\usepackage{graphicx}
+% Template for Elsevier CRC journal article
+% version 1.0 dated 13 October 2009
+
+% This file (c) 2009 Elsevier Ltd.  Modifications may be freely made,
+% provided the edited file is saved under a different name
+
+% This file contains modifications for Procedia Computer Science
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% This template uses the elsarticle.cls document class and the extension package ecrc.sty
+%% For full documentation on usage of elsarticle.cls, consult the documentation "elsdoc.pdf"
+%% Further resources available at http://www.elsevier.com/latex
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% The '1p' and 'times' class options of elsarticle are used for Elsevier CRC
+\documentclass[1p,times]{elsarticle}
+
+%% The `ecrc' package must be called to make the CRC functionality available
+\usepackage{ecrc}
+
+%% The ecrc package defines commands needed for running heads and logos.
+%% For running heads, you can set the journal name, the volume, the starting page and the authors
+
+%% set the volume if you know. Otherwise `00'
+\volume{00}
+
+%% set the starting page if not 1
+\firstpage{1}
+
+%% Give the name of the journal
+\journalname{Procedia Computer Science}
+
+%% Give the author list to appear in the running head
+%% Example \runauth{C.V. Radhakrishnan et al.}
+\runauth{}
+
+%% The choice of journal logo is determined by the \jid and \jnltitlelogo commands.
+%% A user-supplied logo with the name <\jid>logo.pdf will be inserted if present.
+%% e.g. if \jid{yspmi} the system will look for a file yspmilogo.pdf
+%% Otherwise the content of \jnltitlelogo will be set between horizontal lines as a default logo
+
+%% Give the abbreviation of the Journal.
+\jid{procs}
+
+%% Give a short journal name for the dummy logo (if needed)
+\jnltitlelogo{Procedia Computer Science}
+
+%% Hereafter the template follows `elsarticle'.
+%% For more details see the existing template files elsarticle-template-harv.tex and elsarticle-template-num.tex.
+
+%% Elsevier CRC generally uses a numbered reference style
+%% For this, the conventions of elsarticle-template-num.tex should be followed (included below)
+%% If using BibTeX, use the style file elsarticle-num.bst
+
+%% End of ecrc-specific commands
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% The amssymb package provides various useful mathematical symbols
+\usepackage{amssymb}
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
+
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers after \end{frontmatter}.
+%% \usepackage{lineno}
+
+%% natbib.sty is loaded by default. However, natbib options can be
+%% provided with \biboptions{...} command. Following options are
+%% valid:
+
+%%   round  -  round parentheses are used (default)
+%%   square -  square brackets are used   [option]
+%%   curly  -  curly braces are used      {option}
+%%   angle  -  angle brackets are used    <option>
+%%   semicolon  -  multiple citations separated by semi-colon
+%%   colon  - same as semicolon, an earlier confusion
+%%   comma  -  separated by comma
+%%   numbers-  selects numerical citations
+%%   super  -  numerical citations as superscripts
+%%   sort   -  sorts multiple citations according to order in ref. list
+%%   sort&compress   -  like sort, but also compresses numerical citations
+%%   compress - compresses without sorting
+%%
+%% \biboptions{comma,round}
+
+% \biboptions{}
+
+% if you have landscape tables
+\usepackage[figuresright]{rotating}
+
+% put your own definitions here:
+%   \newcommand{\cZ}{\cal{Z}}
+%   \newtheorem{def}{Definition}[section]
+%   ...
+
+% add words to TeX's hyphenation exception list
+%\hyphenation{author another created financial paper re-commend-ed Post-Script}
+
+% declarations for front matter
+
 \begin{document}
+
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for the associated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for the associated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for the associated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%%
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \address{Address\fnref{label3}}
+%% \fntext[label3]{}
+
+\dochead{}
+%% Use \dochead if there is an article header, e.g. \dochead{Short communication}
+
 \title{The CMS Data Aggregation System}
 
-\author{Valentin Kuznetsov}
-\address{Cornell University, Ithaca, New York, USA}
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{<author name>}
+%% \address[label1]{<address>}
+%% \address[label2]{<address>}
+
+%\author[vkuznet]{Valentin Kuznetsov\corref{cor1}}
+\author[vkuznet]{Valentin Kuznetsov}
+\address[vkuznet]{Cornell University, Ithaca, New York, USA}
 \ead{vkuznet@gmail.com}
 
-\author{Dave Evans}
-\address{Fermilab, Batavia, Illinois, USA}
+\author[evans]{Dave Evans}
+\address[evans]{Fermilab, Batavia, Illinois, USA}
 \ead{evansde@fnal.gov}
 
-\author{Simon Metson}
-\address{Bristol University, Bristol, UK}
+\author[metson]{Simon Metson}
+\address[metson]{Bristol University, Bristol, UK}
 \ead{s.metson@bristol.ac.uk}
 
+%\cortext[cor1]{Corresponding autho}
+
 \begin{abstract}
-Meta-data plays signiﬁcant role in large modern enterprises, 
+Meta-data plays significant role in large modern enterprises, 
 research experiments and digital libraries where it comes from a different 
@@
 and explore meta-data associated with petabytes of experimental data. 
-We combine simplpicity of keyword-based search with precision of RDMS
+We combine simplicity of keyword-based search with precision of RDMS
 under the new system. The aggregated information being collected from various sources,
@@
 to motivate this work.
-
 \end{abstract}
 
-\newpage
+\begin{keyword}
+%% keywords here, in the form: keyword \sep keyword
+Meta-data\sep data aggregation\sep HEP.
 
+%% PACS codes here, in the form: \PACS code \sep code
+
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
+\end{keyword}
+
+\end{frontmatter}
+
+%%
+%% Start line numbering here if you want
+%%
+% \linenumbers
+
+%% main text
 \section{Introduction}
@@
 At this scale the ``information discovery'' within heterogeneous, distributed 
-environment becomes a key ingredient of successful data analysis in HEP.
+environment becomes a key ingredient of successful data analysis.
 The data and associated meta-data are produced in variety of forms and digital formats.
@@
 \section{Related Work\label{RelatedWork}}
-Even though the idea of keyword queries over the relational database(s)
+Even though the idea of keyword queries over the relational database
 %querying relation databases via keyword based search
@@
 in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
-i.e. retrieval of only documents that contain all query keywords, has
+i.e. retrieval of documents that contain all query keywords, has
 been discussed. It requires that all keywords are matched in a row from a
@@
 end-users over the generated new database. These and other related 
-works\footnote{For more information about keyword based search over relational 
-databases we redirect users to references in \cite{DBXplorer, QueryAnswer}.} 
+works, see \cite{DBXplorer, QueryAnswer} for more references,
 are based on some kind of the graph over the underlying database and
-moreover use a generated database out of the pristine one.
+use a generated database out of pristine one to perform the work.
 The proximity of the results in those techniques 
@@
 {\it find me total number of files whose size more then 20 but less then 
-100} would be hard or almost impossible using keyword search techniques discussed 
-in \cite{DBXplorer, QueryAnswer}, while it can be easily accomplished
-using SQL syntax.
+100} would be hard or almost impossible in \cite{DBXplorer, QueryAnswer}, 
+while it can be easily accomplished using SQL syntax.
 
@@
 federated databases, see for example \cite{FedDB}. It unifies data coming 
-from different RDMS systems into the federated DB where SQL queries can be placed to 
-search for desired data. Even though it would be interesting to apply technique 
-discussed in \cite{DBS-QL} to a federated database it still requires 
+from different RDMS systems into the federated DB where SQL queries can be user to 
+search desired data. Even though it would be interesting to apply technique 
+discussed in \cite{DBS-QL} to a federated DB it still requires 
 an access to the underlying database schema. This obviously creates a problem
@@
 existing data-services whose nature and access policies are defined by developers.
-Keeping in mind that meta-data can come in different forms and data formats
+Keeping in mind that meta-data can come in different forms and data formats,
 we used a document-oriented database back-end with full support of the query language.
@@
 necessary glue across existing API/notations and user queries. This approach 
-allows us to perform the dynamic data aggregation across various data-services, achieve
+allows us to perform dynamic data aggregation across various data-services, achieve
 precise answers using keyword-based queries
@@
 the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
-It is designed to explore the frontier of the High Energy Physics and provide physicists
-ability to look at conditions presented at the early stage of our Universe.
+It is designed to explore the frontier of the HEP and provide physicists
+the ability to look at conditions presented at the early stage of our Universe.
 More then 3000 physicists from 183 institutions reprepresenting almost 
@@
 for archiving the data coming out from the detector, prompt first pass reconstruction
-and data distribution to Tier-1 centers. The 7 Tier-1 centers
-located in France, Germany, Ialy, Spain, Taiwan, United Kingdom and United States
-keep a portion (copy) of the data delivered by Tier-0 for further processing.
+and data distribution to 7 Tier-1 centers, located around the world.
+%The 7 Tier-1 centers
+%located in France, Germany, Ialy, Spain, Taiwan, United Kingdom and United States
+%keep a portion (copy) of the data delivered by Tier-0 for further processing.
 They provide storage and CPU power for high priority analysis jobs.
-The Tier-2 centers, located at more then 50 sites around the world,
+The 50+ Tier-2 centers
+%, located at more then 50 sites around the world,
 are dedicated for user analysis tasks and production of simulated data.
@@
 are located at CERN, it was never been a requirement in CMS computing
-and data model. For instance, the production teams operated at Tier-1,2
-centers set up local services for data bookkeeping and operational
-tasks. Based on their resources the choice of database back-end was
-up to the site managers. Therefore the CMS software were designed to support different
-DB back-ends and provide tools for migrating data across them.
+and data model. 
+%For instance, the production teams operated at Tier-1,2
+%centers set up local services for data bookkeeping and operational
+%tasks. Based on their resources the choice of database back-end was
+%up to the site managers. Therefore the CMS software were designed to support different
+%DB back-ends and provide tools for migrating data across them.
 
@@
 under one umrella, utilizing their access rules, APIs, security policies, etc., and
-allowing users to perform queries against them. In contrast, approaches
+allowing users to perform queries against them. In contrast, the approaches
 discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
@@
 by existing data-services and infrastructure, whose design, implementation and
-security policies are left up to their developers. Therefore DAS does not requires
+security policies are left up to their developers. Therefore DAS does not require
 a data preservation and transaction capabilities. Instead DAS can be treated
-as proxy or caching layer where information retrieval (data) are triggered dynamically
+as proxy or caching layer where information retrieval are triggered dynamically
 by user queries. Such infrastructure allows data sharing across different queries and
-data aggregation using appropriate mapping. Below we discuss details of
+data aggregation using appropriate mapping. Below we discuss in details the
 DAS architecture.
@@
 The design of the DAS system was based on previous studies of CMS Data 
-Bookkeeping System (DBS) \cite{DBS, DBS07}. We carefully analyzed user
-requests and queries placed into the DBS, their patterns, frequencies and latency. 
-As a result we created a data discovery service \cite{DD}. Originally it used MVC
-architecture and later, being re-factored as pure presentation layer on top of 
-the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
-DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
-we took away relations and details on underlying schema from end-users. 
-This was achieved by using Dijkstras shortest path algorithm 
-to establish necessary table joins based on provided information in user queries and
-knowledge of foreign key relationships from auto-loaded database schema.
-The DBS-QL keys exposed to end-users were choosen based on a jargon used by
-physicists. They're internally mapped into database entitites and attributes, 
-e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
-A quick adoption and wide usage of DBS-QL gave us confidence in choosen approach 
+Bookkeeping System (DBS) \cite{DBS, DBS07}. Teh DBS Data Discovery service
+\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provides
+ability to place queries, similar to SQL, without specific knowledge of 
+underlying DB schema.
+%We carefully analyzed user
+%requests and queries placed into the DBS, their patterns, frequencies and latency. 
+%As a result we created a data discovery service \cite{DD}. Originally it used MVC
+%architecture and later, being re-factored as pure presentation layer on top of 
+%the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
+%DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
+%we took away relations and details on underlying schema from end-users. 
+%This was achieved by using Dijkstras shortest path algorithm 
+%to establish necessary table joins based on provided information in user queries and
+%knowledge of foreign key relationships from auto-loaded database schema.
+%The DBS-QL keys exposed to end-users were choosen based on a jargon used by
+%physicists. They're internally mapped into database entitites and attributes, 
+%e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
+A quick adoption and wide usage of DBS-QL in CMS gave us confidence in choosen approach 
 and provide a basis for building DAS system.
@@
 \centering
-\includegraphics[width=120mm]{DAS_Cache_and_Analytics.pdf}
+\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
 \caption{
@@
 \noindent
-Due to this set of requirements, we rule out the choice of RDMS as DAS back-end(s)
+Based on this set we rule out the choice of RDMS as DAS back-end(s)
 for several reasons. We agreed to support dynamic queries, data
@@
 and auto-sharding. Our prelimiary benchmarks shown that it can sustain
-the desired load and size for our meta-data information. We used 
+the desired load and size of our meta-data information. We used 
 document-oriented database, MongoDB, as a back-end for the three 
@@
 
-The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of the
-following components:
+The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of
 core library with support of pluggable modules for data retrieval;
 caching layer to store data-service output and aggregated results;
-request service to handle user requests;
-mapping DB to keep information about data-service APIs, their
+request service to handle user queries;
+Mapping DB to keep information about data-service APIs, their
 notations and mapping to DAS;
-analytics DB for query analysis and identification of pre-fetch 
+Analytics DB for query analysis and identification of pre-fetch 
 strategies.
@@
 data on demand. This was done by invoking appropriate data-service plugin
-upon the user request. Each plugin consists of DAS map,
-data parser(s) suitable for data-service output format(s) into DAS common
-data format (JSON) and was responsible for data-service API calls.
+upon the user request. Each plugin consists of DAS map, the
+data parser(s) to convert data-service output into DAS records
+and it was responsible for data-service API calls.
 The data-service DAS map contains information about data-service APIs,
@@
 Below you can see a typical DAS Mapping DB record for CMS
-SiteDB data-service \cite{SiteDB}\footnote{We used regular expression 
-patterns, evaluated during run time, to validate provided set of input 
-paremters and DAS keys.}:
+SiteDB data-service \cite{SiteDB}:
+%\footnote{We used regular expression 
+%patterns, evaluated during run time, to validate provided set of input 
+%paremters and DAS keys.}:
 \begin{verbatim}
-{"api": {"params": {"name": ""}, "name": "CMSNametoSE"}, 
+{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
  "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
               "api_param": "name", "das_key": "site"}], 
- "_id": "4ae1f8dfe2194e4581000014", 
- "system": "sitedb", 
  "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
@@
 cases where special care was required. For instance,
-we dealt with data-services which were only accessible via private network.
+to deal with data-services which were only accessible via private network.
 Therefore, a way to propagate user credentials was required, leading
@@
 For instance, the CMS DBS system \cite{DBS}
-used {\it logical\_file\_name} notation to identify the name of the
-logical file, while the same entity was called {\it lfn} in CMS PhEDEx system \cite{PhEDEx}.
+used \verb+logical_file_name+ notation to identify the name of the
+logical file, while the same entity was called \verb+lfn+ 
+in CMS PhEDEx system \cite{PhEDEx}.
 Since both of them were referred to the logical file name, 
-we map both notations into DAS key {\it file}.\footnote{We
-also deal with physical file names, but still used {\it file} to refer to them in
-DAS notations, since simple regular expressions were used to identify the
-meaning of the {\it file} keyword from the provided value.}
-Such keys, e.g. {\it file}, were exposed to the end-users as pre-defined keywords to
-be used in search queries. For instance, end-users were able to type {\it file} in
+we map both notations into DAS key \verb+file+.
+%\footnote{We
+%also deal with physical file names, but still used {\it file} to refer to them in
+%DAS notations, since simple regular expressions were used to identify the
+%meaning of the {\it file} keyword from the provided value.}
+Such keys, e.g. \verb+file+, were exposed to the end-users as pre-defined keywords to
+be used in search queries. For instance, end-users were able to type \verb+file+ in
 order to get all file names known to DAS via data-services as well as
-provide a concrete patterns associated with a file, e.g. {\it file=abc*}.
+provide a concrete patterns associated with a file, e.g. \verb+file = abc*+.
 Under the hood, the appropriate data-service APIs were looked up along with appropriate
-parameters to answer those questions. For example,
-in aforementioned example, using {\it file} in DAS query cause call of DBS 
-{\it listFiles} and PhEDEx {\it fileReplicas} APIs, respectively.
+parameters to answer those questions. In aforementioned example, 
+usage of \verb+file+ in DAS query cause a call of DBS 
+\verb+listFiles+ and PhEDEx \verb+fileReplicas+ APIs, respectively.
 This approach gave power to end-users to express the meaning of their queries, especially
-when they deal with numbers and achieve accuracy of results from DAS. 
-For instance, users were able to type $run>10\,\,and\,\, run < 100$
+when they deal with numbers, and achieve accuracy of results. 
+For instance, users were able to type \verb+run > 10 and run < 100+
 to specify a range of run numbers to be looked up in DAS cache. At the same time
-a full free text based keyword search is under development within MongoDB.
+a full keyword search is under development.
+% within MongoDB.
 
@@
 algorithm:
-\begin{description}
-\item[  1.] Decompose query into set of conditions ($conds$) 
+\begin{enumerate}[1.]
+\item Decompose query into set of conditions ($conds$) 
 and selection keys ($skeys$)
-\item[  2.] Check cache for query hash, if found: look-up results from cache
-\item[  3.] Identify list of APIs which can serve requested data
+\item Check cache for query hash, if found: look-up results from the cache
+\item Identify list of APIs which can serve requested data
 \begin{verbatim}
@@
 \end{verbatim}
-\item[  4.] Iterate over APIs
+Here \verb+mappind_db+ defines translation between input query and data-service
+APIs which can delivery such data.
+\item Iterate over APIs
 \begin{verbatim}
 for url, api, args in apilist:
-    if  superset_in_analytics_found(api, args):
+    if  superset_in_analytics(api, args):
         continue
@@
 \end{verbatim}
-\item[  5.] Perform aggrgation by iterating over the previos step
+Here \verb+superset_in_analytics+ identifies if provided \verb+api, args+
+are subset of already registered in Analytics DB api call(s). The 
+\verb+get_data+ and \verb+parser+ are define methods to fetch and parse
+data, respectively.
+\item Perform aggregation by iterating over records from the previos step
 \begin{verbatim}
-if cache_has(row.match_key):
-   existing_record = get_from_cache(row.match_key)
+if cache_has(row.primary_key):
+   existing_record = get_from_cache(row.primary_key)
    merge(row, existing_record)
@@
 \end{verbatim}
-\item[  6.] Get results from the cache
-\end{description}
-The data was stored in a form of DAS records into the cache, while
-queries were recorded into Analytics DB for further analysis. We
-also had an option to run independent set of daemons, DAS robots, who
+We did use a bulk insert operations to increase throughput. The \verb+primary_key+
+is a key which identify the record defined in Mapping DB, e.g. \verb+file.name+
+for \verb+file+ record.
+\item Get results from the cache
+\end{enumerate}
+%The data was stored in a form of DAS records into the cache, while
+%queries were recorded into Analytics DB for further analysis.
+We also had an option to run independent set of daemons, DAS robots, who
 monitor Analytics DB for most popular requests and pre-fetch the most
@@
 express their queries in a free text-based form. Although we must clarify its
-meaning. We used a pre-defined set of keys known to DAS to describe common
-entities of our data. For example, {\it dataset, file, block, run, event}. Those
-names were used by our users in their daily work operations. So, by typing
-{\it file} we didn't perform a string matching to ``file'' string, but rather
-look-up all file objects in a DAS. Moreover each DAS key has a default attribute.
-For intance, for {\it dataset, block, file} the default attribute 
-was a {\it name}, while for {\it run} and {\it event} the default attribute was 
-a {\it number}. This allows to specify key-value pair as simple as possible, e.g.
-{\it run=123456}, which means run number 123456. At the same time a query 
-key, e.g. {\it site} meant to look-up all site objects. Such ability
+meaning. We used a pre-defined set of keys registered in DAS to describe common
+entities of our data. For example, \verb+dataset+, 
+\verb+file+, \verb+block+, \verb+run+, \verb+event+. Those
+names were used by our users in their daily work operations. So, by typing a
+\verb+file+ we didn't perform a string matching to the 
+``file'' string, but rather look-up all file objects in a DAS. 
+Moreover each DAS key has a default attribute.
+For intance, the default attribute for
+\verb+dataset+, \verb+block+, \verb+file+ 
+was a \verb+name+, while the default attribute for
+\verb+run+ and \verb+event+ was a \verb+number+. 
+This allows to specify key-value pair as simple as possible, e.g.
+\verb+run=123456+, which means run number 123456. At the same time a query 
+key, e.g. \verb+site+ meant to look-up all site objects. Such ability
 to use key-value pairs resolve ambiguity in object values. For instance,
-specifying a run number as {\it run=123456} instruct system to resolve
+specifying a run number as \verb+run=123456+ instructs system to resolve
 data object with run equal to 123456 rather then other objects which may have
@@
 The free text-based queries were transformed into MongoDB query syntax, 
-who was used internally by DAS. It represents a set of selection 
+who was used internally in DAS. It represents a set of selection 
 keys and conditions as a JSON document. For example, a query
 {\it dataset site=T1\_CH\_CERN}
-where transformed into 
+was transformed into 
 {\it \{"fields" : ["dataset"], "spec" : \{"site" : "T1\_CH\_CERN"\}\}}.
@@
 Bookkeeping System (DBS) and data location system (PhEDEx)
-provided information about CMS data files. In former case, DBS 
-stored information about files and their relationship to other 
-physics objects, such as run and luminosity, while PhEDEx kept
-information about the file location, e.g. details about sites 
-holding copy of the files. Upon a user request to find a file, 
-we were able to identify appropriate set of DBS and PhEDEx APIs, 
+provided information about CMS data block objects. In former case, DBS 
+stored information about blocks and their relationships to other 
+physics objects, such as dataset adn files, while PhEDEx kept
+information about its location and replicas. Upon a user request to 
+find a block, we were able to identify appropriate set of DBS and PhEDEx APIs, 
 retrieve data from them, re-format their output to DAS notations 
 and store results into DAS cache performing aggregation of matched 
-objects. For instance, since both data-services returns in their
-output a file name we used as a common key during aggregation.
-This merged results represented a set of DAS records. Each DAS
-record contains a standard DAS header, which identify look-up time,
-request url, input parameters and API method as well as response 
-version, expiration time and checksum of the output. Usage of this 
-information was valuable contribution in debugging process of
+objects using \verb+block.name+ key. The merged results were stored 
+as a set of data records. Here is an example of such aggregated record:
+\begin{verbatim}
+{"das_id": ["4b195404e2194e21c4000030", "4b1953fee2194e21c4000002"], 
+ "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
+            "subscribed": "n", "custodial": "n", "site": "T1_US_Buffer", 
+            "nfiles": "18", "se": "srm.site.com", 
+            "size": "112694171838"}, {....}]},
+           {"name": "a/b/c#123", "nevents": "570200", 
+            "created_by": "USER_DN", "path": "/a/b/c", ...}],
+ "das": [{"expire": 1259952908}, {"expire": 1259954702}]}
+\end{verbatim}
+Here \verb+das_id+ refers to api call records stored in Analytics DB,
+\verb+block+ represents an aggregation block object information from DBS
+and PhEDEx systems and \verb+das+ contains their expiration time stamps.
+Usage of this information was valuable contribution in debugging process of
 data-services themselves, e.g. identification of not-matched records, 
-latency studies, etc. Here is an example of DAS record representing
-aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
-and {\it CMStoSiteName}, with API results merged under the {\it site} key.
-\begin{verbatim}
-{"_id": "4ae1e7ffe2194e4243000003", 
- "site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
-         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
- "das": [{"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
-          "ctime": 0.95439004898071289, "expire": 1256362175.7692671, 
-          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSAMName", 
-          "timestamp": 1256318975.7692671, "system": ["sitedb"], 
-          "api": ["CMStoSAMName"], "version": "", "selection_keys": []}, 
-         {"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
-          "ctime": 1.2158100605010986, 
-          "api": ["CMStoSiteName"], 
-          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSiteName", 
-          "timestamp": 1256318974.7774379, "system": ["sitedb"], 
-          "expire": 1256362174.7774379, "version": "", "selection_keys": []}
-        ]
-}
-\end{verbatim}
+latency studies, etc.
+% Here is an example of DAS record representing
+%aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
+%and {\it CMStoSiteName}, with API results merged under the {\it site} key.
+%\begin{verbatim}
+%{"site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
+%         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
+% "das": [{"ctime": 0.9, "expire": 1256362175, "api": ["CMStoSAMName"],
+%          "url": "https://a.b.com/sitedb/json/index/CMStoSAMName", 
+%          "timestamp": 1256318975, "system": ["sitedb"]},
+%         {"ctime": 1.2, "api": ["CMStoSiteName"], "expire": 1256362174},
+%          "url": "https://a.b.com/sitedb/json/index/CMStoSiteName", 
+%          "timestamp": 1256318974, "system": ["sitedb"]]}
+%\end{verbatim}
 
@@
 \item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
-CMS file transfer system and data-service;
+CMS file transfer system;
 \item SiteDB \cite{SiteDB}, the CMS site resources data-service;
@@
 operates on private network accessible at CMS online operation center.
-Table \ref{db_size} shows the current size of meta-data from aforementioned 
-data-services.
+%Table \ref{db_size} shows the current size of meta-data from aforementioned 
+%data-services.
 %\begin{figure}[htb]
@@
 
-\begin{table*}[hbt]
-\centering
-\begin{tabular}{llllll}\hline
-\hline
+%\begin{table*}[hbt]
+%\centering
+%\begin{tabular}{llllll}\hline
+%\hline
 
-Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
-\hline
-tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
-indexies (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
-\hline
-\hline
-\end{tabular}
-\caption{
-Current size of meta-data for CMS data-services participated in DAS, before
-official data-taking. 
-}
-\label{db_size}
-\end{table*}
-
-At the moment we don't have exact numbers for amount of data which will stored 
-in those data-service and therefor accessible in DAS. Although we project to
-accumulate of the order of 500 GB of meta-data per year during normal LHC and CMS
-operations.\footnote{This number is based on tests 
-performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
+%Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
+%\hline
+%tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
+%indexies (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
+%\hline
+%\hline
+%\end{tabular}
+%\caption{
+%Current size of meta-data for CMS data-services participated in DAS, before
+%official data-taking. 
+%}
+%\label{db_size}
+%\end{table*}
+At the moment their size is quite small ranging from 1 to 40 GB and 2 to 140 GB in
+table and index sizes, respectively. But we estimate to accumulate $\sim$500 GB of 
+meta-data each year during normal CMS operation at LHC.
+%\footnote{This number is based on tests 
+%performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
 %In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
@@
 To test DAS performance we measured elapsed time and CPU/RAM utilization
-required to collect and aggregate block object information from DBS and 
-PhEDEx CMS data-services. Preliminary results of those measurements can
-be found in Table \ref{DAS_benchmark}.
+required to collect and aggregate information from different
+data-services. Table \ref{DAS_benchmark} summarizes the biggest 
+contributors so far, the block information between DBS and 
+PhEDEx CMS data-services.
 
@@
 \hline
-
 System & Format & Size & Records & Elapsed time & Elapsed time \\
@@
 \caption{Time required to fetch, parse and aggregate block information
-from DBS and PhEDEx systems via their APIs. In the case of PhEDEx
+from DBS and PhEDEx systems. In the case of PhEDEx
 system we show total number of fetched records 175884 together with
 number of records merged in DAS 170131. The total number of DAS records 
-were calculated as number of aggregated records plus left over, non-matched records,
-from both systems.}
+were calculated as number of aggregated records plus left overs,
+the non-matched records from both systems.}
 \label{DAS_benchmark}
@@
 
-Please note that initial DAS implementation was done in Python \cite{Python}
+Please note that initial DAS implementation was done in Python 
 and all tests were performed on x64 Linux platform with 16GB of RAM where
 1 CPU was used by DAS server and 1 CPU by MongoDB database server. The CPU utilization
-of DAS server was 67\% in average and RAM usage was below 35MB for
-all tests. All tests were performed multiple times to avoid fluctuations. 
+of DAS server was 67\% in average and RAM usage was below 35MB. 
+All tests were performed multiple times to avoid fluctuations. 
 The elapsed time measured during those tests consists of the 
@@
 the one of the sub-system (either DBS or PhEDEx) since the first one was injected into
-fresh database, while the second required the merging step.
+the fresh database, while the second required the merging step.
 
@@
 The time of PhEDEx component shown in Table \ref{DAS_benchmark} divided 
-among: retrieval + parsing time $\sim$110 sec and merging step $\sim$50 sec.
-
-As you can see from Table \ref{DAS_benchmark},
+among: retrieval + parsing step, $\sim$110 sec, and merging step, $\sim$50 sec.
+As shown in Table \ref{DAS_benchmark},
 the time spent to access individual components (last column) was quite
@@
 get DAS records on disk was about 45 seconds. This was mostly due to I/O and
-conversion between binary data format used by MongoDB, BSON, to DAS
+conversion operations between binary data format used by MongoDB, BSON, to DAS
 data format, JSON.
 
-Based on this results we measured that we can achieve 7000 docs/sec rate
+Based on results of stress tests we measured that we can achieve 7000 docs/sec rate
 for insertion, 3400 docs/sec for aggregation and 8500 docs/sec for reading operation
 in DAS, which is suitable for our needs. Since all operations were performed 
-on a single CPU core a further improvements are expected. Also, a necesarry
-tuning can be done by using native C++ MongoDB driver instead of python one.
+on a single CPU a further improvements are expected. Also, a necessary
+tuning can be done by using C++ driver instead of python one.
 
@@
 prototype in python for DAS system which is under commissioning phase right now 
-and performed performance studies shown in this paper. 
+and presented performance studies in this paper. 
 The CMS experiment is started to collect its data in December of 2009 and 
@@
 meta-data produced by experiment each year will be of the order of
-$\sim500$GB. The DAS should sustain such load and will provide generic 
+500GB. The DAS should sustain such load and will provide generic 
 ``data-discovery'' service for CMS experiment in years to come.
@@
 
+%% References
+%%
+%% Following citation commands can be used in the body text:
+%% Usage of \cite is as follows:
+%%   \cite{key}         ==>>  [#]
+%%   \cite[chap. 2]{key} ==>> [#, chap. 2]
+%%
+
+%% References with BibTeX database:
+
+%\bibliographystyle{elsarticle-num}
+%\bibliography{<your-bib-database>}
+
+%% Authors are advised to use a BibTeX database file for their reference list.
+%% The provided style file elsarticle-num.bst formats references in the required Procedia style
+
+%% For references without a BibTeX database:
+
+% \begin{thebibliography}{00}
+
+%% \bibitem must have the following form:
+%%   \bibitem{key}...
+%%
+
+% \bibitem{}
+
+% \end{thebibliography}
+
 \section*{References}
-\begin{thebibliography}{9}
+\begin{thebibliography}{00}
 \bibitem{DBXplorer}
@@
 ``The CMS Computing Model'',
-CERN-LHCC-2004-035/G-083 (2004);
-A. Fanfani et. al., 
-``Distributed Analysis in CMS'', 
-to be published in Journal of Grid Computing.
+CERN-LHCC-2004-035/G-083 (2004).
+%A. Fanfani et. al., 
+%``Distributed Analysis in CMS'', 
+%to be published in Journal of Grid Computing.
 
@@
 ``A multi-dimensional view on information retrieval of CMS data'',
-Journal of Physics: Conference Series 119, 072013, (2008).
+J. Phys. Conf. Ser, 119, 072013 (2008).
 
@@
 \bibitem{PhEDEx}
-Rehn et. al.,
-``PhEDEx high-throughput data transfer management system'', CHEP 2006;
-R. Egeland et al., 
-``Data transfer infrastructure for CMS data taking'', 
-Proceedings of Science, PoS (ACAT08) 033 (2008);
+%Rehn et. al.,
+%``PhEDEx high-throughput data transfer management system'', CHEP 2006;
+%R. Egeland et al., 
+%``Data transfer infrastructure for CMS data taking'', 
+%Proceedings of Science, PoS (ACAT08) 033 (2008);
 L. Tuura et al., 
@@
 
-\bibitem{CRAFT09}
-https://twiki.cern.ch/twiki/bin/view/CMS/CRAFT09AnalysisInfo
+%\bibitem{CRAFT09}
+%https://twiki.cern.ch/twiki/bin/view/CMS/CRAFT09AnalysisInfo
 
-\bibitem{Python} Python programming language, http://www.python.org
+%\bibitem{Python} Python programming language, http://www.python.org
 
@@
 \end{thebibliography}
-
 \end{document}

### 2010-05-04 doc/paper.tex
Bug fix
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@649 4525493e-7705-40b1-a816-d608a930855b

@@
 due to longer retrieval and parsing time.
-The time of PhEDEx component shown in Table \ref{DAS_benchmark} divided 
+The time of PhEDEx component shown in Table \ref{DAS_benchmark} is divided 
 among: retrieval + parsing step, $\sim$110 sec, and merging step, $\sim$50 sec.

### 2010-05-04 doc/paper.tex
More tune-ups
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@650 4525493e-7705-40b1-a816-d608a930855b

@@
 information retrieval approach to non-relation data sources,
-we can't do so for relational ones, where information can be accessed via
+we can't do so for relational ones, where information is accessed via
 pre-established set of data-services.
@@
 based on studies  of use cases of the CMS particle physics experiment at 
-the LHC where we used a large scale, distributed computing system 
+the LHC where we used a large scale, distributed computing environment 
 to motivate this work.
@@
 %% keywords here, in the form: keyword \sep keyword
-Meta-data\sep data aggregation\sep HEP.
+Meta-data\sep data aggregation\sep information discovery\sep HEP.
 
@@
 The proximity of the results in those techniques 
-are mostly confined to the text-based values stored in a database, since 
+are mostly confined to a text-based values stored in a database, since 
 usage of numerical values in the input search field is mostly useless without 
@@
 the end-users. As a results a human questions were intuitively mapped into 
-simple queries. For example, the question
+simple queries. For instance, the question
 {\it I'm looking for files who contain data taken on certain date and located at
@@
 federated databases, see for example \cite{FedDB}. It unifies data coming 
-from different RDMS systems into the federated DB where SQL queries can be user to 
+from different RDMS systems into the federated DB where SQL queries can be used to 
 search desired data. Even though it would be interesting to apply technique 
@@
 an access to the underlying database schema. This obviously creates a problem
-in distributed heterogeneous environment where different security policies
+in distributed, heterogeneous environment where different security policies
 are in place.
@@
 for archiving the data coming out from the detector, prompt first pass reconstruction
-and data distribution to 7 Tier-1 centers, located around the world.
+and data distribution to Tier-1 centers, located around the world.
 %The 7 Tier-1 centers
@@
 They provide storage and CPU power for high priority analysis jobs.
-The 50+ Tier-2 centers
+The Tier-2 centers
 %, located at more then 50 sites around the world,
@@
 
-A broad variety of data-services being designed and developed to
+A broad variety of data-services have being designed and developed to
 maintain detector and production operations, including detector
@@
 \section{Data Aggregation System\label{DAS}}
-The goal of the Data Aggregation System was to combine existing data-services
+The goal of the Data Aggregation System was to combine and
+aggregate data from existing data-services
 under one umrella, utilizing their access rules, APIs, security policies, etc., and
@@
 enforced a publication step of data into a dedicated database 
-or central repositories \cite{iRODS}. Quite often it is required
+or central repositories \cite{iRODS}. Therefore, quite often it is required
 that data should be converted into uniform data formats in order enable
 aggregation, see \cite{OpenArchive}. In DAS architecture the data are provided
-by existing data-services and infrastructure, whose design, implementation and
-security policies are left up to their developers. Therefore DAS does not require
-a data preservation and transaction capabilities. Instead DAS can be treated
+by existing data-services and infrastructure, whose implementation,
+data representation and security policies are left intact. Consequently,
+DAS does not require a data preservation and transaction capabilities. 
+Instead DAS can be treated
 as proxy or caching layer where information retrieval are triggered dynamically
@@
 \subsection{DAS architecture}
-The design of the DAS system was based on previous studies of CMS Data 
-Bookkeeping System (DBS) \cite{DBS, DBS07}. Teh DBS Data Discovery service
-\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provides
-ability to place queries, similar to SQL, without specific knowledge of 
-underlying DB schema.
+The design of the DAS system was based on previous studies of the CMS Data 
+Bookkeeping System (DBS) \cite{DBS, DBS07}. The DBS Data Discovery service
+\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
+a syntax similar to SQL, without specific knowledge of underlying DB schema.
 %We carefully analyzed user
@@
 \begin{itemize}
-\item support of keyword based search queries with ability to use condition operators;
-\item pluggable interface to existing and up-coming data-service APIs;
-\item support heterogeneous software environment and distributed nature of the data-services;
-\item preserve security policies of individual data-services;
+\item support a keyword based search queries with ability to use condition operators;
+%\item pluggable interface to existing and up-coming data-service APIs;
+\item use existing heterogeneous software environment and distributed nature of the data-services;
+\item preserve access rules and security policies of individual data-services;
 %\item be flexible to adjust to dynamic nature of data-service APIs, 
@@
 multiple data-services;
-\item be transparent to data-services implementations and their data formats;
+\item be transparent to data-services implementation and their data formats;
 %\item support legacy data-services, applications and APIs.
@@
 The DAS architecture diagram. 
-Input query where placed into analytics DB and mapped into set of
+Input query were placed into analytics DB and mapped into a set of
 data-service APIs and their parameters. If superset of the query has been
-found in DAS cache, the appropriate results were delivered back to users.
+found in DAS cache, the appropriate results were delivered back to the users.
 Otherwise a data-service APIs were invoked, triggering DAS cache population. 
-Once data become available and aggregated in DAS cache users were able 
+Once data become available and aggregated in DAS cache, users were able 
 to see the results. This on demand system was supplemented by DAS robots 
@@
 The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of
-core library with support of pluggable modules for data retrieval;
-caching layer to store data-service output and aggregated results;
+the core library with support of pluggable modules for data retrieval;
+caching layer to store data-service output and aggregate results;
 request service to handle user queries;
@@
 \subsection{DAS core}
-The DAS core communicates with different data-services and retrieves
-data on demand. This was done by invoking appropriate data-service plugin
+The DAS core communicated with different data-services and retrieved
+data on demand. This was done by invoking appropriate data-service plugin(s)
 upon the user request. Each plugin consists of DAS map, the
@@
 and it was responsible for data-service API calls.
-The data-service DAS map contains information about data-service APIs,
-input and output parameters and their mapping to DAS keys. 
-This information was stored into DAS Mapping DB which we discuss next.
-Below you can see a typical DAS Mapping DB record for CMS
-SiteDB data-service \cite{SiteDB}:
-%\footnote{We used regular expression 
-%patterns, evaluated during run time, to validate provided set of input 
-%paremters and DAS keys.}:
+The data-service plugin contains a map about data-service APIs,
+input and output parameters and their mapping to DAS keys. For example
 \begin{verbatim}
@@
 \end{verbatim}
-Although the majority of data-services looks alike
+This information was stored into DAS Mapping DB which we discuss next.
+
+Although the majority of data-service plugins looks alike
 and their logic was abstracted, including parsers, there were a few
-cases where special care was required. For instance,
-to deal with data-services which were only accessible via private network.
-Therefore, a way to propagate user credentials was required, leading
-to specific plugin implementation.
+cases where special care was required, e.g. access to data-services 
+accessible via private network. Therefore, a way to propagate 
+user credentials was required, leading to specific plugin implementation.
 
@@
 \subsection{DAS Mapping DB}
-DAS Mapping DB holded the mapping between data-service APIs and DAS keys exposed to
+The Mapping DB holded the mapping between data-service APIs and DAS keys exposed to
 the users. It created a necessary glue between data-services and identifies
@@
 we stored their names, input parameters, together with their type and accept 
-values as regular expression patterns into the Mapping DB. 
+values, as regular expression patterns, into the Mapping DB. 
 Also we stored the mapping between data-service notations and DAS keys. 
 For instance, the CMS DBS system \cite{DBS}
-used \verb+logical_file_name+ notation to identify the name of the
+used a \verb+logical_file_name+ notation to identify the name of the
 logical file, while the same entity was called \verb+lfn+ 
@@
 \verb+listFiles+ and PhEDEx \verb+fileReplicas+ APIs, respectively.
-This approach gave power to end-users to express the meaning of their queries, especially
+This approach gave ability to end-users to express the meaning of their queries, especially
 when they deal with numbers, and achieve accuracy of results. 
@@
 \begin{enumerate}[1.]
-\item Decompose query into set of conditions ($conds$) 
-and selection keys ($skeys$)
-\item Check cache for query hash, if found: look-up results from the cache
+\item Decompose query into set of conditions (\verb+conds+) 
+and selection keys (\verb+skeys+)
+\item Check cache for query hash, if found, look-up results from the cache
 \item Identify list of APIs which can serve requested data
@@
 \end{verbatim}
-Here \verb+superset_in_analytics+ identifies if provided \verb+api, args+
+Here \verb+superset_in_analytics+ verifies if provided \verb+api, args+
 are subset of already registered in Analytics DB api call(s). The 
@@
 \end{verbatim}
-We did use a bulk insert operations to increase throughput. The \verb+primary_key+
+The \verb+primary_key+
 is a key which identify the record defined in Mapping DB, e.g. \verb+file.name+
 for \verb+file+ record.
+We did use a bulk insert operations at this step to increase throughput. 
 \item Get results from the cache
@@
 \subsection{DAS queries}
-Apart from previous DBS-QL syntax we decided to drop its syntax and allow users
+Apart from the previous DBS-QL syntax we decided to drop its syntax and allow users
 express their queries in a free text-based form. Although we must clarify its

### 2010-05-04 src/python/DAS/web/DASCacheModel.py
query is presented in args dict, no needs to record it separately
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@662 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 DAS cache RESTfull model, based on WMCore/WebTools
 """
 
-__revision__ = "$Id: DASCacheModel.py,v 1.22 2009/12/08 15:16:20 valya Exp $"
-__version__ = "$Revision: 1.22 $"
+__revision__ = "$Id: DASCacheModel.py,v 1.23 2009/12/08 15:18:15 valya Exp $"
+__version__ = "$Revision: 1.23 $"
 __author__ = "Valentin Kuznetsov"
 
 # system modules
@@ -164,7 +164,7 @@ class DASCacheModel(RESTModel):
         """
         query = cherrypy.request.params.get("query", None)
         qhash = genkey(query)
-        doc = dict(query=query, qhash=qhash, timestamp=time.time(),
+        doc = dict(qhash=qhash, timestamp=time.time(),
                 method=cherrypy.request.method,
                 path=cherrypy.request.path_info,
                 args=cherrypy.request.params,

### 2010-05-04 src/python/DAS/web/DASCacheModel.py
Added header
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@663 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 DAS cache RESTfull model, based on WMCore/WebTools
 """
 
-__revision__ = "$Id: DASCacheModel.py,v 1.23 2009/12/08 15:18:15 valya Exp $"
-__version__ = "$Revision: 1.23 $"
+__revision__ = "$Id: DASCacheModel.py,v 1.24 2009/12/08 15:38:11 valya Exp $"
+__version__ = "$Revision: 1.24 $"
 __author__ = "Valentin Kuznetsov"
 
 # system modules
@@ -164,7 +164,21 @@ class DASCacheModel(RESTModel):
         """
         query = cherrypy.request.params.get("query", None)
         qhash = genkey(query)
+        headers = cherrypy.request.headers
         doc = dict(qhash=qhash, timestamp=time.time(),
+                header=dict(
+                agent=headers.get('User-Agent', ''),
+                accept=headers.get('Accept', ''),
+                cache=headers.get('Cache-Control', ''),
+                date=headers.get('Date', ''),
+                etag=headers.get('ETag', ''),
+                expect=headers.get('Expect', ''),
+                expires=headers.get('Expires', ''),
+                host=headers.get('Host', ''),
+                last_modified=headers.get('Last-Modified', ''),
+                referer=headers.get('Referer', ''),
+                server=headers.get('Server', ''),
+                ),
                 method=cherrypy.request.method,
                 path=cherrypy.request.path_info,
                 args=cherrypy.request.params,

### 2010-05-04 src/python/DAS/web/DASCacheModel.py
Header's dict can be passed directly
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@664 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 DAS cache RESTfull model, based on WMCore/WebTools
 """
 
-__revision__ = "$Id: DASCacheModel.py,v 1.24 2009/12/08 15:38:11 valya Exp $"
-__version__ = "$Revision: 1.24 $"
+__revision__ = "$Id: DASCacheModel.py,v 1.25 2009/12/08 15:41:13 valya Exp $"
+__version__ = "$Revision: 1.25 $"
 __author__ = "Valentin Kuznetsov"
 
 # system modules
@@ -166,19 +166,7 @@ class DASCacheModel(RESTModel):
         qhash = genkey(query)
         headers = cherrypy.request.headers
         doc = dict(qhash=qhash, timestamp=time.time(),
-                header=dict(
-                agent=headers.get('User-Agent', ''),
-                accept=headers.get('Accept', ''),
-                cache=headers.get('Cache-Control', ''),
-                date=headers.get('Date', ''),
-                etag=headers.get('ETag', ''),
-                expect=headers.get('Expect', ''),
-                expires=headers.get('Expires', ''),
-                host=headers.get('Host', ''),
-                last_modified=headers.get('Last-Modified', ''),
-                referer=headers.get('Referer', ''),
-                server=headers.get('Server', ''),
-                ),
+                headers=cherrypy.request.headers,
                 method=cherrypy.request.method,
                 path=cherrypy.request.path_info,
                 args=cherrypy.request.params,

### 2010-05-04 doc/paper.tex
Fix spelling with ispell
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@665 4525493e-7705-40b1-a816-d608a930855b

@@
 
-%\cortext[cor1]{Corresponding autho}
+%\cortext[cor1]{Corresponding author}
 
@@
 The data and associated meta-data are produced in variety of forms and digital formats.
-They are stored and retrieved from relational and non-releational data-sources, such as 
+They are stored and retrieved from relational and non-relational data-sources, such as 
 RDMS systems, document oriented databases, blogs, twikies, file systems,
@@
 tool(s) to look-up their desired data. A well-known solutions, such as data-services
-and search engines are tighted to a specific data sources and end-users are left 
-with a manual task to bookkeep and relatate information from them.
+and search engines are tighten to a specific data sources and end-users are left 
+with a manual task to bookkeep and relate information from them.
 
@@
 entity (table.column) in the underlying database.
-Therefore to answer presicely the logical questions, such as  
+Therefore to answer precisely the logical questions, such as  
 {\it find me total number of files whose size more then 20 but less then 
@@
 the ability to look at conditions presented at the early stage of our Universe.
-More then 3000 physicists from 183 institutions reprepresenting almost 
+More then 3000 physicists from 183 institutions representing almost 
 38 countries are involved in the design, construction and maintenance of the experiment.
@@
 hardware, mass and storage elements and configuration of the
-clusters. To accomodate efficient data processing CMS uses
+clusters. To accommodate efficient data processing CMS uses
 a multi-Tier distributed model, where specific tasks of data taking,
@@
 %The 7 Tier-1 centers
-%located in France, Germany, Ialy, Spain, Taiwan, United Kingdom and United States
+%located in France, Germany, Italy, Spain, Taiwan, United Kingdom and United States
 %keep a portion (copy) of the data delivered by Tier-0 for further processing.
@@
 The goal of the Data Aggregation System was to combine and
-aggregate data from existing data-services
-under one umrella, utilizing their access rules, APIs, security policies, etc., and
+aggregate meta-data from existing data-services
+under one umbrella, utilizing their access rules, APIs, security policies, etc., and
 allowing users to perform queries against them. In contrast, the approaches
@@
 %we took away relations and details on underlying schema from end-users. 
-%This was achieved by using Dijkstras shortest path algorithm 
+%This was achieved by using Dijkstra's shortest path algorithm 
 %to establish necessary table joins based on provided information in user queries and
 %knowledge of foreign key relationships from auto-loaded database schema.
-%The DBS-QL keys exposed to end-users were choosen based on a jargon used by
-%physicists. They're internally mapped into database entitites and attributes, 
+%The DBS-QL keys exposed to end-users were chosen based on a jargon used by
+%physicists. They're internally mapped into database entities and attributes, 
 %e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
-A quick adoption and wide usage of DBS-QL in CMS gave us confidence in choosen approach 
+A quick adoption and wide usage of DBS-QL in CMS gave us confidence in chosen approach 
 and provide a basis for building DAS system.
@@
 to see the results. This on demand system was supplemented by DAS robots 
-(unix daemons) who were consulting DAS analytics database to enable data 
+(Unix daemons) who were consulting DAS analytics database to enable data 
 population into DAS cache for most common query requests.
@@
 of concrete schema for data storage was problematic. Moreover we didn't impose
-any transactions and data persistency in DAS. This allows us to look 
+any transactions and data persistence in DAS. This allows us to look 
 for alternative IT solutions.
@@
 MongoDB was obvious, due to its support of dynamic queries, 
-full indexies, including inner objects and embedded arrays,
-and auto-sharding. Our prelimiary benchmarks shown that it can sustain
+full indexes, including inner objects and embedded arrays,
+and auto-sharding. Our preliminary benchmarks shown that it can sustain
 the desired load and size of our meta-data information. We used 
@@
 The DAS core communicated with different data-services and retrieved
-data on demand. This was done by invoking appropriate data-service plugin(s)
-upon the user request. Each plugin consists of DAS map, the
+data on demand. This was done by invoking appropriate data-service plug-in(s)
+upon the user request. Each plug-in consists of DAS map, the
 data parser(s) to convert data-service output into DAS records
 and it was responsible for data-service API calls.
-The data-service plugin contains a map about data-service APIs,
+The data-service plug-in contains a map about data-service APIs,
 input and output parameters and their mapping to DAS keys. For example
@@
 
-Although the majority of data-service plugins looks alike
+Although the majority of data-service plug-ins looks alike
 and their logic was abstracted, including parsers, there were a few
@@
 accessible via private network. Therefore, a way to propagate 
-user credentials was required, leading to specific plugin implementation.
+user credentials was required, leading to specific plug-in implementation.
 
@@
 \subsection{DAS Mapping DB}
-The Mapping DB holded the mapping between data-service APIs and DAS keys exposed to
+The Mapping DB hold the mapping between data-service APIs and DAS keys exposed to
 the users. It created a necessary glue between data-services and identifies
@@
 data, respectively.
-\item Perform aggregation by iterating over records from the previos step
+\item Perform aggregation by iterating over records from the previous step
 \begin{verbatim}
@@
 entities of our data. For example, \verb+dataset+, 
-\verb+file+, \verb+block+, \verb+run+, \verb+event+. Those
-names were used by our users in their daily work operations. So, by typing a
-\verb+file+ we didn't perform a string matching to the 
-``file'' string, but rather look-up all file objects in a DAS. 
-Moreover each DAS key has a default attribute.
-For intance, the default attribute for
-\verb+dataset+, \verb+block+, \verb+file+ 
-was a \verb+name+, while the default attribute for
-\verb+run+ and \verb+event+ was a \verb+number+. 
-This allows to specify key-value pair as simple as possible, e.g.
-\verb+run=123456+, which means run number 123456. At the same time a query 
-key, e.g. \verb+site+ meant to look-up all site objects. Such ability
-to use key-value pairs resolve ambiguity in object values. For instance,
-specifying a run number as \verb+run=123456+ instructs system to resolve
-data object with run equal to 123456 rather then other objects which may have
-123456 values, e.g. event number or file size. At the same time, to complement
-the stregth of the DAS we do plan to index data objects and add a keyword-based 
+\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
+was based on analysis of daily work operations of our users, see \cite{DBS07}.
+Each key had a set of attributes with primary one assigned to the key itself.
+For instance, the \verb+block+ key had the following attributes
+\verb+name+, \verb+size+, \verb+replica+ etc., 
+with primary attribute being \verb+name+.
+The choice of attributes were based on a set of meta-data information
+associated with a given DAS key. Therefore end-users where able to specify
+a context in their queries by using appropriate attribute(s). For example,
+for query \verb+block+ we didn't perform a string matching to the
+``block'' string, but rather look-up a block objects in a DAS, while query
+\verb+block.name=abc*+ instructs system to look-up appropriate blocks
+with matched names. At the same time, to complement the strength of the 
+DAS, we do plan to index dynamically retrieved meta-data and add a keyword-based 
 search later.
@@
 Bookkeeping System (DBS) and data location system (PhEDEx)
-provided information about CMS data block objects. In former case, DBS 
+provided information about CMS block meta-data. In former case, DBS 
 stored information about blocks and their relationships to other 
-physics objects, such as dataset adn files, while PhEDEx kept
+physics objects, such as dataset and files, while PhEDEx kept
 information about its location and replicas. Upon a user request to 
-find a block, we were able to identify appropriate set of DBS and PhEDEx APIs, 
+find a block meta-data, we were able to identify appropriate set of DBS and PhEDEx APIs, 
 retrieve data from them, re-format their output to DAS notations 
@@
 Here \verb+das_id+ refers to api call records stored in Analytics DB,
-\verb+block+ represents an aggregation block object information from DBS
+\verb+block+ represents an aggregation block meta-data information from DBS
 and PhEDEx systems and \verb+das+ contains their expiration time stamps.
@@
 %Grid certificates. Instead of propagating user certificate to data-service
-%the DAS cache server works as a proxy to this server and used host certficiate 
+%the DAS cache server works as a proxy to this server and used host certificate 
 %to serve such requests on behalf of the users.
@@
 \section{Results\label{Results}}
-In this section we discuss preliminary restults achieved with DAS system
+In this section we discuss preliminary results achieved with DAS system
 using the following CMS data-services:
@@
 %tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
-%indexies (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
+%indexes (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
 %\hline
@@
 %\end{table*}
-At the moment their size is quite small ranging from 1 to 40 GB and 2 to 140 GB in
-table and index sizes, respectively. But we estimate to accumulate $\sim$500 GB of 
-meta-data each year during normal CMS operation at LHC.
+At the moment, the amount of meta-data stored in all services is equal to
+$\sim$75 GB and $\sim$200 GB in tables and index sizes, respectively. While 
+we estimate to accumulate $\sim$500 GB of the meta-data each year during 
+normal CMS operation at LHC.
 %\footnote{This number is based on tests 
@@
 %HLTMON FEVTHLTALL: 31 million events, 7.4TB
-%Above event numbers have signicant overlap by design, 
+%Above event numbers have significant overlap by design, 
 %2.2 billion unique events, including 524 million cosmic muon
@@
 data-services. Table \ref{DAS_benchmark} summarizes the biggest 
-contributors so far, the block information between DBS and 
+contributors so far, the block meta-data information provided by DBS and 
 PhEDEx CMS data-services.
@@
 
-Please note that initial DAS implementation was done in Python 
-and all tests were performed on x64 Linux platform with 16GB of RAM where
-1 CPU was used by DAS server and 1 CPU by MongoDB database server. The CPU utilization
-of DAS server was 67\% in average and RAM usage was below 35MB. 
+The initial DAS implementation was done in Python language \cite{Python}
+and all tests were performed on x64 Linux platform where
+1 CPU was used by DAS server and 1 CPU by MongoDB database server. 
+%The CPU utilization
+%of DAS server was 67\% in average and RAM usage was below 35MB. 
 All tests were performed multiple times to avoid fluctuations. 
 The elapsed time measured during those tests consists of the 
-following componenets:
+following components:
 \begin{itemize}
@@
 {\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
-is a time required to convert notations used by data-service to DAS ones for every object
-we parse, {\it cache insertion} and {\it indexing time} represents time spend to add objects into
-the DAS cache, {\it aggregation time} is required to merge objects into DAS records based
-on their common key (block name in this case) and {\it output creation time}
+is a time required to convert data-service meta-data to DAS records for every object
+we parse, {\it cache insertion} and {\it indexing time} represents time spent on DAS 
+cache back-end, {\it aggregation time} is required to merge objects into DAS records based
+on their common key (\verb+block.name+ in this case) and {\it output creation time}
 is a time required to write DAS records to disk. Please note that last two
 components, {\it aggregation} and {\it output creation time}, were only applied to
-the one of the sub-system (either DBS or PhEDEx) since the first one was injected into
-the fresh database, while the second required the merging step.
+one of the sub-system, either DBS or PhEDEx, depending on insertion order.
+% since the first one was injected into
+%the fresh database, while the second required the merging step.
 
@@
 The time of PhEDEx component shown in Table \ref{DAS_benchmark} is divided 
-among: retrieval + parsing step, $\sim$110 sec, and merging step, $\sim$50 sec.
-As shown in Table \ref{DAS_benchmark},
+among retrieval and parsing, $\sim$110 sec, and merging steps, $\sim$50 sec,
+respectively. As shown in Table \ref{DAS_benchmark},
 the time spent to access individual components (last column) was quite
@@
 
-Based on results of stress tests we measured that we can achieve 7000 docs/sec rate
-for insertion, 3400 docs/sec for aggregation and 8500 docs/sec for reading operation
-in DAS, which is suitable for our needs. Since all operations were performed 
-on a single CPU a further improvements are expected. Also, a necessary
-tuning can be done by using C++ driver instead of python one.
+Based on results of stress tests we measured that we can achieve throughput of
+7000 docs/sec for insertion, 3400 docs/sec for aggregation and 8500 docs/sec for 
+reading operations in DAS, which are suitable for our needs. Since all 
+operations were performed on a single CPU a further improvements are expected. 
+Also, a necessary tuning can be done by using C++ driver instead of python one.
 
@@
 The CMS experiment is started to collect its data in December of 2009 and 
-we expect to accumulate a few PB of real data each year to tape. 
+we expect to accumulate a few PB of real data each year. 
 In addition, the Monte-Carlo samples will be produced at this scale.
 Based on test studies performed in CMS, we expect that total size of
-meta-data produced by experiment each year will be of the order of
+meta-data produced each year by experiment will be of the order of
 500GB. The DAS should sustain such load and will provide generic 
@@
 
-%\bibitem{Python} Python programming language, http://www.python.org
+\bibitem{Python} Python programming language, http://www.python.org
 

### 2010-05-04 src/python/DAS/services/abstract_service.py
Bug fix
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@668 4525493e-7705-40b1-a816-d608a930855b

@@ -4,8 +4,8 @@
 """
 Abstract interface for DAS service
 """
-__revision__ = "$Id: abstract_service.py,v 1.56 2009/12/11 16:37:56 valya Exp $"
-__version__ = "$Revision: 1.56 $"
+__revision__ = "$Id: abstract_service.py,v 1.57 2009/12/11 18:25:27 valya Exp $"
+__version__ = "$Revision: 1.57 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -208,14 +208,14 @@ class DASAbstractService(object):
         msg  = 'DBSAbstractService::pass_apicall, %s, API=%s, args=%s'\
         % (self.name, api, api_params)
         for row in self.analytics.col.find(spec):
-            input_query = dict(spec=api_params)
-            exist_query = dict(spec=row['apicall']['api_params'])
-            if  compare_specs(input_query, exists_query):
+            input_query = {'spec':api_params}
+            exist_query = {'spec':row['apicall']['api_params']}
+            if  compare_specs(input_query, exist_query):
                 msg += '\nwill re-use existing api call with args=%s'\
                 % row['apicall']['api_params']
                 self.logger.info(msg)
-                return True
-        return False
+                return False
+        return True
 
     def lookup_keys(self, api):
         """
@@ -358,6 +358,6 @@ class DASAbstractService(object):
             if 'required' in args.values():
                 continue
             # check if analytics db has a similar API call
-            if  self.pass_apicall(url, api, args):
+            if  not self.pass_apicall(url, api, args):
                 continue
             yield url, api, args

### 2010-05-04 src/python/DAS/web/das_cacheconfig.py
Increase size of logging.db to 10GB
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@677 4525493e-7705-40b1-a816-d608a930855b

@@ -6,8 +6,8 @@
 DAS cache server web configuration file
 """
 
-__revision__ = "$Id: das_cacheconfig.py,v 1.9 2009/12/13 02:51:07 valya Exp $"
-__version__ = "$Revision: 1.9 $"
+__revision__ = "$Id: das_cacheconfig.py,v 1.10 2009/12/14 17:38:22 valya Exp $"
+__version__ = "$Revision: 1.10 $"
 __author__ = "Valentin Kuznetsov"
 
 #
@@ -61,4 +61,4 @@ active.rest.logLevel = 'INFO'
 # verbose defines level of logger
 active.rest.sleep = 1
 active.rest.verbose = 0
-active.rest.capped_size = 104857600 # 100 MB
+active.rest.capped_size = 10737418240 # 10 GB

### 2010-05-04 doc/paper.tex
Update numbers
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@678 4525493e-7705-40b1-a816-d608a930855b

@@
 \centering
-\begin{tabular}{llllll}\hline
+\begin{tabular}{lllll}\hline
 \hline
-System & Format & Size & Records & Elapsed time & Elapsed time \\
-& & & & no cached data & w/ cached data \\
+System & Format & Records & Elapsed time & Elapsed time \\
+& & & no cached data & w/ cached data \\
 \hline
-DAS (DBS) & XML & 187MB & 368716 & 53 sec & 0.92 sec \\
-DAS (PhEDEx) & XML & 164MB & 175884/170131 & 160 sec & 0.94 sec \\
-DAS (aggregate) & JSON & 357MB & 374468 & 256 sec & 44.8 sec \\
+DAS (DBS) & XML & 382396 & 63 sec & 0.98 sec \\
+DAS (PhEDEx) & XML & 186145/180439 & 164 sec & 0.98 sec \\
+DAS (aggregate) & JSON & 388108 & 234 sec & 2.05/50.7 sec \\
 \hline
@@
 from DBS and PhEDEx systems. In the case of PhEDEx
-system we show total number of fetched records 175884 together with
-number of records merged in DAS 170131. The total number of DAS records 
+system we show total number of fetched records 186145 together with
+number of records merged in DAS 180439. The total number of DAS records 
 were calculated as number of aggregated records plus left overs,
-the non-matched records from both systems.}
+the non-matched records from both systems. The look-up time is presented in
+last column, where for DAS it shows as query look-up time (2.05 sec) 
+and time to get all records to disk (50.7 sec).}
 \label{DAS_benchmark}
@@
 respectively. As shown in Table \ref{DAS_benchmark},
-the time spent to access individual components (last column) was quite
+the look-up time spent to access individual components (last column) was quite
 reasonable, roughly 1 second for both systems. While final time to
-get DAS records on disk was about 45 seconds. This was mostly due to I/O and
+get DAS records on disk was about 50 seconds. This was mostly due to I/O and
 conversion operations between binary data format used by MongoDB, BSON, to DAS
@@
 Based on results of stress tests we measured that we can achieve throughput of
-7000 docs/sec for insertion, 3400 docs/sec for aggregation and 8500 docs/sec for 
+$\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
+$\sim$7600 docs/sec for 
 reading operations in DAS, which are suitable for our needs. Since all 

### 2010-05-04 src/python/DAS/extensions/__init__.py
Added dict_handler c-extension which creates new dict from parsed record according to provided notation mapping. The most called function when we parse XML data. Will need to write similar one for JSON parser
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@679 4525493e-7705-40b1-a816-d608a930855b

@@ -1,0 +1,10 @@
+#!/usr/bin/env python
+#-*- coding: ISO-8859-1 -*-
+
+"""
+DAS extensions
+"""
+
+__revision__ = "$Id: __init__.py,v 1.1 2009/12/15 02:06:00 valya Exp $"
+__version__ = "$Revision: 1.1 $"
+__author__ = "Valentin Kuznetsov"

### 2010-05-04 src/python/DAS/utils/utils.py

@@ -5,8 +5,8 @@
 General set of useful utilities used by DAS
 """
 
-__revision__ = "$Id: utils.py,v 1.48 2009/12/14 15:40:42 valya Exp $"
-__version__ = "$Revision: 1.48 $"
+__revision__ = "$Id: utils.py,v 1.49 2009/12/15 02:06:00 valya Exp $"
+__version__ = "$Revision: 1.49 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -689,14 +689,17 @@ def access(data, elem):
                         for item in result:
                             yield item
 
-def dict_helper(notations, idict):
+def dict_helper(idict, notations):
     """Create new dict for provided notations/dict"""
-    child_dict = {}
-    for kkk, vvv in idict.items():
-        child_dict[notations.get(kkk, kkk)] = vvv
-    return child_dict
+    try:
+        from DAS.extensions.das_speed_utils import _dict_handler
+        return _dict_handler(idict, notations)
+    except:
+        child_dict = {}
+        for kkk, vvv in idict.items():
+            child_dict[notations.get(kkk, kkk)] = vvv
+        return child_dict
 
-#def xml_parser(source, tag, add=None):
 def xml_parser(notations, source, tag, add=None):
     """
     XML parser based on ElementTree module. To reduce memory footprint for
@@ -724,15 +727,11 @@ def xml_parser(notations, source, tag, add=None):
             elem.clear()
             continue
         key = notations.get(elem.tag, elem.tag)
-        row[key] = dict_helper(notations, elem.attrib)
-#        key = elem.tag
-#        row[key] = dict(elem.attrib)
+        row[key] = dict_helper(elem.attrib, notations)
         row.update(sup)
         for child in elem.getchildren():
             child_key  = notations.get(child.tag, child.tag)
-            child_dict = dict_helper(notations, child.attrib)
-#            child_key  = child.tag
-#            child_dict = dict(child.attrib)
+            child_dict = dict_helper(child.attrib, notations)
 
             if  row[key].has_key(child_key):
                 val = row[key][child_key]

### 2010-05-04 doc/paper.tex
Dave fixed the abstract
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@683 4525493e-7705-40b1-a816-d608a930855b

@@
 \begin{abstract}
-Meta-data plays significant role in large modern enterprises, 
-research experiments and digital libraries where it comes from a different 
-sources and distributed in a variety of forms and digital formats. 
+Metadata plays a significant role in large modern enterprises, 
+research experiments and digital libraries where it comes from many different 
+sources and is distributed in a variety of digital formats. 
 It is organized and managed by constantly evolving software using 
 both relational and non-relational data sources. Even though we can apply
-information retrieval approach to non-relation data sources,
+an information retrieval approach to non-relation data sources,
 we can't do so for relational ones, where information is accessed via
-pre-established set of data-services.
+a pre-established set of data-services.
 
@@
 non-relational data sources to answer cross data-service queries 
-and explore meta-data associated with petabytes of experimental data. 
-We combine simplicity of keyword-based search with precision of RDMS
-under the new system. The aggregated information being collected from various sources,
+and explore metadata associated with petabytes of experimental data. 
+We combine the simplicity of keyword-based search with the precision of RDMS
+under the new system. The aggregated information is collected from various sources,
 allowing end-users to place dynamic queries, get precise answers and 
-trigger information retrieval on demand. Such close to real-time system 
-based on studies  of use cases of the CMS particle physics experiment at 
-the LHC where we used a large scale, distributed computing environment 
-to motivate this work.
+trigger information retrieval on demand. Based on the use cases of the CMS experiment, 
+we have performed a set of detailed, large scale tests the results of which 
+we present in this paper.
 \end{abstract}

### 2010-05-04 src/python/DAS/core/qlparser.py
Use adjust_value to do type conversion
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@686 4525493e-7705-40b1-a816-d608a930855b

@@ -9,8 +9,8 @@ tests integrity of DAS-QL queries, conversion routine from DAS-QL
 syntax to MongoDB one.
 """
 
-__revision__ = "$Id: qlparser.py,v 1.30 2009/12/01 15:15:26 valya Exp $"
-__version__ = "$Revision: 1.30 $"
+__revision__ = "$Id: qlparser.py,v 1.31 2009/12/20 15:38:07 valya Exp $"
+__version__ = "$Revision: 1.31 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -21,7 +21,7 @@ import traceback
 
 from itertools import groupby
 from DAS.utils.utils import oneway_permutations, unique_list, add2dict
-from DAS.utils.utils import getarg, genkey
+from DAS.utils.utils import getarg, genkey, adjust_value
 
 import DAS.utils.jsonwrapper as json
 
@@ -350,6 +350,7 @@ class MongoParser(object):
                 if  not value:
                     continue
                 key = prev_word
+                value = adjust_value(value)
                 for system in self.map.list_systems():
                     try:
                         lkeys = self.map.lookup_keys(system, key, 

### 2010-05-04 doc/paper.tex
Work on section 4.6 and 4.7
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@712 4525493e-7705-40b1-a816-d608a930855b

@@
 The European Organization for Nuclear Research, known as CERN, plays a leading
-role in fundamental studies of physics. But it is also known as a place where
+role in fundamental studies of physics. It is also known as a place where
 many innovations in the area of computer science were developed, e.g. World Wide Web.
@@
 RDMS systems, document oriented databases, blogs, twikies, file systems,
-customized applications, etc. 
-Working in such environment requires a lot of expertise and our users, physicists, 
-are always looking for simple, intuitive and flexible
-tool(s) to look-up their desired data. A well-known solutions, such as data-services
-and search engines are tighten to a specific data sources and end-users are left 
-with a manual task to bookkeep and relate information from them.
+customized applications, etc. Working in such environment requires a lot 
+of expertise and easy way to look-up desired information is always a necessity.
+A well-known solutions, e.g. data-services, are tighten to a specific data 
+sources and end-users are left with a manual task to bookkeep and 
+relate information from them.
 
-Here we present a work on Data Aggregation System (DAS) designed for
-CMS High-Energy Experiment at LHC which provides
-ability to query, search and aggregate information from different 
-data-services and trigger information retrieval on demand.
+Here we present a work on Data Aggregation System (DAS) which provides
+ability to search and aggregate information across different 
+data-services. While it was designed for CMS High-Energy Physics 
+experiment at LHC, it can be applied elsewhere.
 
@@
 keyword search over relational data sources.
-Section \ref{DataModel} describes CMS experiment and its data model. In section
+Section \ref{DataModel} describes the CMS experiment and its data model. In section
 \ref{DAS} we present architecture of the DAS system, including discussion of its
@@
 been discussed. It requires that all keywords are matched in a row from a
-single table or joined tables. Authors of \cite{QueryAnswer} made a step forward
-by considering information around matched values and presents customized answers to
-end-users over the generated new database. These and other related 
-works, see \cite{DBXplorer, QueryAnswer} for more references,
-are based on some kind of the graph over the underlying database and
-use a generated database out of pristine one to perform the work.
-The proximity of the results in those techniques 
-are mostly confined to a text-based values stored in a database, since 
-usage of numerical values in the input search field is mostly useless without 
-knowledge of the context of the value. For example, typing a value 100 
-in the search input field leads to plenty of inappropriate matches, 
-such as row ids, and in order to be
-answered correctly requires additional keyword, clarifying its meaning, 
-which should be matched with particular
-entity (table.column) in the underlying database.
-Therefore to answer precisely the logical questions, such as  
+single table or joined tables. Authors of \cite{QueryAnswer} are made a step forward
+and evaluated information around matched values. The results were presented as
+customized templates to end-users. 
+These and other related works, see \cite{DBXplorer, QueryAnswer} 
+for more references, are based on some kind of the graph over the underlying 
+database and generated a new database out of pristine one to perform the work.
+The proximity of the results in those techniques are mostly confined to a 
+text-based values stored in a database, since the usage of numerical ones 
+is mostly useless without knowledge of the context of the input value. 
+For example, typing a value 100 in the search input field can lead to 
+plenty of inappropriate matches, e.g. row ids. In order to be
+answered correctly it requires additional keyword, clarifying its meaning, 
+which should be matched with particular entity (table.column) in the 
+underlying database. Therefore the logical questions, e.g.
 {\it find me total number of files whose size more then 20 but less then 
-100} would be hard or almost impossible in \cite{DBXplorer, QueryAnswer}, 
-while it can be easily accomplished using SQL syntax.
+100} are hard to answer precisely using approaches outlined in \cite{DBXplorer, QueryAnswer}, 
+while it can be easily accomplished using SQL.
 
-To address those limitations an attempt to build a simple, intuitive and 
-flexible query language was introduced in \cite{DBS-QL}.
-It represents a power of SQL while hiding underlying relational schema from 
-the end-users. As a results a human questions were intuitively mapped into 
-simple queries. For instance, the question
+To address those limitations a keyword-based query language was introduced in \cite{DBS-QL}.
+It used a set of pre-defined keys mapped to underlying schema and shortest path
+algorithm which employed a foreign key constrains to build SQL queries. As a result
+the question
 {\it I'm looking for files who contain data taken on certain date and located at
@@
 \end{verbatim}
-This solution represents a common use cases within HEP community and provide
-intuitive mapping between mental model of end-users and underlying database schema.
+without specifying any relational details.
+This solution represents a common use case in HEP community and provides
+an intuitive mapping between mental model of the end-users and 
+underlying database schema.
 
 At the same time the data aggregation is commonly addressed in the context of
-federated databases, see for example \cite{FedDB}. It unifies data coming 
-from different RDMS systems into the federated DB where SQL queries can be used to 
-search desired data. Even though it would be interesting to apply technique 
-discussed in \cite{DBS-QL} to a federated DB it still requires 
-an access to the underlying database schema. This obviously creates a problem
-in distributed, heterogeneous environment where different security policies
-are in place.
+federated databases, see for example \cite{FedDB}. The data from different 
+RDMS systems are stored into the federated DB where SQL queries can be used to 
+search desired data. Apart from many internal obstacles, e.g.
+schema, domain and naming conflicts, the external ones are often
+a barrier to work with in distributed, heterogeneous environment where
+different security policies are in place. At the end an usage of SQL and/or
+any other Query Language will be required to search for data.
+%Even though it would be interesting to apply technique 
+%discussed in \cite{DBS-QL} to a federated DB it still requires 
+%an access to the underlying database schema.
 
 To avoid those limitations we decided to build a keyword-search based system around
-existing data-services whose nature and access policies are defined by developers.
-Keeping in mind that meta-data can come in different forms and data formats,
-we used a document-oriented database back-end with full support of the query language.
-The key ingredients were the setup of the mapping and analytics services which provide
-necessary glue across existing API/notations and user queries. This approach 
+existing data-services whose nature and access policies are intact.
+%The information retrieval were triggered on demand and stored into the cache.
+The key ingredients were the setup of the mapping and analytics services which provided
+a necessary glue across existing API/notations and user queries. This approach 
 allows us to perform dynamic data aggregation across various data-services, achieve
@@
 The CMS, Compact Muon Solenoid, \cite{CMS} 
-is one of the two large general-purpose particle physics detectors built on 
+is one of the two particle physics detectors built on 
 the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
 It is designed to explore the frontier of the HEP and provide physicists
-the ability to look at conditions presented at the early stage of our Universe.
-More then 3000 physicists from 183 institutions representing almost 
-38 countries are involved in the design, construction and maintenance of the experiment.
+the ability to look at the conditions presented at the early stage of our Universe.
+More then 3000 physicists from 183 institutions representing 38 countries 
+are involved in the design, construction and maintenance of the experiment.
 
@@
 interconnected via high throughput networks and operated by means 
-of Grid software. The model allows to cover broad variety of
+of the Grid software. The model allows to cover broad variety of
 hardware, mass and storage elements and configuration of the
@@
 a multi-Tier distributed model, where specific tasks of data taking,
-processing, archival and distribution are assigned to each tier based
+e.g. processing, archival and distribution, are assigned to each tier based
 on CMS data model. For example, the Tier-0 center at CERN is responsible
@@
 
-Once such conglomerate of data-services starts operating an obvious
-question arise: how to find out desired information across multiple data-services
+Once such conglomerate of data-services starts operating the obvious
+question arises: how to find out desired information across multiple data-services
 in our distributed environment? Even though individual data-services were designed
@@
 \section{Data Aggregation System\label{DAS}}
-The goal of the Data Aggregation System was to combine and
-aggregate meta-data from existing data-services
-under one umbrella, utilizing their access rules, APIs, security policies, etc., and
-allowing users to perform queries against them. In contrast, the approaches
+The goal of the Data Aggregation System was to fetch and aggregate meta-data 
+on demand from existing CMS data-services under one umbrella by using their APIs, 
+security policies, etc.
+In contrast, the approaches
 discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
 enforced a publication step of data into a dedicated database 
-or central repositories \cite{iRODS}. Therefore, quite often it is required
-that data should be converted into uniform data formats in order enable
-aggregation, see \cite{OpenArchive}. In DAS architecture the data are provided
+or central repositories \cite{iRODS}. Therefore, quite often it was required
+that data should be converted into common data structure/format in order to enable
+aggregation, see \cite{OpenArchive}. In DAS the data are provided
 by existing data-services and infrastructure, whose implementation,
@@
 DAS does not require a data preservation and transaction capabilities. 
-Instead DAS can be treated
-as proxy or caching layer where information retrieval are triggered dynamically
-by user queries. Such infrastructure allows data sharing across different queries and
-data aggregation using appropriate mapping. Below we discuss in details the
-DAS architecture.
+Instead DAS can be treated as proxy or caching layer where information 
+retrieval and aggregation are triggered dynamically by user queries. 
+Below we discuss in details the DAS architecture.
 
 \subsection{DAS architecture}
-The design of the DAS system was based on previous studies of the CMS Data 
+The design of the DAS system was based on previous studies of CMS Data 
 Bookkeeping System (DBS) \cite{DBS, DBS07}. The DBS Data Discovery service
@@
 of the system to successful information discovery. 
-Starting from this ground we designed DAS system as an
+Starting from this ground we are designed DAS system as an
 additional layer on top of the existing data-services
@@
 \caption{
-The DAS architecture diagram. 
-Input query were placed into analytics DB and mapped into a set of
-data-service APIs and their parameters. If superset of the query has been
-found in DAS cache, the appropriate results were delivered back to the users.
-Otherwise a data-service APIs were invoked, triggering DAS cache population. 
-Once data become available and aggregated in DAS cache, users were able 
-to see the results. This on demand system was supplemented by DAS robots 
-(Unix daemons) who were consulting DAS analytics database to enable data 
-population into DAS cache for most common query requests.
+The DAS architecture diagram. It conists of cacher server,
+analytics and mapping databases. The data in a cache can be shared
+across multiple queries. The information retrieval is triggered
+on demand basis by invoking appropriate data-service APIs. The
+optional DAS robots (Unix daemons) can be used to pre-fetch
+the most populary query requests.
 }
@@
 \noindent
-Based on this set we rule out the choice of RDMS as DAS back-end(s)
-for several reasons. We agreed to support dynamic queries, data
-aggregation on demand and pluggable data-services. Therefore a choice 
-of concrete schema for data storage was problematic. Moreover we didn't impose
-any transactions and data persistence in DAS. This allows us to look 
-for alternative IT solutions.
-Through the analysis of available options, such as a file-based and memory caches, 
+The choice of existing RDMS as DAS back-end(s) was ruled out for several reasons. 
+For instance, we didn't require any transactions and data persistency in DAS;
+a dynamically typing of stored meta-data objects was one of the requirements, etc.
+Those arguments forces us to look for alternative IT solutions.
+Through the careful analysis of available options, such as a file-based and in memory caches, 
 key-value databases, documented-oriented databases we made our choice in favor 
 of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
-MongoDB \cite{MongoDB}. Both systems provide schema free document
-storage, replication and fail-over features, but our choice in favor of 
-MongoDB was obvious, due to its support of dynamic queries, 
+MongoDB \cite{MongoDB}. Both systems are ``schemaless'' databases providing
+arbitrary document structure storage, replication and fail-over features. 
+We made our decision in favor of MongoDB, due to its support of dynamic queries, 
 full indexes, including inner objects and embedded arrays,
@@
 the desired load and size of our meta-data information. We used 
-document-oriented database, MongoDB, as a back-end for the three 
+document-oriented database, MongoDB, as a back-end for three 
 main DAS components: Mapping, Analytics and DAS cache databases, 
@@
 the core library with support of pluggable modules for data retrieval;
-caching layer to store data-service output and aggregate results;
+caching layer to store and aggregate results;
 request service to handle user queries;
-Mapping DB to keep information about data-service APIs, their
-notations and mapping to DAS;
-Analytics DB for query analysis and identification of pre-fetch 
-strategies.
+Mapping DB to keep information about data-service APIs, e.g.
+notations and mapping to DAS keys;
+Analytics DB for query analysis.
 
-\subsection{DAS core}
-The DAS core communicated with different data-services and retrieved
-data on demand. This was done by invoking appropriate data-service plug-in(s)
-upon the user request. Each plug-in consists of DAS map, the
-data parser(s) to convert data-service output into DAS records
-and it was responsible for data-service API calls.
-The data-service plug-in contains a map about data-service APIs,
-input and output parameters and their mapping to DAS keys. For example
+\subsection{DAS core\label{DAS_core}}
+The DAS core library was responsible for data retrieval on demand. 
+It had the following components: input query validator and parser,
+query analyzer, workflow builder, execution engine and data-service 
+plug-ins. The communication between client (web server or CLI) and 
+DAS cache server was done via the REST \cite{REST} interface. 
+%For each GET request we performed the following action.
+Each input query was recorded into Analytics DB for further 
+analysis. It was decomposed into the set of conditions and/or selection keys.
+First the results were looked up in a cache and delivered to the requestor
+if superset of input conditions was found in analytics DB.\footnote{It is important
+to note here that we didn't look-up results in a cache and rather
+compared input query condition set to the records in analytics DB 
+in order to understand if other API calls are covered the input request.}
+Otherwise, the appropriate workflow, a set of API calls, was constructed
+for the execution engine. It invoked the API calls via data-service
+plug-ins and stored results into the cache. At this step we also
+performed necessary data aggregation for records with similar key-value
+pairs. The data-service plug-in implementation were similar and their 
+logic was abstracted, while a few extentions were required, e.g. specific 
+security modules for data-services accessible via private network.
+Finally, the results were retrieved from the cache and delivered to the user.
+
+\subsection{DAS Mapping DB}
+The Mapping DB represents DAS {\it registration} service. It collects
+information about data-service APIs, translates API input/output
+parameters into DAS keys and contains mapping between DAS records
+and their UI representation.
+
+%The Mapping DB hold the mapping between data-service APIs and DAS keys exposed to
+%the users. It created a necessary glue between data-services and identifies
+%how their data can be related to each other. Basically Mapping DB was positioned 
+%as a {\it registration} service with ability to translate notations among data-services.
+Upon identification of APIs which participate in DAS 
+we stored their names, input parameters and associated DAS keys
+into the Mapping DB. For example
 \begin{verbatim}
@@
 \end{verbatim}
-This information was stored into DAS Mapping DB which we discuss next.
+represents DAS mapping record.\footnote{
+The data-service URLs were kept in the configuration file.
+}
+This record can be read as following:
+the \verb+CMSNametoSE+ API provided by 
+\verb+sitedb+ data-service accepts \verb+name+ as input parameter. It
+confirms regular expression pattern, \verb+re.compile('^T[0-3]_')+, and
+corresponds to \verb+site+ DAS key. The \verb+site+ key
+maps into \verb+site.name+ within DAS record.
 
-Although the majority of data-service plug-ins looks alike
-and their logic was abstracted, including parsers, there were a few
-cases where special care was required, e.g. access to data-services 
-accessible via private network. Therefore, a way to propagate 
-user credentials was required, leading to specific plug-in implementation.
-
-%All user queries were written to Analytics DB for further analysis, while
-%mapping between data services were kept in Mapping DB. The data-service 
-%output were re-mapped into DAS notations and, if necessary, aggregated 
-%inside of the DAS cache.
-
-\subsection{DAS Mapping DB}
-The Mapping DB hold the mapping between data-service APIs and DAS keys exposed to
-the users. It created a necessary glue between data-services and identifies
-how their data can be related to each other. Basically Mapping DB was positioned 
-as a {\it registration} service with ability to translate notations among data-services.
-Upon identification of APIs which will participate in DAS activity 
-we stored their names, input parameters, together with their type and accept 
-values, as regular expression patterns, into the Mapping DB. 
-Also we stored the mapping between data-service notations and DAS keys. 
-For instance, the CMS DBS system \cite{DBS}
-used a \verb+logical_file_name+ notation to identify the name of the
-logical file, while the same entity was called \verb+lfn+ 
-in CMS PhEDEx system \cite{PhEDEx}.
-Since both of them were referred to the logical file name, 
-we map both notations into DAS key \verb+file+.
-%\footnote{We
-%also deal with physical file names, but still used {\it file} to refer to them in
-%DAS notations, since simple regular expressions were used to identify the
-%meaning of the {\it file} keyword from the provided value.}
-Such keys, e.g. \verb+file+, were exposed to the end-users as pre-defined keywords to
-be used in search queries. For instance, end-users were able to type \verb+file+ in
-order to get all file names known to DAS via data-services as well as
-provide a concrete patterns associated with a file, e.g. \verb+file = abc*+.
-Under the hood, the appropriate data-service APIs were looked up along with appropriate
-parameters to answer those questions. In aforementioned example, 
-usage of \verb+file+ in DAS query cause a call of DBS 
-\verb+listFiles+ and PhEDEx \verb+fileReplicas+ APIs, respectively.
-This approach gave ability to end-users to express the meaning of their queries, especially
-when they deal with numbers, and achieve accuracy of results. 
-For instance, users were able to type \verb+run > 10 and run < 100+
-to specify a range of run numbers to be looked up in DAS cache. At the same time
-a full keyword search is under development.
-% within MongoDB.
+Therefore the Mapping DB identifies a set of pre-defined DAS keys, 
+e.g. \verb+site, file, run+, etc., to be used by end-users in their queries.
+Each key was mapped into set of APIs whose output was aggregated afterwards.
 
 \subsection{DAS Analytics DB}
-The DAS Analytics DB played a special role. It collected information
+The DAS Analytics DB collected information
 about user requests placed to the system. Each request was recorded. Upon its
 decomposition into set of selection keys and conditions we also recorded
-API name and input parameters into the Analytics database and 
-keep updating their counter information for repeated calls. 
-This allows to keep track of frequency of API calls and establish 
+API name and input parameters into the Analytics DB and 
+kept updating their counter information for repeated calls. 
+This allowed to keep track of frequency of API calls and establish 
 pre-fetch strategies for most common queries.
@@
 and aggregate data upon user requests. We used the following 
-algorithm:
+algorithm (code snippets are written in Python \cite{Python}):
 \begin{enumerate}[1.]
 \item Decompose query into set of conditions (\verb+conds+) 
-and selection keys (\verb+skeys+)
-\item Check cache for query hash, if found, look-up results from the cache
+and selection keys (\verb+skeys+).
+\item Check cache for query hash. If found, look-up results from the cache,
+otherwise proceed to the next step.
 \item Identify list of APIs which can serve requested data
 \begin{verbatim}
-apilist = [r(url, api, args) for r in mapping_db(skeys, conds)]
+apilist = [r for r in mapping_db(skeys, conds)]
 \end{verbatim}
-Here \verb+mappind_db+ defines translation between input query and data-service
-APIs which can delivery such data.
+Here \verb+r+ represents a triplet \verb+url, api, args+ and 
+\verb+mappind_db+ defines translation between input query and data-service
+APIs which can delivery the data.
 \item Iterate over APIs
@@
 \end{verbatim}
-Here \verb+superset_in_analytics+ verifies if provided \verb+api, args+
+Here \verb+superset_in_analytics+ checks if provided \verb+api, args+
 are subset of already registered in Analytics DB api call(s). The 
@@
 \begin{verbatim}
-if cache_has(row.primary_key):
-   existing_record = get_from_cache(row.primary_key)
+if cache_has(daskey):
+   existing_record = get_from_cache(daskey)
    merge(row, existing_record)
@@
 \end{verbatim}
-The \verb+primary_key+
-is a key which identify the record defined in Mapping DB, e.g. \verb+file.name+
+The \verb+daskey+
+is a key which identifies the record defined in Mapping DB, e.g. \verb+file.name+
 for \verb+file+ record.
@@
 \end{enumerate}
-%The data was stored in a form of DAS records into the cache, while
-%queries were recorded into Analytics DB for further analysis.
 We also had an option to run independent set of daemons, DAS robots, who
 monitor Analytics DB for most popular requests and pre-fetch the most
-common data into DAS cache. 
-%The interesting part was being able
-%to identify the superset of api/args out of found one in an Analytics DB.
-
-%The DAS caching system was used to store data-service results and aggregated
-%information in form of DAS records. As it shown in figure \ref{DAS_cache} 
-%the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
-%and aggregate data over there. The independent set of daemons, DAS robots, were
-%constantly monitor Analytics DB for most popular requests and pre-fetch most
-%common data into DAS cache.
-
-%Each user request placed to the system was parsed and set of selection
-%and condition keys were identified. They were mapped into set of
-%data service APIs and their input parameters. DAS
-%cache was checked for a queries with a similar set of selection and condition
-%keys used earlier.\footnote{Since DAS didn't hold
-%any data by default we cannot rely on data being present in a cache.
-%Moreover if we look-up data first, we cannot tell if it was complete
-%set or not, since earlier queries may retrieve only portion of the requested data.}
-%If the superset was found a data were
-%retrieved from the cache, otherwise a data-service APIs were
-%invoked and the results were collected and aggregated inside of DAS cache.
+common data into the cache.
 
 \subsection{DAS queries}
-Apart from the previous DBS-QL syntax we decided to drop its syntax and allow users
-express their queries in a free text-based form. Although we must clarify its
-meaning. We used a pre-defined set of keys registered in DAS to describe common
-entities of our data. For example, \verb+dataset+, 
+All DAS queries were expressed in a free text-based form, either as a 
+set of keywords or key-value pairs, where a pair can represent
+condition, e.g. \verb+key > value+. We used a pre-defined set of keys registered 
+in DAS to describe common entities of our data. For example, \verb+dataset+, 
 \verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
-was based on analysis of daily work operations of our users, see \cite{DBS07}.
-Each key had a set of attributes with primary one assigned to the key itself.
-For instance, the \verb+block+ key had the following attributes
-\verb+name+, \verb+size+, \verb+replica+ etc., 
-with primary attribute being \verb+name+.
-The choice of attributes were based on a set of meta-data information
-associated with a given DAS key. Therefore end-users where able to specify
-a context in their queries by using appropriate attribute(s). For example,
-for query \verb+block+ we didn't perform a string matching to the
-``block'' string, but rather look-up a block objects in a DAS, while query
-\verb+block.name=abc*+ instructs system to look-up appropriate blocks
-with matched names. At the same time, to complement the strength of the 
-DAS, we do plan to index dynamically retrieved meta-data and add a keyword-based 
+was based on the analysis of daily work operations of our users, see \cite{DBS07}.
+Each key represents a DAS record. Every record was stored into the cache
+in JSON format. Due to the ``schemaless'' nature of underlying back-end (MongoDB)
+we were able to store arbitrary structure of DAS records, e.g.
+dictionary, lists, key-value pairs, etc. Therefore every DAS key had a 
+set of attributes representing JSON structure. For instance a document
+\begin{verbatim}
+{'block': {'name':'abc', 'size': 1, 'replica': [...]}}
+\end{verbatim}
+associated with a \verb+block+ key and therefore had the following attributes
+\verb+name+, \verb+size+, \verb+replica+. The set of attributes were based 
+on a set of meta-data information provided by data-services. Our users 
+where able to specify a context in their queries by using appropriate 
+attribute(s). For example, for query \verb+block+ we didn't perform a string 
+matching to the ``block'' string, but rather look-up a block records, while query
+\verb+block.name=abc*+ instructs the system to look-up appropriate blocks
+with matched names. The numerical values, conditions, groupping were
+easy to express too. For example, \verb+run > 10 and run < 100+ query 
+was used to get all run records for provided range of run numbers. 
+At the same time, to complement the strength of the queries, we do plan 
+to index dynamically DAS records and add a keyword-based 
 search later.
 
-The free text-based queries were transformed into MongoDB query syntax, 
+All input queries were transformed into MongoDB query syntax, 
 who was used internally in DAS. It represents a set of selection 
 keys and conditions as a JSON document. For example, a query
-{\it dataset site=T1\_CH\_CERN}
+\verb+dataset file=abc run>10+
 was transformed into 
-{\it \{"fields" : ["dataset"], "spec" : \{"site" : "T1\_CH\_CERN"\}\}}.
-Such syntax was perfect fit for us, since it allows to store all queries
-as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
-MongoDB query syntax is very reach. It supports advanced queries with 
-boolean expressions, conditional operators, regular expressions, etc. providing
-a real power for our end-users.
+\begin{verbatim}
+{"fields": {"dataset"], "spec": {"file": "abc", "run": {"$gt": 10}}}
+\end{verbatim}
+Therefore users queries were stored into Analytics DB similar to other DAS 
+records and used for further analysis.
+%Such syntax was perfect fit for us, since it allows to store all queries
+%as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
+%MongoDB query syntax is very reach. It supports advanced queries with 
+%boolean expressions, conditional operators, regular expressions, etc. providing
+%a real power for our end-users.
 
 \subsection{Data-services and aggregation}
-As we outlined above the DAS Mapping DB becomes an authorative
+As we outlined above the DAS Mapping DB became an authorative
 source of information about all data-services participating in DAS.
-Based on information stored in Mapping DB we were able to 
+Based on the information stored in Mapping DB we were able to 
 retrieve appropriate data from different data-services, re-map them into 
 DAS notations and aggregate them on demand. For instance, the Data 
-Bookkeeping System (DBS) and data location system (PhEDEx)
+Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
 provided information about CMS block meta-data. In former case, DBS 
@@
 information about its location and replicas. Upon a user request to 
-find a block meta-data, we were able to identify appropriate set of DBS and PhEDEx APIs, 
-retrieve data from them, re-format their output to DAS notations 
-and store results into DAS cache performing aggregation of matched 
-objects using \verb+block.name+ key. The merged results were stored 
-as a set of data records. Here is an example of such aggregated record:
+find a block meta-data, we were able to identify appropriate set of 
+DBS and PhEDEx APIs, retrieve data from them, re-format their output 
+into DAS notations and store results into DAS cache. The records
+were merged (aggregated) based on identical key-value pairs, e.g.
+based on \verb+block.name+ key. Here is an example of such aggregated record:
 \begin{verbatim}
@@
 \end{verbatim}
-Here \verb+das_id+ refers to api call records stored in Analytics DB,
+The \verb+das_id+ refers to api call records stored in Analytics DB,
 \verb+block+ represents an aggregation block meta-data information from DBS
@@
 Usage of this information was valuable contribution in debugging process of
-data-services themselves, e.g. identification of not-matched records, 
+data-services themselves, e.g. identification of data-service inconsistencies, 
 latency studies, etc.
-% Here is an example of DAS record representing
-%aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
-%and {\it CMStoSiteName}, with API results merged under the {\it site} key.
-%\begin{verbatim}
-%{"site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
-%         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
-% "das": [{"ctime": 0.9, "expire": 1256362175, "api": ["CMStoSAMName"],
-%          "url": "https://a.b.com/sitedb/json/index/CMStoSAMName", 
-%          "timestamp": 1256318975, "system": ["sitedb"]},
-%         {"ctime": 1.2, "api": ["CMStoSiteName"], "expire": 1256362174},
-%          "url": "https://a.b.com/sitedb/json/index/CMStoSiteName", 
-%          "timestamp": 1256318974, "system": ["sitedb"]]}
-%\end{verbatim}
-
-%We used RESTful model \cite{REST} for all DAS cache server APIs.
-%It allows us to share core library and other components during development of 
-%various DAS clients, e.g. CLI, web interface, robots, etc. and hides
-%complexity of access to individual data-services and their security policies
-%as well as proper organize caching solution.
-%For instance, to access on-line Run Summary information we were forced to use
-%Grid certificates. Instead of propagating user certificate to data-service
-%the DAS cache server works as a proxy to this server and used host certificate 
-%to serve such requests on behalf of the users.
 
@@
 
-%\bibitem{REST}
-%Fielding, Roy Thomas ``Architectural Styles and the Design of 
-%Network-based Software Architectures'', Doctoral dissertation, 2000,
-%University of California, Irvine
+\bibitem{REST}
+Fielding, Roy Thomas ``Architectural Styles and the Design of 
+Network-based Software Architectures'', Doctoral dissertation, 2000,
+University of California, Irvine
+
+\bibitem{Python} Python programming language, http://www.python.org
 
@@
 
-\bibitem{Python} Python programming language, http://www.python.org
-
 %\bibitem{DBSearch}

### 2010-05-04 doc/paper.tex
Finish DAS paper, work on results/summary sections
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@713 4525493e-7705-40b1-a816-d608a930855b

@@
 \item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
-CMS file transfer system;
-\item SiteDB \cite{SiteDB}, the CMS site resources data-service;
+CMS data location and file transfer system;
+\item SiteDB \cite{SiteDB}, the CMS site resources;
 \item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
@@
 \end{itemize}
-Each data-service has its own scope, size and peculiar properties. 
-For instance, the data in PhEDEx are transient due to constant 
-migration of CMS data; the DBS system were divided into dozen 
-of individual instances with own size and lifetime, who has been used by different production 
-teams for various bookkeeping tasks; the RunSummary data-service
-operates on private network accessible at CMS online operation center.
+Each data-service had its own scope, size and peculiar properties. 
+For instance, the data in PhEDEx were transient due to constant 
+migration of CMS data; the DBS system was divided into dozen 
+of individual services (instances) with own size and lifetime 
+used by different production teams for various bookkeeping tasks; 
+the RunSummary data-service operated on private network accessible 
+at CMS online operation center.
 %Table \ref{db_size} shows the current size of meta-data from aforementioned 
@@
 %\end{table*}
-At the moment, the amount of meta-data stored in all services is equal to
-$\sim$75 GB and $\sim$200 GB in tables and index sizes, respectively. While 
-we estimate to accumulate $\sim$500 GB of the meta-data each year during 
+At the moment, the amount of meta-data stored in all services is around
+75 GB and 200 GB in tables and index sizes, respectively. Although we estimate 
+$\sim$500 GB of the meta-data each year during 
 normal CMS operation at LHC.
-%\footnote{This number is based on tests 
-%performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
 %In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
@@
 
-To test DAS performance we measured elapsed time and CPU/RAM utilization
+To test DAS performance we were measured elapsed time and CPU/RAM utilization
 required to collect and aggregate information from different
@@
 were calculated as number of aggregated records plus left overs,
-the non-matched records from both systems. The look-up time is presented in
-last column, where for DAS it shows as query look-up time (2.05 sec) 
-and time to get all records to disk (50.7 sec).}
+the non-matched records from both systems. The look-up time shown in
+last column represents query look-up time (2.05 sec) and time required
+to get all records written to disk (50.7 sec).}
 \label{DAS_benchmark}
@@
 \end{itemize}
-Here the {\it retrieval time} is a time required to access data from remote data-service,
+Here {\it retrieval time} is a time required to access data from data-service,
 {\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
 is a time required to convert data-service meta-data to DAS records for every object
-we parse, {\it cache insertion} and {\it indexing time} represents time spent on DAS 
+we parse, {\it cache insertion} and {\it indexing time} represent time spent on DAS 
 cache back-end, {\it aggregation time} is required to merge objects into DAS records based
@@
 the same performance. Although the time of PhEDEx component was bigger 
-due to longer retrieval and parsing time.
-The time of PhEDEx component shown in Table \ref{DAS_benchmark} is divided 
-among retrieval and parsing, $\sim$110 sec, and merging steps, $\sim$50 sec,
-respectively. As shown in Table \ref{DAS_benchmark},
+due to longer retrieval and parsing time as well as time spent
+during aggregation step.
+%The time of PhEDEx component shown in Table \ref{DAS_benchmark} is divided 
+%among retrieval and parsing, $\sim$110 sec, and merging steps, $\sim$50 sec,
+%respectively. 
+As shown in Table \ref{DAS_benchmark},
 the look-up time spent to access individual components (last column) was quite
-reasonable, roughly 1 second for both systems. While final time to
+reasonable, roughly 1 second for both systems. While the final time to
 get DAS records on disk was about 50 seconds. This was mostly due to I/O and
@@
 
-Based on results of stress tests we measured that we can achieve throughput of
+Based on the results of stress tests we measured that we can achieve throughput of
 $\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
@@
 operations were performed on a single CPU a further improvements are expected. 
-Also, a necessary tuning can be done by using C++ driver instead of python one.
+Also, a necessary tuning can be done by using C/C++ driver instead of python one.
 
@@
 while data pre-fetching of most common queries is in our plans. We developed
-prototype in python for DAS system which is under commissioning phase right now 
-and presented performance studies in this paper. 
+prototype in python for DAS system which is under commissioning phase right now.
+The performance studies of DAS system are presented in this paper. 
 The CMS experiment is started to collect its data in December of 2009 and 

### 2010-05-04 doc/paper.tex
More tune-ups to section 4.6 and 4.7
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@714 4525493e-7705-40b1-a816-d608a930855b

@@
 
-\subsection{DAS Mapping DB}
+\subsection{DAS Mapping DB\label{MappingDB}}
 The Mapping DB represents DAS {\it registration} service. It collects
@@
 in JSON format. Due to the ``schemaless'' nature of underlying back-end (MongoDB)
-we were able to store arbitrary structure of DAS records, e.g.
+we were able to store DAS records of arbitrary structure, e.g.
 dictionary, lists, key-value pairs, etc. Therefore every DAS key had a 
@@
 \begin{verbatim}
-{'block': {'name':'abc', 'size': 1, 'replica': [...]}}
+{"block": {"name": "abc", "size": 1, "replica": [...]}}
 \end{verbatim}
-associated with a \verb+block+ key and therefore had the following attributes
-\verb+name+, \verb+size+, \verb+replica+. The set of attributes were based 
-on a set of meta-data information provided by data-services. Our users 
-where able to specify a context in their queries by using appropriate 
-attribute(s). For example, for query \verb+block+ we didn't perform a string 
-matching to the ``block'' string, but rather look-up a block records, while query
+associated with a \verb+block+ key who had the following attributes
+\verb+name+, \verb+size+, \verb+replica+. The set of attributes is based 
+on a set of meta-data information stored in a record and therefore
+a nested structures were allowed, e.g. \verb+block.replica.size+. Our users 
+where able to specify a context of their queries by using appropriate 
+attribute(s). For example, query \verb+block+ represents search for all
+block records rather then ``block'' string, while query
 \verb+block.name=abc*+ instructs the system to look-up appropriate blocks
 with matched names. The numerical values, conditions, groupping were
-easy to express too. For example, \verb+run > 10 and run < 100+ query 
+easy to express too, e.g. \verb+run > 10 and run < 100+ query 
 was used to get all run records for provided range of run numbers. 
-At the same time, to complement the strength of the queries, we do plan 
-to index dynamically DAS records and add a keyword-based 
-search later.
+To complement the strength of the queries, we do plan to index 
+dynamically DAS records and add a keyword-based 
+search within record content later.
 
-All input queries were transformed into MongoDB query syntax, 
-who was used internally in DAS. It represents a set of selection 
-keys and conditions as a JSON document. For example, a query
-\verb+dataset file=abc run>10+
-was transformed into 
+All input queries were transformed into MongoDB query syntax who is itself
+a JSON record. It contains a set of selection 
+keys (\verb+fields+) and conditions (\verb+spec+). For example, a query
+\verb+dataset, file=abc, run>10+
+was represented as
 \begin{verbatim}
-{"fields": {"dataset"], "spec": {"file": "abc", "run": {"$gt": 10}}}
+{"fields": ["dataset"], "spec": {"file": "abc", "run": {"$gt": 10}}}
 \end{verbatim}
@@
 \subsection{Data-services and aggregation}
-As we outlined above the DAS Mapping DB became an authorative
+As was shown in section \ref{MappingDB} the DAS Mapping DB became an authorative
 source of information about all data-services participating in DAS.
@@
 Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
-provided information about CMS block meta-data. In former case, DBS 
+were provided information about CMS blocks. In former case, DBS 
 stored information about blocks and their relationships to other 
@@
 find a block meta-data, we were able to identify appropriate set of 
-DBS and PhEDEx APIs, retrieve data from them, re-format their output 
+DBS and PhEDEx APIs, retrieve data, re-format their output 
 into DAS notations and store results into DAS cache. The records
@@
 \end{verbatim}
-The \verb+das_id+ refers to api call records stored in Analytics DB,
-\verb+block+ represents an aggregation block meta-data information from DBS
-and PhEDEx systems and \verb+das+ contains their expiration time stamps.
+The \verb+das_id+ part refers to api call records stored in Analytics DB,
+the \verb+block+ part represents an aggregation block meta-data information from DBS
+and PhEDEx systems and the \verb+das+ part contains their expiration time stamps.
 Usage of this information was valuable contribution in debugging process of

### 2010-05-04 src/python/DAS/utils/utils.py
Tune-up json_parser
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@717 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 General set of useful utilities used by DAS
 """
 
-__revision__ = "$Id: utils.py,v 1.52 2009/12/20 17:51:13 valya Exp $"
-__version__ = "$Revision: 1.52 $"
+__revision__ = "$Id: utils.py,v 1.53 2010/01/04 19:01:51 valya Exp $"
+__version__ = "$Revision: 1.53 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -763,8 +763,11 @@ def json_parser(source):
     data as a string object.
     """
     if  type(source) is types.InstanceType: # got data descriptor
-        jsondict = json.load(source)
-        data.close()
+        try:
+            jsondict = json.load(source)
+        except:
+            traceback.print_exc()
+        source.close()
     else:
         data = source
         # to prevent unicode/ascii errors like

### 2010-05-04 src/js/utils.js
Confirm to XHTML validator test
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@757 4525493e-7705-40b1-a816-d608a930855b

@@ -1,4 +1,24 @@
+/*
+ * DAS specific utilities
+ * Author: Valentin Kuznetsov, 2009
+ */ 
+function updateInput() {
+   var uinput = gup('input');
+   if  (uinput) updateTag('input', uinput);
+   else updateTag('input', '');
 
+   var ulimit = gup('limit');
+   if  (ulimit) updateTag('limit', ulimit);
+   else updateTag('limit', '');
+
+   var uview = gup('view');
+   if  (uview) updateTag('view', uview);
+   else updateTag('view', '');
+
+   var ushow = gup('show');
+   if  (ushow) updateTag('show', ushow);
+   else updateTag('show', '');
+}
 /*
  * General set of utilities used in DAS web interface
  * Author: Valentin Kuznetsov, 2009

### 2010-05-04 doc/paper.tex
Tune-up for Restuls/Conclusion section
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@758 4525493e-7705-40b1-a816-d608a930855b

@@
 both relational and non-relational data sources. Even though we can apply
-an information retrieval approach to non-relation data sources,
+an information retrieval approach to non-relational data sources,
 we can't do so for relational ones, where information is accessed via
@@
 role in fundamental studies of physics. It is also known as a place where
-many innovations in the area of computer science were developed, e.g. World Wide Web.
+many innovations in the area of computer science were developed, e.g. The World Wide Web.
 Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
 Physics (HEP), promising to deliver a few PB of data each year. 
-At this scale the ``information discovery'' within heterogeneous, distributed 
+At this scale the ``information discovery'' within a heterogeneous, distributed 
 environment becomes a key ingredient of successful data analysis.
@@
 RDMS systems, document oriented databases, blogs, twikies, file systems,
-customized applications, etc. Working in such environment requires a lot 
-of expertise and easy way to look-up desired information is always a necessity.
+customized applications, etc. Working in such an environment requires a lot 
+of expertise and demands a straightforward, simple
+way to look-up the desired information.
 A well-known solutions, e.g. data-services, are tighten to a specific data 
@@
 
-Here we present a work on Data Aggregation System (DAS) which provides
+Here we present our work on Data Aggregation System (DAS) which provides
 ability to search and aggregate information across different 
-data-services. While it was designed for CMS High-Energy Physics 
-experiment at LHC, it can be applied elsewhere.
+data-services. While designed for CMS High-Energy Physics 
+experiment at LHC, the strategies and technology could be
+used elsewhere.
 
-The rest of this paper is organized as following. 
+The rest of this paper is organized as follows. 
 In section \ref{RelatedWork} we discuss related work in a domain of 
@@
 Section \ref{DataModel} describes the CMS experiment and its data model. In section
-\ref{DAS} we present architecture of the DAS system, including discussion of its
+\ref{DAS} we present the architecture of the DAS system, including discussion of its
 various components. Finally, our results are summarized in section \ref{Results}.
-
-%As was pointed out in \cite{Arms} a mixed content and 
-%mixed meta-data and meta-data consistency should be considered as a whole in design 
-%of the system to successful information discovery.
 
@@
 Even though the idea of keyword queries over the relational database
-%querying relation databases via keyword based search
-%algorithms 
 is not knew it is still under discussion in computer science domain.
-A few alternative solutions has been proposed to address this issue
+A few alternative solutions have been proposed to address this issue
 in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
@@
 been discussed. It requires that all keywords are matched in a row from a
-single table or joined tables. Authors of \cite{QueryAnswer} are made a step forward
-and evaluated information around matched values. The results were presented as
+single table or joined tables. Authors of \cite{QueryAnswer} make a step forward
+and have evaluated information around matched values. The results were presented as
 customized templates to end-users. 
-These and other related works, see \cite{DBXplorer, QueryAnswer} 
-for more references, are based on some kind of the graph over the underlying 
-database and generated a new database out of pristine one to perform the work.
+These and other related works, (see \cite{DBXplorer, QueryAnswer} 
+for more references), are based on an approach that builds some 
+kind of the graph over the underlying 
+database and generates a new database out of the source database to perform the work.
 The proximity of the results in those techniques are mostly confined to a 
@@
 plenty of inappropriate matches, e.g. row ids. In order to be
-answered correctly it requires additional keyword, clarifying its meaning, 
+answered correctly it requires an additional keyword to clarify its meaning, 
 which should be matched with particular entity (table.column) in the 
 underlying database. Therefore the logical questions, e.g.
-{\it find me total number of files whose size more then 20 but less then 
+{\it find me the total number of files whose size more then 20 but less then 
 100} are hard to answer precisely using approaches outlined in \cite{DBXplorer, QueryAnswer}, 
@@
 
-To address those limitations a keyword-based query language was introduced in \cite{DBS-QL}.
-It used a set of pre-defined keys mapped to underlying schema and shortest path
-algorithm which employed a foreign key constrains to build SQL queries. As a result
+To address these limitations a keyword-based query language was introduced in \cite{DBS-QL}.
+It used a set of pre-defined keys mapped to an underlying schema and shortest path
+algorithm which employed foreign key constraints to build SQL queries. As a result
 the question
 {\it I'm looking for files who contain data taken on certain date and located at
-particular site} was represented as simple as \cite{DBS-QL}
+particular site} was represented as simply as \cite{DBS-QL}
 \begin{verbatim}
@@
 without specifying any relational details.
-This solution represents a common use case in HEP community and provides
-an intuitive mapping between mental model of the end-users and 
-underlying database schema.
+This solution represents a common use case in the HEP community and provides
+an intuitive mapping between the mental model of the end-users and 
+the underlying database schema.
 
@@
 RDMS systems are stored into the federated DB where SQL queries can be used to 
-search desired data. Apart from many internal obstacles, e.g.
-schema, domain and naming conflicts, the external ones are often
-a barrier to work with in distributed, heterogeneous environment where
-different security policies are in place. At the end an usage of SQL and/or
-any other Query Language will be required to search for data.
-%Even though it would be interesting to apply technique 
-%discussed in \cite{DBS-QL} to a federated DB it still requires 
-%an access to the underlying database schema.
+search desired data. Apart from many internal obstacles, such as
+schema, domain and naming conflicts, it is external issues that are often
+a large barrier within a distributed, heterogeneous environment where
+different security policies are in place. At the end the usage of SQL and/or
+any other Query Language is required to search for data.
 
-To avoid those limitations we decided to build a keyword-search based system around
-existing data-services whose nature and access policies are intact.
-%The information retrieval were triggered on demand and stored into the cache.
+To avoid these limitations we resolved to build a keyword-search based system around
+existing data-services whose nature and access policies are in place and well understood.
 The key ingredients were the setup of the mapping and analytics services which provided
-a necessary glue across existing API/notations and user queries. This approach 
+the necessary glue across existing API/notations and user queries. This approach 
 allows us to perform dynamic data aggregation across various data-services, achieve
@@
 and organize information retrieval on demand.
-%to store meta-data information 
-%We wanted to expand this approach further and apply it 
-%to handle cross data-service queries from broad variety 
-%of relational and non-relational data-sources within 
-%our distributed heterogeneous environment.
 
 \section{CMS data model\label{DataModel}}
-The CMS, Compact Muon Solenoid, \cite{CMS} 
-is one of the two particle physics detectors built on 
-the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
-It is designed to explore the frontier of the HEP and provide physicists
-the ability to look at the conditions presented at the early stage of our Universe.
+The Compact Muon Solenoid, (CMS) \cite{CMS} 
+is one of the two general purpose particle physics detectors built on 
+the Large Hadron Collider (LHC) at CERN in Switzerland and France. 
+The experiment is designed to explore the frontiers of physics and provide physicists
+with the ability to look at the conditions presented in the early stage of our Universe.
 More then 3000 physicists from 183 institutions representing 38 countries 
@@
 The CMS distributed computing and data model \cite{CMSDataModel} 
-is designed to process
-and efficiently manage a few PBs of data expected each year
-in CMS operation at LHC. The computing resources provided by members of CMS
+is designed to process and efficiently manage the few PBs of data expected each year
+during normal operation of the LHC. The computing resources available to the CMS
 collaboration are geographically distributed, 
 interconnected via high throughput networks and operated by means 
-of the Grid software. The model allows to cover broad variety of
-hardware, mass and storage elements and configuration of the
-clusters. To accommodate efficient data processing CMS uses
-a multi-Tier distributed model, where specific tasks of data taking,
-e.g. processing, archival and distribution, are assigned to each tier based
-on CMS data model. For example, the Tier-0 center at CERN is responsible
+of various Grid software infrastructure. The model is designed to
+incorporate a broad variety of both mass data storage and data processing
+resources into a single coherent system. CMS uses
+multi-Tiered model, where specific tasks related to data taking such as
+processing, archival and distribution, are assigned to each tier based
+on the CMS data model. For example, the Tier-0 center at CERN is responsible
 for archiving the data coming out from the detector, prompt first pass reconstruction
 and data distribution to Tier-1 centers, located around the world.
-%The 7 Tier-1 centers
-%located in France, Germany, Italy, Spain, Taiwan, United Kingdom and United States
-%keep a portion (copy) of the data delivered by Tier-0 for further processing.
-They provide storage and CPU power for high priority analysis jobs.
-The Tier-2 centers
-%, located at more then 50 sites around the world,
-are dedicated for user analysis tasks and production of simulated data.
+The Tier-1 centers provide archival storage and CPU power for high
+priority selection and re-reconstructions jobs and distribution points
+to the Tier 2 resources.
+The Tier-2 centers are dedicated for user analysis tasks and production of simulated data.
 
-A broad variety of data-services have being designed and developed to
+A broad variety of data-services have been designed and developed to
 maintain detector and production operations, including detector
-conditions databases, data location and bookkeeping services,
-data transfer and job monitoring tasks. Even though majority of them
-are located at CERN, it was never been a requirement in CMS computing
-and data model. 
-%For instance, the production teams operated at Tier-1,2
-%centers set up local services for data bookkeeping and operational
-%tasks. Based on their resources the choice of database back-end was
-%up to the site managers. Therefore the CMS software were designed to support different
-%DB back-ends and provide tools for migrating data across them.
+conditions databases, data bookkeeping services,
+data transfer and job monitoring tasks. While the majority of these
+services are located at CERN, they need to be accessed remotely, and
+combined with other distributed services.
 
-Once such conglomerate of data-services starts operating the obvious
-question arises: how to find out desired information across multiple data-services
-in our distributed environment? Even though individual data-services were designed
+Once such a conglomerate of data-services start operating the obvious
+question arises: how do we find the desired information across multiple data-services
+in our distributed environment? Even though the individual data-services were designed
 to answer specific questions about the data they serve, the ability to search and relate
-information among them was always a tedious human task. The growing amount of information
-and desire to make cross-service queries force us to design and develop new
+information among them was remains a tedious human task. The growing amount of information
+and desire to make cross-service queries led us to design and develop the new
 Data Aggregation System.
@@
 \section{Data Aggregation System\label{DAS}}
-The goal of the Data Aggregation System was to fetch and aggregate meta-data 
+The goal of the Data Aggregation System (DAS) was to fetch and aggregate meta-data 
 on demand from existing CMS data-services under one umbrella by using their APIs, 
-security policies, etc.
-In contrast, the approaches
-discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
-enforced a publication step of data into a dedicated database 
-or central repositories \cite{iRODS}. Therefore, quite often it was required
+security policies, etc. In contrast, the approaches
+discussed in \cite{DBXplorer, QueryAnswer, FedDB}
+enforce a publication step of data into a dedicated database 
+or central repository \cite{iRODS}. Therefore, quite often it was required
 that data should be converted into common data structure/format in order to enable
@@
 \cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
-a syntax similar to SQL, without specific knowledge of underlying DB schema.
-%We carefully analyzed user
-%requests and queries placed into the DBS, their patterns, frequencies and latency. 
-%As a result we created a data discovery service \cite{DD}. Originally it used MVC
-%architecture and later, being re-factored as pure presentation layer on top of 
-%the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
-%DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
-%we took away relations and details on underlying schema from end-users. 
-%This was achieved by using Dijkstra's shortest path algorithm 
-%to establish necessary table joins based on provided information in user queries and
-%knowledge of foreign key relationships from auto-loaded database schema.
-%The DBS-QL keys exposed to end-users were chosen based on a jargon used by
-%physicists. They're internally mapped into database entities and attributes, 
-%e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
-A quick adoption and wide usage of DBS-QL in CMS gave us confidence in chosen approach 
-and provide a basis for building DAS system.
+a syntax similar to SQL, without specific knowledge of the underlying DB schema.
+Quick adoption and wide usage of DBS-QL in CMS gave us confidence in the chosen approach 
+and provided a basis for building the DAS system.
 
-As was pointed out in \cite{Arms} a mixed content and 
-mixed meta-data and meta-data consistency should be considered as a whole in design 
-of the system to successful information discovery. 
-Starting from this ground we are designed DAS system as an
+As was pointed out in \cite{Arms} the mixed content and 
+mixed meta-data and meta-data consistency 
+of the underlying services needs to be carefully
+considered in the design of the system if it is to be a successful information 
+discovery service.
+Starting from this ground we are designed the DAS system as an
 additional layer on top of the existing data-services
@@
 \begin{itemize}
-\item support a keyword based search queries with ability to use condition operators;
+\item support a keyword based search queries with ability to use conditional operators;
 %\item pluggable interface to existing and up-coming data-service APIs;
@@
 \caption{
-The DAS architecture diagram. It conists of cacher server,
+The DAS architecture diagram. It conists of cache server,
 analytics and mapping databases. The data in a cache can be shared
@@
 \noindent
-The choice of existing RDMS as DAS back-end(s) was ruled out for several reasons. 
+The choice of existing RDMS as DAS back-end(s) were ruled out for several reasons. 
 For instance, we didn't require any transactions and data persistency in DAS;
 a dynamically typing of stored meta-data objects was one of the requirements, etc.
-Those arguments forces us to look for alternative IT solutions.
+Those arguments forced us to look for alternative IT solutions.
 Through the careful analysis of available options, such as a file-based and in memory caches, 
@@
 full indexes, including inner objects and embedded arrays,
-and auto-sharding. Our preliminary benchmarks shown that it can sustain
-the desired load and size of our meta-data information. We used 
+and auto-sharding. Our preliminary benchmarks have shown that it can sustain
+the desired load and size capable of handling our meta-data information. We used 
 document-oriented database, MongoDB, as a back-end for three 
@@
 \subsection{DAS core\label{DAS_core}}
-The DAS core library was responsible for data retrieval on demand. 
-It had the following components: input query validator and parser,
+The DAS core library is responsible for data retrieval on demand. 
+It has the following components: input query validator and parser,
 query analyzer, workflow builder, execution engine and data-service 
@@
 DAS cache server was done via the REST \cite{REST} interface. 
-%For each GET request we performed the following action.
-Each input query was recorded into Analytics DB for further 
-analysis. It was decomposed into the set of conditions and/or selection keys.
+Each input query was recorded into the Analytics DB for further 
+analysis, and was decomposed into a set of conditions and/or selection keys.
 First the results were looked up in a cache and delivered to the requestor
@@
 performed necessary data aggregation for records with similar key-value
-pairs. The data-service plug-in implementation were similar and their 
-logic was abstracted, while a few extentions were required, e.g. specific 
+pairs. The data-service plug-in implementations are similar and their 
+logic is largely abstracted, apart from a few extentions such as specific
 security modules for data-services accessible via private network.
@@
 \subsection{DAS Mapping DB\label{MappingDB}}
-The Mapping DB represents DAS {\it registration} service. It collects
+The Mapping DB is used to track services known to DAS and how to interact with them.
+%The Mapping DB represents DAS {\it registration} service. 
+It collects
 information about data-service APIs, translates API input/output
@@
 
-%The Mapping DB hold the mapping between data-service APIs and DAS keys exposed to
-%the users. It created a necessary glue between data-services and identifies
-%how their data can be related to each other. Basically Mapping DB was positioned 
-%as a {\it registration} service with ability to translate notations among data-services.
-Upon identification of APIs which participate in DAS 
-we stored their names, input parameters and associated DAS keys
+On identifying API's that are accessible to DAS we store
+their names, input parameters and associated DAS keys
 into the Mapping DB. For example
@@
 represents DAS mapping record.\footnote{
-The data-service URLs were kept in the configuration file.
+The data-service URL's are kept in the configuration file.
 }
@@
 e.g. \verb+site, file, run+, etc., to be used by end-users in their queries.
-Each key was mapped into set of APIs whose output was aggregated afterwards.
+Each key is mapped into set of API calls, whose output is retrieved,
+cached and aggregated as required.
 
 \subsection{DAS Analytics DB}
-The DAS Analytics DB collected information
-about user requests placed to the system. Each request was recorded. Upon its
-decomposition into set of selection keys and conditions we also recorded
-API name and input parameters into the Analytics DB and 
-kept updating their counter information for repeated calls. 
-This allowed to keep track of frequency of API calls and establish 
-pre-fetch strategies for most common queries.
+The DAS Analytics DB collects information on user requests against the system.
+Each request that DAS handles is recorded.
+The Analytics DB tracks:
+\begin{itemize}
+\item The users request
+\item How DAS has mapped the request onto data service API's
+\item The time taken by the remote data services to process the API calls
+\end{itemize}
+By recording this information we can plan pre-fetch strategies for common queries,
+identify issues in remote data services and cross check that DAS resolves requests
+in a deterministic manner.
+
+% Upon its
+%decomposition into set of selection keys and conditions we also recorded
+%API name and input parameters into the Analytics DB and 
+%kept updating their counter information for repeated calls. 
+%This allowed to keep track of frequency of API calls and establish 
+%pre-fetch strategies for most common queries.
 
 \subsection{DAS caching system}
-The DAS caching system was used to dynamically fetch
-and aggregate data upon user requests. We used the following 
+The DAS caching system is used to dynamically fetch
+and aggregate data upon user requests. We use the following 
 algorithm (code snippets are written in Python \cite{Python}):
@@
 \end{enumerate}
-We also had an option to run independent set of daemons, DAS robots, who
+We also have an option to run independent set of daemons, DAS robots, who
 monitor Analytics DB for most popular requests and pre-fetch the most
@@
 \subsection{DAS queries}
-All DAS queries were expressed in a free text-based form, either as a 
+All DAS queries are expressed in a free text-based form, either as a 
 set of keywords or key-value pairs, where a pair can represent
-condition, e.g. \verb+key > value+. We used a pre-defined set of keys registered 
+condition, e.g. \verb+key > value+. We use a pre-defined set of keys registered 
 in DAS to describe common entities of our data. For example, \verb+dataset+, 
 \verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
-was based on the analysis of daily work operations of our users, see \cite{DBS07}.
-Each key represents a DAS record. Every record was stored into the cache
+is based on the analysis of daily work operations of our users, see \cite{DBS07}.
+Each key represents a DAS record. Every record is stored into the cache
 in JSON format. Due to the ``schemaless'' nature of underlying back-end (MongoDB)
-we were able to store DAS records of arbitrary structure, e.g.
-dictionary, lists, key-value pairs, etc. Therefore every DAS key had a 
+we are able to store DAS records of arbitrary structure, e.g.
+dictionary, lists, key-value pairs, etc. Therefore every DAS key has a 
 set of attributes representing JSON structure. For instance a document
@@
 \end{verbatim}
-associated with a \verb+block+ key who had the following attributes
+is associated with a \verb+block+ key who has the following attributes
 \verb+name+, \verb+size+, \verb+replica+. The set of attributes is based 
 on a set of meta-data information stored in a record and therefore
-a nested structures were allowed, e.g. \verb+block.replica.size+. Our users 
-where able to specify a context of their queries by using appropriate 
+a nested structures are allowed, e.g. \verb+block.replica.size+. Our users 
+are able to specify a context of their queries by using appropriate 
 attribute(s). For example, query \verb+block+ represents search for all
@@
 \verb+block.name=abc*+ instructs the system to look-up appropriate blocks
-with matched names. The numerical values, conditions, groupping were
+with matched names. The numerical values, conditions, grouping are
 easy to express too, e.g. \verb+run > 10 and run < 100+ query 
-was used to get all run records for provided range of run numbers. 
+is used to get all run records for provided range of run numbers. 
 To complement the strength of the queries, we do plan to index 
@@
 
-All input queries were transformed into MongoDB query syntax who is itself
+All input queries are transformed into MongoDB query syntax who is itself
 a JSON record. It contains a set of selection 
@@
 \verb+dataset, file=abc, run>10+
-was represented as
+is represented as
 \begin{verbatim}
@@
 \end{verbatim}
-Therefore users queries were stored into Analytics DB similar to other DAS 
+Therefore users queries are stored into Analytics DB similar to other DAS 
 records and used for further analysis.
-%Such syntax was perfect fit for us, since it allows to store all queries
-%as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
-%MongoDB query syntax is very reach. It supports advanced queries with 
-%boolean expressions, conditional operators, regular expressions, etc. providing
-%a real power for our end-users.
 
@@
 Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
-were provided information about CMS blocks. In former case, DBS 
+provide information about CMS blocks. In former case, DBS 
 stored information about blocks and their relationships to other 
 physics objects, such as dataset and files, while PhEDEx kept
-information about its location and replicas. Upon a user request to 
+information about its location and replicas.
+
+Upon a user request to 
 find a block meta-data, we were able to identify appropriate set of 
@@
 were merged (aggregated) based on identical key-value pairs, e.g.
-based on \verb+block.name+ key. Here is an example of such aggregated record:
+based on \verb+block.name+ key. Here is an example of such an aggregated record:
 \begin{verbatim}
@@
 \end{verbatim}
-The \verb+das_id+ part refers to api call records stored in Analytics DB,
+The \verb+das_id+ refers to api call records stored in Analytics DB,
 the \verb+block+ part represents an aggregation block meta-data information from DBS
@@
 \end{itemize}
-Each data-service had its own scope, size and peculiar properties. 
-For instance, the data in PhEDEx were transient due to constant 
-migration of CMS data; the DBS system was divided into dozen 
+Each data-service has its own scope, size and peculiar properties. 
+For instance, the data in PhEDEx are transient due to constant 
+migration of CMS data; the DBS system is divided into dozen 
 of individual services (instances) with own size and lifetime 
 used by different production teams for various bookkeeping tasks; 
-the RunSummary data-service operated on private network accessible 
+the RunSummary data-service is operated on private network accessible 
 at CMS online operation center.
-%Table \ref{db_size} shows the current size of meta-data from aforementioned 
-%data-services.
-%\begin{figure}[htb]
-%\centering
-%\includegraphics[width=100mm]{db_size.pdf}
-%\caption{
-%Current size of meta-data for CMS data-services participated in DAS, before
-%official data-taking. 
-%}
-%\label{db_size}
-%\end{figure}
 
-%\begin{table*}[hbt]
-%\centering
-%\begin{tabular}{llllll}\hline
-%\hline
-
-%Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
-%\hline
-%tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
-%indexes (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
-%\hline
-%\hline
-%\end{tabular}
-%\caption{
-%Current size of meta-data for CMS data-services participated in DAS, before
-%official data-taking. 
-%}
-%\label{db_size}
-%\end{table*}
 At the moment, the amount of meta-data stored in all services is around
-75 GB and 200 GB in tables and index sizes, respectively. Although we estimate 
-$\sim$500 GB of the meta-data each year during 
+75 GB and 200 GB in tables and index sizes, respectively. We estimate to
+collect $\sim$500 GB of the meta-data each year during 
 normal CMS operation at LHC.
@@
 
-To test DAS performance we were measured elapsed time and CPU/RAM utilization
-required to collect and aggregate information from different
+To test DAS performance we ere measured elapsed time and CPU/RAM utilization
+required to fetch, store and aggregate information from different
 data-services. Table \ref{DAS_benchmark} summarizes the biggest 
@@
 \caption{Time required to fetch, parse and aggregate block information
-from DBS and PhEDEx systems. In the case of PhEDEx
-system we show total number of fetched records 186145 together with
-number of records merged in DAS 180439. The total number of DAS records 
-were calculated as number of aggregated records plus left overs,
-the non-matched records from both systems. The look-up time shown in
-last column represents query look-up time (2.05 sec) and time required
-to get all records written to disk (50.7 sec).}
+from DBS and PhEDEx systems. The PhEDEx
+system shows fetched (186145) and merged (180439) records. 
+The total number of DAS records are calculated as number of 
+aggregated records plus left overs, the non-matched records from both systems. 
+The look-up time shown in last column represents query look-up 
+time (2.05 sec) and time required to get all records written 
+to disk (50.7 sec).}
 \label{DAS_benchmark}
@@
 
-The initial DAS implementation was done in Python language \cite{Python}
-and all tests were performed on x64 Linux platform where
+The initial DAS implementation is done in Python language \cite{Python}
+and all tests are performed on x64 Linux platform where
 1 CPU was used by DAS server and 1 CPU by MongoDB database server. 
-%The CPU utilization
-%of DAS server was 67\% in average and RAM usage was below 35MB. 
-All tests were performed multiple times to avoid fluctuations. 
-The elapsed time measured during those tests consists of the 
-following components:
+All tests are performed multiple times to avoid fluctuations. 
+The elapsed time is measured and consists of the following components:
 \begin{itemize}
@@
 is a time required to convert data-service meta-data to DAS records for every object
-we parse, {\it cache insertion} and {\it indexing time} represent time spent on DAS 
+we parse, {\it cache insertion} and {\it indexing time} represents time spent on DAS 
 cache back-end, {\it aggregation time} is required to merge objects into DAS records based
@@
 is a time required to write DAS records to disk. Please note that last two
-components, {\it aggregation} and {\it output creation time}, were only applied to
-one of the sub-system, either DBS or PhEDEx, depending on insertion order.
-% since the first one was injected into
-%the fresh database, while the second required the merging step.
+components, {\it aggregation} and {\it output creation time}, are only applied to
+one of the sub-system, either DBS or PhEDEx, depending on the insertion order.
 
@@
 The individual tests of DAS components (DBS and PhEDEx) shown roughly 
-the same performance. Although the time of PhEDEx component was bigger 
+the same performance. Although the time of PhEDEx component is bigger 
 due to longer retrieval and parsing time as well as time spent
-during aggregation step.
-%The time of PhEDEx component shown in Table \ref{DAS_benchmark} is divided 
-%among retrieval and parsing, $\sim$110 sec, and merging steps, $\sim$50 sec,
-%respectively. 
-As shown in Table \ref{DAS_benchmark},
-the look-up time spent to access individual components (last column) was quite
+during aggregation step. As shown in Table \ref{DAS_benchmark},
+the look-up time spent to access individual components (last column) is quite
 reasonable, roughly 1 second for both systems. While the final time to
-get DAS records on disk was about 50 seconds. This was mostly due to I/O and
+get DAS records on disk is about 50 seconds. This is mostly due to I/O and
 conversion operations between binary data format used by MongoDB, BSON, to DAS
@@
 
-Based on the results of stress tests we measured that we can achieve throughput of
+Based on the results of stress tests we are measured that we achieve throughput of
 $\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
 $\sim$7600 docs/sec for 
-reading operations in DAS, which are suitable for our needs. Since all 
-operations were performed on a single CPU a further improvements are expected. 
+reading operations in DAS, which is suitable for our needs. Since all 
+operations are performed on a single CPU, a further improvements are expected. 
 Also, a necessary tuning can be done by using C/C++ driver instead of python one.
@@
 \section{Summary}
-We presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
-at LHC, CERN, Geneva, Switzerland. It was designed to provide caching and
+We are presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
+at LHC, CERN, Geneva, Switzerland. It is designed to provide caching and
 aggregation layer on top of the existing relational and non-relation data-services
-mostly in real time fashion. All of the data were retrieved on demand basis,
-while data pre-fetching of most common queries is in our plans. We developed
-prototype in python for DAS system which is under commissioning phase right now.
-The performance studies of DAS system are presented in this paper. 
+mostly in real time fashion. All data are retrieved on demand basis,
+while pre-fetching of most common queries is in our plans. We are developed a
+prototype in Python which is under commissioning phase right now.
+The performance studies of the DAS system are presented in this paper. 
 The CMS experiment is started to collect its data in December of 2009 and 
@@
 meta-data produced each year by experiment will be of the order of
-500GB. The DAS should sustain such load and will provide generic 
+500GB. The DAS system should sustain such load and provide generic 
 ``data-discovery'' service for CMS experiment in years to come.

### 2010-05-04 doc/paper.tex
Corrections after spell check
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@760 4525493e-7705-40b1-a816-d608a930855b

@@
 \begin{abstract}
-Metadata plays a significant role in large modern enterprises, 
+Meta-data plays a significant role in large modern enterprises, 
 research experiments and digital libraries where it comes from many different 
@@
 non-relational data sources to answer cross data-service queries 
-and explore metadata associated with petabytes of experimental data. 
+and explore meta-data associated with petabytes of experimental data. 
 We combine the simplicity of keyword-based search with the precision of RDMS
@@
 \caption{
-The DAS architecture diagram. It conists of cache server,
+The DAS architecture diagram. It consists of cache server,
 analytics and mapping databases. The data in a cache can be shared
@@
 optional DAS robots (Unix daemons) can be used to pre-fetch
-the most populary query requests.
+the most popular query requests.
 }
@@
 of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
-MongoDB \cite{MongoDB}. Both systems are ``schemaless'' databases providing
+MongoDB \cite{MongoDB}. Both systems are ``schema-less'' databases providing
 arbitrary document structure storage, replication and fail-over features. 
@@
 analysis, and was decomposed into a set of conditions and/or selection keys.
-First the results were looked up in a cache and delivered to the requestor
+First the results were looked up in a cache and delivered to the requester
 if superset of input conditions was found in analytics DB.\footnote{It is important
@@
 pairs. The data-service plug-in implementations are similar and their 
-logic is largely abstracted, apart from a few extentions such as specific
+logic is largely abstracted, apart from a few extensions such as specific
 security modules for data-services accessible via private network.
@@
 \subsection{DAS Mapping DB\label{MappingDB}}
-The Mapping DB is used to track services known to DAS and how to interact with them.
-%The Mapping DB represents DAS {\it registration} service. 
+The Mapping DB is used to track services known to DAS and how DAS should interact with them.
 It collects
@@
 e.g. \verb+site, file, run+, etc., to be used by end-users in their queries.
-Each key is mapped into set of API calls, whose output is retrieved,
+Each key is mapped onto set of API calls, whose output is retrieved,
 cached and aggregated as required.
@@
    insert_into_cache(row)
-   delete_from_cache(existing_record)
 else:
@@
 \end{enumerate}
-We also have an option to run independent set of daemons, DAS robots, who
-monitor Analytics DB for most popular requests and pre-fetch the most
-common data into the cache.
+%We also have an option to run independent set of daemons, DAS robots, who
+%monitor Analytics DB for most popular requests and pre-fetch the most
+%common data into the cache. 
+We can also automate populating the cache by running an 
+independent set of daemons, DAS robots, which monitor the 
+Analytics DB for suitable requests (for example queries that are 
+popular, periodic in nature or expensive) and pre-fetch data into the cache.
 
@@
 in DAS to describe common entities of our data. For example, \verb+dataset+, 
-\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
-is based on the analysis of daily work operations of our users, see \cite{DBS07}.
+\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. The choice of keys
+is based on the analysis of day to day user operations \cite{DBS07}.
 Each key represents a DAS record. Every record is stored into the cache
-in JSON format. Due to the ``schemaless'' nature of underlying back-end (MongoDB)
-we are able to store DAS records of arbitrary structure, e.g.
+in JSON format. Due to the ``schema-less'' nature of the underlying MongoDB
+back-end we are able to store DAS records of arbitrary structure, e.g.
 dictionary, lists, key-value pairs, etc. Therefore every DAS key has a 
-set of attributes representing JSON structure. For instance a document
+set of attributes describing its JSON structure. For instance a document
 \begin{verbatim}
@@
 \verb+name+, \verb+size+, \verb+replica+. The set of attributes is based 
-on a set of meta-data information stored in a record and therefore
-a nested structures are allowed, e.g. \verb+block.replica.size+. Our users 
-are able to specify a context of their queries by using appropriate 
-attribute(s). For example, query \verb+block+ represents search for all
-block records rather then ``block'' string, while query
+on a set of meta-data information stored in a record, allowing
+for nested structures, e.g. \verb+block.replica.size+. Our users 
+are able to specify the context of their queries by using appropriate 
+attribute(s). For example, a query for \verb+block+ represents a search for all
+block records rather then the ``block'' string, while the query
 \verb+block.name=abc*+ instructs the system to look-up appropriate blocks
-with matched names. The numerical values, conditions, grouping are
-easy to express too, e.g. \verb+run > 10 and run < 100+ query 
-is used to get all run records for provided range of run numbers. 
-To complement the strength of the queries, we do plan to index 
-dynamically DAS records and add a keyword-based 
-search within record content later.
+with matching names. Numerical values, conditions and grouping are
+also easy to express, for example, a query
+of the form \verb+run > 10 and run < 100+ can be used to
+retrieve all run records for the given range of run numbers. 
+As a complement to the queries, we will also be adding a dynamic index
+of DAS records with a keyword-based search within record content later.
 
-All input queries are transformed into MongoDB query syntax who is itself
+All input queries are transformed into the MongoDB query syntax who is itself
 a JSON record. It contains a set of selection 
 keys (\verb+fields+) and conditions (\verb+spec+). For example, a query
-\verb+dataset, file=abc, run>10+
+\verb+dataset, file=abc, run > 10+
 is represented as
@@
 \subsection{Data-services and aggregation}
-As was shown in section \ref{MappingDB} the DAS Mapping DB became an authorative
-source of information about all data-services participating in DAS.
-Based on the information stored in Mapping DB we were able to 
+As discussed in section \ref{MappingDB} the DAS Mapping DB 
+is the authoritative source of information on the relationships between 
+all the data services known to DAS.
+%became an authorative
+%source of information about all data-services participating in DAS.
+Based on the information stored in Mapping DB we are able to 
 retrieve appropriate data from different data-services, re-map them into 
@@
 Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
-provide information about CMS blocks. In former case, DBS 
-stored information about blocks and their relationships to other 
-physics objects, such as dataset and files, while PhEDEx kept
-information about its location and replicas.
+provide information about CMS file block structure and locations.
+DBS associates groups of files into blocks, while PhEDEx tracks the blocks
+location. 
+%In the former case, DBS 
+%stored information about blocks and their relationships to other 
+%physics objects, such as dataset and files, while PhEDEx kept
+%information about its location and replicas.
 
-Upon a user request to 
-find a block meta-data, we were able to identify appropriate set of 
-DBS and PhEDEx APIs, retrieve data, re-format their output 
-into DAS notations and store results into DAS cache. The records
+When a query for block meta-data is made, we are
+%Upon a user request to 
+%find a block meta-data, we were 
+able to identify the appropriate set of 
+DBS and PhEDEx APIs, retrieve the raw data from the services, format it
+for DAS use and add the resulting information to the DAS cache. The records
 were merged (aggregated) based on identical key-value pairs, e.g.
@@
 \end{verbatim}
-The \verb+das_id+ refers to api call records stored in Analytics DB,
-the \verb+block+ part represents an aggregation block meta-data information from DBS
-and PhEDEx systems and the \verb+das+ part contains their expiration time stamps.
+The \verb+das_id+ indicates the records stored in the DAS Analytics DB,
+the \verb+block+ part represents the aggregated block meta-data information from DBS
+and PhEDEx and the \verb+das+ part contains expiration time stamps
+about how long the information is valid.
 Usage of this information was valuable contribution in debugging process of
@@
 \end{itemize}
-Each data-service has its own scope, size and peculiar properties. 
-For instance, the data in PhEDEx are transient due to constant 
-migration of CMS data; the DBS system is divided into dozen 
-of individual services (instances) with own size and lifetime 
-used by different production teams for various bookkeeping tasks; 
+Each data-service has its own scope, size and peculiarities. 
+For instance, the data in PhEDEx is transient due to constant 
+migration of CMS data; the DBS system is divided into several individual 
+instances with sizes and lifetimes driven by specific processing tasks; 
 the RunSummary data-service is operated on private network accessible 
@@
 75 GB and 200 GB in tables and index sizes, respectively. We estimate to
-collect $\sim$500 GB of the meta-data each year during 
-normal CMS operation at LHC.
+collect around 500 GB of meta-data each year during 
+normal CMS operations. The average query rate to DAS is expected to be at the 
+$\leq$10K requests a day.
 %In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
@@
 
-To test DAS performance we ere measured elapsed time and CPU/RAM utilization
+To test DAS performance we measured the elapsed time and CPU/RAM utilization
 required to fetch, store and aggregate information from different
@@
 
-The initial DAS implementation is done in Python language \cite{Python}
-and all tests are performed on x64 Linux platform where
-1 CPU was used by DAS server and 1 CPU by MongoDB database server. 
+The initial DAS implementation is done in Python language \cite{Python}.
+All tests are performed on a 64 bit Linux platform with
+1 CPU for the DAS server and 1 CPU for the MongoDB database server. 
 All tests are performed multiple times to avoid fluctuations. 
@@
 \end{itemize}
-Here {\it retrieval time} is a time required to access data from data-service,
-{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
-is a time required to convert data-service meta-data to DAS records for every object
-we parse, {\it cache insertion} and {\it indexing time} represents time spent on DAS 
-cache back-end, {\it aggregation time} is required to merge objects into DAS records based
-on their common key (\verb+block.name+ in this case) and {\it output creation time}
-is a time required to write DAS records to disk. Please note that last two
+Here {\it retrieval time} is the time required to access data from data-service,
+{\it parsing time} is the time required to read and parse the received data
+from the services, and {\it re-mapping time} is the time required to convert 
+that information into DAS. {\it Cache insertion} and {\it indexing time} 
+represents time spent on DAS back-end caching, {\it aggregation time} is
+the time required to merge objects into DAS records based
+on their common keys (\verb+block.name+ in this case) and {\it output creation time}
+is the time required to write DAS records to disk. Please note that last two
 components, {\it aggregation} and {\it output creation time}, are only applied to
@@
 
-The individual tests of DAS components (DBS and PhEDEx) shown roughly 
-the same performance. Although the time of PhEDEx component is bigger 
-due to longer retrieval and parsing time as well as time spent
+The individual tests of DAS components (DBS and PhEDEx) show roughly 
+the same performance, although the time taken by the PhEDEx component is longer
+due to higher retrieval and parsing time as well as time spent
 during aggregation step. As shown in Table \ref{DAS_benchmark},
@@
 
-Based on the results of stress tests we are measured that we achieve throughput of
+The results of the stress tests have shown that we can achieve a throughput of
 $\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
-$\sim$7600 docs/sec for 
-reading operations in DAS, which is suitable for our needs. Since all 
-operations are performed on a single CPU, a further improvements are expected. 
-Also, a necessary tuning can be done by using C/C++ driver instead of python one.
+$\sim$7600 docs/sec for reading operations in DAS, which is suitable for our 
+needs. Please note, we have relatively small audience of users
+who perform complex queries, while search engines designed for
+a large number of users asking for imprecise, simple data.
+Since all tests were performed on a single CPU, further improvements from
+expanding to multiple CPUs are expected. In addition, more
+performance advantage can be gained by writing dedicated C/C++
+python extensions for data intensive operations.
 
 \section{Summary}
-We are presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
-at LHC, CERN, Geneva, Switzerland. It is designed to provide caching and
-aggregation layer on top of the existing relational and non-relation data-services
-mostly in real time fashion. All data are retrieved on demand basis,
-while pre-fetching of most common queries is in our plans. We are developed a
-prototype in Python which is under commissioning phase right now.
-The performance studies of the DAS system are presented in this paper. 
-The CMS experiment is started to collect its data in December of 2009 and 
-we expect to accumulate a few PB of real data each year. 
-In addition, the Monte-Carlo samples will be produced at this scale.
-Based on test studies performed in CMS, we expect that total size of
-meta-data produced each year by experiment will be of the order of
-500GB. The DAS system should sustain such load and provide generic 
-``data-discovery'' service for CMS experiment in years to come.
+%We are presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
+%at LHC, CERN, Geneva, Switzerland. It is designed to provide caching and
+%aggregation layer on top of the existing relational and non-relation data-services
+%mostly in real time fashion. All data are retrieved on demand basis,
+%while pre-fetching of most common queries is in our plans. We are developed a
+%prototype in Python which is under commissioning phase right now.
+%The performance studies of the DAS system are presented in this paper. 
+%The CMS experiment is started to collect its data in December of 2009 and 
+%we expect to accumulate a few PB of real data each year. 
+%In addition, the Monte-Carlo samples will be produced at this scale.
+%Based on test studies performed in CMS, we expect that total size of
+%meta-data produced each year by experiment will be of the order of
+%500GB. The DAS system should sustain such load and provide generic 
+%``data-discovery'' service for CMS experiment in years to come.
+In this paper we have presented a new data aggregation service (DAS) 
+developed for CMS High-Energy experiment at LHC, CERN, Geneva, Switzerland. 
+We have developed a prototype for DAS, which has been designed to provide 
+caching and aggregation layer on top of the existing relational and 
+non-relational data-services in a close to real time fashion. The data 
+is retrieved on demand, with pre-fetching of common queries, 
+determined from an extensive analytics database, also possible. 
+The DAS system is currently being commissioned for production use by CMS.
+
+Data taking at CMS began in December 2009, and, alongside simulation 
+data, we expect to accumulate petabytes of data yearly. This will have 
+of the order of a terabyte of associated meta-data that will need to be 
+easily queried by physicists. From the performance studies of DAS presented 
+in this paper we expect that the system will be able to sustain such load 
+and will provide generic “data-discovery” service for CMS experiment for 
+years to come.
 
@@
 
-This work was supported by the National Science Foundation and 
+This work was supported by the National Science Foundation, 
+contract No. PHY-0757894, and 
 Department of Energy of the United States of America.
-Fermilab is operated by Fermi Research Alliance, LLC under Contract
-No. DE-AC02-07CH11359 with the United States Department of Energy.
+
+%Fermilab is operated by Fermi Research Alliance, LLC under Contract
+%No. DE-AC02-07CH11359 with the United States Department of Energy.
 

### 2010-05-04 doc/das_paper/paper.tex
Relocate docs for das paper into separate folder
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@851 4525493e-7705-40b1-a816-d608a930855b

--- 2010-05-04 doc/paper.tex
@@
 to answer specific questions about the data they serve, the ability to search and relate
-information among them was remains a tedious human task. The growing amount of information
+information among them remains a tedious human task. The growing amount of information
 and desire to make cross-service queries led us to design and develop the new
@@
 plug-ins. The communication between client (web server or CLI) and 
-DAS cache server was done via the REST \cite{REST} interface. 
-Each input query was recorded into the Analytics DB for further 
-analysis, and was decomposed into a set of conditions and/or selection keys.
-First the results were looked up in a cache and delivered to the requester
-if superset of input conditions was found in analytics DB.\footnote{It is important
+DAS cache server is done via the REST \cite{REST} interface. 
+Each input query is recorded into the Analytics DB for further 
+analysis, and is decomposed into a set of conditions and/or selection keys.
+First the results are looked up in a cache and delivered to the requester
+if superset of input conditions is found in analytics DB.\footnote{It is important
 to note here that we didn't look-up results in a cache and rather
@@
 in order to understand if other API calls are covered the input request.}
-Otherwise, the appropriate workflow, a set of API calls, was constructed
+Otherwise, the appropriate workflow, a set of API calls, is constructed
 for the execution engine. It invoked the API calls via data-service
@@
 security modules for data-services accessible via private network.
-Finally, the results were retrieved from the cache and delivered to the user.
+Finally, the results are retrieved from the cache and delivered to the user.
 
@@
 DAS (PhEDEx) & XML & 186145/180439 & 164 sec & 0.98 sec \\
-DAS (aggregate) & JSON & 388108 & 234 sec & 2.05/50.7 sec \\
+DAS (total) & JSON & 388108 & 234 sec & 2.05/50.7 sec \\
 \hline
@@
 system shows fetched (186145) and merged (180439) records. 
+The last row shows aggregated results and total time
+spent in all components (DBS+PhEDEx+aggregation).
 The total number of DAS records are calculated as number of 
@@
 $\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
-$\sim$7600 docs/sec for reading operations in DAS, which is suitable for our 
+$\sim$7600 docs/sec for reading and writing DAS records to disk,
+which is suitable for our 
 needs. Please note, we have relatively small audience of users
-who perform complex queries, while search engines designed for
-a large number of users asking for imprecise, simple data.
+who perform complex queries and get precise results, while search 
+engines are designed for a large number of users asking for imprecise, simple data.
 Since all tests were performed on a single CPU, further improvements from

### 2010-05-04 doc/data_srv_example.py
Relocate stuff
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@852 4525493e-7705-40b1-a816-d608a930855b

@@ -1,27 +1,0 @@
-#!/usr/bin/env python
-
-from DAS.services.abstract_service import DASAbstractService
-from DAS.utils.utils import map_validator
-
-class MyService(DASAbstractService):
-    def __init__(self, config):
-        DASAbstractService.__init__(self, 'mysrv', config)
-        self.map = {
-            'myapi' : {
-                'keys': ['service_key_to_be_used_in_DAS'],
-                'params' : {'name':''}
-            },
-        }
-        map_validator(self.map)
-
-    def api(self, query, cond_dict=None):
-        """
-        A service worker. It parses input query, invoke service API 
-        and return results in a list with provided row.
-        Return: [{'key':'value', 'key2':'value2'}, ...]
-        """
-        # construct params out of provided query and cond_dict
-        data  = self.getdata(self.url, params)
-        # parse output data here
-        return data
-

### 2010-05-04 src/python/DAS/core/qlparser.py
Bug fix
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@969 4525493e-7705-40b1-a816-d608a930855b

@@ -9,8 +9,8 @@ tests integrity of DAS-QL queries, conversion routine from DAS-QL
 syntax to MongoDB one.
 """
 
-__revision__ = "$Id: qlparser.py,v 1.44 2010/03/01 20:38:02 valya Exp $"
-__version__ = "$Revision: 1.44 $"
+__revision__ = "$Id: qlparser.py,v 1.45 2010/03/02 01:34:01 valya Exp $"
+__version__ = "$Revision: 1.45 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -226,10 +226,11 @@ def fix_operator(query, pos, operators):
         if  pat.match(operator):
             operator = ' %s ' % operator
         idx = query[pos:].find(operator)
-        if  idx != -1 and idx < idx_pos:
+        if  idx == -1:
+            continue
+        if  idx < idx_pos:
             idx_pos = idx
             oper_pos = operator
-            break
     if  idx_pos >= len(query):
         return None, None
     if  idx_pos != -1:

### 2010-05-04 src/python/DAS/utils/utils.py
Change to xml_paser to handle child.text
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@970 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 General set of useful utilities used by DAS
 """
 
-__revision__ = "$Id: utils.py,v 1.75 2010/03/01 18:53:25 valya Exp $"
-__version__ = "$Revision: 1.75 $"
+__revision__ = "$Id: utils.py,v 1.76 2010/03/02 03:38:13 valya Exp $"
+__version__ = "$Revision: 1.76 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -857,7 +857,11 @@ def xml_parser(source, prim_key, tags=[]):
         row[key].update(sup)
         for child in elem.getchildren():
             child_key  = child.tag
-            child_dict = dict_helper(child.attrib, notations)
+            child_data = child.attrib
+            if  not child_data:
+                child_dict = adjust_value(child.text)
+            else:
+                child_dict = dict_helper(child_data, notations)
 
             if  row[key].has_key(child_key):
                 val = row[key][child_key]
@@ -898,7 +902,6 @@ def json_parser(source):
             res  = data.replace('null', '\"null\"')
         else:
             res  = data
-        print "res", res, type(res)
         try:
             jsondict = json.loads(res)
         except:

### 2010-05-04 doc/das_paper/paper.tex
Some work on future work paragraph
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@972 4525493e-7705-40b1-a816-d608a930855b

@@
 %% Example \runauth{C.V. Radhakrishnan et al.}
-\runauth{}
+\runauth{V. Kuznetsov, D. Evans, S. Metson}
 
@@
 
+% mdwlist allows you to create e.g. bulleted lists without extra line spacing
+\usepackage{mdwlist}
+
 % declarations for front matter
@@
 
-\dochead{}
+\dochead{International Conference on Computational Science, ICCS 2010}
 %% Use \dochead if there is an article header, e.g. \dochead{Short communication}
@@
 \address[vkuznet]{Cornell University, Ithaca, New York, USA}
-\ead{vkuznet@gmail.com}
+%\ead{vkuznet@gmail.com}
 
@@
 \address[evans]{Fermilab, Batavia, Illinois, USA}
-\ead{evansde@fnal.gov}
+%\ead{evansde@fnal.gov}
 
@@
 \address[metson]{Bristol University, Bristol, UK}
-\ead{s.metson@bristol.ac.uk}
+%\ead{s.metson@bristol.ac.uk}
 
@@
 keyword search over relational data sources.
-Section \ref{DataModel} describes the CMS experiment and its data model. In section
-\ref{DAS} we present the architecture of the DAS system, including discussion of its
+Section \ref{DataModel} breifly describes the Compact Muon Solenoid (CMS) experiment 
+and its data model. In section \ref{DAS} we present the architecture 
+of the DAS system, including discussion of its
 various components. Finally, our results are summarized in section \ref{Results}.
@@
 within CMS computing infrastructure by imposing the following set of requirements:
-\begin{itemize}
-\item support a keyword based search queries with ability to use conditional operators;
-%\item pluggable interface to existing and up-coming data-service APIs;
-\item use existing heterogeneous software environment and distributed nature of the data-services;
-\item preserve access rules and security policies of individual data-services;
-%\item be flexible to adjust to dynamic nature of data-service APIs, 
-%due to continuously changing detector and software environments;
-\item retrieve data on demand and aggregate them if necessary across
+support a keyword based search queries with ability to use conditional operators;
+use existing heterogeneous software environment and distributed nature of the data-services;
+preserve access rules and security policies of individual data-services;
+retrieve data on demand and aggregate them if necessary across
 multiple data-services;
-\item be transparent to data-services implementation and their data formats;
-%\item support legacy data-services, applications and APIs.
-\end{itemize}
+be transparent to data-services implementation and their data formats.
+%\begin{itemize}
+%\item support a keyword based search queries with ability to use conditional operators;
+%\item use existing heterogeneous software environment and distributed nature of the data-services;
+%\item preserve access rules and security policies of individual data-services;
+%\item retrieve data on demand and aggregate them if necessary across
+%multiple data-services;
+%\item be transparent to data-services implementation and their data formats;
+%\end{itemize}
 \begin{figure}[htb]
 \centering
-\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
+\includegraphics[width=90mm]{DAS_Cache_and_Analytics.pdf}
 \caption{
 The DAS architecture diagram. It consists of cache server,
-analytics and mapping databases. The data in a cache can be shared
+analytics, mapping and cache databases. The data in a cache can be shared
 across multiple queries. The information retrieval is triggered
@@
 optional DAS robots (Unix daemons) can be used to pre-fetch
-the most popular query requests.
+the most popular query requests. The solid, dashed and dotted lines 
+represent outer, inner and optional process communications between
+different DAS components, respectively.
 }
@@
 the core library with support of pluggable modules for data retrieval;
-caching layer to store and aggregate results;
+caching layer to store and aggregate results into DAS Cache and DAS merge
+databases, respectively;
 request service to handle user queries;
@@
 \begin{verbatim}
-{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
- "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
-              "api_param": "name", "das_key": "site"}], 
- "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
-              "key": "site", "map": "site.name"}]}
+{"system": "phedex", "format": "XML", "urn": "fileReplicas",
+ "url": "http://a.b.com/phedex/datasvc/xml/prod/fileReplicas", 
+ "params": {"node": "*", "block": "required", "se": "*"}, 
+ "api2das": [{"pattern": "", "das_key": "file.block.name", 
+    "api_param": "block"}, {"pattern": "re.compile('^T[0-3]_')", 
+    "das_key": "site", "api_param": "node"}], 
+ "daskeys": [{"map": "file.name", "key": "file", "pattern": ""}, 
+             {"map": "file.block.name", "key": "block", "pattern": ""}]}
 \end{verbatim}
-represents DAS mapping record.\footnote{
-The data-service URL's are kept in the configuration file.
-}
+represents DAS mapping record.
 This record can be read as following:
-the \verb+CMSNametoSE+ API provided by 
-\verb+sitedb+ data-service accepts \verb+name+ as input parameter. It
-confirms regular expression pattern, \verb+re.compile('^T[0-3]_')+, and
-corresponds to \verb+site+ DAS key. The \verb+site+ key
-maps into \verb+site.name+ within DAS record.
+the \verb+fileReplicas+ API (urn) provided by given url delivers data in
+XML data format and belongs to \verb+phedex+ system.
+It accepts \verb+node+, \verb+block+ and \verb+se+ input parameters, shown in \verb+params+ part
+of the record. Each record has two maps: \verb+api2das+ for mapping api input parameters into
+DAS keys and \verb+daskeys+ for mapping input DAS keys into DAS record access keys.
 
 Therefore the Mapping DB identifies a set of pre-defined DAS keys, 
-e.g. \verb+site, file, run+, etc., to be used by end-users in their queries.
+e.g. \verb+file, block+, etc., to be used by end-users in their queries.
 Each key is mapped onto set of API calls, whose output is retrieved,
@@
 Each request that DAS handles is recorded.
-The Analytics DB tracks:
-\begin{itemize}
-\item The users request
-\item How DAS has mapped the request onto data service API's
-\item The time taken by the remote data services to process the API calls
-\end{itemize}
+The Analytics DB tracks: the users request; how DAS has mapped the request 
+onto data service API's; the time taken by the remote data services to 
+process the API calls.
+
 By recording this information we can plan pre-fetch strategies for common queries,
@@
 The DAS caching system is used to dynamically fetch
-and aggregate data upon user requests. We use the following 
-algorithm (code snippets are written in Python \cite{Python}):
+and aggregate data upon user requests. It consits of
+DAS cache and DAS merge independent databases. The former one is used to store
+the raw results coming out from data-services, while the later contains aggregated
+records. Each data record in DAS cache/merge DBs contains expiration timestamps 
+provided either by data-service itself or based on default values during 
+data-service registration.
+We use the following algorithm (code snippets are written in Python \cite{Python}):
 \begin{enumerate}[1.]
@@
  "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
-            "subscribed": "n", "custodial": "n", "site": "T1_US_Buffer", 
-            "nfiles": "18", "se": "srm.site.com", 
-            "size": "112694171838"}, {....}]},
+            "subscribed": "n", "site": "T1_US_Buffer", "nfiles": "18", 
+            "se": "a.b.com", "size": "112694171838"}, {....}]},
            {"name": "a/b/c#123", "nevents": "570200", 
             "created_by": "USER_DN", "path": "/a/b/c", ...}],
- "das": [{"expire": 1259952908}, {"expire": 1259954702}]}
+ "das": {"expire": 1259952908}}
 \end{verbatim}
@@
 In this section we discuss preliminary results achieved with DAS system
-using the following CMS data-services:
-\begin{itemize}
-\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
-\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
-CMS data location and file transfer system;
-\item SiteDB \cite{SiteDB}, the CMS site resources;
-\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
-%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
-%\item DQ \cite{DQ}, Data Quality data-service which collects information
-%about detector conditions during data taking;
-%\item Overview \cite{Overview}, collects information about CMS
-%transfer rate, etc.;
-\item Dashboard \cite{Dashboard}, the LHC virtual organization
-monitoring data-service portal;
-\end{itemize}
-Each data-service has its own scope, size and peculiarities. 
-For instance, the data in PhEDEx is transient due to constant 
-migration of CMS data; the DBS system is divided into several individual 
-instances with sizes and lifetimes driven by specific processing tasks; 
-the RunSummary data-service is operated on private network accessible 
-at CMS online operation center.
+using different CMS data-services, see \cite{CMS, CMSDataModel}.
 
-At the moment, the amount of meta-data stored in all services is around
+%using the following CMS data-services:
+%Data Bookkeeping System (DBS) \cite{DBS}, the CMS meta-data service;
+%Physics Experiment Data Export (PhEDEx) \cite {PhEDEx}, the 
+%CMS data location and file transfer system;
+%SiteDB \cite{SiteDB}, the CMS site resources;
+%RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
+%Dashboard \cite{Dashboard}, the LHC virtual organization
+%monitoring data-service portal.
+
+%Each data-service has its own scope, size and peculiarities. 
+%For instance, the data in PhEDEx is transient due to constant 
+%migration of CMS data; the DBS system is divided into several individual 
+%instances with sizes and lifetimes driven by specific processing tasks; 
+%the RunSummary data-service is operated on private network accessible 
+%at CMS online operation center.
+
+At the moment, the amount of meta-data stored in all services participating
+in DAS is around
 75 GB and 200 GB in tables and index sizes, respectively. We estimate to
@@
 \hline
-DAS (DBS) & XML & 382396 & 63 sec & 0.98 sec \\
-DAS (PhEDEx) & XML & 186145/180439 & 164 sec & 0.98 sec \\
-DAS (total) & JSON & 388108 & 234 sec & 2.05/50.7 sec \\
+DBS yield & XML & 386943 & 68 sec & 0.98 sec \\
+PhEDEx yield & XML & 189679 & 107 sec & 0.98 sec \\
+Merge step & JSON & 576622 & 63 sec & 0.9 sec \\
+DAS total & JSON & 392635 & 238 sec & 2.05/50.7 sec \\
 \hline
@@
 \caption{Time required to fetch, parse and aggregate block information
-from DBS and PhEDEx systems. The PhEDEx
-system shows fetched (186145) and merged (180439) records. 
-The last row shows aggregated results and total time
-spent in all components (DBS+PhEDEx+aggregation).
+from DBS and PhEDEx systems. The last row shows aggregated results and 
+total time spent in all DAS components.
 The total number of DAS records are calculated as number of 
@@
         + cache insertion/indexing time 
-        + (aggregation time) + (output creation time)
+        + aggregation time + output creation time
 }
@@
 on their common keys (\verb+block.name+ in this case) and {\it output creation time}
-is the time required to write DAS records to disk. Please note that last two
-components, {\it aggregation} and {\it output creation time}, are only applied to
-one of the sub-system, either DBS or PhEDEx, depending on the insertion order.
-
-%DAS execution time (phedex) 160.606897116 sec
-%DAS execution time (dbs) 52.5349619389 sec
-%DAS execution time 256.0670228 sec, Sat, 28 Nov 2009 15:46:39 GMT
-%
-%DAS execution time (phedex) 0.952513933182 sec
-%DAS execution time (dbs) 0.951184034348 sec
-%DAS execution time 44.7982931137 sec, Sat, 28 Nov 2009 15:52:11 GMT
+is the time required to write DAS records to disk.
+% Please note that last two
+%components, {\it aggregation} and {\it output creation time}, are only applied to
+%one of the sub-system, either DBS or PhEDEx, depending on the insertion order.
 
@@
 the same performance, although the time taken by the PhEDEx component is longer
-due to higher retrieval and parsing time as well as time spent
-during aggregation step. As shown in Table \ref{DAS_benchmark},
+due to more complex records which leads to higher retrieval and parsing time. 
+As shown in Table \ref{DAS_benchmark},
 the look-up time spent to access individual components (last column) is quite
@@
 
-The results of the stress tests have shown that we can achieve a throughput of
-$\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
+The individual benchmarks of MongoDB have shown that it can deliver insert rate of
+20K docs/sec and provide look-up time of $10^-5$ sec per random document.
+The results of DAS stress tests have demonstrated that we can achieve a throughput of
+$\sim$6000 docs/sec for raw cache population, 
+%$\sim$6000 docs/sec for aggregation and 
 $\sim$7600 docs/sec for reading and writing DAS records to disk,
-which is suitable for our 
-needs. Please note, we have relatively small audience of users
+which is suitable for our needs. 
+Please note, we have relatively small audience of users
 who perform complex queries and get precise results, while search 
@@
 python extensions for data intensive operations.
+
+\section{Future work}
+The DAS system was desgined for CMS HEP experiment, while its architecture is
+transparent to participating data-providers and stored meta-data. 
+In the near future we plan to put DAS into production and confirm its scalability, 
+transparency and durability for various CMS data-services. Our goal is to sustain 
+1TB of meta-data per year. In long term, we would like to use DAS across various
+data-services in HEP community and other disciplines. Such work has began under 
+umbrella of DISCOVER Research Service Group at Cornell University \cite{DRSG} 
+who expressed their interest to use DAS in other scientific domains.
 
@@
 
-\bibitem{RunSummary}
-https://cmswbm.web.cern.ch/cmswbm/cmsdb/servlet/RunSummary
+%\bibitem{RunSummary}
+%https://cmswbm.web.cern.ch/cmswbm/cmsdb/servlet/RunSummary
 
-\bibitem{SiteDB}
-https://cmsweb.cern.ch/sitedb/
+%\bibitem{SiteDB}
+%https://cmsweb.cern.ch/sitedb/
 
@@
 
-\bibitem{Dashboard}
-J. Andreeva, et. al,
-``Experiment Dashboard – The Monitoring System for the LHC Experiments'',
-GMW’07, June 25, 2007, Monterey, California, USA.
+%\bibitem{Dashboard}
+%J. Andreeva, et. al,
+%``Experiment Dashboard – The Monitoring System for the LHC Experiments'',
+%GMW’07, June 25, 2007, Monterey, California, USA.
 
@@
 %Proc. of the 31st VLDB Conference, 2005.
+
+\bibitem{DRSG}
+http://drsg.cac.cornell.edu/
 \end{thebibliography}

### 2010-05-04 src/python/DAS/core/das_core.py
Use filters to override fields
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@973 4525493e-7705-40b1-a816-d608a930855b

@@ -13,8 +13,8 @@ It performs the following tasks:
 
 from __future__ import with_statement
 
-__revision__ = "$Id: das_core.py,v 1.62 2010/03/01 20:13:36 valya Exp $"
-__version__ = "$Revision: 1.62 $"
+__revision__ = "$Id: das_core.py,v 1.63 2010/03/02 15:56:47 valya Exp $"
+__version__ = "$Revision: 1.63 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -395,9 +395,11 @@ class DASCore(object):
                 % (query, idx, limit, skey, sorder)
         self.logger.info(msg)
         spec      = query.get('spec', {})
-#        fields    = query.get('fields', None)
+        fields    = query.get('fields', None)
         mapreduce = query.get('mapreduce', None)
-        fields    = query.get('filters', None)
+        filters   = query.get('filters', None)
+        if  filters:
+            fields = filters
         if  mapreduce:
             res = self.rawcache.map_reduce(mapreduce, spec)
         else:

### 2010-05-04 test/qlparser_t.py
Additional test for less/more operators
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@974 4525493e-7705-40b1-a816-d608a930855b

@@ -41,6 +41,11 @@ class testQLParser(unittest.TestCase):
 
     def test_fix_operator(self):
         """Test fix_operator function"""
+        query  = "data lat>1.1 lon<-1.1"
+        expect = "data lat > 1.1 lon < -1.1"
+        result = add_spaces(query, self.operators) 
+        self.assertEqual(expect, result)
+
         query  = "data latidue=1.1 longitude=-1.1"
         expect = "data latidue = 1.1 longitude = -1.1"
         result = add_spaces(query, self.operators) 

### 2010-05-04 src/python/DAS/core/das_mapping_db.py
Use flat-namespace for DAS keys
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@975 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 DAS mapping DB module
 """
 
-__revision__ = "$Id: das_mapping_db.py,v 1.33 2010/02/25 19:11:22 valya Exp $"
-__version__ = "$Revision: 1.33 $"
+__revision__ = "$Id: das_mapping_db.py,v 1.34 2010/03/02 15:59:51 valya Exp $"
+__version__ = "$Revision: 1.34 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -240,8 +240,9 @@ class DASMapping(object):
                 urn = row['urn']
                 for key in row['daskeys']:
                     if  key.has_key('map') and key['key'] == das_key:
-                        mapkeys.append((urn, key['map']))
-        return mapkeys
+                        return key['map']
+#                        mapkeys.append((urn, key['map']))
+#        return mapkeys
 
     def find_apis(self, das_system, map_key):
         """

### 2010-05-04 src/python/DAS/core/qlparser.py

@@ -9,8 +9,8 @@ tests integrity of DAS-QL queries, conversion routine from DAS-QL
 syntax to MongoDB one.
 """
 
-__revision__ = "$Id: qlparser.py,v 1.45 2010/03/02 01:34:01 valya Exp $"
-__version__ = "$Revision: 1.45 $"
+__revision__ = "$Id: qlparser.py,v 1.46 2010/03/02 15:59:51 valya Exp $"
+__version__ = "$Revision: 1.46 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -398,25 +398,23 @@ class MongoParser(object):
                     condlist.append(cdict)
                     continue
                 for system in self.map.list_systems():
-                    for api, mapkey in self.map.find_mapkey(system, key):
-                        prim_key = self.map.primary_key(system, api)
-                        lkeys = self.map.lookup_keys(system, key, 
-                                    api=api, value=value)
-#                        print "system", system, key, prim_key, lkeys
-                        for kkk in lkeys:
-                            cdict = dict(key=kkk, op=oper, value=value)
-                            if  skeys:
-                                if  prim_key in skeys:
-                                    if  cdict not in condlist:
-                                        condlist.append(cdict)
-                            else:
-                                if  prim_key == key:
-                                    if  cdict not in condlist:
-                                        condlist.append(cdict)
+                    mapkey = self.map.find_mapkey(system, key)
+                    if  mapkey:
+                        cdict = dict(key=mapkey, op=oper, value=value)
+                        if  cdict not in condlist:
+                            condlist.append(cdict)
             else:
                 if  word not in skeys and word in self.daskeys:
                     skeys.append(word)
             idx += 1
+        if  not condlist and skeys: # e.g. --query="dataset"
+            for key in skeys:
+                for system, daskeys in self.map.daskeys().items():
+                    if  key in daskeys:
+                        mapkey = self.map.find_mapkey(system, key)
+                        cdict = dict(key=mapkey, op="=", value="*")
+                        condlist.append(cdict)
+                        break
 #        print "\n### condlist", condlist
         spec = mongo_exp(condlist)
 #        print "### spec", spec

### 2010-05-04 src/python/DAS/utils/utils.py
Added _delete method to dotdict
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@981 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 General set of useful utilities used by DAS
 """
 
-__revision__ = "$Id: utils.py,v 1.76 2010/03/02 03:38:13 valya Exp $"
-__version__ = "$Revision: 1.76 $"
+__revision__ = "$Id: utils.py,v 1.77 2010/03/02 20:01:12 valya Exp $"
+__version__ = "$Revision: 1.77 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -67,6 +67,19 @@ class dotdict(dict):
     __setattr__ = dict.__setitem__
     __delattr__ = dict.__delitem__
 
+    def _delete(self, ckey):
+        """
+        Delete provided compound key from dot dictionary
+        """
+        obj  = self
+        keys = ckey.split('.')
+        for key in keys:
+            if  key == keys[-1]:
+                if  obj.has_key(key):
+                    del obj[key]
+            else:
+                obj = obj[key]
+        
     def _get(self, ckey):
         """
         Get value for provided compound key.

### 2010-05-04 doc/das_paper/paper.tex
More update on Future work section
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@982 4525493e-7705-40b1-a816-d608a930855b

@@
 \section{Future work}
-The DAS system was desgined for CMS HEP experiment, while its architecture is
-transparent to participating data-providers and stored meta-data. 
 In the near future we plan to put DAS into production and confirm its scalability, 
-transparency and durability for various CMS data-services. Our goal is to sustain 
-1TB of meta-data per year. In long term, we would like to use DAS across various
-data-services in HEP community and other disciplines. Such work has began under 
-umbrella of DISCOVER Research Service Group at Cornell University \cite{DRSG} 
-who expressed their interest to use DAS in other scientific domains.
+transparency and durability for various CMS data-services. 
+Our interests are plug-able interface for participating data-providers;
+analytics framework and pre-fetch strategies as well as horizontal scaling of the
+system. Our goal is to sustain 1TB of meta-data per year. 
+In a long term, we would like to use DAS across various
+data-services in HEP community and other disciplines. Even though the DAS was developed
+for CMS HEP experiment, its architecture is transparent to participating 
+data-providers and meta-data content. Therefore the generalization of DAS is feasible.
+Such work has began under the umbrella of DISCOVER Research Service Group at 
+Cornell University \cite{DRSG} who expressed their interest to use DAS in other 
+scientific domains.
 

### 2010-05-04 doc/das_paper/paper.tex
New set of changes
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@984 4525493e-7705-40b1-a816-d608a930855b

@@
 way to look-up the desired information.
-A well-known solutions, e.g. data-services, are tighten to a specific data 
-sources and end-users are left with a manual task to bookkeep and 
-relate information from them.
+Well-known solutions, e.g. data-services, are tightly coupled to a specific data 
+sources and end-users are left with the manual tasks of bookkeeping and 
+relating the information from them.
 
-Here we present our work on Data Aggregation System (DAS) which provides
-ability to search and aggregate information across different 
-data-services. While designed for CMS High-Energy Physics 
+Here we present our work on the Data Aggregation System (DAS) which provides
+the ability to search and aggregate information across different 
+data-services. While designed for the CMS High-Energy Physics 
 experiment at LHC, the strategies and technology could be
@@
 The rest of this paper is organized as follows. 
-In section \ref{RelatedWork} we discuss related work in a domain of 
+In section \ref{RelatedWork} we discuss related work in the domain of 
 keyword search over relational data sources.
@@
 Even though the idea of keyword queries over the relational database
-is not knew it is still under discussion in computer science domain.
+is not new, it is still under discussion in computer science domain.
 A few alternative solutions have been proposed to address this issue
@@
 single table or joined tables. Authors of \cite{QueryAnswer} make a step forward
-and have evaluated information around matched values. The results were presented as
+and evaluate information around matched values. The results are presented as
 customized templates to end-users. 
@@
 for more references), are based on an approach that builds some 
-kind of the graph over the underlying 
+kind of graph over the underlying 
 database and generates a new database out of the source database to perform the work.
-The proximity of the results in those techniques are mostly confined to a 
-text-based values stored in a database, since the usage of numerical ones 
-is mostly useless without knowledge of the context of the input value. 
+The proximity of the results in those techniques are mostly confined to 
+text-based values stored in a database, since numerical ones 
+are mostly useless without knowledge of the context of the input value. 
 For example, typing a value 100 in the search input field can lead to 
 plenty of inappropriate matches, e.g. row ids. In order to be
-answered correctly it requires an additional keyword to clarify its meaning, 
+answered correctly, an additional keyword is required to clarify its meaning, 
 which should be matched with particular entity (table.column) in the 
-underlying database. Therefore the logical questions, e.g.
+underlying database. Therefore logical questions, e.g.
 {\it find me the total number of files whose size more then 20 but less then 
 100} are hard to answer precisely using approaches outlined in \cite{DBXplorer, QueryAnswer}, 
-while it can be easily accomplished using SQL.
+while they can be easily accomplished using SQL.
 
 To address these limitations a keyword-based query language was introduced in \cite{DBS-QL}.
-It used a set of pre-defined keys mapped to an underlying schema and shortest path
+It used a set of pre-defined keys mapped to an underlying schema and a shortest path
 algorithm which employed foreign key constraints to build SQL queries. As a result
 the question
-{\it I'm looking for files who contain data taken on certain date and located at
-particular site} was represented as simply as \cite{DBS-QL}
+{\it I'm looking for files which contain data taken on a certain date and located at
+a particular site} was represented as simply as \cite{DBS-QL}
 \begin{verbatim}
@@
 
-At the same time the data aggregation is commonly addressed in the context of
-federated databases, see for example \cite{FedDB}. The data from different 
+At the same time, data aggregation is commonly addressed in the context of
+federated data\-ba\-ses, see for example \cite{FedDB}. The data from different 
 RDMS systems are stored into the federated DB where SQL queries can be used to 
-search desired data. Apart from many internal obstacles, such as
+search for desired data. Apart from many internal obstacles, such as
 schema, domain and naming conflicts, it is external issues that are often
 a large barrier within a distributed, heterogeneous environment where
-different security policies are in place. At the end the usage of SQL and/or
+different security policies are in place. In the end the usage of SQL and/or
 any other Query Language is required to search for data.
@@
 existing data-services whose nature and access policies are in place and well understood.
-The key ingredients were the setup of the mapping and analytics services which provided
+The key ingredients are the mapping and analytics services which provide
 the necessary glue across existing API/notations and user queries. This approach 
@@
 incorporate a broad variety of both mass data storage and data processing
-resources into a single coherent system. CMS uses
+resources into a single coherent system. CMS uses a
 multi-Tiered model, where specific tasks related to data taking such as
@@
 \section{Data Aggregation System\label{DAS}}
-The goal of the Data Aggregation System (DAS) was to fetch and aggregate meta-data 
+The goal of the Data Aggregation System (DAS) is to fetch and aggregate meta-data 
 on demand from existing CMS data-services under one umbrella by using their APIs, 
-security policies, etc. In contrast, the approaches
+security policies, etc. By contrast, the approaches
 discussed in \cite{DBXplorer, QueryAnswer, FedDB}
@@
 data representation and security policies are left intact. Consequently,
-DAS does not require a data preservation and transaction capabilities. 
-Instead DAS can be treated as proxy or caching layer where information 
+DAS does not require data preservation and transaction capabilities. 
+Instead DAS can be treated as a proxy or caching layer where information 
 retrieval and aggregation are triggered dynamically by user queries. 
-Below we discuss in details the DAS architecture.
+Imagine: a user opens a browser and places a query, e.g.
+{\it file block=abc}. The DAS system responds with aggregated infromation about files
+for the provided set of conditions from all CMS data-services. Information
+is either retrieved from the DAS cache or fetched on-demand from relevant
+data-providers. Such a system provides uniform access to all meta-data in CMS 
+via simple, intuitive query interface. It neither forces users to learn 
+the structure of CMS data-services
+nor requires data-providers to agree on common data representations
+and the requirement to have central meta-data repository.
+
+%Below we discuss in details the DAS architecture.
 
 \subsection{DAS architecture}
-The design of the DAS system was based on previous studies of CMS Data 
+The design of the DAS system was based on previous studies of the CMS Data 
 Bookkeeping System (DBS) \cite{DBS, DBS07}. The DBS Data Discovery service
-\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
-a syntax similar to SQL, without specific knowledge of the underlying DB schema.
+\cite{DD} was based on the DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
+a syntax similar to SQL, without requiring specific knowledge of the underlying DB schema.
 Quick adoption and wide usage of DBS-QL in CMS gave us confidence in the chosen approach 
@@
 discovery service.
-Starting from this ground we are designed the DAS system as an
+Starting from here we designed the DAS system as an
 additional layer on top of the existing data-services
-within CMS computing infrastructure by imposing the following set of requirements:
-support a keyword based search queries with ability to use conditional operators;
-use existing heterogeneous software environment and distributed nature of the data-services;
-preserve access rules and security policies of individual data-services;
+within the CMS computing infrastructure by imposing the following set of requirements:
+support keyword based search queries with the ability to use conditional operators;
+use existing heterogeneous software environments and the
+distributed nature of the data-services;
+preserve the access rules and security policies of individual data-services;
 retrieve data on demand and aggregate them if necessary across
@@
 
-\noindent
-The choice of existing RDMS as DAS back-end(s) were ruled out for several reasons. 
+The choice of an existing RDMS as the DAS back-end(s) was ruled out for several reasons. 
 For instance, we didn't require any transactions and data persistency in DAS;
-a dynamically typing of stored meta-data objects was one of the requirements, etc.
+the dynamically typing of stored meta-data objects was one of the requirements, etc.
 Those arguments forced us to look for alternative IT solutions.
-Through the careful analysis of available options, such as a file-based and in memory caches, 
-key-value databases, documented-oriented databases we made our choice in favor 
+Through the careful analysis of available options, such as file-based and in memory caches, 
+key-value databases and documented-oriented databases, we made our choice in favor 
 of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
-MongoDB \cite{MongoDB}. Both systems are ``schema-less'' databases providing
+MongoDB \cite{MongoDB}. Both are ``schema-less'' databases providing
 arbitrary document structure storage, replication and fail-over features. 
@@
 the desired load and size capable of handling our meta-data information. We used 
-document-oriented database, MongoDB, as a back-end for three 
+the document-oriented database, MongoDB, as a back-end for three 
 main DAS components: Mapping, Analytics and DAS cache databases, 
@@
 the core library with support of pluggable modules for data retrieval;
-caching layer to store and aggregate results into DAS Cache and DAS merge
-databases, respectively;
+the caching layer to store and aggregate results into DAS Cache and DAS merge
+databases, respectively; the
 request service to handle user queries;
-Mapping DB to keep information about data-service APIs, e.g.
-notations and mapping to DAS keys;
+the Mapping DB to keep information about data-service APIs, e.g.
+notations and mapping to DAS keys and the
 Analytics DB for query analysis.
@@
 First the results are looked up in a cache and delivered to the requester
-if superset of input conditions is found in analytics DB.\footnote{It is important
-to note here that we didn't look-up results in a cache and rather
-compared input query condition set to the records in analytics DB 
-in order to understand if other API calls are covered the input request.}
+if a superset of the input conditions is found in analytics DB.\footnote{It is important
+to note here that we didn't look-up results in a cache, but rather
+compared the input query condition set to the records in the analytics DB 
+in order to understand if other API calls covered the input request.}
 Otherwise, the appropriate workflow, a set of API calls, is constructed
-for the execution engine. It invoked the API calls via data-service
-plug-ins and stored results into the cache. At this step we also
+for the execution engine. It invokes the API calls via the data-service
+plug-ins and stores the results into the cache. At this step we also
 performed necessary data aggregation for records with similar key-value
@@
  "params": {"node": "*", "block": "required", "se": "*"}, 
- "api2das": [{"pattern": "", "das_key": "file.block.name", 
+ "das2api": [{"pattern": "", "das_key": "file.block.name", 
     "api_param": "block"}, {"pattern": "re.compile('^T[0-3]_')", 
@@
 \end{verbatim}
-represents DAS mapping record.
+represents a DAS mapping record.
 This record can be read as following:
 the \verb+fileReplicas+ API (urn) provided by given url delivers data in
-XML data format and belongs to \verb+phedex+ system.
+XML data format and belongs to the \verb+phedex+ system.
 It accepts \verb+node+, \verb+block+ and \verb+se+ input parameters, shown in \verb+params+ part
-of the record. Each record has two maps: \verb+api2das+ for mapping api input parameters into
-DAS keys and \verb+daskeys+ for mapping input DAS keys into DAS record access keys.
+of the record. Each record has two maps: 
+\verb+daskeys+ for mapping input DAS keys into DAS record access keys and
+\verb+das2api+ for mapping DAS access keys into api input parameters.
 
-Therefore the Mapping DB identifies a set of pre-defined DAS keys, 
+The Mapping DB identifies a set of pre-defined DAS keys, 
 e.g. \verb+file, block+, etc., to be used by end-users in their queries.
@@
 The DAS caching system is used to dynamically fetch
-and aggregate data upon user requests. It consits of
-DAS cache and DAS merge independent databases. The former one is used to store
-the raw results coming out from data-services, while the later contains aggregated
-records. Each data record in DAS cache/merge DBs contains expiration timestamps 
-provided either by data-service itself or based on default values during 
+and aggregate data upon user requests. It consists of the
+DAS cache and the DAS merge independent databases. The former is used to store
+the raw results coming out from data-services, while the latter contains aggregated
+records. Each data record in the DAS cache/merge DBs contains expiration timestamps 
+provided either by the data-service itself or based on default values during 
 data-service registration.
@@
 \end{verbatim}
-Here \verb+r+ represents a triplet \verb+url, api, args+ and 
-\verb+mappind_db+ defines translation between input query and data-service
+Here \verb+r+ represents a triplet \verb+url, api, args+, and 
+\verb+mappind_db+ defines the translation between the input query and the data-service
 APIs which can delivery the data.
@@
 \end{verbatim}
-Here \verb+superset_in_analytics+ checks if provided \verb+api, args+
-are subset of already registered in Analytics DB api call(s). The 
+Here \verb+superset_in_analytics+ checks if the provided \verb+api, args+
+are a subset of api call(s) already registered in the Analytics DB. The 
 \verb+get_data+ and \verb+parser+ are define methods to fetch and parse
@@
 for \verb+file+ record.
-We did use a bulk insert operations at this step to increase throughput. 
+Bulk insert operations are used at this step to increase throughput. 
 \item Get results from the cache
@@
 %common data into the cache. 
-We can also automate populating the cache by running an 
+The cache population can be automated by running an
 independent set of daemons, DAS robots, which monitor the 
@@
 for nested structures, e.g. \verb+block.replica.size+. Our users 
-are able to specify the context of their queries by using appropriate 
+are able to specify the context of their queries by using the appropriate 
 attribute(s). For example, a query for \verb+block+ represents a search for all
@@
 retrieve all run records for the given range of run numbers. 
-As a complement to the queries, we will also be adding a dynamic index
-of DAS records with a keyword-based search within record content later.
+As a complement to the queries, we will soon be adding a dynamic index
+of DAS records with a keyword-based search within record content.
 
-All input queries are transformed into the MongoDB query syntax who is itself
+All input queries are transformed into the MongoDB query syntax which is itself
 a JSON record. It contains a set of selection 
-keys (\verb+fields+) and conditions (\verb+spec+). For example, a query
+keys (\verb+fields+) and conditions (\verb+spec+). For example, the query
 \verb+dataset, file=abc, run > 10+
@@
 %source of information about all data-services participating in DAS.
-Based on the information stored in Mapping DB we are able to 
+Based on the information stored in the Mapping DB we are able to 
 retrieve appropriate data from different data-services, re-map them into 
@@
 for DAS use and add the resulting information to the DAS cache. The records
-were merged (aggregated) based on identical key-value pairs, e.g.
+are merged (aggregated) based on identical key-value pairs, e.g.
 based on \verb+block.name+ key. Here is an example of such an aggregated record:
@@
 about how long the information is valid.
-Usage of this information was valuable contribution in debugging process of
+Usage of this information was a valuable contribution in debugging process of
 data-services themselves, e.g. identification of data-service inconsistencies, 
@@
 \section{Results\label{Results}}
-In this section we discuss preliminary results achieved with DAS system
+In this section we discuss preliminary results achieved with the DAS system
 using different CMS data-services, see \cite{CMS, CMSDataModel}.
@@
 in DAS is around
-75 GB and 200 GB in tables and index sizes, respectively. We estimate to
-collect around 500 GB of meta-data each year during 
-normal CMS operations. The average query rate to DAS is expected to be at the 
+75 GB and 200 GB in tables and index sizes, respectively. We estimate 
+collecting around 500 GB of meta-data each year during 
+normal CMS operations. The average query rate to DAS is expected to be
 $\leq$10K requests a day.
@@
 total time spent in all DAS components.
-The total number of DAS records are calculated as number of 
+The total number of DAS records are calculated as the number of 
 aggregated records plus left overs, the non-matched records from both systems. 
-The look-up time shown in last column represents query look-up 
+The look-up time shown in the last column represents the query look-up 
 time (2.05 sec) and time required to get all records written 
@@
 
-The initial DAS implementation is done in Python language \cite{Python}.
+The initial DAS implementation is done in the Python language \cite{Python}.
 All tests are performed on a 64 bit Linux platform with
@@
 the look-up time spent to access individual components (last column) is quite
-reasonable, roughly 1 second for both systems. While the final time to
+reasonable, roughly 1 second for both systems. The final time to
 get DAS records on disk is about 50 seconds. This is mostly due to I/O and
-conversion operations between binary data format used by MongoDB, BSON, to DAS
-data format, JSON.
+conversion operations from the binary data format used by MongoDB, BSON, 
+to the DAS data format, JSON.
 
-The individual benchmarks of MongoDB have shown that it can deliver insert rate of
-20K docs/sec and provide look-up time of $10^-5$ sec per random document.
+The individual benchmarks of MongoDB have shown that it can deliver an insert rate of
+20K docs/sec and provide a look-up time of $10^{-5}$ sec per random document.
 The results of DAS stress tests have demonstrated that we can achieve a throughput of
@@
 which is suitable for our needs. 
-Please note, we have relatively small audience of users
+Please note, we have a relatively small audience of users
 who perform complex queries and get precise results, while search 
@@
 transparency and durability for various CMS data-services. 
-Our interests are plug-able interface for participating data-providers;
+Our interests are plug-able interface for participating data-providers,
 analytics framework and pre-fetch strategies as well as horizontal scaling of the
@@
 In a long term, we would like to use DAS across various
-data-services in HEP community and other disciplines. Even though the DAS was developed
-for CMS HEP experiment, its architecture is transparent to participating 
+data-services in the HEP community and other disciplines. Even though DAS was developed
+for the CMS experiment, its architecture is transparent to participating 
 data-providers and meta-data content. Therefore the generalization of DAS is feasible.
 Such work has began under the umbrella of DISCOVER Research Service Group at 
-Cornell University \cite{DRSG} who expressed their interest to use DAS in other 
+Cornell University \cite{DRSG} which has expressed their interest to use DAS in other 
 scientific domains.
@@
 \section{Summary}
-%We are presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
-%at LHC, CERN, Geneva, Switzerland. It is designed to provide caching and
-%aggregation layer on top of the existing relational and non-relation data-services
-%mostly in real time fashion. All data are retrieved on demand basis,
-%while pre-fetching of most common queries is in our plans. We are developed a
-%prototype in Python which is under commissioning phase right now.
-%The performance studies of the DAS system are presented in this paper. 
-%The CMS experiment is started to collect its data in December of 2009 and 
-%we expect to accumulate a few PB of real data each year. 
-%In addition, the Monte-Carlo samples will be produced at this scale.
-%Based on test studies performed in CMS, we expect that total size of
-%meta-data produced each year by experiment will be of the order of
-%500GB. The DAS system should sustain such load and provide generic 
-%``data-discovery'' service for CMS experiment in years to come.
 In this paper we have presented a new data aggregation service (DAS) 
-developed for CMS High-Energy experiment at LHC, CERN, Geneva, Switzerland. 
-We have developed a prototype for DAS, which has been designed to provide 
-caching and aggregation layer on top of the existing relational and 
+developed for the CMS High-Energy experiment at LHC, CERN, Geneva, Switzerland. 
+We have developed a prototype for DAS, designed to provide 
+caching and aggregation layers on top of the existing relational and 
 non-relational data-services in a close to real time fashion. The data 

### 2010-05-04 test/das_aggregators_t.py
Unit test for DAS aggregators
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@985 4525493e-7705-40b1-a816-d608a930855b

@@ -1,0 +1,48 @@
+#!/usr/bin/env python
+#pylint: disable-msg=C0301,C0103
+
+"""
+Unit test for DAS aggregators
+"""
+
+import unittest
+from DAS.core.das_aggregators import das_func
+from DAS.utils.utils import dotdict
+
+class testQLParser(unittest.TestCase):
+    """
+    A test class for the DAS qlparser
+    """
+    def testBracketObj(self):                          
+        """test search for bracket objects"""
+        rows = []
+        data = {'block': {'name':'AAA', 'replica': [{'name':'a', 'size':1}, {'name':'b', 'size':10}] }}
+        rows.append(data)
+        data = {'block': {'name':'AAA', 'replica': [{'name':'a', 'size':2}, {'name':'b', 'size':20}] }}
+        rows.append(data)
+
+        expect = 33
+        result = das_func('sum', 'block.replica.size', rows)
+        self.assertEqual(expect, result)
+
+        expect = 4
+        result = das_func('count', 'block.replica.size', rows)
+        self.assertEqual(expect, result)
+
+        expect = 1
+        result = das_func('min', 'block.replica.size', rows)
+        self.assertEqual(expect, result)
+
+        expect = 20
+        result = das_func('max', 'block.replica.size', rows)
+        self.assertEqual(expect, result)
+
+        expect = 20
+        drows  = [dotdict(row) for row in rows]
+        result = das_func('max', 'block.replica.size', drows)
+        self.assertEqual(expect, result)
+#
+# main
+#
+if __name__ == '__main__':
+    unittest.main()

### 2010-05-04 doc/das_paper/paper.tex
Final tupe-up
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@995 4525493e-7705-40b1-a816-d608a930855b

@@
 
-%\author[vkuznet]{Valentin Kuznetsov\corref{cor1}}
-\author[vkuznet]{Valentin Kuznetsov}
+\author[vkuznet]{Valentin Kuznetsov\corref{cor1}}
+%\author[vkuznet]{Valentin Kuznetsov}
 \address[vkuznet]{Cornell University, Ithaca, New York, USA}
-%\ead{vkuznet@gmail.com}
-
+\ead{vkuznet@gmail.com}
 \author[evans]{Dave Evans}
@@
 %\ead{evansde@fnal.gov}
-
 \author[metson]{Simon Metson}
@@
 
-%\cortext[cor1]{Corresponding author}
+\cortext[cor1]{Corresponding author}
 
@@
  "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
-            "subscribed": "n", "site": "T1_US_Buffer", "nfiles": "18", 
-            "se": "a.b.com", "size": "112694171838"}, {....}]},
-           {"name": "a/b/c#123", "nevents": "570200", 
-            "created_by": "USER_DN", "path": "/a/b/c", ...}],
+            "site": "T1_US", "nfiles": "18", "se": "a.b.com"}, {...}]},
+           {"name": "/a/b/c#123", "nevents": "570200", "path":"/a/b/c"}],
  "das": {"expire": 1259952908}}
@@
 The look-up time shown in the last column represents the query look-up 
-time (2.05 sec) and time required to get all records written 
-to disk (50.7 sec).}
+time (2.05s) and time required to get all records written 
+to disk (50.7s).}
 \label{DAS_benchmark}

### 2010-05-04 doc/das_paper/paper.tex
Put Dave's comments
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@996 4525493e-7705-40b1-a816-d608a930855b

@@
 and data distribution to Tier-1 centers, located around the world.
-The Tier-1 centers provide archival storage and CPU power for high
-priority selection and re-reconstructions jobs and distribution points
+The Tier-1 centers provide archival storage, CPU power for high
+priority selection and re-reconstructions, and distribution points
 to the Tier 2 resources.
@@
 For instance, we didn't require any transactions and data persistency in DAS;
-the dynamically typing of stored meta-data objects was one of the requirements, etc.
+the dynamic typing of stored meta-data objects was one of the requirements, etc.
 Those arguments forced us to look for alternative IT solutions.

### 2010-05-04 src/python/DAS/core/das_ql.py
Combine all DAS QL definitions in one comon place
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@997 4525493e-7705-40b1-a816-d608a930855b

@@ -1,0 +1,57 @@
+#!/usr/bin/env python
+
+"""
+DAS QL module include definition of DAS operators,
+filters, aggregators, etc.
+"""
+
+__revision__ = "$Id: das_ql.py,v 1.1 2010/03/04 15:34:35 valya Exp $"
+__version__ = "$Revision: 1.1 $"
+__author__ = "Valentin Kuznetsov"
+
+import inspect
+from DAS.core.das_aggregators import ResultObject
+
+DAS_FILTERS = ['grep']
+DAS_OPERATORS = ['!=', '<=', '<', '>=', '>', '=', 
+                 'between', 'nin', 'in', 'last']
+MONGO_MAP = {
+    '>':'$gt',
+    '<':'$lt',
+    '>=':'$gte',
+    '<=':'$lte',
+    'in':'$in',
+    '!=':'$ne',
+    'nin':'$nin',
+}
+def mongo_operator(das_operator):
+    """
+    Convert DAS operator into MongoDB equivalent
+    """
+    if  MONGO_MAP.has_key(das_operator):
+        return MONGO_MAP[das_operator]
+    return None
+
+def das_filters():
+    """
+    Return list of DAS filters
+    """
+    return DAS_FILTERS
+
+def das_operators():
+    """
+    Return list of DAS operators
+    """
+    return DAS_OPERATORS
+
+def das_aggregators():
+    """
+    Inspect ResultObject class and return its member function,
+    which represents DAS aggregator functions
+    """
+    alist = []
+    for name, ftype in inspect.getmembers(ResultObject):
+        if  name.find("__") != -1:
+            continue
+        alist.append(name)
+    return alist

### 2010-05-04 src/python/DAS/core/das_aggregators.py
Switch to use das_ql module
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@998 4525493e-7705-40b1-a816-d608a930855b

@@ -7,8 +7,11 @@ das_func sink, which by itself hold ResultObject as a result
 holder.
 """
 
+__revision__ = "$Id: das_aggregators.py,v 1.2 2010/03/04 15:43:36 valya Exp $"
+__version__ = "$Revision: 1.2 $"
+__author__ = "Valentin Kuznetsov"
+
 import types
-import inspect
 from DAS.utils.utils import dict_type
 
 class ResultObject(object):
@@ -43,25 +46,13 @@ class ResultObject(object):
             self.result = 99999999999
         if  value < self.result:
             self.result = value
-    def avg(self, value):
-        """Average function for this object"""
-        pass
-    def median(self, value):
-        """Median function for this object"""
-        pass
+#    def avg(self, value):
+#        """Average function for this object"""
+#        pass
+#    def median(self, value):
+#        """Median function for this object"""
+#        pass
     
-def das_aggregators():
-    """
-    Inspect ResultObject class and return its member function,
-    which represents DAS aggregator functions
-    """
-    alist = []
-    for name, ftype in inspect.getmembers(ResultObject):
-        if  name.find("__") != -1:
-            continue
-        alist.append(name)
-    return alist
-
 def coroutine(func):
     """Coroutine helper, to be used as decorator"""
     def start(*args, **kwargs):

### 2010-05-04 setup.py
Use cvs tag as DAS version
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1040 4525493e-7705-40b1-a816-d608a930855b

@@ -12,6 +12,7 @@ from distutils.errors import CCompilerError
 from distutils.errors import DistutilsPlatformError, DistutilsExecError
 from distutils.core import Extension
 from distutils.command.install import INSTALL_SCHEMES
+from subprocess import Popen, PIPE
 
 requirements = []
 try:
@@ -82,8 +83,13 @@ if "--no_ext" in sys.argv:
 else:
     features = {"c-ext": c_ext}
 
-
-version      = "1.0.0" # need to define it somehow
+proc1        = Popen(['cvs', 'status', '-v', 'setup.py'], stdout=PIPE)
+proc2        = Popen(['grep', 'revision'], stdin=proc1.stdout, stdout=PIPE)
+proc3        = Popen(['grep', '-v', 'Repository'], stdin=proc2.stdout, stdout=PIPE)
+proc4        = Popen(['grep', '-v', 'Working'], stdin=proc3.stdout, stdout=PIPE)
+proc5        = Popen(['head', '-1'], stdin=proc4.stdout, stdout=PIPE)
+version      = proc5.communicate()[0].split()[0]
+#version      = "1.0.0" # need to define it somehow
 name         = "DAS"
 description  = "CMS Data Aggregation System"
 readme       ="""

### 2010-05-04 src/python/DAS/web/tools.py
Tune-ups
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1041 4525493e-7705-40b1-a816-d608a930855b

@@ -6,8 +6,8 @@ Web tools.
 """
 
 __license__ = "GPL"
-__revision__ = "$Id: tools.py,v 1.2 2010/02/18 15:08:14 valya Exp $"
-__version__ = "$Revision: 1.2 $"
+__revision__ = "$Id: tools.py,v 1.3 2010/03/14 20:20:58 valya Exp $"
+__version__ = "$Revision: 1.3 $"
 __author__ = "Valentin Kuznetsov"
 __email__ = "vkuznet@gmail.com"
 
@@ -155,7 +155,7 @@ def exposejson (func):
             jsondata = encoder.encode(data)
             return jsondata
         except:
-            Exception("Fail to jsontify obj '%s' type '%s'" % (data, type(data)))
+            Exception("Fail to JSONtify obj '%s' type '%s'" % (data, type(data)))
     wrapper.__doc__ = func.__doc__
     wrapper.__name__ = func.__name__
     wrapper.exposed = True
@@ -204,14 +204,13 @@ def exposedasjson (func):
     """
     def wrapper (self, *args, **kwds):
         encoder = JSONEncoder()
-#        data = runDas(self, func, *args, **kwds)
         data = func(self, *args, **kwds)
         cherrypy.response.headers['Content-Type'] = "application/json"
         try:
             jsondata = encoder.encode(data)
             return jsondata
         except:
-            Exception("Failed to json-ify obj '%s' type '%s'" % (data, type(data)))
+            Exception("Failed to JSONtify obj '%s' type '%s'" % (data, type(data)))
 
     wrapper.__doc__ = func.__doc__
     wrapper.__name__ = func.__name__
@@ -224,10 +223,8 @@ def exposedasplist (func):
     see http://docs.python.org/library/plistlib.html#module-plistlib
     """
     def wrapper (self, *args, **kwds):
-#        data_struct = runDas(self, func, *args, **kwds)
         data_struct = func(self, *args, **kwds)
         plist_str = plistlib.writePlistToString(data_struct)
-        cherrypy.response.headers['ETag'] = data_struct['results'].__str__().__hash__()
         cherrypy.response.headers['Content-Type'] = "application/xml"
         return plist_str
     wrapper.__doc__ = func.__doc__

### 2010-05-04 src/python/DAS/web/das_server.py
Adjust code quality
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1043 4525493e-7705-40b1-a816-d608a930855b

@@ -1,25 +1,24 @@
 #!/usr/bin/env python
+#-*- coding: ISO-8859-1 -*-
+#pylint: disable-msg=W0702,E1101
 """
 DAS server based on CherryPy web framework. We define Root class and
 pass it into CherryPy web server.
 """
 
-__revision__ = "$Id: das_server.py,v 1.3 2010/03/10 01:19:56 valya Exp $"
-__version__ = "$Revision: 1.3 $"
+__revision__ = "$Id: das_server.py,v 1.4 2010/03/15 02:44:09 valya Exp $"
+__version__ = "$Revision: 1.4 $"
 __author__ = "Valentin Kuznetsov"
 
 # system modules
 import os
 import sys
 import yaml
-import types
 import logging 
 from optparse import OptionParser
 
 # CherryPy modules
-import cherrypy
-from cherrypy import quickstart, expose, server, log
-from cherrypy import tree, engine, dispatch, tools
+from cherrypy import log, tree, engine
 from cherrypy import config as cpconfig
 
 # DAS modules
@@ -30,6 +29,9 @@ from DAS.web.DASCacheModel import DASCacheModel
 from DAS.web.das_doc import DocServer
 
 class Root(object):
+    """
+    DAS web server class.
+    """
     def __init__(self, model, config):
         self.model  = model
         self.config = config
@@ -50,19 +52,23 @@ class Root(object):
         except:
             cpconfig.update ({"server.socket_host": '0.0.0.0'})
         try:
-            cpconfig.update ({'tools.expires.secs': int(self.config['expires'])})
+            cpconfig.update ({'tools.expires.secs': 
+                                int(self.config['expires'])})
         except:
             cpconfig.update ({'tools.expires.secs': 300})
         try:
-            cpconfig.update ({'log.screen': bool(self.config['log_screen'])})
+            cpconfig.update ({'log.screen': 
+                                bool(self.config['log_screen'])})
         except:
             cpconfig.update ({'log.screen': True})
         try:
-            cpconfig.update ({'log.access_file': self.config['access_log_file']})
+            cpconfig.update ({'log.access_file': 
+                                self.config['access_log_file']})
         except:
             cpconfig.update ({'log.access_file': None})
         try:
-            cpconfig.update ({'log.error_file': self.config['error_log_file']})
+            cpconfig.update ({'log.error_file': 
+                                self.config['error_log_file']})
         except:
             cpconfig.update ({'log.error_file': None})
         try:
[…]

### 2010-05-04 src/python/DAS/web/tools.py

@@ -6,8 +6,8 @@ Web tools.
 """
 
 __license__ = "GPL"
-__revision__ = "$Id: tools.py,v 1.3 2010/03/14 20:20:58 valya Exp $"
-__version__ = "$Revision: 1.3 $"
+__revision__ = "$Id: tools.py,v 1.4 2010/03/15 02:44:09 valya Exp $"
+__version__ = "$Revision: 1.4 $"
 __author__ = "Valentin Kuznetsov"
 __email__ = "vkuznet@gmail.com"
 
@@ -24,17 +24,13 @@ from wsgiref.handlers import format_date_time
 # cherrypy modules
 import cherrypy
 from cherrypy import log as cplog
-from cherrypy import request
+from cherrypy import expose
 
 # cheetag modules
 from Cheetah.Template import Template
 from Cheetah import Version
 
-try:
-    from json import JSONEncoder
-except:
-    # Prior python 2.6 json comes from simplejson
-    from simplejson import JSONEncoder
+from json import JSONEncoder
 
 class Page(object):
     """
@@ -87,26 +83,28 @@ class TemplatedPage(Page):
         self.debug("Templates are located in: %s" % self.templatedir)
         self.debug("Using Cheetah version: %s" % Version)
 
-    def templatepage(self, file=None, *args, **kwargs):
+    def templatepage(self, ifile=None, *args, **kwargs):
         """
         Template page method.
         """
-        searchList = []
+        search_list = []
         if len(args) > 0:
-            searchList.append(args)
+            search_list.append(args)
         if len(kwargs) > 0:
-            searchList.append(kwargs)
-        templatefile = "%s/%s.tmpl" % (self.templatedir, file)
+            search_list.append(kwargs)
+        templatefile = "%s/%s.tmpl" % (self.templatedir, ifile)
         if os.path.exists(templatefile):
-            template = Template(file=templatefile, searchList=searchList)
+            template = Template(file=templatefile, searchList=search_list)
             return template.respond()
         else:
-            self.warning("%s not found at %s" % (file, self.templatedir))
-            return "Template %s not known" % file
+            self.warning("%s not found at %s" % (ifile, self.templatedir))
+            return "Template %s not known" % ifile
 
 def exposexml (func):
     """CherryPy expose XML decorator"""
+    @expose
     def wrapper (self, *args, **kwds):
+        """Decorator wrapper"""
         data = func (self, *args, **kwds)
         if  type(data) is types.ListType:
             results = data
[…]

### 2010-05-04 src/python/DAS/analytics/analytics_base.py
Core of analytics framework.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1044 4525493e-7705-40b1-a816-d608a930855b

@@ -1,0 +1,177 @@
+#!/usr/bin/env python
+
+import multiprocessing
+import time
+import DAS.core.das_core
+
+class DASAnalyticsBase(multiprocessing.Process):
+    """
+    Basic managed analytics class. This is a process that is run quasi-independent of the core
+    daemon. The main loop automatically handles messages and timing, calling the analysis_run
+    method at a given interval.
+    """
+    stats = {}
+    defaults = {
+                'interval':300,
+                }
+    
+    @classmethod
+    def merge_defaults(cls,config):
+        """
+        Merge the supplied config with the defaults in this class.
+        """
+        default_copy = copy.deepcopy(cls.defaults)
+        default_copy.update(config)
+        return default_copy
+    
+    @staticmethod
+    def validate_config(config):
+        """
+        Stub to be overriden, if validation rules are required for the supplied config.
+        Return false if something is irredeemably wrong.
+        """
+        return True
+    
+    def __init__(self,name,config,das_config,controller_pipe):
+        """
+        Create a new analytic class. Subclasses should generally not override the details
+        of this class, but use the configure() method (called at __init__ time).
+        """
+        multiprocessing.Process.__init__(self)
+        self.config = config
+        self.das_config = das_config
+        self.das_core = DAS.core.das_core.DASCore(das_config)
+        self._pipe = controller_pipe
+        self.name = name
+        self.internal_interval = 1
+        self.analysis_interval = config['interval']       
+        self.log(0,"created")
+        self._ticks = 0
+        self._last_analysis_time = 0
+        self._analysis_calls = 0
+        self._min_analysis_time = 0
+        self._max_analysis_time = 1e30
+        self._total_analysis_time = 0
+        self._time_started = 0
+        self._stop = False
+        self.configure(config)
+    
+    def configure(self,config):
+        """
+        Stub for subclasses to perform any init-time configuration, before the main loop is entered.
+        """
+        pass
+        
+    def analysis_start(self):
+        """
+        Stub for subclasses to perform any actions when the main loop is first started.
+        """
+        pass
+    
+    def analysis_run(self):
+        """
+        Stub for subclasses to perform the main analysis action. Should be called every 'interval' seconds.
+        """
+        pass
+    
+    def run(self):
+        """
+        Actual main loop. This should not be overridden (use analysis_run), as it handles receiving pipe messages,
+        calling the analysis at correct intervals and exiting if necessary.
+        """
+        self.log(0,"started")
+        self._time_started = time.time()
+        self.analysis_start()
[…]

### 2010-05-04 src/python/DAS/analytics/analytics_controller.py

@@ -1,0 +1,448 @@
+#!/usr/bin/env python
+
+import DAS.core.das_robot
+import DAS.core.das_core
+import DAS.utils.das_config
+import multiprocessing
+import os
+import DAS.core.das_core
+import time
+import sys
+import analytics_utils
+import analytics_base
+import analytics_service
+import fnmatch
+
+class DASAnalyticsController(DAS.core.das_robot.Robot):
+    """
+    Analytics Controller
+    
+    This is the sole daemon process started for analytics, which spawns
+    a series of owned processes (using multiprocessing.Process). These
+    processes perform actual functions, and are divided into analysers
+    (which connect to the analytics DB and perform periodic calculations)
+    and services (which are assistance modules to the analytics daemon, such
+    as logging and web access).   
+    """
+    def __init__(self,das_config,analytics_config,daemon=True):
+        """
+        Create a new DASAnalyticController. 
+        das_config is a DAS.utils.das_config.das_config object.
+        analytics_config is a path to a analytics config
+        daemon determines whether to detach and fork.
+        """
+        self.das_config = das_config
+        self.config = self.read_config(analytics_config)
+        self.daemon = daemon
+        self.pidfile = self.config.pidfile
+        
+        self.analysers = {}
+        self.analyser_pipes = {}
+        self.services = {}
+        self.service_pipes = {}
+        self.all_pipes = {}
+        self.subscriptions = []
+        
+        self.current_heartbeat = 0
+        self.last_heartbeat = {}
+        
+        self.master_pipe = analytics_utils.FakePipe()
+        
+        self.timers = []
+        
+        self._stop = False
+    
+    def read_config(self,conffile):
+        """
+        Read the analytics config by executing the supplied file with, using a DASAnalyticConfig object as the global namespace.
+        """
+        config_object = analytics_utils.DASAnalyticsConfig()
+        exec_dict = {'analyser':config_object.analyser,'service':config_object.service}
+        execfile(conffile,exec_dict)
+        config_object.__dict__.update(exec_dict)
+        return config_object
+        
+    def load_modules(self,config):
+        """
+        Import all the modules defined in DASAnalyticsConfig.modules, then load the necessary classes.
+        Set up the pipes for communication, and add any message subscriptions defined by services.
+        Don't actually start them yet.
+        """
+        modules = []
+        for module in config.modules:
+            if '.' in module:
+                modules += [__import__(module,fromlist=[module.split('.')[-1]])]
+            else:
+                modules += [__import__(module)]
+        
+        def get_class(name):
[…]

### 2010-05-04 src/python/DAS/analytics/sample_config.py

@@ -1,0 +1,5 @@
+
+modules = ['services.logger']
+#analyser(classname,analysername,kwarg0=0,kwarg1=1)
+service('Logger')
+service('Logger','CoreLogger',output='/tmp/das_analytics_core_log',subscriptions=[('all','core','log')])

### 2010-05-04 src/python/DAS/analytics/services/logger.py

@@ -1,0 +1,40 @@
+import DAS.analytics.analytics_service
+import sys
+import time
+
+class Logger(DAS.analytics.analytics_service.DASServiceBase):
+    defaults = {
+                'interval':1,
+                'output':'/tmp/das_analytics_log',
+                'subscriptions':[('all','*','log')],
+                'flush_frequency':5
+                }
+    levels = {
+              0: 'INFO',
+              1: 'ERROR',
+              2: 'CRITICAL'
+              }
+    def configure(self,config):
+        self.output = open(config['output'],'w')
+        self.flush_frequency = config['flush_frequency']
+        self.flush_count = 0
+        
+        #self.output = sys.stdout
+    
+    def pipe_log(self,msg):
+        print 'pipe_log'
+        src = msg.get('src','')
+        level = msg.get('level',0)
+        t = msg.get('time',0)
+        text = msg.get('text','')
+        self.output.write('[%s] %s: %s: %s\n'%(time.strftime('%d-%m-%Y %H:%M:%S',time.gmtime(t)),src,self.levels.get(level,'OTHER'),text))
+        self.flush_count += 1
+        if self.flush_count%self.flush_frequency:
+            self.output.flush()
+        
+    def finalise(self):
+        self.output.close()
+        
+    
+        
+    

### 2010-05-04 src/python/DAS/core/das_mongocache.py
Added index on lookup-keys and clean-up code
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1502 4525493e-7705-40b1-a816-d608a930855b

@@ -5,22 +5,20 @@
 DAS mongocache wrapper.
 """
 
-__revision__ = "$Id: das_mongocache.py,v 1.5 2009/09/02 19:56:37 valya Exp $"
-__version__ = "$Revision: 1.5 $"
+__revision__ = "$Id: das_mongocache.py,v 1.6 2009/09/02 20:57:11 valya Exp $"
+__version__ = "$Revision: 1.6 $"
 __author__ = "Valentin Kuznetsov"
 
-import os
 import time
 import types
-import traceback
 
 # DAS modules
-from DAS.utils.utils import genkey, getarg, sort_data
-from DAS.utils.utils import dict_value, merge_dict
+from DAS.utils.utils import getarg, dict_value, merge_dict
 from DAS.core.cache import Cache
 
 # monogo db modules
 from pymongo.connection import Connection
+import pymongo
 
 def mongo_query(query):
     """Take das input query and convert it into mongo DB one"""
@@ -123,7 +121,7 @@ class DASMongocache(Cache):
             res = self.col.find(**query)
         for obj in res:
             row = obj.to_dict()
-            del(row['_id']) # mongoDB add internal _id of the obj, we don't need it
+            del(row['_id']) #mongo add internal _id, we don't need it
             fields = query['fields']
             if  fields:
                 fkeys = [k.split('.')[0] for k in fields]
@@ -141,14 +139,14 @@ class DASMongocache(Cache):
                 % query)
         if  not results:
             return
+        dasheader = header['das']
+        dasheader['selection_keys'] = header['selection_keys']
+        lkeys = header['lookup_keys']
+        index_list = [(key, pymongo.DESCENDING) for key in lkeys]
+        prim_key = lkeys[0] # TODO: what to do with multiple look-up keys
         if  type(results) is types.ListType or \
             type(results) is types.GeneratorType:
             for item in results:
-                dasheader = header['das']
-                dasheader['selection_keys'] = header['selection_keys']
-                # find out if cache contains already doc with primary key
-                lkeys = header['lookup_keys']
-                prim_key = lkeys[0] # TODO: what to do with multiple look-up keys
                 item['das'] = dasheader
                 entry = dict_value(item, prim_key)
                 res = self.col.find_one({prim_key:entry})
@@ -164,6 +162,11 @@ class DASMongocache(Cache):
                         self.col.insert(item)
                 else:
                     self.col.insert(item)
+                if  index_list:
+                    try:
+                        self.col.ensure_index(index_list)
+                    except:
[…]

### 2010-05-04 src/python/DAS/tools/das_cache_client.py
Tune up code to be complaint with new WebTools REST interface
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1503 4525493e-7705-40b1-a816-d608a930855b

@@ -4,8 +4,8 @@
 """
 DAS cache client tools 
 """
-__revision__ = "$Id: das_cache_client.py,v 1.13 2009/07/15 18:59:03 valya Exp $"
-__version__ = "$Revision: 1.13 $"
+__revision__ = "$Id: das_cache_client.py,v 1.14 2009/09/08 13:29:26 valya Exp $"
+__version__ = "$Revision: 1.14 $"
 __author__ = "Valentin Kuznetsov"
 
 import urllib2, urllib
@@ -89,19 +89,36 @@ if __name__ == '__main__':
     else:
         msg = 'You need to provide either input dict or query.'
         raise Exception(msg)
-    path    = '/rest/%s/%s' % (opts.format.lower(), opts.request.upper())
+#    path    = '/rest/%s/%s' % (opts.format.lower(), opts.request.upper())
+    if  opts.request.lower() == 'post':
+        method = 'create'
+    elif opts.request.lower() == 'put':
+        method = 'replace'
+    elif opts.request.lower() == 'delete':
+        method = 'delete'
+    else:
+        method = 'request'
+    path = '/rest/%s' % method
 
     if  host.find('http://') == -1:
         host = 'http://' + host
     url = host + path
-    if  request == 'GET':
+
+    headers={"Accept": "application/json"}
+    format = opts.format.lower()
+    if  format == 'json' or format == 'dasjson':
+        headers={"Accept": "application/json"}
+    elif format == 'xml' or format == 'dasxml':
+        headers={"Accept": "application/xml"}
+    else:
+        headers={"Accept": "application/%s" % format}
+
+    if  request == 'GET' or request == 'DELETE':
         encoded_data=urllib.urlencode(params, doseq=True)
         url += '?%s' % encoded_data
-        req = UrlRequest(request, url=url)
+        req = UrlRequest(request, url=url, headers=headers)
     else:
-#        encoded_data=urllib.urlencode(params, doseq=True)
         encoded_data=json.dumps(params)
-        headers={"Content-type": "application/json"}
         req = UrlRequest(request, url=url, data=encoded_data, headers=headers)
 
     if  debug:

### 2010-05-04 src/python/DAS/web/DASCacheModel.py

@@ -5,8 +5,8 @@
 DAS cache RESTfull model, based on WMCore/WebTools
 """
 
-__revision__ = "$Id: DASCacheModel.py,v 1.12 2009/09/01 01:42:48 valya Exp $"
-__version__ = "$Revision: 1.12 $"
+__revision__ = "$Id: DASCacheModel.py,v 1.13 2009/09/08 13:29:27 valya Exp $"
+__version__ = "$Revision: 1.13 $"
 __author__ = "Valentin Kuznetsov"
 
 # system modules
@@ -39,26 +39,43 @@ def checkargs(func):
     """Decorator to check arguments to REST server"""
     def wrapper (self, *args, **kwds):
         """Wrapper for decorator"""
-        # check request headers. If Content-type is set to JSON
-        # process the body request, extract JSON dict, convert it into
-        # set of parameters passed to HTTP method.
+        # check request headers. For methods POST/PUT
+        # we need to read request body to get parameters
         headers = cherrypy.request.headers
-        if  headers.has_key('Content-type'):
-            content = headers['Content-type']
-            if  content in ['application/json', 'text/json', 'text/x-json']:
-                body = cherrypy.request.body.read()
-                if  args and kwds:
-                    msg  = 'Misleading request.'
-                    msg += 'Headers: %s ' % headers
-                    msg += 'Parameters: %s, %s' % (args, kwds)
-                    return {'status':'fail', 'reason': msg}
-                jsondict = json.loads(body, encoding='latin-1')
-                for key, val in jsondict.items():
-                    kwds[str(key)] = str(val)
+        if  cherrypy.request.method == 'POST' or\
+            cherrypy.request.method == 'PUT':
+            body = cherrypy.request.body.read()
+            if  args and kwds:
+                msg  = 'Misleading request.\n'
+                msg += 'Request: %s\n' % cherrypy.request.method
+                msg += 'Headers: %s\n' % headers
+                msg += 'Parameters: args=%s, kwds=%s\n' % (args, kwds)
+                return {'status':'fail', 'reason': msg}
+            jsondict = json.loads(body, encoding='latin-1')
+            for key, val in jsondict.items():
+                kwds[str(key)] = str(val)
+
+#        headers = cherrypy.request.headers
+#        if  headers.has_key('Content-type'):
+#            content = headers['Content-type']
+#            cherrypy.response.headers['Content-type'] = content
+#            if  content in ['application/json', 'text/json', 'text/x-json']:
[…]

### 2010-05-04 doc/paper.tex
New addition
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1505 4525493e-7705-40b1-a816-d608a930855b

--- 2010-05-04 doc/paper.tex
@@
-% Template for Elsevier CRC journal article
-% version 1.0 dated 13 October 2009
-
-% This file (c) 2009 Elsevier Ltd.  Modifications may be freely made,
-% provided the edited file is saved under a different name
-
-% This file contains modifications for Procedia Computer Science
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% This template uses the elsarticle.cls document class and the extension package ecrc.sty
-%% For full documentation on usage of elsarticle.cls, consult the documentation "elsdoc.pdf"
-%% Further resources available at http://www.elsevier.com/latex
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% The '1p' and 'times' class options of elsarticle are used for Elsevier CRC
-\documentclass[1p,times]{elsarticle}
-
-%% The `ecrc' package must be called to make the CRC functionality available
-\usepackage{ecrc}
-
-%% The ecrc package defines commands needed for running heads and logos.
-%% For running heads, you can set the journal name, the volume, the starting page and the authors
-
-%% set the volume if you know. Otherwise `00'
-\volume{00}
-
-%% set the starting page if not 1
-\firstpage{1}
-
-%% Give the name of the journal
-\journalname{Procedia Computer Science}
-
-%% Give the author list to appear in the running head
-%% Example \runauth{C.V. Radhakrishnan et al.}
-\runauth{}
-
-%% The choice of journal logo is determined by the \jid and \jnltitlelogo commands.
-%% A user-supplied logo with the name <\jid>logo.pdf will be inserted if present.
-%% e.g. if \jid{yspmi} the system will look for a file yspmilogo.pdf
-%% Otherwise the content of \jnltitlelogo will be set between horizontal lines as a default logo
-
-%% Give the abbreviation of the Journal.
-\jid{procs}
-
-%% Give a short journal name for the dummy logo (if needed)
-\jnltitlelogo{Procedia Computer Science}
-
-%% Hereafter the template follows `elsarticle'.
-%% For more details see the existing template files elsarticle-template-harv.tex and elsarticle-template-num.tex.
-
-%% Elsevier CRC generally uses a numbered reference style
-%% For this, the conventions of elsarticle-template-num.tex should be followed (included below)
-%% If using BibTeX, use the style file elsarticle-num.bst
-
-%% End of ecrc-specific commands
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% The amssymb package provides various useful mathematical symbols
-\usepackage{amssymb}
-%% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
-
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers after \end{frontmatter}.
-%% \usepackage{lineno}
-
-%% natbib.sty is loaded by default. However, natbib options can be
-%% provided with \biboptions{...} command. Following options are
-%% valid:
-
-%%   round  -  round parentheses are used (default)
-%%   square -  square brackets are used   [option]
-%%   curly  -  curly braces are used      {option}
-%%   angle  -  angle brackets are used    <option>
-%%   semicolon  -  multiple citations separated by semi-colon
-%%   colon  - same as semicolon, an earlier confusion
-%%   comma  -  separated by comma
-%%   numbers-  selects numerical citations
-%%   super  -  numerical citations as superscripts
-%%   sort   -  sorts multiple citations according to order in ref. list
-%%   sort&compress   -  like sort, but also compresses numerical citations
-%%   compress - compresses without sorting
-%%
-%% \biboptions{comma,round}
-
-% \biboptions{}
-
-% if you have landscape tables
-\usepackage[figuresright]{rotating}
-
-% put your own definitions here:
-%   \newcommand{\cZ}{\cal{Z}}
-%   \newtheorem{def}{Definition}[section]
-%   ...
-
-% add words to TeX's hyphenation exception list
-%\hyphenation{author another created financial paper re-commend-ed Post-Script}
-
-% declarations for front matter
-
+\documentclass[a4paper]{jpconf}
+\usepackage{graphicx}
 \begin{document}
-
-\begin{frontmatter}
-
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for the associated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for the associated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for the associated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%%
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \address{Address\fnref{label3}}
-%% \fntext[label3]{}
-
-\dochead{}
-%% Use \dochead if there is an article header, e.g. \dochead{Short communication}
-
 \title{The CMS Data Aggregation System}
 
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{<author name>}
-%% \address[label1]{<address>}
-%% \address[label2]{<address>}
-
-%\author[vkuznet]{Valentin Kuznetsov\corref{cor1}}
-\author[vkuznet]{Valentin Kuznetsov}
-\address[vkuznet]{Cornell University, Ithaca, New York, USA}
+\author{Valentin Kuznetsov}
+\address{Cornell University, Ithaca, New York, USA}
 \ead{vkuznet@gmail.com}
 
-\author[evans]{Dave Evans}
-\address[evans]{Fermilab, Batavia, Illinois, USA}
+\author{Dave Evans}
+\address{Fermilab, Batavia, Illinois, USA}
 \ead{evansde@fnal.gov}
 
-\author[metson]{Simon Metson}
-\address[metson]{Bristol University, Bristol, UK}
+\author{Simon Matson}
+\address{Bristol University, Bristol, UK}
 \ead{s.metson@bristol.ac.uk}
 
-%\cortext[cor1]{Corresponding author}
+%In a large modern enterprise, information is almost inevitably distributed among several database management systems.
+Despite considerable attention from the research community, relatively few commercial systems have attempted to address this issue.
+This article describes the technology that enables clients of IBM's federated database engine to access and integrate the data and specialized computational capabilities of a wide range of relational and non­relational data sources.
 
 \begin{abstract}
-Metadata plays a significant role in large modern enterprises, 
-research experiments and digital libraries where it comes from many different 
-sources and is distributed in a variety of digital formats. 
-It is organized and managed by constantly evolving software using 
-both relational and non-relational data sources. Even though we can apply
-an information retrieval approach to non-relational data sources,
-we can't do so for relational ones, where information is accessed via
-a pre-established set of data-services.
-
-Here we discuss a new data aggregation system which consumes, 
-indexes and delivers information from different relational and 
-non-relational data sources to answer cross data-service queries 
-and explore metadata associated with petabytes of experimental data. 
-We combine the simplicity of keyword-based search with the precision of RDMS
-under the new system. The aggregated information is collected from various sources,
-allowing end-users to place dynamic queries, get precise answers and 
-trigger information retrieval on demand. Based on the use cases of the CMS experiment, 
-we have performed a set of detailed, large scale tests the results of which 
-we present in this paper.
+A metadata plays significant role in a large modern enterprises, research experiments,
+digital libraries where it comes from different sources and distributed in a 
+variety of forms and digital formats. It is organized and managed by constantly
+evolving software using both relational and nonrelational data sources. There is
+a big demand to access information from multiple sources.
+Here we discuss a new data aggregation system which deliver and cache information 
+from different relational and nonrelational data sources on a concrete example 
+of large scale High-Energy physics experiment.
 \end{abstract}
 
-\begin{keyword}
-%% keywords here, in the form: keyword \sep keyword
-Meta-data\sep data aggregation\sep information discovery\sep HEP.
+\newpage
 
-%% PACS codes here, in the form: \PACS code \sep code
+\section{Introduction}
+The CERN, the European Organization for Nuclear Research, plays a leading
+role in fundamental studies of physics. But apart from scientific community 
+it is known as a place where World Wide Web was born. At that time the 
+information look-up among hyperlinks represents a certain challenge for scientists.
+Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
+Physics (HEP), promising to deliver a few PB of data each year. Scientists are
+facing another set of problems from distributed computing to multi-dimentional
+view on information retrieval. One of the aspect of such research is an efficient
+and at the same time precise look-up of produced meta-data, which comes in variety 
+of forms and digital formats. As was pointed out in \cite{Amr} a mixed content and 
+mixed metadata and metadata consistency should be considered as a whole in design 
+of the system to successful information discovery.
 
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-\end{keyword}
+The CMS, the Compact Muon Solenoid experiment, operated at LHC,
+represents an heterogeneous environment of distributing computing, relational and
+nonrelational data sources where we face this problem. The computing resources
+are distributed among almost 40 countries, 183 institutions and more then 3000 physicists.
+Broad variety of RDMS systems at different data centers collect a various
+metadata information, including data location and transfer, detector conditions,
+calibrations constants, simulation information, etc. Moreover the development
+of those components were done in parallel by different group and technology
+tools. Therefore a wide spectrum of data services and
+data formats represents a problem of information discovery.
 
-\end{frontmatter}
+\section{Related Work}
+Even though the idea of querying relation databases via keyword based search
+algorithm is not knew it is still under significant activity in computer
+sceince domain. A few
+alternative solutions has been proposed to address this issue. On one
+side, the federated DB \cite{FedDB} unify data coming from different
+RDMS into another DB where SQL queries can be placed to search desired
+data. While another approaches of querying relational
+DB(s) using keyword search algorithms has been proposed recently, see
+\cite{DBXplorer, QueryAnswer}.
+In former case, you still face with understanding of the underlying schema and
+imposing relational conditions in your query, while queries can be
+expressed via simple keyword based search, even though exact match
+of provided keywords is expected. Although keyword based search is
+intuitive and easy to use, with respect to HEP domain it is not sufficient. 
+For example, how to deal with numbers? What a single number means in a user input?
+The row ID, the value of the field, the sum of files or their size? In other
+words a further step is required to understand up-coming request without
+sacrifies the output results match. Our users can be classified as proficient
+users who understand semantics of the problem and want precise answers
+for their queries. They are more interesting in precise answers to their
+questions rather ranking output.
 
-%%
-%% Start line numbering here if you want
-%%
-% \linenumbers
-
-%% main text
-\section{Introduction}
-The European Organization for Nuclear Research, known as CERN, plays a leading
-role in fundamental studies of physics. It is also known as a place where
-many innovations in the area of computer science were developed, e.g. The World Wide Web.
-Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
-Physics (HEP), promising to deliver a few PB of data each year. 
-At this scale the ``information discovery'' within a heterogeneous, distributed 
-environment becomes a key ingredient of successful data analysis.
-The data and associated meta-data are produced in variety of forms and digital formats.
-They are stored and retrieved from relational and non-relational data-sources, such as 
-RDMS systems, document oriented databases, blogs, twikies, file systems,
-customized applications, etc. Working in such an environment requires a lot 
-of expertise and demands a straightforward, simple
-way to look-up the desired information.
-A well-known solutions, e.g. data-services, are tighten to a specific data 
-sources and end-users are left with a manual task to bookkeep and 
-relate information from them.
-
-Here we present our work on Data Aggregation System (DAS) which provides
-ability to search and aggregate information across different 
-data-services. While designed for CMS High-Energy Physics 
-experiment at LHC, the strategies and technology could be
-used elsewhere.
-
-The rest of this paper is organized as follows. 
-In section \ref{RelatedWork} we discuss related work in a domain of 
-keyword search over relational data sources.
-Section \ref{DataModel} describes the CMS experiment and its data model. In section
-\ref{DAS} we present the architecture of the DAS system, including discussion of its
-various components. Finally, our results are summarized in section \ref{Results}.
-
-\section{Related Work\label{RelatedWork}}
-Even though the idea of keyword queries over the relational database
-is not knew it is still under discussion in computer science domain.
-A few alternative solutions have been proposed to address this issue
-in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
-i.e. retrieval of documents that contain all query keywords, has
-been discussed. It requires that all keywords are matched in a row from a
-single table or joined tables. Authors of \cite{QueryAnswer} make a step forward
-and have evaluated information around matched values. The results were presented as
-customized templates to end-users. 
-These and other related works, (see \cite{DBXplorer, QueryAnswer} 
-for more references), are based on an approach that builds some 
-kind of the graph over the underlying 
-database and generates a new database out of the source database to perform the work.
-The proximity of the results in those techniques are mostly confined to a 
-text-based values stored in a database, since the usage of numerical ones 
-is mostly useless without knowledge of the context of the input value. 
-For example, typing a value 100 in the search input field can lead to 
-plenty of inappropriate matches, e.g. row ids. In order to be
-answered correctly it requires an additional keyword to clarify its meaning, 
-which should be matched with particular entity (table.column) in the 
-underlying database. Therefore the logical questions, e.g.
-{\it find me the total number of files whose size more then 20 but less then 
-100} are hard to answer precisely using approaches outlined in \cite{DBXplorer, QueryAnswer}, 
-while it can be easily accomplished using SQL.
-
-To address these limitations a keyword-based query language was introduced in \cite{DBS-QL}.
-It used a set of pre-defined keys mapped to an underlying schema and shortest path
-algorithm which employed foreign key constraints to build SQL queries. As a result
-the question
+In \cite{DBS-QL} a simple, intuitive and flexible query language was introduced 
+on top of the CMS data-bookkeeping system. It represented a power of SQL while
+hiding underlying relational schema from the end users. As a results
+a human questions were intuitively mapped into simple queries. For example,
+a question
 {\it I'm looking for files who contain data taken on certain date and located at
-particular site} was represented as simply as \cite{DBS-QL}
+particular site} was represented as simple as
 \begin{verbatim}
@@
 \end{verbatim}
-without specifying any relational details.
-This solution represents a common use case in the HEP community and provides
-an intuitive mapping between the mental model of the end-users and 
-the underlying database schema.
+We wanted to expand this approach and apply such queries across multiple
+data services and different DB back-ends.
 
-At the same time the data aggregation is commonly addressed in the context of
-federated databases, see for example \cite{FedDB}. The data from different 
-RDMS systems are stored into the federated DB where SQL queries can be used to 
-search desired data. Apart from many internal obstacles, such as
-schema, domain and naming conflicts, it is external issues that are often
-a large barrier within a distributed, heterogeneous environment where
-different security policies are in place. At the end the usage of SQL and/or
-any other Query Language is required to search for data.
-
-To avoid these limitations we resolved to build a keyword-search based system around
-existing data-services whose nature and access policies are in place and well understood.
-The key ingredients were the setup of the mapping and analytics services which provided
-the necessary glue across existing API/notations and user queries. This approach 
-allows us to perform dynamic data aggregation across various data-services, achieve
-precise answers using keyword-based queries
-and organize information retrieval on demand.
+Here we discuss a new system, the Data Aggregation System (DAS),
+developed in CMS collaboration to address this issue. We start with
+discussion of CMS data model and existing data-services, see \ref{DataModel}. 
+In section \ref{DAS}
+we provides a detailed description of underlying design and components
+used in DAS system. Our results are summarized in section \ref{Results}.
 
 \section{CMS data model\label{DataModel}}
-The Compact Muon Solenoid, (CMS) \cite{CMS} 
-is one of the two general purpose particle physics detectors built on 
-the Large Hadron Collider (LHC) at CERN in Switzerland and France. 
-The experiment is designed to explore the frontiers of physics and provide physicists
-with the ability to look at the conditions presented in the early stage of our Universe.
-More then 3000 physicists from 183 institutions representing 38 countries 
-are involved in the design, construction and maintenance of the experiment.
-
-The CMS distributed computing and data model \cite{CMSDataModel} 
-is designed to process and efficiently manage the few PBs of data expected each year
-during normal operation of the LHC. The computing resources available to the CMS
-collaboration are geographically distributed, 
-interconnected via high throughput networks and operated by means 
-of various Grid software infrastructure. The model is designed to
-incorporate a broad variety of both mass data storage and data processing
-resources into a single coherent system. CMS uses
-multi-Tiered model, where specific tasks related to data taking such as
-processing, archival and distribution, are assigned to each tier based
-on the CMS data model. For example, the Tier-0 center at CERN is responsible
-for archiving the data coming out from the detector, prompt first pass reconstruction
-and data distribution to Tier-1 centers, located around the world.
-The Tier-1 centers provide archival storage and CPU power for high
-priority selection and re-reconstructions jobs and distribution points
-to the Tier 2 resources.
-The Tier-2 centers are dedicated for user analysis tasks and production of simulated data.
-
-A broad variety of data-services have been designed and developed to
-maintain detector and production operations, including detector
-conditions databases, data bookkeeping services,
-data transfer and job monitoring tasks. While the majority of these
-services are located at CERN, they need to be accessed remotely, and
-combined with other distributed services.
-
-Once such a conglomerate of data-services start operating the obvious
-question arises: how do we find the desired information across multiple data-services
-in our distributed environment? Even though the individual data-services were designed
-to answer specific questions about the data they serve, the ability to search and relate
-information among them was remains a tedious human task. The growing amount of information
-and desire to make cross-service queries led us to design and develop the new
-Data Aggregation System.
-
 \section{Data Aggregation System\label{DAS}}
-The goal of the Data Aggregation System (DAS) was to fetch and aggregate meta-data 
-on demand from existing CMS data-services under one umbrella by using their APIs, 
-security policies, etc. In contrast, the approaches
-discussed in \cite{DBXplorer, QueryAnswer, FedDB}
-enforce a publication step of data into a dedicated database 
-or central repository \cite{iRODS}. Therefore, quite often it was required
-that data should be converted into common data structure/format in order to enable
-aggregation, see \cite{OpenArchive}. In DAS the data are provided
-by existing data-services and infrastructure, whose implementation,
-data representation and security policies are left intact. Consequently,
-DAS does not require a data preservation and transaction capabilities. 
-Instead DAS can be treated as proxy or caching layer where information 
-retrieval and aggregation are triggered dynamically by user queries. 
-Below we discuss in details the DAS architecture.
-
-\subsection{DAS architecture}
-The design of the DAS system was based on previous studies of CMS Data 
-Bookkeeping System (DBS) \cite{DBS, DBS07}. The DBS Data Discovery service
-\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
-a syntax similar to SQL, without specific knowledge of the underlying DB schema.
-Quick adoption and wide usage of DBS-QL in CMS gave us confidence in the chosen approach 
-and provided a basis for building the DAS system.
-
-As was pointed out in \cite{Arms} the mixed content and 
-mixed meta-data and meta-data consistency 
-of the underlying services needs to be carefully
-considered in the design of the system if it is to be a successful information 
-discovery service.
-Starting from this ground we are designed the DAS system as an
-additional layer on top of the existing data-services
-within CMS computing infrastructure by imposing the following set of requirements:
-\begin{itemize}
-\item support a keyword based search queries with ability to use conditional operators;
-%\item pluggable interface to existing and up-coming data-service APIs;
-\item use existing heterogeneous software environment and distributed nature of the data-services;
-\item preserve access rules and security policies of individual data-services;
-%\item be flexible to adjust to dynamic nature of data-service APIs, 
-%due to continuously changing detector and software environments;
-\item retrieve data on demand and aggregate them if necessary across
-multiple data-services;
-\item be transparent to data-services implementation and their data formats;
-%\item support legacy data-services, applications and APIs.
-\end{itemize}
+Data Aggregation System shown in Figure \ref{DAS_arch} and
+consists out of several components:
 \begin{figure}[htb]
 \centering
-\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
+\includegraphics[width=150mm]{DAS_architecture.pdf}
 \caption{
-The DAS architecture diagram. It conists of cache server,
-analytics and mapping databases. The data in a cache can be shared
-across multiple queries. The information retrieval is triggered
-on demand basis by invoking appropriate data-service APIs. The
-optional DAS robots (Unix daemons) can be used to pre-fetch
-the most populary query requests.
+DAS architecture.
+}
+\label{DAS_arch}
+\end{figure}
+
+\begin{itemize}
+\item DAS core system which is support pluggable modules to retrieve data from
+CMS data-services, e.g. Data Bookkeeping System (DBS), data location service (Phedex),
+etc.;
+\item DAS caching system which has dual purpose: to store ``raw'' results coming
+out from data-services, a raw-cache, and aggregated results produced by DAS core into
+hot-cache;
+\item DAS request service who collect and place into queue user requests
+\item DAS mapping DB to store mapping between data-services and DAS notations and
+data-representation;
+\item DAS analytics DB to collect query statistics and establish pre-fetching 
+strategies.
+\end{itemize}
+We will discuss each of those components in more details in up-coming sub-sections.
+The DAS supports a pluggable approach to add a new CMS sub-system into a game by
+identifying appropriate APIs, input and output parameters and keys, data-service
+format and mapping between data-service, DAS and other sub-system. For instance,
+the Data Bookkeeping System (DBS) and data location system (Phedex) store information about
+a file. In former case, a file name was stored as well as its relations to other physics 
+objects, such as run and luminosity, while in later case the Phedex store information 
+about file location. When user specify a file as their selection key, based on
+information store in DAS mapping DB we were able to identify appropriate set of
+data-service APIs, look-up data, re-format their output to DAS notations and
+store it into DAS raw-cache. Even though we support a data retrieval on demand
+basis we also store user queries into DAS analytics DB to identify
+a common patterns. Based on this information we organized a pre-fetching
+populators. Once data were coming to raw-cache an additional step of
+aggregation was applied. If other data with identical keys were presented in
+a raw-cache we merge it them together. For instance upon the following query
+$find file where ... $ a file information from DBS and Phedex systems were
+retrieved. Since both data-services output contained file name as a common
+key, their output where merged and final result was stored into raw-cache.
+This merged (aggregated) results were shipped back to client and stored into
+hot-cache (memcached) to speed up a DAS web client application. Each DAS data
+object contains a standard DAS header, which identify look-up time,
+request url, API, parameters and method as well as response version, expiration
+time and checksum of the output.
+
+\subsection{DAS Query Language}
+Upon success of DBS-QL \ref{DBS-QL} we adopted its syntax to DAS with a few minor
+changes. Here we briefly recall its syntax:
+\begin{equation}
+action\,\,\,
+key_1,\, key_2,\, ...\,\,\, where\,\,\,
+\langle key\rangle\, 
+\langle op\rangle\, 
+\langle value\rangle \,\,\, and|or ...
+\label{QL_syntax}
+\end{equation}
+Each DAS QL expression starts with special keyword indicating an action such as 
+$find,\, plot,\, summary$, followed by a set of selection keys. The action identify
+the behavior of the query, find to look-up data, plot to plot found data, summary to
+apply a summary view over provided selection keys. An optional conditions can be 
+specified via $where$ keyword following by condition expression.
+Each condition expression represented via $key\,\,\, operator\,\,\, value$ triplet, where
+$key$ is DAS selection key, $operator$ is conditional operator, 
+$=,\, >=,\, <,\, in,\, between$ and $value$ is a value applied to selection key.
+We also support brackets as well as $and$ and $or$ logical operators.
+The DAS selection and condition keys were identified based on analysis 
+of terminology used by our users. For instance, $file$ was used to either
+select file name or used in condition to identify file name provided by user.
+Here is a simple example of DAS QL expression:
+$$
+find\,\,\, file,\, run\,\,\, where\,\,\, site=T1\_CH\_CERN
+$$
+which represents a natural language question: 
+{\it find files and runs located at T1\_CH\_CERN site}.
+
+\subsection{DAS Mapping DB}
+As been discussed above we used a DAS mapping DB to map data-service API into
+DAS as well as map provide a relations between different data-services.
+Upon identification of APIs which participate in DAS activity 
+we store a mapping between data-service notations
+into DAS mapping DB. This allows to map user input query parameters into
+data-service API metrics as well as map back data-service output into DAS names.
+For instance, the DBS system used $logical\_file\_name$ notation, while in Phedex
+the same entity was named as $lfn$. Since both of them were referred to
+logical file name, we map both notations into $file$ notation in DAS.\footnote{We
+also deal with physical file names, but still used $file$ to refer to them in
+DAS notation, since simple regular expressions were used to identify $file$
+meaning from the provided value.}
+For each service we provide a mapping from its own data-format (JSON, XML, etc.) into
+common data-format used in DAS.\footnote{We used JSON as a DAS data format}
+We also store a mapping between DAS keywords and data-service API names to
+establish appropriate calls to data-service upon provided DAS query. For example,
+when user specifies $file$ in their DAS query, we invoke DBS listFiles and Phedex
+fileReplicas APIs. Since all mapping kept in RDMS we were able to abstract the 
+DAS core library calls from hardcoded data-service API calls and make it
+transparent to any participating data-services. Here is a logic of DAS core
+library:
+\begin{itemize}
+\item parse input DAS query
+\item map provided selection and condition keys into data-service APIs and
+input parameters
+\item invoke data-service APIs
+\item parse output results by mapping output notation into DAS keys
+\item store and aggregate them into DAS raw-cache
+\end{itemize}
+
+\subsection{DAS Analytics DB}
+
+\subsection{Caching system}
+The DAS caching system was divided into two different caches: raw-cache and hot-cache.
+The purpose of raw-cache was to store results coming from data services for
+as long as specified in data-service response header. While the purpose of hot-cache
+was to keep final DAS results corresponding to provided user query. We use a populator
+services (jobs) to feed raw-cache with highly requested queries. Here we will only
+discuss the raw-cache organization. 
+\begin{figure}[htb]
+\centering
+\includegraphics[width=150mm]{DAS_cache.pdf}
+\caption{
+DAS cache system.
 }
@@
 \end{figure}
+As it shown in figure \ref{DAS_cache} we setup
+dedicated DAS cache request server to manage user requests. Each request placed into
+queue who was used by workers to spawn a data-service request. 
+This allows us to utilze a multi-core technology and place a workers at each 
+core.\footnote{A number of workers per core were adjustable to satisfy requests load.}
+Each worker talk to DAS raw-cache and either stored mapped results over there or merge
+them with existing data object in DAS cache. We evaluated several technologies
+as a choice for DAS cache back-end, including relational MySQL DB \cite{MySQL} 
+and object oriented CouchDB \cite{CouchDB} and MongoDB \cite{MongoDB}.
+Each of this technologies had its own pros and cons. For instance usage
+of MySQL provided a SQL queries (as in federated DBs, see \cite{FedDB}), but
+required auto-schema generation on a fly.\footnote{We did not apply any
+restriction to DAS QL and did not know in advance what our users will be
+querying.} While, using CouchDB and MongoDB, we were able to store
+data-service output naturally as data objects into DB. Finally we decided
+to use MongoDB due to its flexible and
+powerful querying syntax, which were easy to map into DAS QL expressions.
 
-\noindent
-The choice of existing RDMS as DAS back-end(s) were ruled out for several reasons. 
-For instance, we didn't require any transactions and data persistency in DAS;
-a dynamically typing of stored meta-data objects was one of the requirements, etc.
-Those arguments forced us to look for alternative IT solutions.
-Through the careful analysis of available options, such as a file-based and in memory caches, 
-key-value databases, documented-oriented databases we made our choice in favor 
-of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
-MongoDB \cite{MongoDB}. Both systems are ``schemaless'' databases providing
-arbitrary document structure storage, replication and fail-over features. 
-We made our decision in favor of MongoDB, due to its support of dynamic queries, 
-full indexes, including inner objects and embedded arrays,
-and auto-sharding. Our preliminary benchmarks have shown that it can sustain
-the desired load and size capable of handling our meta-data information. We used 
-document-oriented database, MongoDB, as a back-end for three 
-main DAS components: Mapping, Analytics and DAS cache databases, 
-which will be discussed next.
-
-The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of
-the core library with support of pluggable modules for data retrieval;
-caching layer to store and aggregate results;
-request service to handle user queries;
-Mapping DB to keep information about data-service APIs, e.g.
-notations and mapping to DAS keys;
-Analytics DB for query analysis.
-
-\subsection{DAS core\label{DAS_core}}
-The DAS core library is responsible for data retrieval on demand. 
-It has the following components: input query validator and parser,
-query analyzer, workflow builder, execution engine and data-service 
-plug-ins. The communication between client (web server or CLI) and 
-DAS cache server was done via the REST \cite{REST} interface. 
-Each input query was recorded into the Analytics DB for further 
-analysis, and was decomposed into a set of conditions and/or selection keys.
-First the results were looked up in a cache and delivered to the requestor
-if superset of input conditions was found in analytics DB.\footnote{It is important
-to note here that we didn't look-up results in a cache and rather
-compared input query condition set to the records in analytics DB 
-in order to understand if other API calls are covered the input request.}
-Otherwise, the appropriate workflow, a set of API calls, was constructed
-for the execution engine. It invoked the API calls via data-service
-plug-ins and stored results into the cache. At this step we also
-performed necessary data aggregation for records with similar key-value
-pairs. The data-service plug-in implementations are similar and their 
-logic is largely abstracted, apart from a few extentions such as specific
-security modules for data-services accessible via private network.
-Finally, the results were retrieved from the cache and delivered to the user.
-
-\subsection{DAS Mapping DB\label{MappingDB}}
-The Mapping DB is used to track services known to DAS and how to interact with them.
-%The Mapping DB represents DAS {\it registration} service. 
-It collects
-information about data-service APIs, translates API input/output
-parameters into DAS keys and contains mapping between DAS records
-and their UI representation.
-
-On identifying API's that are accessible to DAS we store
-their names, input parameters and associated DAS keys
-into the Mapping DB. For example
-\begin{verbatim}
-{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
- "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
-              "api_param": "name", "das_key": "site"}], 
- "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
-              "key": "site", "map": "site.name"}]}
-\end{verbatim}
-represents DAS mapping record.\footnote{
-The data-service URL's are kept in the configuration file.
-}
-This record can be read as following:
-the \verb+CMSNametoSE+ API provided by 
-\verb+sitedb+ data-service accepts \verb+name+ as input parameter. It
-confirms regular expression pattern, \verb+re.compile('^T[0-3]_')+, and
-corresponds to \verb+site+ DAS key. The \verb+site+ key
-maps into \verb+site.name+ within DAS record.
-
-Therefore the Mapping DB identifies a set of pre-defined DAS keys, 
-e.g. \verb+site, file, run+, etc., to be used by end-users in their queries.
-Each key is mapped into set of API calls, whose output is retrieved,
-cached and aggregated as required.
-
-\subsection{DAS Analytics DB}
-The DAS Analytics DB collects information on user requests against the system.
-Each request that DAS handles is recorded.
-The Analytics DB tracks:
-\begin{itemize}
-\item The users request
-\item How DAS has mapped the request onto data service API's
-\item The time taken by the remote data services to process the API calls
-\end{itemize}
-By recording this information we can plan pre-fetch strategies for common queries,
-identify issues in remote data services and cross check that DAS resolves requests
-in a deterministic manner.
-
-% Upon its
-%decomposition into set of selection keys and conditions we also recorded
-%API name and input parameters into the Analytics DB and 
-%kept updating their counter information for repeated calls. 
-%This allowed to keep track of frequency of API calls and establish 
-%pre-fetch strategies for most common queries.
-
-\subsection{DAS caching system}
-The DAS caching system is used to dynamically fetch
-and aggregate data upon user requests. We use the following 
-algorithm (code snippets are written in Python \cite{Python}):
-\begin{enumerate}[1.]
-\item Decompose query into set of conditions (\verb+conds+) 
-and selection keys (\verb+skeys+).
-\item Check cache for query hash. If found, look-up results from the cache,
-otherwise proceed to the next step.
-\item Identify list of APIs which can serve requested data
-\begin{verbatim}
-apilist = [r for r in mapping_db(skeys, conds)]
-\end{verbatim}
-Here \verb+r+ represents a triplet \verb+url, api, args+ and 
-\verb+mappind_db+ defines translation between input query and data-service
-APIs which can delivery the data.
-\item Iterate over APIs
-\begin{verbatim}
-for url, api, args in apilist:
-    if  superset_in_analytics(api, args):
-        continue
-    for row in parser(get_data(url, api, args)):
-        yield row
-\end{verbatim}
-Here \verb+superset_in_analytics+ checks if provided \verb+api, args+
-are subset of already registered in Analytics DB api call(s). The 
-\verb+get_data+ and \verb+parser+ are define methods to fetch and parse
-data, respectively.
-\item Perform aggregation by iterating over records from the previous step
-\begin{verbatim}
-if cache_has(daskey):
-   existing_record = get_from_cache(daskey)
-   merge(row, existing_record)
-   insert_into_cache(row)
-   delete_from_cache(existing_record)
-else:
-   insert_into_cache(row)
-\end{verbatim}
-The \verb+daskey+
-is a key which identifies the record defined in Mapping DB, e.g. \verb+file.name+
-for \verb+file+ record.
-We did use a bulk insert operations at this step to increase throughput. 
-\item Get results from the cache
-\end{enumerate}
-We also have an option to run independent set of daemons, DAS robots, who
-monitor Analytics DB for most popular requests and pre-fetch the most
-common data into the cache.
-
-\subsection{DAS queries}
-All DAS queries are expressed in a free text-based form, either as a 
-set of keywords or key-value pairs, where a pair can represent
-condition, e.g. \verb+key > value+. We use a pre-defined set of keys registered 
-in DAS to describe common entities of our data. For example, \verb+dataset+, 
-\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
-is based on the analysis of daily work operations of our users, see \cite{DBS07}.
-Each key represents a DAS record. Every record is stored into the cache
-in JSON format. Due to the ``schemaless'' nature of underlying back-end (MongoDB)
-we are able to store DAS records of arbitrary structure, e.g.
-dictionary, lists, key-value pairs, etc. Therefore every DAS key has a 
-set of attributes representing JSON structure. For instance a document
-\begin{verbatim}
-{"block": {"name": "abc", "size": 1, "replica": [...]}}
-\end{verbatim}
-is associated with a \verb+block+ key who has the following attributes
-\verb+name+, \verb+size+, \verb+replica+. The set of attributes is based 
-on a set of meta-data information stored in a record and therefore
-a nested structures are allowed, e.g. \verb+block.replica.size+. Our users 
-are able to specify a context of their queries by using appropriate 
-attribute(s). For example, query \verb+block+ represents search for all
-block records rather then ``block'' string, while query
-\verb+block.name=abc*+ instructs the system to look-up appropriate blocks
-with matched names. The numerical values, conditions, grouping are
-easy to express too, e.g. \verb+run > 10 and run < 100+ query 
-is used to get all run records for provided range of run numbers. 
-To complement the strength of the queries, we do plan to index 
-dynamically DAS records and add a keyword-based 
-search within record content later.
-
-All input queries are transformed into MongoDB query syntax who is itself
-a JSON record. It contains a set of selection 
-keys (\verb+fields+) and conditions (\verb+spec+). For example, a query
-\verb+dataset, file=abc, run>10+
-is represented as
-\begin{verbatim}
-{"fields": ["dataset"], "spec": {"file": "abc", "run": {"$gt": 10}}}
-\end{verbatim}
-Therefore users queries are stored into Analytics DB similar to other DAS 
-records and used for further analysis.
-
-\subsection{Data-services and aggregation}
-As was shown in section \ref{MappingDB} the DAS Mapping DB became an authorative
-source of information about all data-services participating in DAS.
-Based on the information stored in Mapping DB we were able to 
-retrieve appropriate data from different data-services, re-map them into 
-DAS notations and aggregate them on demand. For instance, the Data 
-Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
-provide information about CMS blocks. In former case, DBS 
-stored information about blocks and their relationships to other 
-physics objects, such as dataset and files, while PhEDEx kept
-information about its location and replicas.
-
-Upon a user request to 
-find a block meta-data, we were able to identify appropriate set of 
-DBS and PhEDEx APIs, retrieve data, re-format their output 
-into DAS notations and store results into DAS cache. The records
-were merged (aggregated) based on identical key-value pairs, e.g.
-based on \verb+block.name+ key. Here is an example of such an aggregated record:
-\begin{verbatim}
-{"das_id": ["4b195404e2194e21c4000030", "4b1953fee2194e21c4000002"], 
- "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
-            "subscribed": "n", "custodial": "n", "site": "T1_US_Buffer", 
-            "nfiles": "18", "se": "srm.site.com", 
-            "size": "112694171838"}, {....}]},
-           {"name": "a/b/c#123", "nevents": "570200", 
-            "created_by": "USER_DN", "path": "/a/b/c", ...}],
- "das": [{"expire": 1259952908}, {"expire": 1259954702}]}
-\end{verbatim}
-The \verb+das_id+ refers to api call records stored in Analytics DB,
-the \verb+block+ part represents an aggregation block meta-data information from DBS
-and PhEDEx systems and the \verb+das+ part contains their expiration time stamps.
-Usage of this information was valuable contribution in debugging process of
-data-services themselves, e.g. identification of data-service inconsistencies, 
-latency studies, etc.
+\subsection{REST interface}
+We designed DAS interface to be fully REST complaint \cite{REST}.
+All requests placed into DAS cache server were done via GET/PUT/POST/DELETE methods.
+DAS clients initiate their request via GET method. An appropriate message
+with either resulting data or their absense was send back to client.
+Clients were able to decided upon received message how to proceed, either
+request data via POST/PUT methods, DELETE data in cache or initiate another
+GET request. The REST complaince help us to treat our clients, user quyering
+DAS system via web browser, other CMS applications and command line tools
+in identical manner. Also, the implementation of pre-fetch populator
+tools become as easy and identical as any other clients.
 
 \section{Results\label{Results}}
-In this section we discuss preliminary results achieved with DAS system
-using the following CMS data-services:
-\begin{itemize}
-\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
-\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
-CMS data location and file transfer system;
-\item SiteDB \cite{SiteDB}, the CMS site resources;
-\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
-%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
-%\item DQ \cite{DQ}, Data Quality data-service which collects information
-%about detector conditions during data taking;
-%\item Overview \cite{Overview}, collects information about CMS
-%transfer rate, etc.;
-\item Dashboard \cite{Dashboard}, the LHC virtual organization
-monitoring data-service portal;
-\end{itemize}
-Each data-service has its own scope, size and peculiar properties. 
-For instance, the data in PhEDEx are transient due to constant 
-migration of CMS data; the DBS system is divided into dozen 
-of individual services (instances) with own size and lifetime 
-used by different production teams for various bookkeeping tasks; 
-the RunSummary data-service is operated on private network accessible 
-at CMS online operation center.
-
-At the moment, the amount of meta-data stored in all services is around
-75 GB and 200 GB in tables and index sizes, respectively. We estimate to
-collect $\sim$500 GB of the meta-data each year during 
-normal CMS operation at LHC.
-%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
-%billion events, 330TB in 32 datasets:
-%AlCa/Calibration RAW: 926 million events, 29TB
-%Physics RAW: 1.24 billion events, 147TB
-%PromptReco: 1.24 billion events, 111TB
-%Bulk ALCARECO: 2.34 billion events, 11TB
-%Express FEVT: 105 million events, 24TB
-%Express ALCARECO: 153 million events, 0.99TB
-%HLTMON FEVTHLTALL: 31 million events, 7.4TB
-%Above event numbers have significant overlap by design, 
-%2.2 billion unique events, including 524 million cosmic muon
-%triggers
-% 0.3PB - 50GB meta-data in DBS
-% 3PB/y - x GB
-
-To test DAS performance we ere measured elapsed time and CPU/RAM utilization
-required to fetch, store and aggregate information from different
-data-services. Table \ref{DAS_benchmark} summarizes the biggest 
-contributors so far, the block meta-data information provided by DBS and 
-PhEDEx CMS data-services.
-
-\begin{table*}[hbt]
-\centering
-\begin{tabular}{lllll}\hline
-\hline
-System & Format & Records & Elapsed time & Elapsed time \\
-& & & no cached data & w/ cached data \\
-\hline
-DAS (DBS) & XML & 382396 & 63 sec & 0.98 sec \\
-DAS (PhEDEx) & XML & 186145/180439 & 164 sec & 0.98 sec \\
-DAS (aggregate) & JSON & 388108 & 234 sec & 2.05/50.7 sec \\
-\hline
-\hline
-\end{tabular}
-\caption{Time required to fetch, parse and aggregate block information
-from DBS and PhEDEx systems. The PhEDEx
-system shows fetched (186145) and merged (180439) records. 
-The total number of DAS records are calculated as number of 
-aggregated records plus left overs, the non-matched records from both systems. 
-The look-up time shown in last column represents query look-up 
-time (2.05 sec) and time required to get all records written 
-to disk (50.7 sec).}
-\label{DAS_benchmark}
-\end{table*}
-
-The initial DAS implementation is done in Python language \cite{Python}
-and all tests are performed on x64 Linux platform where
-1 CPU was used by DAS server and 1 CPU by MongoDB database server. 
-All tests are performed multiple times to avoid fluctuations. 
-The elapsed time is measured and consists of the following components:
-\begin{itemize}
-\item[]
-{\it
-Elapsed time = retrieval time + parsing time + re-mapping time 
-        + cache insertion/indexing time 
-        + (aggregation time) + (output creation time)
-}
-\end{itemize}
-Here {\it retrieval time} is a time required to access data from data-service,
-{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
-is a time required to convert data-service meta-data to DAS records for every object
-we parse, {\it cache insertion} and {\it indexing time} represents time spent on DAS 
-cache back-end, {\it aggregation time} is required to merge objects into DAS records based
-on their common key (\verb+block.name+ in this case) and {\it output creation time}
-is a time required to write DAS records to disk. Please note that last two
-components, {\it aggregation} and {\it output creation time}, are only applied to
-one of the sub-system, either DBS or PhEDEx, depending on the insertion order.
-
-%DAS execution time (phedex) 160.606897116 sec
-%DAS execution time (dbs) 52.5349619389 sec
-%DAS execution time 256.0670228 sec, Sat, 28 Nov 2009 15:46:39 GMT
-%
-%DAS execution time (phedex) 0.952513933182 sec
-%DAS execution time (dbs) 0.951184034348 sec
-%DAS execution time 44.7982931137 sec, Sat, 28 Nov 2009 15:52:11 GMT
-
-The individual tests of DAS components (DBS and PhEDEx) shown roughly 
-the same performance. Although the time of PhEDEx component is bigger 
-due to longer retrieval and parsing time as well as time spent
-during aggregation step. As shown in Table \ref{DAS_benchmark},
-the look-up time spent to access individual components (last column) is quite
-reasonable, roughly 1 second for both systems. While the final time to
-get DAS records on disk is about 50 seconds. This is mostly due to I/O and
-conversion operations between binary data format used by MongoDB, BSON, to DAS
-data format, JSON.
-
-Based on the results of stress tests we are measured that we achieve throughput of
-$\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
-$\sim$7600 docs/sec for 
-reading operations in DAS, which is suitable for our needs. Since all 
-operations are performed on a single CPU, a further improvements are expected. 
-Also, a necessary tuning can be done by using C/C++ driver instead of python one.
 
 \section{Summary}
-We are presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
-at LHC, CERN, Geneva, Switzerland. It is designed to provide caching and
-aggregation layer on top of the existing relational and non-relation data-services
-mostly in real time fashion. All data are retrieved on demand basis,
-while pre-fetching of most common queries is in our plans. We are developed a
-prototype in Python which is under commissioning phase right now.
-The performance studies of the DAS system are presented in this paper. 
-The CMS experiment is started to collect its data in December of 2009 and 
-we expect to accumulate a few PB of real data each year. 
-In addition, the Monte-Carlo samples will be produced at this scale.
-Based on test studies performed in CMS, we expect that total size of
-meta-data produced each year by experiment will be of the order of
-500GB. The DAS system should sustain such load and provide generic 
-``data-discovery'' service for CMS experiment in years to come.
 
-\section{Acknowledgments}
+\section{Acknowledgements}
 
-This work was supported by the National Science Foundation and 
-Department of Energy of the United States of America. 
-Fermilab is operated by Fermi Research Alliance, LLC under Contract
+This work was supported by the National Science Foundation and Department of Energy of the United States of America. Fermilab is operated by Fermi Research Alliance, LLC under Contract
 No. DE-AC02-07CH11359 with the United States Department of Energy.
 
-%% References
-%%
-%% Following citation commands can be used in the body text:
-%% Usage of \cite is as follows:
-%%   \cite{key}         ==>>  [#]
-%%   \cite[chap. 2]{key} ==>> [#, chap. 2]
-%%
+\section*{References}
+\begin{thebibliography}{9}
+\bibitem{DBS} A. Afaq, et. al. ``The CMS Dataset Bookkeeping Service'', CHEP 2007 
+\bibitem{DBS07} A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
+``A multi-dimensional view on information retrieval of CMS data'', CHEP 2007
+\bibitem{DBS-QL} V. Kuznetsov, D. Riley, A. Afaq, V. Sekhri, Y. Guo, L. Lueking,
+``The CMS DBS Query Language'', CHEP 2009
+\bibitem{DD} https://cmsweb.cern.ch/dbs\_discovery
 
-%% References with BibTeX database:
-
-%\bibliographystyle{elsarticle-num}
-%\bibliography{<your-bib-database>}
-
-%% Authors are advised to use a BibTeX database file for their reference list.
-%% The provided style file elsarticle-num.bst formats references in the required Procedia style
-
-%% For references without a BibTeX database:
-
-% \begin{thebibliography}{00}
-
-%% \bibitem must have the following form:
-%%   \bibitem{key}...
-%%
-
-% \bibitem{}
-
-% \end{thebibliography}
-
-\section*{References}
-\begin{thebibliography}{00}
+\bibitem{Arms}
+C. R. Arms, W. Y. Arms, ``Mixed Content and Mixed Metadata 
+Information Discovery in a Messy World'',
+chapter from ``Metadata in Practice'', ALA Editions, 2004
 \bibitem{DBXplorer}
-S. Agrawal, S. Chaudhuri, G. Das,
-``DBXplorer: A System for Keyword-Based Search over Relational Databases'',
-ICDE 2002, pp. 5-16.
+Sanjay Agrawal, Surajit Chaudhuri, Gautam Das: DBXplorer: A System for
+Keyword-Based Search over Relational Databases. ICDE 2002: 5-16
 
 \bibitem{QueryAnswer}
-G. Koutrika, A. Simitsis, Y. E. Ioannidis,
-``Pr\'{e}cis: The Essence of a Query Answer'',
-ICDE 2006, pp. 69-78.
-
-\bibitem{DBS-QL} 
-V. Kuznetsov, D. Riley, A. Afaq, V. Sekhri, Y. Guo, L. Lueking,
-``The CMS DBS Query Language'', CHEP 2009.
-
-%\bibitem{AMI}
-%Altas AMI web portal,
-%http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html
+Georgia Koutrika, Alkis Simitsis, Yannis E. Ioannidis: Pr\'{e}cis: The Essence of
+a Query Answer. ICDE 2006: 69-78
 
@@
 http://www.ibm.com/developerworks/data/library/techarticle/0203haas/0203haas.html
-
-\bibitem{CMS} 
-R. Adolphi et al., 
-``The CMS experiment at the CERN LHC'',
-JINST 0803, S08004 (2008).
-
-\bibitem{CMSDataModel} 
-C. Grandi, D.Stickland, L.Taylor et al.,
-``The CMS Computing Model'',
-CERN-LHCC-2004-035/G-083 (2004).
-%A. Fanfani et. al., 
-%``Distributed Analysis in CMS'', 
-%to be published in Journal of Grid Computing.
-
-\bibitem{iRODS}
-The Integrated Rule-Oriented Data System,
-http://www.irods.org/
-
-\bibitem{OpenArchive}
-Open Archives Initiative,
-http://www.openarchives.org/
-
-\bibitem{DBS} 
-A. Afaq, et. al.,
-``The CMS Dataset Bookkeeping Service'', 
-J. Phys. Conf. Ser, 119, 072001 (2008).
-
-\bibitem{DBS07} 
-A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
-``A multi-dimensional view on information retrieval of CMS data'',
-J. Phys. Conf. Ser, 119, 072013 (2008).
-
-\bibitem{DD} https://cmsweb.cern.ch/dbs\_discovery
-
-\bibitem{Arms}
-C. R. Arms, W. Y. Arms,
- ``Mixed Content and Mixed Metadata 
-Information Discovery in a Messy World'',
-chapter from ``Metadata in Practice'', ALA Editions, 2004.
 
@@
 
-\bibitem{Python} Python programming language, http://www.python.org
+\end{thebibliography}
 
-\bibitem{PhEDEx}
-%Rehn et. al.,
-%``PhEDEx high-throughput data transfer management system'', CHEP 2006;
-%R. Egeland et al., 
-%``Data transfer infrastructure for CMS data taking'', 
-%Proceedings of Science, PoS (ACAT08) 033 (2008);
-L. Tuura et al., 
-``Scaling CMS data transfer system for LHC start-up'', 
-J. Phys. Conf. Ser, 119, 072030 (2008).
-
-\bibitem{RunSummary}
-https://cmswbm.web.cern.ch/cmswbm/cmsdb/servlet/RunSummary
-
-\bibitem{SiteDB}
-https://cmsweb.cern.ch/sitedb/
-
-%\bibitem{LumiDB}
-%https://twiki.cern.ch/twiki/bin/view/CMS/CMS-DMWM-DBS-Luminosity
-
-%\bibitem{DQ}
-%Data-Quality DB reference
-
-%\bibitem{Overview}
-%https://cmsweb.cern.ch/overview/
-
-\bibitem{Dashboard}
-J. Andreeva, et. al,
-``Experiment Dashboard – The Monitoring System for the LHC Experiments'',
-GMW’07, June 25, 2007, Monterey, California, USA.
-
-%\bibitem{CRAFT09}
-%https://twiki.cern.ch/twiki/bin/view/CMS/CRAFT09AnalysisInfo
-
-%\bibitem{DBSearch}
-%D. Konopnicki, O. Shmueli,
-%``Database-Inspired Search'', 
-%Proc. of the 31st VLDB Conference, 2005.
-\end{thebibliography}
 \end{document}

### 2010-05-04 src/python/DAS/core/qlparser.py
Added input query into output QL dict
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1506 4525493e-7705-40b1-a816-d608a930855b

@@ -65,8 +65,8 @@ find intlumi,dataset where site=T2_UK or hlt=OK
  'unique_keys': ['dataset', 'intlumi', 'lumi', 'run']}
 """
 
-__revision__ = "$Id: qlparser.py,v 1.16 2009/09/02 19:52:55 valya Exp $"
-__version__ = "$Revision: 1.16 $"
+__revision__ = "$Id: qlparser.py,v 1.17 2009/09/09 18:31:38 valya Exp $"
+__version__ = "$Revision: 1.17 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -751,7 +751,7 @@ class QLParser(QLLexer):
         query = self.fix_reserved_keywords(query)
         query, namespace = self.find_namespaces(query)
         query, functions = self.das_functions(query)
-        rdict = {}
+        rdict = dict(query=query)
         rdict['functions']       = functions
         rdict['conditions']      = self.conditions(query)
         order_by_list, order_by  = self.order_by(query)

### 2010-05-04 src/python/DAS/core/das_core.py
Drop off raw-cache look-up from individual data-services. Instead let's then either return if objects exists in a cache or update a cache. Add to core as a final step a look-up of query from raw-cache once all data-services complete their work. It has dual benefit, obviously speed up, since query will goes to DB only once per request and output will not contain intermediate objects, e.g. DBS add block object, then phedex update it, so we will not see DBS one.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1508 4525493e-7705-40b1-a816-d608a930855b

@@ -12,8 +12,8 @@ combine them together for presentation layer (CLI or WEB).
 
 from __future__ import with_statement
 
-__revision__ = "$Id: das_core.py,v 1.28 2009/09/01 01:42:44 valya Exp $"
-__version__ = "$Revision: 1.28 $"
+__revision__ = "$Id: das_core.py,v 1.29 2009/09/09 18:40:36 valya Exp $"
+__version__ = "$Revision: 1.29 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -75,6 +75,7 @@ class DASCore(object):
         self.viewmgr = DASViewManager(dasconfig)
 
         dasroot = os.environ['DAS_ROOT']
+        self.rawcache = None
         # load from configuration what will be used as a raw/cold cache
         if  dasconfig.has_key('rawcache') and dasconfig['rawcache']:
             klass   = dasconfig['rawcache']
@@ -86,6 +87,9 @@ class DASCore(object):
             setattr(self, 'rawcache', eval(klassobj))
             dasconfig['rawcache'] = self.rawcache
             self.logger.info('DASCore::__init__, rawcache=%s' % klass)
+        else:
+            msg = 'DAS configuration file does not provide rawcache'
+            raise Exception(msg)
 
         # load from configuration what will be used as a hot cache
         if  dasconfig.has_key('hotcache') and dasconfig['hotcache']:
@@ -301,25 +305,6 @@ class DASCore(object):
             if  self.cache.incache(query):
                 self.cache.remove_from_cache(query)
 
-#    def in_raw_cache(self, uinput):
-#        """
-#        Check if input query is presented in raw cache
-#        """
-#        query    = self.viewanalyzer(uinput)
-#        params   = self.qlparser.params(query)
-#        services = params['unique_services']
-#        daslist  = params['daslist']
-#        for qdict in daslist:
-#            for service in services:
-#                if  qdict.has_key(service):
-#                    squery = qdict[service]
-                    # raw cache uses "query @ service" syntax to store queries
-                    # see services/abstrace_service.py call method
-#                    qqq = '%s @ %s' % (squery, service)
-#                    if  self.rawcache.incache(qqq):
-#                        return True
-#        return False
-
     def get_params(self, uinput):
         """
         The purpose of this method is to parse input query and return
@@ -352,6 +337,7 @@ class DASCore(object):
         ulist    = params['unique_keys']
         services = params['unique_services']
[…]

### 2010-05-04 test/das_analytics_t.py
New Class to deal with DAS Analytics DB, simple schema: System, Api, ApiParam, ApiCounter. DB is generated/loaded via SQLAlchemy. Provide associative unit test
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1512 4525493e-7705-40b1-a816-d608a930855b

@@ -1,0 +1,84 @@
+#!/usr/bin/env python
+#pylint: disable-msg=C0301,C0103
+
+"""
+Unit test for DAS analyticsdb class
+"""
+
+import os
+import time
+import unittest
+from DAS.utils.das_config import das_readconfig
+from DAS.utils.logger import DASLogger
+from DAS.core.das_analytics_db import DASAnalytics
+
+class testDASAnalyticsMgr(unittest.TestCase):
+    """
+    A test class for the DAS analyticsdb class
+    """
+    def setUp(self):
+        """
+        set up DAS core module
+        """
+        debug    = 0
+        self.db  = 'test_analytics.db'
+        config   = das_readconfig()
+        logger   = DASLogger(verbose=debug, stdout=debug)
+        config['logger']  = logger
+        config['verbose'] = debug
+        config['analytics_db_engine'] = 'sqlite:///%s' % self.db
+        self.mgr = DASAnalytics(config)
+
+    def tearDown(self):
+        """Invoke after each test"""
+        try:
+            os.remove(self.db)
+        except:
+            pass
+
+    def test_system(self):                          
+        """test methods for system table"""
+        self.mgr.create_db()
+        self.mgr.add_system('dbs')
+        res = self.mgr.list_systems()
+        self.assertEqual(['dbs'], res)
+        res = self.mgr.delete_system('dbs')
+        self.assertEqual(True, res)
+
+    def test_api(self):                          
+        """test methods for api table"""
+        self.mgr.create_db()
+        system = 'dbs'
+        self.mgr.add_system(system)
+
+        api = 'listBlocks'
+        params = {'apiversion':'DBS_2_0_8',
+                  'block_name':'*', 'storage_element_name':'*',
+                  'user_type':'NORMAL'}
+        self.mgr.add_api(system, api, params)
+
+        res = self.mgr.list_systems()
+        self.assertEqual([system], res)
+
+        res = self.mgr.list_apis(system)
+        self.assertEqual([api], res)
+
+        self.mgr.add_api(system, api, params)
+        res = self.mgr.api_counter(api)
+        self.assertEqual(2, res) # we invoke API twice, so should get 2
+
+        res = self.mgr.api_params(api)
+        self.assertEqual(params, res)
+
+        res = self.mgr.delete_api(api)
+        self.assertEqual(True, res)
+
+        res = self.mgr.delete_system(system)
+        self.assertEqual(True, res)
+
+#
+# main
+#
+if __name__ == '__main__':
+    unittest.main()
+

### 2010-05-04 doc/paper.tex
Adding Simon's cache and analytics picture, since it's better describe cache/analytics flow. Change text wrt that.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1514 4525493e-7705-40b1-a816-d608a930855b

@@
 \centering
-\includegraphics[width=150mm]{DAS_cache.pdf}
+\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
 \caption{
@@
 \end{figure}
-As it shown in figure \ref{DAS_cache} we setup
-dedicated DAS cache request server to manage user requests. Each request placed into
-queue who was used by workers to spawn a data-service request. 
-This allows us to utilze a multi-core technology and place a workers at each 
-core.\footnote{A number of workers per core were adjustable to satisfy requests load.}
-Each worker talk to DAS raw-cache and either stored mapped results over there or merge
-them with existing data object in DAS cache. We evaluated several technologies
+As it shown in figure \ref{DAS_cache} the DAS cache server has dual purpose.
+On one side it talks to DAS Analytics DB to trigger jobs for worker robot
+who update cache with most popular queries, on another it talks to DAS
+raw-cache back-end to either stored mapped results or merge them with 
+existing DAS objects in a cache. We should note that we use on-demand
+approach to answer user queries, but at the same time with a help of
+DAS Analytics DB we force worker robots (populators) to pre-fetch
+the most popular queries up-front.
+
+We evaluated several technologies
 as a choice for DAS cache back-end, including relational MySQL DB \cite{MySQL} 
 and object oriented CouchDB \cite{CouchDB} and MongoDB \cite{MongoDB}.
-Each of this technologies had its own pros and cons. For instance usage
+Each of them had its own pros and cons. For instance usage
 of MySQL provided a SQL queries (as in federated DBs, see \cite{FedDB}), but
@@
 querying.} While, using CouchDB and MongoDB, we were able to store
-data-service output naturally as data objects into DB. Finally we decided
+data-service output naturally as documents objects into DB. Finally we decided
 to use MongoDB due to its flexible and

### 2010-05-04 doc/paper.tex
More refinements on content flow
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1515 4525493e-7705-40b1-a816-d608a930855b

@@
 \section{Data Aggregation System\label{DAS}}
-Data Aggregation System shown in Figure \ref{DAS_arch} and
-consists out of several components:
+Our goal was to create scalable and pluggable system for existing data-services
+within CMS computing infrastructure. We knew that all our data are delivered 
+via data-services who take care of their availability. Therefore, we treat 
+DAS as aggregation and cache service for existing data-services who should
+take care of the following:
+\begin{itemize}
+\item an individual data-service can have an own set of security policies;
+\item a dynamic nature of data-service APIs, due to continiously changing
+detector and software environments;
+\item a broad variety of data-formats;
+\item distributed nature of data-services;
+\item heterogeneous software environment.
+\item support flexible and intuitive Query Language (QL) to easy look-up our
+data;
+\item pluggable interface to existing and up-coming data-service APIs;
+\end{itemize}
+
 \begin{figure}[htb]
@@
 
+Based on this set of constrains and requirements we come up with the following
+DAS components, see Figure \ref{DAS_arch}:
 \begin{itemize}
@@
 hot-cache;
-\item DAS request service who collect and place into queue user requests
+\item DAS request service to handle user requests
 \item DAS mapping DB to store mapping between data-services and DAS notations and
@@
 \end{itemize}
-We will discuss each of those components in more details in up-coming sub-sections.
+We will discuss individual components in more details in up-coming sub-sections.
+
+\subsection{Dealing with data-services}
 The DAS supports a pluggable approach to add a new CMS sub-system into a game by
@@
 
-\subsection{DAS Query Language}
+We designed DAS interface to be fully REST complaint \cite{REST}. It is not
+only simplify development of DAS clients (CLI, web interface, populator, etc.)
+but also consider DAS itself as own sub-component, i.e. place queries against
+DAS content, rather then data-service one.
+
+\subsection{Query Language\label{DAS-QL}}
 Upon success of DBS-QL \ref{DBS-QL} we adopted its syntax to DAS with a few minor
@@
 
-\subsection{DAS Mapping DB}
+\subsection{Mapping}
 As been discussed above we used a DAS mapping DB to map data-service API into
@@
 
-\subsection{DAS Analytics DB}
-
-\subsection{Caching system}
+\subsection{Caching system and Analytics}
 The DAS caching system was divided into two different caches: raw-cache and hot-cache.
@@
 powerful querying syntax, which were easy to map into DAS QL expressions.
-
-\subsection{REST interface}
-We designed DAS interface to be fully REST complaint \cite{REST}.
-All requests placed into DAS cache server were done via GET/PUT/POST/DELETE methods.
-DAS clients initiate their request via GET method. An appropriate message
-with either resulting data or their absense was send back to client.
-Clients were able to decided upon received message how to proceed, either
-request data via POST/PUT methods, DELETE data in cache or initiate another
-GET request. The REST complaince help us to treat our clients, user quyering
-DAS system via web browser, other CMS applications and command line tools
-in identical manner. Also, the implementation of pre-fetch populator
-tools become as easy and identical as any other clients.
 

### 2010-05-04 src/python/DAS/core/das_analytics_db.py
Added Query into Analytics DB. So now, I can invoke add_api(query, system, api, params) and perform look-up of most querying APIs, their parameters and associative queries. Added new method to look-up all queries, can be extended to look-up queries for given system and/or api. Adjust unit test
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1516 4525493e-7705-40b1-a816-d608a930855b

@@ -7,8 +7,8 @@ DAS analytics DB
 
 from __future__ import with_statement
 
-__revision__ = "$Id: das_analytics_db.py,v 1.1 2009/09/09 20:59:38 valya Exp $"
-__version__ = "$Revision: 1.1 $"
+__revision__ = "$Id: das_analytics_db.py,v 1.2 2009/09/10 19:21:20 valya Exp $"
+__version__ = "$Revision: 1.2 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -36,6 +36,21 @@ class System(Base):
     def __repr__(self):
         """String representation of System ORM object"""
         return "<System('%s')>" % self.name
+
+class Query(Base):
+    """Query ORM"""
+    __tablename__ = 'queries'
+    __table_args__ = {'mysql_engine':'InnoDB'}
+
+    id = Column(Integer, primary_key=True)
+    query = Column(String(10), nullable=False, unique=True)
+
+    def __init__(self, query):
+        self.query = query
+
+    def __repr__(self):
+        """String representation of Query ORM object"""
+        return "<Query('%s')>" % self.query
 
 class Api(Base):
     """Api ORM"""
@@ -89,11 +104,13 @@ class ApiCounter(Base):
     system_id   = Column(Integer, ForeignKey('systems.id'))
     api_id      = Column(Integer, ForeignKey('apis.id'))
     param_id    = Column(Integer, ForeignKey('apiparams.id'))
+    query_id    = Column(Integer, ForeignKey('queries.id'))
     counter     = Column(Integer, nullable=False)
 
     system      = relation(System, order_by=System.id)
     api         = relation(Api, order_by=Api.id)
     param       = relation(ApiParam, order_by=ApiParam.id)
+    query       = relation(Query, order_by=Query.id)
 
     def __init__(self):
         self.counter = 1
@@ -132,9 +149,10 @@ class DASAnalytics(object):
         metadata = MetaData()
         Base.metadata.create_all(self.engine)
 
-    def add_api(self, system, api, args):
+    def add_api(self, query, system, api, args):
         """
-        Add API into analytics DB. Here args is a dict of API parameters.
+        Add API info to analytics DB. 
+        Here args is a dict of API parameters.
         """
         session = self.session()
         try:
@@ -145,6 +163,12 @@ class DASAnalytics(object):
                 aobj = Api(api)
                 aobj.system = sobj
                 session.add(aobj)
+                session.commit()
+            try:
[…]

### 2010-05-04 src/python/DAS/core/das_core.py
One minor change to analytics db relying on schema-based DB to keep it for the record, before converting to Mongo back-end. Put initialization of Analytics class into core and pass it to services. Plug add_api call into api call of services.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1517 4525493e-7705-40b1-a816-d608a930855b

@@ -12,8 +12,8 @@ combine them together for presentation layer (CLI or WEB).
 
 from __future__ import with_statement
 
-__revision__ = "$Id: das_core.py,v 1.29 2009/09/09 18:40:36 valya Exp $"
-__version__ = "$Revision: 1.29 $"
+__revision__ = "$Id: das_core.py,v 1.30 2009/09/11 13:26:56 valya Exp $"
+__version__ = "$Revision: 1.30 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -25,6 +25,7 @@ import traceback
 from DAS.core.qlparser import QLParser
 from DAS.core.das_viewmanager import DASViewManager
 from DAS.core.das_mapping import DASMapping
+from DAS.core.das_analytics_db import DASAnalytics
 from DAS.utils.das_config import das_readconfig
 from DAS.utils.logger import DASLogger
 
@@ -71,6 +72,9 @@ class DASCore(object):
 
         dasmapping = DASMapping(dasconfig)
         dasconfig['dasmapping'] = dasmapping
+
+        dasanalytics = DASAnalytics(dasconfig)
+        dasconfig['dasanalytics'] = dasanalytics
 
         self.viewmgr = DASViewManager(dasconfig)
 

### 2010-05-04 src/python/DAS/services/abstract_service.py
Change regexp pattern for querying values.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1553 4525493e-7705-40b1-a816-d608a930855b

@@ -4,8 +4,8 @@
 """
 Abstract interface for DAS service
 """
-__revision__ = "$Id: abstract_service.py,v 1.35 2009/09/14 20:37:56 valya Exp $"
-__version__ = "$Revision: 1.35 $"
+__revision__ = "$Id: abstract_service.py,v 1.36 2009/09/21 15:28:12 valya Exp $"
+__version__ = "$Revision: 1.36 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -274,7 +274,7 @@ class DASAbstractService(object):
                 if  not nkey:
                     nkey = key
                 if  oper == '=' and type(value) is types.StringType:
-                    value = re.compile('%s.*' % value)
+                    value = re.compile('.*%s.*' % value)
                 if  key != 'date': # we skip date for Mongo to look-up,
                     # due to variety of return formats
                     cdict = dict(key=nkey, op=oper, value=value)

### 2010-05-04 doc/paper.tex
Bug fix, footebnote
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1556 4525493e-7705-40b1-a816-d608a930855b

@@
 \end{verbatim}
-We wanted to expand this approach and apply such queries across multiple
-data services and different DB back-ends.
+We wanted to expand this approach to handle broad variety of relational and
+non-relational data-source in CMS within our distributed non-homogeneous environment.
 
-Here we discuss a new system, the Data Aggregation System (DAS),
-developed in CMS collaboration to address this issue. We start with
-discussion of CMS data model and existing data-services, see \ref{DataModel}. 
-In section \ref{DAS}
-we provides a detailed description of underlying design and components
-used in DAS system. Our results are summarized in section \ref{Results}.
+%Here we discuss a new system, the Data Aggregation System (DAS),
+%developed in CMS collaboration to address this issue. We start with
+%discussion of CMS data model and existing data-services, see \ref{DataModel}. 
+%In section \ref{DAS}
+%we provides a detailed description of underlying design and components
+%used in DAS system. Our results are summarized in section \ref{Results}.
 
 \section{CMS data model\label{DataModel}}
+The CMS distributed computing and data model \cite{CMSDataModel} 
+is designed to process
+and efficiently manage a PBs of data expected to be produced each year
+at LHC. The computing resources provided by members of CMS
+collaboration are geographically distributed, 
+interconnected via high throughput networks and operated by means 
+of Grid software. The model allows to cover broad variety of
+hardware, mass and storage elements and configuration of
+clusters. To accomodate the efficient data processing we use 
+a multi-Tier distributed model, where specific tasks of data taking,
+processing, archival and distribution assigned to each tier based
+on CMS data model. For example, the Tier-0 center at CERN responsible
+for data archives coming out of detector, prompt first pass reconstruction
+and data distribution to Tier-1 centers. The 7 Tier-1 centers
+located in France, Germany, Ialy, Spain, Taiwan, United Kingdom and United States
+keep a portion (copy) of data delivered by Tier-0 for further processing.
+They provide storage and high priority analysis jobs at their facilities.
+The Tier-2 centers, located at more then 50 sites around the world,
+are dedicated for user analysis tasks and production of simulated data.
+
+A broad variety of data-services being designed and developed to
+maintain detector and production operations, including detector
+conditions databases, data location and bookkeeping services,
+data transfer and job monitoring tasks. Even though majority of them
+are located at CERN, it was never been a requirement in CMS computing
+and data model. For instance, the production teams operated at Tier-1,2
+centers set up local services for data bookkeeping and operational
+tasks. Based on their budget resources the choice of back-end DB were
+up to site managers. Therefore we designed our software to support different
+DB back-ends and provide tools for migrating data across them.
+
+Once such conglomerate of data-services starts operating an obvious
+question arise: how to find out desired information across multiple data-services
+in our distributed environment? Even though individual data-services were designed
+to answer specific question about data they serve the ability to relate and search
+information among them was human task. The growing amount of information
+and desire to make cross-service queries force us to design and develop a new
+type of system, the Data Aggregation System.
+
 \section{Data Aggregation System\label{DAS}}
-Our goal was to create scalable and pluggable system for existing data-services
-within CMS computing infrastructure. We knew that all our data are delivered 
-via data-services who take care of their availability. Therefore, we treat 
-DAS as aggregation and cache service for existing data-services who should
-take care of the following:
+The design of DAS system was based on previous studies of CMS 
+data bookkeeping system (DBS) usage patterns \cite{DBS07}. We carefully analyzed user
+queries placed to DBS, their patterns, frequencies and latency. As a result of this
+we created a data discovery service \cite{DD} which came through several iteration
+of UI. At the end the end-user interface was based on DBS Query Language \cite{DBS-QL}
+with broad variety of template for presentation layer. A quick adoption
+and wide spread of DBS QL give us confidence of choosen approach and QL syntax.
+
+Starting from those ground accomplishments we created DAS system as
+scalable and pluggable layer sitting on top of existing data-services
+within CMS computing infrastructure. Data availability were deligated to individual
+data-services. Therefore, we treat DAS as aggregation and cache service 
+with the following characteristics:
 \begin{itemize}
-\item an individual data-service can have an own set of security policies;
-\item a dynamic nature of data-service APIs, due to continiously changing
-detector and software environments;
-\item a broad variety of data-formats;
-\item distributed nature of data-services;
-\item heterogeneous software environment.
 \item support flexible and intuitive Query Language (QL) to easy look-up our
@@
 \item pluggable interface to existing and up-coming data-service APIs;
+\item support heterogeneous software environment and distributed nature of data-services;
+\item preserve security policies of individual data-services while provide unique
+authentication schema for end-users;
+\item be flexible to adjust to dynamic nature of data-service APIs, 
+due to continuously changing detector and software environments;
+\item be transparent to data-service data formats.
 \end{itemize}
 
+%\begin{figure}[htb]
+%\centering
+%\includegraphics[width=150mm]{DAS_architecture.pdf}
+%\caption{
+%DAS architecture.
+%}
+%\label{DAS_arch}
+%\end{figure}
 \begin{figure}[htb]
 \centering
-\includegraphics[width=150mm]{DAS_architecture.pdf}
+\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
 \caption{
-DAS architecture.
+DAS architecture diagram. It consists of DAS cache server, mapping and analytics DBs.
+Once query was placed to DAS an associated
+data-service API was invoked, triggering in DAS cache population. Once data become
+available in DAS cache user were able to see the results. This so-called on demand
+query retrieval system, where data were retrieved and store into cache from
+back-end data service. We use DAS analytics DB to establish cache population
+for most common queries by using DAS robots.
 }
-\label{DAS_arch}
+\label{DAS_cache}
 \end{figure}
 
-Based on this set of constrains and requirements we come up with the following
-DAS components, see Figure \ref{DAS_arch}:
+\noindent
+Figure \ref{DAS_cache} summarize a overview of DAS architecture who consist of the
+following components:
 \begin{itemize}
-\item DAS core system which is support pluggable modules to retrieve data from
-CMS data-services, e.g. Data Bookkeeping System (DBS), data location service (Phedex),
-etc.;
-\item DAS caching system which has dual purpose: to store ``raw'' results coming
-out from data-services, a raw-cache, and aggregated results produced by DAS core into
-hot-cache;
-\item DAS request service to handle user requests
-\item DAS mapping DB to store mapping between data-services and DAS notations and
-data-representation;
-\item DAS analytics DB to collect query statistics and establish pre-fetching 
+\item DAS core library with support of pluggable modules for data retrieval;
+\item DAS caching layer with long-term cache for aggregated results and short-term
+cache for presentation layer;
+\item DAS request service to handle user requests;
+\item DAS mapping DB with DAS/data-services mapping, notations mapping and
+mapping of DAS objects into data-presentation layer;
+\item DAS analytics DB with query statistics to enable pre-fetching 
 strategies.
 \end{itemize}
-We will discuss individual components in more details in up-coming sub-sections.
+Further we proceed with a discussion of individual DAS components.
 
 \subsection{Dealing with data-services}
-The DAS supports a pluggable approach to add a new CMS sub-system into a game by
-identifying appropriate APIs, input and output parameters and keys, data-service
-format and mapping between data-service, DAS and other sub-system. For instance,
-the Data Bookkeeping System (DBS) and data location system (Phedex) store information about
-a file. In former case, a file name was stored as well as its relations to other physics 
-objects, such as run and luminosity, while in later case the Phedex store information 
-about file location. When user specify a file as their selection key, based on
-information store in DAS mapping DB we were able to identify appropriate set of
+DAS provided pluggable system to add a new CMS data-service. To do that we
+identify appropriate APIs, input and output parameters and keys, data-service
+format and mapping between data-service, DAS and other sub-systems. 
+This information were stored into DAS mapping DB. For instance,
+the Data Bookkeeping System (DBS) and data location system (Phedex) both
+provide information about CMS data files. In former case, DBS keeps information between
+files and other physics  objects, such as run and luminosity, while Phedex provides 
+information about file location, such as site and details about storage elements.
+Upon user request to find information about a file, based on forementioned
+information we were able to identify appropriate set of
 data-service APIs, look-up data, re-format their output to DAS notations and
-store it into DAS raw-cache. Even though we support a data retrieval on demand
+store results into DAS cache.\footnote{Even though we supported a data retrieval on demand
 basis we also store user queries into DAS analytics DB to identify
 a common patterns. Based on this information we organized a pre-fetching
-populators. Once data were coming to raw-cache an additional step of
-aggregation was applied. If other data with identical keys were presented in
-a raw-cache we merge it them together. For instance upon the following query
-$find file where ... $ a file information from DBS and Phedex systems were
-retrieved. Since both data-services output contained file name as a common
-key, their output where merged and final result was stored into raw-cache.
-This merged (aggregated) results were shipped back to client and stored into
-hot-cache (memcached) to speed up a DAS web client application. Each DAS data
+strategies.}
+Once data were retreived from data-services an additional step to aggregate
+matched objects in DAS cache were applied. For instance in above examples
+file information came from DBS and Phedex systems. 
+Since both data-services output contained file name as a common
+key, their output where merged and final result was stored into DAS cache.
+This merged (aggregated) results represented a DAS data object. Each DAS data
 object contains a standard DAS header, which identify look-up time,
@@
 \subsection{Query Language\label{DAS-QL}}
-Upon success of DBS-QL \ref{DBS-QL} we adopted its syntax to DAS with a few minor
+Upon success of DBS-QL \cite{DBS-QL} we adopted its syntax to DAS with a few minor
 changes. Here we briefly recall its syntax:
@@
 \section{Results\label{Results}}
+DAS was deployed with the following CMS data-services:
+\begin{itemize}
+\item DBS \cite{DBS}, Data Bookkeeping System which collect information
+about CMS meta-data;
+\item PhEDEx \cite{Phedex}, Physics Experiment Data Export project which
+provides the data placement and the file transfer system for the CMS experiment;
+\item RunSummary DB \cite{RunSummary} collects information about run and triggers
+conditions during data-taking;
+\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
+\item DQ \cite{DQ}, Data Quality data-service which collects information
+about detector conditions during data taking;
+\item Overview \cite{Overview}, collects information about CMS
+transfer rate, etc.;
+\item Dashboard \cite{Dashboard} project for LHC experiments aims to 
+provide a single entry point to the monitoring data collected from the 
+distributed computing systems of the LHC virtual organization.
+\end{itemize}
+Each data-service has its own scope, size and lifetime. For instance the data
+in Phedex are transient due to constant migration of CMS data. The DBS system
+were divided into dozen of individual DBs, whose size varied at the level of
+few tens of GBs.
 
@@
 
+\bibitem{Phedex}
+Rehn et. al.,
+``PhEDEx high-throughput data transfer management system'', CHEP06, Mumbai, India.
+
+\bibitem{RunSummary}
+
+\bibitem{SiteDB}
+
+\bibitem{LumiDB}
+https://twiki.cern.ch/twiki/bin/view/CMS/CMS-DMWM-DBS-Luminosity
+
+\bibitem{DQ}
+
+\bibitem{Overview}
+https://cmsweb.cern.ch/overview/
+
+\bibitem{Dashboard}
+J. Andreeva, et. al,
+``Experiment Dashboard – The Monitoring System for the LHC Experiments'',
+GMW’07, June 25, 2007, Monterey, California, USA.
+
+\bibitem{DBSearch}
+D. Konopnicki, O. Shmueli,
+``Database-Inspired Search'', 
+Proc. of the 31st VLDB Conference, 2005.
 \end{thebibliography}

### 2010-05-04 doc/db_size.py
Plot and script how to get db size of meta-data services
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1657 4525493e-7705-40b1-a816-d608a930855b

@@ -1,0 +1,63 @@
+#!/usr/bin/env python
+import sys
+import matplotlib.pyplot as plt
+
+# DBS instance
+dbs_info = [
+('CMS_DBS_PROD_GLOBAL', 12.9034424, 22.1050415),
+('CMS_DBS_CAF_ANALYSIS_01', 0.894, 1.961),
+('CMS_DBS_PH_ANALYSIS_01', 0.601135254, 1.27581787),
+('CMS_DBS_PH_ANALYSIS_02', 0.969543457, 2.47174072),
+('CMS_DBS_PROD_TIER0', 2.01916504, 4.98065186),
+('CMS_DBS_PROD_LOCAL_01', 2.72540283, 4.42938232),
+('CMS_DBS_PROD_LOCAL_02', 1.2432251, 3.29675293),
+('CMS_DBS_PROD_LOCAL_03', 1.20812988, 2.66101074 ), 
+('CMS_DBS_PROD_LOCAL_04', 1.9777832, 4.06530762), 
+('CMS_DBS_PROD_LOCAL_05', 0.003723145, 0.010925293 ), 
+('CMS_DBS_PROD_LOCAL_06', 0.681152344, 1.80340576), 
+('CMS_DBS_PROD_LOCAL_07', 2.19720459, 7.07037354), 
+('CMS_DBS_PROD_LOCAL_08', 0.201416016, 0.397583008), 
+('CMS_DBS_PROD_LOCAL_09', 0.399047852, .720947266),
+]
+dbs_tables = sum(i[1] for i in dbs_info)
+dbs_index = sum(i[2] for i in dbs_info)
+db_info = [
+('dashboard', 40.82977294921875, 135.9593505859375),
+('phedex', 5.51031494, 12.5614014),
+('dbs', dbs_tables, dbs_index),
+('sitedb', 0.001, 0.001),
+('runsum', 5.0, 5.0), # need real numbers
+]
+
+db_names = [i[0].replace('CMS_DBS_','') for i in db_info]
+x = [i for i in range(0, len(db_info))]
+
+y_table=[i[1] for i in db_info]
+y_index=[i[2] for i in db_info]
+
+#dbs_tables=[i[1] for i in db_info if i[0].find('DBS') != -1]
+#dbs_indexies=[i[2] for i in db_info if i[0].find('DBS') != -1]
+#print "DBS tables size", sum(dbs_tables)
+#print "DBS index size", sum(dbs_indexies)
+
[…]

### 2010-05-04 doc/paper.tex
More changes from the flight
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1658 4525493e-7705-40b1-a816-d608a930855b

@@
 \begin{abstract}
-A meta-data plays significant role in a large modern enterprises, research experiments,
-digital libraries where it comes from different sources and distributed in a 
-variety of forms and digital formats. It is organized and managed by constantly
-evolving software using both relational and non-relational data sources. There is
-a big demand to enable information discovery from multiple sources.
-Here we discuss a new data aggregation system which consume and deliver information 
-from different relational and non-relational data sources on a concrete example 
-of large scale, distributed system of CMS physics experiment at LHC.
+%A meta-data plays significant role in a large modern enterprises, research experiments,
+%digital libraries where it comes from different sources and distributed in a 
+%variety of forms and digital formats. It is organized and managed by constantly
+%evolving software using both relational and non-relational data sources. There is
+%a big demand to enable information discovery from multiple sources.
+%Here we discuss a new data aggregation system which consume and deliver information 
+%from different relational and non-relational data sources on a concrete example 
+%of large scale, distributed system of CMS physics experiment at LHC.
+
+Meta-data plays signiﬁcant role in a large modern enterprises, 
+research experiments and digital libraries where it comes from different 
+sources and distributed in a variety of forms and digital formats. 
+It is organized and managed by constantly evolving software using 
+both relational and non-relational data sources. 
+Even though we can access it via set of data-services, twikies, blogs, etc.,
+there is no coherent way for information discovery from different sources
+in heterogeneous, distributed environment with different security policies enforced.
+
+Here we discuss a new data aggregation system which consumes, 
+indexes and delivers information from different relational and 
+non-relational data sources to answer cross data-service queries 
+to explore metadata associated with petabytes of experimental data. 
+We combine simplpicity of keyword-based search with precision of RDMS
+under a new system where aggregated information being collected from various sources 
+allowing end-users to place dynamic queries, get precise answers and 
+trigger information retrieval on demand. Such close to real-time system 
+based on studies  of use cases of the CMS particle physics experiment at 
+the LHC which uses a large scale, distributed computing system to motivate the work.
+
 \end{abstract}
@@
 The CERN, the European Organization for Nuclear Research, plays a leading
-role in fundamental studies of physics. Apart from scientific community 
-it is known as a place where World Wide Web was born. At that time the 
-information look-up among hyperlinks represents a certain challenge for scientists.
+role in fundamental studies of physics. It is also known as a place where
+World Wide Web was born to help researches share information among each other
+via concept of hypertext.
 Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
-Physics (HEP), promising to deliver a few PB of data each year. Scientists are
-facing another set of problems from distributed computing to multi-dimensional
-view on information retrieval. One of the aspect of such research is an efficient
-and at the same time precise look-up of produced meta-data, which comes in variety 
-of forms and digital formats. As was pointed out in \cite{Arms} a mixed content and 
-mixed meta-data and meta-data consistency should be considered as a whole in design 
-of the system to successful information discovery.
+Physics (HEP), promising to deliver a few PB of data each year. 
+At this scale scientists are facing a new set of problems in an area of
+``information discovery'' within heterogeneous, distributed environment.
+The data and associated meta-data are produced in variety of forms and digital formats.
+They are stored into relational and non-releational data-sources, such as 
+RDMS systems, document oriented databases, blogs, twikies, customized applications, etc. 
+and its access is restricted by broad variety of security policies. 
+Working in such environment is not easy and require a lot of knowledge, but users
+(in this case physicists) are always looking for simple, intuitive and flexible
+tool to look-up their desired data. A well-known solutions, such as web-services,
+blogs, twikies and search engines targeted and tighted to specific data
+sources and end-users are left with a way to bookkeep and relatate information
+from them.
 
-The CMS, the Compact Muon Solenoid experiment, operated at LHC,
-represents an heterogeneous environment of distributing computing, relational and
-non-relational data sources where we are facing this problem. The computing resources
-are distributed among almost 40 countries, 183 institutions and more than 3000 physicists.
-Broad variety of RDMS systems at different data centers collect a various
-meta-data information, including data location and transfer, detector conditions,
-calibrations constants, simulation information, etc. Moreover the development
-of those components were done in parallel by different group and technology
-tools. Therefore a wide spectrum of data services and
-data formats represents a challenge of efficient information discovery.
+Here we present a work on Data Aggregation System (DAS) designed for
+CMS High-Energy Experiment operated at LHC which provide
+ability to query, search and aggregate information from different 
+data-services as well as serves as a caching layer for them, 
+preserving their security policies.
 
-Here we present a work on Data Aggregation System (DAS), the system designed
-to provide ability to query, search and aggregate information from different 
-data-services as well as serves as a caching layer for them, 
-preserving their security policies. We organize
+We organize
 our discussion as following. In section \ref{RelatedWork} we discuss
-related work in domain of keyword search over relation data sources.
-The section \ref{DataModel} provides review of CMS data model. In section
+related work in a domain of keyword search over relational data sources.
+The section \ref{DataModel} describes CMS experiment and its data model. In section
 \ref{DAS} we outline architecture of the DAS system, including discussion of its
 various components. Finally our results are summarized in section \ref{Results}.
+
+%As was pointed out in \cite{Arms} a mixed content and 
+%mixed meta-data and meta-data consistency should be considered as a whole in design 
+%of the system to successful information discovery.
 
@@
 in last several years. The federated DB \cite{FedDB} by IBM unifies data coming 
-from different RDMS into another DB where SQL queries can be placed to search desired
+from different RDMS into federated DB where SQL queries can be placed to search desired
 data. The other approaches \cite{DBXplorer, QueryAnswer}, influenced by great success
 of search engines, explored ability to use keyword search algorithms over relataional RDMS
-by indexing DB content. In former case, you still face a problem with understanding 
-of the underlying schema and imposing relational conditions in your query, while 
+by indexing DB content. In former case, a problem with understanding 
+of underlying schema and imposing relational conditions in a query still exists, while 
 in later queries can be expressed via simple keyword based search, even though exact match
@@
 intuitive and easy to use, with respect to HEP and other domains it is not sufficient. 
-For example, it does provide provide ability to place conditions expressions between
-keys (DB entities/attributes) and their values. The proximity of results becomes very poor
-or impossible in case of numbers, aggregate functions, etc. To address this issues
+For example, it does not provide ability to place condition expressions,
+%keys (DB entities/attributes) and their numerical values, 
+e.g. $a > 1$, since
+it only index a text based values. Therefore proximity of results becomes very poor
+or impossible in case of numbers, aggregate functions, etc. To address those issues
 an attempt to build a simple, intuitive and flexible query language was introduced
-in \cite{DBS-QL}. It represented a power of SQL while
-hiding underlying relational schema from the end users. As a results
+in \cite{DBS-QL, AMI}.
+%\footnote{The ATLAS experiment at CERN, has been developed
+%AMI web portal \cite{AMI} with similar functionality.}
+It represented a power of SQL while
+hiding underlying relational schema from the end-users. As a results
 a human questions were intuitively mapped into simple queries. For example,
@@
 \end{verbatim}
-We wanted to expand this approach to handle broad variety of relational and
-non-relational data-source in CMS within our distributed non-homogeneous environment.
+Such approach not only address common use cases within HEP community, but also
+represented an intuitive mapping between mental model of our users and 
+underlying database back-end.
 
-%Here we discuss a new system, the Data Aggregation System (DAS),
-%developed in CMS collaboration to address this issue. We start with
-%discussion of CMS data model and existing data-services, see \ref{DataModel}. 
-%In section \ref{DAS}
-%we provides a detailed description of underlying design and components
-%used in DAS system. Our results are summarized in section \ref{Results}.
+We wanted to expand it further and apply to handle cross data-service queries and
+cover broad variety of relational and non-relational data-sources in 
+CMS within our distributed heterogeneous environment.
+%At the same time we wanted to simplify it up to the level that users will be
+%able to place free text-based queries (keyword based search) to look-up their desired data.
 
 \section{CMS data model\label{DataModel}}
+The CMS, the Compact Muon Solenoid experiment \cite{CMS} 
+is one of the two large general-purpose particle physics detectors built on 
+the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
+It is designed to explore the frontier of High Energy Physics and provide physicsists
+ability to look at conditions which were present at early stage of our Universe.
+More then 3000 physicists from 183 institutions reprepresenting almost 
+38 countries are involved in design, construction and maintenance of the experiment.
+
 The CMS distributed computing and data model \cite{CMSDataModel} 
 is designed to process
-and efficiently manage a PBs of data expected to be produced each year
+and efficiently manage a few PBs of data expected to be produced each year
 at LHC. The computing resources provided by members of CMS
@@
 to answer specific questions about data they serve, the ability to relate and search
-information among them was human task. The growing amount of information
+information among them was tedious human task. The growing amount of information
 and desire to make cross-service queries force us to design and develop a new
@@
 The design of DAS system was based on previous studies of CMS 
-data bookkeeping system (DBS) usage patterns \cite{DBS, DBS07}. We carefully analyzed user
-queries placed to DBS, their patterns, frequencies and latency. As a result of this
-we created a data discovery service \cite{DD} which came through several iteration
-of UI. At the end the end-user interface was based on DBS Query Language (DBS-QL)\cite{DBS-QL}
-with broad variety of template for presentation layer. A quick adoption
-and wide spread of DBS-QL give us confidence of choosen approach and QL syntax.
+data bookkeeping system (DBS) use cases \cite{DBS, DBS07}. We carefully analyzed user
+queries placed to DBS, their patterns, frequencies and latency. As a result
+we created a data discovery service \cite{DD} which came through several iterations
+of design choices. At the end the end-user interface was based on DBS 
+Query Language (DBS-QL)\cite{DBS-QL} which supports dynamic queries against
+DBS back-end.
+% and had syntax very closed to free-text based search queries. 
+The DBS-QL \cite{DBS-QL} syntax was based on SQL query language with one important
+difference that we took away from end-users any knowledge about relations of underlying schema.
+It was achieved by usage of short-path algorithm to establish necessary
+table joins based on provided information in user queries and auto-loaded
+schema from DB back-end and knowledge of foreign key relationships.
+The DBS-QL keys exposed to end-users were choosen based on jargon used by
+physicists and internally they're mapped 
+into database entitites and attributes, e.g. run key refers to
+run number attribute of run summary table.
+A quick adoption and wide spread of DBS-QL give us confidence of choosen approach 
+and provide a basis for building DAS system.
 
-Starting from those ground accomplishments we created DAS system as
-scalable and pluggable layer sitting on top of the existing data-services
-within CMS computing infrastructure. Data availability were deligated to individual
-data-services. Therefore, we treat DAS as aggregation and cache service 
-with the following characteristics:
+As was pointed out in \cite{Arms} a mixed content and 
+mixed meta-data and meta-data consistency should be considered as a whole in design 
+of the system to successful information discovery. 
+Starting from this ground we designed DAS system as
+additional layer on top of the existing data-services
+within CMS computing infrastructure by requiring the following:
 \begin{itemize}
@@
 data-service API was invoked, triggering DAS cache population. Once data become
-available in DAS cache user were able to see the results. This on demand
-query retrieval system, where data were retrieved and store into cache from
-back-end data service. We use DAS analytics DB to establish cache population
-for most common queries by using DAS robots.
+available in DAS cache users were able to see the results. This on demand
+system, where data were retrieved and store into cache from
+back-end data services, was supplemented by DAS robots (unix daemons) who
+were consulting DAS analytics database to enable data population into DAS
+cache for most common queries.
 }
@@
 \noindent
-Figure \ref{DAS_cache} summarize a overview of DAS architecture who consist of the
+The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of the
 following components:
 \begin{itemize}
-\item DAS core library with support of pluggable modules for data retrieval;
-\item DAS caching layer with long-term cache for aggregated results and short-term
-cache for presentation layer;
-\item DAS request service to handle user requests;
-\item DAS mapping DB with DAS/data-services mapping, notations mapping and
-mapping of DAS objects into data-presentation layer;
-\item DAS analytics DB with query statistics to enable pre-fetching 
+\item core library with support of pluggable modules for data retrieval;
+\item caching layer to store data-service output and aggregated results;
+\item request service to handle user requests;
+\item mapping DB to keep information about data-service APIs, their
+notations and mapping to DAS;
+\item analytics DB with query statistics to enable pre-fetching 
 strategies.
 \end{itemize}
-Further we proceed with a discussion of individual DAS components.
 
-\subsection{Dealing with data-services}
-DAS provided pluggable system to add a new CMS data-service. To do that we
-identify appropriate APIs, input and output parameters and keys, data-service
-format and mapping between data-service, DAS and other sub-systems. 
-This information were stored into DAS mapping DB. For instance,
-the Data Bookkeeping System (DBS) and data location system (Phedex) both
-provide information about CMS data files. In former case, DBS keeps information between
-files and other physics  objects, such as run and luminosity, while Phedex provides 
-information about file location, such as site and details about storage elements.
-Upon user request to find information about a file, based on forementioned
-information we were able to identify appropriate set of
-data-service APIs, look-up data, re-format their output to DAS notations and
-store results into DAS cache.\footnote{Even though we supported a data retrieval on demand
-basis we also store user queries into DAS analytics DB to identify
-a common patterns. Based on this information we organized a pre-fetching
-strategies.}
-Once data were retreived from data-services an additional step to aggregate
-matched objects in DAS cache were applied. For instance in above examples
-file information came from DBS and Phedex systems. 
-Since both data-services output contained file name as a common
-key, their output where merged and final result was stored into DAS cache.
-This merged (aggregated) results represented a DAS data object. Each DAS data
-object contains a standard DAS header, which identify look-up time,
-request url, API, parameters and method as well as response version, expiration
-time and checksum of the output.
+The DAS core communicates with different data-services and retrieve
+data on demand. This was done by invoking appropriate data-service APIs
+upon user requests. All user queries were written to Analytics DB
+for further analysis. The data-service output where re-mapped into
+DAS notations and if necessary aggregated inside of DAS cache.
+Since we didn't know a-priory what kind of data users will ask, 
+what kind of queries they will place against the system and what
+level of aggregation will be applied the choice
+of RDMS as DAS back-end was problematic\footnote{In addition DAS 
+didn't impose any requirements for transactions and persistency of the data.} 
+and we look around for other options. Through the analysis of available IT solutions, 
+such as file-based and memory caches, 
+key-value databases, documented-oriented databases we made our choice in favor 
+of the last technology. Among of them we evaluated CouchDB \cite{CouchDB} and 
+MongoDB \cite{MongoDB}. Both systems provide schema free document
+storage, replication and fail-over features, but our choice in favor of 
+MongoDB was obvious because its support of dynamic queries, 
+full indexies, including on inner objects and embedded arrays,
+and auto-sharding. Our prelimiary benchmarks shown that it can sustain
+the desired load and size for our meta-data information. We use MongoDB as a back-end 
+for the three main components: Mapping, Analytics and DAS cache databases.
 
-We designed DAS interface to be fully REST complaint \cite{REST}. It is not
-only simplify development of DAS clients (CLI, web interface, robot, etc.)
-but also consider DAS itself as own sub-component, i.e. place queries against
-DAS content, rather then data-service one.
-
-\subsection{Query Language\label{DAS-QL}}
-Upon success of DBS-QL \cite{DBS-QL} we adopted its syntax to DAS with a few minor
-changes. Here we briefly recall its syntax:
-\begin{equation}
-action\,\,\,
-key_1,\, key_2,\, ...\,\,\, where\,\,\,
-\langle key\rangle\, 
-\langle op\rangle\, 
-\langle value\rangle \,\,\, and|or ...
-\label{QL_syntax}
-\end{equation}
-Each DAS QL expression starts with special keyword indicating an action such as 
-$find,\, plot,\, summary$, followed by a set of selection keys. The action identify
-the behavior of the query, find to look-up data, plot to plot found data, summary to
-apply a summary view over provided selection keys. An optional conditions can be 
-specified via $where$ keyword following by condition expression.
-Each condition expression represented via $key\,\,\, operator\,\,\, value$ triplet, where
-$key$ is DAS selection key, $operator$ is conditional operator, 
-$=,\, >=,\, <,\, in,\, between$ and $value$ is a value applied to selection key.
-We also support brackets as well as $and$ and $or$ logical operators.
-The DAS selection and condition keys were identified based on analysis 
-of terminology used by our users. For instance, $file$ was used to either
-select file name or used in condition to identify file name provided by user.
-Here is a simple example of DAS QL expression:
-$$
-find\,\,\, file,\, run\,\,\, where\,\,\, site=T1\_CH\_CERN
-$$
-which represents a natural language question: 
-{\it find files and runs located at T1\_CH\_CERN site}.
-
-\subsection{Mapping}
-As been discussed above we used a DAS mapping DB to map data-service API into
-DAS as well as map provide a relations between different data-services.
-Upon identification of APIs which participate in DAS activity 
-we store a mapping between data-service notations
-into DAS mapping DB. This allows to map user input query parameters into
+DAS Mapping DB holds mapping between data-service APIs and DAS keys. 
+Upon identification of APIs which will participate in DAS activity 
+we store to Mapping DB the API name, input parameters, their type and accept 
+values as regular expression patterns as well as mapping between data-service 
+notations and DAS keys. It allows us to map user input query parameters into
 data-service API metrics as well as map back data-service output into DAS names.
-For instance, the DBS system used $logical\_file\_name$ notation, while in Phedex
-the same entity was named as $lfn$. Since both of them were referred to
-logical file name, we map both notations into $file$ notation in DAS.\footnote{We
-also deal with physical file names, but still used $file$ to refer to them in
-DAS notation, since simple regular expressions were used to identify $file$
+For instance, the DBS system \cite{DBS}
+used {\it logical\_file\_name} notation, while in PhEDEx
+\cite{PhEDEx}
+the same entity was named as {\it lfn}. Since both of them were referred to
+logical file name, we map both notations into DAS key {\it file}.\footnote{We
+also deal with physical file names, but still used {\it file} to refer to them in
+DAS notation, since simple regular expressions were used to identify {\it file}
 meaning from the provided value.}
 For each service we provide a mapping from its own data-format (JSON, XML, etc.) into
-common data-format used in DAS.\footnote{We used JSON as a DAS data format}
-We also store a mapping between DAS keywords and data-service API names to
+common data-format used in DAS.\footnote{We used JSON as a DAS data format.}
+We also store a mapping between DAS keys and data-service API names to
 establish appropriate calls to data-service upon provided DAS query. For example,
-when user specifies $file$ in their DAS query, we invoke DBS listFiles and Phedex
-fileReplicas APIs. Since all mapping kept in RDMS we were able to abstract the 
-DAS core library calls from hard coded data-service API calls and make it
-transparent to any participating data-services. Here is a logic of DAS core
-library:
-\begin{itemize}
-\item parse input DAS query
-\item map provided selection and condition keys into data-service APIs and
-input parameters
-\item invoke data-service APIs
-\item parse output results by mapping output notation into DAS keys
-\item store and aggregate them into DAS raw-cache
-\end{itemize}
+when user specifies {\it file} in their DAS query, we invoke DBS 
+{\it listFiles} and PhEDEx {\it fileReplicas} APIs.
 
-\subsection{Caching system and Analytics}
-The DAS caching system was divided into two different caches: raw-cache and hot-cache.
-The purpose of raw-cache was to store results coming from data services for
-as long as specified in data-service response header. While the purpose of hot-cache
-was to keep final DAS results corresponding to provided user query. We use a robot
-services (jobs) to feed raw-cache with highly requested queries. Here we will only
-discuss the raw-cache organization. 
-%\begin{figure}[htb]
-%\centering
-%\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
-%\caption{
-%DAS cache system.
-%}
-%\label{DAS_cache}
-%\end{figure}
-As it shown in figure \ref{DAS_cache} the DAS cache server has dual purpose.
-On one side it talks to DAS Analytics DB to trigger jobs for worker robot
-who update cache with most popular queries, on another it talks to DAS
-raw-cache back-end to either stored mapped results or merge them with 
-existing DAS objects in a cache. We should note that we use on-demand
-approach to answer user queries, but at the same time with a help of
-DAS Analytics DB we force worker robots (robots) to pre-fetch
-the most popular queries up-front.
+The DAS Analytics DB plays special role. It collects information
+about user requests placed to the system. Each request was recorded. Upon its
+decomposition into set of selection keys and conditions we also record
+API name and API input parameters and keep updating their counter for repeated
+calls. This allows to keep track of frequency of API calls and establish pre-fetch
+strategy for most requested APIs.
 
-We evaluated several technologies
-as a choice for DAS cache back-end, including relational MySQL DB \cite{MySQL} 
-and object oriented CouchDB \cite{CouchDB} and MongoDB \cite{MongoDB}.
-Each of them had its own pros and cons. For instance usage
-of MySQL provided a SQL queries (as in federated DBs, see \cite{FedDB}), but
-required auto-schema generation on a fly.\footnote{We did not apply any
-restriction to DAS QL and did not know in advance what our users will be
-querying.} While, using CouchDB and MongoDB, we were able to store
-data-service output naturally as documents objects into DB. Finally we decided
-to use MongoDB due to its flexible and
-powerful querying syntax, which were easy to map into DAS QL expressions.
+The DAS caching system was used to store data-service results and aggregated
+information in form of DAS records. As it shown in figure \ref{DAS_cache} 
+the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
+and aggregate data over there. The independent set of daemons, DAS robots, are
+constantly monitor Analytics DB for most popular requests and pre-fetch most
+common data in DAS cache.
+
+Each user request placed to the system was parsed and set of selection
+and condition keys were identified. They were mapped into set of
+data service APIs and their input parameters. DAS
+cache was checked for a similar set of selection and condition
+keys which can be placed earlier.\footnote{Since DAS doesn't hold
+any data by default we cannot rely on data being present in a cache.
+Moreover if we look-up for data first, we cannot tell if it was complete
+set since earlier queries may retrieve only portion of the requested data.}
+If the superset was found a data were
+retrieved from the cache, otherwise a data-service APIs were
+invoked and the results were collected and aggregated in DAS cache.
+
+%Now we present overall logic of DAS workflow and details of individual components.
+%Upon request to DAS the following steps were executed: 
+%\begin{itemize}
+%\item parse input DAS query
+%\item map provided selection and condition keys into data-service APIs and
+%input parameters
+%\item look-up similar queries in DAS cache 
+%\item invoke data-service APIs
+%\item parse output results by mapping output notation into DAS keys
+%\item store and aggregate them into DAS raw-cache
+%\end{itemize}
+
+Apart from previous DBS-QL syntax we decided to fully support free text-based queries for end-users.
+All of them were transformed into MongoDB query syntax, who was used internally in
+DAS core layer. It represents a set of selection keys and conditions as a JSON document.
+For example, a query
+{\it dataset site=T1\_CH\_CERN}
+where transformed into 
+{\it \{"fields" : ["dataset"], "spec" : \{"site" : "T1\_CH\_CERN"\}\}}.
+Such syntax was perfect fit for us, since it allows us to store all queries
+as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
+MongoDB query syntax is very reach. It supports advanced queries with 
+boolean expressions, conditional operators, regular expression, etc. providing
+a real power for our end-users.
+
+\subsection{Data-services and aggregation}
+DAS provided a pluggable interface to add new CMS data-services. 
+Each data-service plugin consists of its DAS map and individual data parser. 
+The data-service DAS map consists of information about data-service APIs,
+input and output parameters, data-service
+format and their mapping to DAS keys as well as relation to other sub-systems. 
+The information was stored into DAS Mapping DB. Below you can see a typical
+record of SiteDB \cite{SiteDB} data-service:
+\begin{verbatim}
+{"api": {"params": {"name": ""}, "name": "CMSNametoSE"}, 
+ "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
+              "api_param": "name", 
+              "das_key": "site"}], 
+ "_id": "4ae1f8dfe2194e4581000014", 
+ "system": "sitedb", 
+ "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
+              "key": "site", 
+              "map": "site.name"}]}
+\end{verbatim}
+It contains information about API name and parameters, mapping between API input
+parameter and DAS key, CMS system name and set of DAS keys this API is covering.
+We also used regular expression patterns, evaluated by python during run time
+execution, to verify provided set of input paremters and DAS keys.
+
+The DAS Mapping DB becomes an authorative
+source of information about all data-services participating in DAS and their
+relationships. Based on this information we were able to retrieve appropriate data
+from different data-services, re-map them into DAS notations and
+aggregate them on demand.
+For instance, the Data Bookkeeping System (DBS) and data location system (PhEDEx)
+provided information about CMS data files. In former case, DBS holds information about
+files and their relationship to other physics objects, such as run and luminosity, 
+while PhEDEx stores information about the file location, such as sites which
+holds copy of the files and details about site storage elements.
+Upon a user request to find a file, we were able to identify appropriate set of
+DBS and PhEDEx APIs, look-up data, re-format their output to DAS notations and
+store results into DAS cache.
+Once data were retreived from data-services an additional step to aggregate
+matched objects in DAS cache were applied. For instance, both data-service outputs 
+contained a file name in them. We use it as a common key to merge records in DAS cache.
+This merged (aggregated) results represent DAS data objects. Each DAS data
+object contains a standard DAS header, which identify look-up time,
+request url, input parameters and API method as well as response version, expiration
+time and checksum of the output. Here is an example of individual DAS record:
+\begin{verbatim}
+{"_id": "4ae1e7ffe2194e4243000003", 
+ "site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
+         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
+ "das": [{"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
+          "ctime": 0.95439004898071289, "expire": 1256362175.7692671, 
+          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSAMName", 
+          "timestamp": 1256318975.7692671, "system": ["sitedb"], 
+          "api": ["CMStoSAMName"], "version": "", "selection_keys": []}, 
+         {"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
+          "ctime": 1.2158100605010986, 
+          "api": ["CMStoSiteName"], 
+          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSiteName", 
+          "timestamp": 1256318974.7774379, "system": ["sitedb"], 
+          "expire": 1256362174.7774379, "version": "", "selection_keys": []}
+        ]
+}
+\end{verbatim}
+This DAS record represents aggregated information between two SiteDB APIs, 
+{\it CMStoSAMName} and {\it CMStoSiteName}, where API results are merged 
+under {\it site} key. The information provided by {\it das} key shows 
+DAS header, which contains information about API name, location, execution 
+time and lifetime of the data.
+Usage of this information was valuable contribution in debugging process of
+data-services themselves, e.g. identification of not-matched records, 
+latency studies, etc.
+
+We used RESTful model \cite{REST} for all DAS cache server APIs.
+It allows us to share core library and other components during development of 
+various DAS clients, e.g. CLI, web interface, robots, etc. and hides
+complexity of access to individual data-services and their security policies
+as well as proper organize caching solution.
+For instance, to access on-line Run Summary information we were forced to use
+Grid certificates. Instead of propagating user certificate to data-service
+the DAS cache server works as a proxy to this server and used host certficiate 
+to serve such requests on behalf of user requests.
 
 \section{Results\label{Results}}
-DAS was deployed with the following CMS data-services:
+DAS was deployed to work with the following CMS data-services:
 \begin{itemize}
@@
 about CMS meta-data;
-\item PhEDEx \cite{Phedex}, Physics Experiment Data Export project which
+\item PhEDEx \cite{PhEDEx}, Physics Experiment Data Export project which
 provides the data placement and the file transfer system for the CMS experiment;
+\item SiteDB \cite{SiteDB} is a catalogue of all CMS computing Tiers. 
+It records the CMS resources at the site, the resources pledged for the 
+future, and it keeps track of CMS personnel at each site, including the 
+roles and duties they fulfill in the collaboration;
 \item RunSummary DB \cite{RunSummary} collects information about run and triggers
 conditions during data-taking;
-\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
-\item DQ \cite{DQ}, Data Quality data-service which collects information
-about detector conditions during data taking;
-\item Overview \cite{Overview}, collects information about CMS
-transfer rate, etc.;
+%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
+%\item DQ \cite{DQ}, Data Quality data-service which collects information
+%about detector conditions during data taking;
+%\item Overview \cite{Overview}, collects information about CMS
+%transfer rate, etc.;
 \item Dashboard \cite{Dashboard} project for LHC experiments aims to 
@@
 Each data-service has its own scope, size and lifetime. For instance the data
-in Phedex are transient due to constant migration of CMS data. The DBS system
-were divided into dozen of individual DBs, whose size varied at the level of
-few tens of GBs.
+in PhEDEx are transient due to constant migration of CMS data with daily DB size of
+around few tens of GBs. The DBS system were divided into dozen of individual instances, 
+who has been used by different production teams for various bookkeeping tasks.
+Figure \ref{db_size} shows current size of meta-data from aforementioned data-service.
+\begin{figure}[htb]
+\centering
+\includegraphics[width=100mm]{db_size.pdf}
+\caption{
+Database size of CMS data-services participating in DAS in November 2009, before
+data taking.
+}
+\label{db_size}
+\end{figure}
+%Typical
+%size of individual DBS instance was around of fwe GBs, with overall DBS size of
+%around of 50 GBs.\footnote{This include table and index sizes.}
+
+At the moment we don't have exact numbers for amount of data which will stored to DAS,
+but based on test performed during CRAFT09 \cite{CRAFT09}
+exercise we produced 0.3PB of data which
+translate into $\sim50$ GB of meta-data stored into DBS system. Therefore
+we can anticipate to collect $\sim500$GB of meta-data per year during data-taking period
+at LHC.
+%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
+%billion events, 330TB in 32 datasets:
+%AlCa/Calibration RAW: 926 million events, 29TB
+%Physics RAW: 1.24 billion events, 147TB
+%PromptReco: 1.24 billion events, 111TB
+%Bulk ALCARECO: 2.34 billion events, 11TB
+%Express FEVT: 105 million events, 24TB
+%Express ALCARECO: 153 million events, 0.99TB
+%HLTMON FEVTHLTALL: 31 million events, 7.4TB
+%Above event numbers have signicant overlap by design, 
+%2.2 billion unique events, including 524 million cosmic muon
+%triggers
+% 0.3PB - 50GB meta-data in DBS
+% 3PB/y - x GB
+
+We did preliminary studies of DAS performance based on information stored
+in 2 systems, DBS and PhEDEx. We benchmark time required to collect and 
+aggregate block objects information from both system, see Table \ref{DAS_benchmark}.
+
+\begin{table*}[hbt]
+\centering
+\begin{tabular}{llllll}\hline
+\hline
+
+System & Format & Size & Records & Elapsed time & Elapsed time \\
+& & & & no cached data & w/ cached data \\
+\hline
+DAS (DBS) & XML & 187MB & 346820 & 165 sec & 2.53 sec \\
+DAS (PhEDEx) & JSON & 164MB & 230035 & 177 sec & 2.42 sec \\
+DAS (aggregate) & JSON & 357MB & 348866 & 386 sec & 41.7 sec \\
+\hline
+\hline
+\end{tabular}
+\caption{DAS benchmark results. We performed a stress test with 2 CMS system, which
+both deliver block object information. The total number of DAS records is calculated
+as number of aggregated records and left over from both systems.}
+\label{DAS_benchmark}
+\end{table*}
+
+We measured elapsed time as the following:
+\begin{itemize}
+\item[]
+{\it
+Elapsed time = retrieval time + parsing time + remapping time 
+        + cache insertion/indexing time 
+        + (aggregation time) + (output creation time)
+}
+\end{itemize}
+Here the {\it retrieval time} is a time required to access data from remote data-services,
+{\it parsing time} is a time required to read and parse received data, {\it remapping time}
+is a time to convert notations used by data-service to DAS ones for every object
+we parse, {\it cache insertion} and {\it indexing time} represents time spend to add objects into
+the DAS cache, {\it aggregation time} is required to merge objects into DAS records based
+on their common key (block name in this case) and {\it output creation time}
+is a time required to write DAS records to disk. Please note that last two
+components, {\it aggregation} and {\it output creation time}, were only applied to
+final DAS step and not to individual DAS components.
+
+The individual DAS components (DBS and PhEDEx) shown roughly the same performance.
+For example, in case of PhEDEx the total time spent in DAS was 177 seconds, where
+majority of the time was spent on retrieval, 83 seconds, and parsing, 29 seconds.
+The remaining 65 seconds was distributed among remapping, cache insertion, indexing. The
+numbers for DBS system were alike. As you can see from Table \ref{DAS_benchmark},
+the time spent to read DAS records (last column) from the cache was quite
+reasonable, roughly 2.5 seconds from both systems. While final time to
+get DAS records on disk was below 42 seconds. This was mostly due to I/O operations.
+
+Based on this results we measured that we can achieve 900 docs/sec for creation of DAS
+in a cache and 8500 docs/sec for reading DAS records back, which is suitable for our
+needs.\footnote{We've been satisfied with such performance,
+even though our measurements shown that it can still be speed up by factor of
+10 if we will use C++ DB driver.}
 
 \section{Summary}
+We presented new data aggregation service (DAS) developed for CMS High-Energy experiment
+at LHC, CERN, Geneva, Switzerland. It was designed to provide caching and
+aggregation layer on top of existing relational and non-relation data-service
+mostly in real time fashion. All the data were retrieved on demand basis,
+while data pre-fetching of most common queries is in our plans. The DAS system 
+is under commissioning phase right now and prelimiary performance studies
+show it can accommodate needs of CMS experiment as general purpose meta-data
+service. The CMS experiment is started to
+collect data from November 2009 and we expect to write about PB of data each
+year to tape. In addition, the Monte-Carlo samples will be produced at this scale.
+Based on test studies performed in CMS, it can translate to the order of
+$\sim500$GB of meta-data each year. The DAS should sustain such load and provide
+interface as data discovery service in CMS.
 
@@
 
-\bibitem{CMSDataModel} A. Fanfani et. al.,
+\bibitem{AMI}
+Altas AMI web portal:
+http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html
+
+\bibitem{CMS} CMS Collaboration R. Adolphi et al., The CMS experiment at the CERN LHC, JINST, 0803, S08004 (2008)
+
+\bibitem{CMSDataModel} 
+.Grandi, D.Stickland, L.Taylor et al., The CMS Computing Model, CERN-LHCC-2004- 035/G-083 (2004);
+A. Fanfani et. al.,
 ``Distributed Analysis in CMS'', to be published in Journal of Grid Computing.
@@
 
-\bibitem{Phedex}
+\bibitem{PhEDEx}
 Rehn et. al.,
 ``PhEDEx high-throughput data transfer management system'', CHEP06, Mumbai, India.
+R. Egeland et al., Data transfer infrastructure for CMS data taking, Proceedings of Science,
+PoS(ACAT08)033 (2008) ;
+L. Tuura et al., Scaling CMS data transfer system for LHC start-up, J.Phys.Conf.Ser, 119, 072030 (2008)
 
 \bibitem{RunSummary}
+https://cmswbm.web.cern.ch/cmswbm/cmsdb/servlet/RunSummary
 
 \bibitem{SiteDB}
+https://cmsweb.cern.ch/sitedb/
 
-\bibitem{LumiDB}
-https://twiki.cern.ch/twiki/bin/view/CMS/CMS-DMWM-DBS-Luminosity
+%\bibitem{LumiDB}
+%https://twiki.cern.ch/twiki/bin/view/CMS/CMS-DMWM-DBS-Luminosity
 
-\bibitem{DQ}
+%\bibitem{DQ}
+%Data-Quality DB reference
 
-\bibitem{Overview}
-https://cmsweb.cern.ch/overview/
+%\bibitem{Overview}
+%https://cmsweb.cern.ch/overview/
 

### 2010-05-04 src/python/DAS/utils/utils.py
Work around ObjectId in JSON, see http://groups.google.com/group/mongodb-user/browse_thread/thread/b96f121ec5eaba72
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1661 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 General set of useful utilities used by DAS
 """
 
-__revision__ = "$Id: utils.py,v 1.35 2009/10/23 19:37:15 valya Exp $"
-__version__ = "$Revision: 1.35 $"
+__revision__ = "$Id: utils.py,v 1.36 2009/11/03 16:28:11 valya Exp $"
+__version__ = "$Revision: 1.36 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -22,6 +22,7 @@ import time
 import types
 import traceback
 from itertools import groupby
+from pymongo.objectid import ObjectId
 
 class dict_of_none (dict):
     """Define new dict type whose missing keys always assigned to None"""
@@ -113,6 +114,11 @@ def genkey(query):
         # prior python 2.5
         keyhash = md5.new()
     if  type(query) is types.DictType:
+        if  query.has_key('spec') and query['spec'].has_key('_id'):
+            val = query['spec']['_id']
+            if  isinstance(val, ObjectId):
+                val = val.url_encode()
+                query['spec']['_id'] = val
         query = json.dumps(query)
     keyhash.update(query)
     return keyhash.hexdigest()

### 2010-05-04 doc/paper.tex
More changes to improve layout of the paper and my bad English
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1669 4525493e-7705-40b1-a816-d608a930855b

@@
 non-relational data sources to answer cross data-service queries 
-to explore metadata associated with petabytes of experimental data. 
+and explore meta-data associated with petabytes of experimental data. 
 We combine simplpicity of keyword-based search with precision of RDMS
@@
 in last several years. The federated DB \cite{FedDB} by IBM unifies data coming 
-from different RDMS into federated DB where SQL queries can be placed to search desired
-data. The other approaches \cite{DBXplorer, QueryAnswer}, influenced by great success
-of search engines, explored ability to use keyword search algorithms over relataional RDMS
-by indexing DB content. In former case, a problem with understanding 
+from different RDMS systems into federated DB where SQL queries can be placed to 
+search desired data. The other approaches \cite{DBXplorer, QueryAnswer}, influenced by 
+great success of search engines, explored ability to use keyword search algorithms 
+over relataional RDMS by indexing DB content. In former case, a problem with understanding 
 of underlying schema and imposing relational conditions in a query still exists, while 
-in later queries can be expressed via simple keyword based search, even though exact match
-of provided keywords is expected. Although keyword based search is
-intuitive and easy to use, with respect to HEP and other domains it is not sufficient. 
-For example, it does not provide ability to place condition expressions,
-%keys (DB entities/attributes) and their numerical values, 
-e.g. $a > 1$, since
-it only index a text based values. Therefore proximity of results becomes very poor
-or impossible in case of numbers, aggregate functions, etc. To address those issues
+in later queries can be expressed via simple keyword based search, even though the exact match
+of provided keywords is expected. Although those solutions exist,
+with respect to HEP and other domains they are not sufficient.
+In a case of federeated database exact knowledge of schema is required. This means
+that end-users should be aware of database internal sctructure in order to use SQL.
+While the proposed keyword search based systems
+work in confined environment and do not provide ability to use condition expressions
+over relational data. For instance, as quoted in \cite{QueryAnswer}, if someone asks about 
+{\it Woody Allen} a set of database records can be found to satisfy this request.
+Such approach cannot be used to answer the logical questions, such as 
+{\it find me total number of files whose size more then 20 but less then 100 taken from january
+to february of 2009} which is easy to address using SQL syntax.
+
+To address those issues
 an attempt to build a simple, intuitive and flexible query language was introduced
@@
 
-We wanted to expand it further and apply to handle cross data-service queries and
-cover broad variety of relational and non-relational data-sources in 
-CMS within our distributed heterogeneous environment.
-%At the same time we wanted to simplify it up to the level that users will be
-%able to place free text-based queries (keyword based search) to look-up their desired data.
+We wanted to expand this approach further and apply it 
+to handle cross data-service queries from broad variety 
+of relational and non-relational data-sources within 
+our distributed heterogeneous environment.
 
 \section{CMS data model\label{DataModel}}
-The CMS, the Compact Muon Solenoid experiment \cite{CMS} 
+The CMS, Compact Muon Solenoid, experiment \cite{CMS} 
 is one of the two large general-purpose particle physics detectors built on 
@@
 in our distributed environment? Even though individual data-services were designed
-to answer specific questions about data they serve, the ability to relate and search
+to answer specific questions about data they serve, the ability to search and relate
 information among them was tedious human task. The growing amount of information
@@
 \section{Data Aggregation System\label{DAS}}
-The design of DAS system was based on previous studies of CMS 
-data bookkeeping system (DBS) use cases \cite{DBS, DBS07}. We carefully analyzed user
-queries placed to DBS, their patterns, frequencies and latency. As a result
+The design of DAS system was based on previous studies of CMS Data 
+Bookkeeping System (DBS) use cases \cite{DBS, DBS07}. We carefully analyzed user
+requests and queries placed to DBS, their patterns, frequencies and latency. As a result
 we created a data discovery service \cite{DD} which came through several iterations
@@
 Query Language (DBS-QL)\cite{DBS-QL} which supports dynamic queries against
-DBS back-end.
-% and had syntax very closed to free-text based search queries. 
-The DBS-QL \cite{DBS-QL} syntax was based on SQL query language with one important
-difference that we took away from end-users any knowledge about relations of underlying schema.
-It was achieved by usage of short-path algorithm to establish necessary
-table joins based on provided information in user queries and auto-loaded
-schema from DB back-end and knowledge of foreign key relationships.
+DBS back-end. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language with one 
+important difference, we took away any knowledge about relations of underlying schema
+from end-users. It was achieved by usage of Dijkstras shortest path algorithm 
+to establish necessary table joins based on provided information in user queries and
+knowledge of foreign key relationships from auto-loaded database schema.
 The DBS-QL keys exposed to end-users were choosen based on jargon used by
-physicists and internally they're mapped 
-into database entitites and attributes, e.g. run key refers to
-run number attribute of run summary table.
-A quick adoption and wide spread of DBS-QL give us confidence of choosen approach 
+physicists. They're internally mapped into database entitites and attributes, 
+e.g. {\it run} DBS-QL key refers to run number attribute of run summary table.
+A quick adoption and wide usage of DBS-QL gave us confidence of choosen approach 
 and provide a basis for building DAS system.
@@
 DAS architecture diagram. It consists of DAS cache server, mapping and analytics DBs.
-Once query was placed to DAS an associated
-data-service API was invoked, triggering DAS cache population. Once data become
-available in DAS cache users were able to see the results. This on demand
+Once query was placed to DAS the associated
+data-service APIs were invoked, triggering DAS cache population. Once data become
+available and aggregated in DAS cache users were able to see the results. This on demand
 system, where data were retrieved and store into cache from
@@
 storage, replication and fail-over features, but our choice in favor of 
-MongoDB was obvious because its support of dynamic queries, 
+MongoDB was obvious because of its support of dynamic queries, 
 full indexies, including on inner objects and embedded arrays,
@@
 notations and DAS keys. It allows us to map user input query parameters into
-data-service API metrics as well as map back data-service output into DAS names.
+data-service API metrics as well as map back data-service output into DAS notations.
 For instance, the DBS system \cite{DBS}
@@
 We also store a mapping between DAS keys and data-service API names to
-establish appropriate calls to data-service upon provided DAS query. For example,
+establish appropriate calls to data-services upon provided DAS query. For example,
 when user specifies {\it file} in their DAS query, we invoke DBS 
-{\it listFiles} and PhEDEx {\it fileReplicas} APIs.
+{\it listFiles} and PhEDEx {\it fileReplicas} APIs, respectively.
 
-The DAS Analytics DB plays special role. It collects information
+The DAS Analytics DB plays a special role. It collects information
 about user requests placed to the system. Each request was recorded. Upon its
 decomposition into set of selection keys and conditions we also record
-API name and API input parameters and keep updating their counter for repeated
-calls. This allows to keep track of frequency of API calls and establish pre-fetch
-strategy for most requested APIs.
+API name and API input parameters into Analytics database and 
+keep updating their counter information for repeated calls. 
+This allows to keep track of frequency of API calls and establish 
+pre-fetch strategies for most common queries.
 
@@
 contained a file name in them. We use it as a common key to merge records in DAS cache.
-This merged (aggregated) results represent DAS data objects. Each DAS data
-object contains a standard DAS header, which identify look-up time,
+This merged (aggregated) results represent DAS records. Each DAS
+record contains a standard DAS header, which identify look-up time,
 request url, input parameters and API method as well as response version, expiration
@@
 This DAS record represents aggregated information between two SiteDB APIs, 
-{\it CMStoSAMName} and {\it CMStoSiteName}, where API results are merged 
-under {\it site} key. The information provided by {\it das} key shows 
+{\it CMStoSAMName} and {\it CMStoSiteName}, where API results were merged 
+under the {\it site} key. The information provided by {\it das} key shows 
 DAS header, which contains information about API name, location, execution 
@@
 the DAS cache server works as a proxy to this server and used host certficiate 
-to serve such requests on behalf of user requests.
+to serve such requests on behalf of the users.
 
@@
 We did preliminary studies of DAS performance based on information stored
-in 2 systems, DBS and PhEDEx. We benchmark time required to collect and 
+in 2 systems, DBS and PhEDEx and benchmark time required to collect and 
 aggregate block objects information from both system, see Table \ref{DAS_benchmark}.
@@
 \end{tabular}
-\caption{DAS benchmark results. We performed a stress test with 2 CMS system, which
-both deliver block object information. The total number of DAS records is calculated
-as number of aggregated records and left over from both systems.}
+\caption{DAS benchmark results. 
+We are measured a time to fetch, parse and aggregate block information
+from DBS and PhEDEx systems via their APIs. In both cases a full block 
+table content, 347K and 230K records, respectively, were used to populate
+DAS cache. The total number of DAS records were calculated
+as number of aggregated records plus left over, non-matched records,
+from both systems.}
 \label{DAS_benchmark}
@@
 
-We measured elapsed time as the following:
+We measured elapsed time of this operation as the following:
 \begin{itemize}
@@
 components, {\it aggregation} and {\it output creation time}, were only applied to
-final DAS step and not to individual DAS components.
+final DAS step and not to time spent in individual data-services.
 
@@
 get DAS records on disk was below 42 seconds. This was mostly due to I/O operations.
+All operations were performed on a single CPU core and obviously can be
+paralelized in a future.
 
-Based on this results we measured that we can achieve 900 docs/sec for creation of DAS
-in a cache and 8500 docs/sec for reading DAS records back, which is suitable for our
+Based on this results we measured that we can achieve 900 docs/sec rate
+for inserting/aggregating data in DAS cache and 8500 docs/sec for reading 
+DAS records back, which is suitable for our
 needs.\footnote{We've been satisfied with such performance,
 even though our measurements shown that it can still be speed up by factor of
-10 if we will use C++ DB driver.}
+10 if we will use C++ DB driver instead of python one used in this prototype system.}
 

### 2010-05-04 src/python/DAS/utils/das_config.py
Added new option to specify bulk_update size
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1670 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 Config utilities
 """
 
-__revision__ = "$Id: das_config.py,v 1.28 2009/10/14 15:19:05 valya Exp $"
-__version__ = "$Revision: 1.28 $"
+__revision__ = "$Id: das_config.py,v 1.29 2009/11/05 18:13:59 valya Exp $"
+__version__ = "$Revision: 1.29 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -48,6 +48,7 @@ def das_readconfig(dasconfig=None):
     configdict['mongocache_dbhost'] = config.get('mongocache', 'dbhost', 'localhost')
     configdict['mongocache_dbport'] = int(config.get('mongocache', 'dbport', '27017'))
     configdict['mongocache_dbname'] = config.get('mongocache', 'dbname', 'das')
+    configdict['mongocache_bulkupdate_size'] = config.getint('mongocache', 'bulkupdate_size')
     configdict['mongocache_lifetime'] = config.getint('mongocache', 'lifetime')
 
     configdict['filecache_dir'] = config.get('filecache', 'dir', '')
@@ -124,6 +125,7 @@ def das_writeconfig():
     config.set('mongocache', 'dbhost', 'localhost')
     config.set('mongocache', 'dbport', '27017')
     config.set('mongocache', 'dbname', 'das')
+    config.set('mongocache', 'bulkupdate_size', 5000)
 
     config.add_section('filecache')
     dbdir  = os.path.join(os.environ['DAS_ROOT'], 'cache')

### 2010-05-04 src/python/DAS/core/das_mongocache.py
For cache_size use bulkupdate_size parameter passed via config
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1671 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 DAS mongocache wrapper.
 """
 
-__revision__ = "$Id: das_mongocache.py,v 1.26 2009/11/03 16:30:00 valya Exp $"
-__version__ = "$Revision: 1.26 $"
+__revision__ = "$Id: das_mongocache.py,v 1.27 2009/11/05 18:24:48 valya Exp $"
+__version__ = "$Revision: 1.27 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -252,6 +252,7 @@ class DASMongocache(Cache):
         self.dbhost  = config['mongocache_dbhost']
         self.dbport  = config['mongocache_dbport']
         self.limit   = config['mongocache_lifetime']
+        self.cache_size = config['mongocache_bulkupdate_size']
         self.dbname  = getarg(config, 'mongocache_dbname', 'das')
         self.colname = 'cache'
         self.logger  = config['logger']
@@ -264,7 +265,6 @@ class DASMongocache(Cache):
         self.conn    = Connection(self.dbhost, self.dbport)
         self.db      = self.conn[self.dbname]
         self.col     = self.db[self.colname]
-        self.cache_size = 100 # TODO: I can pass it via config
         
     def is_expired(self, query):
         """

### 2010-05-04 src/python/DAS/tools/das_cli.py
Re-factor chunk of DAS workflow excutaion to make it work within profiler
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1672 4525493e-7705-40b1-a816-d608a930855b

@@ -4,8 +4,8 @@
 """
 DAS command line interface
 """
-__revision__ = "$Id: das_cli.py,v 1.17 2009/10/20 15:00:55 valya Exp $"
-__version__ = "$Revision: 1.17 $"
+__revision__ = "$Id: das_cli.py,v 1.18 2009/11/06 00:05:08 valya Exp $"
+__version__ = "$Revision: 1.18 $"
 __author__ = "Valentin Kuznetsov"
 
 import time
@@ -75,6 +75,21 @@ def nooutput(input_results):
     """Just iterate over generator, but don't print it out"""
     for elem in input_results:
         pass
+
+def run(DAS, query, idx, limit, nooutput, plain, debug):
+    """
+    Execute DAS workflow for given set of parameters.
+    We use this function in main and in profiler.
+    """
+    results = DAS.result(query, idx, limit)
+    if  not nooutput:
+        if  plain:
+            for item in results:
+                print item
+        else:
+            dump(results, idx)
+    else:
+        nooutput(results)
 #
 # main
 #
@@ -109,52 +124,24 @@ if __name__ == '__main__':
         keys.sort()
         for key in keys:
             print key
-#    elif  opts.view:
-#        if  opts.view == 'all':
-#            view = None
-#        else:
-#            view = opts.view
-#        for name, query in DAS.get_view(view).items():
-#            print 'view name: %s' % name
-#            print 'DAS query: %s' % query
-#    elif  opts.createview:
-#        vlist = opts.createview.split(',')
-#        name  = vlist[0]
-#        query = ','.join(vlist[1:])
-#        print "Creating a view '%s' with query '%s'" % (name, query)
-#        DAS.create_view(name, query)
-#    elif  opts.updateview:
-#        vlist = opts.updateview.split(',')
-#        name  = vlist[0]
-#        query = ','.join(vlist[1:])
-#        print "Updating a view '%s' with query '%s'" % (name, query)
-#        DAS.update_view(name, query)
-#    elif  opts.deleteview:
-#        name = opts.deleteview
-#        DAS.delete_view(name)
-#        print "View '%s' has been deleted" % name
     elif query:
+
+        idx    = opts.idx
+        limit  = opts.limit
+        output = opts.nooutput
+        plain  = opts.plain
 
         if  opts.profile:
             import hotshot                   # Python profiler
             import hotshot.stats             # profiler statistics
-            print "Start DAS in profile mode"
             profiler = hotshot.Profile("profile.dat")
-            profiler.run("DAS.result(query)")
[…]

### 2010-05-04 src/python/DAS/core/qlparser.py
Change assignment to fields/spec while parsing the free text-based query, such that passing block will go to block.name=* rather then to fields. This will allow to use loose constrains to be match this patterns
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1735 4525493e-7705-40b1-a816-d608a930855b

@@ -9,8 +9,8 @@ tests integrity of DAS-QL queries, conversion routine from DAS-QL
 syntax to MongoDB one.
 """
 
-__revision__ = "$Id: qlparser.py,v 1.29 2009/11/18 17:06:16 valya Exp $"
-__version__ = "$Revision: 1.29 $"
+__revision__ = "$Id: qlparser.py,v 1.30 2009/12/01 15:15:26 valya Exp $"
+__version__ = "$Revision: 1.30 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -361,9 +361,26 @@ class MongoParser(object):
                                     condlist.append(cdict)
                     except:
                         pass
+            else:
+                if  word not in skeys:
+                    skeys.append(word)
             idx += 1
         spec = mongo_exp(condlist)
-        mongo_query = dict(fields=skeys, spec=spec)
+        # loop over skeys and add '*' value for each key into spec
+        for key in skeys:
+            insert = 0
+            for system in self.map.list_systems():
+                try:
+                    lkeys = self.map.lookup_keys(system, key)
+                    for nkey in lkeys:
+                        if  not spec.has_key(nkey):
+                            spec[nkey] = '*'
+                            insert = 1
+                except:
+                    pass
+            if  not insert:
+                spec[key] = '*'
+        mongo_query = dict(fields=None, spec=spec)
         self.analytics.add_query(query, mongo_query)
         return mongo_query
 

### 2010-05-04 doc/paper.tex
Added new section DAS queries
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1736 4525493e-7705-40b1-a816-d608a930855b

@@
 \section{Related Work\label{RelatedWork}}
-Even though the idea of querying relation databases via keyword based search
-algorithms is not knew it is still under significant activity in computer
-science domain. A few alternative solutions has been proposed to address this issue
-in last several years. The federated DB \cite{FedDB} by IBM unifies data coming 
-from different RDMS systems into federated DB where SQL queries can be placed to 
-search desired data. The other approaches \cite{DBXplorer, QueryAnswer}, influenced by 
-great success of search engines, explored ability to use keyword search algorithms 
-over relataional RDMS by indexing DB content. In former case, a problem with understanding 
-of underlying schema and imposing relational conditions in a query still exists, while 
-in later queries can be expressed via simple keyword based search, even though the exact match
-of provided keywords is expected. Although those solutions exist,
-with respect to HEP and other domains they are not sufficient.
-In a case of federeated database exact knowledge of schema is required. This means
-that end-users should be aware of database internal sctructure in order to use SQL.
-While the proposed keyword search based systems
-work in confined environment and do not provide ability to use condition expressions
-over relational data. For instance, as quoted in \cite{QueryAnswer}, if someone asks about 
-{\it Woody Allen} a set of database records can be found to satisfy this request.
-Such approach cannot be used to answer the logical questions, such as 
-{\it find me total number of files whose size more then 20 but less then 100 taken from january
-to february of 2009} which is easy to address using SQL syntax.
+Even though the idea of keyword queries over the relational database(s)
+%querying relation databases via keyword based search
+%algorithms 
+is not knew it is still under discussion in computer science domain.
+A few alternative solutions has been proposed to address this issue
+in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
+i.e. retrieval of only documents that contain all query keywords, has
+been discussed. It requires that all keywords are matched in a row from
+single table or joined tables. While \cite{QueryAnswer} make a step forward
+by considering information around matched values and presents customized answers to
+end-users over the generated new database. Those and other works\footnote{
+For more information about
+keyword based search over relational databases we redirect users to reference in
+\cite{DBXplorer, QueryAnswer}.} 
+are based on some kind of the graph over the underlying database and
+moreover use a generated database out of the pristine one.
+The proximity of the results in those techniques 
+are mostly confined to the text-based values stored in a database, since the
+keyword based search with numerical values 
+need the precise knowledge of the related entity in a schema. For example, typing
+100 can lead to inappropriate matches, such as row ids, and in order to be
+answered correctly it requires additional keyword, clarifying its context, 
+which should be matched with particular
+entity (table.column) in the underlying database.
+Therefore such approaches cannot be used to answer the logical questions, such as  
+{\it find me total number of files whose size more then 20 but less then 
+100 taken from january to february of 2009} which can be easily accomplished
+using SQL syntax.
 
-To address those issues
-an attempt to build a simple, intuitive and flexible query language was introduced
-in \cite{DBS-QL, AMI}.
-%\footnote{The ATLAS experiment at CERN, has been developed
-%AMI web portal \cite{AMI} with similar functionality.}
-It represented a power of SQL while
-hiding underlying relational schema from the end-users. As a results
-a human questions were intuitively mapped into simple queries. For example,
-the question
+To address those limitations an attempt to build a simple, intuitive and 
+flexible query language was introduced in \cite{DBS-QL}\footnote{A similar but less
+sophisticated approach was used in \cite{AMI}}.
+It represents a power of SQL while hiding underlying relational schema from 
+the end-users. As a results a human questions were intuitively mapped into 
+simple queries. For example, the question
 {\it I'm looking for files who contain data taken on certain date and located at
@@
 \noindent
+Due to this set of requirements, we rule out the choice of RDMS as DAS back-ends
+for several reasons. We agreed to support dynamic queries, data
+aggregation on demand and pluggable data-services. Therefore a choice 
+of concrete schema for data storage was problematic. Moreover we didn't impose
+any transactions and data persistency in DAS. This allows us to look 
+for alternative IT solutions.
+Through the analysis of available options, such as a file-based and memory caches, 
+key-value databases, documented-oriented databases we made our choice in favor 
+of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
+MongoDB \cite{MongoDB}. Both systems provide schema free document
+storage, replication and fail-over features, but our choice in favor of 
+MongoDB was obvious, due to its support of dynamic queries, 
+full indexies, including inner objects and embedded arrays,
+and auto-sharding. Our prelimiary benchmarks shown that it can sustain
+the desired load and size for our meta-data information. We used 
+document-oriented database, MongoDB, as a back-end for the three 
+main DAS components: Mapping, Analytics and DAS cache databases, 
+which will be discussed next.
+
 The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of the
 following components:
-\begin{itemize}
-\item core library with support of pluggable modules for data retrieval;
-\item caching layer to store data-service output and aggregated results;
-\item request service to handle user requests;
-\item mapping DB to keep information about data-service APIs, their
+core library with support of pluggable modules for data retrieval;
+caching layer to store data-service output and aggregated results;
+request service to handle user requests;
+mapping DB to keep information about data-service APIs, their
 notations and mapping to DAS;
-\item analytics DB with query statistics to enable pre-fetching 
+analytics DB for query analysis and identification of pre-fetch 
 strategies.
-\end{itemize}
 
+\subsection{DAS core}
 The DAS core communicates with different data-services and retrieves
-data on demand. This was done by invoking appropriate data-service APIs
-upon user requests. All user queries were written to Analytics DB
-for further analysis. The data-service output were re-mapped into
-DAS notations and, if necessary, aggregated inside of DAS cache.
-Since we didn't know a-priory what kind of data users will ask, 
-what kind of queries they will place against the system and what
-level of aggregation will be applied the choice
-of RDMS as DAS back-end was problematic\footnote{In addition DAS 
-didn't impose any requirements on transactions and persistency of the data.} 
-and we look around for other options. Through the analysis of available IT solutions, 
-such as file-based and memory caches, 
-key-value databases, documented-oriented databases we made our choice in favor 
-of the last technology. Among of them we evaluated CouchDB \cite{CouchDB} and 
-MongoDB \cite{MongoDB}. Both systems provide schema free document
-storage, replication and fail-over features, but our choice in favor of 
-MongoDB was obvious because of its support of dynamic queries, 
-full indexies, including on inner objects and embedded arrays,
-and auto-sharding. Our prelimiary benchmarks shown that it can sustain
-the desired load and size for our meta-data information. We use MongoDB as a back-end 
-for the three main components: Mapping, Analytics and DAS cache databases.
-
-DAS Mapping DB holds mapping between data-service APIs and DAS keys. 
-Upon identification of APIs which will participate in DAS activity 
-we store to Mapping DB the API name, input parameters, their type and accept 
-values as regular expression patterns as well as mapping between data-service 
-notations and DAS keys. It allows us to map user input query parameters into
-data-service API metrics as well as map back data-service output into DAS notations.
-For instance, the DBS system \cite{DBS}
-used {\it logical\_file\_name} notation, while in PhEDEx
-\cite{PhEDEx}
-the same entity was named as {\it lfn}. Since both of them were referred to
-the logical file name, we map both notations into DAS key {\it file}.\footnote{We
-also deal with physical file names, but still used {\it file} to refer to them in
-DAS notationis, since simple regular expressions were used to identify the
-meaning of the {\it file} keyword from the provided value.}
-For each service we provide a mapping from its own data-format (JSON, XML, etc.) into
-common data-format used in DAS.\footnote{We used JSON as a DAS data format.}
-We also store a mapping between DAS keys and data-service API names to
-establish appropriate calls to data-services upon provided DAS query. For example,
-when user specifies {\it file} in their DAS query, we invoke DBS 
-{\it listFiles} and PhEDEx {\it fileReplicas} APIs, respectively.
-
-The DAS Analytics DB plays a special role. It collects information
-about user requests placed to the system. Each request was recorded. Upon its
-decomposition into set of selection keys and conditions we also record
-API name and API input parameters into Analytics database and 
-keep updating their counter information for repeated calls. 
-This allows to keep track of frequency of API calls and establish 
-pre-fetch strategies for most common queries.
-
-The DAS caching system was used to store data-service results and aggregated
-information in form of DAS records. As it shown in figure \ref{DAS_cache} 
-the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
-and aggregate data over there. The independent set of daemons, DAS robots, are
-constantly monitor Analytics DB for most popular requests and pre-fetch most
-common data into DAS cache.
-
-Each user request placed to the system was parsed and set of selection
-and condition keys were identified. They were mapped into set of
-data service APIs and their input parameters. DAS
-cache was checked for a similar set of selection and condition
-keys which can be used earlier.\footnote{Since DAS doesn't hold
-any data by default we cannot rely on data being present in a cache.
-Moreover if we look-up data first, we cannot tell if it was complete
-set or not, since earlier queries may retrieve only portion of the requested data.}
-If the superset was found a data were
-retrieved from the cache, otherwise a data-service APIs were
-invoked and the results were collected and aggregated in DAS cache.
-
-%Now we present overall logic of DAS workflow and details of individual components.
-%Upon request to DAS the following steps were executed: 
-%\begin{itemize}
-%\item parse input DAS query
-%\item map provided selection and condition keys into data-service APIs and
-%input parameters
-%\item look-up similar queries in DAS cache 
-%\item invoke data-service APIs
-%\item parse output results by mapping output notation into DAS keys
-%\item store and aggregate them into DAS raw-cache
-%\end{itemize}
-
-Apart from previous DBS-QL syntax we decided to fully support free text-based queries for end-users.
-All of them were transformed into MongoDB query syntax, who was used internally in
-DAS core layer. It represents a set of selection keys and conditions as a JSON document.
-For example, a query
-{\it dataset site=T1\_CH\_CERN}
-where transformed into 
-{\it \{"fields" : ["dataset"], "spec" : \{"site" : "T1\_CH\_CERN"\}\}}.
-Such syntax was perfect fit for us, since it allows to store all queries
-as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
-MongoDB query syntax is very reach. It supports advanced queries with 
-boolean expressions, conditional operators, regular expressions, etc. providing
-a real power for our end-users.
-
-\subsection{Data-services and aggregation}
-DAS provided an interface to add new CMS data-services via data-service plugins.
-Each plugin consists of its DAS map and data parser suitable for data-service output
-format(s).
+data on demand. This was done by invoking appropriate data-service plugin
+upon the user request. Each plugin consists of DAS map,
+data parser(s) suitable for data-service output format(s) into DAS common
+data format (JSON) and was responsible for data-service API calls.
 The data-service DAS map contains information about data-service APIs,
-input and output parameters, data-service
-format and their mapping to DAS keys as well as relation to other sub-systems. 
+input and output parameters and their mapping to DAS keys. 
+This information was stored into DAS Mapping DB which we discuss next.
 Below you can see a typical DAS Mapping DB record for CMS
 SiteDB data-service \cite{SiteDB}\footnote{We used regular expression 
-patterns, evaluated by python during run time
-execution, to validate provided set of input paremters and DAS keys.}:
+patterns, evaluated during run time, to validate provided set of input 
+paremters and DAS keys.}:
 \begin{verbatim}
@@
 \end{verbatim}
-This information was stored into DAS Mapping DB.
+Although the majority of data-services looks alike
+and their logic was abstracted, including parsers, there were a few
+cases where special care was required. For instance,
+we dealt with data-services which were only accessible via private network.
+Therefore, a way to propagate user credentials was required, leading
+to specific plugin implementation.
 
-The DAS Mapping DB becomes an authorative
-source of information about all data-services participating in DAS and their
-relationships. Based information stored there we were able to retrieve appropriate data
-from different data-services, re-map them into DAS notations and
-aggregate them on demand.
-For instance, the Data Bookkeeping System (DBS) and data location system (PhEDEx)
-provided information about CMS data files. In former case, DBS holds information about
-files and their relationship to other physics objects, such as run and luminosity, 
-while PhEDEx stores information about the file location, such as sites which
-holds copy of the files and details about site storage elements.
-Upon a user request to find a file, we were able to identify appropriate set of
-DBS and PhEDEx APIs, look-up data, re-format their output to DAS notations and
-store results into DAS cache.
-Once data were retreived from data-services an additional step to aggregate
-matched objects in DAS cache were applied. For instance, both data-service outputs 
-contained a file name in them. We use it as a common key to merge records in DAS cache.
-This merged (aggregated) results represent DAS records. Each DAS
+%All user queries were written to Analytics DB for further analysis, while
+%mapping between data services were kept in Mapping DB. The data-service 
+%output were re-mapped into DAS notations and, if necessary, aggregated 
+%inside of the DAS cache.
+
+\subsection{DAS Mapping DB}
+DAS Mapping DB holded the mapping between data-service APIs and DAS keys exposed to
+the users. It created a necessary glue between data-services and identifies
+how their data can be related to each other. Basically Mapping DB was positioned 
+as a {\it registration} service with ability to translate notations among data-services.
+Upon identification of APIs which will participate in DAS activity 
+we stored their names, input parameters, together with their type and accept 
+values as regular expression patterns into the Mapping DB. 
+Also we stored the mapping between data-service notations and DAS keys. 
+For instance, the CMS DBS system \cite{DBS}
+used {\it logical\_file\_name} notation to identify the name of the
+logical file, while the same entity was called {\it lfn} in CMS PhEDEx system \cite{PhEDEx}.
+Since both of them were referred to the logical file name, 
+we map both notations into DAS key {\it file}.\footnote{We
+also deal with physical file names, but still used {\it file} to refer to them in
+DAS notations, since simple regular expressions were used to identify the
+meaning of the {\it file} keyword from the provided value.}
+Such keys, e.g. {\it file}, were exposed to the end-users as pre-defined keywords to
+be used in search queries. For instance, end-users were able to type {\it file} in
+order to get all file names known to DAS via data-services as well as
+provide a concrete patterns associated with a file, e.g. {\it file=abc*}.
+Under the hood, the appropriate data-service APIs were looked up along with appropriate
+parameters to answer those questions. For example,
+in aforementioned example, using {\it file} in DAS query cause call of DBS 
+{\it listFiles} and PhEDEx {\it fileReplicas} APIs, respectively.
+This approach gave power to end-users to express the meaning of their queries, especially
+when they deal with numbers and achieve accuracy of results from DAS. 
+For instance, users were able to type $run>10\,\,and\,\, run < 100$
+to specify a range of run numbers to be looked up in DAS cache. At the same time
+a full free text based keyword search is under development within MongoDB.
+
+\subsection{DAS Analytics DB}
+The DAS Analytics DB played a special role. It collected information
+about user requests placed to the system. Each request was recorded. Upon its
+decomposition into set of selection keys and conditions we also recorded
+API name and input parameters into the Analytics database and 
+keep updating their counter information for repeated calls. 
+This allows to keep track of frequency of API calls and establish 
+pre-fetch strategies for most common queries.
+
+\subsection{DAS caching system}
+The DAS caching system was used to store data-service results and aggregated
+information in form of DAS records. As it shown in figure \ref{DAS_cache} 
+the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
+and aggregate data over there. The independent set of daemons, DAS robots, were
+constantly monitor Analytics DB for most popular requests and pre-fetch most
+common data into DAS cache.
+
+Each user request placed to the system was parsed and set of selection
+and condition keys were identified. They were mapped into set of
+data service APIs and their input parameters. DAS
+cache was checked for a queries with a similar set of selection and condition
+keys used earlier.\footnote{Since DAS didn't hold
+any data by default we cannot rely on data being present in a cache.
+Moreover if we look-up data first, we cannot tell if it was complete
+set or not, since earlier queries may retrieve only portion of the requested data.}
+If the superset was found a data were
+retrieved from the cache, otherwise a data-service APIs were
+invoked and the results were collected and aggregated inside of DAS cache.
+
+\subsection{DAS queries}
+Apart from previous DBS-QL syntax we decided to drop its syntax and allow users
+express their queries in a free text-based form. Although we must clarify its
+meaning. We used a pre-defined set of keys known to DAS to describe common
+entities of our data. For example, {\it dataset, file, block, run, event}. Those
+names were used by our users in their daily work operations. So, by typing
+{\it file} we didn't perform a string matching to ``file'' string, but rather
+look-up all file objects in a DAS. Moreover each DAS key has a default attribute.
+For intance, for {\it dataset, block, file} the default attribute 
+was a {\it name}, while for {\it run} and {\it event} the default attribute was 
+a {\it number}. This allows to specify key-value pair as simple as possible, e.g.
+{\it run=123456}, which means run number 123456. At the same time a queyr 
+key, e.g. {\it site} meant to look-up all site objects. Such ability
+to use key-value pairs resolve ambiguity in object values. For instance,
+specifying a run number as {\it run=123456} instruct system to resolve
+data object with run equal to 123456 rather then other objects which may have
+123456 values, e.g. event number or file size. At the same time, to complement
+the stregth of the DAS we do plan to index data objects and add a keyword-based 
+search later.
+
+The free text-based queries were transformed into MongoDB query syntax, 
+who was used internally by DAS. It represents a set of selection 
+keys and conditions as a JSON document. For example, a query
+{\it dataset site=T1\_CH\_CERN}
+where transformed into 
+{\it \{"fields" : ["dataset"], "spec" : \{"site" : "T1\_CH\_CERN"\}\}}.
+Such syntax was perfect fit for us, since it allows to store all queries
+as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
+MongoDB query syntax is very reach. It supports advanced queries with 
+boolean expressions, conditional operators, regular expressions, etc. providing
+a real power for our end-users.
+
+\subsection{Data-services and aggregation}
+As we outlined above the DAS Mapping DB becomes an authorative
+source of information about all data-services participating in DAS.
+Based on information stored in Mapping DB we were able to 
+retrieve appropriate data from different data-services, re-map them into 
+DAS notations and aggregate them on demand. For instance, the Data 
+Bookkeeping System (DBS) and data location system (PhEDEx)
+provided information about CMS data files. In former case, DBS 
+stored information about files and their relationship to other 
+physics objects, such as run and luminosity, while PhEDEx kept
+information about the file location, e.g. details about sites 
+holding copy of the files. Upon a user request to find a file, 
+we were able to identify appropriate set of DBS and PhEDEx APIs, 
+retrieve data from them, re-format their output to DAS notations 
+and store results into DAS cache performing aggregation of matched 
+objects. For instance, since both data-services returns in their
+output a file name we used as a common key during aggregation.
+This merged results represented a set of DAS records. Each DAS
 record contains a standard DAS header, which identify look-up time,
-request url, input parameters and API method as well as response version, expiration
-time and checksum of the output. Here is an example of individual DAS record:
+request url, input parameters and API method as well as response 
+version, expiration time and checksum of the output. Usage of this 
+information was valuable contribution in debugging process of
+data-services themselves, e.g. identification of not-matched records, 
+latency studies, etc. Here is an example of DAS record representing
+aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
+and {\it CMStoSiteName}, with API results merged under the {\it site} key.
 \begin{verbatim}
@@
 \end{verbatim}
-This DAS record represents aggregated information between two SiteDB APIs, 
-{\it CMStoSAMName} and {\it CMStoSiteName}, where API results were merged 
-under the {\it site} key. The information provided by {\it das} key shows 
-DAS header, which contains information about API name, location, execution 
-time and lifetime of the data.
-Usage of this information was valuable contribution in debugging process of
-data-services themselves, e.g. identification of not-matched records, 
-latency studies, etc.
 
-We used RESTful model \cite{REST} for all DAS cache server APIs.
-It allows us to share core library and other components during development of 
-various DAS clients, e.g. CLI, web interface, robots, etc. and hides
-complexity of access to individual data-services and their security policies
-as well as proper organize caching solution.
-For instance, to access on-line Run Summary information we were forced to use
-Grid certificates. Instead of propagating user certificate to data-service
-the DAS cache server works as a proxy to this server and used host certficiate 
-to serve such requests on behalf of the users.
+%We used RESTful model \cite{REST} for all DAS cache server APIs.
+%It allows us to share core library and other components during development of 
+%various DAS clients, e.g. CLI, web interface, robots, etc. and hides
+%complexity of access to individual data-services and their security policies
+%as well as proper organize caching solution.
+%For instance, to access on-line Run Summary information we were forced to use
+%Grid certificates. Instead of propagating user certificate to data-service
+%the DAS cache server works as a proxy to this server and used host certficiate 
+%to serve such requests on behalf of the users.
 
 \section{Results\label{Results}}
-DAS was deployed to work with the following CMS data-services:
+In this section we discuss preliminary restults achieved with DAS system
+using the following CMS data-services:
 \begin{itemize}
-\item DBS \cite{DBS}, Data Bookkeeping System, which collect information
-about CMS meta-data;
-\item PhEDEx \cite{PhEDEx}, Physics Experiment Data Export project, which
-provides the data placement and the file transfer system for the CMS experiment;
-\item SiteDB \cite{SiteDB} which 
-records the CMS resources at the site, the resources pledged for the 
-future, and it keeps track of CMS personnel at each site, including the 
-roles and duties they fulfill in the collaboration;
-\item RunSummary DB \cite{RunSummary} which collects information about run and triggers
-conditions during data-taking;
+\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
+\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
+CMS file transfer system and data-service;
+\item SiteDB \cite{SiteDB}, the CMS site resources data-service;
+\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
 %\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
@@
 
-We did preliminary studies of DAS performance based on information stored
-in 2 systems, DBS and PhEDEx and benchmark time required to collect and 
-aggregate block objects information from both system, see Table \ref{DAS_benchmark}.
+To test DAS performance we measured elapsed time and CPU/RAM utilization
+required to collect and aggregate block object information from DBS and 
+PhEDEx CMS data-services. Preliminary results of those measurements can
+be found in Table \ref{DAS_benchmark}.
 
@@
 \end{tabular}
-\caption{DAS benchmark results. 
-We are measured a time to fetch, parse and aggregate block information
-from DBS and PhEDEx systems via their APIs. In both cases a full block 
-table content, 347K and 230K records, respectively, were used to populate
-DAS cache. The total number of DAS records were calculated
-as number of aggregated records plus left over, non-matched records,
+\caption{Time required to fetch, parse and aggregate block information
+from DBS and PhEDEx systems via their APIs. In the case of PhEDEx
+system we show total number of fetched records 175884 together with
+number of records merged in DAS 170131. The total number of DAS records 
+were calculated as number of aggregated records plus left over, non-matched records,
 from both systems.}
@@
 
-We measured elapsed time of this operation as the following:
+Please note that initial DAS implementation was done in Python \cite{Python}
+and all tests were performed on x64 Linux platform with 16GB of RAM where
+1 CPU was used by DAS server and 1 CPU by MongoDB database server. The CPU utilization
+of DAS server was 67\% in average and RAM usage was below 35MB for
+all tests. All tests were performed multiple times to avoid fluctuations. 
+The elapsed time measured during those tests consists of the 
+following componenets:
 \begin{itemize}
@@
 components, {\it aggregation} and {\it output creation time}, were only applied to
-final DAS step and not to time spent in individual data-services.
+the one of the sub-system (either DBS or PhEDEx) since the first one was injected into
+fresh database, while the second required the merging step.
 
-The individual DAS components (DBS and PhEDEx) shown roughly the same performance.
-For example, in case of PhEDEx the total time spent in DAS was 177 seconds, where
-majority of the time was spent on retrieval, 83 seconds, and parsing, 29 seconds.
-The remaining 65 seconds was distributed among remapping, cache insertion, indexing. The
-numbers for DBS system were alike. As you can see from Table \ref{DAS_benchmark},
-the time spent to read DAS records (last column) from the cache was quite
-reasonable, roughly 2.5 seconds from both systems. While final time to
-get DAS records on disk was below 42 seconds. This was mostly due to I/O operations.
-All operations were performed on a single CPU core and obviously can be
-paralelized in a future.
+%DAS execution time (phedex) 160.606897116 sec
+%DAS execution time (dbs) 52.5349619389 sec
+%DAS execution time 256.0670228 sec, Sat, 28 Nov 2009 15:46:39 GMT
+%
+%DAS execution time (phedex) 0.952513933182 sec
+%DAS execution time (dbs) 0.951184034348 sec
+%DAS execution time 44.7982931137 sec, Sat, 28 Nov 2009 15:52:11 GMT
 
-Based on this results we measured that we can achieve 900 docs/sec rate
-for inserting/aggregating data in DAS cache and 8500 docs/sec for reading 
-DAS records back, which is suitable for our
-needs.\footnote{We've been satisfied with such performance,
-even though our measurements shown that it can still be speed up by factor of
-10 if we will use C++ DB driver instead of python one used in this prototype system.}
+The individual tests of DAS components (DBS and PhEDEx) shown roughly 
+the same performance. Although the time of PhEDEx component was bigger 
+due to longer retrieval and parsing time.
+The time of PhEDEx component shown in Table \ref{DAS_benchmark} divided 
+among: retrieval + parsing time $\sim$110 sec and merging step $\sim$50 sec.
+
+As you can see from Table \ref{DAS_benchmark},
+the time spent to access individual components (last column) was quite
+reasonable, roughly 1 second for both systems. While final time to
+get DAS records on disk was about 45 seconds. This was mostly due to I/O and
+conversion between binary data format used by MongoDB, BSON, to DAS
+data format, JSON.
+
+Based on this results we measured that we can achieve 7000 docs/sec rate
+for insertion, 3400 docs/sec for aggregation and 8500 docs/sec for reading operation
+in DAS, which is suitable for our needs. Since all operations were performed 
+on a single CPU core a further improvements are expected. Also, a necesarry
+tuning can be done by using native C++ MongoDB driver instead of python one.
 
 \section{Summary}
-We presented new data aggregation service (DAS) developed for CMS High-Energy experiment
+We presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
 at LHC, CERN, Geneva, Switzerland. It was designed to provide caching and
 aggregation layer on top of the existing relational and non-relation data-services
-mostly in real time fashion. All the data were retrieved on demand basis,
+mostly in real time fashion. All of the data were retrieved on demand basis,
 while data pre-fetching of most common queries is in our plans. We developed
@@
 \begin{thebibliography}{9}
+\bibitem{DBXplorer}
+S. Agrawal, S. Chaudhuri, G. Das,
+``DBXplorer: A System for Keyword-Based Search over Relational Databases'',
+ICDE 2002, pp. 5-16.
+
+\bibitem{QueryAnswer}
+G. Koutrika, A. Simitsis, Y. E. Ioannidis,
+``Pr\'{e}cis: The Essence of a Query Answer'',
+ICDE 2006, pp. 69-78.
+
+\bibitem{DBS-QL} 
+V. Kuznetsov, D. Riley, A. Afaq, V. Sekhri, Y. Guo, L. Lueking,
+``The CMS DBS Query Language'', CHEP 2009.
+
+\bibitem{AMI}
+Altas AMI web portal,
+http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html
+
 \bibitem{FedDB}
@@
 
-\bibitem{DBXplorer}
-Sanjay Agrawal, Surajit Chaudhuri, Gautam Das: DBXplorer: A System for
-Keyword-Based Search over Relational Databases. ICDE 2002: 5-16
-
-\bibitem{QueryAnswer}
-Georgia Koutrika, Alkis Simitsis, Yannis E. Ioannidis: Pr\'{e}cis: The Essence of
-a Query Answer. ICDE 2006: 69-78
-
-\bibitem{DBS-QL} V. Kuznetsov, D. Riley, A. Afaq, V. Sekhri, Y. Guo, L. Lueking,
-``The CMS DBS Query Language'', CHEP 2009
-
-\bibitem{AMI}
-Altas AMI web portal:
-http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html
-
-\bibitem{CMS} CMS Collaboration R. Adolphi et al., The CMS experiment at the CERN LHC, JINST, 0803, S08004 (2008)
+\bibitem{CMS} 
+R. Adolphi et al., 
+``The CMS experiment at the CERN LHC'',
+JINST 0803, S08004 (2008).
 
 \bibitem{CMSDataModel} 
-.Grandi, D.Stickland, L.Taylor et al., The CMS Computing Model, CERN-LHCC-2004- 035/G-083 (2004);
+C. Grandi, D.Stickland, L.Taylor et al.,
+``The CMS Computing Model'',
+CERN-LHCC-2004-035/G-083 (2004);
 A. Fanfani et. al., 
-``Distributed Analysis in CMS'', to be published in Journal of Grid Computing.
+``Distributed Analysis in CMS'', 
+to be published in Journal of Grid Computing.
 
-\bibitem{DBS} A. Afaq, et. al. ``The CMS Dataset Bookkeeping Service'', CHEP 2007 
-J.Phys.Conf.Ser, 119, 072001 (2008)
+\bibitem{DBS} 
+A. Afaq, et. al.,
+``The CMS Dataset Bookkeeping Service'', 
+J. Phys. Conf. Ser, 119, 072001 (2008).
 
-\bibitem{DBS07} A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
-``A multi-dimensional view on information retrieval of CMS data'', CHEP 2007
+\bibitem{DBS07} 
+A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
+``A multi-dimensional view on information retrieval of CMS data'',
+Journal of Physics: Conference Series 119, 072013, (2008).
 
@@
 
-\bibitem{DBSearch}
-D. Konopnicki, O. Shmueli,
-``Database-Inspired Search'', 
-Proc. of the 31st VLDB Conference, 2005.
+\bibitem{Python} Python programming language, http://www.python.org
+
+%\bibitem{DBSearch}
+%D. Konopnicki, O. Shmueli,
+%``Database-Inspired Search'', 
+%Proc. of the 31st VLDB Conference, 2005.
 \end{thebibliography}

### 2010-05-04 doc/paper.tex
More corrections to chapter 1 and 2.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1737 4525493e-7705-40b1-a816-d608a930855b

@@
 To address those limitations an attempt to build a simple, intuitive and 
-flexible query language was introduced in \cite{DBS-QL}\footnote{A similar but less
-sophisticated approach was used in \cite{AMI}}.
+flexible query language was introduced in \cite{DBS-QL}.
 It represents a power of SQL while hiding underlying relational schema from 
@@
 
-\bibitem{AMI}
-Altas AMI web portal,
-http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html
+%\bibitem{AMI}
+%Altas AMI web portal,
+%http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html
 

### 2010-05-04 test/das_mongocache_t.py
Adjusted compare_specs and unit test for it
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1738 4525493e-7705-40b1-a816-d608a930855b

@@ -14,7 +14,7 @@ from pymongo.connection import Connection
 
 from DAS.utils.das_config import das_readconfig
 from DAS.utils.logger import DASLogger
-from DAS.core.das_mongocache import DASMongocache, transform_keys, loose
+from DAS.core.das_mongocache import DASMongocache, loose
 from DAS.core.das_mongocache import update_item, convert2pattern, compare_specs
 
 class testDASMongocache(unittest.TestCase):
@@ -34,13 +34,6 @@ class testDASMongocache(unittest.TestCase):
         connection = Connection("localhost", 27017)
         connection.drop_database('das') 
         self.dasmongocache = DASMongocache(config)
-
-    def test_transform_keys(self):
-        """Test transform_keys function"""
-        query  = dict(fields=None, spec={'test.name':1, 'site.name':1})
-        expect = dict(fields=None, spec={'test:name':1, 'site:name':1})
-        result = transform_keys(query, '.', ':')
-        self.assertEqual(expect, result)
 
     def test_loose(self):
         """Test loose function"""
@@ -150,6 +143,11 @@ class testDASMongocache(unittest.TestCase):
         exist_query = dict(fields=None, spec={'test':{'$in':[1,2,3]}})
         result = compare_specs(input_query, exist_query)
         self.assertEqual(True, result) # exist_query is a superset 
+
+        input_query = dict(fields=None, spec={'test':'aaa'})
+        exist_query = dict(fields=None, spec={'test':'aaa', 'test2':'bbb'})
+        result = compare_specs(input_query, exist_query)
+        self.assertEqual(False, result) # different set of spec keys
 
     def test_convert2pattern(self):
         """

### 2010-05-04 doc/paper.tex
Adjusted to Eslevier style. Work on text to make it fit in 10 pages
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1745 4525493e-7705-40b1-a816-d608a930855b

@@
 \section{Data Aggregation System\label{DAS}}
+The goal of the Data Aggregation System was to combine existing data-services
+under one umrella, utilizing their access rules, APIs, security policies, etc., and
+allowing users to perform queries against them. In contrast, the approaches
+discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
+enforced a publication step of data into a dedicated database 
+or central repositories \cite{iRODS}. Quite often it is required
+that data should be converted into uniform data formats in order enable
+aggregation, see \cite{OpenArchive}. In DAS architecture the data are provided
+by existing data-services and infrastructure, whose design, implementation and
+security policies are left up to their developers. Therefore DAS does not require
+a data preservation and transaction capabilities. Instead DAS can be treated
+as proxy or caching layer where information retrieval are triggered dynamically
+by user queries. Such infrastructure allows data sharing across different queries and
+data aggregation using appropriate mapping. Below we discuss in details the
+DAS architecture.
+
+\subsection{DAS architecture}
 The design of the DAS system was based on previous studies of CMS Data 
-Bookkeeping System (DBS) \cite{DBS, DBS07}. We carefully analyzed user
-requests and queries placed into the DBS, their patterns, frequencies and latency. 
-As a result we created a data discovery service \cite{DD}. Originally it used MVC
-architecture and later, being re-factored as pure presentation layer on top of 
-the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
-DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
-we took away relations and details on underlying schema from end-users. 
-This was achieved by usaging Dijkstras shortest path algorithm 
-to establish necessary table joins based on provided information in user queries and
-knowledge of foreign key relationships from auto-loaded database schema.
-The DBS-QL keys exposed to end-users were choosen based on a jargon used by
-physicists. They're internally mapped into database entitites and attributes, 
-e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
-A quick adoption and wide usage of DBS-QL gave us confidence in choosen approach 
+Bookkeeping System (DBS) \cite{DBS, DBS07}. Teh DBS Data Discovery service
+\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provides
+ability to place queries, similar to SQL, without specific knowledge of 
+underlying DB schema.
+%We carefully analyzed user
+%requests and queries placed into the DBS, their patterns, frequencies and latency. 
+%As a result we created a data discovery service \cite{DD}. Originally it used MVC
+%architecture and later, being re-factored as pure presentation layer on top of 
+%the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
+%DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
+%we took away relations and details on underlying schema from end-users. 
+%This was achieved by using Dijkstras shortest path algorithm 
+%to establish necessary table joins based on provided information in user queries and
+%knowledge of foreign key relationships from auto-loaded database schema.
+%The DBS-QL keys exposed to end-users were choosen based on a jargon used by
+%physicists. They're internally mapped into database entitites and attributes, 
+%e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
+A quick adoption and wide usage of DBS-QL in CMS gave us confidence in choosen approach 
 and provide a basis for building DAS system.
@@
 \centering
-\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
+\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
 \caption{
@@
 \noindent
-Due to this set of requirements, we rule out the choice of RDMS as DAS back-ends
+Based on this set we rule out the choice of RDMS as DAS back-end(s)
 for several reasons. We agreed to support dynamic queries, data
@@
 Below you can see a typical DAS Mapping DB record for CMS
-SiteDB data-service \cite{SiteDB}\footnote{We used regular expression 
-patterns, evaluated during run time, to validate provided set of input 
-paremters and DAS keys.}:
+SiteDB data-service \cite{SiteDB}:
+%\footnote{We used regular expression 
+%patterns, evaluated during run time, to validate provided set of input 
+%paremters and DAS keys.}:
 \begin{verbatim}
-{"api": {"params": {"name": ""}, "name": "CMSNametoSE"}, 
+{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
  "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
-              "api_param": "name", 
-              "das_key": "site"}], 
- "_id": "4ae1f8dfe2194e4581000014", 
- "system": "sitedb", 
+              "api_param": "name", "das_key": "site"}], 
  "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
-              "key": "site", 
-              "map": "site.name"}]}
+              "key": "site", "map": "site.name"}]}
 \end{verbatim}
@@
 \subsection{DAS caching system}
-The DAS caching system was used to store data-service results and aggregated
-information in form of DAS records. As it shown in figure \ref{DAS_cache} 
-the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
-and aggregate data over there. The independent set of daemons, DAS robots, were
-constantly monitor Analytics DB for most popular requests and pre-fetch most
+The DAS caching system was used to dynamically fetch
+and aggregate data upon user requests. We used the following 
+algorithm:
+\begin{enumerate}[1.]
+\item Decompose query into set of conditions ($conds$) 
+and selection keys ($skeys$)
+\item Check cache for query hash, if found: look-up results from the cache
+\item Identify list of APIs which can serve requested data
+\begin{verbatim}
+apilist = [r(url, api, args) for r in mapping_db(skeys, conds)]
+\end{verbatim}
+Here \verb+mappind_db+ defines translation between input query and data-service
+APIs which can delivery such data.
+\item Iterate over APIs
+\begin{verbatim}
+for url, api, args in apilist:
+    if  superset_in_analytics(api, args):
+        continue
+    for row in parser(get_data(url, api, args)):
+        yield row
+\end{verbatim}
+Here \verb+superset_in_analytics+ identifies if provided \verb+api, args+
+are subset of already registered in Analytics DB api call(s). The 
+\verb+get_data+ and \verb+parser+ are define method to fetch and parse
+data, respectively.
+\item Perform aggregation by iterating over records from the previos step
+\begin{verbatim}
+if cache_has(row.match_key):
+   existing_record = get_from_cache(row.primary_key)
+   merge(row, existing_record)
+   insert_into_cache(row)
+   delete_from_cache(existing_record)
+else:
+   insert_into_cache(row)
+\end{verbatim}
+We did use a bulk insert operations to increase throughput. The \verb+primary_key+
+is a key which identify the record defined in Mapping DB, e.g. for 
+\verb+file+ record it is a \verb+file.name+.
+\item Get results from the cache
+\end{enumerate}
+The data was stored in a form of DAS records into the cache, while
+queries were recorded into Analytics DB for further analysis. We
+also had an option to run independent set of daemons, DAS robots, who
+monitor Analytics DB for most popular requests and pre-fetch the most
 common data into DAS cache. 
+%The interesting part was being able
+%to identify the superset of api/args out of found one in an Analytics DB.
 
-Each user request placed to the system was parsed and set of selection
-and condition keys were identified. They were mapped into set of
-data service APIs and their input parameters. DAS
-cache was checked for a queries with a similar set of selection and condition
-keys used earlier.\footnote{Since DAS didn't hold
-any data by default we cannot rely on data being present in a cache.
-Moreover if we look-up data first, we cannot tell if it was complete
-set or not, since earlier queries may retrieve only portion of the requested data.}
-If the superset was found a data were
-retrieved from the cache, otherwise a data-service APIs were
-invoked and the results were collected and aggregated inside of DAS cache.
+%The DAS caching system was used to store data-service results and aggregated
+%information in form of DAS records. As it shown in figure \ref{DAS_cache} 
+%the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
+%and aggregate data over there. The independent set of daemons, DAS robots, were
+%constantly monitor Analytics DB for most popular requests and pre-fetch most
+%common data into DAS cache.
+
+%Each user request placed to the system was parsed and set of selection
+%and condition keys were identified. They were mapped into set of
+%data service APIs and their input parameters. DAS
+%cache was checked for a queries with a similar set of selection and condition
+%keys used earlier.\footnote{Since DAS didn't hold
+%any data by default we cannot rely on data being present in a cache.
+%Moreover if we look-up data first, we cannot tell if it was complete
+%set or not, since earlier queries may retrieve only portion of the requested data.}
+%If the superset was found a data were
+%retrieved from the cache, otherwise a data-service APIs were
+%invoked and the results were collected and aggregated inside of DAS cache.
 
@@
 \begin{verbatim}
-{"_id": "4ae1e7ffe2194e4243000003", 
- "site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
+{"site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
          {"sitename": "CERN", "name": "T1_CH_CERN"}], 
- "das": [{"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
-          "ctime": 0.95439004898071289, "expire": 1256362175.7692671, 
-          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSAMName", 
-          "timestamp": 1256318975.7692671, "system": ["sitedb"], 
-          "api": ["CMStoSAMName"], "version": "", "selection_keys": []}, 
-         {"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
-          "ctime": 1.2158100605010986, 
-          "api": ["CMStoSiteName"], 
-          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSiteName", 
-          "timestamp": 1256318974.7774379, "system": ["sitedb"], 
-          "expire": 1256362174.7774379, "version": "", "selection_keys": []}
-        ]
-}
+ "das": [{"ctime": 0.9, "expire": 1256362175, "api": ["CMStoSAMName"],
+          "url": "https://a.b.com/sitedb/json/index/CMStoSAMName", 
+          "timestamp": 1256318975, "system": ["sitedb"]},
+         {"ctime": 1.2, "api": ["CMStoSiteName"], "expire": 1256362174},
+          "url": "https://a.b.com/sitedb/json/index/CMStoSiteName", 
+          "timestamp": 1256318974, "system": ["sitedb"]]}
 \end{verbatim}
@@
 operates on private network accessible at CMS online operation center.
-Figure \ref{db_size} shows the current size of meta-data from aforementioned 
-data-services.
-\begin{figure}[htb]
-\centering
-\includegraphics[width=100mm]{db_size.pdf}
-\caption{
-Current size of meta-data for CMS data-services participated in DAS, before
-official data-taking. 
-%We anticipate to collect $\sim$500GB of meta-data
-%per year during normal LHC operations.
-}
-\label{db_size}
-\end{figure}
+%Table \ref{db_size} shows the current size of meta-data from aforementioned 
+%data-services.
+%\begin{figure}[htb]
+%\centering
+%\includegraphics[width=100mm]{db_size.pdf}
+%\caption{
+%Current size of meta-data for CMS data-services participated in DAS, before
+%official data-taking. 
+%}
+%\label{db_size}
+%\end{figure}
 
+%\begin{table*}[hbt]
+%\centering
+%\begin{tabular}{llllll}\hline
+%\hline
+
+%Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
+%\hline
+%tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
+%indexies (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
+%\hline
+%\hline
+%\end{tabular}
+%\caption{
+%Current size of meta-data for CMS data-services participated in DAS, before
+%official data-taking. 
+%}
+%\label{db_size}
+%\end{table*}
 At the moment we don't have exact numbers for amount of data which will stored 
@@
 accumulate of the order of 500 GB of meta-data per year during normal LHC and CMS
-operations.\footnote{This number is based on tests 
-performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
+operations.
+%\footnote{This number is based on tests 
+%performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
 %In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
@@
 
-Please note that initial DAS implementation was done in Python \cite{Python}
+Please note that initial DAS implementation was done in Python 
 and all tests were performed on x64 Linux platform with 16GB of RAM where
@@
 among: retrieval + parsing time $\sim$110 sec and merging step $\sim$50 sec.
-
 As you can see from Table \ref{DAS_benchmark},
@@
 ``The CMS Computing Model'',
-CERN-LHCC-2004-035/G-083 (2004);
-A. Fanfani et. al., 
-``Distributed Analysis in CMS'', 
-to be published in Journal of Grid Computing.
+CERN-LHCC-2004-035/G-083 (2004).
+%A. Fanfani et. al., 
+%``Distributed Analysis in CMS'', 
+%to be published in Journal of Grid Computing.
+
+\bibitem{iRODS}
+The Integrated Rule-Oriented Data System,
+http://www.irods.org/
+
+\bibitem{OpenArchive}
+Open Archives Initiative,
+http://www.openarchives.org/
 

### 2010-05-04 doc/db_size.py

@@ -21,6 +21,7 @@ dbs_info = [
 ]
 dbs_tables = sum(i[1] for i in dbs_info)
 dbs_index = sum(i[2] for i in dbs_info)
+print "Total DBS size:", dbs_tables, dbs_index
 db_info = [
 ('dashboard', 40.82977294921875, 135.9593505859375),
 ('phedex', 5.51031494, 12.5614014),

### 2010-05-04 doc/paper.tex
Correction to final part of the paper
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1749 4525493e-7705-40b1-a816-d608a930855b

@@
 \section{Data Aggregation System\label{DAS}}
-The goal of the Data Aggregation System was to combine existing data-services
+The goal of the Data Aggregation System was to combine and
+aggregate meta-data from existing data-services
 under one umrella, utilizing their access rules, APIs, security policies, etc., and
@@
 \end{verbatim}
-Here \verb+superset_in_analytics+ identifies if provided \verb+api, args+
+Here \verb+superset_in_analytics+ verifies if provided \verb+api, args+
 are subset of already registered in Analytics DB api call(s). The 
-\verb+get_data+ and \verb+parser+ are define method to fetch and parse
+\verb+get_data+ and \verb+parser+ are define methods to fetch and parse
 data, respectively.
@@
 \begin{verbatim}
-if cache_has(row.match_key):
+if cache_has(row.primary_key):
    existing_record = get_from_cache(row.primary_key)
@@
 \end{verbatim}
-We did use a bulk insert operations to increase throughput. The \verb+primary_key+
-is a key which identify the record defined in Mapping DB, e.g. for 
-\verb+file+ record it is a \verb+file.name+.
+The \verb+primary_key+
+is a key which identify the record defined in Mapping DB, e.g. \verb+file.name+
+for \verb+file+ record.
+We did use a bulk insert operations at this step to increase throughput. 
 \item Get results from the cache
 \end{enumerate}
-The data was stored in a form of DAS records into the cache, while
-queries were recorded into Analytics DB for further analysis. We
-also had an option to run independent set of daemons, DAS robots, who
+%The data was stored in a form of DAS records into the cache, while
+%queries were recorded into Analytics DB for further analysis.
+We also had an option to run independent set of daemons, DAS robots, who
 monitor Analytics DB for most popular requests and pre-fetch the most
@@
 \subsection{DAS queries}
-Apart from previous DBS-QL syntax we decided to drop its syntax and allow users
+Apart from the previous DBS-QL syntax we decided to drop its syntax and allow users
 express their queries in a free text-based form. Although we must clarify its
-meaning. We used a pre-defined set of keys known to DAS to describe common
-entities of our data. For example, {\it dataset, file, block, run, event}. Those
-names were used by our users in their daily work operations. So, by typing
-{\it file} we didn't perform a string matching to ``file'' string, but rather
-look-up all file objects in a DAS. Moreover each DAS key has a default attribute.
-For intance, for {\it dataset, block, file} the default attribute 
-was a {\it name}, while for {\it run} and {\it event} the default attribute was 
-a {\it number}. This allows to specify key-value pair as simple as possible, e.g.
-{\it run=123456}, which means run number 123456. At the same time a query 
-key, e.g. {\it site} meant to look-up all site objects. Such ability
-to use key-value pairs resolve ambiguity in object values. For instance,
-specifying a run number as {\it run=123456} instruct system to resolve
-data object with run equal to 123456 rather then other objects which may have
-123456 values, e.g. event number or file size. At the same time, to complement
-the stregth of the DAS we do plan to index data objects and add a keyword-based 
+meaning. We used a pre-defined set of keys registered in DAS to describe common
+entities of our data. For example, \verb+dataset+, 
+\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
+was based on analysis of daily work operations of our users, see \cite{DBS07}.
+Each key had a set of attributes with primary one assigned to the key itself.
+For intance, the \verb+block+ key had the following attributes
+\verb+name+, \verb+size+, \verb+replica+ etc., 
+with primary attribute being \verb+name+.
+The choice of attributes were based on a set of meta-data information
+associated with a given DAS key. Therefore end-users where able to specify
+a context in their queries by using appropriate attribute(s). For example,
+for query \verb+block+ we didn't perform a string matching to the
+``block'' string, but rather look-up a block objects in a DAS, while query
+\verb+block.name=abc*+ instructs system to look-up appropriate blocks
+with matched names. At the same time, to complement the stregth of the 
+DAS, we do plan to index dynamically retrieved meta-data and add a keyword-based 
 search later.
@@
 Bookkeeping System (DBS) and data location system (PhEDEx)
-provided information about CMS data files. In former case, DBS 
-stored information about files and their relationship to other 
-physics objects, such as run and luminosity, while PhEDEx kept
-information about the file location, e.g. details about sites 
-holding copy of the files. Upon a user request to find a file, 
-we were able to identify appropriate set of DBS and PhEDEx APIs, 
+provided information about CMS block meta-data. In former case, DBS 
+stored information about blocks and their relationships to other 
+physics objects, such as dataset and files, while PhEDEx kept
+information about its location and replicas. Upon a user request to 
+find a block meta-data, we were able to identify appropriate set of DBS and PhEDEx APIs, 
 retrieve data from them, re-format their output to DAS notations 
 and store results into DAS cache performing aggregation of matched 
-objects. For instance, since both data-services returns in their
-output a file name we used as a common key during aggregation.
-This merged results represented a set of DAS records. Each DAS
-record contains a standard DAS header, which identify look-up time,
-request url, input parameters and API method as well as response 
-version, expiration time and checksum of the output. Usage of this 
-information was valuable contribution in debugging process of
+objects using \verb+block.name+ key. The merged results were stored 
+as a set of data records. Here is an example of such aggregated record:
+\begin{verbatim}
+{"das_id": ["4b195404e2194e21c4000030", "4b1953fee2194e21c4000002"], 
+ "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
+            "subscribed": "n", "custodial": "n", "site": "T1_US_Buffer", 
+            "nfiles": "18", "se": "srm.site.com", 
+            "size": "112694171838"}, {....}]},
+           {"name": "a/b/c#123", "nevents": "570200", 
+            "created_by": "USER_DN", "path": "/a/b/c", ...}],
+ "das": [{"expire": 1259952908}, {"expire": 1259954702}]}
+\end{verbatim}
+Here \verb+das_id+ refers to api call records stored in Analytics DB,
+\verb+block+ represents an aggregation block meta-data information from DBS
+and PhEDEx systems and \verb+das+ contains their expiration time stamps.
+Usage of this information was valuable contribution in debugging process of
 data-services themselves, e.g. identification of not-matched records, 
-latency studies, etc. Here is an example of DAS record representing
-aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
-and {\it CMStoSiteName}, with API results merged under the {\it site} key.
-\begin{verbatim}
-{"site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
-         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
- "das": [{"ctime": 0.9, "expire": 1256362175, "api": ["CMStoSAMName"],
-          "url": "https://a.b.com/sitedb/json/index/CMStoSAMName", 
-          "timestamp": 1256318975, "system": ["sitedb"]},
-         {"ctime": 1.2, "api": ["CMStoSiteName"], "expire": 1256362174},
-          "url": "https://a.b.com/sitedb/json/index/CMStoSiteName", 
-          "timestamp": 1256318974, "system": ["sitedb"]]}
-\end{verbatim}
+latency studies, etc.
+% Here is an example of DAS record representing
+%aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
+%and {\it CMStoSiteName}, with API results merged under the {\it site} key.
+%\begin{verbatim}
+%{"site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
+%         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
+% "das": [{"ctime": 0.9, "expire": 1256362175, "api": ["CMStoSAMName"],
+%          "url": "https://a.b.com/sitedb/json/index/CMStoSAMName", 
+%          "timestamp": 1256318975, "system": ["sitedb"]},
+%         {"ctime": 1.2, "api": ["CMStoSiteName"], "expire": 1256362174},
+%          "url": "https://a.b.com/sitedb/json/index/CMStoSiteName", 
+%          "timestamp": 1256318974, "system": ["sitedb"]]}
+%\end{verbatim}
 
@@
 %\end{table*}
-At the moment we don't have exact numbers for amount of data which will stored 
-in those data-service and therefor accessible in DAS. Although we project to
-accumulate of the order of 500 GB of meta-data per year during normal LHC and CMS
-operations.
+At the moment, the amount of meta-data stored in all services is equal to
+$\sim$75 GB and $\sim$200 GB in tables and index sizes, respectively. While 
+we estimate to accumulate $\sim$500 GB of the meta-data each year during 
+normal CMS operation at LHC.
 %\footnote{This number is based on tests 
@@
 To test DAS performance we measured elapsed time and CPU/RAM utilization
-required to collect and aggregate block object information from DBS and 
-PhEDEx CMS data-services. Preliminary results of those measurements can
-be found in Table \ref{DAS_benchmark}.
+required to collect and aggregate information from different
+data-services. Table \ref{DAS_benchmark} summarizes the biggest 
+contributors so far, the block meta-data information provided by DBS and 
+PhEDEx CMS data-services.
 
@@
 
-Please note that initial DAS implementation was done in Python 
-and all tests were performed on x64 Linux platform with 16GB of RAM where
-1 CPU was used by DAS server and 1 CPU by MongoDB database server. The CPU utilization
-of DAS server was 67\% in average and RAM usage was below 35MB for
-all tests. All tests were performed multiple times to avoid fluctuations. 
+The initial DAS implementation was done in Python language \cite{Python}
+and all tests were performed on x64 Linux platform where
+1 CPU was used by DAS server and 1 CPU by MongoDB database server. 
+%The CPU utilization
+%of DAS server was 67\% in average and RAM usage was below 35MB. 
+All tests were performed multiple times to avoid fluctuations. 
 The elapsed time measured during those tests consists of the 
@@
 {\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
-is a time required to convert notations used by data-service to DAS ones for every object
-we parse, {\it cache insertion} and {\it indexing time} represents time spend to add objects into
-the DAS cache, {\it aggregation time} is required to merge objects into DAS records based
-on their common key (block name in this case) and {\it output creation time}
+is a time required to convert data-service meta-data to DAS records for every object
+we parse, {\it cache insertion} and {\it indexing time} represents time spent on DAS 
+cache back-end, {\it aggregation time} is required to merge objects into DAS records based
+on their common key (\verb+block.name+ in this case) and {\it output creation time}
 is a time required to write DAS records to disk. Please note that last two
 components, {\it aggregation} and {\it output creation time}, were only applied to
-the one of the sub-system (either DBS or PhEDEx) since the first one was injected into
-fresh database, while the second required the merging step.
+one of the sub-system, either DBS or PhEDEx, depending on insertion order.
+% since the first one was injected into
+%the fresh database, while the second required the merging step.
 
@@
 due to longer retrieval and parsing time.
-The time of PhEDEx component shown in Table \ref{DAS_benchmark} divided 
-among: retrieval + parsing time $\sim$110 sec and merging step $\sim$50 sec.
-As you can see from Table \ref{DAS_benchmark},
+The time of PhEDEx component shown in Table \ref{DAS_benchmark} is divided 
+among retrieval and parsing, $\sim$110 sec, and merging steps, $\sim$50 sec,
+respectively. As shown in Table \ref{DAS_benchmark},
 the time spent to access individual components (last column) was quite
@@
 get DAS records on disk was about 45 seconds. This was mostly due to I/O and
-conversion between binary data format used by MongoDB, BSON, to DAS
+conversion operations between binary data format used by MongoDB, BSON, to DAS
 data format, JSON.
 
-Based on this results we measured that we can achieve 7000 docs/sec rate
-for insertion, 3400 docs/sec for aggregation and 8500 docs/sec for reading operation
-in DAS, which is suitable for our needs. Since all operations were performed 
-on a single CPU core a further improvements are expected. Also, a necesarry
-tuning can be done by using native C++ MongoDB driver instead of python one.
+Based on results of stress tests we measured that we can achieve throughput of
+7000 docs/sec for insertion, 3400 docs/sec for aggregation and 8500 docs/sec for 
+reading operations in DAS, which are suitable for our needs. Since all 
+operations were performed on a single CPU a further improvements are expected. 
+Also, a necessary tuning can be done by using C++ driver instead of python one.
 
@@
 prototype in python for DAS system which is under commissioning phase right now 
-and performed performance studies shown in this paper. 
+and presented performance studies in this paper. 
 The CMS experiment is started to collect its data in December of 2009 and 
-we expect to accumulate a few PB of real data each year to tape. 
+we expect to accumulate a few PB of real data each year. 
 In addition, the Monte-Carlo samples will be produced at this scale.
 Based on test studies performed in CMS, we expect that total size of
-meta-data produced by experiment each year will be of the order of
-$\sim500$GB. The DAS should sustain such load and will provide generic 
+meta-data produced each year by experiment will be of the order of
+500GB. The DAS should sustain such load and will provide generic 
 ``data-discovery'' service for CMS experiment in years to come.

### 2010-05-04 src/python/DAS/services/abstract_service.py
Check for expired apicall records during pass_apicall API. Since some left over can be found and pass_apicall can find them
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1751 4525493e-7705-40b1-a816-d608a930855b

@@ -4,8 +4,8 @@
 """
 Abstract interface for DAS service
 """
-__revision__ = "$Id: abstract_service.py,v 1.54 2009/12/02 15:44:56 valya Exp $"
-__version__ = "$Revision: 1.54 $"
+__revision__ = "$Id: abstract_service.py,v 1.55 2009/12/07 18:40:48 valya Exp $"
+__version__ = "$Revision: 1.55 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -202,6 +202,8 @@ class DASAbstractService(object):
         """
         Filter provided apicall wrt existing apicall records in Analytics DB.
         """
+        spec = {'apicall.expire':{'$lt' : int(time.time())}}
+        self.analytics.col.remove(spec)
         spec = {'apicall.url':url, 'apicall.api':api}
         msg  = 'DBSAbstractService::pass_apicall, %s, API=%s, args=%s'\
         % (self.name, api, api_params)

### 2010-05-04 src/python/DAS/core/das_mongocache.py
Add ability to view map/reduce function definitions on a web
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1804 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 DAS mongocache wrapper.
 """
 
-__revision__ = "$Id: das_mongocache.py,v 1.52 2009/12/22 19:26:29 valya Exp $"
-__version__ = "$Revision: 1.52 $"
+__revision__ = "$Id: das_mongocache.py,v 1.53 2009/12/23 15:18:45 valya Exp $"
+__version__ = "$Revision: 1.53 $"
 __author__ = "Valentin Kuznetsov"
 
 import re
@@ -474,6 +474,18 @@ class DASMongocache(Cache):
         self.logger.debug(fmap)
         self.logger.debug(freduce)
         return result
+
+    def get_map_reduce(self, name=None):
+        """
+        Return definition of map/reduce functions for provided name
+        or gives full list.
+        """
+        spec = {}
+        if  name:
+            spec = {'name':name}
+        result = self.mrcol.find(spec)
+        for row in result:
+            yield row
 
     def update_cache(self, query, results, header):
         """

### 2010-05-04 doc/paper.tex
Work on Introduction and Related Work sections
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1805 4525493e-7705-40b1-a816-d608a930855b

@@
 
-%\cortext[cor1]{Corresponding autho}
+%\cortext[cor1]{Corresponding author}
 
 \begin{abstract}
-Meta-data plays significant role in large modern enterprises, 
-research experiments and digital libraries where it comes from a different 
-sources and distributed in a variety of forms and digital formats. 
+Metadata plays a significant role in large modern enterprises, 
+research experiments and digital libraries where it comes from many different 
+sources and is distributed in a variety of digital formats. 
 It is organized and managed by constantly evolving software using 
 both relational and non-relational data sources. Even though we can apply
-information retrieval approach to non-relation data sources,
+an information retrieval approach to non-relation data sources,
 we can't do so for relational ones, where information is accessed via
-pre-established set of data-services.
+a pre-established set of data-services.
 
@@
 The data and associated meta-data are produced in variety of forms and digital formats.
-They are stored and retrieved from relational and non-releational data-sources, such as 
+They are stored and retrieved from relational and non-relational data-sources, such as 
 RDMS systems, document oriented databases, blogs, twikies, file systems,
-customized applications, etc. 
-Working in such environment requires a lot of expertise and our users, physicists, 
-are always looking for simple, intuitive and flexible
-tool(s) to look-up their desired data. A well-known solutions, such as data-services
-and search engines are tighted to a specific data sources and end-users are left 
-with a manual task to bookkeep and relatate information from them.
-
-Here we present a work on Data Aggregation System (DAS) designed for
-CMS High-Energy Experiment at LHC which provides
-ability to query, search and aggregate information from different 
-data-services and trigger information retrieval on demand.
+customized applications, etc. Working in such environment requires a lot 
+of expertise and easy way to look-up desired information is always a necessity.
+A well-known solutions, e.g. data-services, are tighten to a specific data 
+sources and end-users are left with a manual task to bookkeep and 
+relate information from them.
+
+Here we present a work on Data Aggregation System (DAS) which provides
+ability to search and aggregate information across different 
+data-services. While it was designed for CMS High-Energy Physics 
+experiment at LHC, it can be applied elsewhere.
 
@@
 been discussed. It requires that all keywords are matched in a row from a
-single table or joined tables. Authors of \cite{QueryAnswer} made a step forward
-by considering information around matched values and presents customized answers to
-end-users over the generated new database. These and other related 
-works, see \cite{DBXplorer, QueryAnswer} for more references,
-are based on some kind of the graph over the underlying database and
-use a generated database out of pristine one to perform the work.
-The proximity of the results in those techniques 
-are mostly confined to a text-based values stored in a database, since 
-usage of numerical values in the input search field is mostly useless without 
-knowledge of the context of the value. For example, typing a value 100 
-in the search input field leads to plenty of inappropriate matches, 
-such as row ids, and in order to be
-answered correctly requires additional keyword, clarifying its meaning, 
-which should be matched with particular
-entity (table.column) in the underlying database.
-Therefore to answer presicely the logical questions, such as  
+single table or joined tables. Authors of \cite{QueryAnswer} are made a step forward
+and evaluated information around matched values. The results were presented as
+customized templates to end-users. 
+These and other related works, see \cite{DBXplorer, QueryAnswer} 
+for more references, are based on some kind of the graph over the underlying 
+database and generated a new database out of pristine one to perform the work.
+The proximity of the results in those techniques are mostly confined to a 
+text-based values stored in a database, since the usage of numerical ones 
+is mostly useless without knowledge of the context of the input value. 
+For example, typing a value 100 in the search input field can lead to 
+plenty of inappropriate matches, e.g. row ids. In order to be
+answered correctly it requires additional keyword, clarifying its meaning, 
+which should be matched with particular entity (table.column) in the 
+underlying database. Therefore the logical questions, e.g.
 {\it find me total number of files whose size more then 20 but less then 
-100} would be hard or almost impossible in \cite{DBXplorer, QueryAnswer}, 
-while it can be easily accomplished using SQL syntax.
-
-To address those limitations an attempt to build a simple, intuitive and 
-flexible query language was introduced in \cite{DBS-QL}.
-It represents a power of SQL while hiding underlying relational schema from 
-the end-users. As a results a human questions were intuitively mapped into 
-simple queries. For instance, the question
+100} are hard to answer precisely using approaches outlined in \cite{DBXplorer, QueryAnswer}, 
+while it can be easily accomplished using SQL.
+
+To address those limitations a keyword-based query language was introduced in \cite{DBS-QL}.
+It used a set of pre-defined keys mapped to underlying schema and shortest path
+algorithm which employed a foreign key constrains to build SQL queries. As a result
+the question
 {\it I'm looking for files who contain data taken on certain date and located at
@@
 aggregate meta-data from existing data-services
-under one umrella, utilizing their access rules, APIs, security policies, etc., and
+under one umbrella, utilizing their access rules, APIs, security policies, etc., and
 allowing users to perform queries against them. In contrast, the approaches
@@
 Each key had a set of attributes with primary one assigned to the key itself.
-For intance, the \verb+block+ key had the following attributes
+For instance, the \verb+block+ key had the following attributes
 \verb+name+, \verb+size+, \verb+replica+ etc., 
@@
 \verb+block.name=abc*+ instructs system to look-up appropriate blocks
-with matched names. At the same time, to complement the stregth of the 
+with matched names. At the same time, to complement the strength of the 
 DAS, we do plan to index dynamically retrieved meta-data and add a keyword-based 
@@
 The elapsed time measured during those tests consists of the 
-following componenets:
+following components:
 \begin{itemize}

### 2010-05-04 doc/paper.tex
Work on Section 3, 4, 4.1
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1806 4525493e-7705-40b1-a816-d608a930855b

@@
 the core library with support of pluggable modules for data retrieval;
-caching layer to store data-service output and aggregate results;
+caching layer to store and aggregate results;
 request service to handle user queries;
-Mapping DB to keep information about data-service APIs, their
-notations and mapping to DAS;
-Analytics DB for query analysis and identification of pre-fetch 
-strategies.
+Mapping DB to keep information about data-service APIs, e.g.
+notations and mapping to DAS keys;
+Analytics DB for query analysis.
 

### 2010-05-04 doc/paper.tex
Work on section 4.2
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1807 4525493e-7705-40b1-a816-d608a930855b

@@
 \subsection{DAS core}
-The DAS core communicated with different data-services and retrieved
-data on demand. This was done by invoking appropriate data-service plug-in(s)
-upon the user request. Each plug-in consists of DAS map, the
-data parser(s) to convert data-service output into DAS records
-and it was responsible for data-service API calls.
-The data-service plug-in contains a map about data-service APIs,
-input and output parameters and their mapping to DAS keys. For example
+The DAS core library was responsible for data retrieval on demand. 
+It had the following components: input query validator and parser,
+query analyzer, workflow builder, execution engine and data-service 
+plug-ins. The communication between client (web server or CLI) and 
+DAS cache server was done via the REST \cite{REST} interface. 
+%For each GET request we performed the following action.
+Each input query was recorded into Analytics DB for further 
+analysis. It was decomposed into the set of conditions and/or selection keys.
+First the results were looked up in a cache and delivered to the requestor
+if superset of input conditions was found in analytics DB.\footnote{It is important
+to note here that we didn't look-up results in a cache and rather
+compared input query condition set to the records in analytics DB 
+in order to understand if other API calls are covered the input request.}
+Otherwise, the appropriate workflow, a set of API calls, was constructed
+for the execution engine. It invoked the API calls via data-service
+plug-ins and store results into the cache. Finally, the results were
+retrieved from the cache and delivered to the user.
+
+The mapping between user queries, the DAS keys used in a query, 
+and API names/parameters was stored in a Mapping DB. Such mapping was 
+used by data-service plug-ins. For example
 \begin{verbatim}
@@
 \end{verbatim}
-This information was stored into DAS Mapping DB which we discuss next.
-
-Although the majority of data-service plug-ins looks alike
-and their logic was abstracted, including parsers, there were a few
-cases where special care was required, e.g. access to data-services 
-accessible via private network. Therefore, a way to propagate 
-user credentials was required, leading to specific plug-in implementation.
-
-%All user queries were written to Analytics DB for further analysis, while
-%mapping between data services were kept in Mapping DB. The data-service 
-%output were re-mapped into DAS notations and, if necessary, aggregated 
-%inside of the DAS cache.
+represents the mapping between \verb+site+ DAS keyword and
+\verb+CMSNametoSE+ API provided by SiteDB data-service (the data-service
+URLs were kept in the configuration file).
+The data-service plug-in implementation were similar and their logic was 
+abstracted, while a few extentions were required, e.g. specific 
+security modules for data-services accessible via private network.
 
@@
 
-%\bibitem{REST}
-%Fielding, Roy Thomas ``Architectural Styles and the Design of 
-%Network-based Software Architectures'', Doctoral dissertation, 2000,
-%University of California, Irvine
+\bibitem{REST}
+Fielding, Roy Thomas ``Architectural Styles and the Design of 
+Network-based Software Architectures'', Doctoral dissertation, 2000,
+University of California, Irvine
 

### 2010-05-04 doc/paper.tex
Work on section 4.3 and 4.4
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1808 4525493e-7705-40b1-a816-d608a930855b

@@
 
-\subsection{DAS core}
+\subsection{DAS core\label{DAS_core}}
 The DAS core library was responsible for data retrieval on demand. 
@@
 for the execution engine. It invoked the API calls via data-service
-plug-ins and store results into the cache. Finally, the results were
-retrieved from the cache and delivered to the user.
-
-The mapping between user queries, the DAS keys used in a query, 
-and API names/parameters was stored in a Mapping DB. Such mapping was 
-used by data-service plug-ins. For example
+plug-ins and stored results into the cache. At this step we also
+performed necessary data aggregation for records with similar key-value
+pairs. The data-service plug-in implementation were similar and their 
+logic was abstracted, while a few extentions were required, e.g. specific 
+security modules for data-services accessible via private network.
+Finally, the results were retrieved from the cache and delivered to the user.
+
+\subsection{DAS Mapping DB}
+The Mapping DB represents DAS {\it registration} service. It collects
+information about data-service APIs, translates API input/output
+parameters into DAS keys and contains mapping between DAS records
+and their UI representation.
+
+%The Mapping DB hold the mapping between data-service APIs and DAS keys exposed to
+%the users. It created a necessary glue between data-services and identifies
+%how their data can be related to each other. Basically Mapping DB was positioned 
+%as a {\it registration} service with ability to translate notations among data-services.
+Upon identification of APIs which participate in DAS 
+we stored their names, input parameters and associated DAS keys
+into the Mapping DB. For example
 \begin{verbatim}
@@
 \end{verbatim}
-represents the mapping between \verb+site+ DAS keyword and
-\verb+CMSNametoSE+ API provided by SiteDB data-service (the data-service
-URLs were kept in the configuration file).
-The data-service plug-in implementation were similar and their logic was 
-abstracted, while a few extentions were required, e.g. specific 
-security modules for data-services accessible via private network.
-
-\subsection{DAS Mapping DB}
-The Mapping DB hold the mapping between data-service APIs and DAS keys exposed to
-the users. It created a necessary glue between data-services and identifies
-how their data can be related to each other. Basically Mapping DB was positioned 
-as a {\it registration} service with ability to translate notations among data-services.
-Upon identification of APIs which will participate in DAS activity 
-we stored their names, input parameters, together with their type and accept 
-values, as regular expression patterns, into the Mapping DB. 
-Also we stored the mapping between data-service notations and DAS keys. 
-For instance, the CMS DBS system \cite{DBS}
-used a \verb+logical_file_name+ notation to identify the name of the
-logical file, while the same entity was called \verb+lfn+ 
-in CMS PhEDEx system \cite{PhEDEx}.
-Since both of them were referred to the logical file name, 
-we map both notations into DAS key \verb+file+.
-%\footnote{We
-%also deal with physical file names, but still used {\it file} to refer to them in
-%DAS notations, since simple regular expressions were used to identify the
-%meaning of the {\it file} keyword from the provided value.}
-Such keys, e.g. \verb+file+, were exposed to the end-users as pre-defined keywords to
-be used in search queries. For instance, end-users were able to type \verb+file+ in
-order to get all file names known to DAS via data-services as well as
-provide a concrete patterns associated with a file, e.g. \verb+file = abc*+.
-Under the hood, the appropriate data-service APIs were looked up along with appropriate
-parameters to answer those questions. In aforementioned example, 
-usage of \verb+file+ in DAS query cause a call of DBS 
-\verb+listFiles+ and PhEDEx \verb+fileReplicas+ APIs, respectively.
-This approach gave ability to end-users to express the meaning of their queries, especially
-when they deal with numbers, and achieve accuracy of results. 
-For instance, users were able to type \verb+run > 10 and run < 100+
-to specify a range of run numbers to be looked up in DAS cache. At the same time
-a full keyword search is under development.
-% within MongoDB.
+represents DAS mapping record.\footnote{
+The data-service URLs were kept in the configuration file.
+}
+This record can be read as following:
+the \verb+CMSNametoSE+ API provided by 
+\verb+sitedb+ data-service accepts \verb+name+ as input parameter. It
+confirms regular expression pattern, \verb+re.compile('^T[0-3]_')+, and
+corresponds to \verb+site+ DAS key. The \verb+site+ key
+maps into \verb+site.name+ within DAS record.
+
+Therefore the Mapping DB identifies a set of pre-defined DAS keys, 
+e.g. \verb+site, file, run+, etc., to be used by end-users in their queries.
+Each key was mapped into set of APIs whose output was aggregated afterwards.
 
 \subsection{DAS Analytics DB}
-The DAS Analytics DB played a special role. It collected information
+The DAS Analytics DB was used to collect information
 about user requests placed to the system. Each request was recorded. Upon its
 decomposition into set of selection keys and conditions we also recorded
-API name and input parameters into the Analytics database and 
-keep updating their counter information for repeated calls. 
-This allows to keep track of frequency of API calls and establish 
+API name and input parameters into the Analytics DB and 
+kept updating their counter information for repeated calls. 
+This allowed to keep track of frequency of API calls and establish 
 pre-fetch strategies for most common queries.
@@
 \verb+block.name=abc*+ instructs system to look-up appropriate blocks
-with matched names. At the same time, to complement the strength of the 
+with matched names. The numerical values, conditions, groupping were
+easy to express too. For example, \verb+run > 10 and run < 100+ query 
+was used to get all run records for provided range of run numbers. 
+At the same time, to complement the strength of the 
 DAS, we do plan to index dynamically retrieved meta-data and add a keyword-based 

### 2010-05-04 doc/paper.tex
Work on section 4.5
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1809 4525493e-7705-40b1-a816-d608a930855b

@@
 \subsection{DAS Analytics DB}
-The DAS Analytics DB was used to collect information
+The DAS Analytics DB collected information
 about user requests placed to the system. Each request was recorded. Upon its
@@
 The \verb+primary_key+
-is a key which identify the record defined in Mapping DB, e.g. \verb+file.name+
+is a key which identifies the record defined in Mapping DB, e.g. \verb+file.name+
 for \verb+file+ record.
@@
 \end{enumerate}
-%The data was stored in a form of DAS records into the cache, while
-%queries were recorded into Analytics DB for further analysis.
 We also had an option to run independent set of daemons, DAS robots, who
@@
 common data into DAS cache.
-%The interesting part was being able
-%to identify the superset of api/args out of found one in an Analytics DB.
-
-%The DAS caching system was used to store data-service results and aggregated
-%information in form of DAS records. As it shown in figure \ref{DAS_cache} 
-%the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
-%and aggregate data over there. The independent set of daemons, DAS robots, were
-%constantly monitor Analytics DB for most popular requests and pre-fetch most
-%common data into DAS cache.
-
-%Each user request placed to the system was parsed and set of selection
-%and condition keys were identified. They were mapped into set of
-%data service APIs and their input parameters. DAS
-%cache was checked for a queries with a similar set of selection and condition
-%keys used earlier.\footnote{Since DAS didn't hold
-%any data by default we cannot rely on data being present in a cache.
-%Moreover if we look-up data first, we cannot tell if it was complete
-%set or not, since earlier queries may retrieve only portion of the requested data.}
-%If the superset was found a data were
-%retrieved from the cache, otherwise a data-service APIs were
-%invoked and the results were collected and aggregated inside of DAS cache.
 

### 2010-05-04 src/python/DAS/web/DASSearch.py
Allow to choose between json formatting of DAS records and the one provided by yaml which close to human readable ascii text. Will experiement with different records to get better idea which is the best choice. JSON is still a default formatting for DAS records
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1843 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 DAS web interface, based on WMCore/WebTools
 """
 
-__revision__ = "$Id: DASSearch.py,v 1.34 2010/01/06 21:18:12 valya Exp $"
-__version__ = "$Revision: 1.34 $"
+__revision__ = "$Id: DASSearch.py,v 1.35 2010/01/07 21:29:02 valya Exp $"
+__version__ = "$Revision: 1.35 $"
 __author__ = "Valentin Kuznetsov"
 
 # system modules
@@ -17,6 +17,8 @@ import types
 import urllib
 import cherrypy
 import traceback
+
+import yaml
 
 from itertools import groupby
 from cherrypy import expose
@@ -145,9 +147,15 @@ class DASSearch(TemplatedPage):
         Return DAS mapping record about provided API.
         """
         record = self.dasmgr.mapping.api_info(name)
-        jsoncode = {'jsoncode': json2html(record, "")}
-        page  = "<b>DAS mapping record</b>"
-        page += self.templatepage('das_json', **jsoncode)
+        show   = kwargs.get('show', 'json')
+        page   = "<b>DAS mapping record</b>"
+        if  show == 'json':
+            jsoncode = {'jsoncode': json2html(record, "")}
+            page += self.templatepage('das_json', **jsoncode)
+        else:
+            code  = yaml.dump(record, width=100, indent=4, 
+                        default_flow_style=False)
+            page += self.templatepage('das_code', code=code)
         return self.page(page, response_div=False)
 
     @expose
@@ -238,69 +246,83 @@ class DASSearch(TemplatedPage):
         """
         Retieve all records id's.
         """
-        recordid = None
-        format = ''
-        if  args:
-            recordid = args[0]
-            spec = {'_id':recordid}
-            fields = None
-            query = dict(fields=fields, spec=spec)
-            if  len(args) == 2:
-                format = args[1]
-        elif  kwargs and kwargs.has_key('_id'):
-            spec = {'_id': kwargs['_id']}
-            fields = None
-            query = dict(fields=fields, spec=spec)
-        else: # return all ids
-            query = dict(fields=None, spec={})
+        try:
+            recordid = None
+            format = ''
+            if  args:
+                recordid = args[0]
+                spec = {'_id':recordid}
+                fields = None
+                query = dict(fields=fields, spec=spec)
+                if  len(args) == 2:
+                    format = args[1]
+            elif  kwargs and kwargs.has_key('_id'):
[…]

### 2010-05-04 doc/paper.tex
Incorporate Dave's changes
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1844 4525493e-7705-40b1-a816-d608a930855b

@@
 \end{verbatim}
-The \verb+primary_key+
+The \verb+daskey+
 is a key which identifies the record defined in Mapping DB, e.g. \verb+file.name+
@@
 monitor Analytics DB for most popular requests and pre-fetch the most
-common data into DAS cache.
+common data into the cache.
 
 \subsection{DAS queries}
-Apart from the previous DBS-QL syntax we decided to drop its syntax and allow users
-express their queries in a free text-based form. Although we must clarify its
-meaning. We used a pre-defined set of keys registered in DAS to describe common
-entities of our data. For example, \verb+dataset+, 
+All DAS queries were expressed in a free text-based form, either as a 
+set of keywords or key-value pairs, where a pair can represent
+condition, e.g. \verb+key > value+. We used a pre-defined set of keys registered 
+in DAS to describe common entities of our data. For example, \verb+dataset+, 
 \verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
-was based on analysis of daily work operations of our users, see \cite{DBS07}.
-Each key had a set of attributes with primary one assigned to the key itself.
-For instance, the \verb+block+ key had the following attributes
-\verb+name+, \verb+size+, \verb+replica+ etc., 
-with primary attribute being \verb+name+.
-The choice of attributes were based on a set of meta-data information
-associated with a given DAS key. Therefore end-users where able to specify
-a context in their queries by using appropriate attribute(s). For example,
-for query \verb+block+ we didn't perform a string matching to the
-``block'' string, but rather look-up a block objects in a DAS, while query
-\verb+block.name=abc*+ instructs system to look-up appropriate blocks
+was based on the analysis of daily work operations of our users, see \cite{DBS07}.
+Each key represents a DAS record. Every record was stored into the cache
+in JSON format. Due to the ``schemaless'' nature of underlying back-end (MongoDB)
+we were able to store DAS records of arbitrary structure, e.g.
+dictionary, lists, key-value pairs, etc. Therefore every DAS key had a 
+set of attributes representing JSON structure. For instance a document
+\begin{verbatim}
+{"block": {"name": "abc", "size": 1, "replica": [...]}}
+\end{verbatim}
+associated with a \verb+block+ key who had the following attributes
+\verb+name+, \verb+size+, \verb+replica+. The set of attributes is based 
+on a set of meta-data information stored in a record and therefore
+a nested structures were allowed, e.g. \verb+block.replica.size+. Our users 
+where able to specify a context of their queries by using appropriate 
+attribute(s). For example, query \verb+block+ represents search for all
+block records rather then ``block'' string, while query
+\verb+block.name=abc*+ instructs the system to look-up appropriate blocks
 with matched names. The numerical values, conditions, groupping were
-easy to express too. For example, \verb+run > 10 and run < 100+ query 
+easy to express too, e.g. \verb+run > 10 and run < 100+ query 
 was used to get all run records for provided range of run numbers. 
-At the same time, to complement the strength of the 
-DAS, we do plan to index dynamically retrieved meta-data and add a keyword-based 
-search later.
+To complement the strength of the queries, we do plan to index 
+dynamically DAS records and add a keyword-based 
+search within record content later.
 
-The free text-based queries were transformed into MongoDB query syntax, 
-who was used internally in DAS. It represents a set of selection 
-keys and conditions as a JSON document. For example, a query
-{\it dataset site=T1\_CH\_CERN}
-was transformed into 
-{\it \{"fields" : ["dataset"], "spec" : \{"site" : "T1\_CH\_CERN"\}\}}.
-Such syntax was perfect fit for us, since it allows to store all queries
-as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
-MongoDB query syntax is very reach. It supports advanced queries with 
-boolean expressions, conditional operators, regular expressions, etc. providing
-a real power for our end-users.
+All input queries were transformed into MongoDB query syntax who is itself
+a JSON record. It contains a set of selection 
+keys (\verb+fields+) and conditions (\verb+spec+). For example, a query
+\verb+dataset, file=abc, run>10+
+was represented as
+\begin{verbatim}
+{"fields": ["dataset"], "spec": {"file": "abc", "run": {"$gt": 10}}}
+\end{verbatim}
+Therefore users queries were stored into Analytics DB similar to other DAS 
+records and used for further analysis.
+%Such syntax was perfect fit for us, since it allows to store all queries
+%as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
+%MongoDB query syntax is very reach. It supports advanced queries with 
+%boolean expressions, conditional operators, regular expressions, etc. providing
+%a real power for our end-users.
 
 \subsection{Data-services and aggregation}
-As we outlined above the DAS Mapping DB becomes an authorative
+As was shown in section \ref{MappingDB} the DAS Mapping DB became an authorative
 source of information about all data-services participating in DAS.
-Based on information stored in Mapping DB we were able to 
+Based on the information stored in Mapping DB we were able to 
 retrieve appropriate data from different data-services, re-map them into 
 DAS notations and aggregate them on demand. For instance, the Data 
-Bookkeeping System (DBS) and data location system (PhEDEx)
-provided information about CMS block meta-data. In former case, DBS 
+Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
+were provided information about CMS blocks. In former case, DBS 
 stored information about blocks and their relationships to other 
@@
 information about its location and replicas. Upon a user request to 
-find a block meta-data, we were able to identify appropriate set of DBS and PhEDEx APIs, 
-retrieve data from them, re-format their output to DAS notations 
-and store results into DAS cache performing aggregation of matched 
-objects using \verb+block.name+ key. The merged results were stored 
-as a set of data records. Here is an example of such aggregated record:
+find a block meta-data, we were able to identify appropriate set of 
+DBS and PhEDEx APIs, retrieve data, re-format their output 
+into DAS notations and store results into DAS cache. The records
+were merged (aggregated) based on identical key-value pairs, e.g.
+based on \verb+block.name+ key. Here is an example of such aggregated record:
 \begin{verbatim}
@@
 \end{verbatim}
-Here \verb+das_id+ refers to api call records stored in Analytics DB,
-\verb+block+ represents an aggregation block meta-data information from DBS
-and PhEDEx systems and \verb+das+ contains their expiration time stamps.
+The \verb+das_id+ part refers to api call records stored in Analytics DB,
+the \verb+block+ part represents an aggregation block meta-data information from DBS
+and PhEDEx systems and the \verb+das+ part contains their expiration time stamps.
 Usage of this information was valuable contribution in debugging process of
-data-services themselves, e.g. identification of not-matched records, 
+data-services themselves, e.g. identification of data-service inconsistencies, 
 latency studies, etc.
-% Here is an example of DAS record representing
-%aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
-%and {\it CMStoSiteName}, with API results merged under the {\it site} key.
-%\begin{verbatim}
-%{"site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
-%         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
-% "das": [{"ctime": 0.9, "expire": 1256362175, "api": ["CMStoSAMName"],
-%          "url": "https://a.b.com/sitedb/json/index/CMStoSAMName", 
-%          "timestamp": 1256318975, "system": ["sitedb"]},
-%         {"ctime": 1.2, "api": ["CMStoSiteName"], "expire": 1256362174},
-%          "url": "https://a.b.com/sitedb/json/index/CMStoSiteName", 
-%          "timestamp": 1256318974, "system": ["sitedb"]]}
-%\end{verbatim}
-
-%We used RESTful model \cite{REST} for all DAS cache server APIs.
-%It allows us to share core library and other components during development of 
-%various DAS clients, e.g. CLI, web interface, robots, etc. and hides
-%complexity of access to individual data-services and their security policies
-%as well as proper organize caching solution.
-%For instance, to access on-line Run Summary information we were forced to use
-%Grid certificates. Instead of propagating user certificate to data-service
-%the DAS cache server works as a proxy to this server and used host certificate 
-%to serve such requests on behalf of the users.
 
@@
 
+\bibitem{Python} Python programming language, http://www.python.org
+
 \bibitem{PhEDEx}

### 2010-05-04 doc/paper.tex
Final round of correction based on Simon/Dave comments.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1857 4525493e-7705-40b1-a816-d608a930855b

@@
 \verb+name+, \verb+size+, \verb+replica+. The set of attributes is based 
-on a set of meta-data information stored in a record and therefore
-a nested structures are allowed, e.g. \verb+block.replica.size+. Our users 
-are able to specify a context of their queries by using appropriate 
-attribute(s). For example, query \verb+block+ represents search for all
-block records rather then ``block'' string, while query
+on a set of meta-data information stored in a record, allowing
+for nested structures, e.g. \verb+block.replica.size+. Our users 
+are able to specify the context of their queries by using appropriate 
+attribute(s). For example, a query for \verb+block+ represents a search for all
+block records rather then the ``block'' string, while the query
 \verb+block.name=abc*+ instructs the system to look-up appropriate blocks
-with matched names. The numerical values, conditions, grouping are
-easy to express too, e.g. \verb+run > 10 and run < 100+ query 
-is used to get all run records for provided range of run numbers. 
-To complement the strength of the queries, we do plan to index 
-dynamically DAS records and add a keyword-based 
-search within record content later.
+with matching names. Numerical values, conditions and grouping are
+also easy to express, for example, a query
+of the form \verb+run > 10 and run < 100+ can be used to
+retrieve all run records for the given range of run numbers. 
+As a complement to the queries, we will als be adding a dynamic index
+of DAS records with a keyword-based search within record content later.
 
-All input queries are transformed into MongoDB query syntax who is itself
+All input queries are transformed into the MongoDB query syntax who is itself
 a JSON record. It contains a set of selection 
 keys (\verb+fields+) and conditions (\verb+spec+). For example, a query
-\verb+dataset, file=abc, run>10+
+\verb+dataset, file=abc, run > 10+
 is represented as
@@
 \end{itemize}
-Each data-service had its own scope, size and peculiar properties. 
-For instance, the data in PhEDEx were transient due to constant 
-migration of CMS data; the DBS system was divided into dozen 
-of individual services (instances) with own size and lifetime 
-used by different production teams for various bookkeeping tasks; 
-the RunSummary data-service operated on private network accessible 
+Each data-service has its own scope, size and peculiarities. 
+For instance, the data in PhEDEx is transient due to constant 
+migration of CMS data; the DBS system is divided into several individual 
+instances with sizes and lifetimes driven by specific processing tasks; 
+the RunSummary data-service is operated on private network accessible 
 at CMS online operation center.
-%Table \ref{db_size} shows the current size of meta-data from aforementioned 
-%data-services.
-%\begin{figure}[htb]
-%\centering
-%\includegraphics[width=100mm]{db_size.pdf}
-%\caption{
-%Current size of meta-data for CMS data-services participated in DAS, before
-%official data-taking. 
-%}
-%\label{db_size}
-%\end{figure}
 
-%\begin{table*}[hbt]
-%\centering
-%\begin{tabular}{llllll}\hline
-%\hline
-
-%Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
-%\hline
-%tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
-%indexes (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
-%\hline
-%\hline
-%\end{tabular}
-%\caption{
-%Current size of meta-data for CMS data-services participated in DAS, before
-%official data-taking. 
-%}
-%\label{db_size}
-%\end{table*}
 At the moment, the amount of meta-data stored in all services is around
-75 GB and 200 GB in tables and index sizes, respectively. Although we estimate 
-$\sim$500 GB of the meta-data each year during 
-normal CMS operation at LHC.
+75 GB and 200 GB in tables and index sizes, respectively. We estimate to
+collect around 500 GB of meta-data each year during 
+normal CMS operations. The average query rate to DAS is expected to be at the 
+$\leq$10K requests a day.
 %In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
@@
 
-To test DAS performance we were measured elapsed time and CPU/RAM utilization
-required to collect and aggregate information from different
+To test DAS performance we measured the elapsed time and CPU/RAM utilization
+required to fetch, store and aggregate information from different
 data-services. Table \ref{DAS_benchmark} summarizes the biggest 
@@
 
-The initial DAS implementation was done in Python language \cite{Python}
-and all tests were performed on x64 Linux platform where
-1 CPU was used by DAS server and 1 CPU by MongoDB database server. 
-%The CPU utilization
-%of DAS server was 67\% in average and RAM usage was below 35MB. 
-All tests were performed multiple times to avoid fluctuations. 
-The elapsed time measured during those tests consists of the 
-following components:
+The initial DAS implementation is done in Python language \cite{Python}.
+All tests are performed on a 64 bit Linux platform with
+1 CPU for the DAS server and 1 CPU for the MongoDB database server. 
+All tests are performed multiple times to avoid fluctuations. 
+The elapsed time is measured and consists of the following components:
 \begin{itemize}
@@
 \end{itemize}
-Here {\it retrieval time} is a time required to access data from data-service,
-{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
-is a time required to convert data-service meta-data to DAS records for every object
-we parse, {\it cache insertion} and {\it indexing time} represent time spent on DAS 
-cache back-end, {\it aggregation time} is required to merge objects into DAS records based
-on their common key (\verb+block.name+ in this case) and {\it output creation time}
-is a time required to write DAS records to disk. Please note that last two
-components, {\it aggregation} and {\it output creation time}, were only applied to
-one of the sub-system, either DBS or PhEDEx, depending on insertion order.
-% since the first one was injected into
-%the fresh database, while the second required the merging step.
+Here {\it retrieval time} is the time required to access data from data-service,
+{\it parsing time} is the time required to read and parse the received data
+from the services, and {\it re-mapping time} is the time required to convert 
+that information into DAS. {\it Cache insertion} and {\it indexing time} 
+represents time spent on DAS back-end caching, {\it aggregation time} is
+the time required to merge objects into DAS records based
+on their common keys (\verb+block.name+ in this case) and {\it output creation time}
+is the time required to write DAS records to disk. Please note that last two
+components, {\it aggregation} and {\it output creation time}, are only applied to
+one of the sub-system, either DBS or PhEDEx, depending on the insertion order.
 
@@
 
-The individual tests of DAS components (DBS and PhEDEx) shown roughly 
-the same performance. Although the time of PhEDEx component was bigger 
-due to longer retrieval and parsing time as well as time spent
-during aggregation step.
-%The time of PhEDEx component shown in Table \ref{DAS_benchmark} is divided 
-%among retrieval and parsing, $\sim$110 sec, and merging steps, $\sim$50 sec,
-%respectively. 
-As shown in Table \ref{DAS_benchmark},
-the look-up time spent to access individual components (last column) was quite
+The individual tests of DAS components (DBS and PhEDEx) show roughly 
+the same performance, although the time taken by the PhEDEx component is longer
+due to higher retrieval and parsing time as well as time spent
+during aggregation step. As shown in Table \ref{DAS_benchmark},
+the look-up time spent to access individual components (last column) is quite
 reasonable, roughly 1 second for both systems. While the final time to
-get DAS records on disk was about 50 seconds. This was mostly due to I/O and
+get DAS records on disk is about 50 seconds. This is mostly due to I/O and
 conversion operations between binary data format used by MongoDB, BSON, to DAS
@@
 
-Based on the results of stress tests we measured that we can achieve throughput of
+The results of the stress tests have shown that we can achieve a throughput of
 $\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
-$\sim$7600 docs/sec for 
-reading operations in DAS, which are suitable for our needs. Since all 
-operations were performed on a single CPU a further improvements are expected. 
-Also, a necessary tuning can be done by using C/C++ driver instead of python one.
+$\sim$7600 docs/sec for reading operations in DAS, which is suitable for our 
+needs. Please note, we have relatively small audience of users
+who perform complex queries, while search engines designed for
+a large number of users asking for imprecise, simple data.
+Since all tests were performed on a single CPU, further improvements from
+expanding to multiple CPUs are expected. In addition, more
+performance advantage can be gained by writing dedicated C/C++
+python extensions for data intensive operations.
 
 \section{Summary}
-We presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
-at LHC, CERN, Geneva, Switzerland. It was designed to provide caching and
-aggregation layer on top of the existing relational and non-relation data-services
-mostly in real time fashion. All of the data were retrieved on demand basis,
-while data pre-fetching of most common queries is in our plans. We developed
-prototype in python for DAS system which is under commissioning phase right now.
-The performance studies of DAS system are presented in this paper. 
-The CMS experiment is started to collect its data in December of 2009 and 
-we expect to accumulate a few PB of real data each year. 
-In addition, the Monte-Carlo samples will be produced at this scale.
-Based on test studies performed in CMS, we expect that total size of
-meta-data produced each year by experiment will be of the order of
-500GB. The DAS should sustain such load and will provide generic 
-``data-discovery'' service for CMS experiment in years to come.
+%We are presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
+%at LHC, CERN, Geneva, Switzerland. It is designed to provide caching and
+%aggregation layer on top of the existing relational and non-relation data-services
+%mostly in real time fashion. All data are retrieved on demand basis,
+%while pre-fetching of most common queries is in our plans. We are developed a
+%prototype in Python which is under commissioning phase right now.
+%The performance studies of the DAS system are presented in this paper. 
+%The CMS experiment is started to collect its data in December of 2009 and 
+%we expect to accumulate a few PB of real data each year. 
+%In addition, the Monte-Carlo samples will be produced at this scale.
+%Based on test studies performed in CMS, we expect that total size of
+%meta-data produced each year by experiment will be of the order of
+%500GB. The DAS system should sustain such load and provide generic 
+%``data-discovery'' service for CMS experiment in years to come.
+In this paper we have presented a new data aggregation service (DAS) 
+developed for CMS High-Energy experiment at LHC, CERN, Geneva, Switzerland. 
+We have developed a prototype for DAS, which has been designed to provide 
+caching and aggregation layer on top of the existing relational and 
+non-relational data-services in a close to real time fashion. The data 
+is retrieved on demand, with pre-fetching of common queries, 
+determined from an extensive analytics database, also possible. 
+The DAS system is currently being commissioned for production use by CMS.
+
+Data taking at CMS began in December 2009, and, alongside simulation 
+data, we expect to accumulate petabytes of data yearly. This will have 
+of the order of a terabyte of associated metadata that will need to be 
+easily queried by physicists. From the performance studies of DAS presented 
+in this paper we expect that the system will be able to sustain such load 
+and will provide generic “data-discovery” service for CMS experiment for 
+years to come.
 

### 2010-05-04 src/python/DAS/utils/utils.py
Change to support HTTP headers while making URL calls
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@2026 4525493e-7705-40b1-a816-d608a930855b

@@ -5,8 +5,8 @@
 General set of useful utilities used by DAS
 """
 
-__revision__ = "$Id: utils.py,v 1.70 2010/02/19 17:29:04 valya Exp $"
-__version__ = "$Revision: 1.70 $"
+__revision__ = "$Id: utils.py,v 1.71 2010/02/19 22:28:56 valya Exp $"
+__version__ = "$Revision: 1.71 $"
 __author__ = "Valentin Kuznetsov"
 
 import os
@@ -985,3 +985,13 @@ def extract_http_error(err):
         pass
     return msg
 
+def make_headers(data_format):
+    """
+    Create HTTP header based on input parameters
+    """
+    headers = {}
+    if  data_format.lower() == 'json':
+        headers.update({'Accept':'text/json;application/json'})
+    elif data_format.lower() == 'xml':
+        headers.update({'Accept':'text/xml;application/xml'})
+    return headers

### 2010-05-04 doc/das_paper/paper.tex
Change style from 1p to 3p; enlarge picture due to style change
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1042 4525493e-7705-40b1-a816-d608a930855b
New set of changes wrt reviewers feedback
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@2028 4525493e-7705-40b1-a816-d608a930855b

--- 2010-05-04 doc/das_paper/paper.tex
@@
 way to look-up the desired information.
-Well-known solutions, e.g. data-services, are tightly coupled to a specific data 
-sources and end-users are left with the manual tasks of bookkeeping and 
-relating the information from them.
+A well-known solutions, e.g. data-services, are tighten to a specific data 
+sources and end-users are left with a manual task to bookkeep and 
+relate information from them.
 
-Here we present our work on the Data Aggregation System (DAS) which provides
-the ability to search and aggregate information across different 
-data-services. While designed for the CMS High-Energy Physics 
+Here we present our work on Data Aggregation System (DAS) which provides
+ability to search and aggregate information across different 
+data-services. While designed for CMS High-Energy Physics 
 experiment at LHC, the strategies and technology could be
@@
 The rest of this paper is organized as follows. 
-In section \ref{RelatedWork} we discuss related work in the domain of 
+In section \ref{RelatedWork} we discuss related work in a domain of 
 keyword search over relational data sources.
@@
 Even though the idea of keyword queries over the relational database
-is not new, it is still under discussion in computer science domain.
+is not knew it is still under discussion in computer science domain.
 A few alternative solutions have been proposed to address this issue
@@
 single table or joined tables. Authors of \cite{QueryAnswer} make a step forward
-and evaluate information around matched values. The results are presented as
+and have evaluated information around matched values. The results were presented as
 customized templates to end-users. 
@@
 for more references), are based on an approach that builds some 
-kind of graph over the underlying 
+kind of the graph over the underlying 
 database and generates a new database out of the source database to perform the work.
-The proximity of the results in those techniques are mostly confined to 
-text-based values stored in a database, since numerical ones 
-are mostly useless without knowledge of the context of the input value. 
+The proximity of the results in those techniques are mostly confined to a 
+text-based values stored in a database, since the usage of numerical ones 
+is mostly useless without knowledge of the context of the input value. 
 For example, typing a value 100 in the search input field can lead to 
 plenty of inappropriate matches, e.g. row ids. In order to be
-answered correctly, an additional keyword is required to clarify its meaning, 
+answered correctly it requires an additional keyword to clarify its meaning, 
 which should be matched with particular entity (table.column) in the 
-underlying database. Therefore logical questions, e.g.
+underlying database. Therefore the logical questions, e.g.
 {\it find me the total number of files whose size more then 20 but less then 
 100} are hard to answer precisely using approaches outlined in \cite{DBXplorer, QueryAnswer}, 
-while they can be easily accomplished using SQL.
+while it can be easily accomplished using SQL.
 
 To address these limitations a keyword-based query language was introduced in \cite{DBS-QL}.
-It used a set of pre-defined keys mapped to an underlying schema and a shortest path
+It used a set of pre-defined keys mapped to an underlying schema and shortest path
 algorithm which employed foreign key constraints to build SQL queries. As a result
 the question
-{\it I'm looking for files which contain data taken on a certain date and located at
-a particular site} was represented as simply as \cite{DBS-QL}
+{\it I'm looking for files who contain data taken on certain date and located at
+particular site} was represented as simply as \cite{DBS-QL}
 \begin{verbatim}
@@
 
-At the same time, data aggregation is commonly addressed in the context of
-federated data\-ba\-ses, see for example \cite{FedDB}. The data from different 
+At the same time the data aggregation is commonly addressed in the context of
+federated databases, see for example \cite{FedDB}. The data from different 
 RDMS systems are stored into the federated DB where SQL queries can be used to 
-search for desired data. Apart from many internal obstacles, such as
+search desired data. Apart from many internal obstacles, such as
 schema, domain and naming conflicts, it is external issues that are often
 a large barrier within a distributed, heterogeneous environment where
-different security policies are in place. In the end the usage of SQL and/or
+different security policies are in place. At the end the usage of SQL and/or
 any other Query Language is required to search for data.
@@
 existing data-services whose nature and access policies are in place and well understood.
-The key ingredients are the mapping and analytics services which provide
+The key ingredients were the setup of the mapping and analytics services which provided
 the necessary glue across existing API/notations and user queries. This approach 
@@
 incorporate a broad variety of both mass data storage and data processing
-resources into a single coherent system. CMS uses a
+resources into a single coherent system. CMS uses
 multi-Tiered model, where specific tasks related to data taking such as
@@
 \section{Data Aggregation System\label{DAS}}
-The goal of the Data Aggregation System (DAS) is to fetch and aggregate meta-data 
+The goal of the Data Aggregation System (DAS) was to fetch and aggregate meta-data 
 on demand from existing CMS data-services under one umbrella by using their APIs, 
-security policies, etc. By contrast, the approaches
+security policies, etc. In contrast, the approaches
 discussed in \cite{DBXplorer, QueryAnswer, FedDB}
@@
 data representation and security policies are left intact. Consequently,
-DAS does not require data preservation and transaction capabilities. 
-Instead DAS can be treated as a proxy or caching layer where information 
+DAS does not require a data preservation and transaction capabilities. 
+Instead DAS can be treated as proxy or caching layer where information 
 retrieval and aggregation are triggered dynamically by user queries. 
-Imagine: a user opens a browser and places a query, e.g.
-{\it file block=abc}. The DAS system responds with aggregated infromation about files
-for the provided set of conditions from all CMS data-services. Information
-is either retrieved from the DAS cache or fetched on-demand from relevant
-data-providers. Such a system provides uniform access to all meta-data in CMS 
-via simple, intuitive query interface. It neither forces users to learn 
-the structure of CMS data-services
-nor requires data-providers to agree on common data representations
-and the requirement to have central meta-data repository.
-
-%Below we discuss in details the DAS architecture.
+Below we discuss in details the DAS architecture.
 
 \subsection{DAS architecture}
-The design of the DAS system was based on previous studies of the CMS Data 
+The design of the DAS system was based on previous studies of CMS Data 
 Bookkeeping System (DBS) \cite{DBS, DBS07}. The DBS Data Discovery service
-\cite{DD} was based on the DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
-a syntax similar to SQL, without requiring specific knowledge of the underlying DB schema.
+\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
+a syntax similar to SQL, without specific knowledge of the underlying DB schema.
 Quick adoption and wide usage of DBS-QL in CMS gave us confidence in the chosen approach 
@@
 discovery service.
-Starting from here we designed the DAS system as an
+Starting from this ground we are designed the DAS system as an
 additional layer on top of the existing data-services
-within the CMS computing infrastructure by imposing the following set of requirements:
-support keyword based search queries with the ability to use conditional operators;
-use existing heterogeneous software environments and the
-distributed nature of the data-services;
-preserve the access rules and security policies of individual data-services;
+within CMS computing infrastructure by imposing the following set of requirements:
+support a keyword based search queries with ability to use conditional operators;
+use existing heterogeneous software environment and distributed nature of the data-services;
+preserve access rules and security policies of individual data-services;
 retrieve data on demand and aggregate them if necessary across
@@
 
-The choice of an existing RDMS as the DAS back-end(s) was ruled out for several reasons. 
+\noindent
+The choice of existing RDMS as DAS back-end(s) were ruled out for several reasons. 
 For instance, we didn't require any transactions and data persistency in DAS;
-the dynamically typing of stored meta-data objects was one of the requirements, etc.
+a dynamically typing of stored meta-data objects was one of the requirements, etc.
 Those arguments forced us to look for alternative IT solutions.
-Through the careful analysis of available options, such as file-based and in memory caches, 
-key-value databases and documented-oriented databases, we made our choice in favor 
+Through the careful analysis of available options, such as a file-based and in memory caches, 
+key-value databases, documented-oriented databases we made our choice in favor 
 of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
-MongoDB \cite{MongoDB}. Both are ``schema-less'' databases providing
+MongoDB \cite{MongoDB}. Both systems are ``schema-less'' databases providing
 arbitrary document structure storage, replication and fail-over features. 
@@
 the desired load and size capable of handling our meta-data information. We used 
-the document-oriented database, MongoDB, as a back-end for three 
+document-oriented database, MongoDB, as a back-end for three 
 main DAS components: Mapping, Analytics and DAS cache databases, 
@@
 the core library with support of pluggable modules for data retrieval;
-the caching layer to store and aggregate results into DAS Cache and DAS merge
-databases, respectively; the
+caching layer to store and aggregate results into DAS Cache and DAS merge
+databases, respectively;
 request service to handle user queries;
-the Mapping DB to keep information about data-service APIs, e.g.
-notations and mapping to DAS keys and the
+Mapping DB to keep information about data-service APIs, e.g.
+notations and mapping to DAS keys;
 Analytics DB for query analysis.
@@
 First the results are looked up in a cache and delivered to the requester
-if a superset of the input conditions is found in analytics DB.\footnote{It is important
-to note here that we didn't look-up results in a cache, but rather
-compared the input query condition set to the records in the analytics DB 
-in order to understand if other API calls covered the input request.}
+if superset of input conditions is found in analytics DB.\footnote{It is important
+to note here that we didn't look-up results in a cache and rather
+compared input query condition set to the records in analytics DB 
+in order to understand if other API calls are covered the input request.}
 Otherwise, the appropriate workflow, a set of API calls, is constructed
-for the execution engine. It invokes the API calls via the data-service
-plug-ins and stores the results into the cache. At this step we also
+for the execution engine. It invoked the API calls via data-service
+plug-ins and stored results into the cache. At this step we also
 performed necessary data aggregation for records with similar key-value
@@
  "params": {"node": "*", "block": "required", "se": "*"}, 
- "das2api": [{"pattern": "", "das_key": "file.block.name", 
+ "api2das": [{"pattern": "", "das_key": "file.block.name", 
     "api_param": "block"}, {"pattern": "re.compile('^T[0-3]_')", 
@@
 \end{verbatim}
-represents a DAS mapping record.
+represents DAS mapping record.
 This record can be read as following:
 the \verb+fileReplicas+ API (urn) provided by given url delivers data in
-XML data format and belongs to the \verb+phedex+ system.
+XML data format and belongs to \verb+phedex+ system.
 It accepts \verb+node+, \verb+block+ and \verb+se+ input parameters, shown in \verb+params+ part
-of the record. Each record has two maps: 
-\verb+daskeys+ for mapping input DAS keys into DAS record access keys and
-\verb+das2api+ for mapping DAS access keys into api input parameters.
+of the record. Each record has two maps: \verb+api2das+ for mapping api input parameters into
+DAS keys and \verb+daskeys+ for mapping input DAS keys into DAS record access keys.
 
-The Mapping DB identifies a set of pre-defined DAS keys, 
+Therefore the Mapping DB identifies a set of pre-defined DAS keys, 
 e.g. \verb+file, block+, etc., to be used by end-users in their queries.
@@
 The DAS caching system is used to dynamically fetch
-and aggregate data upon user requests. It consists of the
-DAS cache and the DAS merge independent databases. The former is used to store
-the raw results coming out from data-services, while the latter contains aggregated
-records. Each data record in the DAS cache/merge DBs contains expiration timestamps 
-provided either by the data-service itself or based on default values during 
+and aggregate data upon user requests. It consits of
+DAS cache and DAS merge independent databases. The former one is used to store
+the raw results coming out from data-services, while the later contains aggregated
+records. Each data record in DAS cache/merge DBs contains expiration timestamps 
+provided either by data-service itself or based on default values during 
 data-service registration.
@@
 \end{verbatim}
-Here \verb+r+ represents a triplet \verb+url, api, args+, and 
-\verb+mappind_db+ defines the translation between the input query and the data-service
+Here \verb+r+ represents a triplet \verb+url, api, args+ and 
+\verb+mappind_db+ defines translation between input query and data-service
 APIs which can delivery the data.
@@
 \end{verbatim}
-Here \verb+superset_in_analytics+ checks if the provided \verb+api, args+
-are a subset of api call(s) already registered in the Analytics DB. The 
+Here \verb+superset_in_analytics+ checks if provided \verb+api, args+
+are subset of already registered in Analytics DB api call(s). The 
 \verb+get_data+ and \verb+parser+ are define methods to fetch and parse
@@
 for \verb+file+ record.
-Bulk insert operations are used at this step to increase throughput. 
+We did use a bulk insert operations at this step to increase throughput. 
 \item Get results from the cache
@@
 %common data into the cache. 
-The cache population can be automated by running an
+We can also automate populating the cache by running an 
 independent set of daemons, DAS robots, which monitor the 
@@
 for nested structures, e.g. \verb+block.replica.size+. Our users 
-are able to specify the context of their queries by using the appropriate 
+are able to specify the context of their queries by using appropriate 
 attribute(s). For example, a query for \verb+block+ represents a search for all
@@
 retrieve all run records for the given range of run numbers. 
-As a complement to the queries, we will soon be adding a dynamic index
-of DAS records with a keyword-based search within record content.
+As a complement to the queries, we will also be adding a dynamic index
+of DAS records with a keyword-based search within record content later.
 
-All input queries are transformed into the MongoDB query syntax which is itself
+All input queries are transformed into the MongoDB query syntax who is itself
 a JSON record. It contains a set of selection 
-keys (\verb+fields+) and conditions (\verb+spec+). For example, the query
+keys (\verb+fields+) and conditions (\verb+spec+). For example, a query
 \verb+dataset, file=abc, run > 10+
@@
 %source of information about all data-services participating in DAS.
-Based on the information stored in the Mapping DB we are able to 
+Based on the information stored in Mapping DB we are able to 
 retrieve appropriate data from different data-services, re-map them into 
@@
 for DAS use and add the resulting information to the DAS cache. The records
-are merged (aggregated) based on identical key-value pairs, e.g.
+were merged (aggregated) based on identical key-value pairs, e.g.
 based on \verb+block.name+ key. Here is an example of such an aggregated record:
@@
 about how long the information is valid.
-Usage of this information was a valuable contribution in debugging process of
+Usage of this information was valuable contribution in debugging process of
 data-services themselves, e.g. identification of data-service inconsistencies, 
@@
 \section{Results\label{Results}}
-In this section we discuss preliminary results achieved with the DAS system
+In this section we discuss preliminary results achieved with DAS system
 using different CMS data-services, see \cite{CMS, CMSDataModel}.
@@
 in DAS is around
-75 GB and 200 GB in tables and index sizes, respectively. We estimate 
-collecting around 500 GB of meta-data each year during 
-normal CMS operations. The average query rate to DAS is expected to be
+75 GB and 200 GB in tables and index sizes, respectively. We estimate to
+collect around 500 GB of meta-data each year during 
+normal CMS operations. The average query rate to DAS is expected to be at the 
 $\leq$10K requests a day.
@@
 total time spent in all DAS components.
-The total number of DAS records are calculated as the number of 
+The total number of DAS records are calculated as number of 
 aggregated records plus left overs, the non-matched records from both systems. 
-The look-up time shown in the last column represents the query look-up 
+The look-up time shown in last column represents query look-up 
 time (2.05 sec) and time required to get all records written 
@@
 
-The initial DAS implementation is done in the Python language \cite{Python}.
+The initial DAS implementation is done in Python language \cite{Python}.
 All tests are performed on a 64 bit Linux platform with
@@
 the look-up time spent to access individual components (last column) is quite
-reasonable, roughly 1 second for both systems. The final time to
+reasonable, roughly 1 second for both systems. While the final time to
 get DAS records on disk is about 50 seconds. This is mostly due to I/O and
-conversion operations from the binary data format used by MongoDB, BSON, 
-to the DAS data format, JSON.
+conversion operations between binary data format used by MongoDB, BSON, to DAS
+data format, JSON.
 
-The individual benchmarks of MongoDB have shown that it can deliver an insert rate of
-20K docs/sec and provide a look-up time of $10^{-5}$ sec per random document.
+The individual benchmarks of MongoDB have shown that it can deliver insert rate of
+20K docs/sec and provide look-up time of $10^-5$ sec per random document.
 The results of DAS stress tests have demonstrated that we can achieve a throughput of
@@
 which is suitable for our needs. 
-Please note, we have a relatively small audience of users
+Please note, we have relatively small audience of users
 who perform complex queries and get precise results, while search 
@@
 \section{Future work}
-In the near future we plan to put DAS into production and confirm its scalability, 
-transparency and durability for various CMS data-services. 
-Our interests are plug-able interface for participating data-providers,
-analytics framework and pre-fetch strategies as well as horizontal scaling of the
-system. Our goal is to sustain 1TB of meta-data per year. 
-In a long term, we would like to use DAS across various
-data-services in the HEP community and other disciplines. Even though DAS was developed
-for the CMS experiment, its architecture is transparent to participating 
-data-providers and meta-data content. Therefore the generalization of DAS is feasible.
-Such work has began under the umbrella of DISCOVER Research Service Group at 
-Cornell University \cite{DRSG} which has expressed their interest to use DAS in other 
-scientific domains.
+The DAS system was desgined for CMS HEP experiment, while its architecture is
+transparent to consumed meta-data, data formats and participating data-providers. 
+In the current year we plan to put DAS into production and confirm its scalability, 
+transparency and durability for various CMS data-services. Our goal is to sustain 
+in DAS 1TB of meta-data per year. The DISCOVER Research Service Group at Cornell 
+University \cite{DRSG} is expressed its interests to apply DAS for other 
+scientific domains and turn it into generic open-source project.
 
 \section{Summary}
+%We are presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
+%at LHC, CERN, Geneva, Switzerland. It is designed to provide caching and
+%aggregation layer on top of the existing relational and non-relation data-services
+%mostly in real time fashion. All data are retrieved on demand basis,
+%while pre-fetching of most common queries is in our plans. We are developed a
+%prototype in Python which is under commissioning phase right now.
+%The performance studies of the DAS system are presented in this paper. 
+%The CMS experiment is started to collect its data in December of 2009 and 
+%we expect to accumulate a few PB of real data each year. 
+%In addition, the Monte-Carlo samples will be produced at this scale.
+%Based on test studies performed in CMS, we expect that total size of
+%meta-data produced each year by experiment will be of the order of
+%500GB. The DAS system should sustain such load and provide generic 
+%``data-discovery'' service for CMS experiment in years to come.
 In this paper we have presented a new data aggregation service (DAS) 
-developed for the CMS High-Energy experiment at LHC, CERN, Geneva, Switzerland. 
-We have developed a prototype for DAS, designed to provide 
-caching and aggregation layers on top of the existing relational and 
+developed for CMS High-Energy experiment at LHC, CERN, Geneva, Switzerland. 
+We have developed a prototype for DAS, which has been designed to provide 
+caching and aggregation layer on top of the existing relational and 
 non-relational data-services in a close to real time fashion. The data 

### 2010-05-04 doc/das_paper/procs-sample.tex
New set of changes wrt reviewers feedback
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@2028 4525493e-7705-40b1-a816-d608a930855b

--- 2010-05-04 doc/das_paper/paper.tex
@@
 
-%% The '1p' and 'times' class options of elsarticle are used for Elsevier CRC
-\documentclass[1p,times]{elsarticle}
+%% The '3p' and 'times' class options of elsarticle are used for Elsevier CRC
+\documentclass[3p,times]{elsarticle}
 
@@
 %% Example \runauth{C.V. Radhakrishnan et al.}
-\runauth{V. Kuznetsov, D. Evans, S. Metson}
+\runauth{G.D. van Albada}
 
@@
 
-\title{The CMS Data Aggregation System}
+\title{Click \emph{here}, type the title of your paper, Capitalize first letter}
 
@@
 
-\author[vkuznet]{Valentin Kuznetsov\corref{cor1}}
-%\author[vkuznet]{Valentin Kuznetsov}
-\address[vkuznet]{Cornell University, Ithaca, New York, USA}
-\ead{vkuznet@gmail.com}
-\author[evans]{Dave Evans}
-\address[evans]{Fermilab, Batavia, Illinois, USA}
-%\ead{evansde@fnal.gov}
-\author[metson]{Simon Metson}
-\address[metson]{Bristol University, Bristol, UK}
-%\ead{s.metson@bristol.ac.uk}
+\author{Dick van Albada\corref{cor1}\fnref{ref2}}
+\fntext[ref2]{Corresponding author}
+\cortext[cor1]{\ead {G.D.vanAlbada@uva.nl}}
 
-\cortext[cor1]{Corresponding author}
+\address{Computational Science, University of Amsterdam}
 
 \begin{abstract}
-Meta-data plays a significant role in large modern enterprises, 
-research experiments and digital libraries where it comes from many different 
-sources and is distributed in a variety of digital formats. 
-It is organized and managed by constantly evolving software using 
-both relational and non-relational data sources. Even though we can apply
-an information retrieval approach to non-relational data sources,
-we can't do so for relational ones, where information is accessed via
-a pre-established set of data-services.
-
-Here we discuss a new data aggregation system which consumes, 
-indexes and delivers information from different relational and 
-non-relational data sources to answer cross data-service queries 
-and explore meta-data associated with petabytes of experimental data. 
-We combine the simplicity of keyword-based search with the precision of RDMS
-under the new system. The aggregated information is collected from various sources,
-allowing end-users to place dynamic queries, get precise answers and 
-trigger information retrieval on demand. Based on the use cases of the CMS experiment, 
-we have performed a set of detailed, large scale tests the results of which 
-we present in this paper.
+Put your abstract here - do not forget to add relevant keywords. This sample file uses
+the text from the MS Word template to allow some comparison of the formats.
+%% Text of abstract
 \end{abstract}
@@
 %% keywords here, in the form: keyword \sep keyword
-Meta-data\sep data aggregation\sep information discovery\sep HEP.
-
+template \sep latex \sep Procedia
 %% PACS codes here, in the form: \PACS code \sep code
@@
 %% or \MSC[2008] code \sep code (2000 is the default)
+
 \end{keyword}
@@
 %% main text
-\section{Introduction}
-The European Organization for Nuclear Research, known as CERN, plays a leading
-role in fundamental studies of physics. It is also known as a place where
-many innovations in the area of computer science were developed, e.g. The World Wide Web.
-Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
-Physics (HEP), promising to deliver a few PB of data each year. 
-At this scale the ``information discovery'' within a heterogeneous, distributed 
-environment becomes a key ingredient of successful data analysis.
-The data and associated meta-data are produced in variety of forms and digital formats.
-They are stored and retrieved from relational and non-relational data-sources, such as 
-RDMS systems, document oriented databases, blogs, twikies, file systems,
-customized applications, etc. Working in such an environment requires a lot 
-of expertise and demands a straightforward, simple
-way to look-up the desired information.
-Well-known solutions, e.g. data-services, are tightly coupled to a specific data 
-sources and end-users are left with the manual tasks of bookkeeping and 
-relating the information from them.
+\section{intro}
+\label{intro}
+(10 pt) Here introduce the paper, and put a nomenclature if necessary, in a box with the same font size as the rest of the paper.
+The paragraphs continue from here and are only separated by headings, subheadings, images and formulae.
+The section headings are arranged by numbers, bold and 10 pt.
+Here follows further instructions for authors.
+\subsection{Structure}
+\label{structure}
+Files should be in MS Word format only and should be formatted for direct printing.
+Figures and tables should be embedded and not supplied separately.
+Please make sure that you use as much as possible normal fonts in your documents.
+Special fonts, such as fonts used in the Far East (Japanese, Chinese, Korean, etc.) may cause problems during processing.
+To avoid unnecessary errors you are strongly advised to use the 'spellchecker' function of MS Word.
+Follow this order when typing manuscripts: Title, Authors, Affiliations, Abstract, Keywords, Main text (including figures and tables), Acknowledgements, References, Appendix.
+Collate acknowledgements in a separate section at the end of the article and do not include them on the title page, as a footnote to the title or otherwise.
+\\
+Bulleted lists may be included and should look like this:
+\begin{itemize*}
+	\item First point
+  \item Second point
+  \item And so on
+\end{itemize*}
+Ensure that you return to the 'Els-body-text' style, the style that you will mainly be using for large blocks of text, when you have completed your bulleted list. 
+Please do not alter the formatting and style layouts which have been set up in this template document.
+As indicated in the template, papers should be prepared in single column format suitable for direct printing onto A4 paper (8.3in x 11.7in/210mm x 297mm).
+Do not number pages on the front, as page numbers will be added separately for the preprints and the Proceedings.
+Leave a line clear between paragraphs.
+All the required style templates are provided in this document with the appropriate name supplied, e.g.
+choose 1.
+Els1st-order-head for your first order heading text, els-abstract-text for the abstract text etc.
+\subsection{Tables}
+\label{table}
+All tables should be numbered with Arabic numerals.
+Headings should be placed above tables, left justified.
+Leave one line space between the heading and the table.
+Only horizontal lines should be used within a table, to distinguish the column headings from the body of the table, and immediately above and below the table.
+Tables must be embedded into the text and not supplied separately.
+Below is an example which authors may find useful.
 
-Here we present our work on the Data Aggregation System (DAS) which provides
-the ability to search and aggregate information across different 
-data-services. While designed for the CMS High-Energy Physics 
-experiment at LHC, the strategies and technology could be
-used elsewhere.
+%% The Appendices part is started with the command \appendix;
+%% appendix sections are then done as normal sections
+%% \appendix
 
-The rest of this paper is organized as follows. 
-In section \ref{RelatedWork} we discuss related work in the domain of 
-keyword search over relational data sources.
-Section \ref{DataModel} breifly describes the Compact Muon Solenoid (CMS) experiment 
-and its data model. In section \ref{DAS} we present the architecture 
-of the DAS system, including discussion of its
-various components. Finally, our results are summarized in section \ref{Results}.
-
-\section{Related Work\label{RelatedWork}}
-Even though the idea of keyword queries over the relational database
-is not new, it is still under discussion in computer science domain.
-A few alternative solutions have been proposed to address this issue
-in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
-i.e. retrieval of documents that contain all query keywords, has
-been discussed. It requires that all keywords are matched in a row from a
-single table or joined tables. Authors of \cite{QueryAnswer} make a step forward
-and evaluate information around matched values. The results are presented as
-customized templates to end-users. 
-These and other related works, (see \cite{DBXplorer, QueryAnswer} 
-for more references), are based on an approach that builds some 
-kind of graph over the underlying 
-database and generates a new database out of the source database to perform the work.
-The proximity of the results in those techniques are mostly confined to 
-text-based values stored in a database, since numerical ones 
-are mostly useless without knowledge of the context of the input value. 
-For example, typing a value 100 in the search input field can lead to 
-plenty of inappropriate matches, e.g. row ids. In order to be
-answered correctly, an additional keyword is required to clarify its meaning, 
-which should be matched with particular entity (table.column) in the 
-underlying database. Therefore logical questions, e.g.
-{\it find me the total number of files whose size more then 20 but less then 
-100} are hard to answer precisely using approaches outlined in \cite{DBXplorer, QueryAnswer}, 
-while they can be easily accomplished using SQL.
-
-To address these limitations a keyword-based query language was introduced in \cite{DBS-QL}.
-It used a set of pre-defined keys mapped to an underlying schema and a shortest path
-algorithm which employed foreign key constraints to build SQL queries. As a result
-the question
-{\it I'm looking for files which contain data taken on a certain date and located at
-a particular site} was represented as simply as \cite{DBS-QL}
-\begin{verbatim}
-find file where date 2009-01-02 11:59 CET and site = T2
-\end{verbatim}
-without specifying any relational details.
-This solution represents a common use case in the HEP community and provides
-an intuitive mapping between the mental model of the end-users and 
-the underlying database schema.
-
-At the same time, data aggregation is commonly addressed in the context of
-federated data\-ba\-ses, see for example \cite{FedDB}. The data from different 
-RDMS systems are stored into the federated DB where SQL queries can be used to 
-search for desired data. Apart from many internal obstacles, such as
-schema, domain and naming conflicts, it is external issues that are often
-a large barrier within a distributed, heterogeneous environment where
-different security policies are in place. In the end the usage of SQL and/or
-any other Query Language is required to search for data.
-
-To avoid these limitations we resolved to build a keyword-search based system around
-existing data-services whose nature and access policies are in place and well understood.
-The key ingredients are the mapping and analytics services which provide
-the necessary glue across existing API/notations and user queries. This approach 
-allows us to perform dynamic data aggregation across various data-services, achieve
-precise answers using keyword-based queries
-and organize information retrieval on demand.
-
-\section{CMS data model\label{DataModel}}
-The Compact Muon Solenoid, (CMS) \cite{CMS} 
-is one of the two general purpose particle physics detectors built on 
-the Large Hadron Collider (LHC) at CERN in Switzerland and France. 
-The experiment is designed to explore the frontiers of physics and provide physicists
-with the ability to look at the conditions presented in the early stage of our Universe.
-More then 3000 physicists from 183 institutions representing 38 countries 
-are involved in the design, construction and maintenance of the experiment.
-
-The CMS distributed computing and data model \cite{CMSDataModel} 
-is designed to process and efficiently manage the few PBs of data expected each year
-during normal operation of the LHC. The computing resources available to the CMS
-collaboration are geographically distributed, 
-interconnected via high throughput networks and operated by means 
-of various Grid software infrastructure. The model is designed to
-incorporate a broad variety of both mass data storage and data processing
-resources into a single coherent system. CMS uses a
-multi-Tiered model, where specific tasks related to data taking such as
-processing, archival and distribution, are assigned to each tier based
-on the CMS data model. For example, the Tier-0 center at CERN is responsible
-for archiving the data coming out from the detector, prompt first pass reconstruction
-and data distribution to Tier-1 centers, located around the world.
-The Tier-1 centers provide archival storage, CPU power for high
-priority selection and re-reconstructions, and distribution points
-to the Tier 2 resources.
-The Tier-2 centers are dedicated for user analysis tasks and production of simulated data.
-
-A broad variety of data-services have been designed and developed to
-maintain detector and production operations, including detector
-conditions databases, data bookkeeping services,
-data transfer and job monitoring tasks. While the majority of these
-services are located at CERN, they need to be accessed remotely, and
-combined with other distributed services.
-
-Once such a conglomerate of data-services start operating the obvious
-question arises: how do we find the desired information across multiple data-services
-in our distributed environment? Even though the individual data-services were designed
-to answer specific questions about the data they serve, the ability to search and relate
-information among them remains a tedious human task. The growing amount of information
-and desire to make cross-service queries led us to design and develop the new
-Data Aggregation System.
-
-\section{Data Aggregation System\label{DAS}}
-The goal of the Data Aggregation System (DAS) is to fetch and aggregate meta-data 
-on demand from existing CMS data-services under one umbrella by using their APIs, 
-security policies, etc. By contrast, the approaches
-discussed in \cite{DBXplorer, QueryAnswer, FedDB}
-enforce a publication step of data into a dedicated database 
-or central repository \cite{iRODS}. Therefore, quite often it was required
-that data should be converted into common data structure/format in order to enable
-aggregation, see \cite{OpenArchive}. In DAS the data are provided
-by existing data-services and infrastructure, whose implementation,
-data representation and security policies are left intact. Consequently,
-DAS does not require data preservation and transaction capabilities. 
-Instead DAS can be treated as a proxy or caching layer where information 
-retrieval and aggregation are triggered dynamically by user queries. 
-Imagine: a user opens a browser and places a query, e.g.
-{\it file block=abc}. The DAS system responds with aggregated infromation about files
-for the provided set of conditions from all CMS data-services. Information
-is either retrieved from the DAS cache or fetched on-demand from relevant
-data-providers. Such a system provides uniform access to all meta-data in CMS 
-via simple, intuitive query interface. It neither forces users to learn 
-the structure of CMS data-services
-nor requires data-providers to agree on common data representations
-and the requirement to have central meta-data repository.
-
-%Below we discuss in details the DAS architecture.
-
-\subsection{DAS architecture}
-The design of the DAS system was based on previous studies of the CMS Data 
-Bookkeeping System (DBS) \cite{DBS, DBS07}. The DBS Data Discovery service
-\cite{DD} was based on the DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
-a syntax similar to SQL, without requiring specific knowledge of the underlying DB schema.
-Quick adoption and wide usage of DBS-QL in CMS gave us confidence in the chosen approach 
-and provided a basis for building the DAS system.
-
-As was pointed out in \cite{Arms} the mixed content and 
-mixed meta-data and meta-data consistency 
-of the underlying services needs to be carefully
-considered in the design of the system if it is to be a successful information 
-discovery service.
-Starting from here we designed the DAS system as an
-additional layer on top of the existing data-services
-within the CMS computing infrastructure by imposing the following set of requirements:
-support keyword based search queries with the ability to use conditional operators;
-use existing heterogeneous software environments and the
-distributed nature of the data-services;
-preserve the access rules and security policies of individual data-services;
-retrieve data on demand and aggregate them if necessary across
-multiple data-services;
-be transparent to data-services implementation and their data formats.
-%\begin{itemize}
-%\item support a keyword based search queries with ability to use conditional operators;
-%\item use existing heterogeneous software environment and distributed nature of the data-services;
-%\item preserve access rules and security policies of individual data-services;
-%\item retrieve data on demand and aggregate them if necessary across
-%multiple data-services;
-%\item be transparent to data-services implementation and their data formats;
-%\end{itemize}
-\begin{figure}[htb]
-\centering
-\includegraphics[width=90mm]{DAS_Cache_and_Analytics.pdf}
-\caption{
-The DAS architecture diagram. It consists of cache server,
-analytics, mapping and cache databases. The data in a cache can be shared
-across multiple queries. The information retrieval is triggered
-on demand basis by invoking appropriate data-service APIs. The
-optional DAS robots (Unix daemons) can be used to pre-fetch
-the most popular query requests. The solid, dashed and dotted lines 
-represent outer, inner and optional process communications between
-different DAS components, respectively.
-}
-\label{DAS_cache}
-\end{figure}
-
-The choice of an existing RDMS as the DAS back-end(s) was ruled out for several reasons. 
-For instance, we didn't require any transactions and data persistency in DAS;
-the dynamic typing of stored meta-data objects was one of the requirements, etc.
-Those arguments forced us to look for alternative IT solutions.
-Through the careful analysis of available options, such as file-based and in memory caches, 
-key-value databases and documented-oriented databases, we made our choice in favor 
-of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
-MongoDB \cite{MongoDB}. Both are ``schema-less'' databases providing
-arbitrary document structure storage, replication and fail-over features. 
-We made our decision in favor of MongoDB, due to its support of dynamic queries, 
-full indexes, including inner objects and embedded arrays,
-and auto-sharding. Our preliminary benchmarks have shown that it can sustain
-the desired load and size capable of handling our meta-data information. We used 
-the document-oriented database, MongoDB, as a back-end for three 
-main DAS components: Mapping, Analytics and DAS cache databases, 
-which will be discussed next.
-
-The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of
-the core library with support of pluggable modules for data retrieval;
-the caching layer to store and aggregate results into DAS Cache and DAS merge
-databases, respectively; the
-request service to handle user queries;
-the Mapping DB to keep information about data-service APIs, e.g.
-notations and mapping to DAS keys and the
-Analytics DB for query analysis.
-
-\subsection{DAS core\label{DAS_core}}
-The DAS core library is responsible for data retrieval on demand. 
-It has the following components: input query validator and parser,
-query analyzer, workflow builder, execution engine and data-service 
-plug-ins. The communication between client (web server or CLI) and 
-DAS cache server is done via the REST \cite{REST} interface. 
-Each input query is recorded into the Analytics DB for further 
-analysis, and is decomposed into a set of conditions and/or selection keys.
-First the results are looked up in a cache and delivered to the requester
-if a superset of the input conditions is found in analytics DB.\footnote{It is important
-to note here that we didn't look-up results in a cache, but rather
-compared the input query condition set to the records in the analytics DB 
-in order to understand if other API calls covered the input request.}
-Otherwise, the appropriate workflow, a set of API calls, is constructed
-for the execution engine. It invokes the API calls via the data-service
-plug-ins and stores the results into the cache. At this step we also
-performed necessary data aggregation for records with similar key-value
-pairs. The data-service plug-in implementations are similar and their 
-logic is largely abstracted, apart from a few extensions such as specific
-security modules for data-services accessible via private network.
-Finally, the results are retrieved from the cache and delivered to the user.
-
-\subsection{DAS Mapping DB\label{MappingDB}}
-The Mapping DB is used to track services known to DAS and how DAS should interact with them.
-It collects
-information about data-service APIs, translates API input/output
-parameters into DAS keys and contains mapping between DAS records
-and their UI representation.
-
-On identifying API's that are accessible to DAS we store
-their names, input parameters and associated DAS keys
-into the Mapping DB. For example
-\begin{verbatim}
-{"system": "phedex", "format": "XML", "urn": "fileReplicas",
- "url": "http://a.b.com/phedex/datasvc/xml/prod/fileReplicas", 
- "params": {"node": "*", "block": "required", "se": "*"}, 
- "das2api": [{"pattern": "", "das_key": "file.block.name", 
-    "api_param": "block"}, {"pattern": "re.compile('^T[0-3]_')", 
-    "das_key": "site", "api_param": "node"}], 
- "daskeys": [{"map": "file.name", "key": "file", "pattern": ""}, 
-             {"map": "file.block.name", "key": "block", "pattern": ""}]}
-\end{verbatim}
-represents a DAS mapping record.
-This record can be read as following:
-the \verb+fileReplicas+ API (urn) provided by given url delivers data in
-XML data format and belongs to the \verb+phedex+ system.
-It accepts \verb+node+, \verb+block+ and \verb+se+ input parameters, shown in \verb+params+ part
-of the record. Each record has two maps: 
-\verb+daskeys+ for mapping input DAS keys into DAS record access keys and
-\verb+das2api+ for mapping DAS access keys into api input parameters.
-
-The Mapping DB identifies a set of pre-defined DAS keys, 
-e.g. \verb+file, block+, etc., to be used by end-users in their queries.
-Each key is mapped onto set of API calls, whose output is retrieved,
-cached and aggregated as required.
-
-\subsection{DAS Analytics DB}
-The DAS Analytics DB collects information on user requests against the system.
-Each request that DAS handles is recorded.
-The Analytics DB tracks: the users request; how DAS has mapped the request 
-onto data service API's; the time taken by the remote data services to 
-process the API calls.
-
-By recording this information we can plan pre-fetch strategies for common queries,
-identify issues in remote data services and cross check that DAS resolves requests
-in a deterministic manner.
-
-% Upon its
-%decomposition into set of selection keys and conditions we also recorded
-%API name and input parameters into the Analytics DB and 
-%kept updating their counter information for repeated calls. 
-%This allowed to keep track of frequency of API calls and establish 
-%pre-fetch strategies for most common queries.
-
-\subsection{DAS caching system}
-The DAS caching system is used to dynamically fetch
-and aggregate data upon user requests. It consists of the
-DAS cache and the DAS merge independent databases. The former is used to store
-the raw results coming out from data-services, while the latter contains aggregated
-records. Each data record in the DAS cache/merge DBs contains expiration timestamps 
-provided either by the data-service itself or based on default values during 
-data-service registration.
-We use the following algorithm (code snippets are written in Python \cite{Python}):
-\begin{enumerate}[1.]
-\item Decompose query into set of conditions (\verb+conds+) 
-and selection keys (\verb+skeys+).
-\item Check cache for query hash. If found, look-up results from the cache,
-otherwise proceed to the next step.
-\item Identify list of APIs which can serve requested data
-\begin{verbatim}
-apilist = [r for r in mapping_db(skeys, conds)]
-\end{verbatim}
-Here \verb+r+ represents a triplet \verb+url, api, args+, and 
-\verb+mappind_db+ defines the translation between the input query and the data-service
-APIs which can delivery the data.
-\item Iterate over APIs
-\begin{verbatim}
-for url, api, args in apilist:
-    if  superset_in_analytics(api, args):
-        continue
-    for row in parser(get_data(url, api, args)):
-        yield row
-\end{verbatim}
-Here \verb+superset_in_analytics+ checks if the provided \verb+api, args+
-are a subset of api call(s) already registered in the Analytics DB. The 
-\verb+get_data+ and \verb+parser+ are define methods to fetch and parse
-data, respectively.
-\item Perform aggregation by iterating over records from the previous step
-\begin{verbatim}
-if cache_has(daskey):
-   existing_record = get_from_cache(daskey)
-   merge(row, existing_record)
-   insert_into_cache(row)
-else:
-   insert_into_cache(row)
-\end{verbatim}
-The \verb+daskey+
-is a key which identifies the record defined in Mapping DB, e.g. \verb+file.name+
-for \verb+file+ record.
-Bulk insert operations are used at this step to increase throughput. 
-\item Get results from the cache
-\end{enumerate}
-%We also have an option to run independent set of daemons, DAS robots, who
-%monitor Analytics DB for most popular requests and pre-fetch the most
-%common data into the cache. 
-The cache population can be automated by running an
-independent set of daemons, DAS robots, which monitor the 
-Analytics DB for suitable requests (for example queries that are 
-popular, periodic in nature or expensive) and pre-fetch data into the cache.
-
-\subsection{DAS queries}
-All DAS queries are expressed in a free text-based form, either as a 
-set of keywords or key-value pairs, where a pair can represent
-condition, e.g. \verb+key > value+. We use a pre-defined set of keys registered 
-in DAS to describe common entities of our data. For example, \verb+dataset+, 
-\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. The choice of keys
-is based on the analysis of day to day user operations \cite{DBS07}.
-Each key represents a DAS record. Every record is stored into the cache
-in JSON format. Due to the ``schema-less'' nature of the underlying MongoDB
-back-end we are able to store DAS records of arbitrary structure, e.g.
-dictionary, lists, key-value pairs, etc. Therefore every DAS key has a 
-set of attributes describing its JSON structure. For instance a document
-\begin{verbatim}
-{"block": {"name": "abc", "size": 1, "replica": [...]}}
-\end{verbatim}
-is associated with a \verb+block+ key who has the following attributes
-\verb+name+, \verb+size+, \verb+replica+. The set of attributes is based 
-on a set of meta-data information stored in a record, allowing
-for nested structures, e.g. \verb+block.replica.size+. Our users 
-are able to specify the context of their queries by using the appropriate 
-attribute(s). For example, a query for \verb+block+ represents a search for all
-block records rather then the ``block'' string, while the query
-\verb+block.name=abc*+ instructs the system to look-up appropriate blocks
-with matching names. Numerical values, conditions and grouping are
-also easy to express, for example, a query
-of the form \verb+run > 10 and run < 100+ can be used to
-retrieve all run records for the given range of run numbers. 
-As a complement to the queries, we will soon be adding a dynamic index
-of DAS records with a keyword-based search within record content.
-
-All input queries are transformed into the MongoDB query syntax which is itself
-a JSON record. It contains a set of selection 
-keys (\verb+fields+) and conditions (\verb+spec+). For example, the query
-\verb+dataset, file=abc, run > 10+
-is represented as
-\begin{verbatim}
-{"fields": ["dataset"], "spec": {"file": "abc", "run": {"$gt": 10}}}
-\end{verbatim}
-Therefore users queries are stored into Analytics DB similar to other DAS 
-records and used for further analysis.
-
-\subsection{Data-services and aggregation}
-As discussed in section \ref{MappingDB} the DAS Mapping DB 
-is the authoritative source of information on the relationships between 
-all the data services known to DAS.
-%became an authorative
-%source of information about all data-services participating in DAS.
-Based on the information stored in the Mapping DB we are able to 
-retrieve appropriate data from different data-services, re-map them into 
-DAS notations and aggregate them on demand. For instance, the Data 
-Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
-provide information about CMS file block structure and locations.
-DBS associates groups of files into blocks, while PhEDEx tracks the blocks
-location. 
-%In the former case, DBS 
-%stored information about blocks and their relationships to other 
-%physics objects, such as dataset and files, while PhEDEx kept
-%information about its location and replicas.
-
-When a query for block meta-data is made, we are
-%Upon a user request to 
-%find a block meta-data, we were 
-able to identify the appropriate set of 
-DBS and PhEDEx APIs, retrieve the raw data from the services, format it
-for DAS use and add the resulting information to the DAS cache. The records
-are merged (aggregated) based on identical key-value pairs, e.g.
-based on \verb+block.name+ key. Here is an example of such an aggregated record:
-\begin{verbatim}
-{"das_id": ["4b195404e2194e21c4000030", "4b1953fee2194e21c4000002"], 
- "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
-            "site": "T1_US", "nfiles": "18", "se": "a.b.com"}, {...}]},
-           {"name": "/a/b/c#123", "nevents": "570200", "path":"/a/b/c"}],
- "das": {"expire": 1259952908}}
-\end{verbatim}
-The \verb+das_id+ indicates the records stored in the DAS Analytics DB,
-the \verb+block+ part represents the aggregated block meta-data information from DBS
-and PhEDEx and the \verb+das+ part contains expiration time stamps
-about how long the information is valid.
-Usage of this information was a valuable contribution in debugging process of
-data-services themselves, e.g. identification of data-service inconsistencies, 
-latency studies, etc.
-
-\section{Results\label{Results}}
-In this section we discuss preliminary results achieved with the DAS system
-using different CMS data-services, see \cite{CMS, CMSDataModel}.
-
-%using the following CMS data-services:
-%Data Bookkeeping System (DBS) \cite{DBS}, the CMS meta-data service;
-%Physics Experiment Data Export (PhEDEx) \cite {PhEDEx}, the 
-%CMS data location and file transfer system;
-%SiteDB \cite{SiteDB}, the CMS site resources;
-%RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
-%Dashboard \cite{Dashboard}, the LHC virtual organization
-%monitoring data-service portal.
-
-%Each data-service has its own scope, size and peculiarities. 
-%For instance, the data in PhEDEx is transient due to constant 
-%migration of CMS data; the DBS system is divided into several individual 
-%instances with sizes and lifetimes driven by specific processing tasks; 
-%the RunSummary data-service is operated on private network accessible 
-%at CMS online operation center.
-
-At the moment, the amount of meta-data stored in all services participating
-in DAS is around
-75 GB and 200 GB in tables and index sizes, respectively. We estimate 
-collecting around 500 GB of meta-data each year during 
-normal CMS operations. The average query rate to DAS is expected to be
-$\leq$10K requests a day.
-%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
-%billion events, 330TB in 32 datasets:
-%AlCa/Calibration RAW: 926 million events, 29TB
-%Physics RAW: 1.24 billion events, 147TB
-%PromptReco: 1.24 billion events, 111TB
-%Bulk ALCARECO: 2.34 billion events, 11TB
-%Express FEVT: 105 million events, 24TB
-%Express ALCARECO: 153 million events, 0.99TB
-%HLTMON FEVTHLTALL: 31 million events, 7.4TB
-%Above event numbers have significant overlap by design, 
-%2.2 billion unique events, including 524 million cosmic muon
-%triggers
-% 0.3PB - 50GB meta-data in DBS
-% 3PB/y - x GB
-
-To test DAS performance we measured the elapsed time and CPU/RAM utilization
-required to fetch, store and aggregate information from different
-data-services. Table \ref{DAS_benchmark} summarizes the biggest 
-contributors so far, the block meta-data information provided by DBS and 
-PhEDEx CMS data-services.
-
-\begin{table*}[hbt]
-\centering
-\begin{tabular}{lllll}\hline
-\hline
-System & Format & Records & Elapsed time & Elapsed time \\
-& & & no cached data & w/ cached data \\
-\hline
-DBS yield & XML & 386943 & 68 sec & 0.98 sec \\
-PhEDEx yield & XML & 189679 & 107 sec & 0.98 sec \\
-Merge step & JSON & 576622 & 63 sec & 0.9 sec \\
-DAS total & JSON & 392635 & 238 sec & 2.05/50.7 sec \\
-\hline
-\hline
-\end{tabular}
-\caption{Time required to fetch, parse and aggregate block information
-from DBS and PhEDEx systems. The last row shows aggregated results and 
-total time spent in all DAS components.
-The total number of DAS records are calculated as the number of 
-aggregated records plus left overs, the non-matched records from both systems. 
-The look-up time shown in the last column represents the query look-up 
-time (2.05s) and time required to get all records written 
-to disk (50.7s).}
-\label{DAS_benchmark}
-\end{table*}
-
-The initial DAS implementation is done in the Python language \cite{Python}.
-All tests are performed on a 64 bit Linux platform with
-1 CPU for the DAS server and 1 CPU for the MongoDB database server. 
-All tests are performed multiple times to avoid fluctuations. 
-The elapsed time is measured and consists of the following components:
-\begin{itemize}
-\item[]
-{\it
-Elapsed time = retrieval time + parsing time + re-mapping time 
-        + cache insertion/indexing time 
-        + aggregation time + output creation time
-}
-\end{itemize}
-Here {\it retrieval time} is the time required to access data from data-service,
-{\it parsing time} is the time required to read and parse the received data
-from the services, and {\it re-mapping time} is the time required to convert 
-that information into DAS. {\it Cache insertion} and {\it indexing time} 
-represents time spent on DAS back-end caching, {\it aggregation time} is
-the time required to merge objects into DAS records based
-on their common keys (\verb+block.name+ in this case) and {\it output creation time}
-is the time required to write DAS records to disk.
-% Please note that last two
-%components, {\it aggregation} and {\it output creation time}, are only applied to
-%one of the sub-system, either DBS or PhEDEx, depending on the insertion order.
-
-The individual tests of DAS components (DBS and PhEDEx) show roughly 
-the same performance, although the time taken by the PhEDEx component is longer
-due to more complex records which leads to higher retrieval and parsing time. 
-As shown in Table \ref{DAS_benchmark},
-the look-up time spent to access individual components (last column) is quite
-reasonable, roughly 1 second for both systems. The final time to
-get DAS records on disk is about 50 seconds. This is mostly due to I/O and
-conversion operations from the binary data format used by MongoDB, BSON, 
-to the DAS data format, JSON.
-
-The individual benchmarks of MongoDB have shown that it can deliver an insert rate of
-20K docs/sec and provide a look-up time of $10^{-5}$ sec per random document.
-The results of DAS stress tests have demonstrated that we can achieve a throughput of
-$\sim$6000 docs/sec for raw cache population, 
-%$\sim$6000 docs/sec for aggregation and 
-$\sim$7600 docs/sec for reading and writing DAS records to disk,
-which is suitable for our needs. 
-Please note, we have a relatively small audience of users
-who perform complex queries and get precise results, while search 
-engines are designed for a large number of users asking for imprecise, simple data.
-Since all tests were performed on a single CPU, further improvements from
-expanding to multiple CPUs are expected. In addition, more
-performance advantage can be gained by writing dedicated C/C++
-python extensions for data intensive operations.
-
-\section{Future work}
-In the near future we plan to put DAS into production and confirm its scalability, 
-transparency and durability for various CMS data-services. 
-Our interests are plug-able interface for participating data-providers,
-analytics framework and pre-fetch strategies as well as horizontal scaling of the
-system. Our goal is to sustain 1TB of meta-data per year. 
-In a long term, we would like to use DAS across various
-data-services in the HEP community and other disciplines. Even though DAS was developed
-for the CMS experiment, its architecture is transparent to participating 
-data-providers and meta-data content. Therefore the generalization of DAS is feasible.
-Such work has began under the umbrella of DISCOVER Research Service Group at 
-Cornell University \cite{DRSG} which has expressed their interest to use DAS in other 
-scientific domains.
-
-\section{Summary}
-In this paper we have presented a new data aggregation service (DAS) 
-developed for the CMS High-Energy experiment at LHC, CERN, Geneva, Switzerland. 
-We have developed a prototype for DAS, designed to provide 
-caching and aggregation layers on top of the existing relational and 
-non-relational data-services in a close to real time fashion. The data 
-is retrieved on demand, with pre-fetching of common queries, 
-determined from an extensive analytics database, also possible. 
-The DAS system is currently being commissioned for production use by CMS.
-
-Data taking at CMS began in December 2009, and, alongside simulation 
-data, we expect to accumulate petabytes of data yearly. This will have 
-of the order of a terabyte of associated meta-data that will need to be 
-easily queried by physicists. From the performance studies of DAS presented 
-in this paper we expect that the system will be able to sustain such load 
-and will provide generic “data-discovery” service for CMS experiment for 
-years to come.
-
-\section{Acknowledgments}
-
-This work was supported by the National Science Foundation, 
-contract No. PHY-0757894, and 
-Department of Energy of the United States of America.
-
-%Fermilab is operated by Fermi Research Alliance, LLC under Contract
-%No. DE-AC02-07CH11359 with the United States Department of Energy.
+%% \section{}
+%% \label{}
 
@@
 
-%\bibliographystyle{elsarticle-num}
-%\bibliography{<your-bib-database>}
+\bibliographystyle{elsarticle-num}
+\bibliography{<your-bib-database>}
 
@@
 
-\section*{References}
-\begin{thebibliography}{00}
-\bibitem{DBXplorer}
-S. Agrawal, S. Chaudhuri, G. Das,
-``DBXplorer: A System for Keyword-Based Search over Relational Databases'',
-ICDE 2002, pp. 5-16.
+\end{document}
 
-\bibitem{QueryAnswer}
-G. Koutrika, A. Simitsis, Y. E. Ioannidis,
-``Pr\'{e}cis: The Essence of a Query Answer'',
-ICDE 2006, pp. 69-78.
-
-\bibitem{DBS-QL} 
-V. Kuznetsov, D. Riley, A. Afaq, V. Sekhri, Y. Guo, L. Lueking,
-``The CMS DBS Query Language'', CHEP 2009.
-
-%\bibitem{AMI}
-%Altas AMI web portal,
-%http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html
-
-\bibitem{FedDB}
-L. Haas, E. Lin,
-``IBM Federated Database Technology'', \\
-http://www.ibm.com/developerworks/data/library/techarticle/0203haas/0203haas.html
-
-\bibitem{CMS} 
-R. Adolphi et al., 
-``The CMS experiment at the CERN LHC'',
-JINST 0803, S08004 (2008).
-
-\bibitem{CMSDataModel} 
-C. Grandi, D.Stickland, L.Taylor et al.,
-``The CMS Computing Model'',
-CERN-LHCC-2004-035/G-083 (2004).
-%A. Fanfani et. al., 
-%``Distributed Analysis in CMS'', 
-%to be published in Journal of Grid Computing.
-
-\bibitem{iRODS}
-The Integrated Rule-Oriented Data System,
-http://www.irods.org/
-
-\bibitem{OpenArchive}
-Open Archives Initiative,
-http://www.openarchives.org/
-
-\bibitem{DBS} 
-A. Afaq, et. al.,
-``The CMS Dataset Bookkeeping Service'', 
-J. Phys. Conf. Ser, 119, 072001 (2008).
-
-\bibitem{DBS07} 
-A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
-``A multi-dimensional view on information retrieval of CMS data'',
-J. Phys. Conf. Ser, 119, 072013 (2008).
-
-\bibitem{DD} https://cmsweb.cern.ch/dbs\_discovery
-
-\bibitem{Arms}
-C. R. Arms, W. Y. Arms,
- ``Mixed Content and Mixed Metadata 
-Information Discovery in a Messy World'',
-chapter from ``Metadata in Practice'', ALA Editions, 2004.
-
-\bibitem{MySQL}
-http://www.mysql.com/
-
-\bibitem{CouchDB}
-http://couchdb.apache.org/
-
-\bibitem{MongoDB}
-http://www.mongodb.org/
-
-\bibitem{REST}
-Fielding, Roy Thomas ``Architectural Styles and the Design of 
-Network-based Software Architectures'', Doctoral dissertation, 2000,
-University of California, Irvine
-
-\bibitem{Python} Python programming language, http://www.python.org
-
-\bibitem{PhEDEx}
-%Rehn et. al.,
-%``PhEDEx high-throughput data transfer management system'', CHEP 2006;
-%R. Egeland et al., 
-%``Data transfer infrastructure for CMS data taking'', 
-%Proceedings of Science, PoS (ACAT08) 033 (2008);
-L. Tuura et al., 
-``Scaling CMS data transfer system for LHC start-up'', 
-J. Phys. Conf. Ser, 119, 072030 (2008).
-
-%\bibitem{RunSummary}
-%https://cmswbm.web.cern.ch/cmswbm/cmsdb/servlet/RunSummary
-
-%\bibitem{SiteDB}
-%https://cmsweb.cern.ch/sitedb/
-
-%\bibitem{LumiDB}
-%https://twiki.cern.ch/twiki/bin/view/CMS/CMS-DMWM-DBS-Luminosity
-
-%\bibitem{DQ}
-%Data-Quality DB reference
-
-%\bibitem{Overview}
-%https://cmsweb.cern.ch/overview/
-
-%\bibitem{Dashboard}
-%J. Andreeva, et. al,
-%``Experiment Dashboard – The Monitoring System for the LHC Experiments'',
-%GMW’07, June 25, 2007, Monterey, California, USA.
-
-%\bibitem{CRAFT09}
-%https://twiki.cern.ch/twiki/bin/view/CMS/CRAFT09AnalysisInfo
-
-%\bibitem{DBSearch}
-%D. Konopnicki, O. Shmueli,
-%``Database-Inspired Search'', 
-%Proc. of the 31st VLDB Conference, 2005.
-
-\bibitem{DRSG}
-http://drsg.cac.cornell.edu/
-\end{thebibliography}
-\end{document}
+%%
+%% End of file `procs-template.tex'.

### 2010-05-04 test/utils_t.py
Added _delete method to dotdict
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@2079 4525493e-7705-40b1-a816-d608a930855b

@@ -94,6 +94,12 @@ class testUtils(unittest.TestCase):
         result = ddict._get('a.M.Z')
         self.assertEqual(expect, result)
 
+        res = {'a': {'b': {'c':1, 'd':2}}}
+        ddict = dotdict(res)
+        expect = {'a': {'b': {'c':1}}}
+        ddict._delete('a.b.d')
+        self.assertEqual(expect, ddict)
+
     def test_merge_dict(self):
         """Test merge_dict"""
         dict1  = {'block':{'name':'AAA', 'b':{'c':1}, 'size':2}, 'das':{'system':'dbs'}}

### 2010-05-04 doc/das_paper/paper.tex
Added use case into Section 4
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@2081 4525493e-7705-40b1-a816-d608a930855b

@@
 retrieval and aggregation are triggered dynamically by user queries. 
-Below we discuss in details the DAS architecture.
+Imaging, a user opens a browser and places a query, e.g.
+{\it file block=abc}. The DAS system responds with aggregated infromation about files
+for provided set of conditions from all CMS data-services. Information
+is either retrieved from the DAS cache or fetched on-demand from relevant
+data-providers. Such system provides a uniform access to all meta-data in CMS 
+via simple, intuitive query interface. It neither enforces users to learn 
+the structure of CMS data-services
+%, their intefaces, data-formats and access policies
+or requires data-providers to agree on common data representation
+and requirement to have central meta-data repository.
+
+%Below we discuss in details the DAS architecture.
 
@@
 \section{Future work}
-The DAS system was desgined for CMS HEP experiment, while its architecture is
-transparent to consumed meta-data, data formats and participating data-providers. 
-In the current year we plan to put DAS into production and confirm its scalability, 
-transparency and durability for various CMS data-services. Our goal is to sustain 
-in DAS 1TB of meta-data per year. The DISCOVER Research Service Group at Cornell 
-University \cite{DRSG} is expressed its interests to apply DAS for other 
-scientific domains and turn it into generic open-source project.
+In the near future we plan to put DAS into production and confirm its scalability, 
+transparency and durability for various CMS data-services. 
+Our interests are plug-able interface for participating data-providers;
+analytics framework and pre-fetch strategies as well as horizontal scaling of the
+system. Our goal is to sustain 1TB of meta-data per year. 
+In a long term, we would like to use DAS across various
+data-services in HEP community and other disciplines. Even though the DAS was developed
+for CMS HEP experiment, its architecture is transparent to participating 
+data-providers and meta-data content. Therefore the generalization of DAS is feasible.
+Such work has began under the umbrella of DISCOVER Research Service Group at 
+Cornell University \cite{DRSG} who expressed their interest to use DAS in other 
+scientific domains.
 

### 2010-11-16 doc/chep_paper/das_chep2010_proceedings.tex
Files for DAS CHEP paper, fixes #663
Signed-off-by: Valentin Kuznetsov <vkuznet@gmail.com>


git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@10888 4525493e-7705-40b1-a816-d608a930855b

\documentclass[a4paper]{jpconf}
\usepackage{graphicx}

\begin{document}
\title{Data Aggregation System - a system for information retrieval on demand over relational and non-relational distributed data sources}
\author{G Ball$^1$, V Kuznetsov$^2$, D Evans$^3$ and S Metson$^4$}

\address{$^1$ Imperial College London, London, UK}
\address{$^2$ Cornell University, Ithaca, New York, USA}
\address{$^3$ Fermilab, Batavia, Illinois, USA}
\address{$^4$ Bristol University, Bristol, UK}

\ead{gordon.ball@cern.ch, vkuznet@gmail.com}

\begin{abstract}
We present the Data Aggregation System, a system for information retrieval and aggregation from heterogenous sources of relational and non-relational data for the Compact Muon Solenoid experiment on the CERN Large Hadron Collider. The experiment currently has a number of organically-developed data sources, including front-ends to a number of different relational databases and non-database data services which do not share common data structures or APIs, and cannot at this stage be readily converged. DAS provides an single interface for querying all these services, a caching layer to speed up access to expensive underlying calls and the ability to merge records from different data services pertaining to a single primary key.
\end{abstract}

\section{Introduction}

In 2010, data taking at the CMS experiment began in earnest, with approximately $42pb^{-1}$ of proton-proton collision data recorded at the time of writing. At full data-taking rates, this is expected to produce around 5PiB of raw and reconstructed data per year. Associated with this is around 1TiB per year of metadata, such as data locations, dataset descriptions and machine conditions.

Each of these metadata aspects are stored into a specialised system, using a variety of different technologies. Each provides specialised interfaces to query the specific metadata stored. In the pre-data-taking epoch, this situation was satisfactory as most systems were used in isolation and only then by experts. However, in the data-taking era users regularly need to perform queries across multiple data services (for example, to look up the recorded luminosity, machine conditions and software configuration for a particular dataset) which prove very expensive to serve under the existing model. DAS simplifies data look-up to end users by providing a single point of access and provides a caching layer to data services.

Although designed in the context of the CMS experiment, DAS itself does not make any experiment-specific assumptions about the data services it handles and is generalisable to other similar situations~\cite{drsg}.

\section{Architecture \& Implementation}

\begin{figure}
\begin{center}
\includegraphics[width=0.6\textwidth]{DAS_Cache_and_Analytics}
\end{center}
\caption{\label{dasarch}DAS architecture, showing the relationships between \emph{web server}, \emph{cache server}, \emph{analytics}, \emph{data services} and MongoDB (\emph{cache} and \emph{merge}).}
\end{figure}

DAS is designed as an additional layer on top of a heterogenous ecosystem of existing CMS data services.

The architecture of DAS was designed as a series of independent components, which can scale from all running on a single node to multiple nodes and duplicate components, as necessary (Fig.~\ref{dasarch}).

The \emph{web server} handles user sessions, whether they originate from a browser or automated scripts. Queries made here are passed onto the \emph{cache server} for processing, during which time the \emph{web server} periodically polls for the current status (displayed to the user by AJAX) until the request is completed. An MD5 hash of the input query is used to identify the correct output (and hence multiple users making identical queries will only result in a single worker thread doing the processing). Methods are provided here for output in machine-readable formats as well as human-readable formatting.

The \emph{cache server} consists of a pool of worker threads which handle the DAS queries received from the web front-end. Each query is handled entirely by a single thread. A detailed description of the workflow is given in \cite{das2010}. The worker thread first parses a text-based query from the DAS Query Language (Sec.~\ref{dasql}) into an internal representation. The query is then analysed to identify the set of data service APIs which are relevant. The \emph{cache} is checked for items of data matching the query, or data that would form a superset of the query. If any data cannot be found, the relevant service (Sec.~\ref{services}) is called to fetch and transform the data, and insert it into the raw cache (Sec.~\ref{cachemerge}). Once all the data is in the raw cache, records sharing common keys are merged together into a single document, which is stored in the merge cache. If any additional data processing has been specified in the query, it is performed on the merged document before being returned.

The \emph{cache} consists of one or more MongoDB~\cite{mongodb} shards. MongoDB is a document store which natively stores the JSON documents DAS uses as its internal representation. This is used both for the primary record stores (the \emph{raw} and \emph{merged} caches) and for the ancillary databases required by DAS, such as storing the server logs, analytics data and mapping between keys. MongoDB also provides GridFS~\cite{gridfs}, which DAS uses to store otherwise oversize documents, typically the result of merging large numbers of raw documents together.

The \emph{analytics server} provides a facility for scheduling regular tasks. The rest of DAS operates only when triggered by user input, and although some clean-up is performed at run-time (such as discarded expired records) it is also necessary to have a facility for running asynchronous operations. Analytics consists of a task scheduler and a pool of workers which execute the tasks. This is done independently of the \emph{cache server}, so heavy analytics tasks will not interfere with end-user use.

DAS assumes that the data in the cache can be recreated from the original data sources at any time, and thus it can function with a limited amount of space by deleting old data, nor does it require backup of this space. The design assumes the total number of records to be $O(5M)$, and MongoDB is capable of effectively functioning with collections of this size. Furthermore,  only a fraction would be expected to be resident in the cache at any one time.

DAS is designed as an entirely read-only system, with no ability to write data back to the underlying services, which simplifies the requirements for authentication and validation that this would otherwise require.

A random subset of results retrieved by DAS are examined to learn which data members are returned by which underlying queries, so that the web interface can suggest which input DAS keys will find the desired result.

The non-MongoDB parts of DAS are entirely written in python. Using python does come with a performance penalty but it allows for a fast development cycle, access to a number of useful libraries and is consistent with other CMS web projects.

\subsection{\label{dasql}DAS Query Language}

DAS queries are made in a custom text-based language. Work originally centred on an extended version of the query language already developed for the CMS Dataset Bookkeeping Service~\cite{dbsql} but the syntax was found unsuitable. The DAS syntax broadly resembles using pipes and commands in a UNIX shell, albeit with entirely dissimilar implementation.

Queries are parsed using the PLY~\cite{ply} parser for tokenising and lexical analysis, with a caching layer for previously parsed queries. The query is internally represented as a python dictionary containing the conditions and any additional operations.

A query is structured:

\begin{verbatim}
conditions | filters | aggregators OR map-reduce 
\end{verbatim}

To make a query, a user must know which \emph{DAS key} describes the data they are seeking. DAS keys describe a single logical object, descriptions of which may exist in multiple data services. Examples of DAS keys used for CMS include \textbf{dataset}, \textbf{run\_number} and \textbf{person}. In most cases these are the same names used by the data services, but DAS hides small differences in data service terminologies (eg, \textbf{run\_number} is variously called \textbf{runNumber} and \textbf{Run} by different services).

The \emph{conditions} consist of one or more DAS keys, optionally followed by an operator and operand. Examples of queries include:

\begin{verbatim}
city=Geneva
dataset site=institute.ac.uk
run in [123, 456]
\end{verbatim}

The conditions determine the total data for the query, but since this may still be a significant volume of data (or contain a significant number of fields of no interest to the user), further operations can be performed on the data.

\emph{Filters} are commands that either eliminate complete data records or prune data members within a record. At this time only two are provided; \emph{grep} works similarly to the UNIX command of the same name, filtering all but the given data members from the output, and \emph{unique} eliminates duplicate records. This is implemented by converting the filter conditions into MongoDB query format.

\emph{Aggregators} are functions which run over all the selected (and filtered, if applicable) records, summarising their contents. The basic aggregators available are simple numerical functions like \emph{sum}, \emph{count} and \emph{avg}, which perform the named aggregation over all data members of a given name. Aggregation is implemented using python coroutines.

An example of a ``complete'' query might be:

\begin{verbatim}
block dataset=/a/b/c* | grep block.size | sum(block.size), max(block.size)
\end{verbatim}

The set of available filters and aggregators is likely to grow in accordance with user needs. For more complex use-cases, it is possible to run a javascript map-reduce function on the MongoDB cache server. DAS QL does not allow functions to be specified, so such functions are written in advance and stored in the cache, and their execution requested with an alias.

\subsection{\label{services}Services}

Almost any source of information can be incorporated in DAS, such as an SQL database cursor, a command called in a subshell or an HTML page scraper, but in practice most data services used by DAS are APIs accessed over HTTP and returning data in JSON or XML format.

Each data service is described in a mapping document (supplied in YAML~\cite{yaml} format) which describes the functions available, the mapping from input DAS keys to actual arguments, and finally a mapping from the returned keys back to DAS keys. For a service accessed over HTTP and using a standard format, adding a new service only requires this mapping.

For more complex cases, such as those for which authentication is required, or for different source types a plugin must be written which translates DAS requests into the necessary calls, and then transforms the output into DAS JSON documents.

The mapping document may also specify presentation information which is used to produce more human-friendly output by mapping key names to descriptions.

When a query for which no exact match exists in the cache is received, the set of input DAS keys is considered to determine which services and their respective APIs to invoke. If the query consists only of keys with associated conditions, the relevant APIs are all those for which the accepted DAS keys are either equal or a superset of the input keys. If no API matches the set of input queries, it is decomposed into multiple sub-queries.

Before an API query is made, a check is made for existence of a superset query already in the cache, if wildcarding is supported by the service.

\subsection{\label{cachemerge}Caching \& Merging}

Data in DAS is principally stored in two seperate collections; the raw and merged caches.

Two types of data are stored in the raw cache; the documents returned by the respective APIs and documents describing the current status of queries. The latter serves as the central record for a query, holding the original query, description of the APIs called, status of the processing and a common key with all the results documents.

Once all the data for a query is available in the raw cache, a merged record is created for each primary DAS key. Merged documents may exceed the maximum size possible for MongoDB records, in which case they are stored using GridFS and a reference to them stored in the merged cache instead.

\subsection{\label{analytics}Analytics}

The DAS analytics system is a daemon that schedules and executes small tasks that access the DAS document store. This can be used to perform prosaic maintainance tasks, such as cleaning out expired data or pruning still-valid data if space becomes limited, but the main function is to analyse the queries received by DAS to provide information to developers and to perform automated cache optimisation.

The caching aspect of DAS is meant to reduce the latency users experience when performing complex, cross-service queries, but since the data in question usually has TTL of less than an hour users will relatively often find that the data they request is not in the cache and must be fetched on demand.

The analytics system hence tries to identify the most popular user queries and ensure that they are always resident in the cache and thus available for real-time retrieval, whatever the actual fetch latency may have been.

It is vital to the success of DAS, both in terms of user experience and system load, that the majority of all queries can be served from the cache. To ensure this, we must have a good pre-fetching strategy that maintains itself based on the actions of users.

Tasks that run on the analytics system may be divided into query analysis and cache populators. The former run infrequently, performing expensive (both in terms of processing and MongoDB access) analysis operations. Having identified the set of queries they wish to maintain, they spawn simpler populator tasks that schedule themselves to pre-empt each expiry of the specified data and perform an update, until the next time the analyser runs and the priorities may be altered.

A simple example of an analyser is designated \emph{ValueHotspot}, and is designed to run every few hours, identifying popular arguments to a given DAS key. Upon each invocation, the set of queries issued since the last invocation which reference the specified key is identified, and a count of each argument made. This is then combined with summary documents from previous invocations (containing the same information for those epochs) to calculate a 30-day moving average. The top $15\%$ most-used arguments are selected, which based on past results represent about $50\%$ of all queries are then kept in the cache until the analyser next runs.

There are a large number of possible future analysis strategies (such as the time structure throughout a week of certain types of queries, automatically pre-fetching information about new datasets as they are released or considering a metric based on total preventible latency rather than total numbers of calls), but some experience with actual users is needed to determine what will actually prove useful.

\section{Benchmarks}

\begin{figure}
\begin{center}
\includegraphics[width=0.6\textwidth]{das_vs_mongo_fully_random}
\end{center}
\caption{\label{dasbench}Benchmark showing the average latency for fetching a single randomly-selected document from a wildcard query, with a given number of parallel clients. The cache contains $5\times10^7$ records. The red line shows the performance of the complete DAS stack, and the blue line the time spent by MongoDB.}
\end{figure}

To benchmark DAS performance, we used a 64-bit linux node with 8 cores (each 2.33GhZ) and 16GiB of RAM, which we would expect to be typical of the hardware DAS would use in production. All DAS systems and MongoDB share this node.

Testing was performed with the cache pre-populated with approximately $5\times10^7$ records (100 times the expected volume), consisting of \emph{block} records from the CMS PhEDEx~\cite{phedex} and DBS systems. We test with up to 500 clients making requests in parallel. Each client requests one random record from the entire result set (which will test the time required for DAS to process the query but largely exclude the time required to serialise a large output set), using queries of the form \verb block=/[a-z]* (past experience with the DBS system shows wildcard queries are very common, even when the user knows the complete string in question). Fig.~\ref{dasbench} shows the benchmark results.

\section{Future Work}

We are currently in process of deploying DAS in the production environment for beta testing. The main development focus at this time is to increase the scope of possible queries, both by adding additional services and increasing the number of APIs supported for existing services. Support currently exists to some degree for the CMS services DBS, PhEDEx, SiteDB, Tier-0, LumiDB, RunRegistry, Dashboard and Overview. Most of these however only have a few APIs described by the DAS mapping, and to ensure usefulness to users we need to support as many of the queries supported by their original interfaces as possible.

\ack

This work was supported by the National Science Foundation, contract No. PHY-0757894, and the Department of Energy of the United States of America.

This work was supported by the Science and Technology Facilities Council, UK.

\appendix

\section*{References}
\begin{thebibliography}{9}
\bibitem{drsg} DISCOVER RSG {\it http://drsg.cac.cornell.edu/content/drsg-pilot-projects}
\bibitem{das2010} Kuznetsov V, Evans D and Metson S 2010 The CMS data aggregation system {\it Procedia Comp. Sci.} {\bf 1} 1529-37
\bibitem{mongodb} MongoDB, a scalable, high-performance, open-source, document-orientated database {\it http://www.mongodb.org/}
\bibitem{gridfs} MongoDB GridFS {\it http://www.mongodb.org/display/DOCS/GridFS}
\bibitem{dbsql} Kuznetsov V, Riley D, Afaq A, Sekhri V, Guo Y and Lueking L 2009 The CMS DBS query language {\it CHEP}
\bibitem{ply} Python Lex-Yacc {\it http://www.dabeaz.com/ply/}
\bibitem{yaml} YAML ain't markup language {\it http://www.yaml.org/}
\bibitem{phedex} Rehn J, Barrass T, Bonacorsi D, Hernandez J, Semeniouk I, Tuura L and Wu Y 2006 PhEDEx high-throughput data transfer management system {\it CHEP}
\end{thebibliography}
\end{document}

### 2011-01-03 doc/chep_paper/das_chep2010_proceedings.tex
Update DAS CHEP paper; re-factor UI to make external links; fixes #880
Signed-off-by: Valentin Kuznetsov <vkuznet@gmail.com>


git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@11242 4525493e-7705-40b1-a816-d608a930855b

@@
 \title{Data Aggregation System - a system for information retrieval on demand over relational and non-relational distributed data sources}
-\author{G Ball$^1$, V Kuznetsov$^2$, D Evans$^3$ and S Metson$^4$}
+\author{G. Ball$^1$, V. Kuznetsov$^2$, D. Evans$^3$ and S. Metson$^4$}
 
@@
 We present the Data Aggregation System, a system for information retrieval and aggregation from heterogenous sources of relational and non-relational data for the Compact Muon Solenoid experiment on the CERN Large Hadron Collider.
-The experiment currently has a number of organically-developed data sources, including front-ends to a number of different relational databases and non-database data services which do not share common data structures or APIs, and cannot at this stage be readily converged.
-DAS provides an single interface for querying all these services, a caching layer to speed up access to expensive underlying calls and the ability to merge records from different data services pertaining to a single primary key.
+The experiment currently has a number of organically-developed data sources, including front-ends to a number of different relational databases and non-database data services which do not share common data structures or APIs {\it (Application Programming Interfaces)}, and cannot at this stage be readily converged.
+DAS provides a single interface for querying all these services, a caching layer to speed up access to expensive underlying calls and the ability to merge records from different data services pertaining to a single primary key.
 \end{abstract}
@@
 In 2010, data taking at the CMS experiment began in earnest, with approximately $42pb^{-1}$ of proton-proton collision data recorded at the time of writing.
-At full data-taking rates, this is expected to produce around 5PiB of raw and reconstructed data per year.
-Associated with this is around 1TiB per year of metadata, such as data locations, dataset descriptions and machine conditions.
+At full data-taking rates, this is expected to produce around 5PB of raw and reconstructed data per year.
+Associated with this is around 1TB per year of metadata, such as data locations, dataset descriptions and machine conditions.
 
@@
 The architecture of DAS was designed as a series of independent components, which can scale from all running on a single node to multiple nodes and duplicate components, as necessary (Fig.~\ref{dasarch}).
+An initial description of the architectur, choice of technology and comparison with similar systems was given in \cite{das2010}.
+Here we present further development of the query language (Sec.~\ref{dasql}), analytics (Sec.~\ref{analytics}) and initial benchmark results (Sec.~\ref{benchmarks}).
 
 The \emph{web server} handles user sessions, whether they originate from a browser or automated scripts.
-Queries made here are passed onto the \emph{cache server} for processing, during which time the \emph{web server} periodically polls for the current status (displayed to the user by AJAX) until the request is completed.
+Queries made here are passed onto the \emph{cache server} for processing, during which time the \emph{web server} periodically polls for the current status and displays it to the user by AJAX {\it (Asynchronous Javascript and XML)}, until the request is completed.
 An MD5 hash of the input query is used to identify the correct output (and hence multiple users making identical queries will only result in a single worker thread doing the processing).
@@
 Each query is handled entirely by a single thread.
-A detailed description of the workflow is given in \cite{das2010}.
 The worker thread first parses a text-based query from the DAS Query Language (Sec.~\ref{dasql}) into an internal representation.
@@
 The \emph{cache} consists of one or more MongoDB~\cite{mongodb} shards.
-MongoDB is a document store which natively stores the JSON documents DAS uses as its internal representation.
+MongoDB is a document store which natively stores the JSON {\it (Javascript Object Notation)} documents DAS uses as its internal representation.
 This is used both for the primary record stores (the \emph{raw} and \emph{merged} caches) and for the ancillary databases required by DAS, such as storing the server logs, analytics data and mapping between keys.
 MongoDB also provides GridFS~\cite{gridfs}, which DAS uses to store otherwise oversize documents, typically the result of merging large numbers of raw documents together.
+The choice of a document store instead of a traditional relational database was driven by the need to be able to store and subsequently query individual members of the deeply-nested data structures that DAS handles, the structure of which are not known until run-time.
+We tested the CouchDB~\cite{couchdb} and MongoDB document stores for this purpose; we found the latter to have better performance in our use-case, better scalability across multiple nodes and a query model more compatible with our own.
 
@@
 
-\section{Benchmarks}
+\section{\label{benchmarks}Benchmarks}
 
@@
 The cache contains $5\times10^7$ records.
-The red line shows the performance of the complete DAS stack, and the blue line the time spent by MongoDB.}
+The red triangles show the performance of the complete DAS stack, and the blue stars the time spent by MongoDB.}
 \end{figure}
 
-To benchmark DAS performance, we used a 64-bit linux node with 8 cores (each 2.33GhZ) and 16GiB of RAM, which we would expect to be typical of the hardware DAS would use in production.
+To benchmark DAS performance, we used a 64-bit linux node with 8 cores (each 2.33GhZ) and 16GB of RAM, which we would expect to be typical of the hardware DAS would use in production.
 All DAS systems and MongoDB share this node.
@@
 \bibitem{gridfs} MongoDB GridFS {\it http://www.mongodb.org/display/DOCS/GridFS}
+\bibitem{couchdb} Apache CouchDB {\it http://couchdb.apache.org/}
 \bibitem{dbsql} Kuznetsov V, Riley D, Afaq A, Sekhri V, Guo Y and Lueking L 2009 The CMS DBS query language {\it CHEP}

### 2011-01-03 src/python/DAS/core/das_mapping_db.py

@@ -39,8 +39,11 @@ class DASMapping(object):
         
         self.create_db()
 
-        self.notationcache = {}
+        self.presentationcache = {}    # to be filled at run time
+        self.reverse_presentation = {} # to be filled at run time
+        self.notationcache = {}        # to be filled at run time
         self.init_notationcache()
+        self.init_presentationcache()
 
     # ===============
     # Management APIs
@@ -56,6 +59,22 @@ class DASMapping(object):
                     self.notationcache[key] += [ (row['api'], row['map']) ]
                 else:
                     self.notationcache[key] = [ (row['api'], row['map']) ]
+
+    def init_presentationcache(self):
+        """
+        Initialize presentation cache by reading presentation map.
+        """
+        query = {'presentation':{'$ne':None}}
+        data  = self.col.find_one(query)
+        if  data:
+            self.presentationcache = data['presentation']
+            for daskey, uilist in self.presentationcache.items():
+                for row in uilist:
+                    link = None
+                    if  row.has_key('link'):
+                        link = row['link']
+                    self.reverse_presentation[row['ui']] = \
+                            (daskey, row['das'], link)
 
     def create_db(self):
         """
@@ -463,9 +482,13 @@ class DASMapping(object):
         Return web UI presentation keys for provided DAS keyword.
         For example once asked for block we present block.name, block.size, etc.
         """
-        query = {'presentation':{'$ne':None}}
-        for row in self.col.find(query):
-            data = row['presentation']
-            if  data.has_key(daskey):
-                return data[daskey]
+        if  self.presentationcache.has_key(daskey):
+            return self.presentationcache[daskey]
         return [daskey]
+    def daskey_from_presentation(self, uikey):
+        """
+        Return triplet (DAS key, DAS access key, link)
+        associated with provided UI key.
+        """
+        if  self.reverse_presentation.has_key(uikey):
+            return self.reverse_presentation[uikey]

### 2011-01-03 src/python/DAS/web/das_web.py

@@ -50,7 +50,7 @@ DAS_WEB_INPUTS = ['input', 'idx', 'limit', 'show', 'collection', 'name',
 
 def make_links(key, values):
     """
-    Make new link out of provided key/value pair.
+    Make new link for provided key/values.
     """
     for val in values:
         uinput = urllib.quote('%s=%s' % (key, val))
@@ -58,11 +58,14 @@ def make_links(key, values):
         url = """<a href="%s">%s</a>""" % (quote(url), val)
         yield url
 
-def key_values(gen):
+def adjust_values(func, gen):
     """
-    Helper function to group by values for identical key.
-    It can be extended further to provide links for specific values
-    of known keys, e.g. make links for dataset, run, etc.
+    Helper function to adjust values in UI.
+    It groups values for identical key, make links for provided mapped function,
+    represent "Number of" keys as integers and represents size values in GB format.
+    The mapped function is the one from das_mapping_db which convert
+    UI key into triplet of das key, das access key and link, see 
+    das_mapping_db:daskey_from_presentation
     """
     rdict = {}
     for uikey, value in [k for k, g in groupby(gen)]:
@@ -75,22 +78,12 @@ def key_values(gen):
             rdict[uikey] = [val]
     page = ""
     for key, val in rdict.items():
-        if  key == 'CMSName':
-            value = make_links('site', val)
-        elif  key == 'Primary dataset':
-            value = make_links('primary_dataset', val)
-        elif  key == 'Run number':
-            value = make_links('run', val)
-        elif  key == 'Dataset':
-            value = make_links('dataset', val)
-        elif  key == 'Block name':
-            value = make_links('block', val)
-        elif  key == 'File name':
-            value = make_links('file', val)
-        elif  key == 'Block size' or key == 'File size':
+        daskey, _, link = func(key)
+        if  daskey and link:
+            value = make_links(daskey, val)
+        elif  key.find('size') != -1:
             value = [size_format(val[-1])]
-        elif  key == 'Number of events' or key == 'Number of files' or\
-            key == 'Number of triggers':
+        elif  key.find('Number of ') != -1:
             try:
                 vvv = val[-1].split('.')
                 if  vvv[-1] == '0':
@@ -639,7 +632,9 @@ class DASWebService(DASWebManager):
                     pkey  = row['das']['primary_key']
                     page += '<b>DAS key:</b> %s<br />' % pkey.split('.')[0]
             gen   = self.convert2ui(row)
[…]

### 2011-03-16 doc/chep_paper/das_chep2010_proceedings.tex
Update DAS CHEP paper with comments provided by referees
Signed-off-by: Valentin Kuznetsov <vkuznet@gmail.com>


git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@11969 4525493e-7705-40b1-a816-d608a930855b

@@
 The architecture of DAS was designed as a series of independent components, which can scale from all running on a single node to multiple nodes and duplicate components, as necessary (Fig.~\ref{dasarch}).
-An initial description of the architectur, choice of technology and comparison with similar systems was given in \cite{das2010}.
+An initial description of the architecture, choice of technology and comparison with similar systems was given in \cite{das2010}.
 Here we present further development of the query language (Sec.~\ref{dasql}), analytics (Sec.~\ref{analytics}) and initial benchmark results (Sec.~\ref{benchmarks}).
@@
 
-Testing was performed with the cache pre-populated with approximately $5\times10^7$ records (100 times the expected volume), consisting of \emph{block} records from the CMS PhEDEx~\cite{phedex} and DBS systems.
+%Testing was performed with the cache pre-populated with approximately $5\times10^7$ records (100 times the expected volume), consisting of \emph{block} records from the CMS PhEDEx~\cite{phedex} and DBS systems.
 We test with up to 500 clients making requests in parallel.
 Each client requests one random record from the entire result set (which will test the time required for DAS to process the query but largely exclude the time required to serialise a large output set), using queries of the form \verb block=/[a-z]* (past experience with the DBS system shows wildcard queries are very common, even when the user knows the complete string in question).
+Fig.~\ref{dasbench} shows the benchmark results.
+Testing was performed with the cache pre-populated with approximately 
+$5\times10^7$ records (100 times the expected volume), consisting 
+of \emph{block} records from the CMS PhEDEx~\cite{phedex} and DBS 
+systems. Each \emph{block} record is a JSON document with common keys, such as
+\emph{name, nevents, nfiles, replica}, etc.
+We test with up to 500 clients making requests in parallel. 
+Each client requests a random wildcard query, e.g. \verb block=/a* , and fetches one random
+record from the returned set (which will test the time required for DAS to parse the 
+query and access the cache but largely exclude the time required to serialise a large 
+output set). No additional operations were done during this step.
 Fig.~\ref{dasbench} shows the benchmark results.

### 2011-03-17 src/python/DAS/core/das_core.py
Use initial timeout as 5 sec with max limit of 30 sec; fixes #1032
Signed-off-by: Valentin Kuznetsov <vkuznet@gmail.com>


git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@11982 4525493e-7705-40b1-a816-d608a930855b

@@ -41,32 +41,22 @@ from DAS.utils.das_timer import das_timer, get_das_timer
 import DAS.utils.jsonwrapper as json
 import DAS.core.das_aggregators as das_aggregator
 
-def dasheader(system, query, api, url, ctime, expire):
+def dasheader(system, query, expire, api=None, url=None, ctime=None):
     """
     Return DAS header (dict) wrt DAS specifications, see
     https://twiki.cern.ch/twiki/bin/view/CMS/
         DMWMDataAggregationService#DAS_data_service_compliance
     """
-    systems = []
-    if  system:
-        if  system == 'das': # init for das record
-            systems = []
-        else:
-            systems = [system]
-    apis = []
-    if  api:
-        apis = [api]
-    urls = []
-    if  url:
-        urls = [url]
-    ctimes = []
-    if  ctime:
-        ctimes = [ctime]
     query   = encode_mongo_query(query)
-    dasdict = dict(system=systems, timestamp=time.time(),
-                url=urls, ctime=ctimes, qhash=genkey(query), 
-                expire=expire_timestamp(expire), urn=apis,
-                api=apis, status="requested")
+    if  not api:
+        dasdict = dict(system=[system], timestamp=time.time(),
+                    qhash=genkey(query), expire=expire_timestamp(expire),
+                    status="requested")
+    else:
+        dasdict = dict(system=[system], timestamp=time.time(),
+                    url=[url], ctime=[ctime], qhash=genkey(query), 
+                    expire=expire_timestamp(expire), urn=[api],
+                    api=[api], status="requested")
     return dict(das=dasdict)
 
 class DASCore(object):
@@ -395,11 +385,10 @@ class DASCore(object):
         params = self.mongoparser.params(query)
         services = params['services']
         self.logger.info('DASCore::call, services = %s' % services)
-        qhash = genkey(query)
         das_timer('das_record', self.verbose)
         # initial expire tstamp 1 day (long enough to be overwriten by data-srv)
         expire = expire_timestamp(time.time()+1*24*60*60)
-        header = dasheader("das", query, api="", url="", ctime=0, expire=expire)
+        header = dasheader("das", query, expire)
         header['lookup_keys'] = []
         self.rawcache.insert_query_record(query, header)
         das_timer('das_record', self.verbose)
@@ -416,12 +405,12 @@ class DASCore(object):
             traceback.print_exc()
             return 0
[…]

### 2011-03-17 src/python/DAS/core/das_mongocache.py

@@ -408,7 +408,8 @@ class DASMongocache(object):
         Find if cache has query whose specs are identical to provided query.
         """
         enc_query = encode_mongo_query(query)
-        cond = {'das.qhash': {"$exists":True}, "query.spec": enc_query['spec']}
+        cond = {'das.qhash': {"$exists":True}, "query.spec": enc_query['spec'],
+                        'das.system':'das'}
         return self.col.find_one(cond)
 
     def das_record(self, query):
@@ -424,42 +425,50 @@ class DASMongocache(object):
         """
         return self.col.find({'das_id': das_id})
 
-    def add_to_record(self, query, info):
+    def add_to_record(self, query, info, system=None):
         """
         Add to existing DAS record provided info
         """
         enc_query = encode_mongo_query(query)
-        self.col.update({'query': enc_query}, {'$set': info}, upsert=True)
+        if  system:
+            self.col.update({'query': enc_query, 'das.system':system}, 
+                            {'$set': info}, upsert=True)
+        else:
+            self.col.update({'query': enc_query}, 
+                            {'$set': info}, upsert=True)
 
-    def update_das_record(self, query, status, header=None):
+    def update_query_record(self, query, status, header=None):
         """
         Update DAS record for provided query.
         """
         enc_query = encode_mongo_query(query)
         if  header:
-            record = self.col.find_one({'query': enc_query})
-            if  header['das']['expire'] < record['das']['expire']:
+            system = header['das']['system']
+            dasrecord = self.col.find_one({'query': enc_query, 'das.system': 'das'})
+            sysrecord = self.col.find_one({'query': enc_query, 'das.system': system})
+            if  header['das']['expire'] < dasrecord['das']['expire']:
                 expire = header['das']['expire']
             else:
-                expire = record['das']['expire']
+                expire = dasrecord['das']['expire']
             api = header['das']['api']
             url = header['das']['url']
-            if  set(api) & set(record['das']['api']) == set(api) and \
-                set(url) & set(record['das']['url']) == set(url):
-                self.col.update({'_id':ObjectId(record['_id'])}, 
+            if  set(api) & set(sysrecord['das']['api']) == set(api) and \
+                set(url) & set(sysrecord['das']['url']) == set(url):
+                self.col.update({'_id':ObjectId(sysrecord['_id'])}, 
                      {'$set': {'das.expire':expire, 'das.status':status}})
             else:
[…]

### 2011-03-17 src/python/DAS/services/abstract_service.py

@@ -239,7 +239,7 @@ class DASAbstractService(object):
                 % self.name
         msg += ' query=%s, api=%s, args=%s' % (query, api, args)
         self.logger.debug(msg)
-        header  = dasheader(self.name, query, api, url, ctime, expire)
+        header  = dasheader(self.name, query, expire, api, url, ctime)
         header['lookup_keys'] = self.lookup_keys(api)
 
         # check that apicall record is present in analytics DB

### 2011-03-17 src/python/DAS/tools/das_client.py

@@ -90,7 +90,8 @@ def get_data(host, query, idx, limit, debug):
         pid = data
     else:
         pid = None
-    count = 3 # initial waiting time in seconds
+    count = 5  # initial waiting time in seconds
+    timeout = 30 # final waiting time in seconds
     while pid:
         params.update({'pid':data})
         encoded_data = urllib.urlencode(params, doseq=True)
@@ -108,10 +109,10 @@ def get_data(host, query, idx, limit, debug):
         else:
             pid = None
         time.sleep(count)
-        if  count < 60:
+        if  count < timeout:
             count *= 2
         else:
-            count = 60
+            count = timeout
     return data
 
 def main():

### 2011-03-17 src/python/DAS/web/das_web.py

@@ -186,7 +186,7 @@ class DASWebService(DASWebManager):
     """
     def __init__(self, config={}):
         DASWebManager.__init__(self, config)
-        self.next       = 3000 # initial next update status in miliseconds
+        self.next       = 5000 # initial next update status in miliseconds
         self.cachesrv   = config['cache_server_url']
         self.base       = config['url_base']
         logfile         = config['logfile']
@@ -1026,7 +1026,7 @@ class DASWebService(DASWebManager):
         """
         Place AJAX request to obtain status about given query
         """
-        limit = 60000 # 1 minute, max check status limit
+        limit = 30000 # 1 minute, max check status limit
         next  = int(next)
         if  next < limit and next*2 < limit:
             next *= 2

### 2011-03-17 src/python/DAS/web/das_web_srv.py

@@ -188,7 +188,7 @@ class DASWebService(DASWebManager):
     def __init__(self, config={}):
         DASWebManager.__init__(self, config)
         self.base    = config['url_base']
-        self.next    = 3000 # initial next update status in miliseconds
+        self.next    = 5000 # initial next update status in miliseconds
         self.engine  = config.get('engine', None)
         logfile      = config['logfile']
         loglevel     = config['loglevel']
@@ -675,7 +675,7 @@ class DASWebService(DASWebManager):
         """
         Place AJAX request to obtain status about given process pid.
         """
-        limit = 60000 # 1 minute, max check status limit
+        limit = 30000 # 1 minute, max check status limit
         next  = int(next)
         if  next < limit and next*2 < limit:
             next *= 2

### 2012-10-02 doc/thesis/latex/report_sept.tex
added draft of september report; detailed instructions of deployment on CERN VM and debian/ubuntu

\documentclass[a4paper,11pt,draft]{article}
\special{papersize=210mm,297mm}

% twocolumn
%\usepackage{fullpage} 
\usepackage[cm]{fullpage}
%\usepackage[margin=0.5in]{geometry}
% \usepackage{times}
\usepackage{color}
\usepackage{titlesec}

\bibliographystyle{alpha}

% usually first paragraph is NOT indented, so this is commented out: 
% \usepackage{indentfirst}

\begin{document}

% Article top matter
\title{Master thesis problem statement: Keyword search over heterogeneous sources} %\LaTeX is a macro for printing the Latex logo
\author{Vidmantas Zemleris}  %\texttt formats the text to a typewriter style font
\date{\today}  %\today is replaced with the current date

%\maketitle

%
 \centerline{\Large \bf Searching heterogeneous data-sources: Master thesis problem statement} %% Paper title
 \medskip

 \centerline{Vidmantas Zemleris, \today}  %% Author name(s)
 \medskip

\section{Introduction}
At large scientific collaborations like the CMS Experiment at CERN's LHC that includes more than 3000 collaborators data usually resides on a fair number of autonomous and heterogeneous proprietary systems each serving it's own purpose
\footnote{For instance, at CERN,  due to many reasons (e.g. research and need of freedom, politics of institutes involved) software projects usually evolve in independent fashion resulting in fair number of proprietary systems\cite{Koch00CERN}. Further high turnover makes it harder extending these systems}. As data stored on one system may be related to data residing on other systems%
	\footnote{For example, datasets containing physics events are registered at DBS, while the physical location of files is tracked by Phedex which also takes care of their transfers within the worldwide grid storage}%
, users are in need of a centralized and easy-to-use solution for locating and combining data from all these multiple services.

Using a highly structured language like SQL is problematic because users need to know not only the language but also where to find the information and also lots of technical details like schema. A data integration system based on simple structured queries is already in place. Still {\color{red}various improvements including support} for less restricted keyword queries and improvements to system's usability and performance still have to be researched.

% {\color{red}A remarkable opportunity we have is possibility to test ideas on a sample of ~3000 people working at CMS.}

\section{Case study: the CMS Experiment at CERN}
Users' information need may vary greatly depending on their role, however most of the time they are interested in locating full set of entities matching some selection criteria, e.g.:
%\footnote{in contrary to exploratory approach where  a subset of best matches is enough}
   \begin{itemize}
  		\item find all \textit{files} from \textit{dataset(s)} matching wild-card query each containing some of the 'interesting' \textit{runs} from a list provided (Release validation teams)
         \item find (all) \textit{datasets} related some specific physics phenomena\footnote{In case of dataset this data is present in filename or run} together with conditions describing how this data was recorded by detector or simulated which are present in separate autonomous system than the datasets (Physicists)
         % TODO:is this a good example? conditions is a separate query now...
         \item find all \textit{datasets} matching some pattern stored at a given \textit{site} (filtering on entities stored on separate services)
   \end{itemize}                		
% In such dynamic environment it is preferred to choose local-as-view approach for information integration instead of creating a global schemata\cite{Koch00CERN}.

For more use-cases of data retrieval at CMS Experiment see \cite{CMS_data08}.

%Following a similar approach a 
\subsection*{The Data Aggregation System}

The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it queries the data-sources, merges the results, and caches them for subsequent uses.

Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]) operators. The combined query results could be later 'pipped' for further filtering and aggregation (min, avg, etc), e.g.:

{\footnotesize 
\begin{verbatim}
dataset=*RelVal* | grep dataset.nevents >1000 | avg(dataset.size), median(dataset.size)
\end{verbatim}
}

The query above would return average and median datasets sizes  of ones containing  \textit{RelVal} in their name having more than 1000 events.

Queries could be run either from web browser or through  command line interface where the results could  be fed into another application (e.g. program doing physics analysis or automatic software release validation).

% \textbf{\color{red}Examples of more general uses: Digital libraries, Stock markets, [ Weather, Flight costs]}

\section{Problem statement}
\subsection{Ease of use: Query Language over the mediated schema}
	Even a very simple structured query language that also contain entity names over the mediated schema may seem {\color{red}hard} to learn, especially in the beginning.
   On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources). 
   So some sort of guiding helping user to build the query shall be useful.

	Further a minimum set of search predicates is imposed by APIs (mainly because of performance reasons) and user has to be at least informed what is he expected to provide.

Also to consider:
\begin{itemize}         
       \item input ambiguity, typos
       \item {\color{red} what about ranked search?}
\end{itemize}

\subsection{Performance}
\subsubsection{Handling distributed search efficiently}

\subsubsection{Other performance issues}
	     \begin{itemize}
	                		\item more intelligent caching needed:
	                			\begin{itemize}
	                			\item given an entity received from provider, determining if it is useful to cache for long-term
	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
	                			\item could make use of even older information warning user
	                			\item could also have different validity dates for certain fields. if certain field is not explicitly requested, even a very old cache could be used.
	                			\end{itemize}
	                		\item query rewriting then used with grep=smf then the same item is available as selection key. check how many instances.
	                		\item pre-fetching common (sub-)queries: determining manually and/or automatically
	                		\item scale testing - if we are storing lots of historical info. MongoDB is not so performant if DB cant fit in memory. One of well known solutions is installing SSD.
                        \item ? with current API's for aggregation service has to fetch ALL records not just the first page..
                        \item ??? do we have cases then multiple APIs of the same service could be called for the same data, can we choose just one???
                        \item ? showing partial results
        \end{itemize}

\section{Preliminary Literature review}

% Overview: keyword search is good for non-structured documents, it is not [as] effective with structured sources\cite{Levy96}, therefore the keyword search on relational databases approach is good then no specific APIs exist and it's there are no resources to do \textbf{a full integration} .

\textbf{\color{red}TODO: Overview}

\subsection{Keyword search over DBs}
The problem of Keyword search over Relational Databases (or also semi-structured sources like XML) has received a significant attention by the research community over the last decade. There exist many papers and a couple of quite comprehensive PhD. dissertations\cite{PhD_2011, PhD_2012} describing performance optimization details, etc.

There was a number of variations of the problem including returning of only Top-k results expecting that users only care about the quality of limited number of results but not completeness of results, ... \textbf{\color{red}describe this in details}

\subsubsection*{Ranking Query Templates based on keyword query}
A simple way to access relational database could  be through a set of predefined named query templates (SQL with selection parameters or operators still to be specified) exposed to a user as a Form that the user has to fill in.

\cite{forms_kws} proposes alternative approach for processing keyword queries over relational databases: given a keyword query, instead of returning database tuples one could rank query forms that best matches the query for user to choose the right one (if they are properly named this is fairly easy). The ranking is done by \textbf{\color{red}TODO}.

An interesting feature of this approach is that a Query Template is functionally similar to any autonomous web service (which given the parameters would in turn execute that query on its database).  In case of the Data Aggregation System, a user after entering a keyword query could be provided with a ranked list of structured queries (attribute=value) that could be processed given data source constraints (e.g. parameters required) and if needed he could refine his search (e.g. provide more parameters).

\subsubsection*{Semantic approach}

Researchers from \textit{ETHZ} have been investigating Generation of SQL for Business users in collaboration with \textit{Credit Suisse}.  To converting natural language queries (that in addition to keyword search could convey some semantic structure) into SQL statements they employed some natural language processing and used meta-data describing the schema and the domain (ontologies at multiple levels).

Even on a large data-warehouse of ~220GB data with a complex schema of 300 tables they reported fairly promising results that proves that if good meta-data is available, generating even quite complex SQL  (n-way joins with aggregations, etc) is quite feasible for computer making it 'much easier for business users to interactively explore highly-complex data warehouses' \cite[p.932]{ethz2012}.

\subsubsection*{Automatic mapping between distributed services}
{\color{red}an interesting approach but not for CERN...}

\subsubsection*{\color{red}Question Answering}

% See Appendix #1, for evaluation of they Demo system.

\subsection{Searching Heterogeneous web services}
{\color{red}There doesn't seem to be much of recent papers going towards this direction (except from specific search systems like flight fare comparisons), however these are based on structured arguments.}

In 1990s there were was much research on search based service integration systems, for instance \textit{Information Manifold} following Local-as-View approach and \textit{TSIMMIS} developed at Stanford following Global-as-View approach.

The \textit{Information Manifold} is quite resembling to DAS {\color{red},  so it could be useful checking it's papers. TODO: finish}

{\color{red}Performance: Answering Queries using Views}

\begin{verbatim}
Some other approaches could include:

- creating virtual documents offline by joining tables, for instantaneous search results by employing IR techniques
(Indexing Relational Database Content Offline for Efficient Keyword-Based Search, 2003)

- (Efficient Keyword Search Across Heterogeneous Relational Databases, 2007)
 : combines schema matching and structure discovery techniques to ﬁnd approximate foreign-key joins across heterogeneous databases
\end{verbatim}

\section{Possible approaches}

\subsection{Ease of use}
\begin{verbatim}
- javascript interpreter
    : field + description -- dropdown searcheable list

- if entered keyword query suggest a structured query, optionally with description in natural language
- select entity (e.g. dataset) --> give all available keys from forms, however there are still combinations which are enforced
\end{verbatim}
%(could use Natural Language Generation)

\subsection{Performance}
\subsubsection{Continuous view maintenance at Data providers with large DBs}
Use either \textit{materialized refresh fast views with query rewriting} (Oracle; completely transparent for proprietary apps)\cite{Oracle11}
 or some other continuous view maintenance tool (e.g. DBToaster\footnote{http://www.dbtoaster.org that is being developed at EPFL}) to improve performance of heavy queries containing joins and/or aggregations.

\subsection{Integrating distributed information efficiently}
Bloom-join (which could be quite transparent and implemented even on DB side [pure SQL is possible for MySQL, to check for Oracle]- take a query and bit-vector as parameter) , lazy pagination (and order required for aggregation) - this is not yet supported by any of the data-service APIs

%\newline
integration at DAS level. (at source DB could be more performant a little)

\subsection{More intelligent caching}
Some of the data entities instances changing very rarely, for example in DBS system old datasets would never change, while new ones are constantly added (still some of their attributes may change, in this example validity).

Cached copy may be shown before hand while up to date results could be retrieved on user's request.

\textbf{Decide if to cache or not...}

An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.

\begin{verbatim}
Also: Caching of intermediary results, not only the queries
\end{verbatim}

\section{Work status}
TODO

\begin{verbatim}
- obtained DB copy of biggest data provider DBS (currently 80 GB + 200 GB indexes)
- preliminary literature review (to be continued deeper + waiting for book arrival: Principles of Data Integration)

--- some analysis of DAS logs

Upcomming Work items:
- couple of fairly simple prototypes of UI/access patters for simpler DAS querying
- check on performance improvements after creation of materialized view(s)
- interview (more) DAS users
\end{verbatim}

% ---  Bibliography
\bibliographystyle{unsrt}
\addcontentsline{toc}{chapter}{Bibliography}

%\nocite{*}
\thispagestyle{empty}
\bibliography{refs}

\end{document}

### 2012-10-03 doc/thesis/latex/report_sept.tex
minor changes

@@
 %\usepackage[margin=0.5in]{geometry}
-% \usepackage{times}
+%\usepackage{times}
 \usepackage{color}
-\usepackage{titlesec}
+%\usepackage[small,compact]{titlesec}
 
@@
 \section{Problem statement}
+
+\textbf{\color{red}DO WE also have proprietary big databases that could be useful to be searched, but the standard integration work would be to heavy (or sub-parts of existing systems that are not covered by APIs, and could be accessible through DBs directly -- these potentially more expensive queries could processed with lower priority)}
+
 \subsection{Ease of use: Query Language over the mediated schema}
@@
 	     \begin{itemize}
+	     					\item  {\color{red}maybe we could have prioritization in general, the heavier your query is the more you have to wait in favour of  the light queries}: 
+	     					(we could have some sort of query cost evaluation based on history or manually predefined scores per API)
+
 	                		\item more intelligent caching needed:
@@
 	                			\end{itemize}
+
 	                		\item query rewriting then used with grep=smf then the same item is available as selection key. check how many instances.
@@
 
-\textbf{\color{red}TODO: Overview}
+\textbf{\color{red}TODO: Overview; evaluate performance}
 
@@
 The problem of Keyword search over Relational Databases (or also semi-structured sources like XML) has received a significant attention by the research community over the last decade.
-There exist many papers and a couple of quite comprehensive PhD.
-dissertations\cite{PhD_2011, PhD_2012} describing performance optimization details, etc.
 
-There was a number of variations of the problem including returning of only Top-k results expecting that users only care about the quality of limited number of results but not completeness of results, ...
-\textbf{\color{red}describe this in details}
+The basic approach would first build an inverted index on database tables (usually only text columns).
+Then after finding all occurrences of the keywords, would try to construct join paths (based on Foreign keys) that would unite tuples containing the keywords.
+
+% In addition to many papers the PhD dissertations \cite{PhD_2011, PhD_2012} describe the approaches in details including performance optimization details (e.g. generating materialized views), etc. 
+% 
+A number of variations exist:  returning only the Top-k results by some ranking function, SQAK and SODA would allow generating even more complex queries including aggregations.
 
@@
 \cite{forms_kws} proposes alternative approach for processing keyword queries over relational databases: given a keyword query, instead of returning database tuples one could rank query forms that best matches the query for user to choose the right one (if they are properly named this is fairly easy).
-The ranking is done by \textbf{\color{red}TODO}.
+The ranking is based on checking matching of keywords to table names in templates and to column values (could be implemented with inverted index).\textbf{\color{red}TODO: more detailed and our limitations (after reading keyword cleaning)}.
 
@@
 
-\subsubsection*{Semantic approach}
+\subsubsection*{Keyword query cleaning}
+{\color{red} From first sight looks fairly cool. Build index with some powerful fulltext search engine (e.g. Xapian) and use it to generate mapping to structured queries}
 
-Researchers from \textit{ETHZ} have been investigating Generation of SQL for Business users in collaboration with \textit{Credit Suisse}.
-To converting natural language queries (that in addition to keyword search could convey some semantic structure) into SQL statements they employed some natural language processing and used meta-data describing the schema and the domain (ontologies at multiple levels).
+\subsubsection*{Meta-data approach}
 
-Even on a large data-warehouse of ~220GB data with a complex schema of 300 tables they reported fairly promising results that proves that if good meta-data is available, generating even quite complex SQL  (n-way joins with aggregations, etc) is quite feasible for computer making it 'much easier for business users to interactively explore highly-complex data warehouses' \cite[p.932]{ethz2012}.
+With a goal to bridge the increasing gap between high-level (conceptual, business) and low level (physical) representations of data, researchers from \textit{ETHZ} have been investigating Generation of SQL for Business users at \textit{Credit Suisse}.
+For converting natural language queries
+% (that in addition to keyword search could convey some semantic structure)
+ into SQL statements, in addition to what used by earlier approaches they used meta-data describing the schema (at multiple representation levels) and the domain (ontologies) and some natural language processing.
 
-\subsubsection*{Automatic mapping between distributed services}
-{\color{red}an interesting approach but not for CERN...}
+Even on a large data-warehouse of ~220GB data with a complex schema of 400+ tables they reported that if good meta-data is available, generating even quite complex SQL  (n-way joins with aggregations, etc) is quite feasible for computer.
+That would making it 'much easier for business users to interactively explore highly-complex data warehouses' \cite[p.932]{ethz2012}.
 
-\subsubsection*{\color{red}Question Answering}
+\subsubsection*{Then there is no access to index data terms}
+{\color{red} TODO: Describe Keymantic\cite{Keymantic10}. This is a suboptimal solution as indexing terms improves results. }
+
+%\subsubsection*{Automatic mapping between distributed services}
+%{\color{red} TODO: an interesting approach but not for CERN...}
+
+%\subsubsection*{\color{red}Question Answering}
 
@@
 
-\subsection{Searching Heterogeneous web services}
+\subsection{Searching 'Deep Web' and Heterogeneous web services}
 {\color{red}There doesn't seem to be much of recent papers going towards this direction (except from specific search systems like flight fare comparisons), however these are based on structured arguments.}
@@
 
+{\color{red}Check deep web (also recent research exist)}
 {\color{red}Performance: Answering Queries using Views}
@@
 
-- if entered keyword query suggest a structured query, optionally with description in natural language
+- if entered keyword query suggest a structured query
+	based on 'Query forms' and 'Keyword cleaning' approaches
+	--> evaluate different approaches (HMM, etc) and implementations
+	optionally with description in natural language
 - select entity (e.g. dataset) --> give all available keys from forms, however there are still combinations which are enforced

### 2012-10-05 doc/thesis/latex/report_sept.tex
minor additions to report

@@
 Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]) operators.
-The combined query results could be later 'pipped' for further filtering and aggregation (min, avg, etc), e.g.:
+The combined query results could be later 'piped' for further filtering and aggregation (min, avg, etc), e.g.:
 
@@
 
+\subsection{Data integration}
+\subsubsection*{Virtual integration}
+In Virtual Integration, multiple data-sources are integrated with help of a mediator containing source descriptions and probably some schema mappings, while data is left at the sources.
+This is mainly the approach taken by our DAS.
+
+\subsubsection*{Large scale search}
+[multiple papers from Google] discusses various ways for implementing data integration in terms of large-scale search engine (Google) Virtual integration vs.
+surfacing.
+They also present ways for integrating systems without human intervention through use of statistical 'mediator'.
+
+There two approaches to web scale search for deep-web (Google mostly cares about web forms):
+
+\textit{Runtime query reformulation} - 'leaves data at the sources and routes queries to appropriate services'\cite[p. 1]{paygo_integration}
+
+\textit{Deep-web surfacing} - tries to add content from the deep-web into search index.
+There are algorithms which allow to iteratively choose input parameters to forms to surface most of the 'hidden' data without loosing much (i.e.
+if choosing parameters in not smart way the web form could yield as many results as a cross product of all input combinations).
+
+PayGo approach - there is NO single mediated schema over which users pose queries.
+Queries are routed to to the relevant sources.
+Statistical methods used to model uncertainty at all levels: queries, mappings and underlying data.
+
+\subsubsection*{Then there is no access to index data terms}
+{\color{red} TODO: Describe Keymantic\cite{Keymantic10}.
+This is a very suboptimal solution as indexing terms improves results (mention evaluation from SODA paper).
+It works only then keywords map entity names.
+
+some hybrid approaches: 
+index if exists
+regexp
+some string similarity measure based on historical data (e.g. even edit-distance would work for many items, like site)}
+
 \subsection{Keyword search over DBs}
@@
 % 
-A number of variations exist:  returning only the Top-k results by some ranking function, SQAK and SODA would allow generating even more complex queries including aggregations.
+A number of problem variations exist:  returning only the ranked Top-k results vs.
+returning ranked list of possible queries, while some systems would even allow generating more complex queries including aggregations, etc (SQAK, SODA\cite{ethz2012}).
 
@@
 \subsubsection*{Keyword query cleaning}
-{\color{red} From first sight looks fairly cool. Build index with some powerful fulltext search engine (e.g. Xapian) and use it to generate mapping to structured queries}
+Keyword queries are often ambiguous, may contain misspellings or multiple keywords that refer to the same attribute value,  therefore \cite{kw_cleaning} suggested to perform query cleaning before proceeding to subsequent more computationally expensive steps (e.g.
+exploring all the possible join paths).
+
+Further employing some machine learning method like HMM\cite{kw_cleaning_hmm} would allow to incorporate user's feedback (even the fact that user has chosen n-th result as a query to be executed is a good clue).
+
+{\color{red} In terms of virtual integration of services via system like our DAS, one could employ query cleaning process to suggest possible queries.
+
+Inverted index could be built with some powerful full-text search engine (e.g. Xapian) and used to generate mapping to structured queries}
 
@@
 That would making it 'much easier for business users to interactively explore highly-complex data warehouses' \cite[p.932]{ethz2012}.
-
-\subsubsection*{Then there is no access to index data terms}
-{\color{red} TODO: Describe Keymantic\cite{Keymantic10}. This is a suboptimal solution as indexing terms improves results. }
 
@@
 
-{\color{red}Check deep web (also recent research exist)}
-{\color{red}Performance: Answering Queries using Views}
+%{\color{red}Check deep web (also recent research exist)}
+%{\color{red}Performance: Answering Queries using Views}
 
@@
 
+\pagebreak
+\appendix
+\section{DAS Query logs}
+
+non parsible queries: 98923
+
+\subsection*{Most common query patterns (Feb 2012-June 2012)}
+\begin{verbatim}
+total valid queries: 569408
+
+50.16% (285605) :	dataset dataset.name=?
+13.54% (77071)  :	site dataset.name=?
+8.96%  (51035) 	:	file dataset.name=?
+5.88%  (33504) 	:	run dataset.name=?
+2.59%  (14739) 	:	release dataset.name=?
+2.11%  (12036) 	:	config dataset.name=?
+1.65%  (9420) 	:	dataset run.run_number=?
+1.34%  (7642) 	:	block dataset.name=?
+1.24%  (7084) 	:	dataset site.name=?
+1.15%  (6562) 	:	dataset dataset.name=? release.name=?
+1.10%  (6287) 	:	parent dataset.name=?
+0.96%  (5488) 	:	file file.name=?
+0.92%  (5245) 	:	dataset dataset.name=? status.name=?
+0.77%  (4363) 	:	dataset release.name=?
+0.75%  (4257) 	:	file dataset.name=? run.run_number=?
+0.68%  (3874) 	:	run run.run_number=?
+0.53%  (2994) 	:	site site.name=?
+0.45%  (2576) 	:	site file.name=?
+0.45%  (2556) 	:	dataset file.name=?
+0.43%  (2438) 	:	lumi file.name=?
+0.40%  (2282) 	:	dataset dataset.name=? site.name=?
+0.35%  (1999) 	:	file block.name=?
+0.35%  (1970) 	:	dataset dataset.name=? run.run_number=?
+0.29%  (1640) 	:	group dataset.name=?
+0.29%  (1631) 	:	lumi run.run_number=?
+0.25%  (1402) 	:	child dataset.name=?
+0.22%  (1268) 	:	run file.name=?
+0.21%  (1204) 	:	block block.name=?
+0.19%  (1088) 	:	release release.name=?
+0.16%  (936) 	:	site block.name=?
+0.12%  (703) 	:	file dataset.name=? lumi.number=? run.run_number=?
+0.10%  (594) 	:	parent file.name=?
+0.08%  (455) 	:	dataset block.name=?
+0.06%  (365) 	:	dataset dataset.name=? datatype.name=? release.name=?
+0.06%  (360) 	:	file dataset.name=? site.name=?
+
+Interesting non structured queries:
+py dataset=/DYJetsToLL_M-50_TuneZ2Star_8TeV-madgraph-tarball/Summer12_DR53X-PU_S10_START53_V7A-v1/AODSIM
+T!_CERN
+*herwig*/AODSIM
+
+\end{verbatim}
+
 \end{document}

### 2012-10-09 doc/thesis/latex/literature_review.tex
some changes to report (still early)

--- 2012-10-05 doc/thesis/latex/report_sept.tex
@@
-\documentclass[a4paper,11pt,draft]{article}
-\special{papersize=210mm,297mm}
-
-% twocolumn
-%\usepackage{fullpage} 
-\usepackage[cm]{fullpage}
-%\usepackage[margin=0.5in]{geometry}
-%\usepackage{times}
-\usepackage{color}
-%\usepackage[small,compact]{titlesec}
-
-\bibliographystyle{alpha}
-
-% usually first paragraph is NOT indented, so this is commented out: 
-% \usepackage{indentfirst}
-
-\begin{document}
-
-% Article top matter
-\title{Master thesis problem statement: Keyword search over heterogeneous sources} %\LaTeX is a macro for printing the Latex logo
-\author{Vidmantas Zemleris}  %\texttt formats the text to a typewriter style font
-\date{\today}  %\today is replaced with the current date
-
-%\maketitle
-
-%
- \centerline{\Large \bf Searching heterogeneous data-sources: Master thesis problem statement} %% Paper title
- \medskip
-
- \centerline{Vidmantas Zemleris, \today}  %% Author name(s)
- \medskip
-
-\section{Introduction}
-At large scientific collaborations like the CMS Experiment at CERN's LHC that includes more than 3000 collaborators data usually resides on a fair number of autonomous and heterogeneous proprietary systems each serving it's own purpose
-\footnote{For instance, at CERN,  due to many reasons (e.g.
-research and need of freedom, politics of institutes involved) software projects usually evolve in independent fashion resulting in fair number of proprietary systems\cite{Koch00CERN}.
-Further high turnover makes it harder extending these systems}.
-As data stored on one system may be related to data residing on other systems%
-	\footnote{For example, datasets containing physics events are registered at DBS, while the physical location of files is tracked by Phedex which also takes care of their transfers within the worldwide grid storage}%
-, users are in need of a centralized and easy-to-use solution for locating and combining data from all these multiple services.
-
-Using a highly structured language like SQL is problematic because users need to know not only the language but also where to find the information and also lots of technical details like schema.
-A data integration system based on simple structured queries is already in place.
-Still {\color{red}various improvements including support} for less restricted keyword queries and improvements to system's usability and performance still have to be researched.
-
-% {\color{red}A remarkable opportunity we have is possibility to test ideas on a sample of ~3000 people working at CMS.}
-
-\section{Case study: the CMS Experiment at CERN}
-Users' information need may vary greatly depending on their role, however most of the time they are interested in locating full set of entities matching some selection criteria, e.g.:
-%\footnote{in contrary to exploratory approach where  a subset of best matches is enough}
-   \begin{itemize}
-  		\item find all \textit{files} from \textit{dataset(s)} matching wild-card query each containing some of the 'interesting' \textit{runs} from a list provided (Release validation teams)
-         \item find (all) \textit{datasets} related some specific physics phenomena\footnote{In case of dataset this data is present in filename or run} together with conditions describing how this data was recorded by detector or simulated which are present in separate autonomous system than the datasets (Physicists)
-         % TODO:is this a good example? conditions is a separate query now...
-         \item find all \textit{datasets} matching some pattern stored at a given \textit{site} (filtering on entities stored on separate services)
-   \end{itemize}                		
-% In such dynamic environment it is preferred to choose local-as-view approach for information integration instead of creating a global schemata\cite{Koch00CERN}.
-
-For more use-cases of data retrieval at CMS Experiment see \cite{CMS_data08}.
-
-%Following a similar approach a 
-\subsection*{The Data Aggregation System}
-
-The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it queries the data-sources, merges the results, and caches them for subsequent uses.
-
-Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]) operators.
-The combined query results could be later 'piped' for further filtering and aggregation (min, avg, etc), e.g.:
-
-{\footnotesize 
-\begin{verbatim}
-dataset=*RelVal* | grep dataset.nevents >1000 | avg(dataset.size), median(dataset.size)
-\end{verbatim}
-}
-
-The query above would return average and median datasets sizes  of ones containing  \textit{RelVal} in their name having more than 1000 events.
-
-Queries could be run either from web browser or through  command line interface where the results could  be fed into another application (e.g.
-program doing physics analysis or automatic software release validation).
-
-% \textbf{\color{red}Examples of more general uses: Digital libraries, Stock markets, [ Weather, Flight costs]}
-
-\section{Problem statement}
-
-\textbf{\color{red}DO WE also have proprietary big databases that could be useful to be searched, but the standard integration work would be to heavy (or sub-parts of existing systems that are not covered by APIs, and could be accessible through DBs directly -- these potentially more expensive queries could processed with lower priority)}
-
-\subsection{Ease of use: Query Language over the mediated schema}
-	Even a very simple structured query language that also contain entity names over the mediated schema may seem {\color{red}hard} to learn, especially in the beginning.
-   On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources).
-   So some sort of guiding helping user to build the query shall be useful.
-
-	Further a minimum set of search predicates is imposed by APIs (mainly because of performance reasons) and user has to be at least informed what is he expected to provide.
-
-Also to consider:
-\begin{itemize}         
-       \item input ambiguity, typos
-       \item {\color{red} what about ranked search?}
-\end{itemize}
-
-\subsection{Performance}
-\subsubsection{Handling distributed search efficiently}
-
-\subsubsection{Other performance issues}
-	     \begin{itemize}
-	     					\item  {\color{red}maybe we could have prioritization in general, the heavier your query is the more you have to wait in favour of  the light queries}: 
-	     					(we could have some sort of query cost evaluation based on history or manually predefined scores per API)
-
-	                		\item more intelligent caching needed:
-	                			\begin{itemize}
-	                			\item given an entity received from provider, determining if it is useful to cache for long-term
-	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
-	                			\item could make use of even older information warning user
-	                			\item could also have different validity dates for certain fields. if certain field is not explicitly requested, even a very old cache could be used.
-	                			\end{itemize}
-
-	                		\item query rewriting then used with grep=smf then the same item is available as selection key. check how many instances.
-	                		\item pre-fetching common (sub-)queries: determining manually and/or automatically
-	                		\item scale testing - if we are storing lots of historical info. MongoDB is not so performant if DB cant fit in memory. One of well known solutions is installing SSD.
-                        \item ? with current API's for aggregation service has to fetch ALL records not just the first page..
-                        \item ??? do we have cases then multiple APIs of the same service could be called for the same data, can we choose just one???
-                        \item ? showing partial results
-        \end{itemize}
-
 \section{Preliminary Literature review}
@@
 
-\subsection{Data integration}
-\subsubsection*{Virtual integration}
-In Virtual Integration, multiple data-sources are integrated with help of a mediator containing source descriptions and probably some schema mappings, while data is left at the sources.
-This is mainly the approach taken by our DAS.
+\subsection{Searching 'Deep Web' and Heterogeneous web services}
 
-\subsubsection*{Large scale search}
+In 1990s there were was much research on search based service integration systems, for instance \textit{Information Manifold} following Local-as-View approach and \textit{TSIMMIS} developed at Stanford following Global-as-View approach.
+The \textit{Information Manifold} is quite resembling to DAS {\color{red},  so it could be useful checking it's papers.
+TODO: finish}
+
+\subsubsection*{Deep web search at Google}
 [multiple papers from Google] discusses various ways for implementing data integration in terms of large-scale search engine (Google) Virtual integration vs.
@@
 
-{\color{red} In terms of virtual integration of services via system like our DAS, one could employ query cleaning process to suggest possible queries.
-
-Inverted index could be built with some powerful full-text search engine (e.g. Xapian) and used to generate mapping to structured queries}
-
 \subsubsection*{Meta-data approach}
@@
 
-\subsection{Searching 'Deep Web' and Heterogeneous web services}
-{\color{red}There doesn't seem to be much of recent papers going towards this direction (except from specific search systems like flight fare comparisons), however these are based on structured arguments.}
+% {\color{red}There doesn't seem to be much of recent papers going towards this direction (except from specific search systems like flight fare comparisons), however these are based on structured arguments.}
 
-In 1990s there were was much research on search based service integration systems, for instance \textit{Information Manifold} following Local-as-View approach and \textit{TSIMMIS} developed at Stanford following Global-as-View approach.
+% Some other approaches could include: - creating virtual documents offline by joining tables, for instantaneous search results by employing IR techniques
+% (Indexing Relational Database Content Offline for Efficient Keyword-Based Search, 2003)
 
-The \textit{Information Manifold} is quite resembling to DAS {\color{red},  so it could be useful checking it's papers. TODO: finish}
-
-%{\color{red}Check deep web (also recent research exist)}
-%{\color{red}Performance: Answering Queries using Views}
-
-\begin{verbatim}
-Some other approaches could include:
-
-- creating virtual documents offline by joining tables, for instantaneous search results by employing IR techniques
-(Indexing Relational Database Content Offline for Efficient Keyword-Based Search, 2003)
-
-- (Efficient Keyword Search Across Heterogeneous Relational Databases, 2007)
- : combines schema matching and structure discovery techniques to ﬁnd approximate foreign-key joins across heterogeneous databases
-\end{verbatim}
-
-\section{Possible approaches}
-
-\subsection{Ease of use}
-\begin{verbatim}
-- javascript interpreter
-    : field + description -- dropdown searcheable list
-
-- if entered keyword query suggest a structured query
-	based on 'Query forms' and 'Keyword cleaning' approaches
-	--> evaluate different approaches (HMM, etc) and implementations
-	optionally with description in natural language
-- select entity (e.g. dataset) --> give all available keys from forms, however there are still combinations which are enforced
-\end{verbatim}
-%(could use Natural Language Generation)
-
-\subsection{Performance}
-\subsubsection{Continuous view maintenance at Data providers with large DBs}
-Use either \textit{materialized refresh fast views with query rewriting} (Oracle; completely transparent for proprietary apps)\cite{Oracle11}
- or some other continuous view maintenance tool (e.g.
-DBToaster\footnote{http://www.dbtoaster.org that is being developed at EPFL}) to improve performance of heavy queries containing joins and/or aggregations.
-
-\subsection{Integrating distributed information efficiently}
-Bloom-join (which could be quite transparent and implemented even on DB side [pure SQL is possible for MySQL, to check for Oracle]- take a query and bit-vector as parameter) , lazy pagination (and order required for aggregation) - this is not yet supported by any of the data-service APIs
-
-%\newline
-integration at DAS level. (at source DB could be more performant a little)
-
-\subsection{More intelligent caching}
-Some of the data entities instances changing very rarely, for example in DBS system old datasets would never change, while new ones are constantly added (still some of their attributes may change, in this example validity).
-
-Cached copy may be shown before hand while up to date results could be retrieved on user's request.
-
-\textbf{Decide if to cache or not...}
-
-An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.
-
-\begin{verbatim}
-Also: Caching of intermediary results, not only the queries
-\end{verbatim}
-
-\section{Work status}
-TODO
-
-\begin{verbatim}
-- obtained DB copy of biggest data provider DBS (currently 80 GB + 200 GB indexes)
-- preliminary literature review (to be continued deeper + waiting for book arrival: Principles of Data Integration)
-
---- some analysis of DAS logs
-
-Upcomming Work items:
-- couple of fairly simple prototypes of UI/access patters for simpler DAS querying
-- check on performance improvements after creation of materialized view(s)
-- interview (more) DAS users
-\end{verbatim}
-
-% ---  Bibliography
-\bibliographystyle{unsrt}
-\addcontentsline{toc}{chapter}{Bibliography}
-
-%\nocite{*}
-\thispagestyle{empty}
-\bibliography{refs}
-
-\pagebreak
-\appendix
-\section{DAS Query logs}
-
-non parsible queries: 98923
-
-\subsection*{Most common query patterns (Feb 2012-June 2012)}
-\begin{verbatim}
-total valid queries: 569408
-
-50.16% (285605) :	dataset dataset.name=?
-13.54% (77071)  :	site dataset.name=?
-8.96%  (51035) 	:	file dataset.name=?
-5.88%  (33504) 	:	run dataset.name=?
-2.59%  (14739) 	:	release dataset.name=?
-2.11%  (12036) 	:	config dataset.name=?
-1.65%  (9420) 	:	dataset run.run_number=?
-1.34%  (7642) 	:	block dataset.name=?
-1.24%  (7084) 	:	dataset site.name=?
-1.15%  (6562) 	:	dataset dataset.name=? release.name=?
-1.10%  (6287) 	:	parent dataset.name=?
-0.96%  (5488) 	:	file file.name=?
-0.92%  (5245) 	:	dataset dataset.name=? status.name=?
-0.77%  (4363) 	:	dataset release.name=?
-0.75%  (4257) 	:	file dataset.name=? run.run_number=?
-0.68%  (3874) 	:	run run.run_number=?
-0.53%  (2994) 	:	site site.name=?
-0.45%  (2576) 	:	site file.name=?
-0.45%  (2556) 	:	dataset file.name=?
-0.43%  (2438) 	:	lumi file.name=?
-0.40%  (2282) 	:	dataset dataset.name=? site.name=?
-0.35%  (1999) 	:	file block.name=?
-0.35%  (1970) 	:	dataset dataset.name=? run.run_number=?
-0.29%  (1640) 	:	group dataset.name=?
-0.29%  (1631) 	:	lumi run.run_number=?
-0.25%  (1402) 	:	child dataset.name=?
-0.22%  (1268) 	:	run file.name=?
-0.21%  (1204) 	:	block block.name=?
-0.19%  (1088) 	:	release release.name=?
-0.16%  (936) 	:	site block.name=?
-0.12%  (703) 	:	file dataset.name=? lumi.number=? run.run_number=?
-0.10%  (594) 	:	parent file.name=?
-0.08%  (455) 	:	dataset block.name=?
-0.06%  (365) 	:	dataset dataset.name=? datatype.name=? release.name=?
-0.06%  (360) 	:	file dataset.name=? site.name=?
-
-Interesting non structured queries:
-py dataset=/DYJetsToLL_M-50_TuneZ2Star_8TeV-madgraph-tarball/Summer12_DR53X-PU_S10_START53_V7A-v1/AODSIM
-T!_CERN
-*herwig*/AODSIM
-
-\end{verbatim}
-
-\end{document}
+% TODO: (Efficient Keyword Search Across Heterogeneous Relational Databases, 2007) : combines schema matching and structure discovery techniques to ﬁnd approximate foreign-key joins across heterogeneous databases

### 2012-10-09 doc/thesis/latex/proposed_solutions.tex
some changes to report (still early)

\section{Proposed solutions}
\subsection{Ease of use}

\begin{itemize}
\item Interactive guidance on how to compose queries to ease the learning (could also include some examples of popular queries):
\begin{enumerate}
\item What entity are you searching for?
\item How could you identify it (i.e. what do you know)?
\item After seeing first results: What do you want to do with the results? (list items; select only specific columns; do aggregation; use in command line/another program)
\end{enumerate}

\item (?) JavaScript-based query interpreter to ease query writing: could suggest search attribute and entity attribute names. also taking into account some possible ambiguous namings (auto completion)

\item Keyword query --> ranked list of matching structured queries:
	\begin{itemize}
	\item map keywords to entity name and selection parameters (quite resembling 'Query forms' and 'Keyword cleaning' approaches)
	\item
		{\color{red} Inverted index could be built with help of some good full-text search engine (e.g. Xapian\footnote{http://xapian.org/}) and used to generate mapping to structured queries. If no match found (e.g. new entry not yet in our cache) some pattern matching could be used to try to corresponding guess search attribute}
	\item ranking could be based on: how closely keywords are matching some or all required parameters, popularity of certain query and users feedback

	\item {\color{red}	--> evaluate different approaches (HMM, etc); could some natural language processing help?}
	\item special attention to \textbf{wilcard} queries

	\end{itemize}
\end{itemize}

\subsection{Performance}
\subsubsection{Continuous view maintenance at Data providers with large DBs}
Use either \textit{materialized refresh fast views with query rewriting} (Oracle; completely transparent for proprietary apps)\cite{Oracle11}
 or some other continuous view maintenance tool (e.g. DBToaster\footnote{http://www.dbtoaster.org that is being developed at EPFL}) to improve performance of heavy queries containing joins and/or aggregations.

\subsubsection{More intelligent caching}
Some of the data entities instances changing very rarely, for example in DBS system old datasets would never change, while new ones are constantly added (still some of their attributes may change, in this example validity).

Cached copy may be shown before hand while up to date results could be retrieved on user's request.

An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.

Also consider:
                			\begin{itemize}
                			\item given an entity received from provider, determining if it is useful to cache for long-term
%	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
                			\item could cache even \textit{expired} data, but warning user
                			\item could also have different validity dates for certain fields. if no volatile fields are not explicitly requested, even a very old cache could be used. 
                			\item Can we automatically figure which fields are static and which are changing?
                			\item pre-fetching common (sub-)queries: determining manually and/or automatically

                			\item \textbf{Deciding if to cache or not...}

                			\end{itemize}

\subsection{Integrating distributed information efficiently}
Bloom-join (which could be quite transparent and implemented even on DB side [pure SQL is possible for MySQL, to check for Oracle]- take a query and bit-vector as parameter) , lazy pagination (and order required for aggregation) - this is not yet supported by any of the data-service APIs

%\newline
% integration at DAS level. (at source DB could be more performant a little)

### 2012-10-09 doc/thesis/latex/report_sept.tex
some changes to report (still early)

--- 2012-10-02 doc/thesis/latex/report_sept.tex
@@
 \section{Case study: the CMS Experiment at CERN}
-Users' information need may vary greatly depending on their role, however most of the time they are interested in locating full set of entities matching some selection criteria, e.g.:
+Users' information need may vary greatly depending on their role, however most of the time they are interested in locating a \textit{full set} of entities matching some selection criteria, e.g.:
 %\footnote{in contrary to exploratory approach where  a subset of best matches is enough}
@@
 
-The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it queries the data-sources, merges the results, and caches them for subsequent uses.
+The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it queries the data-sources, merges the results, and caches them for subsequent use.
+DAS uses \textit{Boolean retrieval model} as users are often interested in retrieving ALL the items matching their query.
 
 Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]) operators.
-The combined query results could be later 'pipped' for further filtering and aggregation (min, avg, etc), e.g.:
+The combined query results could be later 'piped' for further filtering and aggregation (min, avg, etc), e.g.:
 
@@
 \section{Problem statement}
+
+DAS is based on \textit{Virtual Integration} where data is left at the sources allowing data to be volatile (e.g.
+new records gathered or existing records updated; service descriptions may be changed; new services added), while some subsets of the data could be fairly static (e.g.
+datasets stored at DBS do not change often).
+Further as the total data that is being queried could be fairly large (with growth of order of 1TB per year) and the data-services may have certain constraints (e.g.
+only certain queries allowed) as their main goal is supporting production (data taking from the CMS detector).
+
+Therefore, it is important to balance between returning locally cached query results quickly   and between issuing queries to slow data-services to get up to date results.
+It is also not less important to provide users with easy to use interface having a fairly flat learning curve, while at the same time not loosing the functionality of fairly complex queries.
+% expressive power.
+\textbf{\color{red}Another issue is about handling distributed queries, where either selection criteria or selection values could come from distributed sources -- the biggest problem could be in the APIs they [DO NOT] provide}.
+
+%TODO: Further it has to comply with dataset constraints
+
+% \textbf{\color{red}DO WE also have proprietary big databases that could be useful to be searched, but the standard integration work would be to heavy (or sub-parts of existing systems that are not covered by APIs, and could be accessible through DBs directly -- these potentially more expensive queries could processed with lower priority)}
+
 \subsection{Ease of use: Query Language over the mediated schema}
@@
 
+   Supporting non-structured keyword queries is also worth investigation as many users reporting they are missing this Google-like search experience.
+{\color{red}Because DAS uses Boolean IR we could only rank specific structured queries.}
+
 	Further a minimum set of search predicates is imposed by APIs (mainly because of performance reasons) and user has to be at least informed what is he expected to provide.
 
-Also to consider:
+Also to consider: input ambiguity and typos, some queries are more common than others.
+
+\subsection{Performance}
+
 \begin{itemize}
-       \item input ambiguity, typos
-       \item {\color{red} what about ranked search?}
+	     					\item  Query prioritization: {\color{red}maybe we could have prioritization in general, the heavier your query is the more you have to wait in favour of  the light queries}: 
+	     					(we could have some sort of query cost evaluation based on history or manually predefined scores per API)
+
+                		\item Explore more intelligent caching
+
+	                		% TODO: \item query rewriting then used with grep=smf then the same item is available as selection key. check how many instances.
+
+					\item Could displaying Partial Results improve performance?
+					\begin{itemize}
+					     \item given current APIs to do aggregation over data sources, DAS has to fetch ALL records matching the query instead of only the first page.
+%                        \item ??? do we have cases then multiple APIs of the same service could be called for the same data, can we choose just one???
+                        \item {\color{red}e.g. intermediary aggregation results while still calculating}
+
 					\end{itemize}
 
-\subsection{Performance}
-\subsubsection{Handling distributed search efficiently}
-
-\subsubsection{Other performance issues}
-	     \begin{itemize}
-	                		\item more intelligent caching needed:
-	                			\begin{itemize}
-	                			\item given an entity received from provider, determining if it is useful to cache for long-term
-	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
-	                			\item could make use of even older information warning user
-	                			\item could also have different validity dates for certain fields. if certain field is not explicitly requested, even a very old cache could be used.
-	                			\end{itemize}
-	                		\item query rewriting then used with grep=smf then the same item is available as selection key. check how many instances.
-	                		\item pre-fetching common (sub-)queries: determining manually and/or automatically
-	                		\item scale testing - if we are storing lots of historical info. MongoDB is not so performant if DB cant fit in memory. One of well known solutions is installing SSD.
-                        \item ? with current API's for aggregation service has to fetch ALL records not just the first page..
-                        \item ??? do we have cases then multiple APIs of the same service could be called for the same data, can we choose just one???
-                        \item ? showing partial results
+	                	 \item scale testing - if we are storing lots of historical info.
+MongoDB is not so performant if DB cant fit in memory.
+One of well known solutions is installing SSD.
 \end{itemize}
 
-\section{Preliminary Literature review}
+\subsection{Handling distributed search efficiently}
 
-% Overview: keyword search is good for non-structured documents, it is not [as] effective with structured sources\cite{Levy96}, therefore the keyword search on relational databases approach is good then no specific APIs exist and it's there are no resources to do \textbf{a full integration} .
+\pagebreak
+\include{proposed_solutions}
 
-\textbf{\color{red}TODO: Overview}
-
-\subsection{Keyword search over DBs}
-The problem of Keyword search over Relational Databases (or also semi-structured sources like XML) has received a significant attention by the research community over the last decade.
-There exist many papers and a couple of quite comprehensive PhD.
-dissertations\cite{PhD_2011, PhD_2012} describing performance optimization details, etc.
-
-There was a number of variations of the problem including returning of only Top-k results expecting that users only care about the quality of limited number of results but not completeness of results, ...
-\textbf{\color{red}describe this in details}
-
-\subsubsection*{Ranking Query Templates based on keyword query}
-A simple way to access relational database could  be through a set of predefined named query templates (SQL with selection parameters or operators still to be specified) exposed to a user as a Form that the user has to fill in.
-
-\cite{forms_kws} proposes alternative approach for processing keyword queries over relational databases: given a keyword query, instead of returning database tuples one could rank query forms that best matches the query for user to choose the right one (if they are properly named this is fairly easy).
-The ranking is done by \textbf{\color{red}TODO}.
-
-An interesting feature of this approach is that a Query Template is functionally similar to any autonomous web service (which given the parameters would in turn execute that query on its database).
-In case of the Data Aggregation System, a user after entering a keyword query could be provided with a ranked list of structured queries (attribute=value) that could be processed given data source constraints (e.g.
-parameters required) and if needed he could refine his search (e.g.
-provide more parameters).
-
-\subsubsection*{Semantic approach}
-
-Researchers from \textit{ETHZ} have been investigating Generation of SQL for Business users in collaboration with \textit{Credit Suisse}.
-To converting natural language queries (that in addition to keyword search could convey some semantic structure) into SQL statements they employed some natural language processing and used meta-data describing the schema and the domain (ontologies at multiple levels).
-
-Even on a large data-warehouse of ~220GB data with a complex schema of 300 tables they reported fairly promising results that proves that if good meta-data is available, generating even quite complex SQL  (n-way joins with aggregations, etc) is quite feasible for computer making it 'much easier for business users to interactively explore highly-complex data warehouses' \cite[p.932]{ethz2012}.
-
-\subsubsection*{Automatic mapping between distributed services}
-{\color{red}an interesting approach but not for CERN...}
-
-\subsubsection*{\color{red}Question Answering}
-
-% See Appendix #1, for evaluation of they Demo system.
-
-\subsection{Searching Heterogeneous web services}
-{\color{red}There doesn't seem to be much of recent papers going towards this direction (except from specific search systems like flight fare comparisons), however these are based on structured arguments.}
-
-In 1990s there were was much research on search based service integration systems, for instance \textit{Information Manifold} following Local-as-View approach and \textit{TSIMMIS} developed at Stanford following Global-as-View approach.
-
-The \textit{Information Manifold} is quite resembling to DAS {\color{red},  so it could be useful checking it's papers. TODO: finish}
-
-{\color{red}Performance: Answering Queries using Views}
-
-\begin{verbatim}
-Some other approaches could include:
-
-- creating virtual documents offline by joining tables, for instantaneous search results by employing IR techniques
-(Indexing Relational Database Content Offline for Efficient Keyword-Based Search, 2003)
-
-- (Efficient Keyword Search Across Heterogeneous Relational Databases, 2007)
- : combines schema matching and structure discovery techniques to ﬁnd approximate foreign-key joins across heterogeneous databases
-\end{verbatim}
-
-\section{Possible approaches}
-
-\subsection{Ease of use}
-\begin{verbatim}
-- javascript interpreter
-    : field + description -- dropdown searcheable list
-
-- if entered keyword query suggest a structured query, optionally with description in natural language
-- select entity (e.g. dataset) --> give all available keys from forms, however there are still combinations which are enforced
-\end{verbatim}
-%(could use Natural Language Generation)
-
-\subsection{Performance}
-\subsubsection{Continuous view maintenance at Data providers with large DBs}
-Use either \textit{materialized refresh fast views with query rewriting} (Oracle; completely transparent for proprietary apps)\cite{Oracle11}
- or some other continuous view maintenance tool (e.g.
-DBToaster\footnote{http://www.dbtoaster.org that is being developed at EPFL}) to improve performance of heavy queries containing joins and/or aggregations.
-
-\subsection{Integrating distributed information efficiently}
-Bloom-join (which could be quite transparent and implemented even on DB side [pure SQL is possible for MySQL, to check for Oracle]- take a query and bit-vector as parameter) , lazy pagination (and order required for aggregation) - this is not yet supported by any of the data-service APIs
-
-%\newline
-integration at DAS level. (at source DB could be more performant a little)
-
-\subsection{More intelligent caching}
-Some of the data entities instances changing very rarely, for example in DBS system old datasets would never change, while new ones are constantly added (still some of their attributes may change, in this example validity).
-
-Cached copy may be shown before hand while up to date results could be retrieved on user's request.
-
-\textbf{Decide if to cache or not...}
-
-An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.
-
-\begin{verbatim}
-Also: Caching of intermediary results, not only the queries
-\end{verbatim}
+\pagebreak
+\include{literature_review}
 
@@
 
+\pagebreak
+\appendix
+\section{DAS Query logs}
+
+non parsible queries: 98923
+
+\subsection*{Most common query patterns (Feb 2012-June 2012)}
+\begin{verbatim}
+total valid queries: 569408
+
+50.16% (285605) :	dataset dataset.name=?
+13.54% (77071)  :	site dataset.name=?
+8.96%  (51035) 	:	file dataset.name=?
+5.88%  (33504) 	:	run dataset.name=?
+2.59%  (14739) 	:	release dataset.name=?
+2.11%  (12036) 	:	config dataset.name=?
+1.65%  (9420) 	:	dataset run.run_number=?
+1.34%  (7642) 	:	block dataset.name=?
+1.24%  (7084) 	:	dataset site.name=?
+1.15%  (6562) 	:	dataset dataset.name=? release.name=?
+1.10%  (6287) 	:	parent dataset.name=?
+0.96%  (5488) 	:	file file.name=?
+0.92%  (5245) 	:	dataset dataset.name=? status.name=?
+0.77%  (4363) 	:	dataset release.name=?
+0.75%  (4257) 	:	file dataset.name=? run.run_number=?
+0.68%  (3874) 	:	run run.run_number=?
+0.53%  (2994) 	:	site site.name=?
+0.45%  (2576) 	:	site file.name=?
+0.45%  (2556) 	:	dataset file.name=?
+0.43%  (2438) 	:	lumi file.name=?
+0.40%  (2282) 	:	dataset dataset.name=? site.name=?
+0.35%  (1999) 	:	file block.name=?
+0.35%  (1970) 	:	dataset dataset.name=? run.run_number=?
+0.29%  (1640) 	:	group dataset.name=?
+0.29%  (1631) 	:	lumi run.run_number=?
+0.25%  (1402) 	:	child dataset.name=?
+0.22%  (1268) 	:	run file.name=?
+0.21%  (1204) 	:	block block.name=?
+0.19%  (1088) 	:	release release.name=?
+0.16%  (936) 	:	site block.name=?
+0.12%  (703) 	:	file dataset.name=? lumi.number=? run.run_number=?
+0.10%  (594) 	:	parent file.name=?
+0.08%  (455) 	:	dataset block.name=?
+0.06%  (365) 	:	dataset dataset.name=? datatype.name=? release.name=?
+0.06%  (360) 	:	file dataset.name=? site.name=?
+
+Interesting non structured queries:
+T!_CERN
+*herwig*/AODSIM
+
+\end{verbatim}
+% py dataset=/DYJetsToLL_M-50_TuneZ2Star_8TeV-madgraph-tarball/Summer12_DR53X-PU_S10_START53_V7A-v1/AODSIM
+
 \end{document}

### 2012-10-11 doc/thesis/latex/das_logs.tex
updated problem definition, solutions and literature review. review to be finished

--- 2012-10-09 doc/thesis/latex/report_sept.tex
@@
-\documentclass[a4paper,11pt,draft]{article}
-\special{papersize=210mm,297mm}
-
-% twocolumn
-%\usepackage{fullpage} 
-\usepackage[cm]{fullpage}
-%\usepackage[margin=0.5in]{geometry}
-%\usepackage{times}
-\usepackage{color}
-%\usepackage[small,compact]{titlesec}
-
-\bibliographystyle{alpha}
-
-% usually first paragraph is NOT indented, so this is commented out: 
-% \usepackage{indentfirst}
-
-\begin{document}
-
-% Article top matter
-\title{Master thesis problem statement: Keyword search over heterogeneous sources} %\LaTeX is a macro for printing the Latex logo
-\author{Vidmantas Zemleris}  %\texttt formats the text to a typewriter style font
-\date{\today}  %\today is replaced with the current date
-
-%\maketitle
-
-%
- \centerline{\Large \bf Searching heterogeneous data-sources: Master thesis problem statement} %% Paper title
- \medskip
-
- \centerline{Vidmantas Zemleris, \today}  %% Author name(s)
- \medskip
-
-\section{Introduction}
-At large scientific collaborations like the CMS Experiment at CERN's LHC that includes more than 3000 collaborators data usually resides on a fair number of autonomous and heterogeneous proprietary systems each serving it's own purpose
-\footnote{For instance, at CERN,  due to many reasons (e.g.
-research and need of freedom, politics of institutes involved) software projects usually evolve in independent fashion resulting in fair number of proprietary systems\cite{Koch00CERN}.
-Further high turnover makes it harder extending these systems}.
-As data stored on one system may be related to data residing on other systems%
-	\footnote{For example, datasets containing physics events are registered at DBS, while the physical location of files is tracked by Phedex which also takes care of their transfers within the worldwide grid storage}%
-, users are in need of a centralized and easy-to-use solution for locating and combining data from all these multiple services.
-
-Using a highly structured language like SQL is problematic because users need to know not only the language but also where to find the information and also lots of technical details like schema.
-A data integration system based on simple structured queries is already in place.
-Still {\color{red}various improvements including support} for less restricted keyword queries and improvements to system's usability and performance still have to be researched.
-
-% {\color{red}A remarkable opportunity we have is possibility to test ideas on a sample of ~3000 people working at CMS.}
-
-\section{Case study: the CMS Experiment at CERN}
-Users' information need may vary greatly depending on their role, however most of the time they are interested in locating a \textit{full set} of entities matching some selection criteria, e.g.:
-%\footnote{in contrary to exploratory approach where  a subset of best matches is enough}
-   \begin{itemize}
-  		\item find all \textit{files} from \textit{dataset(s)} matching wild-card query each containing some of the 'interesting' \textit{runs} from a list provided (Release validation teams)
-         \item find (all) \textit{datasets} related some specific physics phenomena\footnote{In case of dataset this data is present in filename or run} together with conditions describing how this data was recorded by detector or simulated which are present in separate autonomous system than the datasets (Physicists)
-         % TODO:is this a good example? conditions is a separate query now...
-         \item find all \textit{datasets} matching some pattern stored at a given \textit{site} (filtering on entities stored on separate services)
-   \end{itemize}                		
-% In such dynamic environment it is preferred to choose local-as-view approach for information integration instead of creating a global schemata\cite{Koch00CERN}.
-
-For more use-cases of data retrieval at CMS Experiment see \cite{CMS_data08}.
-
-%Following a similar approach a 
-\subsection*{The Data Aggregation System}
-
-The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it queries the data-sources, merges the results, and caches them for subsequent use.
-DAS uses \textit{Boolean retrieval model} as users are often interested in retrieving ALL the items matching their query.
-
-Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]) operators.
-The combined query results could be later 'piped' for further filtering and aggregation (min, avg, etc), e.g.:
-
-{\footnotesize 
-\begin{verbatim}
-dataset=*RelVal* | grep dataset.nevents >1000 | avg(dataset.size), median(dataset.size)
-\end{verbatim}
-}
-
-The query above would return average and median datasets sizes  of ones containing  \textit{RelVal} in their name having more than 1000 events.
-
-Queries could be run either from web browser or through  command line interface where the results could  be fed into another application (e.g.
-program doing physics analysis or automatic software release validation).
-
-% \textbf{\color{red}Examples of more general uses: Digital libraries, Stock markets, [ Weather, Flight costs]}
-
-\section{Problem statement}
-
-DAS is based on \textit{Virtual Integration} where data is left at the sources allowing data to be volatile (e.g.
-new records gathered or existing records updated; service descriptions may be changed; new services added), while some subsets of the data could be fairly static (e.g.
-datasets stored at DBS do not change often).
-Further as the total data that is being queried could be fairly large (with growth of order of 1TB per year) and the data-services may have certain constraints (e.g.
-only certain queries allowed) as their main goal is supporting production (data taking from the CMS detector).
-
-Therefore, it is important to balance between returning locally cached query results quickly   and between issuing queries to slow data-services to get up to date results.
-It is also not less important to provide users with easy to use interface having a fairly flat learning curve, while at the same time not loosing the functionality of fairly complex queries.
-% expressive power.
-\textbf{\color{red}Another issue is about handling distributed queries, where either selection criteria or selection values could come from distributed sources -- the biggest problem could be in the APIs they [DO NOT] provide}.
-
-%TODO: Further it has to comply with dataset constraints
-
-% \textbf{\color{red}DO WE also have proprietary big databases that could be useful to be searched, but the standard integration work would be to heavy (or sub-parts of existing systems that are not covered by APIs, and could be accessible through DBs directly -- these potentially more expensive queries could processed with lower priority)}
-
-\subsection{Ease of use: Query Language over the mediated schema}
-	Even a very simple structured query language that also contain entity names over the mediated schema may seem {\color{red}hard} to learn, especially in the beginning.
-   On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources).
-   So some sort of guiding helping user to build the query shall be useful.
-
-   Supporting non-structured keyword queries is also worth investigation as many users reporting they are missing this Google-like search experience.
-{\color{red}Because DAS uses Boolean IR we could only rank specific structured queries.}
-
-	Further a minimum set of search predicates is imposed by APIs (mainly because of performance reasons) and user has to be at least informed what is he expected to provide.
-
-Also to consider: input ambiguity and typos, some queries are more common than others.
-
-\subsection{Performance}
-
-\begin{itemize}
-	     					\item  Query prioritization: {\color{red}maybe we could have prioritization in general, the heavier your query is the more you have to wait in favour of  the light queries}: 
-	     					(we could have some sort of query cost evaluation based on history or manually predefined scores per API)
-
-                		\item Explore more intelligent caching
-
-	                		% TODO: \item query rewriting then used with grep=smf then the same item is available as selection key. check how many instances.
-
-					\item Could displaying Partial Results improve performance?
-					\begin{itemize}
-					     \item given current APIs to do aggregation over data sources, DAS has to fetch ALL records matching the query instead of only the first page.
-%                        \item ??? do we have cases then multiple APIs of the same service could be called for the same data, can we choose just one???
-                        \item {\color{red}e.g. intermediary aggregation results while still calculating}
-
-					\end{itemize}
-
-	                	 \item scale testing - if we are storing lots of historical info.
-MongoDB is not so performant if DB cant fit in memory.
-One of well known solutions is installing SSD.
-\end{itemize}
-
-\subsection{Handling distributed search efficiently}
-
-\pagebreak
-\include{proposed_solutions}
-
-\pagebreak
-\include{literature_review}
-
-\section{Work status}
-TODO
-
-\begin{verbatim}
-- obtained DB copy of biggest data provider DBS (currently 80 GB + 200 GB indexes)
-- preliminary literature review (to be continued deeper + waiting for book arrival: Principles of Data Integration)
-
---- some analysis of DAS logs
-
-Upcomming Work items:
-- couple of fairly simple prototypes of UI/access patters for simpler DAS querying
-- check on performance improvements after creation of materialized view(s)
-- interview (more) DAS users
-\end{verbatim}
-
-% ---  Bibliography
-\bibliographystyle{unsrt}
-\addcontentsline{toc}{chapter}{Bibliography}
-
-%\nocite{*}
-\thispagestyle{empty}
-\bibliography{refs}
-
-\pagebreak
 \appendix
@@
 
-non parsible queries: 98923
+\subsection*{Most common query patterns (from 2011-06-21 to 2012-10-01)}
+\begin{verbatim}
+total valid queries: 569,408
+not well formed queries (e.g. free text or with typing mistakes): 98,923
 
-\subsection*{Most common query patterns (Feb 2012-June 2012)}
-\begin{verbatim}
-total valid queries: 569408
-
-50.16% (285605) :	dataset dataset.name=?
+50.16% (285605):	dataset dataset.name=?
 13.54% (77071) :	site dataset.name=?
@@
 
-Interesting non structured queries:
+Interesting non-valid queries:
 T!_CERN
 *herwig*/AODSIM
+--> keyword search
+lumi file = (file dataset=/RelValProdTTbar/JobRobot-MC_42_V12_JobRobot-v1/GEN-SIM-RECO)
+file,lumi dataset=/RelValProdTTbar/JobRobot-MC_42_V12_JobRobot-v1/GEN-SIM-RECO
+--> Users may like more complex combined queries
+file dataset=/MuEG/Run2011B-PromptReco-v1/AOD, file.size >1
+file dataset=/MinimumBias/Run2010A-valskim-v6/RAW-RECO* | grep run between  [138923, 144086] 
+--> Users mixing up the post and pre filters
+\end{verbatim}
+% 
+% Possible keyword search test cases:
+% run duration run_number between [195100,195200] 
+%
+% py dataset=/DYJetsToLL_M-50_TuneZ2Star_8TeV-madgraph-tarball/Summer12_DR53X-PU_S10_START53_V7A-v1/AODSIM
+%dataset dataset=*Run2011* datatype=data dataset=*RAW*
 
+\appendix
+\section{Data providers statistics}
+\begin{verbatim}
+(Some of the largest ones)
+
+DBS:
+    DB  size: 80GB + 200GB indexes, not many changes to existing  (old) records
+    Largest tables: 
+      Dataset (164K rows) -> Block (2M) -> Files (31M) -> FileRunLumi (902M) <- Runs (65K)
+
+Phedex: ~7GB, more often changes to existing (even old) records
+    change rate: 2,359,934 file transfers last month (from site A to site B; 
+                 change rate on the DB to be found out)
 \end{verbatim}
-% py dataset=/DYJetsToLL_M-50_TuneZ2Star_8TeV-madgraph-tarball/Summer12_DR53X-PU_S10_START53_V7A-v1/AODSIM
-
-\end{document}

### 2012-10-11 doc/thesis/latex/literature_review.tex
updated problem definition, solutions and literature review. review to be finished

@@
 
-\textbf{\color{red}TODO: Overview; evaluate performance}
+\subsection{Overview}
 
-\subsection{Searching 'Deep Web' and Heterogeneous web services}
+The problem of keyword search over \textit{relational and other structured databases} is widely covered,  including even keyword search directly over distributed relational databases\cite[ch.16]{principles_data_integration} and some approaches for automatic schema matching has been discussed also.
+Such search is however very ambiguous and has quite large search space over the possible join paths, therefore most approaches rely on returning the top-k results.
 
-In 1990s there were was much research on search based service integration systems, for instance \textit{Information Manifold} following Local-as-View approach and \textit{TSIMMIS} developed at Stanford following Global-as-View approach.
-The \textit{Information Manifold} is quite resembling to DAS {\color{red},  so it could be useful checking it's papers.
-TODO: finish}
+Regarding the Boolean retrieval that we are the most interested, keyword query could be translated into list of best structured queries (e.g.
+SODA system proposing SQL over complex schemas, while Query Forms would propose mappings from keywords to parameters in SQL templates, Keymantic tries to achieve this without having access to Data itself).
+% TODO: Keymantic would propose SQL queries knowing only schema but no index
+
+%  TODO: Further, {\color{red} in theory,} Google's Deep-web crawl techniques initially designed for accessing content hidden behind html forms could be used to effectively index the contents even of  data providers that do not provide a specific interface to do so.
+
+{\color{red}Current research trends seem to fall on methods of automatic integration (e.g.
+statistical mediator), and pay-as-you-go integration.
+(TODO: see future chapter of \cite{principles_data_integration})}
+
+%\textbf{\color{red}TODO: Overview; evaluate performance}
+
+\subsection{Searching 'Deep Web' and web services}
+
+In 1990s there were was much research on search based service integration systems, for instance \textit{Information Manifold} quite resembling DAS {\color{red},  so it could be useful checking it's papers.}.
+% following Local-as-View approach
+%and \textit{TSIMMIS} developed at Stanford following Global-as-View approach. The \textit{Information Manifold} is
+
+\subsubsection*{No access to index data terms}
+
+\cite{Keymantic10, semantics_without_access} explores the case then there is no possibility to index the data terms, e.g.
+then a DB is behind a wrapper (e.g.
+accessible only through a \textit{Web form} in “Hidden Web” or \textit{a web-service}) then crawling is generally not possible.
+
+In Keymantic\cite{Keymantic10} a 'keyword query is processed as follows: First, all keywords that correspond to metadata items (e.g., field names) are extracted. The remaining
+keywords are considered as possible input fields. Second, the likelihood of a remaining keyword to a metadata item is computed in order to rank different options to execute the keyword query on the
+“Hidden Web” database'\cite[p.942]{ethz2012}.
 
 \subsubsection*{Deep web search at Google}
-[multiple papers from Google] discusses various ways for implementing data integration in terms of large-scale search engine (Google) Virtual integration vs.
-surfacing.
+[multiple papers from Google] discusses various ways for implementing data integration in terms of large-scale search engine (Google): Virtual integration vs.
+Surfacing.
 They also present ways for integrating systems without human intervention through use of statistical 'mediator'.
@@
 
-\textit{Runtime query reformulation} - 'leaves data at the sources and routes queries to appropriate services'\cite[p. 1]{paygo_integration}
+\textit{Runtime query reformulation} - 'leaves data at the sources and routes queries to appropriate services'\cite[p. 1]{webscale_paygo}
 
 \textit{Deep-web surfacing} - tries to add content from the deep-web into search index.
-There are algorithms which allow to iteratively choose input parameters to forms to surface most of the 'hidden' data without loosing much (i.e.
-if choosing parameters in not smart way the web form could yield as many results as a cross product of all input combinations).
+There are algorithms which allow to iteratively choose input parameters to forms to surface a considerable part of the 'hidden' data without large overhead%
+	\footnote{i.e.
+if choosing parameters in not smart way, a web form with just a couple of free inputs (or even dropdowns that's easier case), could yield as many results as a cross product of all input combinations.}.
 
-PayGo approach - there is NO single mediated schema over which users pose queries.
-Queries are routed to to the relevant sources.
-Statistical methods used to model uncertainty at all levels: queries, mappings and underlying data.
+\textit{PayGo approach}\cite{webscale_paygo}  - there is NO single mediated schema over which users pose queries.
+Queries are routed to to the relevant sources with help statistical methods that are used to model uncertainty at all levels: queries, mappings and underlying data.
 
-\subsubsection*{Then there is no access to index data terms}
-{\color{red} TODO: Describe Keymantic\cite{Keymantic10}.
+%{\color{red} TODO: Describe Keymantic\cite{Keymantic10}.
 This is a very suboptimal solution as indexing terms improves results (mention evaluation from SODA paper).
 It works only then keywords map entity names.
+some hybrid approaches:
+%index if exists, regexp, some string similarity measure based on historical data (e.g. even edit-distance would work for many items, like site)}
 
-some hybrid approaches: 
-index if exists
-regexp
-some string similarity measure based on historical data (e.g. even edit-distance would work for many items, like site)}
-
-\subsection{Keyword search over DBs}
+\subsection{Keyword search over relational Database}
 The problem of Keyword search over Relational Databases (or also semi-structured sources like XML) has received a significant attention by the research community over the last decade.
@@
 
-\subsubsection*{Meta-data approach}
+\subsubsection*{SODA: Meta-data approach}
 
-With a goal to bridge the increasing gap between high-level (conceptual, business) and low level (physical) representations of data, researchers from \textit{ETHZ} have been investigating Generation of SQL for Business users at \textit{Credit Suisse}.
+With a goal to bridge the increasing gap between high-level (conceptual, business) and low level (physical) representations of data, researchers from \textit{ETHZ} have been investigating Generation of SQL for Business users over a very complex data warehouse at \textit{Credit Suisse}.
 For converting natural language queries
 % (that in addition to keyword search could convey some semantic structure)
- into SQL statements, in addition to what used by earlier approaches they used meta-data describing the schema (at multiple representation levels) and the domain (ontologies) and some natural language processing.
+ into SQL statements, in addition to what used by earlier approaches they used meta-data describing the schema at both physical, conceptual and logical levels extended with DBpedia (for synonyms, etc) and domain ontologies (to capture business concepts like 'wealthy customer').
+%and some natural language processing
 
-Even on a large data-warehouse of ~220GB data with a complex schema of 400+ tables they reported that if good meta-data is available, generating even quite complex SQL  (n-way joins with aggregations, etc) is quite feasible for computer.
-That would making it 'much easier for business users to interactively explore highly-complex data warehouses' \cite[p.932]{ethz2012}.
+Even on a large data-warehouse of ~220GB data with a complex schema of 400+ tables they reported that if good meta-data is available, generating even fairly complex SQL  (e.g.
+n-way joins with aggregations) is quite feasible for a computer.
+That would make it 'much easier for business users to interactively explore highly-complex data warehouses' \cite[p.932]{ethz2012}.
+The users also reported system's potential a) for analysing the schema and learning patterns about it and b) as tool to help documenting legacy systems.
+
+\subsection{Keyword search: integration on demand}
+{\color{red}TODO?: \cite[ch.16]{principles_data_integration}}
 

### 2012-10-11 doc/thesis/latex/proposed_solutions.tex
updated problem definition, solutions and literature review. review to be finished

@@
 \item How could you identify it (i.e. what do you know)?
-\item After seeing first results: What do you want to do with the results? (list items; select only specific columns; do aggregation; use in command line/another program)
+\item What do you want to do with these results (after seeing the results)? Possible options: list items; select only specific columns; do aggregation; use in command line/another program
 \end{enumerate}
 
+\item Given a keyword query, suggest the best matching structured queries (as a ranked list)%
+ 	\footnote{As DAS uses Boolean IR model we could only rank specific structured queries, but not the results itself.}:
+	\begin{itemize}
+	\item map keywords to entity names and selection attributes (quite resembling 'Query forms' and 'Keyword cleaning' approaches)
+	\item build an inverted index of attribute values with some good full-text search toolkit\footnote{e.g.
+Xapian (http://xapian.org/) implemented in C++ which that is known for its flexibility, or Sphinx (http://sphinxsearch.com/)}.
+If no match found (e.g.
+new entry not yet in our cache) pattern matching could be used to make a best guess.
+	\item ranking could be based on: how closely keywords are matching (some or all) required API's input parameters, popularity of certain queries and users feedback
+
+	\item {\color{red} evaluate different user feedback approaches (e.g. HMM)
+	\item multi-wildcard queries (e.g.
+dataset=*Zmm*special*RECO*) are common at CMS that will need special attention\footnote{Matching to structured query will be harder and guessing even worse.
+Also for performance reasons some web-services will introduce limitations on the format of wilcard queries, while DAS wants still to be flexible at least for datasets}.}
+
+\item making use of direct access to database {\color{red}(if available; in theory)}
+\begin{itemize}
+\item one could do incremental indexing of a target database  \footnote{E.g.
+an inverted index of values in Oracle/MySQL DB tables can be built to keep the list of possible terms.
+For instance, Sphinx full text search engine can access DBs directly and also supports incremental indexing but it needs a bit of manual configuration {http://sphinxsearch.com/docs/current.html\#delta-updates} }, then try to map database table columns into API parameters (manually or even automatically by applying some of Data Integration approaches)
+\item then this index + mapping could be used to improve the keyword search (better mapping from keywords to API inputs) or even structured search (e.g. query cleaning/term auto-completion)
+\item problem: the owners of services may not want to provide direct access to DB because of no trust, security or performance issues
+\item may need to exclude very large tables
+\end{itemize}
+	\end{itemize}
 \item (?) JavaScript-based query interpreter to ease query writing: could suggest search attribute and entity attribute names.
@@
 
-\item Keyword query --> ranked list of matching structured queries:
-	\begin{itemize}
-	\item map keywords to entity name and selection parameters (quite resembling 'Query forms' and 'Keyword cleaning' approaches)
-	\item
-		{\color{red} Inverted index could be built with help of some good full-text search engine (e.g.
-Xapian\footnote{http://xapian.org/}) and used to generate mapping to structured queries.
-If no match found (e.g.
-new entry not yet in our cache) some pattern matching could be used to try to corresponding guess search attribute}
-	\item ranking could be based on: how closely keywords are matching some or all required parameters, popularity of certain query and users feedback
-
-	\item {\color{red}	--> evaluate different approaches (HMM, etc); could some natural language processing help?}
-	\item special attention to \textbf{wilcard} queries
-
-	\end{itemize}
 \end{itemize}
@@
 \subsection{Performance}
-\subsubsection{Continuous view maintenance at Data providers with large DBs}
+
+%%%%%  -------------------------------------------
+
+\subsubsection*{More intelligent caching}
+Some of the data entities instances changing very rarely, for example in DBS system old datasets would never change, while new ones are constantly added (still some of their attributes may change, in this example validity).
+%
+Cached copy may be shown by default while up to date results could be retrieved on user's request.
+An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.
+
+On the other hand some result items may not make sense to be cached, as there may be too many of valid input parameter combinations (e.g. now there are 900M of \textit{(run, file)} combinations).
+
+Also consider:
+                			\begin{itemize}
+                			%\item given an entity received from provider, determining if it is useful to cache for long-term
+%	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
+                			%\item could cache even \textit{expired} data, but warning user
+                			\item could also have different validity dates for certain fields. if no volatile fields are not explicitly requested, even a very old cache could be used. 
+                			\item Can we automatically figure which fields are static and which are changing?
+                			\item pre-fetching common (sub-)queries: determining manually and/or automatically
+                			\end{itemize}
+
+%%%%%  -------------------------------------------
+\subsubsection*{Continuous view maintenance at Data providers with large DBs}
+In the cases when new records are coming not the existing ones are not changing much, continuous view maintenance that computes only differences from earlier results could be a good solution to improve the performance.
+
 Use either \textit{materialized refresh fast views with query rewriting} (Oracle; completely transparent for proprietary apps)\cite{Oracle11}
@@
 
-\subsubsection{More intelligent caching}
-Some of the data entities instances changing very rarely, for example in DBS system old datasets would never change, while new ones are constantly added (still some of their attributes may change, in this example validity).
+For example currently in DBS (80GB + 280GB indexes) to process a query 'find files where run in [r1, r2, r3] and dataset=X' one needs to join all the tables below.
 
-Cached copy may be shown before hand while up to date results could be retrieved on user's request.
+\begin{verbatim}
+Dataset (164K rows) -> Block (2M) -> Files (31M) -> FileRunLumi (902M) <- Runs (65K)
+\end{verbatim}
 
-An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.
+However having a materialized view of all these tables joined together would allow answering such queries much quicker.
+Given low change rates (in comparison to data already present), maintaining the view should be also comparatively cheap with only expense of just couple of times of space.
 
-Also consider:
-                			\begin{itemize}
-                			\item given an entity received from provider, determining if it is useful to cache for long-term
-%	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
-                			\item could cache even \textit{expired} data, but warning user
-                			\item could also have different validity dates for certain fields. if no volatile fields are not explicitly requested, even a very old cache could be used. 
-                			\item Can we automatically figure which fields are static and which are changing?
-                			\item pre-fetching common (sub-)queries: determining manually and/or automatically
-
-                			\item \textbf{Deciding if to cache or not...}
-
-                			\end{itemize}
+%%%%%  -------------------------------------------
 

### 2012-10-11 doc/thesis/latex/report_sept.tex
updated problem definition, solutions and literature review. review to be finished

@@
 A data integration system based on simple structured queries is already in place.
-Still {\color{red}various improvements including support} for less restricted keyword queries and improvements to system's usability and performance still have to be researched.
+Various improvements including support for less restricted keyword queries, improvements to system's usability and performance still have to be researched.
 
@@
          % TODO:is this a good example? conditions is a separate query now...
-         \item find all \textit{datasets} matching some pattern stored at a given \textit{site} (filtering on entities stored on separate services)
+         \item find all \textit{datasets} matching some pattern stored at a given \textit{site} 
+         	  \begin{small}(filtering attributes from separate services)\end{small}
    \end{itemize}                		
@@
 
-The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it queries the data-sources, merges the results, and caches them for subsequent use.
+The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it determines  data-sources are able to answer\footnote{This is done by a mapping between flat mediated schema ('DAS keys') into web-service methods and their arguments.
+Then system queries all services that could provide a result of expected type with given parameters}, queries them, merges the results and caches them for subsequent use.
 DAS uses \textit{Boolean retrieval model} as users are often interested in retrieving ALL the items matching their query.
@@
 Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]) operators.
-The combined query results could be later 'piped' for further filtering and aggregation (min, avg, etc), e.g.:
+The combined query results could be later 'piped' for further filtering, sorting or aggregation (min, max, avg, sum, count, median), e.g.:
 
-{\footnotesize 
+{\small 
 \begin{verbatim}
@@
 \section{Problem statement}
+%TODO: Further it has to comply with dataset constraints
+
+% \textbf{\color{red}DO WE also have proprietary big databases that could be useful to be searched, but the standard integration work would be to heavy (or sub-parts of existing systems that are not covered by APIs, and could be accessible through DBs directly -- these potentially more expensive queries could processed with lower priority)}
+
+\subsection{Ease of use}
+
+For an IR system with wide variety of users, it important to provide an easy to use interface with fairly flat learning curve, while at the same time not loosing support of fairly complex queries.
+%
+	Even a simple structured query language containing entity names over the mediated schema may seem hard to learn%
+	\footnote{On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources).}%
+	, especially in the beginning. Therefore, it shall be useful to guide new users interactively through the process of building of their query.
+%   
+   Supporting non-structured keyword queries is also worth investigation as quite many users reported they are missing this Google-like search experience.
+
+% TODO:	Further a minimum set of search predicates is imposed by APIs (mainly because of performance reasons) and user has to be informed what extra is he expected to provide.
+
+% TODO:	Even in structured query mode users are prone to typos, use either autocompletion where possible or suggest alternative queries.
+
+\subsection{Performance}
 
@@
 Therefore, it is important to balance between returning locally cached query results quickly   and between issuing queries to slow data-services to get up to date results.
-It is also not less important to provide users with easy to use interface having a fairly flat learning curve, while at the same time not loosing the functionality of fairly complex queries.
-% expressive power.
-\textbf{\color{red}Another issue is about handling distributed queries, where either selection criteria or selection values could come from distributed sources -- the biggest problem could be in the APIs they [DO NOT] provide}.
-
-%TODO: Further it has to comply with dataset constraints
-
-% \textbf{\color{red}DO WE also have proprietary big databases that could be useful to be searched, but the standard integration work would be to heavy (or sub-parts of existing systems that are not covered by APIs, and could be accessible through DBs directly -- these potentially more expensive queries could processed with lower priority)}
-
-\subsection{Ease of use: Query Language over the mediated schema}
-	Even a very simple structured query language that also contain entity names over the mediated schema may seem {\color{red}hard} to learn, especially in the beginning.
-   On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources).
-   So some sort of guiding helping user to build the query shall be useful.
-
-   Supporting non-structured keyword queries is also worth investigation as many users reporting they are missing this Google-like search experience.
-{\color{red}Because DAS uses Boolean IR we could only rank specific structured queries.}
-
-	Further a minimum set of search predicates is imposed by APIs (mainly because of performance reasons) and user has to be at least informed what is he expected to provide.
-
-Also to consider: input ambiguity and typos, some queries are more common than others.
-
-\subsection{Performance}
 
 \begin{itemize}
-	     					\item  Query prioritization: {\color{red}maybe we could have prioritization in general, the heavier your query is the more you have to wait in favour of  the light queries}: 
-	     					(we could have some sort of query cost evaluation based on history or manually predefined scores per API)
+						\item Low data providers performance (e.g. on DBS some queries requires joining many large tables)
+	     				\item At the moment all queries are put into one pool and has to wait until some threads are available.
+If a couple of heavy queries were submitted before hand, even very light queries would have to wait long.
+Explore more advanced Query prioritization
+	     					(e.g. we could have some query cost model based on history or predefined scores per API)
 
-                		\item Explore more intelligent caching
+                		\item Currently result items are cached for a fairly short period of time (5min-1h) and then completely discarded, however many entities are not changing that often - Explore more intelligent caching
 
@@
 %                        \item ??? do we have cases then multiple APIs of the same service could be called for the same data, can we choose just one???
-                        \item {\color{red}e.g. intermediary aggregation results while still calculating}
+                        \item 
+                        \begin{small}
+                        {\color{red}e.g.
+displaying intermediary results of aggregation while still calculating with statistical bounds -- however statistical bounds are much meaningful only if items are well snuffled/randomized -- people could stop the query if they see a non-sense}
+                        \end{small}
 
@@
 
-	                	 \item scale testing - if we are storing lots of historical info.
+	              %\item {\color{red}scale testing - if we are storing lots of historical info.
 MongoDB is not so performant if DB cant fit in memory.
 One of well known solutions is installing SSD.
+}
 \end{itemize}
 
-\subsection{Handling distributed search efficiently}
+\subsection{Other}
+\subsubsection*{Handling distributed search efficiently}
+{\color{red} People may be filtering data items where selection criteria could be present on two separate autonomous sources, meaning with the current APIs available the results would have to be fetched separately from each source and ones not contained in both filtered out only afterwards that is suboptimal.}
 
-\pagebreak
+\subsubsection*{Generic connector access relational databases}
+{\color{red}Generic connector able to access relational databases with minimal manual integration could be useful for integrating proprietary systems that were not yet integrated and there's no resources available to build APIs.
+A possible use-case could be the \textit{Prep} database, but this still has to be discussed.
+
+Depending on needs it could be either of exploratory or API/Query Forms based approach, where fake APIs are defined in terms of SQL queries (or even better in a simplified form that would generate SQL by taking schema into account)}
+
 \include{proposed_solutions}
 
-\pagebreak
 \include{literature_review}
@@
 \section{Work status}
-TODO
 
@@
 - obtained DB copy of biggest data provider DBS (currently 80 GB + 200 GB indexes)
-- preliminary literature review (to be continued deeper + waiting for book arrival: Principles of Data Integration)
-
---- some analysis of DAS logs
+- preliminary literature review (to be continued deeper)
+- initial analysis of DAS logs
 
@@
 - couple of fairly simple prototypes of UI/access patters for simpler DAS querying
-- check on performance improvements after creation of materialized view(s)
-- interview (more) DAS users
+- check performance improvements after creation of materialized view(s)
+- interview more DAS users
 \end{verbatim}
@@
 \thispagestyle{empty}
+\begin{small}
 \bibliography{refs}
+\end{small}
 
 \pagebreak
-\appendix
-\section{DAS Query logs}
-
-non parsible queries: 98923
-
-\subsection*{Most common query patterns (Feb 2012-June 2012)}
-\begin{verbatim}
-total valid queries: 569408
-
-50.16% (285605) :	dataset dataset.name=?
-13.54% (77071)  :	site dataset.name=?
-8.96%  (51035) 	:	file dataset.name=?
-5.88%  (33504) 	:	run dataset.name=?
-2.59%  (14739) 	:	release dataset.name=?
-2.11%  (12036) 	:	config dataset.name=?
-1.65%  (9420) 	:	dataset run.run_number=?
-1.34%  (7642) 	:	block dataset.name=?
-1.24%  (7084) 	:	dataset site.name=?
-1.15%  (6562) 	:	dataset dataset.name=? release.name=?
-1.10%  (6287) 	:	parent dataset.name=?
-0.96%  (5488) 	:	file file.name=?
-0.92%  (5245) 	:	dataset dataset.name=? status.name=?
-0.77%  (4363) 	:	dataset release.name=?
-0.75%  (4257) 	:	file dataset.name=? run.run_number=?
-0.68%  (3874) 	:	run run.run_number=?
-0.53%  (2994) 	:	site site.name=?
-0.45%  (2576) 	:	site file.name=?
-0.45%  (2556) 	:	dataset file.name=?
-0.43%  (2438) 	:	lumi file.name=?
-0.40%  (2282) 	:	dataset dataset.name=? site.name=?
-0.35%  (1999) 	:	file block.name=?
-0.35%  (1970) 	:	dataset dataset.name=? run.run_number=?
-0.29%  (1640) 	:	group dataset.name=?
-0.29%  (1631) 	:	lumi run.run_number=?
-0.25%  (1402) 	:	child dataset.name=?
-0.22%  (1268) 	:	run file.name=?
-0.21%  (1204) 	:	block block.name=?
-0.19%  (1088) 	:	release release.name=?
-0.16%  (936) 	:	site block.name=?
-0.12%  (703) 	:	file dataset.name=? lumi.number=? run.run_number=?
-0.10%  (594) 	:	parent file.name=?
-0.08%  (455) 	:	dataset block.name=?
-0.06%  (365) 	:	dataset dataset.name=? datatype.name=? release.name=?
-0.06%  (360) 	:	file dataset.name=? site.name=?
-
-Interesting non structured queries:
-T!_CERN
-*herwig*/AODSIM
-
-\end{verbatim}
-% py dataset=/DYJetsToLL_M-50_TuneZ2Star_8TeV-madgraph-tarball/Summer12_DR53X-PU_S10_START53_V7A-v1/AODSIM
+\include{das_logs}
 

### 2012-10-11 doc/thesis/latex/literature_review.tex
updated problem definition, solutions and literature review. review to be finished

@@
 
-\textit{PayGo approach}\cite{webscale_paygo}  - there is NO single mediated schema over which users pose queries.
+\textit{PayGo approach}\cite{webscale_paygo}  - With this approach, 'a system starts with
+very few (or inaccurate) semantic mappings and these mappings
+are improved over time as deemed necessary'.
+% TODO: cite{bootstraping}
+there is NO single mediated schema over which users pose queries.
 Queries are routed to to the relevant sources with help statistical methods that are used to model uncertainty at all levels: queries, mappings and underlying data.

### 2012-10-11 doc/thesis/latex/proposed_solutions.tex
updated problem definition, solutions and literature review. review to be finished

@@
 %%%%%  -------------------------------------------
-\subsubsection*{Continuous view maintenance at Data providers with large DBs}
+\subsubsection*{Bad data providers performance}
 In the cases when new records are coming not the existing ones are not changing much, continuous view maintenance that computes only differences from earlier results could be a good solution to improve the performance.

### 2012-10-11 doc/thesis/latex/proposed_solutions.tex
applied valentins patch

@@
 
-	\item {\color{red} evaluate different user feedback approaches (e.g. HMM)
+	\item {\color{red} TODO: evaluate different user feedback approaches (e.g. HMM) {\color{green} please elaborate how?}
 	\item multi-wildcard queries (e.g.
@@
 also taking into account some possible ambiguous namings (auto completion)
+
++\item {\color{green} Build series of queries which will answer the given user question, e.g.
+user asks {\it I want to find files which contains runs taken with magnetic field 4T for given dataset at site X}.
+This should be decomposed somehow (subject to research) into series of quries, e.g.
+Find datasets at site X; find files for given dataset list; find runs for found files; filter runs with magnetic field 4T}
 
@@
                 			\item Can we automatically figure which fields are static and which are changing?
+                			 {\color{green} Yes, if we enfore APIs to record modification timestamp}
+                			 {\color{red} I also meant some automatic algorithm, however that would be more complex and less reliable}
                 			\item pre-fetching common (sub-)queries: determining manually and/or automatically

### 2012-10-11 doc/thesis/latex/report_sept.tex
updated problem definition, solutions and literature review. review to be finished
applied valentins patch

@@
 	Even a simple structured query language containing entity names over the mediated schema may seem hard to learn%
-	\footnote{On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources).}%
+%	\footnote{On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources).}%
+	\footnote{On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} {\color{green} Yes, one of the task for DAS is to provide consistent naming convention for end-users} (even though there may exist slight differences in their naming on different data-sources).}%
 	, especially in the beginning. Therefore, it shall be useful to guide new users interactively through the process of building of their query.
@@
 \begin{itemize}
-						\item Low data providers performance (e.g. on DBS some queries requires joining many large tables)
+						\item Bad data providers performance (e.g. on DBS some queries requires joining many very large tables)
 	     				\item At the moment all queries are put into one pool and has to wait until some threads are available.
@@
 displaying intermediary results of aggregation while still calculating with statistical bounds -- however statistical bounds are much meaningful only if items are well snuffled/randomized -- people could stop the query if they see a non-sense}
+                          {\color{green} Partial results are problematic to construct in DAS, since DAS does the aggregation across multiple dataprovider.
+To do so DAS needs to fetch all results.
+Although it would be possible to do partial results for data coming from a single data-source.}
                         \end{small}
@@
 {\color{red} People may be filtering data items where selection criteria could be present on two separate autonomous sources, meaning with the current APIs available the results would have to be fetched separately from each source and ones not contained in both filtered out only afterwards that is suboptimal.}
++{\color{green} Please elaborate more, I don't really follow.}
 
@@
 Depending on needs it could be either of exploratory or API/Query Forms based approach, where fake APIs are defined in terms of SQL queries (or even better in a simplified form that would generate SQL by taking schema into account)}
-
++{\color{green} Indeed, an dthat was the topic of research for another DAS student, Liang Dong.
+The idea was to generalize DBS-QL to any DB back-end and provide common package which will do a job.
+The work is done over here https://github.com/vkuznet/PyQueryBuilder.
+Feel free to check it out and we can discuss it more.}
++
 \include{proposed_solutions}

### 2012-10-18 doc/thesis/latex/literature_review.tex
fixes and some test to wildcard matching; updated report (literature review, proposed solutions etc)

--- 2012-10-11 doc/thesis/latex/literature_review.tex
@@
 
-The problem of keyword search over \textit{relational and other structured databases} is widely covered,  including even keyword search directly over distributed relational databases\cite[ch.16]{principles_data_integration} and some approaches for automatic schema matching has been discussed also.
-Such search is however very ambiguous and has quite large search space over the possible join paths, therefore most approaches rely on returning the top-k results.
+% Data Integration is about balancing between Virtual integration and data warehousing.
 
-Regarding the Boolean retrieval that we are the most interested, keyword query could be translated into list of best structured queries (e.g.
-SODA system proposing SQL over complex schemas, while Query Forms would propose mappings from keywords to parameters in SQL templates, Keymantic tries to achieve this without having access to Data itself).
-% TODO: Keymantic would propose SQL queries knowing only schema but no index
+Since the late 1990's, several Enterprise Information Integration\footnote{%
+	Enterprise Information Integration (EII) is about 'integrating data from 
+	multiple sources 	\textit{without} having to to first load data into
+	 a central warehouse'\cite[p.1]{eii_2005}}
+ (EII) products have appeared in the market (e.g.
+\textit{Information Manifold} by AT\&T Lab) and an significant experience has been accumulated on: data integration formalisms, ways of describing of heterogeneous data sources and their abilities (e.g.
+RDBMS vs web form), query optimization (combining sources efficiently, source overlap, data quality, etc)\cite{eii_2005}.
+%
+Recent research in data integration mostly focused on methods of automatic, uncertainty and feedback  driven (e.g.
+though statistical mediator) semi-automatic pay-as-you-go approaches\cite[ch.19]{principles_data_integration}).
 
-%  TODO: Further, {\color{red} in theory,} Google's Deep-web crawl techniques initially designed for accessing content hidden behind html forms could be used to effectively index the contents even of  data providers that do not provide a specific interface to do so.
+The problem of keyword search over structured sources received significant attention within the last decade.
+Keyword search over relational, XML and other databases was explored form a number of perspectives: returning top-k ranked result tuples vs suggesting structured queries as SQL, performance optimization, user feedback mechanisms, as well as keyword searching over distributed sources.
+It is worth noting that when there is no need for 100\% result exactness keyword search combined with probabilistic schema matching allows lightweight (pay-as-you-go) data integration with minimal human effort upfront\cite[ch.16]{principles_data_integration} but could improve its quality with time by using users' feedback.
 
-{\color{red}Current research trends seem to fall on methods of automatic integration (e.g.
-statistical mediator), and pay-as-you-go integration.
-(TODO: see future chapter of \cite{principles_data_integration})}
+In the setting of Boolean Keyword-based Retrieval over heterogeneous sources a keyword query could be translated into a ranked list of structured queries (similarly as SODA system that propose SQL, Query Forms that propose SQL templates or Keymantic trying to achieve this without accessing the data).
+In the extreme case of having no control over a web-service that do not publish its contents, techniques like Google's Deep-web crawl could index a subset of its contents enabling keyword search to some extent.
 
-%\textbf{\color{red}TODO: Overview; evaluate performance}
+\subsection{Composing multiple data sources efficiently: Query translation using Logical Views}
+The \textit{Information Manifold}\cite{Levy96} is virtual integration (EII) system that uses CARIN dialect of \textit{description logics} (could be though as simplified datalog) for queries and for source descriptions.
+It describes each source as logical view over the global (mediated) schema, augmented with source capabilities (e.g.
+what are possible and the required input parameters for a view to do selections\footnote{this makes these logical views quite similar to source APIs used by DAS, with difference that DAS currently only describe the parameters APIs and only partially the results (as an entity type)}).
+%
+This allows designing algorithms that could allow efficiently answering complex queries that require composing multiple the data sources by finding maximally contained rewriting of (conjunctive) input query in terms of logical views provided by sources (that is, finding an optimal way to compose the sources).
 
-\subsection{Searching 'Deep Web' and web services}
+For example (based on \cite{integr_views2000}), consider such sources as expressed as views (on the left) in terms of global predicates of the mediated schema (on the right) in datalog notation:
+%
+{\small
+\begin{verbatim}
+v1(E, P, M) :- emp(E) & phone(E, P) & mgr(E, M). # employees, their phones and managers
+v2(E, O, D) :- emp(E) & office(E, O) & dept(E, D). # offices and departments of employees
+v3(E, P) :- emp(E) & phone(E, P) & dept(E, toy_dept). # phones of employees only in Toys dept.
+\end{verbatim}
+}
+%
+Suppose we wanted to know Sally's phone and office. We express this by datalog in global predicates:
+{\small\begin{verbatim}
+q1(P,O) :- phone(sally,P) & office(sally,O).
+\end{verbatim}}
+There are two minimal solutions (as the sources could be incomplete the full solutionis union of the two):
+{\small
+\begin{verbatim}
+answer1(P,O) :- v1(sally,P,M) & v2(sally,O,D).
+answer2(P,O) :- v3(sally,P) & v2(sally,O,D).
+\end{verbatim}}
+Notice that the expansions of these solutions (e.g.
+answer1\_exp) are not equivalent to $q_1$, but the conjunctive queries that are closest and still contained in $q_1$ (as these are the only usable views at sources):
+{\footnotesize\begin{verbatim}
+answer1_exp(P,O) :- emp(sally) & phone(sally,P) & mgr(sally,M) & emp(sally) & office(sally,O) & dept(sally,D).
+\end{verbatim}
+}
+After this \textit{Information Manifold} would find an executable order that adheres the capabilities of the sources, by iteratively considering any sources whose input parameters are satisfied. 
+% \subsection{Keyword search: integration on demand}
+% {\color{red}TODO?: \cite[ch.16]{principles_data_integration}}
 
-In 1990s there were was much research on search based service integration systems, for instance \textit{Information Manifold} quite resembling DAS {\color{red},  so it could be useful checking it's papers.}.
-% following Local-as-View approach
-%and \textit{TSIMMIS} developed at Stanford following Global-as-View approach. The \textit{Information Manifold} is
-
-\subsubsection*{No access to index data terms}
-
-\cite{Keymantic10, semantics_without_access} explores the case then there is no possibility to index the data terms, e.g.
-then a DB is behind a wrapper (e.g.
-accessible only through a \textit{Web form} in “Hidden Web” or \textit{a web-service}) then crawling is generally not possible.
-
-In Keymantic\cite{Keymantic10} a 'keyword query is processed as follows: First, all keywords that correspond to metadata items (e.g., field names) are extracted. The remaining
-keywords are considered as possible input fields. Second, the likelihood of a remaining keyword to a metadata item is computed in order to rank different options to execute the keyword query on the
-“Hidden Web” database'\cite[p.942]{ethz2012}.
-
-\subsubsection*{Deep web search at Google}
-[multiple papers from Google] discusses various ways for implementing data integration in terms of large-scale search engine (Google): Virtual integration vs.
-Surfacing.
-They also present ways for integrating systems without human intervention through use of statistical 'mediator'.
-
-There two approaches to web scale search for deep-web (Google mostly cares about web forms):
-
-\textit{Runtime query reformulation} - 'leaves data at the sources and routes queries to appropriate services'\cite[p. 1]{webscale_paygo}
-
-\textit{Deep-web surfacing} - tries to add content from the deep-web into search index.
-There are algorithms which allow to iteratively choose input parameters to forms to surface a considerable part of the 'hidden' data without large overhead%
-	\footnote{i.e.
-if choosing parameters in not smart way, a web form with just a couple of free inputs (or even dropdowns that's easier case), could yield as many results as a cross product of all input combinations.}.
-
-\textit{PayGo approach}\cite{webscale_paygo}  - there is NO single mediated schema over which users pose queries.
-Queries are routed to to the relevant sources with help statistical methods that are used to model uncertainty at all levels: queries, mappings and underlying data.
-
-%{\color{red} TODO: Describe Keymantic\cite{Keymantic10}.
-This is a very suboptimal solution as indexing terms improves results (mention evaluation from SODA paper).
-It works only then keywords map entity names.
-some hybrid approaches:
-%index if exists, regexp, some string similarity measure based on historical data (e.g. even edit-distance would work for many items, like site)}
-
-\subsection{Keyword search over relational Database}
+\subsection{Keyword search over a Relational Database}
 The problem of Keyword search over Relational Databases (or also semi-structured sources like XML) has received a significant attention by the research community over the last decade.
@@
 \cite{forms_kws} proposes alternative approach for processing keyword queries over relational databases: given a keyword query, instead of returning database tuples one could rank query forms that best matches the query for user to choose the right one (if they are properly named this is fairly easy).
-The ranking is based on checking matching of keywords to table names in templates and to column values (could be implemented with inverted index).\textbf{\color{red}TODO: more detailed and our limitations (after reading keyword cleaning)}.
+The ranking is based on checking matching of keywords to table names in templates and to column values.
 
@@
 
-\subsection{Keyword search: integration on demand}
-{\color{red}TODO?: \cite[ch.16]{principles_data_integration}}
+\subsection{Keyword search over web services}
+
+\subsubsection*{What if there is no access to index data terms}
+\cite{Keymantic10, semantics_without_access} explores the case then there is no possibility to index the data terms, e.g.
+then a DB is behind a wrapper (e.g.
+accessible only through a \textit{Web form} in “Hidden Web” or \textit{a web-service}) then crawling is generally not possible.
+%
+In Keymantic\cite{Keymantic10} a 'keyword query is processed as follows: First, all keywords that  correspond to metadata items (e.g., field names) are extracted. The remaining
+keywords are considered as possible input fields. Second, the likelihood of a remaining keyword to a metadata item is computed in order to rank different options to execute the keyword query on the
+“Hidden Web” database'\cite[p.942]{ethz2012}.
+
+\subsubsection*{Deep web search at Google}
+%[multiple papers from Google] discusses various ways for implementing data integration in terms of large-scale search engine (Google): Virtual integration vs.
+Surfacing.
+They also present ways for integrating systems without human intervention through use of statistical 'mediator'.
+
+There are two approaches to web scale search over deep-web (here Google mostly cares about web forms): \marginpar{\scriptsize\color{red}TODO: how does google use PayGo?} 
+%\textit{Runtime query reformulation} - 'leaves data at the sources and routes queries to appropriate services'\cite[p. 1]{webscale_paygo}
+
+\textit{Deep-web surfacing} - surface deep-web (e.g.
+web forms) adding their results into the standard search index  easily allowing to using existing IR technology that scales well.
+There exist algorithms which allow to iteratively choose input parameters to the forms to surface a considerable part of the 'hidden' data without large overhead%
+	\footnote{i.e.
+if choosing parameters in not smart way, a web form with just a couple of free inputs (or even dropdowns that's easier case), could yield as many results as a cross product of all input combinations.}.
+
+\textit{Pay-as-you-go approach}\cite{webscale_paygo} - With this approach, 'a system starts with
+very few (or inaccurate) semantic mappings and these mappings are improved over time as deemed necessary';
+% TODO: cite{bootstraping}
+there is NO single mediated schema over which users pose queries: queries are routed to to the relevant sources with help statistical methods that are used to model uncertainty at all levels: queries, mappings and underlying data.
+
+%{\color{red} TODO: Describe Keymantic\cite{Keymantic10}.
+This is a very suboptimal solution as indexing terms improves results (mention evaluation from SODA paper).
+It works only then keywords map entity names.
+some hybrid approaches:
+%index if exists, regexp, some string similarity measure based on historical data (e.g. even edit-distance would work for many items, like site)}
 

### 2012-10-18 doc/thesis/latex/proposed_solutions.tex
fixes and some test to wildcard matching; updated report (literature review, proposed solutions etc)

--- 2012-10-11 doc/thesis/latex/proposed_solutions.tex
@@
 	\begin{itemize}
-	\item map keywords to entity names and selection attributes (quite resembling 'Query forms' and 'Keyword cleaning' approaches)
-	\item build an inverted index of attribute values with some good full-text search toolkit\footnote{e.g.
-Xapian (http://xapian.org/) implemented in C++ which that is known for its flexibility, or Sphinx (http://sphinxsearch.com/)}.
-If no match found (e.g.
-new entry not yet in our cache) pattern matching could be used to make a best guess.
+	\item the algorithm will generate mappings between each keyword to either an entity name, selection attribute name or its value (resembling 'Query forms' and 'Keyword cleaning' approaches)
 	\item ranking could be based on: how closely keywords are matching (some or all) required API's input parameters, popularity of certain queries and users feedback	
-
-	\item {\color{red} evaluate different user feedback approaches (e.g. HMM)
+	\item build an inverted index of attribute values with some good full-text search toolkit%
+				\footnote{e.g. Xapian (http://xapian.org/) implemented in C++ which that is known for its flexibility, or Sphinx (http://sphinxsearch.com/)}. 
+	If no match found (e.g. new entry not yet in our cache) pattern matching could be used to make a best guess
+	\item the system could improve its results from users' feedback (e.g. by HMM, will need to decide)
 	\item multi-wildcard queries (e.g.
 dataset=*Zmm*special*RECO*) are common at CMS that will need special attention\footnote{Matching to structured query will be harder and guessing even worse.
-Also for performance reasons some web-services will introduce limitations on the format of wilcard queries, while DAS wants still to be flexible at least for datasets}.}
+Also for performance reasons some web-services will introduce limitations on the format of wilcard queries, while DAS wants still to be flexible at least for datasets}.
 
-\item making use of direct access to database {\color{red}(if available; in theory)}
+\item knowing all possible APIs parameter values would be beneficial ({\color{red}but not yet sure if available}%
+		\footnote{problem: the owners of services may not want to provide direct access to DB because of no trust, security or performance issues.
+Otherwise, specific APIs could be developed to return possible parameter values but this is quite an overhead}) and could be used to improve the keyword search (better mapping from keywords to API inputs) or even the structured search (e.g.
+query cleaning/term auto-completion)
 \begin{itemize}
-\item one could do incremental indexing of a target database  \footnote{E.g.
+\item having direct access to DB, one could incrementally index target database\footnote{%
+			E.g.
 an inverted index of values in Oracle/MySQL DB tables can be built to keep the list of possible terms.
-For instance, Sphinx full text search engine can access DBs directly and also supports incremental indexing but it needs a bit of manual configuration {http://sphinxsearch.com/docs/current.html\#delta-updates} }, then try to map database table columns into API parameters (manually or even automatically by applying some of Data Integration approaches)
-\item then this index + mapping could be used to improve the keyword search (better mapping from keywords to API inputs) or even structured search (e.g. query cleaning/term auto-completion)
-\item problem: the owners of services may not want to provide direct access to DB because of no trust, security or performance issues
-\item may need to exclude very large tables
+For instance, Sphinx full text search engine can access DBs directly and also supports incremental indexing but it needs a bit of manual configuration {http://sphinxsearch.com/docs/current.html\#delta-updates} }%
+	tables mapping it's columns into API parameters (manually or automatically).
+
+%\item may need to exclude very large tables
 \end{itemize}
@@
 also taking into account some possible ambiguous namings (auto completion)
+\end{itemize}
 
-\end{itemize}
+\subsection{Supporting complex queries}
+As most entities in the mediated schema (27 of them) are already mapped to multiple services (see the mapping here: \url{https://cmsweb.cern.ch/das/services}) these taken as a graph could be used to build (some of) the more complex queries.
 
@@
 \subsubsection*{More intelligent caching}
-Some of the data entities instances changing very rarely, for example in DBS system old datasets would never change, while new ones are constantly added (still some of their attributes may change, in this example validity).
+Some of the data entities instances changes very rarely.
+For example in DBS system old datasets would never change, while new entities are constantly added (still some of their attributes may change, e.g.
+validity of a dataset).
 %
@@
 An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.
-
+%
 On the other hand some result items may not make sense to be cached, as there may be too many of valid input parameter combinations (e.g. now there are 900M of \textit{(run, file)} combinations).
 
-Also consider:
+Also consider that:
                 			\begin{itemize}
@@
                 			%\item could cache even \textit{expired} data, but warning user
-                			\item could also have different validity dates for certain fields. if no volatile fields are not explicitly requested, even a very old cache could be used. 
-                			\item Can we automatically figure which fields are static and which are changing?
-                			\item pre-fetching common (sub-)queries: determining manually and/or automatically
+                			\item we could also have different validity dates for certain fields. if no volatile fields are not explicitly requested, even a very old cache could be used. 
+                			% \item Can we automatically figure which fields are static and which are changing?
+			 {\color{green} Yes, if we enfore APIs to record modification timestamp} {\color{red} I also meant some automatic algorithm, however that would be more complex and less reliable}
+                			\item pre-fetching common queries: determining manually and/or automatically
                 			\end{itemize}
@@
 %%%%%  -------------------------------------------
-\subsubsection*{Bad data providers performance}
-In the cases when new records are coming not the existing ones are not changing much, continuous view maintenance that computes only differences from earlier results could be a good solution to improve the performance.
+\subsubsection*{Data providers performance issues}
+In the cases when new records are coming but the existing ones are not changing much, continuous view maintenance that computes only differences from earlier results could be a good solution to improve the performance especially for often used queries.
 
@@
 
-For example currently in DBS (80GB + 280GB indexes) to process a query 'find files where run in [r1, r2, r3] and dataset=X' one needs to join all the tables below.
+This is exactly the case for the most popular expensive query (see \hyperref[run_dataset_heavy]{query \#\ref{run_dataset_heavy} in \ref{appendix_das_cli_logs} DAS log analysis}) over DBS system (80GB + 280GB indexes): 'find files where run in [r1, r2, r3] and dataset=X' that needs to join all the biggest tables in that database.
 
@@
 
-However having a materialized view of all these tables joined together would allow answering such queries much quicker.
+Having a materialized view with all these tables joined together would allow answering such queries much quicker.
 Given low change rates (in comparison to data already present), maintaining the view should be also comparatively cheap with only expense of just couple of times of space.
@@
 
-\subsection{Integrating distributed information efficiently}
+\subsubsection*{Integrating distributed information efficiently}
 Bloom-join (which could be quite transparent and implemented even on DB side [pure SQL is possible for MySQL, to check for Oracle]- take a query and bit-vector as parameter) , lazy pagination (and order required for aggregation) - this is not yet supported by any of the data-service APIs

### 2012-10-18 doc/thesis/latex/report_sept.tex
fixes and some test to wildcard matching; updated report (literature review, proposed solutions etc)

@@
 \usepackage[cm]{fullpage}
+
+\addtolength{\topmargin}{-5mm}
+\addtolength{\topskip}{-5mm}
+
 %\usepackage[margin=0.5in]{geometry}
-%\usepackage{times}
+%\usepackage{sans}
+\usepackage{paralist}
 \usepackage{color}
+
+% enable Hyperlinks
+\usepackage{hyperref}
+% fancy verbatim
+\usepackage{fancyvrb}
+\usepackage{appendix}
+
+%\usepackage{alltt}
+
 %\usepackage[small,compact]{titlesec}
@@
 \section{Problem statement}
-%TODO: Further it has to comply with dataset constraints
-
-% \textbf{\color{red}DO WE also have proprietary big databases that could be useful to be searched, but the standard integration work would be to heavy (or sub-parts of existing systems that are not covered by APIs, and could be accessible through DBs directly -- these potentially more expensive queries could processed with lower priority)}
-
-\subsection{Ease of use}
+\subsection{User Interface and ease of use}
 
@@
 %
-	Even a simple structured query language containing entity names over the mediated schema may seem hard to learn%
+	Even a simple structured query language and entity names over the mediated schema  at first may seem hard to learn%
 %	\footnote{On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources).}%
-	\footnote{On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} {\color{green} Yes, one of the task for DAS is to provide consistent naming convention for end-users} (even though there may exist slight differences in their naming on different data-sources).}%
-	, especially in the beginning. Therefore, it shall be useful to guide new users interactively through the process of building of their query.
+	\footnote{on the other hand, at CERN, the names in the mediated schema are refer to entities that in the real-world are fairly consistently named (though there may exist slight differences in their naming, whereas DAS is to provide the 'golden' naming convention).}.
+Therefore, it shall be useful to guide new users interactively through the process of building of their query.
 %   
    Supporting non-structured keyword queries is also worth investigation as quite many users reported they are missing this Google-like search experience.
+
+\subsection{Supporting complex queries}
+Currently DAS can only process queries that could be directly mapped to data providers' APIs, but not the queries where results of one API need to be fed into another APIs 
+	%\footnote{this is partially because of performance reasons {\color{red}and the fact that DAS is mainly schema-less (the mediated schema is only in terms of entities, their mappings to APIs and API parameters)}}%
+	\footnote{except a couple of manually hard-coded "views" in a virtual \textit{"combined"} data provider}%
+. %In such case, the user  is forced to write some code to submit and process the sub-queries himself. 
+Further, to be flexible DAS only keeps a list of 'mediated entities' and how they could be retrieved, but do not enforce any predefined schema (i.e. entity fields/attributes) nor exposes it\footnote{%
+	Actually the fields of each API  (an entity could be mapped to a number of these) could be inferred 
+	from past queries or by predefined 'bootstrap' queries. Currently the field lists may differ slightly depending on the parameters}
+before querying for some particular entity(-ies). 
+%
+This makes it fairly hard to execute complex queries  without a priori knowledge.
+User has to know what queries and how have to be combined (depends on the mediated schema and what APIs are available), and has either to combine the results by hand or to write a program does that.
+%
+Some of the questions to be answered:
+
+	\begin{compactitem}
+		\item could we select  meaningful and sufficiently efficient composition of services even using less detailed source descriptions than those used in Information Manifold (see Literature Review)?
+		\item could we use the current descriptions that only define entity the result and API parameters?
+%		\item how to present the results (as a sequence of DAS queries or implemented within DAS)
+	\end{compactitem}
+
+%+\item {\color{green} Build series of queries which will answer the given user question, e.g.
+user asks {\it I want to find files which contains runs taken with magnetic field 4T for given dataset at site X}.
+This should be decomposed somehow (subject to research) into series of quries, e.g.
+Find datasets at site X; find files for given dataset list; find runs for found files; filter runs with magnetic field 4T}
+
+% It would be nice to {\color{green} Build series of queries which will answer the given user question, e.g.
+user asks {\it I want to find files which contains runs taken with magnetic field 4T for given dataset at site X}.
+This should be decomposed somehow (subject to research) into series of quries, e.g.
+Find datasets at site X; find files for given dataset list; find runs for found files; filter runs with magnetic field 4T}
 
@@
 datasets stored at DBS do not change often).
-Further as the total data that is being queried could be fairly large (with growth of order of 1TB per year) and the data-services may have certain constraints (e.g.
-only certain queries allowed) as their main goal is supporting production (data taking from the CMS detector).
-
-Therefore, it is important to balance between returning locally cached query results quickly   and between issuing queries to slow data-services to get up to date results.
-
-\begin{itemize}
-						\item Bad data providers performance (e.g. on DBS some queries requires joining many very large tables)
+%
+Further as the total data that is stored at data-providers could be fairly large (order of 1TB per year) and as their main goal is supporting production (data taking from the CMS detector), the providers may impose some constraints (e.g.
+only certain queries allowed).
+%
+Therefore, it is important to balance between returning locally cached query results quickly   and between issuing queries to slow data-services to get fresh results.
+%
+Issues to be addressed includes:
+\begin{compactitem}
+						\item Low data providers' performance (e.g. on DBS some queries requires joining many very large tables)
 	     				\item At the moment all queries are put into one pool and has to wait until some threads are available.
-If a couple of heavy queries were submitted before hand, even very light queries would have to wait long.
+If a couple of heavy queries were submitted earlier, even light queries would have to wait long.
 Explore more advanced Query prioritization
@@
 
-					\item Could displaying Partial Results improve performance?
-					\begin{itemize}
-					     \item given current APIs to do aggregation over data sources, DAS has to fetch ALL records matching the query instead of only the first page.
-%                        \item ??? do we have cases then multiple APIs of the same service could be called for the same data, can we choose just one???
-                        \item 
-                        \begin{small}
-                        {\color{red}e.g.
+					\item Since DAS does aggregation across multiple data providers, given their current APIs (with no ordering and limit functionality), DAS has to fetch ALL records matching the query instead of only the first page.
+Although it would be possible to do partial results for data coming from a single data-source (while still loading in background, or the providers (or us) would have to modify their APIs).
+
+					\item Efficient distributed search: as filtering criteria may reside on two or more autonomous sources, given the current APIs much more items than in the final result may need to be fetched from each source.
+Items that do not meet all selection criteria could be filtered out only afterwards in DAS.
+To improve the performance, data provider's APIs for the most used queries may need to be redesigned.
+                        % \item {\color{red}e.g.
 displaying intermediary results of aggregation while still calculating with statistical bounds -- however statistical bounds are much meaningful only if items are well snuffled/randomized -- people could stop the query if they see a non-sense}
-                          {\color{green} Partial results are problematic to construct in DAS, since DAS does the aggregation across multiple dataprovider.
-To do so DAS needs to fetch all results.
-Although it would be possible to do partial results for data coming from a single data-source.}
-                        \end{small}
-
-					\end{itemize}
-
 	              %\item {\color{red}scale testing - if we are storing lots of historical info.
@@
 }
-\end{itemize}
+\end{compactitem}
 
-\subsection{Other}
-\subsubsection*{Handling distributed search efficiently}
-{\color{red} People may be filtering data items where selection criteria could be present on two separate autonomous sources, meaning with the current APIs available the results would have to be fetched separately from each source and ones not contained in both filtered out only afterwards that is suboptimal.}
-+{\color{green} Please elaborate more, I don't really follow.}
-
-\subsubsection*{Generic connector access relational databases}
+\subsection{Generic connector accessing relational databases}
 {\color{red}Generic connector able to access relational databases with minimal manual integration could be useful for integrating proprietary systems that were not yet integrated and there's no resources available to build APIs.
@@
 +
-\include{proposed_solutions}
+
+\input{proposed_solutions}
 

### 2012-10-18 src/python/DAS/core/dataset_match_wildcards.py

@@ -1,84 +1,169 @@
-"""
-The current algorithm is simple
-1) Fetch all the matches (regexp from MongoDB)
-2) for each of them check is what wildcard (*) matched has a slash [running again the regexp in python]
-	if so: we will replace * in initial pattern with '*/*'
-	otherwise: we leave it as it was
-
-	track all these possible replacements and their counts, and finally apply them
-
-	# possible tune ups (however would not match super new datasets (could be fixed by loading datasets very often and Incrementaly!):
-	if all matches for a certain replacement option contain the same string:
-		replace it by that string simplifying the expression further and saving time at the provider!
-
-	e.g. for '*Zmm*special*RECO*' would give: /RelValZmm/*/*special*RECO*
-	while '*Zmm*' would still give: */*/*Zmm*
-"""
 import pymongo
 from pymongo import Connection
 
+# TODO: use global connection
 conn = Connection('localhost', 8230)
 
 import re
-pattern = '*Zmm*special*RECO*'
 
 
 REPLACE_IF_STRINGS_SAME = 1
-
-# TODO: ^ $
-pat_re = pattern.replace('*', '(.*)')
-pat_re_compiled = re.compile(pat_re, re.IGNORECASE)
-
-# DAS regexp? actually it's only 8MB ;)
-
-# 3seconds
-results = [item for item in conn.dbs.cms_dbs_prod_global.find() if pat_re_compiled.match(item['dataset'])]
-
-db = conn.dbs
-# TODO: escape regexp!!!
-# takes around 2secs for 700results but that is quite normal (Zmm) because of MongoDB
-r = db.cms_dbs_prod_global.find({ 'dataset' : { '$regex' : pat_re, '$options': 'i' } } );
-r = [item['dataset'] for item in r]
-
-# TODO: we could also get rid of empty * (if none of them have items)
-# also if all entries all the same, we could replace by the string
-
-# now match the positions of slash
-options = {}
-options_matches = {}
-for item in r:
-	groups = pat_re_compiled.match(item).groups()
-	print groups
-	replacements = tuple([ ('/' in group) and '*/*' or '*'  for group in groups])
-	options[replacements] = options.get(replacements, 0) + 1
-	# add this into list of possible options
-	updated = options_matches.get(replacements, [])
-	updated.append(groups)
-	options_matches[replacements] = updated
+DEBUG = 0
 
 
 
-for option, count in options.items():
-	print option, count
-	pat = pattern
[…]

### 2012-10-18 doc/thesis/latex/proposed_solutions.tex
slight modifications to report

@@
 \subsection{Supporting complex queries}
-As most entities in the mediated schema (27 of them) are already mapped to multiple services (see the mapping here: \url{https://cmsweb.cern.ch/das/services}) these taken as a graph could be used to build (some of) the more complex queries.
+As most entities in the mediated schema (27 of them) are already mapped to multiple services (see the mapping here: \url{https://cmsweb.cern.ch/das/services}) in theory these could be taken as a graph could to build (some of) the more complex queries.
+{\color{red}At the moment this still needs some further consideration/planning, as the source descriptions we have are not so detailed as for example \textit{Information Manifold's}.}
 

### 2012-10-18 doc/thesis/latex/report_sept.tex
slight modifications to report

@@
 A possible use-case could be the \textit{Prep} database, but this still has to be discussed.
+%
+Depending on the needs it could be either of exploratory or API/Query Forms based approach, where fake APIs are defined in terms of SQL queries (or even better in a simplified form that would generate SQL by taking schema into account)}
+{
+There was a project\footnote{\url{https://github.com/vkuznet/PyQueryBuilder}}   at CERN which tried to generalize DBS-Query Language for any database.}
 
-Depending on needs it could be either of exploratory or API/Query Forms based approach, where fake APIs are defined in terms of SQL queries (or even better in a simplified form that would generate SQL by taking schema into account)}
-+{\color{green} Indeed, an dthat was the topic of research for another DAS student, Liang Dong.
+% \color{green} Indeed, an dthat was the topic of research for another DAS student, Liang Dong.
 The idea was to generalize DBS-QL to any DB back-end and provide common package which will do a job.
-The work is done over here https://github.com/vkuznet/PyQueryBuilder.
+The work is done over here .
 Feel free to check it out and we can discuss it more.}
-+
 

### 2012-10-18 doc/thesis/latex/proposed_solutions.tex
slight modifications to report

@@
 	\item the algorithm will generate mappings between each keyword to either an entity name, selection attribute name or its value (resembling 'Query forms' and 'Keyword cleaning' approaches)
-	\item ranking could be based on: how closely keywords are matching (some or all) required API's input parameters, popularity of certain queries and users feedback	
+	\item ranking could be based on: how closely keywords are matching (some or all) required API's input parameters, popularity of certain queries, users' feedback	 and user's personal profile (e.g.
+search history, department, etc)
 	\item build an inverted index of attribute values with some good full-text search toolkit%

### 2012-10-18 doc/thesis/latex/report_sept.tex
slight modifications to report

@@
 \usepackage{appendix}
+\usepackage[usenames,dvipsnames]{xcolor}
 
@@
 
+{\color{gray}
 \subsection{Generic connector accessing relational databases}
-{\color{red}Generic connector able to access relational databases with minimal manual integration could be useful for integrating proprietary systems that were not yet integrated and there's no resources available to build APIs.
+Generic connector able to access relational databases with minimal manual integration could be useful for integrating proprietary systems that were not yet integrated and there's no resources available to build APIs.
 A possible use-case could be the \textit{Prep} database, but this still has to be discussed.
 %
-Depending on the needs it could be either of exploratory or API/Query Forms based approach, where fake APIs are defined in terms of SQL queries (or even better in a simplified form that would generate SQL by taking schema into account)}
-{
-There was a project\footnote{\url{https://github.com/vkuznet/PyQueryBuilder}}   at CERN which tried to generalize DBS-Query Language for any database.}
+Depending on the needs it could be either of exploratory or API/Query Forms based approach, where fake APIs are defined in terms of SQL queries (or even better in a simplified form that would generate SQL by taking schema into account).
+There was a project\footnote{\url{https://github.com/vkuznet/PyQueryBuilder}}   at CERN which tried to generalize DBS-Query Language for any database that could be used for this.}
 

### 2012-10-19 doc/thesis/latex/literature_review.tex
some cleanup of literature review and complex queries

@@
 The problem of keyword search over structured sources received significant attention within the last decade.
-Keyword search over relational, XML and other databases was explored form a number of perspectives: returning top-k ranked result tuples vs suggesting structured queries as SQL, performance optimization, user feedback mechanisms, as well as keyword searching over distributed sources.
-It is worth noting that when there is no need for 100\% result exactness keyword search combined with probabilistic schema matching allows lightweight (pay-as-you-go) data integration with minimal human effort upfront\cite[ch.16]{principles_data_integration} but could improve its quality with time by using users' feedback.
+Keyword search over relational, XML and other databases was explored from a number of perspectives: returning top-k ranked result tuples vs suggesting structured queries as SQL, performance optimization, user feedback mechanisms, as well as keyword searching over distributed sources.
+If there is no need for 100\% result exactness keyword search combined with probabilistic schema matching provides lightweight (pay-as-you-go) exploratory data integration with almost no human effort upfront still improving its results' quality with time adjusting to users' feedback\cite[ch.16]{principles_data_integration}.
+On the other hand, SODA has shown that if enough meta-data is in place, fairly complex queries over large and complex warehouses could be answered.
 
 In the setting of Boolean Keyword-based Retrieval over heterogeneous sources a keyword query could be translated into a ranked list of structured queries (similarly as SODA system that propose SQL, Query Forms that propose SQL templates or Keymantic trying to achieve this without accessing the data).
-In the extreme case of having no control over a web-service that do not publish its contents, techniques like Google's Deep-web crawl could index a subset of its contents enabling keyword search to some extent.
+In the extreme case of having no control over a web-service that do not publish its contents, techniques like Google's Deep-web surfacing could index a subset of its contents enabling keyword search to some extent.
 
-\subsection{Composing multiple data sources efficiently: Query translation using Logical Views}
+\subsection{Composing multiple data sources efficiently: Query translation using Logical Views\label{IM_query_translation}}
 The \textit{Information Manifold}\cite{Levy96} is virtual integration (EII) system that uses CARIN dialect of \textit{description logics} (could be though as simplified datalog) for queries and for source descriptions.
@@
 \subsection{Keyword search over a Relational Database}
-The problem of Keyword search over Relational Databases (or also semi-structured sources like XML) has received a significant attention by the research community over the last decade.
+%The problem of Keyword search over Relational Databases (or also semi-structured sources like XML) has received a significant attention by the research community over the last decade.
 
@@
 % 
-A number of problem variations exist:  returning only the ranked Top-k results vs.
+%A number of problem variations exist:  returning only the ranked Top-k results vs.
 returning ranked list of possible queries, while some systems would even allow generating more complex queries including aggregations, etc (SQAK, SODA\cite{ethz2012}).
@@
 
+{\color{gray}
 \subsubsection*{Deep web search at Google}
@@
 there is NO single mediated schema over which users pose queries: queries are routed to to the relevant sources with help statistical methods that are used to model uncertainty at all levels: queries, mappings and underlying data.
-
-%{\color{red} TODO: Describe Keymantic\cite{Keymantic10}.
-This is a very suboptimal solution as indexing terms improves results (mention evaluation from SODA paper).
-It works only then keywords map entity names.
-some hybrid approaches:
-%index if exists, regexp, some string similarity measure based on historical data (e.g. even edit-distance would work for many items, like site)}
-
-%\subsubsection*{Automatic mapping between distributed services}
-%{\color{red} TODO: an interesting approach but not for CERN...}
-
-%\subsubsection*{\color{red}Question Answering}
+}
 
 % See Appendix #1, for evaluation of they Demo system.
-
-% {\color{red}There doesn't seem to be much of recent papers going towards this direction (except from specific search systems like flight fare comparisons), however these are based on structured arguments.}
-
-% Some other approaches could include: - creating virtual documents offline by joining tables, for instantaneous search results by employing IR techniques
-% (Indexing Relational Database Content Offline for Efficient Keyword-Based Search, 2003)
-
-% TODO: (Efficient Keyword Search Across Heterogeneous Relational Databases, 2007) : combines schema matching and structure discovery techniques to ﬁnd approximate foreign-key joins across heterogeneous databases

### 2012-10-19 doc/thesis/latex/problem_statement.tex
some cleanup of literature review and complex queries

--- 2012-10-18 doc/thesis/latex/report_sept.tex
@@
-\documentclass[a4paper,11pt,draft]{article}
-\special{papersize=210mm,297mm}
-
-% twocolumn
-%\usepackage{fullpage} 
-\usepackage[cm]{fullpage}
-
-\addtolength{\topmargin}{-5mm}
-\addtolength{\topskip}{-5mm}
-
-%\usepackage[margin=0.5in]{geometry}
-%\usepackage{sans}
-\usepackage{paralist}
-\usepackage{color}
-
-% enable Hyperlinks
-\usepackage{hyperref}
-% fancy verbatim
-\usepackage{fancyvrb}
-\usepackage{appendix}
-\usepackage[usenames,dvipsnames]{xcolor}
-
-%\usepackage{alltt}
-
-%\usepackage[small,compact]{titlesec}
-
-\bibliographystyle{alpha}
-
-% usually first paragraph is NOT indented, so this is commented out: 
-% \usepackage{indentfirst}
-
-\begin{document}
-
-% Article top matter
-\title{Master thesis problem statement: Keyword search over heterogeneous sources} %\LaTeX is a macro for printing the Latex logo
-\author{Vidmantas Zemleris}  %\texttt formats the text to a typewriter style font
-\date{\today}  %\today is replaced with the current date
-
-%\maketitle
-
-%
- \centerline{\Large \bf Searching heterogeneous data-sources: Master thesis problem statement} %% Paper title
- \medskip
-
- \centerline{Vidmantas Zemleris, \today}  %% Author name(s)
- \medskip
-
-\section{Introduction}
-At large scientific collaborations like the CMS Experiment at CERN's LHC that includes more than 3000 collaborators data usually resides on a fair number of autonomous and heterogeneous proprietary systems each serving it's own purpose
-\footnote{For instance, at CERN,  due to many reasons (e.g.
-research and need of freedom, politics of institutes involved) software projects usually evolve in independent fashion resulting in fair number of proprietary systems\cite{Koch00CERN}.
-Further high turnover makes it harder extending these systems}.
-As data stored on one system may be related to data residing on other systems%
-	\footnote{For example, datasets containing physics events are registered at DBS, while the physical location of files is tracked by Phedex which also takes care of their transfers within the worldwide grid storage}%
-, users are in need of a centralized and easy-to-use solution for locating and combining data from all these multiple services.
-
-Using a highly structured language like SQL is problematic because users need to know not only the language but also where to find the information and also lots of technical details like schema.
-A data integration system based on simple structured queries is already in place.
-Various improvements including support for less restricted keyword queries, improvements to system's usability and performance still have to be researched.
-
-% {\color{red}A remarkable opportunity we have is possibility to test ideas on a sample of ~3000 people working at CMS.}
-
-\section{Case study: the CMS Experiment at CERN}
-Users' information need may vary greatly depending on their role, however most of the time they are interested in locating a \textit{full set} of entities matching some selection criteria, e.g.:
-%\footnote{in contrary to exploratory approach where  a subset of best matches is enough}
-   \begin{itemize}
-  		\item find all \textit{files} from \textit{dataset(s)} matching wild-card query each containing some of the 'interesting' \textit{runs} from a list provided (Release validation teams)
-         \item find (all) \textit{datasets} related some specific physics phenomena\footnote{In case of dataset this data is present in filename or run} together with conditions describing how this data was recorded by detector or simulated which are present in separate autonomous system than the datasets (Physicists)
-         % TODO:is this a good example? conditions is a separate query now...
-         \item find all \textit{datasets} matching some pattern stored at a given \textit{site} 
-         	  \begin{small}(filtering attributes from separate services)\end{small}
-   \end{itemize}                		
-% In such dynamic environment it is preferred to choose local-as-view approach for information integration instead of creating a global schemata\cite{Koch00CERN}.
-
-For more use-cases of data retrieval at CMS Experiment see \cite{CMS_data08}.
-
-%Following a similar approach a 
-\subsection*{The Data Aggregation System}
-
-The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it determines  data-sources are able to answer\footnote{This is done by a mapping between flat mediated schema ('DAS keys') into web-service methods and their arguments.
-Then system queries all services that could provide a result of expected type with given parameters}, queries them, merges the results and caches them for subsequent use.
-DAS uses \textit{Boolean retrieval model} as users are often interested in retrieving ALL the items matching their query.
-
-Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]) operators.
-The combined query results could be later 'piped' for further filtering, sorting or aggregation (min, max, avg, sum, count, median), e.g.:
-
-{\small 
-\begin{verbatim}
-dataset=*RelVal* | grep dataset.nevents >1000 | avg(dataset.size), median(dataset.size)
-\end{verbatim}
-}
-
-The query above would return average and median datasets sizes  of ones containing  \textit{RelVal} in their name having more than 1000 events.
-
-Queries could be run either from web browser or through  command line interface where the results could  be fed into another application (e.g.
-program doing physics analysis or automatic software release validation).
-
-% \textbf{\color{red}Examples of more general uses: Digital libraries, Stock markets, [ Weather, Flight costs]}
-
 \section{Problem statement}
@@
 	Even a simple structured query language and entity names over the mediated schema  at first may seem hard to learn%
-%	\footnote{On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources).}%
 	\footnote{on the other hand, at CERN, the names in the mediated schema are refer to entities that in the real-world are fairly consistently named (though there may exist slight differences in their naming, whereas DAS is to provide the 'golden' naming convention).}.
@@
 %
-Some of the questions to be answered:
+To be answered:
 
 	\begin{compactitem}
-		\item could we select  meaningful and sufficiently efficient composition of services even using less detailed source descriptions than those used in Information Manifold (see Literature Review)?
-		\item could we use the current descriptions that only define entity the result and API parameters?
-%		\item how to present the results (as a sequence of DAS queries or implemented within DAS)
+		\item how should these queries be entered by users?
+
+		\item could we use the current source descriptions that only define API result type and API parameters as entities of mediated schema to select meaningful and sufficiently efficient composition of services?
 	\end{compactitem}
@@
 Find datasets at site X; find files for given dataset list; find runs for found files; filter runs with magnetic field 4T}
-
-% It would be nice to {\color{green} Build series of queries which will answer the given user question, e.g.
-user asks {\it I want to find files which contains runs taken with magnetic field 4T for given dataset at site X}.
-This should be decomposed somehow (subject to research) into series of quries, e.g.
-Find datasets at site X; find files for given dataset list; find runs for found files; filter runs with magnetic field 4T}
-
-% TODO:	Further a minimum set of search predicates is imposed by APIs (mainly because of performance reasons) and user has to be informed what extra is he expected to provide.
-
-% TODO:	Even in structured query mode users are prone to typos, use either autocompletion where possible or suggest alternative queries.
 
@@
 Feel free to check it out and we can discuss it more.}
-
-\input{proposed_solutions}
-
-\include{literature_review}
-
-\section{Work status}
-
-\begin{verbatim}
-- obtained DB copy of biggest data provider DBS (currently 80 GB + 200 GB indexes)
-- preliminary literature review (to be continued deeper)
-- initial analysis of DAS logs
-
-Upcomming Work items:
-- couple of fairly simple prototypes of UI/access patters for simpler DAS querying
-- check performance improvements after creation of materialized view(s)
-- interview more DAS users
-\end{verbatim}
-
-% ---  Bibliography
-\bibliographystyle{unsrt}
-\addcontentsline{toc}{chapter}{Bibliography}
-
-%\nocite{*}
-\thispagestyle{empty}
-\begin{small}
-\bibliography{refs}
-\end{small}
-
-\pagebreak
-\include{das_logs}
-
-\end{document}

### 2012-10-19 doc/thesis/latex/proposed_solutions.tex
some cleanup of literature review and complex queries

@@
 \subsection{Supporting complex queries}
-As most entities in the mediated schema (27 of them) are already mapped to multiple services (see the mapping here: \url{https://cmsweb.cern.ch/das/services}) in theory these could be taken as a graph could to build (some of) the more complex queries.
-{\color{red}At the moment this still needs some further consideration/planning, as the source descriptions we have are not so detailed as for example \textit{Information Manifold's}.}
+{\color{red}At the moment this still needs some further analysis/consideration, however it seems even with current data-source definitions could be used for building more complex queries.}
+
+DAS Service definitions contain for each data service (API): mapping of its parameters and result type into entities of mediated schema (but NOT ALL their fields are defined a priori). ]
+From theoretical (correctness) stand point, this seems mappable  into \textit{Information Manifold's} data source definitions described on Ch.\ref{IM_query_translation} of Literature review.
+
+Other issues to think about contain:
+\begin{itemize}
+\item how these queries shall be entered by the user and how ambiguity shall be handled especially if queries entered in some simplified (semi-structured) language like DAS Query Language currently is
+\item any optimizations/heuristics if multiple possible solutions exist
+\item handling complex and heavy queries with large number of results, e.g. lazy loading with pagination (at least at the level of service composition as APIs do not yet support pagination/sorting)
+\end{itemize}
 

### 2012-10-19 doc/thesis/latex/report_sept.tex
some cleanup of literature review and complex queries

--- 2012-10-18 doc/thesis/latex/report_sept.tex
@@
 
-\addtolength{\topmargin}{-5mm}
-\addtolength{\topskip}{-5mm}
+\addtolength{\topmargin}{-10mm}
+\addtolength{\topskip}{-10mm}
 
@@
 
-\section{Problem statement}
-\subsection{User Interface and ease of use}
-
-For an IR system with wide variety of users, it important to provide an easy to use interface with fairly flat learning curve, while at the same time not loosing support of fairly complex queries.
-%
-	Even a simple structured query language and entity names over the mediated schema  at first may seem hard to learn%
-%	\footnote{On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources).}%
-	\footnote{on the other hand, at CERN, the names in the mediated schema are refer to entities that in the real-world are fairly consistently named (though there may exist slight differences in their naming, whereas DAS is to provide the 'golden' naming convention).}.
-Therefore, it shall be useful to guide new users interactively through the process of building of their query.
-%   
-   Supporting non-structured keyword queries is also worth investigation as quite many users reported they are missing this Google-like search experience.
-
-\subsection{Supporting complex queries}
-Currently DAS can only process queries that could be directly mapped to data providers' APIs, but not the queries where results of one API need to be fed into another APIs 
-	%\footnote{this is partially because of performance reasons {\color{red}and the fact that DAS is mainly schema-less (the mediated schema is only in terms of entities, their mappings to APIs and API parameters)}}%
-	\footnote{except a couple of manually hard-coded "views" in a virtual \textit{"combined"} data provider}%
-. %In such case, the user  is forced to write some code to submit and process the sub-queries himself. 
-Further, to be flexible DAS only keeps a list of 'mediated entities' and how they could be retrieved, but do not enforce any predefined schema (i.e. entity fields/attributes) nor exposes it\footnote{%
-	Actually the fields of each API  (an entity could be mapped to a number of these) could be inferred 
-	from past queries or by predefined 'bootstrap' queries. Currently the field lists may differ slightly depending on the parameters}
-before querying for some particular entity(-ies). 
-%
-This makes it fairly hard to execute complex queries  without a priori knowledge.
-User has to know what queries and how have to be combined (depends on the mediated schema and what APIs are available), and has either to combine the results by hand or to write a program does that.
-%
-Some of the questions to be answered:
-
-	\begin{compactitem}
-		\item could we select  meaningful and sufficiently efficient composition of services even using less detailed source descriptions than those used in Information Manifold (see Literature Review)?
-		\item could we use the current descriptions that only define entity the result and API parameters?
-%		\item how to present the results (as a sequence of DAS queries or implemented within DAS)
-	\end{compactitem}
-
-%+\item {\color{green} Build series of queries which will answer the given user question, e.g.
-user asks {\it I want to find files which contains runs taken with magnetic field 4T for given dataset at site X}.
-This should be decomposed somehow (subject to research) into series of quries, e.g.
-Find datasets at site X; find files for given dataset list; find runs for found files; filter runs with magnetic field 4T}
-
-% It would be nice to {\color{green} Build series of queries which will answer the given user question, e.g.
-user asks {\it I want to find files which contains runs taken with magnetic field 4T for given dataset at site X}.
-This should be decomposed somehow (subject to research) into series of quries, e.g.
-Find datasets at site X; find files for given dataset list; find runs for found files; filter runs with magnetic field 4T}
-
-% TODO:	Further a minimum set of search predicates is imposed by APIs (mainly because of performance reasons) and user has to be informed what extra is he expected to provide.
-
-% TODO:	Even in structured query mode users are prone to typos, use either autocompletion where possible or suggest alternative queries.
-
-\subsection{Performance}
-
-DAS is based on \textit{Virtual Integration} where data is left at the sources allowing data to be volatile (e.g.
-new records gathered or existing records updated; service descriptions may be changed; new services added), while some subsets of the data could be fairly static (e.g.
-datasets stored at DBS do not change often).
-%
-Further as the total data that is stored at data-providers could be fairly large (order of 1TB per year) and as their main goal is supporting production (data taking from the CMS detector), the providers may impose some constraints (e.g.
-only certain queries allowed).
-%
-Therefore, it is important to balance between returning locally cached query results quickly   and between issuing queries to slow data-services to get fresh results.
-%
-Issues to be addressed includes:
-\begin{compactitem}
-						\item Low data providers' performance (e.g. on DBS some queries requires joining many very large tables)
-	     				\item At the moment all queries are put into one pool and has to wait until some threads are available.
-If a couple of heavy queries were submitted earlier, even light queries would have to wait long.
-Explore more advanced Query prioritization
-	     					(e.g. we could have some query cost model based on history or predefined scores per API)
-
-                		\item Currently result items are cached for a fairly short period of time (5min-1h) and then completely discarded, however many entities are not changing that often - Explore more intelligent caching
-
-	                		% TODO: \item query rewriting then used with grep=smf then the same item is available as selection key. check how many instances.
-
-					\item Since DAS does aggregation across multiple data providers, given their current APIs (with no ordering and limit functionality), DAS has to fetch ALL records matching the query instead of only the first page.
-Although it would be possible to do partial results for data coming from a single data-source (while still loading in background, or the providers (or us) would have to modify their APIs).
-
-					\item Efficient distributed search: as filtering criteria may reside on two or more autonomous sources, given the current APIs much more items than in the final result may need to be fetched from each source.
-Items that do not meet all selection criteria could be filtered out only afterwards in DAS.
-To improve the performance, data provider's APIs for the most used queries may need to be redesigned.
-                        % \item {\color{red}e.g.
-displaying intermediary results of aggregation while still calculating with statistical bounds -- however statistical bounds are much meaningful only if items are well snuffled/randomized -- people could stop the query if they see a non-sense}
-	              %\item {\color{red}scale testing - if we are storing lots of historical info.
-MongoDB is not so performant if DB cant fit in memory.
-One of well known solutions is installing SSD.
-}
-\end{compactitem}
-
-\subsection{Generic connector accessing relational databases}
-{\color{red}Generic connector able to access relational databases with minimal manual integration could be useful for integrating proprietary systems that were not yet integrated and there's no resources available to build APIs.
-A possible use-case could be the \textit{Prep} database, but this still has to be discussed.
-
-Depending on needs it could be either of exploratory or API/Query Forms based approach, where fake APIs are defined in terms of SQL queries (or even better in a simplified form that would generate SQL by taking schema into account)}
-+{\color{green} Indeed, an dthat was the topic of research for another DAS student, Liang Dong.
-The idea was to generalize DBS-QL to any DB back-end and provide common package which will do a job.
-The work is done over here https://github.com/vkuznet/PyQueryBuilder.
-Feel free to check it out and we can discuss it more.}
-+
+\input{problem_statement}
 
@@
 \section{Work status}
-
+{\color{red}
 \begin{verbatim}
@@
 \end{verbatim}
-
+}
 % ---  Bibliography

### 2012-10-19 doc/thesis/latex/problem_statement.tex
improvements to the text

@@
 
-For an IR system with wide variety of users, it important to provide an easy to use interface with fairly flat learning curve, while at the same time not loosing support of fairly complex queries.
+For an IR system with a wide variety of users, it important to provide an easy to use interface with fairly flat learning curve, while not loosing support of fairly complex queries.
 %
-	Even a simple structured query language and entity names over the mediated schema  at first may seem hard to learn%
-	\footnote{on the other hand, at CERN, the names in the mediated schema are refer to entities that in the real-world are fairly consistently named (though there may exist slight differences in their naming, whereas DAS is to provide the 'golden' naming convention).}.
-Therefore, it shall be useful to guide new users interactively through the process of building of their query.
+	Even a simple structured query language plus entity names over the mediated schema  at first may seem hard to learn%
+	\footnote{Actually, the names in DAS mediated schema refer to entities fairly consistently named in the real-world.}.
+Therefore, it shall be useful to guide new users interactively through the process of building the query.
 %   
-   Supporting non-structured keyword queries is also worth investigation as quite many users reported they are missing this Google-like search experience.
+   Supporting non-structured keyword queries is also worth investigation as quite many users reported missing this Google-like search experience.
 
@@
 	%\footnote{this is partially because of performance reasons {\color{red}and the fact that DAS is mainly schema-less (the mediated schema is only in terms of entities, their mappings to APIs and API parameters)}}%
-	\footnote{except a couple of manually hard-coded "views" in a virtual \textit{"combined"} data provider}%
+	\footnote{Except a couple of manually hard-coded "views" in a virtual \textit{"combined"} data provider}%
 . %In such case, the user  is forced to write some code to submit and process the sub-queries himself. 
@@
 This makes it fairly hard to execute complex queries  without a priori knowledge.
-User has to know what queries and how have to be combined (depends on the mediated schema and what APIs are available), and has either to combine the results by hand or to write a program does that.
+User has to know what queries and how have to be combined (depends on the mediated schema and what APIs are available), and has either to combine the results by hand or to write a program doing that.
 %
-To be answered:
-
+To be researched:
 	\begin{compactitem}
@@
 
-		\item could we use the current source descriptions that only define API result type and API parameters as entities of mediated schema to select meaningful and sufficiently efficient composition of services?
+		\item could we use current source descriptions that only define API result type and API parameters as entities of mediated schema to select meaningful and sufficiently efficient composition of services?
 	\end{compactitem}
@@
 
-DAS is based on \textit{Virtual Integration} where data is left at the sources allowing data to be volatile (e.g.
-new records gathered or existing records updated; service descriptions may be changed; new services added), while some subsets of the data could be fairly static (e.g.
-datasets stored at DBS do not change often).
-%
-Further as the total data that is stored at data-providers could be fairly large (order of 1TB per year) and as their main goal is supporting production (data taking from the CMS detector), the providers may impose some constraints (e.g.
-only certain queries allowed).
-%
-Therefore, it is important to balance between returning locally cached query results quickly   and between issuing queries to slow data-services to get fresh results.
+DAS is based on \textit{Virtual Integration} where data is left at the sources allowing data to be volatile, e.g.
+new records gathered or existing ones updated, service's schema changed.
+Still large amounts of the data are fairly static.
+% %the providers may impose some constraints (e.g. only certain queries allowed).}
+As APIs of some data providers' are comparatively slow\footnote{This is partially because of large amounts of data stored at the providers (order of ~1TB per year in total) and as their main goal is supporting data taking from the CMS detector their systems were not optimzed for exploratory queries} it is important to balance between quickly returning items from local cache and getting fresh results.
+Another approach we are taking is spotting the performance issues at the services (e.g.
+at DBS some queries requires joining many very large tables, while most of their existing data stays static).
 %
@@
 \begin{compactitem}
-						\item Low data providers' performance (e.g. on DBS some queries requires joining many very large tables)
 	     				\item At the moment all queries are put into one pool and has to wait until some threads are available.
@@
 Explore more advanced Query prioritization
-	     					(e.g. we could have some query cost model based on history or predefined scores per API)
+	     					(e.g. we could have a query cost model)
+						%  based on history or predefined scores per API
 
@@
 
-					\item Since DAS does aggregation across multiple data providers, given their current APIs (with no ordering and limit functionality), DAS has to fetch ALL records matching the query instead of only the first page.
-Although it would be possible to do partial results for data coming from a single data-source (while still loading in background, or the providers (or us) would have to modify their APIs).
-
+					\item Since DAS does aggregation across multiple data providers, given their current APIs (with no \textit{ordering} and \textit{paging} of results), DAS has to fetch ALL records matching the query instead of only the first page\footnote{Although it would be possible to show partial results for data coming from a single data-source (still loading in background, or the providers' APIs would have to be modified)}.
 					\item Efficient distributed search: as filtering criteria may reside on two or more autonomous sources, given the current APIs much more items than in the final result may need to be fetched from each source.
@@
 
-{\color{gray}
-\subsection{Generic connector accessing relational databases}
-Generic connector able to access relational databases with minimal manual integration could be useful for integrating proprietary systems that were not yet integrated and there's no resources available to build APIs.
+\subsection{\color{gray}Generic connector accessing relational databases}
+
+{\color{gray}A generic connector accessing relational databases with minimal human effort could be useful for integrating proprietary systems that were not yet integrated when there's no resources to build the APIs.
 A possible use-case could be the \textit{Prep} database, but this still has to be discussed.
 %
-Depending on the needs it could be either of exploratory or API/Query Forms based approach, where fake APIs are defined in terms of SQL queries (or even better in a simplified form that would generate SQL by taking schema into account).
-There was a project\footnote{\url{https://github.com/vkuznet/PyQueryBuilder}}   at CERN which tried to generalize DBS-Query Language for any database that could be used for this.}
-
+%Depending on the needs it could be either of exploratory or API/Query Forms based approach, where fake APIs are defined in terms of SQL queries (or even better in a simplified form that would generate SQL by taking schema into account).
+There was a project\footnote{\url{https://github.com/vkuznet/PyQueryBuilder}}   at CERN which tried to generalize DBS-Query Language for any database that could be used as DB side mediator.
+}
 % \color{green} Indeed, an dthat was the topic of research for another DAS student, Liang Dong.

### 2012-10-19 doc/thesis/latex/proposed_solutions.tex
improvements to the text

@@
 
+\item JavaScript-based query interpreter with auto completion to ease query writing: could suggest searchable entities and maybe even entity attribute names.
+
 \item Given a keyword query, suggest the best matching structured queries (as a ranked list)%
@@
 	\item build an inverted index of attribute values with some good full-text search toolkit%
-				\footnote{e.g. Xapian (http://xapian.org/) implemented in C++ which that is known for its flexibility, or Sphinx (http://sphinxsearch.com/)}. 
+				\footnote{e.g. Xapian (http://xapian.org/) that is known for its flexibility, or Sphinx (http://sphinxsearch.com/)}. 
 	If no match found (e.g. new entry not yet in our cache) pattern matching could be used to make a best guess
-	\item the system could improve its results from users' feedback (e.g. by HMM, will need to decide)
-	\item multi-wildcard queries (e.g.
-dataset=*Zmm*special*RECO*) are common at CMS that will need special attention\footnote{Matching to structured query will be harder and guessing even worse.
-Also for performance reasons some web-services will introduce limitations on the format of wilcard queries, while DAS wants still to be flexible at least for datasets}.
+	\item the system could improve its results from users' feedback through machine learning (e.g. HMM)
+	\item multi-wildcard queries (e.g. dataset=*Zmm*special*RECO*) are common at CMS that will need special attention\footnote{%
+		Matching a wildcard query into a schema field will be harder.
+Also for performance reasons some web-services will introduce limitations on the format of wilcard queries, while DAS wants to allow more flexible queries, at least for \textit{datasets}}.
 
-\item knowing all possible APIs parameter values would be beneficial ({\color{red}but not yet sure if available}%
+\item knowing all possible APIs' parameter values, in addition to historical values seen so far, would improve quality of the keyword search (better mapping from keywords to these values) or even the structured search e.g.
+through query cleaning \& auto-completion.
+{\color{red}Not yet known if available}%
 		\footnote{problem: the owners of services may not want to provide direct access to DB because of no trust, security or performance issues.
-Otherwise, specific APIs could be developed to return possible parameter values but this is quite an overhead}) and could be used to improve the keyword search (better mapping from keywords to API inputs) or even the structured search (e.g.
-query cleaning/term auto-completion)
+Otherwise, specific APIs could be developed to return possible parameter values but this is quite an overhead}.
 \begin{itemize}
@@
 an inverted index of values in Oracle/MySQL DB tables can be built to keep the list of possible terms.
-For instance, Sphinx full text search engine can access DBs directly and also supports incremental indexing but it needs a bit of manual configuration {http://sphinxsearch.com/docs/current.html\#delta-updates} }%
-	tables mapping it's columns into API parameters (manually or automatically).
+For instance, Sphinx full text search engine can access DBs directly and also supports incremental indexing but it needs a bit of manual configuration {http://sphinxsearch.com/docs/current.html\#delta-updates} }
+	tables semi-automatically mapping it's columns into API parameters.
 
@@
 	\end{itemize}
-\item (?) JavaScript-based query interpreter to ease query writing: could suggest search attribute and entity attribute names.
-also taking into account some possible ambiguous namings (auto completion)
 \end{itemize}
@@
 \subsection{Supporting complex queries}
-{\color{red}At the moment this still needs some further analysis/consideration, however it seems even with current data-source definitions could be used for building more complex queries.}
+{\color{red}At the moment this still needs some further analysis/consideration, however it seems even the existing data-source definitions could be used for building more complex queries.}
 
-DAS Service definitions contain for each data service (API): mapping of its parameters and result type into entities of mediated schema (but NOT ALL their fields are defined a priori). ]
-From theoretical (correctness) stand point, this seems mappable  into \textit{Information Manifold's} data source definitions described on Ch.\ref{IM_query_translation} of Literature review.
+DAS Service definitions contain: for each data service (API), mapping of its parameters and result type into entities of mediated schema (but NOT ALL their fields are defined a priori).
+Even if DAS definitions do not describe every attribute of API result, as all input parameters are attributed to a mediated schema entity, these definitions seem mappable  into \textit{Information Manifold's} data source definitions described on Sec.\ref{IM_query_translation} of Literature review.
 
-Other issues to think about contain:
+Other issues to consider:
 \begin{itemize}
@@
 \subsubsection*{More intelligent caching}
-Some of the data entities instances changes very rarely.
-For example in DBS system old datasets would never change, while new entities are constantly added (still some of their attributes may change, e.g.
+Some of the data instances changes very rarely.
+For example in DBS system old datasets would never change, while new entities are constantly added (some of their attributes may change, but only rarely, e.g.
 validity of a dataset).
@@
 
-Also consider that:
-                			\begin{itemize}
-                			%\item given an entity received from provider, determining if it is useful to cache for long-term
-%	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
-                			%\item could cache even \textit{expired} data, but warning user
-                			\item we could also have different validity dates for certain fields. if no volatile fields are not explicitly requested, even a very old cache could be used. 
-                			% \item Can we automatically figure which fields are static and which are changing?
-			 {\color{green} Yes, if we enfore APIs to record modification timestamp} {\color{red} I also meant some automatic algorithm, however that would be more complex and less reliable}
-                			\item pre-fetching common queries: determining manually and/or automatically
-                			\end{itemize}
+Other way to boost performance is pre-fetching most common queries from time to time (determined manually and/or automatically).
+Also it's possible have different validity dates for certain fields.
+if no volatile fields are not explicitly requested, even a very old cache could be used
 
 %%%%%  -------------------------------------------
-\subsubsection*{Data providers performance issues}
-In the cases when new records are coming but the existing ones are not changing much, continuous view maintenance that computes only differences from earlier results could be a good solution to improve the performance especially for often used queries.
-
-Use either \textit{materialized refresh fast views with query rewriting} (Oracle; completely transparent for proprietary apps)\cite{Oracle11}
- or some other continuous view maintenance tool (e.g.
-DBToaster\footnote{http://www.dbtoaster.org that is being developed at EPFL}) to improve performance of heavy queries containing joins and/or aggregations.
+\subsubsection*{Data providers' performance issues}
+In the cases when new records are coming but the existing ones are not changing much, continuous view maintenance that computes only differences from earlier results could be a good solution to improve the performance especially for popular and heavy queries containing joins and/or aggregations.
 
@@
 Having a materialized view with all these tables joined together would allow answering such queries much quicker.
-Given low change rates (in comparison to data already present), maintaining the view should be also comparatively cheap with only expense of just couple of times of space.
+Given low change rates (in comparison to data already present), maintaining the view should be also comparatively cheap with the only expense of just couple of times of storage space (storage is bound by the size of the largest table anyway).
+
+For Oracle \textit{materialized refresh fast views with query rewriting} would be completely transparent and would not need any changes to proprietary systems, but it has a couple of limitations on the queries\cite{Oracle11}.
+Otherwise some other continuous view maintenance tool (e.g.
+DBToaster\footnote{http://www.dbtoaster.org that is being developed at EPFL}) could be used, however this wouldn't be as transparent.
 
@@
 \subsubsection*{Integrating distributed information efficiently}
-Bloom-join (which could be quite transparent and implemented even on DB side [pure SQL is possible for MySQL, to check for Oracle]- take a query and bit-vector as parameter) , lazy pagination (and order required for aggregation) - this is not yet supported by any of the data-service APIs
-
-%\newline
-% integration at DAS level. (at source DB could be more performant a little)
+Bloom-join (which could be quite transparent and implemented even on DB side (efficient pure SQL is doable for MySQL, and probably for Oracle too) quite almost transparently (a helper function could take the initial query and bloom-filter's bit-vector as parameters).

### 2012-10-19 doc/thesis/latex/report_sept.tex
improvements to the text

@@
 
-\addtolength{\topmargin}{-10mm}
-\addtolength{\topskip}{-10mm}
+%\addtolength{\topmargin}{-10mm}
+%\addtolength{\topskip}{-10mm}
 
@@
 \begin{document}
+%\addtolength{\bottommargin}{-10mm}
+\addtolength{\textheight}{10mm}
 
@@
 \footnote{For instance, at CERN,  due to many reasons (e.g.
-research and need of freedom, politics of institutes involved) software projects usually evolve in independent fashion resulting in fair number of proprietary systems\cite{Koch00CERN}.
-Further high turnover makes it harder extending these systems}.
-As data stored on one system may be related to data residing on other systems%
-	\footnote{For example, datasets containing physics events are registered at DBS, while the physical location of files is tracked by Phedex which also takes care of their transfers within the worldwide grid storage}%
+research-orientedness and need of freedom, politics of institutes involved) software projects usually evolve in independent fashion resulting in fair number of proprietary systems\cite{Koch00CERN}.
+Further high turnover makes it harder to extend these systems}.
+As data stored on one system may be related to data residing on others%
+	\footnote{For example, datasets containing physics events are registered at DBS, while the physical location of files is tracked by Phedex which also takes care of their transfers within the grid storage worldwide}%
 , users are in need of a centralized and easy-to-use solution for locating and combining data from all these multiple services.
@@
 Using a highly structured language like SQL is problematic because users need to know not only the language but also where to find the information and also lots of technical details like schema.
-A data integration system based on simple structured queries is already in place.
-Various improvements including support for less restricted keyword queries, improvements to system's usability and performance still have to be researched.
+An enterprise information integration system based on simple structured queries is already in place.
+Various improvements including: adding support for more complex queries and for less restricted keyword search, improvements to system's usability and performance still have to be researched.
 
@@
   		\item find all \textit{files} from \textit{dataset(s)} matching wild-card query each containing some of the 'interesting' \textit{runs} from a list provided (Release validation teams)
-         \item find (all) \textit{datasets} related some specific physics phenomena\footnote{In case of dataset this data is present in filename or run} together with conditions describing how this data was recorded by detector or simulated which are present in separate autonomous system than the datasets (Physicists)
-         % TODO:is this a good example? conditions is a separate query now...
          \item find all \textit{datasets} matching some pattern stored at a given \textit{site} 
          	  \begin{small}(filtering attributes from separate services)\end{small}
+         \item find (all) \textit{datasets} related to some specific physics phenomena\footnote{In case of dataset, this data is present in the \textit{filename} or in the \textit{run} entity} together with conditions describing how this data was recorded by detector or simulated which are present in another autonomous system than the datasets (Physicists)
+         % TODO:is this a good example? conditions is a separate query now...
    \end{itemize}                		
@@
 
-The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it determines  data-sources are able to answer\footnote{This is done by a mapping between flat mediated schema ('DAS keys') into web-service methods and their arguments.
+The \textit{CMS Data Aggregation System (DAS)}\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it determines  data-sources are able to answer\footnote{This is done by a mapping between flat mediated schema entities ('DAS keys') into web-service methods and their arguments.
 Then system queries all services that could provide a result of expected type with given parameters}, queries them, merges the results and caches them for subsequent use.
@@
 
-Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]) operators.
+Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]).
 The combined query results could be later 'piped' for further filtering, sorting or aggregation (min, max, avg, sum, count, median), e.g.:
@@
 {\small 
-\begin{verbatim}
+\begin{Verbatim}[commandchars=\\\{\}]
+\footnotesize# returns average and median dataset sizes of ones containing \textit{RelVal} in their name having more than 1000 events.
 dataset=*RelVal* | grep dataset.nevents >1000 | avg(dataset.size), median(dataset.size)
-\end{verbatim}
+\end{Verbatim}
 }
 
-The query above would return average and median datasets sizes  of ones containing  \textit{RelVal} in their name having more than 1000 events.
-
-Queries could be run either from web browser or through  command line interface where the results could  be fed into another application (e.g.
+Queries are executed either from web browser or through a command line interface where the results could be fed into another application (e.g.
 program doing physics analysis or automatic software release validation).
-
 % \textbf{\color{red}Examples of more general uses: Digital libraries, Stock markets, [ Weather, Flight costs]}
-
 \input{problem_statement}

### 2012-10-19 doc/thesis/latex/compile.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+
+# http://amath.colorado.edu/documentation/LaTeX/reference/faq/bibstyles.html
+rm ./report_sept.pdf
+pdflatex report_sept
+biblex report_sept
+pdflatex report_sept
+pdflatex report_sept
+
+killall evince
+evince ./report_sept.pdf &

### 2012-10-19 doc/thesis/latex/literature_review.tex
probably final version of sept report

@@
 
-Since the late 1990's, several Enterprise Information Integration\footnote{%
+Since the late 1990's, several \textit{Enterprise Information Integration}\footnote{%
 	Enterprise Information Integration (EII) is about 'integrating data from 
@@
  (EII) products have appeared in the market (e.g.
-\textit{Information Manifold} by AT\&T Lab) and an significant experience has been accumulated on: data integration formalisms, ways of describing of heterogeneous data sources and their abilities (e.g.
+\textit{Information Manifold} by AT\&T Lab) and an significant experience has been accumulated on data integration formalisms, ways of describing heterogeneous data sources and their abilities (e.g.
 RDBMS vs web form), query optimization (combining sources efficiently, source overlap, data quality, etc)\cite{eii_2005}.
 %
-Recent research in data integration mostly focused on methods of automatic, uncertainty and feedback  driven (e.g.
-though statistical mediator) semi-automatic pay-as-you-go approaches\cite[ch.19]{principles_data_integration}).
+Recent research in  Enterprise Information Integration mostly focused on approaches minimizing  human efforts on source integration, e.g.
+on uncertainty based self-improving systems\cite[ch.19]{principles_data_integration}.
 
@@
 Keyword search over relational, XML and other databases was explored from a number of perspectives: returning top-k ranked result tuples vs suggesting structured queries as SQL, performance optimization, user feedback mechanisms, as well as keyword searching over distributed sources.
-If there is no need for 100\% result exactness keyword search combined with probabilistic schema matching provides lightweight (pay-as-you-go) exploratory data integration with almost no human effort upfront still improving its results' quality with time adjusting to users' feedback\cite[ch.16]{principles_data_integration}.
-On the other hand, SODA has shown that if enough meta-data is in place, fairly complex queries over large and complex warehouses could be answered.
+%
+If there is no need for 100\% result exactness, keyword search combined with probabilistic schema matching provides lightweight exploratory data integration with almost no human effort upfront with ability to improve the results by adjusting to users' feedback\cite[ch.16]{principles_data_integration}.
+On the other hand, the \textit{SODA} system has shown that if enough meta-data is in place, even quite complex queries given in bussiness terms could be answered over a large and complex warehouse.
 
-In the setting of Boolean Keyword-based Retrieval over heterogeneous sources a keyword query could be translated into a ranked list of structured queries (similarly as SODA system that propose SQL, Query Forms that propose SQL templates or Keymantic trying to achieve this without accessing the data).
+In the setting of Boolean Keyword-based Retrieval over heterogeneous sources a keyword query could be translated into a ranked list of structured queries (similarly as \textit{SODA} system that propose SQL, \textit{Query Forms} approach that propose SQL templates or Keymantic trying to achieve this without accessing the data).
 In the extreme case of having no control over a web-service that do not publish its contents, techniques like Google's Deep-web surfacing could index a subset of its contents enabling keyword search to some extent.
@@
 \subsection{Composing multiple data sources efficiently: Query translation using Logical Views\label{IM_query_translation}}
-The \textit{Information Manifold}\cite{Levy96} is virtual integration (EII) system that uses CARIN dialect of \textit{description logics} (could be though as simplified datalog) for queries and for source descriptions.
-It describes each source as logical view over the global (mediated) schema, augmented with source capabilities (e.g.
-what are possible and the required input parameters for a view to do selections\footnote{this makes these logical views quite similar to source APIs used by DAS, with difference that DAS currently only describe the parameters APIs and only partially the results (as an entity type)}).
+The \textit{Information Manifold}\cite{Levy96} is virtual integration (EII) system that represents  both it's \textit{queries} and \textit{source descriptions} through a dialect of \textit{description logics} (could be thought as \textit{datalog}).
+It describes each source as a \textit{logical view} over the global (mediated) schema, augmented with source capabilities (e.g.
+what are possible and the required input parameters for the source to return results\footnote{this makes these logical views quite similar to source APIs used by DAS, with difference that DAS currently only describes the parameters APIs and only partially the results}).
 %
-This allows designing algorithms that could allow efficiently answering complex queries that require composing multiple the data sources by finding maximally contained rewriting of (conjunctive) input query in terms of logical views provided by sources (that is, finding an optimal way to compose the sources).
+This allows designing algorithms that could efficiently answer complex queries that require composing multiple the data sources that is done by finding maximally contained rewriting of the (conjunctive) input query in terms of logical views representing the sources (that is, finding an optimal way to compose the sources).
 
-For example (based on \cite{integr_views2000}), consider such sources as expressed as views (on the left) in terms of global predicates of the mediated schema (on the right) in datalog notation:
+For example, consider such sources expressed as views (on the left) in terms of global predicates of the mediated schema (on the right) in datalog notation (based on \cite{integr_views2000}):
 %
@@
 %
-Suppose we wanted to know Sally's phone and office. We express this by datalog in global predicates:
+Suppose we wanted to know Sally's phone and office. We express this again in datalog over global predicates:
 {\small\begin{verbatim}
@@
 \end{verbatim}}
-There are two minimal solutions (as the sources could be incomplete the full solutionis union of the two):
+There are two minimal solutions (as sources could be incomplete, the full solution is union of the two):
 {\small
@@
 Notice that the expansions of these solutions (e.g.
-answer1\_exp) are not equivalent to $q_1$, but the conjunctive queries that are closest and still contained in $q_1$ (as these are the only usable views at sources):
-{\footnotesize\begin{verbatim}
-answer1_exp(P,O) :- emp(sally) & phone(sally,P) & mgr(sally,M) & emp(sally) & office(sally,O) & dept(sally,D).
-\end{verbatim}
+answer1\_exp) are not equivalent to $q_1$, but only the conjunctive queries that are closest and still contained in $q_1$ (as they are the only usable views provided by the sources):
+{\footnotesize\begin{Verbatim}[commandchars=\\\{\}]
+answer1_exp(P,O) :- emp(sally) & \textbf{\underline{phone}(sally,P)} & mgr(sally,M) & emp(sally) & \textbf{\underline{office}(sally,O)} & dept(sally,D).
+\end{Verbatim}
 }
-After this \textit{Information Manifold} would find an executable order that adheres the capabilities of the sources, by iteratively considering any sources whose input parameters are satisfied. 
+After this the \textit{Information Manifold} would find an executable order that adheres the capabilities of the sources, by iteratively considering any sources whose input parameters are satisfied. 
 % \subsection{Keyword search: integration on demand}
@@
 
-The basic approach would first build an inverted index on database tables (usually only text columns).
-Then after finding all occurrences of the keywords, would try to construct join paths (based on Foreign keys) that would unite tuples containing the keywords.
-
+The basic approach would first build an inverted index on values in database specifying where each comes from (usually only on text columns), as well as schema items (table and column names). 
+%
+Then, given a keyword query, the occurrences of the keywords would serve as entry points for the search, from where it would try to construct join paths (based on Foreign keys) that would unite tuples containing these keywords.
 % In addition to many papers the PhD dissertations \cite{PhD_2011, PhD_2012} describe the approaches in details including performance optimization details (e.g. generating materialized views), etc. 
@@
 \subsubsection*{Ranking Query Templates based on keyword query}
-A simple way to access relational database could  be through a set of predefined named query templates (SQL with selection parameters or operators still to be specified) exposed to a user as a Form that the user has to fill in.
+A simple way to access relational database could  be through a predefined set of named query templates (SQL with selection parameters or operators still to be specified) exposed to a user as a Form that the user has to fill in.
 
-\cite{forms_kws} proposes alternative approach for processing keyword queries over relational databases: given a keyword query, instead of returning database tuples one could rank query forms that best matches the query for user to choose the right one (if they are properly named this is fairly easy).
-The ranking is based on checking matching of keywords to table names in templates and to column values.
+\cite{forms_kws} proposes to use this for answering keyword queries: instead of returning database tuples one could rank query forms and the user could choose the intended one (if they are properly named this is fairly easy).
+The ranking is based on how well the given keywords match table names and column values contained in each template.
 
-An interesting feature of this approach is that a Query Template is functionally similar to any autonomous web service (which given the parameters would in turn execute that query on its database).
-In case of the Data Aggregation System, a user after entering a keyword query could be provided with a ranked list of structured queries (attribute=value) that could be processed given data source constraints (e.g.
+Actually, a \textit{Query Template} is functionally similar to any autonomous web service (which given the parameters would in turn execute that query on its database), with the only difference that access to potentially complex web service is often more expensive than to a database entry.
+In case of DAS, a user after providing a query, could be provided with a ranked list of structured queries (attribute=value) that could be processed given data source constraints (e.g.
 parameters required) and if needed he could refine his search (e.g.
@@
 
-\subsubsection*{What if there is no access to index data terms}
+\subsubsection*{What if there is no access to index data terms?}
 \cite{Keymantic10, semantics_without_access} explores the case then there is no possibility to index the data terms, e.g.
@@
 %
-In Keymantic\cite{Keymantic10} a 'keyword query is processed as follows: First, all keywords that  correspond to metadata items (e.g., field names) are extracted. The remaining
-keywords are considered as possible input fields. Second, the likelihood of a remaining keyword to a metadata item is computed in order to rank different options to execute the keyword query on the
-“Hidden Web” database'\cite[p.942]{ethz2012}.
-
-{\color{gray}
-\subsubsection*{Deep web search at Google}
-%[multiple papers from Google] discusses various ways for implementing data integration in terms of large-scale search engine (Google): Virtual integration vs.
-Surfacing.
-They also present ways for integrating systems without human intervention through use of statistical 'mediator'.
-
-There are two approaches to web scale search over deep-web (here Google mostly cares about web forms): \marginpar{\scriptsize\color{red}TODO: how does google use PayGo?} 
-%\textit{Runtime query reformulation} - 'leaves data at the sources and routes queries to appropriate services'\cite[p. 1]{webscale_paygo}
-
-\textit{Deep-web surfacing} - surface deep-web (e.g.
-web forms) adding their results into the standard search index  easily allowing to using existing IR technology that scales well.
-There exist algorithms which allow to iteratively choose input parameters to the forms to surface a considerable part of the 'hidden' data without large overhead%
-	\footnote{i.e.
-if choosing parameters in not smart way, a web form with just a couple of free inputs (or even dropdowns that's easier case), could yield as many results as a cross product of all input combinations.}.
-
-\textit{Pay-as-you-go approach}\cite{webscale_paygo} - With this approach, 'a system starts with
-very few (or inaccurate) semantic mappings and these mappings are improved over time as deemed necessary';
-% TODO: cite{bootstraping}
-there is NO single mediated schema over which users pose queries: queries are routed to to the relevant sources with help statistical methods that are used to model uncertainty at all levels: queries, mappings and underlying data.
-}
+In Keymantic\cite{Keymantic10} a keyword query is processed as follows: First, all keywords that  correspond to metadata items (e.g., field names) are extracted.
+The remaining keywords are considered as possible parameters to the input fields in the web form.
+Second, the likelihood of a remaining keyword to matching a metadata item is computed in order to rank different options of executing this keyword query on the "Hidden Web"\cite[p.942]{ethz2012}.
 

### 2012-10-19 doc/thesis/latex/report_sept.tex
probably final version of sept report

@@
-\documentclass[a4paper,11pt,draft]{article}
+\documentclass[a4paper,11pt]{article}
+% ,draft
 \special{papersize=210mm,297mm}
@@
 %\usepackage{sans}
+%\usepackage{lmodern}
 \usepackage{paralist}
@@
 
+\include{literature_review}
+
 \input{proposed_solutions}
 
-\include{literature_review}
-
+\newpage
 \section{Work status}
-{\color{red}
+{\begin{footnotesize}
 \begin{verbatim}
-- obtained DB copy of biggest data provider DBS (currently 80 GB + 200 GB indexes)
+- set-up of development environment
 - preliminary literature review (to be continued deeper)
 - initial analysis of DAS logs
+- obtained DB copy of biggest data provider DBS (currently 80 GB + 200 GB indexes)
+- some work done on solving wildcard query restrictions
 
 Upcomming Work items:
+- benchmark data provider's performance (per API based on historical queries)
 - couple of fairly simple prototypes of UI/access patters for simpler DAS querying
 - check performance improvements after creation of materialized view(s)
-- interview more DAS users
+- look into possible implementations for combined queries
+
 \end{verbatim}
+\end{footnotesize}
 }
@@
 \pagebreak
-\include{das_logs}
+\input{milestones}
+
+\pagebreak
+\input{das_logs}
+
+{\color{gray}
+\subsubsection*{Not so relevant from Literature Review: Deep web search at Google}
+%[multiple papers from Google] discusses various ways for implementing data integration in terms of large-scale search engine (Google): Virtual integration vs.
+Surfacing.
+They also present ways for integrating systems without human intervention through use of statistical 'mediator'.
+
+There are two approaches to web scale search over deep-web (here Google mostly cares about web forms): 
+%\marginpar{\scriptsize\color{red}TODO: how does google use PayGo?} 
+%\textit{Runtime query reformulation} - 'leaves data at the sources and routes queries to appropriate services'\cite[p. 1]{webscale_paygo}
+
+\textit{Deep-web surfacing} - surface deep-web (e.g.
+web forms) adding their results into the standard search index  easily allowing to using existing IR technology that scales well.
+There exist algorithms which allow to iteratively choose input parameters to the forms to surface a considerable part of the 'hidden' data without large overhead%
+	\footnote{i.e.
+if choosing parameters in not smart way, a web form with just a couple of free inputs (or even dropdowns that's easier case), could yield as many results as a cross product of all input combinations.}.
+
+\textit{Pay-as-you-go approach}\cite{webscale_paygo} - With this approach, 'a system starts with
+very few (or inaccurate) semantic mappings and these mappings are improved over time as deemed necessary';
+% TODO: cite{bootstraping}
+there is NO single mediated schema over which users pose queries: queries are routed to the relevant sources with help statistical methods that are used to model uncertainty at all levels: queries, mappings and underlying data.
+}
 

### 2012-10-19 doc/thesis/latex/report_sept.tex
final version sept

@@
 An enterprise information integration system based on simple structured queries is already in place.
-Various improvements including: adding support for more complex queries and for less restricted keyword search, improvements to system's usability and performance still have to be researched.
+Various improvements  still have to be researched: adding support for more complex queries and for less restricted keyword search, improvements to system's usability and performance.
 

### 2012-10-22 doc/thesis/latex/das_logs.tex

@@
-\appendix
-\section{DAS Query logs}
+%\appendix
+\begin{appendix}
+\section{DAS Query logs (from 2011-06-21 to 2012-10-01)}
 
-\subsection*{Most common query patterns (from 2011-06-21 to 2012-10-01)}
-\begin{verbatim}
+From the logs it can be seen that queries requiring heavy joins are quite common (100K queries through command line interface) that makes it worth investigating the possible performance optimizations.
+
+\subsection{Common query patterns through Web browser}
 total valid queries: 569,408
-not well formed queries (e.g. free text or with typing mistakes): 98,923
+\\
+not well formed queries (e.g. free text, typing mistakes, spam): 98,923
 
-50.16% (285605):	dataset dataset.name=?
+\begin{Verbatim}[commandchars=\\\{\},numbers=left,numbersep=4pt]
+50.16% (285605)	:	dataset dataset.name=?
 13.54% (77071)	:	site dataset.name=?
@@
 2.11%  (12036)	:	config dataset.name=?
-1.65%  (9420) 	:	dataset run.run_number=?
+\textcolor{red}{1.65%  (9420) 	:	dataset run.run_number=?}
 1.34%  (7642) 	:	block dataset.name=?
@@
 0.77%  (4363) 	:	dataset release.name=?
-0.75%  (4257) 	:	file dataset.name=? run.run_number=?
+\textcolor{red}{0.75%  (4257) 	:	file dataset.name=? run.run_number=?}
 0.68%  (3874) 	:	run run.run_number=?
@@
 0.43%  (2438) 	:	lumi file.name=?
-0.40%  (2282) 	:	dataset dataset.name=? site.name=?
+\textcolor{red}{0.40%  (2282) 	:	dataset dataset.name=? site.name=?}
 0.35%  (1999) 	:	file block.name=?
@@
 0.29%  (1631) 	:	lumi run.run_number=?
-0.25%  (1402) 	:	child dataset.name=?
-0.22%  (1268) 	:	run file.name=?
-0.21%  (1204) 	:	block block.name=?
-0.19%  (1088) 	:	release release.name=?
-0.16%  (936) 	:	site block.name=?
-0.12%  (703) 	:	file dataset.name=? lumi.number=? run.run_number=?
-0.10%  (594) 	:	parent file.name=?
-0.08%  (455) 	:	dataset block.name=?
-0.06%  (365) 	:	dataset dataset.name=? datatype.name=? release.name=?
-0.06%  (360) 	:	file dataset.name=? site.name=?
+\end{Verbatim}
 
+\begin{verbatim}
 Interesting non-valid queries:
-T!_CERN
-*herwig*/AODSIM
---> keyword search
+* keyword search: *herwig*/AODSIM
+* Users may like more complex combined queries:
     lumi file = (file dataset=/RelValProdTTbar/JobRobot-MC_42_V12_JobRobot-v1/GEN-SIM-RECO)
     file,lumi dataset=/RelValProdTTbar/JobRobot-MC_42_V12_JobRobot-v1/GEN-SIM-RECO
---> Users may like more complex combined queries
+* Users mixing up the post and pre filters:
     file dataset=/MuEG/Run2011B-PromptReco-v1/AOD, file.size >1
     file dataset=/MinimumBias/Run2010A-valskim-v6/RAW-RECO* | grep run between  [138923, 144086] 
---> Users mixing up the post and pre filters
 \end{verbatim}
+\newpage
+\subsection{Common query patterns through Command Line\label{appendix_das_cli_logs}}
+total valid queries: 6,9M\\
+non valid queries: 76K
+
+\begin{Verbatim}[commandchars=\\\{\},numbers=left,numbersep=4pt]
+39.56% (2735728):	dataset dataset.name=?
+22.80% (1576886):	dataset dataset.name=? status.name=?
+12.49% (863636) :	file dataset.name=?
+12.44% (860626) :	run run.run_number=?
+4.60%  (318248) :	site dataset.name=?
+2.18%  (150845) :	run dataset.name=?
+\textcolor{red}{1.71%  (118404) :	file dataset.name=? run.run_number=?}\label{run_dataset_heavy}
+1.32%  (90988)  :	file block.name=?
+1.26%  (87266)  :	file file.name=?
+0.50%  (34482)  :	block site.name=?
+0.35%  (24049)  :	lumi file.name=?
+0.25%  (17282)  :	release dataset.name=?
+0.18%  (12556)  :	parent file.name=?
+0.09%  (6341)   :	file dataset.name=? lumi.number=? run.run_number=?
+\textcolor{red}{0.06%  (4306)   :	file dataset.name=? site.name=?}\label{site_dataset_distr_heavy}
+0.05%  (3547)   :	dataset dataset.name=? primary_dataset.name=? release.name=? tier.name=?
+0.05%  (3352)   :	dataset file.name=?
+0.01%  (996)    :	parent dataset.name=?
+0.01%  (755)    :	dataset site.name=?
+\end{Verbatim}
+
 % 
@@
 
-\appendix
+\end{appendix}
+\begin{appendix}
+%\appendix
 \section{Data providers statistics}
@@
 \end{verbatim}
+\end{appendix}

### 2012-10-22 doc/thesis/latex/literature_review.tex

@@
 
-In the setting of Boolean Keyword-based Retrieval over heterogeneous sources a keyword query could be translated into a ranked list of structured queries (similarly as \textit{SODA} system that propose SQL, \textit{Query Forms} approach that propose SQL templates or Keymantic trying to achieve this without accessing the data).
+In the setting of Boolean Keyword-based Retrieval over heterogeneous sources, a keyword query could be translated into a ranked list of structured queries (similarly as \textit{SODA} system that propose SQL, \textit{Query Forms} approach that propose SQL templates or Keymantic trying to achieve this without accessing the data).
 In the extreme case of having no control over a web-service that do not publish its contents, techniques like Google's Deep-web surfacing could index a subset of its contents enabling keyword search to some extent.
@@
 Second, the likelihood of a remaining keyword to matching a metadata item is computed in order to rank different options of executing this keyword query on the "Hidden Web"\cite[p.942]{ethz2012}.
-
+Because of no access to the actual data, results of this method were reported considerably worse on queries containing data terms, even if all metadata (e.g.
+business ontology) is given\cite{ethz2012}, therefore it is helpful to have at least some insight on the data behind a webservice.
 % See Appendix #1, for evaluation of they Demo system.

### 2012-12-20 doc/thesis/paper/procs-template-generic.tex
initial structure of the paper

--- 2010-05-04 doc/das_paper/procs-sample.tex
@@
 % Template for Elsevier CRC journal article
-% version 1.0 dated 13 October 2009
+% version 1.2 dated 09 May 2011
 
-% This file (c) 2009 Elsevier Ltd.  Modifications may be freely made,
+% This file (c) 2009-2011 Elsevier Ltd.  Modifications may be freely made,
 % provided the edited file is saved under a different name
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Changes since version 1.1
+% - added "procedia" option compliant with ecrc.sty version 1.2a
+%   (makes the layout approximately the same as the Word CRC template)
+% - added example for generating copyright line in abstract
+
+%-----------------------------------------------------------------------------------
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%-----------------------------------------------------------------------------------
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                                          %%
+%% Important note on usage                                  %%
+%% -----------------------                                  %%
+%% This file should normally be compiled with PDFLaTeX      %%
+%% Using standard LaTeX should work but may produce clashes %%
+%%                                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 %% The '3p' and 'times' class options of elsarticle are used for Elsevier CRC
-\documentclass[3p,times]{elsarticle}
+%% The 'procedia' option causes ecrc to approximate to the Word template
+\documentclass[3p,times,procedia]{elsarticle}
 
@@
 %% Example \runauth{C.V. Radhakrishnan et al.}
-\runauth{G.D. van Albada}
+\runauth{Author name}
 
@@
 \jnltitlelogo{Procedia Computer Science}
+
+%% Provide the copyright line to appear in the abstract
+%% Usage:
+%   \CopyrightLine[<text-before-year>]{<year>}{<restt-of-the-copyright-text>}
+%   \CopyrightLine[Crown copyright]{2011}{Published by Elsevier Ltd.}
+%   \CopyrightLine{2011}{Elsevier Ltd. All rights reserved}
+\CopyrightLine{2012}{The Authors. Published by Elsevier B.V.\newline Selection and/or peer-review under responsibility of the [Organiser Name]}
 
@@
 
-% mdwlist allows you to create e.g. bulleted lists without extra line spacing
-\usepackage{mdwlist}
-
 % declarations for front matter
@@
 
-\dochead{International Conference on Computational Science, ICCS 2010}
+\dochead{Conference Title}
 %% Use \dochead if there is an article header, e.g. \dochead{Short communication}
+%% \dochead can also be used to include a conference title, if directed by the editors
+%% e.g. \dochead{17th International Conference on Dynamical Processes in Excited States of Solids}
 
-\title{Click \emph{here}, type the title of your paper, Capitalize first letter}
+\title{Click here, type the title of your paper, Capitalize first letter}
 
@@
 
-\author{Dick van Albada\corref{cor1}\fnref{ref2}}
-\fntext[ref2]{Corresponding author}
-\cortext[cor1]{\ead {G.D.vanAlbada@uva.nl}}
+\author[a]{First Author} 
+\author[b]{Second Authorb}
+\author[a,b]{Third Author}
 
-\address{Computational Science, University of Amsterdam}
+\address[a]{First affiliation, Address, City and Postcode, Country}
+\address[b]{Second affiliation, Address, City and Postcode, Country}
 
 \begin{abstract}
-Put your abstract here - do not forget to add relevant keywords. This sample file uses
-the text from the MS Word template to allow some comparison of the formats.
 %% Text of abstract
+Click here and insert your abstract text.
 \end{abstract}
@@
 \begin{keyword}
+Type your keywords here, separated by semicolons ;
+
 %% keywords here, in the form: keyword \sep keyword
-template \sep latex \sep Procedia
+
 %% PACS codes here, in the form: \PACS code \sep code
@@
 
+\correspondingauthor[*]{Corresponding author. Tel.: +0-000-000-0000 ; fax: +0-000-000-0000 .}
+\email{author@institute.xxx}
+
 %%
@@
 %% main text
-\section{intro}
-\label{intro}
-(10 pt) Here introduce the paper, and put a nomenclature if necessary, in a box with the same font size as the rest of the paper.
+\section{Main Text}
+\label{main}
+
+Here introduce the paper, and put a nomenclature if necessary, in a box with the same font size as the rest of the paper.
 The paragraphs continue from here and are only separated by headings, subheadings, images and formulae.
@@
 Here follows further instructions for authors.
-\subsection{Structure}
-\label{structure}
-Files should be in MS Word format only and should be formatted for direct printing.
+
+\begin{nomenclature}
+\begin{deflist}[A]
+\defitem{A}\defterm{radius of}
+\defitem{B}\defterm{position of}
+\defitem{C}\defterm{further nomenclature continues down the page inside the text box}
+\end{deflist}
+\end{nomenclature}
+
+\subsection{ Structure}
+For this Procedia the files must be in MS Word format only and should be formatted for direct printing.
+Figures and tables should be embedded and not supplied separately.
+Please make sure that you use as much as possible normal fonts in your documents.
+
+Special fonts, such as fonts used in the Far East (Japanese, Chinese, Korean, etc.) may cause problems during processing.
+
+To avoid unnecessary errors you are strongly advised to use the `spellchecker' function of MS Word.
+Follow this order when typing manuscripts: Title, Authors, Affiliations, Abstract, Keywords, Main text (including figures and tables), Acknowledgements, References, and Appendix.
+Collate acknowledgements in a separate section at the end of the article and do not include them on the title page, as a footnote to the title or otherwise.
+
+For this Procedia the files must be in MS Word format only and should be formatted for direct printing.
 Figures and tables should be embedded and not supplied separately.
@@
 Special fonts, such as fonts used in the Far East (Japanese, Chinese, Korean, etc.) may cause problems during processing.
-To avoid unnecessary errors you are strongly advised to use the 'spellchecker' function of MS Word.
-Follow this order when typing manuscripts: Title, Authors, Affiliations, Abstract, Keywords, Main text (including figures and tables), Acknowledgements, References, Appendix.
-Collate acknowledgements in a separate section at the end of the article and do not include them on the title page, as a footnote to the title or otherwise.
-\\
+
+To avoid unnecessary errors you are strongly advised to use the `spellchecker' function of MS Word.
+Follow this order when typing manuscripts: Title, Authors, Affiliations, Abstract, Keywords, Main text (including figures and tables), Acknowledgements,
+
+References, Appendix. Collate acknowledgements in a separate section at the end of the article and do not include them on the title page, as a footnote to the title or otherwise.
+
 Bulleted lists may be included and should look like this:
-\begin{itemize*}
+\begin{itemize}[]
 \item First point
@@
 \item And so on
-\end{itemize*}
-Ensure that you return to the 'Els-body-text' style, the style that you will mainly be using for large blocks of text, when you have completed your bulleted list. 
-Please do not alter the formatting and style layouts which have been set up in this template document.
-As indicated in the template, papers should be prepared in single column format suitable for direct printing onto A4 paper (8.3in x 11.7in/210mm x 297mm).
+\end{itemize}
+
+Ensure that you return to the `Els-body-text' style, the style that you will mainly be using for large blocks of text, when you have completed your bulleted list.
+
+Please do not alter the formatting and style layouts which have been set up in this template document. As indicated in the template, 
+Papers should be prepared in single column format suitable for direct printing onto paper size (192 x 262) mm.
 Do not number pages on the front, as page numbers will be added separately for the preprints and the Proceedings.
 Leave a line clear between paragraphs.
-All the required style templates are provided in this document with the appropriate name supplied, e.g.
+All the required style templates are provided in this document with the appropriate name supplied e.g.
 choose 1.
 Els1st-order-head for your first order heading text, els-abstract-text for the abstract text etc.
-\subsection{Tables}
-\label{table}
+
+\subsection{ Tables}
+
 All tables should be numbered with Arabic numerals.
@@
 
-%% The Appendices part is started with the command \appendix;
-%% appendix sections are then done as normal sections
-%% \appendix
+\begin{table}[h]
+\caption{An example of a table.}
+\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}lll@{}}
+\hline
+An example of a column heading & Column A (t) & Column B (T)\\
+\hline
+And an entry &   1 &  2\\
+And another entry  & 3 &  4\\
+And another entry &  5 &  6\\
+\hline
+\end{tabular*}
+\end{table}
 
-%% \section{}
-%% \label{}
+\subsection{ Construction of references}
+
+References should be added at the end of the paper, and its corresponding citation will be added in the order of their appearance in the text.
+Authors should ensure that every reference in the text appears in the list of references and vice versa.
+Indicate references by [1], [2-3] in the text.
+The actual authors can be referred to, but the reference citation(s) must always be given.
+Some examples of how your references should be listed are given at the end of this template in the `References' section, which will allow you to assemble your reference list according to the correct format and font size.
+
+\subsection{Section headings}
+Section headings should be left justified, with the first letter capitalized and numbered consecutively, starting with the Introduction.
+Sub-section headings should be in capital and lower-case italic letters, numbered 1.1, 1.2, etc, and left justified, with second and subsequent lines indented.
+You may need to insert a page break to keep a heading with its text.
+
+\subsection{General guidelines for the preparation of your text}
+Avoid hyphenation at the end of a line.
+Symbols denoting vectors and matrices should be indicated in bold type.
+Scalar variable names should normally be expressed using italics.
+Weights and measures should be expressed in SI units.
+Please title your files in this order conferenceacrynom\_authorslastname.pdf
+
+\section{ Author Artwork}
+All figures should be numbered with Arabic numerals (1,2,...n).
+All photographs, schemas, graphs and diagrams are to be referred to as figures.
+Line drawings should be good quality scans or true electronic output.
+Low-quality scans are not acceptable.
+Figures must be embedded into the text and not supplied separately.
+Lettering and symbols should be clearly defined either in the caption or in a legend provided as part of the figure.
+Figures should be placed at the top or bottom of a page wherever possible, as close as possible to the first reference to them in the paper.
+
+\begin{figure}[h]
+\centerline{\includegraphics{fx1}\hspace*{5mm}\includegraphics{fx1}}
+\caption{(a) first picture; (b) second picture.}
+\end{figure}
+
+The figure number and caption should be typed below the illustration in 8 pt and left justified.
+For more guidelines and information to help you submit high quality artwork please visit: http://www.elsevier.comwps/find/authorsview.
+authors/authorartworkinstructions.
+Artwork has no text along the side of it in the main body of the text.
+However, if two images fit next to each other, these may be placed next to each other to save space, see Fig 1.
+They must be numbered consecutively, all figures, and all tables respectively.
+
+\subsection{Footnotes}
+Footnotes should be avoided if possible.
+Necessary footnotes should be denoted in the text by consecutive superscript letters.
+The footnotes should be typed single spaced, and in smaller type size (8 pt), at the foot of the page in which they are mentioned, and separated from the main text by a short line extending at the foot of the column.
+The `Els-footnote' style is available in this template for the text of the footnote.
+
+Equations and formulae should be typed and numbered consecutively with Arabic numerals in parentheses on the right hand side of the page (if referred to explicitly in the text),
+\begin{equation}
+\begin{array}{rcl}
+\displaystyle X_r &=& \displaystyle\dot{Q}^{"}_{rad}\left/\left(\dot{Q}^{"}_{rad} + \dot{Q}^{"}_{conv}\right)\right.\\[6pt]
+\displaystyle \rho &=& \displaystyle\frac{\vec{E}}{J_c(T={\rm const.})\cdot\left(P\cdot\left(\displaystyle\frac{\vec{E}}{E_c}\right)^m+(1-P)\right)}
+\end{array}
+\end{equation}
+
+They should also be separated from the surrounding text by one space.
+
+\section{Online licence}
+All authors must Transfer the Online licence before the article can be published.
+This transfer agreement enables Elsevier to protect the copyrighted material for the authors, but does not relinquish the authors' proprietary rights.
+The copyright transfer covers the exclusive rights to reproduce and distribute the article, including reprints, photographic reproductions, microfilm or any other reproductions of similar nature and translations.
+Authors are responsible for obtaining from the copyright holder permission to reproduce any figures for which copyright exists.
+
+The citation must be used in following style: \cite{article-minimal}, \cite{article-full}, \cite{article-crossref}, \cite{whole-journal} and \cite{inbook-minimal}.
+
+\section*{Acknowledgements}
+
+These and the Reference headings are in bold but have no numbers. Text below continues as normal.
 
@@
 \bibliographystyle{elsarticle-num}
-\bibliography{<your-bib-database>}
+\bibliography{xampl}
 
@@
 
+%% The Appendices part is started with the command \appendix;
+%% appendix sections are then done as normal sections
+%% \appendix
+
+%% \section{}
+%% \label{}
+
+\appendix
+\section{An example appendix}
+Authors including an appendix section should do so after References section. Multiple appendices should all have headings in the style used above. They will automatically be ordered A, B, C etc.
+
+\subsection{Example of a sub-heading within an appendix}
+There is also the option to include a subheading within the Appendix if you wish.
+
 \end{document}

### 2012-12-20 doc/thesis/paper/procs-template_new.tex
initial structure of the paper

--- 2012-12-20 doc/thesis/paper/procs-template-generic.tex
@@
 %% The `ecrc' package must be called to make the CRC functionality available
-\usepackage{ecrc}
+\usepackage{ecrc, fleqn}
 
@@
 %   \CopyrightLine{2011}{Elsevier Ltd. All rights reserved}
-\CopyrightLine{2012}{The Authors. Published by Elsevier B.V.\newline Selection and/or peer-review under responsibility of the [Organiser Name]}
+%\CopyrightLine{2012}{The Authors.
+Published by Elsevier B.V.\newline Selection and/or peer-review under responsibility of Scientific Programme Committee of the 4th International Conference on Software Development for Enhancing Accessibility and Fighting Info-exclusion (DSAI 2012)}
 
@@
 
-\dochead{Conference Title}
+\dochead{International Conference on Computational Science, ICCS 2013}
 %% Use \dochead if there is an article header, e.g. \dochead{Short communication}
@@
 \author[b]{Second Authorb}
-\author[a,b]{Third Author}
+\author[a,b]{Third Author\corref{cor1}}
 
@@
 %% Text of abstract
-Click here and insert your abstract text.
+Click here and insert your abstract text.\vskip11pt
 \end{abstract}
@@
 \end{keyword}
-
+ \cortext[cor1]{Corresponding author. Tel.: +0-000-000-0000 ; fax: +0-000-000-0000 .}
 \end{frontmatter}
 
-\correspondingauthor[*]{Corresponding author. Tel.: +0-000-000-0000 ; fax: +0-000-000-0000 .}
+%\correspondingauthor[*]{Corresponding author. Tel.: +0-000-000-0000 ; fax: +0-000-000-0000 .}
 \email{author@institute.xxx}
@@
 
-Here introduce the paper, and put a nomenclature if necessary, in a box with the same font size as the rest of the paper.
+(10 pt) Here introduce the paper, and put a nomenclature if necessary, in a box with the same font size as the rest of the paper.
 The paragraphs continue from here and are only separated by headings, subheadings, images and formulae.
@@
 
-\begin{nomenclature}
-\begin{deflist}[A]
-\defitem{A}\defterm{radius of}
-\defitem{B}\defterm{position of}
-\defitem{C}\defterm{further nomenclature continues down the page inside the text box}
-\end{deflist}
-\end{nomenclature}
-
 \subsection{ Structure}
-For this Procedia the files must be in MS Word format only and should be formatted for direct printing.
-Figures and tables should be embedded and not supplied separately.
-Please make sure that you use as much as possible normal fonts in your documents.
-
-Special fonts, such as fonts used in the Far East (Japanese, Chinese, Korean, etc.) may cause problems during processing.
-
-To avoid unnecessary errors you are strongly advised to use the `spellchecker' function of MS Word.
-Follow this order when typing manuscripts: Title, Authors, Affiliations, Abstract, Keywords, Main text (including figures and tables), Acknowledgements, References, and Appendix.
-Collate acknowledgements in a separate section at the end of the article and do not include them on the title page, as a footnote to the title or otherwise.
-
-For this Procedia the files must be in MS Word format only and should be formatted for direct printing.
+Files should be in MS Word format only and should be formatted for direct printing.
 Figures and tables should be embedded and not supplied separately.
@@
 Special fonts, such as fonts used in the Far East (Japanese, Chinese, Korean, etc.) may cause problems during processing.
-
-To avoid unnecessary errors you are strongly advised to use the `spellchecker' function of MS Word.
-Follow this order when typing manuscripts: Title, Authors, Affiliations, Abstract, Keywords, Main text (including figures and tables), Acknowledgements,
-
-References, Appendix. Collate acknowledgements in a separate section at the end of the article and do not include them on the title page, as a footnote to the title or otherwise.
+To avoid unnecessary errors you are strongly advised to use the ‘spellchecker’ function of MS Word.
+Follow this order when typing manuscripts: Title, Authors, Affiliations, Abstract, Keywords, Main text (including figures and tables), Acknowledgements, References, Appendix.
+Collate acknowledgements in a separate section at the end of the article and do not include them on the title page, as a footnote to the title or otherwise.
 
@@
 
-Please do not alter the formatting and style layouts which have been set up in this template document. As indicated in the template, 
-Papers should be prepared in single column format suitable for direct printing onto paper size (192 x 262) mm.
+Please do not alter the formatting and style layouts which have been set up in this template document.
+As indicated in the template, Papers should be prepared in single column format suitable for direct printing onto paper size (192 x 262) mm.
 Do not number pages on the front, as page numbers will be added separately for the preprints and the Proceedings.
 Leave a line clear between paragraphs.
-All the required style templates are provided in this document with the appropriate name supplied e.g.
+All the required style templates are provided in this document with the appropriate name supplied, e.g.
 choose 1.
@@
 
-\subsection{ Construction of references}
+\subsection{Construction of references}
 
-References should be added at the end of the paper, and its corresponding citation will be added in the order of their appearance in the text.
+References should be listed at the end of the paper, and numbered in the order of their appearance in the text.
 Authors should ensure that every reference in the text appears in the list of references and vice versa.
-Indicate references by [1], [2-3] in the text.
-The actual authors can be referred to, but the reference citation(s) must always be given.
-Some examples of how your references should be listed are given at the end of this template in the `References' section, which will allow you to assemble your reference list according to the correct format and font size.
+Indicate references by numbers in the text.
+In the text the number of the reference should be given in square brackets [??].
+The actual authors can be referred to, but the reference number(s) must always be given.
+
+Some examples of how your references should be listed are given at the end of this template in the ‘References’ section which will allow you to assemble your reference list according to the correct format and font size.
+There is a shortened form for last page number.
+e.g., 51–9, and that for more than 6 authors the first 6 should be listed followed by “et al.”
 
 \subsection{Section headings}
+
 Section headings should be left justified, with the first letter capitalized and numbered consecutively, starting with the Introduction.
@@
 
+\subsection{Footnote}
+Footnotes should be avoided if possible.
+Necessary footnotes should be denoted in the text by consecutive superscript letters.
+The footnotes should be typed single spaced, and in smaller type size (8pt), at the foot of the page in which they are mentioned, and separated from the main text by a short line extending at the foot of the column.
+The ‘Els-footnote’ style is available in this template for the text of the footnote.
+
 \section{ Author Artwork}
@@
 
+The figure number and caption should be typed below the illustration in 9pt and left justified.
+For more guidelines and information to help you submit high quality artwork please visit: http://www.elsevier.com/wps/find/ authorsview.authors/authorartworkinstructions.
+Artwork has no text along the side of it in the main body of the text.
+However, if two images fit next to each other, these may be placed next to each other to save space, see Fig 1.
+They must be numbered consecutively, all figures, and all tables respectively.
+
 \begin{figure}[h]
@@
 
-The figure number and caption should be typed below the illustration in 8 pt and left justified.
-For more guidelines and information to help you submit high quality artwork please visit: http://www.elsevier.comwps/find/authorsview.
-authors/authorartworkinstructions.
-Artwork has no text along the side of it in the main body of the text.
-However, if two images fit next to each other, these may be placed next to each other to save space, see Fig 1.
-They must be numbered consecutively, all figures, and all tables respectively.
-
-\subsection{Footnotes}
-Footnotes should be avoided if possible.
-Necessary footnotes should be denoted in the text by consecutive superscript letters.
-The footnotes should be typed single spaced, and in smaller type size (8 pt), at the foot of the page in which they are mentioned, and separated from the main text by a short line extending at the foot of the column.
-The `Els-footnote' style is available in this template for the text of the footnote.
-
 Equations and formulae should be typed and numbered consecutively with Arabic numerals in parentheses on the right hand side of the page (if referred to explicitly in the text),
 \begin{equation}
-\begin{array}{rcl}
-\displaystyle X_r &=& \displaystyle\dot{Q}^{"}_{rad}\left/\left(\dot{Q}^{"}_{rad} + \dot{Q}^{"}_{conv}\right)\right.\\[6pt]
-\displaystyle \rho &=& \displaystyle\frac{\vec{E}}{J_c(T={\rm const.})\cdot\left(P\cdot\left(\displaystyle\frac{\vec{E}}{E_c}\right)^m+(1-P)\right)}
-\end{array}
+Rt = K EP = 93.02 (\pm 9.62) – 13.45
 \end{equation}
@@
 
-\section{Online licence}
-All authors must Transfer the Online licence before the article can be published.
+\section{Copyright}
+All authors must sign the Transfer of Copyright agreement before the article can be published.
 This transfer agreement enables Elsevier to protect the copyrighted material for the authors, but does not relinquish the authors' proprietary rights.

### 2012-12-21 doc/thesis/latex/milestones.tex
elsevier styles; minor changes

\begin{appendix}
\section{Project Milestones}

Initially written up by Valentin (technical lead of the project from US):

\begin{footnotesize}
\begin{Verbatim}[commandchars=\\\{\}]
Sep:
   - familiarize with techniques, research activities in the field
   - setup all tools and learn code to do actual work
   - describe and put on paper CMS use case
   - propose several approaches to target the project
Oct:
   - pick-up couple of approaches for development
   - develop initial prototype in high-level
       - outline workflow
   - math proof of choosen approaches
   - present results at EPFL & CMS DMWM team
Nov:
   - made a choice among target approach and start working on real prototype
   - it would be nice to see math foundations in place, including 
     inital benchmarks, time estimates, stats, etc.
   - mid-term write-up document about current achievements, 
     description of choosen approach and its implementation milestones
   - at this stage work with couple of data-providers (or their limited scope) 
     to demonstrate capabilties of the chosen approach
   - I do not expect at this stage any optimizations, but would like to hear 
     where those can be done, including strategies, technical improvements, etc.
   - do not limit yourself to CMS use case, explore possibility to apply
       choosen approach outside of CMS field
     - a toy prototype with non CMS data-providers is desired to have
            - demonstrate and discuss if it can be applied elsewhere
   - present results at EPFL & CMS DMWM team
   - Peter's action: based on the work presented, make a decision about conference
     submission (see below Jan milestone for approval)
Dec:
   - expand existing work from couple of data-providers into full set of CMS  data-services
   - work on integration with existing DAS code and prepare patch for  production deployment
   - plenty of testing, including but not limited to:
     - unit tests
     - integration tests
     - end-user UI studies
           - good to have prototype and give it to end-users
           - study and observe user behavior
   - present final strategy at EPFL & CMS DMWM team, including benchmarks, stats, user UI studies
Jan:
   - work on optimization (if required)
   - integration with DAS and production deployment
   - prepare main sections of write-up document
     - target approach & math foundation
     - prototype description
     - user studies, benchmarks, stats
   - submit write-up for EPFL & CMS review
     - if it is done, Peter will need to get some approval for conference presentation
Feb:
   - final document in place describing work, achievements, results
   - make conclusion about contribution to the field
     - discuss innovation work
     - discuss engineering work
     - discuss CMS use case and achievements
     - discuss possible application outside the field
Mar 15:
    - thesis delivery
Apr:
   - defense
\end{Verbatim}
\end{footnotesize}

\end{appendix}

### 2012-12-21 doc/thesis/latex/proposed_solutions.tex
elsevier styles; minor changes

@@
 Also it's possible have different validity dates for certain fields.
-if no volatile fields are not explicitly requested, even a very old cache could be used
+If no volatile fields are explicitly requested, even a very old cache could be used.
 
@@
 \subsubsection*{Integrating distributed information efficiently}
-Bloom-join (which could be quite transparent and implemented even on DB side (efficient pure SQL is doable for MySQL, and probably for Oracle too) quite almost transparently (a helper function could take the initial query and bloom-filter's bit-vector as parameters).
+Bloom-join (which could be quite transparent and implemented even on DB side (efficient pure SQL is doable for MySQL, and probably for Oracle too) quite transparently (a helper function could take the initial query and bloom-filter's bit-vector as parameters).

### 2012-12-21 doc/thesis/paper/literature_review.tex
elsevier styles; minor changes

--- 2012-10-22 doc/thesis/latex/literature_review.tex
@@
-\section{Preliminary Literature review}
-
 % Overview: keyword search is good for non-structured documents, it is not [as] effective with structured sources\cite{Levy96}, therefore the keyword search on relational databases approach is good then no specific APIs exist and it's there are no resources to do \textbf{a full integration} .
@@
 on uncertainty based self-improving systems\cite[ch.19]{principles_data_integration}.
+{\color{red} cite from survey: \cite{IIHet_survey08}}
 
@@
 
-\subsection{Composing multiple data sources efficiently: Query translation using Logical Views\label{IM_query_translation}}
-The \textit{Information Manifold}\cite{Levy96} is virtual integration (EII) system that represents  both it's \textit{queries} and \textit{source descriptions} through a dialect of \textit{description logics} (could be thought as \textit{datalog}).
-It describes each source as a \textit{logical view} over the global (mediated) schema, augmented with source capabilities (e.g.
-what are possible and the required input parameters for the source to return results\footnote{this makes these logical views quite similar to source APIs used by DAS, with difference that DAS currently only describes the parameters APIs and only partially the results}).
+\subsection{Keyword search over web services}
+
+\subsubsection*{What if there is no access to index data terms?}
+\cite{Keymantic10, semantics_without_access} explores the case then there is no possibility to index the data terms, e.g.
+then a DB is behind a wrapper (e.g.
+accessible only through a \textit{Web form} in “Hidden Web” or \textit{a web-service}) then crawling is generally not possible.
 %
-This allows designing algorithms that could efficiently answer complex queries that require composing multiple the data sources that is done by finding maximally contained rewriting of the (conjunctive) input query in terms of logical views representing the sources (that is, finding an optimal way to compose the sources).
-
-For example, consider such sources expressed as views (on the left) in terms of global predicates of the mediated schema (on the right) in datalog notation (based on \cite{integr_views2000}):
-%
-{\small
-\begin{verbatim}
-v1(E, P, M) :- emp(E) & phone(E, P) & mgr(E, M). # employees, their phones and managers
-v2(E, O, D) :- emp(E) & office(E, O) & dept(E, D). # offices and departments of employees
-v3(E, P) :- emp(E) & phone(E, P) & dept(E, toy_dept). # phones of employees only in Toys dept.
-\end{verbatim}
-}
-%
-Suppose we wanted to know Sally's phone and office. We express this again in datalog over global predicates:
-{\small\begin{verbatim}
-q1(P,O) :- phone(sally,P) & office(sally,O).
-\end{verbatim}}
-There are two minimal solutions (as sources could be incomplete, the full solution is union of the two):
-{\small
-\begin{verbatim}
-answer1(P,O) :- v1(sally,P,M) & v2(sally,O,D).
-answer2(P,O) :- v3(sally,P) & v2(sally,O,D).
-\end{verbatim}}
-Notice that the expansions of these solutions (e.g.
-answer1\_exp) are not equivalent to $q_1$, but only the conjunctive queries that are closest and still contained in $q_1$ (as they are the only usable views provided by the sources):
-{\footnotesize\begin{Verbatim}[commandchars=\\\{\}]
-answer1_exp(P,O) :- emp(sally) & \textbf{\underline{phone}(sally,P)} & mgr(sally,M) & emp(sally) & \textbf{\underline{office}(sally,O)} & dept(sally,D).
-\end{Verbatim}
-}
-After this the \textit{Information Manifold} would find an executable order that adheres the capabilities of the sources, by iteratively considering any sources whose input parameters are satisfied. 
-% \subsection{Keyword search: integration on demand}
-% {\color{red}TODO?: \cite[ch.16]{principles_data_integration}}
+In Keymantic\cite{Keymantic10} a keyword query is processed as follows: First, all keywords that  correspond to metadata items (e.g., field names) are extracted.
+The remaining keywords are considered as possible parameters to the input fields in the web form.
+Second, the likelihood of a remaining keyword to matching a metadata item is computed in order to rank different options of executing this keyword query on the "Hidden Web"\cite[p.942]{ethz2012}.
+Because of no access to the actual data, results of this method were reported considerably worse on queries containing data terms, even if all metadata (e.g.
+business ontology) is given\cite{ethz2012}, therefore it is helpful to have at least some insight on the data behind a webservice.
+% See Appendix #1, for evaluation of they Demo system.
 
@@
 The users also reported system's potential a) for analysing the schema and learning patterns about it and b) as tool to help documenting legacy systems.
-
-\subsection{Keyword search over web services}
-
-\subsubsection*{What if there is no access to index data terms?}
-\cite{Keymantic10, semantics_without_access} explores the case then there is no possibility to index the data terms, e.g.
-then a DB is behind a wrapper (e.g.
-accessible only through a \textit{Web form} in “Hidden Web” or \textit{a web-service}) then crawling is generally not possible.
-%
-In Keymantic\cite{Keymantic10} a keyword query is processed as follows: First, all keywords that  correspond to metadata items (e.g., field names) are extracted.
-The remaining keywords are considered as possible parameters to the input fields in the web form.
-Second, the likelihood of a remaining keyword to matching a metadata item is computed in order to rank different options of executing this keyword query on the "Hidden Web"\cite[p.942]{ethz2012}.
-Because of no access to the actual data, results of this method were reported considerably worse on queries containing data terms, even if all metadata (e.g.
-business ontology) is given\cite{ethz2012}, therefore it is helpful to have at least some insight on the data behind a webservice.
-% See Appendix #1, for evaluation of they Demo system.

### 2013-03-07 doc/thesis/paper/sub_chapters/cms_case.tex
early draft: still editing

At the start of this project, an enterprise information integration system based on simple structured queries was already in place (described below), which suffered from a number of performance and usability problems. %
The author of this work evaluated possible approaches to these problems: improvements to system's usability through redesigning the user interface and allowing less restricted keyword search, analysis of performance bottlenecks {\color{red}and the possibility of supporting more complex queries}.\marginpar{{\color{red}only to some extent - secondary goal. we are bound by performance issues first}}

% {\color{red}A remarkable opportunity we have is possibility to test ideas on a sample of ~3000 people working at CMS.}

%Following a similar approach a 
\subsubsection*{The EII system used at CMS, CERN}

The \textit{CMS Data Aggregation System (DAS)}\cite{Kuznetsov2010, Kuznetsov2011} allows integrated access to a number of proprietary data-sources by processing users' queries on demand - it determines which  data-sources are able to answer\footnote{This is done by a mapping between flat mediated schema entities ('DAS keys') into web-service methods and their arguments. Then system queries all services that could provide a result of expected type with given parameters}, queries them, merges the results and caches them for subsequent use. DAS uses the \textit{Boolean retrieval model} as users are often interested in retrieving ALL the items matching their query.

Currently the queries are formed by specifying what entity the user is interested in (dataset, file, etc) and providing selection criteria (e.g. attribute=value, name BETWEEN [v1, v2]). The combined query results could be later 'piped' for further filtering, sorting or aggregation (min, max, avg, sum, count, median), e.g.:

{\small 
\begin{Verbatim}[commandchars=\\\{\}]
\footnotesize# find average and median dataset sizes of \textit{RelVal} datasets with more than 1000 events
dataset=*RelVal* | grep dataset.nevents >1000 | avg(dataset.size), median(dataset.size)
\end{Verbatim}
}

Queries are executed either from web browser or through a command line interface where the results could be fed into another application (e.g. programs doing physics analysis or automatic software release validation).

### 2013-03-07 doc/thesis/paper/sub_chapters/excluded.tex
early draft: still editing

--- 2012-10-19 doc/thesis/latex/literature_review.tex
@@
-\section{Preliminary Literature review}
+{\color{red}
+At large scientific collaborations like the CMS Experiment at CERN's Large Hadron Collider, joining together more than 3000 collaborators, data usually resides on a fair number of autonomous and heterogeneous proprietary systems each serving it's own purpose
+\footnote{For instance, at CERN,  due to many reasons (e.g.
+research-orientedness and need of freedom, politics of institutes involved) software projects usually evolve in independent fashion resulting in fair number of proprietary systems\cite{Koch00CERN}.
+Further high turnover makes it harder to extend these systems}.
+As data stored on one system may be related to data residing on others%
+	\footnote{For example, datasets containing physics events are registered at DBS, while the physical location of files is tracked by Phedex which also takes care of their transfers within the grid storage worldwide}%
+, users are in need of a centralized and easy-to-use solution for locating and combining data from all these multiple services.
 
-% Overview: keyword search is good for non-structured documents, it is not [as] effective with structured sources\cite{Levy96}, therefore the keyword search on relational databases approach is good then no specific APIs exist and it's there are no resources to do \textbf{a full integration} .
-
-\subsection{Overview}
-
-% Data Integration is about balancing between Virtual integration and data warehousing.
-
-Since the late 1990's, several \textit{Enterprise Information Integration}\footnote{%
-	Enterprise Information Integration (EII) is about 'integrating data from 
-	multiple sources 	\textit{without} having to to first load data into
-	 a central warehouse'\cite[p.1]{eii_2005}}
- (EII) products have appeared in the market (e.g.
-\textit{Information Manifold} by AT\&T Lab) and an significant experience has been accumulated on data integration formalisms, ways of describing heterogeneous data sources and their abilities (e.g.
-RDBMS vs web form), query optimization (combining sources efficiently, source overlap, data quality, etc)\cite{eii_2005}.
-%
-Recent research in  Enterprise Information Integration mostly focused on approaches minimizing  human efforts on source integration, e.g.
-on uncertainty based self-improving systems\cite[ch.19]{principles_data_integration}.
-
-The problem of keyword search over structured sources received significant attention within the last decade.
-Keyword search over relational, XML and other databases was explored from a number of perspectives: returning top-k ranked result tuples vs suggesting structured queries as SQL, performance optimization, user feedback mechanisms, as well as keyword searching over distributed sources.
-%
-If there is no need for 100\% result exactness, keyword search combined with probabilistic schema matching provides lightweight exploratory data integration with almost no human effort upfront with ability to improve the results by adjusting to users' feedback\cite[ch.16]{principles_data_integration}.
-On the other hand, the \textit{SODA} system has shown that if enough meta-data is in place, even quite complex queries given in bussiness terms could be answered over a large and complex warehouse.
-
-In the setting of Boolean Keyword-based Retrieval over heterogeneous sources a keyword query could be translated into a ranked list of structured queries (similarly as \textit{SODA} system that propose SQL, \textit{Query Forms} approach that propose SQL templates or Keymantic trying to achieve this without accessing the data).
-In the extreme case of having no control over a web-service that do not publish its contents, techniques like Google's Deep-web surfacing could index a subset of its contents enabling keyword search to some extent.
+Using a highly structured language like SQL is problematic because users need to know not only the language but also where to find the information and also lots of technical details like schema. %
+}
 
@@
 % {\color{red}TODO?: \cite[ch.16]{principles_data_integration}}
-
-\subsection{Keyword search over a Relational Database}
-%The problem of Keyword search over Relational Databases (or also semi-structured sources like XML) has received a significant attention by the research community over the last decade.
-
-The basic approach would first build an inverted index on values in database specifying where each comes from (usually only on text columns), as well as schema items (table and column names). 
-%
-Then, given a keyword query, the occurrences of the keywords would serve as entry points for the search, from where it would try to construct join paths (based on Foreign keys) that would unite tuples containing these keywords.
-% In addition to many papers the PhD dissertations \cite{PhD_2011, PhD_2012} describe the approaches in details including performance optimization details (e.g. generating materialized views), etc. 
-% 
-%A number of problem variations exist:  returning only the ranked Top-k results vs.
-returning ranked list of possible queries, while some systems would even allow generating more complex queries including aggregations, etc (SQAK, SODA\cite{ethz2012}).
-
-\subsubsection*{Ranking Query Templates based on keyword query}
-A simple way to access relational database could  be through a predefined set of named query templates (SQL with selection parameters or operators still to be specified) exposed to a user as a Form that the user has to fill in.
-
-\cite{forms_kws} proposes to use this for answering keyword queries: instead of returning database tuples one could rank query forms and the user could choose the intended one (if they are properly named this is fairly easy).
-The ranking is based on how well the given keywords match table names and column values contained in each template.
-
-Actually, a \textit{Query Template} is functionally similar to any autonomous web service (which given the parameters would in turn execute that query on its database), with the only difference that access to potentially complex web service is often more expensive than to a database entry.
-In case of DAS, a user after providing a query, could be provided with a ranked list of structured queries (attribute=value) that could be processed given data source constraints (e.g.
-parameters required) and if needed he could refine his search (e.g.
-provide more parameters).
-
-\subsubsection*{Keyword query cleaning}
-Keyword queries are often ambiguous, may contain misspellings or multiple keywords that refer to the same attribute value,  therefore \cite{kw_cleaning} suggested to perform query cleaning before proceeding to subsequent more computationally expensive steps (e.g.
-exploring all the possible join paths).
-
-Further employing some machine learning method like HMM\cite{kw_cleaning_hmm} would allow to incorporate user's feedback (even the fact that user has chosen n-th result as a query to be executed is a good clue).
-
-\subsubsection*{SODA: Meta-data approach}
-
-With a goal to bridge the increasing gap between high-level (conceptual, business) and low level (physical) representations of data, researchers from \textit{ETHZ} have been investigating Generation of SQL for Business users over a very complex data warehouse at \textit{Credit Suisse}.
-For converting natural language queries
-% (that in addition to keyword search could convey some semantic structure)
- into SQL statements, in addition to what used by earlier approaches they used meta-data describing the schema at both physical, conceptual and logical levels extended with DBpedia (for synonyms, etc) and domain ontologies (to capture business concepts like 'wealthy customer').
-%and some natural language processing
-
-Even on a large data-warehouse of ~220GB data with a complex schema of 400+ tables they reported that if good meta-data is available, generating even fairly complex SQL  (e.g.
-n-way joins with aggregations) is quite feasible for a computer.
-That would make it 'much easier for business users to interactively explore highly-complex data warehouses' \cite[p.932]{ethz2012}.
-The users also reported system's potential a) for analysing the schema and learning patterns about it and b) as tool to help documenting legacy systems.
-
-\subsection{Keyword search over web services}
-
-\subsubsection*{What if there is no access to index data terms?}
-\cite{Keymantic10, semantics_without_access} explores the case then there is no possibility to index the data terms, e.g.
-then a DB is behind a wrapper (e.g.
-accessible only through a \textit{Web form} in “Hidden Web” or \textit{a web-service}) then crawling is generally not possible.
-%
-In Keymantic\cite{Keymantic10} a keyword query is processed as follows: First, all keywords that  correspond to metadata items (e.g., field names) are extracted.
-The remaining keywords are considered as possible parameters to the input fields in the web form.
-Second, the likelihood of a remaining keyword to matching a metadata item is computed in order to rank different options of executing this keyword query on the "Hidden Web"\cite[p.942]{ethz2012}.
-
-% See Appendix #1, for evaluation of they Demo system.

### 2013-03-07 doc/thesis/paper/sub_chapters/keymantic.tex
early draft: still editing

--- 2012-12-21 doc/thesis/paper/literature_review.tex
@@
-% Overview: keyword search is good for non-structured documents, it is not [as] effective with structured sources\cite{Levy96}, therefore the keyword search on relational databases approach is good then no specific APIs exist and it's there are no resources to do \textbf{a full integration} .
-
-\subsection{Overview}
-
-% Data Integration is about balancing between Virtual integration and data warehousing.
-
-Since the late 1990's, several \textit{Enterprise Information Integration}\footnote{%
-	Enterprise Information Integration (EII) is about 'integrating data from 
-	multiple sources 	\textit{without} having to to first load data into
-	 a central warehouse'\cite[p.1]{eii_2005}}
- (EII) products have appeared in the market (e.g.
-\textit{Information Manifold} by AT\&T Lab) and an significant experience has been accumulated on data integration formalisms, ways of describing heterogeneous data sources and their abilities (e.g.
-RDBMS vs web form), query optimization (combining sources efficiently, source overlap, data quality, etc)\cite{eii_2005}.
-%
-Recent research in  Enterprise Information Integration mostly focused on approaches minimizing  human efforts on source integration, e.g.
-on uncertainty based self-improving systems\cite[ch.19]{principles_data_integration}.
-{\color{red} cite from survey: \cite{IIHet_survey08}}
-
-The problem of keyword search over structured sources received significant attention within the last decade.
-Keyword search over relational, XML and other databases was explored from a number of perspectives: returning top-k ranked result tuples vs suggesting structured queries as SQL, performance optimization, user feedback mechanisms, as well as keyword searching over distributed sources.
-%
-If there is no need for 100\% result exactness, keyword search combined with probabilistic schema matching provides lightweight exploratory data integration with almost no human effort upfront with ability to improve the results by adjusting to users' feedback\cite[ch.16]{principles_data_integration}.
-On the other hand, the \textit{SODA} system has shown that if enough meta-data is in place, even quite complex queries given in bussiness terms could be answered over a large and complex warehouse.
-
-In the setting of Boolean Keyword-based Retrieval over heterogeneous sources, a keyword query could be translated into a ranked list of structured queries (similarly as \textit{SODA} system that propose SQL, \textit{Query Forms} approach that propose SQL templates or Keymantic trying to achieve this without accessing the data).
-In the extreme case of having no control over a web-service that do not publish its contents, techniques like Google's Deep-web surfacing could index a subset of its contents enabling keyword search to some extent.
-
-\subsection{Keyword search over web services}
-
-\subsubsection*{What if there is no access to index data terms?}
 \cite{Keymantic10, semantics_without_access} explores the case then there is no possibility to index the data terms, e.g.
@@
 % See Appendix #1, for evaluation of they Demo system.
-
-\subsection{Keyword search over a Relational Database}
-%The problem of Keyword search over Relational Databases (or also semi-structured sources like XML) has received a significant attention by the research community over the last decade.
-
-The basic approach would first build an inverted index on values in database specifying where each comes from (usually only on text columns), as well as schema items (table and column names). 
-%
-Then, given a keyword query, the occurrences of the keywords would serve as entry points for the search, from where it would try to construct join paths (based on Foreign keys) that would unite tuples containing these keywords.
-% In addition to many papers the PhD dissertations \cite{PhD_2011, PhD_2012} describe the approaches in details including performance optimization details (e.g. generating materialized views), etc. 
-% 
-%A number of problem variations exist:  returning only the ranked Top-k results vs.
-returning ranked list of possible queries, while some systems would even allow generating more complex queries including aggregations, etc (SQAK, SODA\cite{ethz2012}).
-
-\subsubsection*{Ranking Query Templates based on keyword query}
-A simple way to access relational database could  be through a predefined set of named query templates (SQL with selection parameters or operators still to be specified) exposed to a user as a Form that the user has to fill in.
-
-\cite{forms_kws} proposes to use this for answering keyword queries: instead of returning database tuples one could rank query forms and the user could choose the intended one (if they are properly named this is fairly easy).
-The ranking is based on how well the given keywords match table names and column values contained in each template.
-
-Actually, a \textit{Query Template} is functionally similar to any autonomous web service (which given the parameters would in turn execute that query on its database), with the only difference that access to potentially complex web service is often more expensive than to a database entry.
-In case of DAS, a user after providing a query, could be provided with a ranked list of structured queries (attribute=value) that could be processed given data source constraints (e.g.
-parameters required) and if needed he could refine his search (e.g.
-provide more parameters).
-
-\subsubsection*{Keyword query cleaning}
-Keyword queries are often ambiguous, may contain misspellings or multiple keywords that refer to the same attribute value,  therefore \cite{kw_cleaning} suggested to perform query cleaning before proceeding to subsequent more computationally expensive steps (e.g.
-exploring all the possible join paths).
-
-Further employing some machine learning method like HMM\cite{kw_cleaning_hmm} would allow to incorporate user's feedback (even the fact that user has chosen n-th result as a query to be executed is a good clue).
-
-\subsubsection*{SODA: Meta-data approach}
-
-With a goal to bridge the increasing gap between high-level (conceptual, business) and low level (physical) representations of data, researchers from \textit{ETHZ} have been investigating Generation of SQL for Business users over a very complex data warehouse at \textit{Credit Suisse}.
-For converting natural language queries
-% (that in addition to keyword search could convey some semantic structure)
- into SQL statements, in addition to what used by earlier approaches they used meta-data describing the schema at both physical, conceptual and logical levels extended with DBpedia (for synonyms, etc) and domain ontologies (to capture business concepts like 'wealthy customer').
-%and some natural language processing
-
-Even on a large data-warehouse of ~220GB data with a complex schema of 400+ tables they reported that if good meta-data is available, generating even fairly complex SQL  (e.g.
-n-way joins with aggregations) is quite feasible for a computer.
-That would make it 'much easier for business users to interactively explore highly-complex data warehouses' \cite[p.932]{ethz2012}.
-The users also reported system's potential a) for analysing the schema and learning patterns about it and b) as tool to help documenting legacy systems.

### 2013-03-07 doc/thesis/paper/sub_chapters/literature_overview_short.tex
early draft: still editing

% Overview: keyword search is good for non-structured documents, it is not [as] effective with structured sources\cite{Levy96}, therefore the keyword search on relational databases approach is good then no specific APIs exist and it's there are no resources to do \textbf{a full integration} .

%\section{Overview}

% Data Integration is about balancing between Virtual integration and data warehousing.

Since the late 1990's, several \textit{Enterprise Information Integration}\footnote{%
	Enterprise Information Integration (EII) is about 'integrating data from 
	multiple sources 	\textit{without} having to to first load data into
	 a central warehouse'\cite[p.1]{eii_2005}}
 (EII) products have appeared in the market (e.g. \textit{Information Manifold} by AT\&T Lab) and an significant experience has been accumulated on data integration formalisms, ways of describing heterogeneous data sources and their abilities (e.g. RDBMS vs web form), query optimization (combining sources efficiently, source overlap, data quality, etc)\cite{eii_2005}. 
%
Recent research in  Enterprise Information Integration mostly focused on approaches minimizing  human efforts on source integration, e.g. on uncertainty based self-improving systems\cite[ch.19]{principles_data_integration}.\marginpar{\color{red} cite from survey: \cite{IIHet_survey08}}

The problem of keyword search over structured sources received a significant attention within the last decade. Keyword search over relational and other structured databases was explored from a number of perspectives: returning top-k ranked result tuples vs suggesting structured queries as SQL, performance optimization, user feedback mechanisms, keyword searching over distributed sources, up to lightweight exploratory\footnote{because of probabilistic nature of schema mappings, it do not provide 100\% result exactness} probabilistic data integration based on users-feedback minimizing the upfront human effort required\cite[ch.16]{principles_data_integration}. 
%
%If there is no need for 100\% result exactness, keyword search combined with probabilistic schema matching provides lightweight exploratory data integration with almost no human effort upfront with ability to self-improve given users' feedback\cite[ch.16]{principles_data_integration}. 
%
On the other extreme, the \textit{SODA}{ethz2012} system has proved that if enough meta-data is in place, even quite complex queries given in bussiness terms could be answered over a large and complex warehouse.

Meanwhile, to the best of our knowledge, the \textit{Boolean Keyword-based search over heterogeneous sources}, which is our main focus, received much less attention from the research community. 
%
{\color{magenta}As the users are interested in complete answer sets, the standard approach is translating the keywords into a ranked list of structured queries (e.g.\textit{SODA} system that propose SQL over large datawarehouse, or KEYRY/Keymantic trying to achieve this without accessing the data).}
%
We have identified these three categories of related research contributions: heuristic based-keyword search, machine learning-based keyword search, and {\color{magenta}less mature works} attemping to process multi-domain natural language queries.
%
%\marginpar{\color{red}TODO: describe keyword search}
%In the extreme case of having no control over a web-service that do not publish its contents, techniques like Google's Deep-web surfacing could index a subset of its contents enabling keyword search to some extent.
% \textit{Query Forms} approach that propose SQL templates

### 2013-03-07 doc/thesis/paper/sub_chapters/literature_review.tex
early draft: still editing

--- 2012-12-21 doc/thesis/paper/literature_review.tex
@@
 
-\subsection{Overview}
+\section{Overview}
 
@@
 %
-If there is no need for 100\% result exactness, keyword search combined with probabilistic schema matching provides lightweight exploratory data integration with almost no human effort upfront with ability to improve the results by adjusting to users' feedback\cite[ch.16]{principles_data_integration}.
+If there is no need for 100\% result exactness, keyword search combined with probabilistic schema matching provides lightweight exploratory data integration with almost no human effort upfront with ability to self-improve given users' feedback\cite[ch.16]{principles_data_integration}.
+%
 On the other hand, the \textit{SODA} system has shown that if enough meta-data is in place, even quite complex queries given in bussiness terms could be answered over a large and complex warehouse.
@@
 
-\subsection{Keyword search over web services}
+\section{Keyword search over web services}
 
-\subsubsection*{What if there is no access to index data terms?}
+\subsubsection*{Then there is no access to index all data terms...}
 \cite{Keymantic10, semantics_without_access} explores the case then there is no possibility to index the data terms, e.g.
@@
 
-\subsection{Keyword search over a Relational Database}
+{\color{red}TODO:KEYRY}
+
+\section{Keyword search over a Relational Database}
 %The problem of Keyword search over Relational Databases (or also semi-structured sources like XML) has received a significant attention by the research community over the last decade.
